### 2013-08-30 jacobian_calc_CMAME.tex
Set up repository

%%
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%%
%%
%% $Id: elsarticle-template-num.tex 4 2009-10-24 08:22:58Z rishi $
%%
%%
\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}

%Package for adding notes/comments/and tracking changes
\usepackage[inline]{trackchanges}
\addeditor{JTF}
\addeditor{Michael}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

\journal{Computer Methods in Applied Mechanics and Engineering}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
\author{Michael D. Brothers}
\author{John T. Foster\corref{cor1}}
\author{Harry R. Millwater}
\address{Mechanical Engineering Department, The University of Texas at San Antonio}

%% \ead{email address}
%% \ead[url]{home page}
\cortext[cor1]{Corresponding Author: john.foster@utsa.edu}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{A comparison of different methods for calculating tangent-stiffness matrices in a massively parallel computational peridynamics code.}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\begin{abstract} %% Text of abstract Shown is a retrospective comparative study of tangent-stiffness
matrix calculation methods including a newly developed Complex Taylor Series Expansion
(CTSE)-based ``complex-step'' method alongside established methods: forward-difference,
central-difference and automatic differentiation.  Tangent-stiffness matrix is a term from
computational mechanics which here refers to the Jacobian matrix as used in first-order
Newton-Raphson methods for the solution of non-linear algebraic systems of equations developed
to describe physical or theoretical systems studied in math, science, engineering and business.
To perform the comparative study, the above tangent-stiffness calculation methods were applied
in a massively parallel computational peridynamics code developed at Sandia National Labs,
called \emph{Peridigm}. In the comparative study, for each datum for each run, complex-step was
multiple orders of magnitude more accurate the finite-difference methods, according to a common
comparison made to the automatic-differentiation method datum. However, for the implementation
contemporary to the study, complex-step was also the slowest method among the four for each
datum for each run. A mathematical definition of \change[Michael]{the methods}{complex-step and
and automatic differentiation}, description of the computer implementations and justification
of the comparative study precede results for clarity. The intended audience of this paper
includes researchers, professionals and students, \change[Michael]{therefore conceptually explicit language
appealing to multiple levels of understanding is used}{therefor readers are directed to
selected background sources in the text for well-known concepts while less familiar concepts
are derived in-paper with an attempt made to lay steps out plainly for those in disparate fields
or who are new to their studies.}  \end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Newton's method \sep Newton-Raphson \sep numeric differentiation \sep complex-step \sep finite-difference \sep automatic-differentiation \sep Jacobian \sep tangent-stiffness
%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\MSC [2010] 65D25 
\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text
\section{Acknowledgements}
This work was funded in part by a grant from the United States Airforce Office of Scientific
Research. This work is affiliated with the Center for Simulation Visualization and Real-Time
Prediction at the University of Texas at San Antonio.

\section{Introduction}
%\subsection{Motivation}
\label{sec:intro}

In order to maintain the quadratic convergence properties of the first-order Newton's method in
quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate,
algorithmically consistent tangent-stiffness matrices. For an extremely small class of nonlinear
material models, these consistent tangent-stiffness operators can be derived analytically; however,
most often in practice, they are found through numerical approximation of derivatives.

A goal of this comparative study, was to develop and evaluate a new, accurate, and practical method
for calculating tangent-stiffness matrices against established methods.  This new method is based on
a complex number Taylor series expansion and refered to as CTSE or the ``complex-step'' method.  The
distinction of `accurate' is defined by comparison of the new method to popular finite differencing
techniques used for computing tangent-stiffness matrices and with the exact algorithmically
consistent derivatives computed with \emph{automatic differentiation}.

It was thought that in addition to comparative information regarding the new method, comparisons of
exclusively finite-difference, central-difference and automatic differentiation could prove valuable
for developers or analysts not necessarily considering complex-step but needing data to help decide
whether or not to invest in automatic-differentiation vs. finite-difference.  A comparison and
discrimination of the methods was to be achieved through application and the measurement of results
in-situ. The results would largely be treated as deterministic consequences of method choice. There
was not an attempt made to mathematically prove the suitability of one method over another of the
methods. The study was retrospective, as explained in section {SECTION} of the instant paper.

The scope of the application component of the study was limited to the Linear Peridynamic Solid
material model implemented in the computational peridynamics code, \textit{Peridigm}. Identifying a
specific application provided a practical framework for implementing the new method and solving
engineering problems to generate the data needed to compare the methods. In particular,
\emph{Peridigm} was chosen cause it combined several helpful characteristics: The reliance on
Newton's method and tangent-stiffness matrices for solving quasi-static implicit problems, prior
inclusion of finite-difference, central-difference and automatic-differentiation based methods
needed for comparison to a new method that would be developed, and being largely based in the Ansi
standard version of the popular C++ programming language.

The aim of this paper is to introduce new information, but also to compile prior art to serve as a
general reference for solving non-linear systems. \remove[Michael]{The former is mainly intended for
an audience of experienced researchers and professionals who are seeking novel tools, the later
is intended for an audience of undergraduate to graduate level students who would benefit from a
source detailing basic methodology.  According to this aim, concepts are presented in multiple
levels of detail to accommodate both a broad and deep audience.} After giving background on the
\add[Michael]{possibly less familiar} differentiation techniques underlying the methods to be
discussed here, presented in this paper is: a description of tangent-stiffness matrices and how to
use them to solve non-linear systems, detailed directions for producing tangent-stiffness matrices
with each the methods identified above \remove[Michael]{preceded by background as it becomes
necessary}, a description and justification of a new method, called 'complex-step' for calculating
tangent-stiffness matrices, a description of implementing complex-step in the software package used
in the comparative study, a description of the parameters of the comparative study conducted to rank
the methods, a presentation and analysis of the results of the comparative study, and
conclusions and thoughts on potential future work. Finally, the reader is referred to the
corresponding author's website for c++ source-code and data necessary to reproduce the results
presented here. Additionally included on the website is a C+ template library with classes for
solving non-linear systems using the techniques dicussed here, as well as two validation and
verification example problemsi which show how to use the library.

\subsection{Differentiation Techniques}

This subsection contains background information on the differentiation techniques underlying the
tangent-stiffness matrix calculation methods compared in the study. Since first-order finite
difference techniques are considered well known, they are not described where. Instead the
reader is referred to \cite[Chap. 4.1.3]{chapra2010}.

\subsubsection{Basis of New Method, 'Complex Step', in CTSE}  
\note[Michael]{TODO: cite lyness and moler, complete section}

During a literature review in preparation of material distinct from the instant paper, one of the
instant authors identified a work presenting an application of the prior work of applied
mathematicians Lyness and Moler to the numerical approximations of the first derivatives of real
functions. First called Complex Taylor Series Expansion (CTSE) in (Lyness and Moler) and later by
\cite{voorhees2011complex}, where Dr. Millwater is one of the  instant authors, the use of the term and
its acronym are adopted out of deference to their  originators. Lyness and Moler described CTSE
based on  approximating the Taylor Series Expansion of a function F, where $F:R^1 \rightarrow R^1$  about a
complex point and solving for the first derivative. The application of Lyness and Moler's work
demonstrated by Squire and Trap showed that there was empirical evidence that a CTSE based numerical
derivative formula could resist the accuracy reducing phenomenon of subtractive cancellation in
circumstances where central-difference and forward-difference were  demonstrated to succumb to
subtractive-cancellation.

\remove[Michael]{
\subsubsection{First-order Finite-difference}

The types of first-order finite-difference that will be discussed are forward-difference and
central-difference. CTSE and these finite-difference techniques are similar in that they have been
developed by algebraically solving approximated Taylor series expansions for first derivative terms.
}

\subsubsection{Differentiation Technique: Automatic-differentiation} 
\label{ADsubsection}

Automatic-differentiation or 'AD' is a computerized method for computing exact derivatives based the
chain-rule from calculus. AD takes advantage of the fact that any
mathematical function executed on a computer, no matter how complicated, is a "composition of simple
operations" (add, multiply, power, transcendental and the like) each having known analytical derivatives
\cite{ref-sacado-presentation}. For reference, the AD
implementation used in the study is the "Sacado" package from the "Trilinos" library developed out of Sandia
National Labs \cite{ref-Sacado}.

The way an AD system works is by first evaluating the innermost function of the composition, then presenting
that function's output as input to the next level function until all levels are complete, in a way no
different from how a normal computer program or human would evaluate composition functions.
AD departs in that as each nested function is evaluated, the function's
partial derivative with respect to the designated variables of the given input is also calculated.
This is possible because the elementary math functions are hard-coded into the
AD source-code along with their analytical derivatives, and linked by special instructions, so that when
the elementary math functions are called upon for computation, their partial derivatives may be computed and
stored in a sequence. The AD system then multiples the final sequence of partial derivatives together to
produce the exact equivalent to taking a partial derivative of the corresponding composition
function made up of the elementary functions with respect to a designated variable at a particular
value. It is obviously, but bears mentioning, that the AD system could simply store one value for partial
derivatives, modifying it as appropriate for every function evaluation rather than keeping a sequence.
In the literature, the AD scheme described here is called 'forward automatic-differentiation'. For
brevity, only forward AD will be covered since it is pertinent to the study, however the reader is referred
to the introduction section of \cite{ref-AD-methods} and its references list for further information on AD,
particularly \cite{ref-on-AD} which is foundational.

To make the above description complete, this subsection concludes with an example forward AD process, with 
an emphasis on how general purpose information known to the computer is specialized with additional
information from the problem and used to carry out AD:

\begin{enumerate}
%Set up example problem. The example is meant to walk the reader through the forward AD algorithm
%while encouraging them to rely on their understanding of the chain rule
\item Take an example composition function $f(x) = (sin(cos(x)))^2$\\

\item Suppose we want to know $\frac{df}{dx} \mid_{x_0}$ where $x_0$ is a particular value of $x$. \\

\begin{enumerate}
%This first block of subitems is meant to establish that the values of x that the user cares about
%about evaluating functions for are included within the set of numbers that functions g,h,k are allowed 
%to deal with, and that nested functions can take each other as input, where s may not necessarily equal
%x.
\item Given that: \\
\label{given}
\begin{enumerate}
\item $S \in R^1$ 
\item $X \in S$ 
\item $g, h, k \in H: S \rightarrow S$ 
\item $\forall s \in S : g(s) = {s}^2$, $h(s) = sin(s)$, $k(s) = cos(s)$ 
\item
\label{composition}
$\forall x \in X : f(x) = g(h(k(x)))$ \\
\end{enumerate}

\item Given that the computer is programmed with some mathematical definitions:\\
\begin{enumerate}
\item
\begin{enumerate}
\item
%The computer has at least definitions for these functions and variables
$u, v, w \in H: R^3 \rightarrow R^1$ \\
\item
$s,a,b,x \in R^1$\\
\end{enumerate}

\item
%AD doesn't return symbolic epressions for derivitives that can be re-evaluated.
%Instead, generic versions of functions and evaluations of their partials are defined together.
particular values can be identified: \\
$s = s_0,..,s_n,...s_{\infty} \mid n=[0,\infty)$ \\ 
and similarly for the other variables $a,b,x$ \\
\item
functions $u,v,w$ and their partial derivatives w.r.t $s$ are defined such that: \\
\begin{tabular}{l | l}
\multicolumn{2}{c}{for $a, b, s$ equal to $a_n, b_n, s_n$:} \\ \hline
$u \mid _{a_n, b_n, s_n} = a_n \cdot s_n^{b_n}$ & $\frac{\partial{u}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot s_n^{b_n - 1}$ \\
$v \mid _{a_n, b_n, s_n} = a_n \cdot sin(b_n \cdot s_n)$ & $\frac{\partial{v}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot cos(b_n \cdot s_n)$\\ 
$w \mid _{a_n, b_n, s_n} = a_n \cdot cos(b_n \cdot s_n)$ & $\frac{\partial{w}}{\partial{s}} \mid _{a_n, b_n, s_n} = -a_n \cdot b_n \cdot sin(b_n \cdot s_n)$\\
\end{tabular} 
\end{enumerate}
%a comment
\item 
\label{abcvalues}
%General definitions of functions are specialized by arguments. 
Given that it is possible to describe $f(x) = (sin(cos(x)))^2$ in terms the computer understands by inputting
$f(x)$ such that the computer stores an equivalent statement $f(x) = u(a, b, s) \mid_{arguments}$,
iff the arguments of $u,v,w$ are chosen such that $u,v,w$ approximate $g,h,k$ as follows:\\
\begin{tabular}{l l l l | c}
	Function & a & b & s & Approximates Function\\ \hline 
	$u$ & $1$ & $2$ & $v$ & $g$\\ \hline 
	$v$ & $1$ & $1$ & $w$ & $h$\\ \hline 
	$w$ & $1$ & $1$ & $x$ & $k$\\ \hline 	
\end{tabular}
\end{enumerate}

\item 
\label{differentiation}
%AD is eqivalent to the chain rule, when the computer has definitions for functions and partials
%that can be adapted to match the compostition function.
It follows from \ref{composition} that we can evaluate $\frac{d}{d x}f(x) \mid_{x_0}$ with the chain rule: \\ \\
$\frac{d}{d x}f(x) = \frac{d}{d x} \cdot g(h(k(x))) \mid_{x_0}$ \\
$\frac{d}{d x}f(x) =  \frac{d{g}}{d{h}} \cdot \frac{d{h}}{d{k}}
\cdot \frac{d{k}}{d{x}}\mid_{x_0}$ \\ \\
From the rest of \ref{given} it follows that we can approximate  $\frac{d}{d x}f(x) \mid_{x_0}$ by
specializing the computer's general forms of $u, v, w$ 
according to \ref{abcvalues}, with parameters $a, b$ chosen for each function and held as constant, and
evaluating, such that the total derivative our original function w.r.t $x$ where $x = x_0$ is approximated
by the partial derivative of our equivalent statement, $f(x) = u(a, b, s) \mid_{arguments}$,  w.r.t $s$ 
where $s = x_0$. \\

\end{enumerate}
\label{doingthework}
%This shows how information flows through the nested functions in one expression, like a table of contents
For completeness we write out the computer's steps to evaluate \ref{differentiation} under the conditions of
\ref{abcvalues} with a particular value of $s = x_0$, in equation format: \\ \\
$\frac{\partial}{\partial x}f(x) \mid _{x_0} = \frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot \frac{ds}{dx}$\\ \\
Repeating above in tabular format: \\ \\
%This shows the same information but is ordered to show progression, like chapters
\begin{tabular}{l l l}
Current Evaluation & $s$ Value & Partial Derivative \\ \hline
$w(a, b, s)\mid_{1, 1, s}$ & $x_0$ & $\frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
$v(a, b, s)\mid_{1, 1, s}$ & $w(a, b, s)\mid_{1, 1, x_0}$ & $\frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
$u(a, b, s)\mid_{1, 2, s}$ & $v(a, b, s)\mid_{1, 1, w(a, b, s)\mid_{1, 1, x_0}}$ & $\frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot 1$ \\
\end{tabular} \\ \\
%This explains what the equation and table above mean.
Because the computer knew the analytical forms of the partial derivatives of each of $u,v,w$ beforehand,
all it needed to do was:
\begin{enumerate}
\item to evaluate each of $u,v,w$ according to \ref{abcvalues}, in order from $w \rightarrow
v \rightarrow u$, 
\item to remember the values for $x_0$ and the output of each function evaluation besides  $u$, 
\item then to use $x_0$ and the output of the function evaluations as input for the corresponding partial
derivative function  evaluations, and
\item store the individual partials.
\end{enumerate}
Lastly, to compute the partial derivative of the entire composition function, the computer multiplies
the individual partials together in observance of the chain-rule.

Some things to note about AD are that no approximation of derivatives is being made because the analytical 
forms of the partials of the elementary math functions are defined alongside them. Accuracy of AD is then limited
by the precision of the AD system's definition of the elementary math functions and their partials.

\subsection{Tangent-stiffness Matrices}

This subsection gives a working definition of a tangent-stiffness matrix and background on a
Newton-Raphson method which uses a tangent-stiffness matrix to help solve nonlinear systems of
equations, similarly to the the solution method used by the software package used in the study.

\subsubsection{What is a Tangent-stiffness Matrix}

It is important to identify what a tangent-stiffness matrix is and what it is used for to help show
the motivation for the work discussed here. It is helpful to begin with the thought that a
tangent-stiffness matrix is a type of slope for vector valued functions of vector variables at
a single point in the space of the vector valued variable. A tangent-stiffness matrix is similar to the
"$m$" in the scalar function $y = mx + b$, except that $y$ and $x$ are vectors while $m$ is
a matrix. \\

In detail, a tangent-stiffness matrix (or operator) can be described as a collection of the first
order partial derivatives of a vector valued function w.r.t each degree of freedom of the vector
valued function for a given value of the vector variable.  The tangent-stiffness matrix comprises a
linear operator that can be used to transform a difference in the vector variable into a difference
in the value of the vector valued function via an inner product between the tangent-stiffness matrix
and the vector variable.  As an example in a solid mechanics system, suppose some dependent
variables represented force components of force vectors at nodes of discretization, and independent
variables represented displacement components for those same nodes. Suppose that as is possible in
peridynamic nodes each dependent variable is a function of all of the independent variables. For
this situation, the tangent-stiffness matrix would be the first derivatives of every force component
variable with respect to every displacement component variable, with the rows ordering the force
components and the columns ordering the displacement components.  \\

How is a tangent-stiffness matrix computed? The mathematical formula for a tangent stiffness matrix
can be expressed in indicial notation as:\\\\ $\frac{\partial F_i}{\partial X_j}\mid_{X_0,...,X_n}$
\\\\ $F_i$ is the i'th component of the vector valued function, $X_j$ is the j'th component of the
vector variable, and a particular value of the vector variable is chosen $X_0,...,X_n$. One then
evaluates the expression for each combination $i, j$ corresponding to an element at $row, column$ in
the tangent-stiffness matrix. By inspection, the elements of a tangent-stiffness matrix can be
estimated using the CTSE, AD and finite-difference techniques for functions $F:R^1 \rightarrow R^1$,
since taking partial derivatives entails holding all but a single independent variable of the
function constant, and each component of a vector valued function can be evaluated independently of
the other components. \\

For reference, expressions for calculating tangent stiffness matrices with CTSE, AD,
central-difference and forward-difference will be shown and explained.

\subsubsection{Solve a Nonlinear System with a Tangent-stiffness Matrix} \label{solve_a_system}
\remove[Michael]{ 
An accurate tangent-stiffness matrix is a main part of the Newton-Raphson method for iteratively
solving non-linear systems for equilibrium solutions. The method is derived by a linearisation using
Taylor expansion of the vector valued function modelling the system in a way reminiscent of how
forward-difference is derived from the Taylor expansion of a scalar valued function. Here the two
derivations are shown side by side: The difference between the two is that in the forward difference
derivation, it is assumed that the derivative term is unknown and can be determined from knowing
change in location, while in the Newton-Raphson derivation, it is assumed that the change in
location is unknown and can be determined from knowing the value of the derivative term.
Additionally we could derive a version of Newton Raphson for CTSE directly: The Newton Raphson
method is applied in the following manner: It is clear to see how if the tangent-stiffness matrix
used in the Newton Raphson method is inaccurate, predicted changes in location will be inaccurate.
However, non-linear functions don't have constant slopes by definition, meaning that a perfectly
accurate tangent-stiffness matrix at the current guess location is not a guarantee of a one step
solution path to the solution location since slope changes over the interval guess to updated guess.
}
The reader is referred to a succinct explanation of the Newton Raphson method, which shows the
role of a tangent-stiffness matrix or Jacobian in that solution method \cite[chap.
13]{young2009}.

\section{Methods and Materials} 
\subsection{Material: Perdynamics Code}
\subsubsection{\emph{Perdigm} and Peridynamics}
\subsubsection{Implementing Complex-Step in \emph{Peridigm}}

\subsection{The Comparative Study}

\subsubsection{Justification: Goals, Assumptions, and Metrics}

A goal of the study was to rank complex-step against forward-difference and central difference on
the basis of accuracy. AD was ommited from the comparison because it served as the standard of
accuracy for the other methods in the absence of appropriate analaytical forms for the Jacobian
associated with the nonlinear system used solved in the study. The assumption that AD is accurate
enough to serve as a standard is supported by AD's implementation as a computerized chain rule as
explained in \ref{ADsubsection}. For a concrete comparison of each of the methods including AD to a
known analytical Jacobian, refer to the corresponding author's website, specifically the C++
source-code example \emph{BeamAndSpringSystem}.

Because it makes no sense to compare tangent-stiffness matrices from different load steps, or from
different iterations, it was necessary to solve one load step and conclude each iteration within
that load step by updating guesses only from the the results of the AD method, and to subsequently
feed all four methods the same updated guess the following iteration. This decision precluded a
comparison of convergence rate, since if the methods were allowed to solve a problem at their
invidual pace, differences in accuracy would produce differences in guess updates and therefor the
number and nature of iterations perfomed. Different guesses from iterations started with different
previous guesses could not be data for a valid comparison of tangent-stiffness accuracy between
methods.  However a separate convergence rate comparison study could be done given some
modifications to the materials for the study described in this paper, and of course, the resources
to perform that study.  Instead, in this study, speed of iteration was measured because it could be
so done at the same time as Accuracy was being measured given a single tangent-stiffness
calculation, which was attractive due to limited resources.  Additionally, having each of the
methods physically operate in the same process, single or parallel, allowed the comparison of
tangent-stiffness matrices as loaded from random access memory rather than from the hard-disk,
giving the benefit of vastly greater speed and the benefit of simplicty compared to some other
solution speculatively involving dynamic file management on files which for one of the components of
the study would be on the order of $1XE11$ bytes, for which "vastly" is not hyperbole.

The other goal was to rank the four methods, that is including AD, on the basis of speed. It was
assumed that the order that each method was evaluated was unimportant, that evaluating each method
sucessively within each solver iteration did not affect their performance individually and that
speed of computation did not change with time. These assumptions allowed the test program to run the
same problem with each of the methods at effectively the same time and generate an equal volume of
data from each method. It was also assumed that the Jacobian matrix calculation time for each of the
methods did not vary based on what values the independent variables held, as they do from iteration
to iteration as the guessed equilibrium solution is updated. This assumption allowed calculation
time measurements to be averaged over all iterations within a solution attempt. The purpose of
averaging calculation time measurements was to informally address the extraneous variables of
evaluation order and computer system load due to other user's prcocesses or processes not associated
with the study.

The goals of collecting accuracy and speed measurements were achieved by developing and implementing
metrics within the simulation program used in the study. The metric used to measure the accuracy of
a tangent-stiffness matrix was the Frobenius norm of the element-wise difference between the
tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix
produced by the AD based method, given that both methods were set upon the same problem. The lower
the value of this metric, the closer the other method's Jacobian was to the AD Jacobian, and
therefore the more accurate the method was. The expression for the accuracy metric:\\ $D =
\sqrt{\sum_{i=0}^{n-1}\sum_{j=0}^{n-1} (J_{AD}(i,j) - J_{M}(i,j))^2}$\\ Where $D$ is distance, n is
the number of degrees of freedom, $J_{AD}$ is the Jacobian matrix produced by the AD based method, M
stands for other \emph{M}ethod, and the root of the summed squared element-wise differences between
the matrices is taken.  The metric used to compare the speed of the methods was the time in seconds
taken to resume execution in the calling method after instructing the implementation of the method
being evaluated to run. Another basis of comparison, but not quite a metric, called computational
efficiency and used in the study is based on the ratio of calculation time to number of Jacobian matrix
elements.

\subsubsection{Materials: Computers, Other Software}

\subsubsection{Methods: Reduction of Data}

\section{Results and Discussion}
\subsection{Speed Data}
\subsection{Accuracy Data}
\subsection{Efficiency Data}

\section{Conclusions and Further Work}
\subsection{Thoughts on Results}
\subsection{Thoughts on \emph{Perdigm}}
\subsection{Final Thoughts on Complex-Step}

\section{Unfinished table}

\begin{center}
    \begin{tabular}{  p{1cm}  p{1cm}  p{1cm}  p{1cm}  p{2cm}  p{3cm} p{2cm} }
    \hline
    \multicolumn{7}{c}{Full Test Matrix} \\ \hline
    Cores & Target & Rows & horizon (m) & Nonzero Rows & Calc. Time (s) & Accuracy Diff. (MPa)\\ 
    \multicolumn{5}{c}{} & CS, CD, FD, AD & CS, CD, FD \\ \hline 
	1 & 1E3 & 3000 & .3 & 1994454  & 3.5, 2.7, 1.7, 1.6 & 1.92E-10, 1.21E-4, .137 \\ \hline
	1 & 2E3 & 5400 & .24 & 4252230  & 6.2, 4.9, 3.2, 2.9 & 2.28E-10, 9.94E-4, .148 \\ \hline
	1 & 3E3 & 8526 & .21 & 7792128  & 11.2, 8.9, 5.7, 5.2 & 2.38E-10, 1.59E-4, .145\\ \hline
	1 & 4E3 & 12288 & .19 & 15389352  & 26.7, 21.0, 13.4, 12.2 & 2.33E-10, 4.61E-4, .12 \\ \hline
	1 & 6E3 & 18468 & .16 & 20027466  & 28.1, 22.2, 14.4, 13.1 & 2.76E-10, 1.05E-3, .145 \\ \hline
	1 & 8E3 & 24000 & .15 & 31388724  & 47.6, 37.7, 24.2, 21.9 & 2.64E-10, 1.65E-3, .133 \\ \hline
	1 & 1.2E4 & 33396 & .131 & 40060026  & 55.6, 44.1, 28.4, 25.9 & 3.03E-10, 1.92E-3, .148 \\ \hline
	1 & 1.6E4 & 50700 & .12 & 83778480  & 138.9, 109.6, 70.2, 64.0 & 3.63E-10, 1.64E-3, .123 \\ \hline
	1 & 3.2E4 & 96768 & .095 & 165140640  & 277.3, 218.1, 139.4, 126.5 & 3.26E-10, 2.18E-3, .128 \\ \hline
	32 & 1E6 & 3E6 & .03 & 1666968710  & 336.1, 277.7, 200.6, 233.1 & 6.21E-10, 1.52E-2, .176 \\ \hline
	64 & 1E6 & 3E6 & .03 & 1666962752  & 169.9, 140.7, 102.0, 119.7 & 6.20E-10, 1.50E-2, .177 \\ \hline
	96 & 1E6 & 3E6 & .03 & 1666760396  & 114.7, 95.0, 69.1, 79.7 & 6.18E-10, 1.50E-2, .177 \\ \hline
	128 & 1E6 & 3E6 & .03 & 1666892660  & 86.4, 71.8, 52.4, 58.8 & 6.16E-10, 1.47E-2, .177 \\ \hline
    \end{tabular}
\end{center}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
\appendix

\section{Extra}
\label{}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}         ==>>  [#]
%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

%% References with bibTeX database:

\bibliographystyle{elsarticle-num}
\bibliography{all}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use elsarticle-num.bst.

\end{document}
%%
%% End of file `elsarticle-template-num.tex'.

### 2013-09-02 jacobian_calc_CMAME.tex
Cleaned up methods, results table

@@
 \subsubsection{Justification: Goals, Assumptions, and Metrics} 
-
 A goal of the study was to rank complex-step against forward-difference and central difference on
@@
 accuracy for the other methods in the absence of appropriate analaytical forms for the Jacobian
-associated with the nonlinear system used solved in the study. The assumption that AD is accurate
+associated with the nonlinear system solved in the study. The assumption that AD is accurate
 enough to serve as a standard is supported by AD's implementation as a computerized chain rule as
@@
 
-Because it makes no sense to compare tangent-stiffness matrices from different load steps, or from
-different iterations, it was necessary to solve one load step and conclude each iteration within
-that load step by updating guesses only from the the results of the AD method, and to subsequently
-feed all four methods the same updated guess the following iteration. This decision precluded a
-comparison of convergence rate, since if the methods were allowed to solve a problem at their
-invidual pace, differences in accuracy would produce differences in guess updates and therefor the
-number and nature of iterations perfomed. Different guesses from iterations started with different
-previous guesses could not be data for a valid comparison of tangent-stiffness accuracy between
-methods.  However a separate convergence rate comparison study could be done given some
-modifications to the materials for the study described in this paper, and of course, the resources
-to perform that study.  Instead, in this study, speed of iteration was measured because it could be
-so done at the same time as Accuracy was being measured given a single tangent-stiffness
-calculation, which was attractive due to limited resources.  Additionally, having each of the
-methods physically operate in the same process, single or parallel, allowed the comparison of
-tangent-stiffness matrices as loaded from random access memory rather than from the hard-disk,
-giving the benefit of vastly greater speed and the benefit of simplicty compared to some other
-solution speculatively involving dynamic file management on files which for one of the components of
-the study would be on the order of $1XE11$ bytes, for which "vastly" is not hyperbole.
+Because it makes no sense to compare tangent-stiffness matrices from different problems, load steps,
+or from different iterations, it was necessary to solve one load step and conclude each iteration
+within that load step by updating displacement guess only from the the results of the AD method, and
+to subsequently feed all four methods the same updated guess the following iteration. This decision
+precluded a comparison of convergence rate, since if the methods were allowed to solve a problem at
+their individual pace, differences in accuracy would produce differences in guess updates and
+therefor the number and nature of iterations perfomed. Different guesses from iterations started
+with different previous guesses could not be data for a valid comparison of tangent-stiffness
+accuracy between methods.  However a separate convergence rate comparison study could be done given
+some modifications to the materials for the study described in this paper, and of course, the
+resources to perform that study.   Additionally, having each of the methods physically operate in
+the same process, single or parallel, allowed the comparison of tangent-stiffness matrices as loaded
+from random access memory rather than from the hard-disk.  Loading from RAM gives the benefit of
+vastly greater speed and the simpler programming compared to some other solution speculatively
+involving dynamic file management on files which for one of the components of the study would be on
+the order of $1XE10$ bytes for which "vastly" is not hyperbole. However, the price of running the
+methods together in the same process and avoiding the use of the hard-disk was that at least two
+tangent-stiffness matrices had to be stored in RAM during simulation, which meant that special
+high-memory compute nodes were needed for the $1E6$ peridynamic element problem series.
 
-The other goal was to rank the four methods, that is including AD, on the basis of speed. It was
+The other goal was to rank the four methods, that is including AD, on the basis of speed. Instead of
+convergence rate being measured, speed of iteration was measured because it could be so done at the
+same time as accuracy was being measured given a single tangent-stiffness calculation.  It was
 assumed that the order that each method was evaluated was unimportant, that evaluating each method
@@
 averaging calculation time measurements was to informally address the extraneous variables of
-evaluation order and computer system load due to other user's prcocesses or processes not associated
-with the study.
+evaluation order and computer system load due to other user's processes or processes not associated
+with the study. It begs the question that if it is assumed that the value of independent variables
+had no effect on iteration speed, than why couldn't each of the method have been run independently
+so that convergence data were recorded, so long as system being solved was the same? The answer is
+that this would have made accuracy comparisons invalid if measured concurrently with convergence
+rate, while it would cost additional time and resources if convergence rate were measured in
+serparate runs. Instead, while not part of the study, the example programs available on the
+corresponding author's website allow the reader to make a comparison of the methods for themselves
+on the basis of accuracy and convergence rate for two example nonlinear systems. In the examples,
+notice how convergence rate decreases when the step size selected for forward-difference or
+complex-step is set to larger numbers such that accuracy relative to AD is decreased. While likely
+not the best possible implementations of these methods, they each were coded with the same skill
+level such that a comparison of their relative strengths is reasonable from the standpoint of
+intuition.
 
-The goals of collecting accuracy and speed measurements were achieved by developing and implementing
-metrics within the simulation program used in the study. The metric used to measure the accuracy of
-a tangent-stiffness matrix was the Frobenius norm of the element-wise difference between the
-tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix
-produced by the AD based method, given that both methods were set upon the same problem. The lower
-the value of this metric, the closer the other method's Jacobian was to the AD Jacobian, and
-therefore the more accurate the method was. The expression for the accuracy metric:\\ $D =
-\sqrt{\sum_{i=0}^{n-1}\sum_{j=0}^{n-1} (J_{AD}(i,j) - J_{M}(i,j))^2}$\\ Where $D$ is distance, n is
-the number of degrees of freedom, $J_{AD}$ is the Jacobian matrix produced by the AD based method, M
-stands for other \emph{M}ethod, and the root of the summed squared element-wise differences between
-the matrices is taken.  The metric used to compare the speed of the methods was the time in seconds
-taken to resume execution in the calling method after instructing the implementation of the method
-being evaluated to run. Another basis of comparison, but not quite a metric, called computational
-efficiency and used in the study is based on the ratio of calculation time to number of Jacobian matrix
-elements.
+Once the selection of what to measure, why to measured it, and if it could be measured was
+determined, the task was to find a way to take measurements that satisfied our goals. The goals of collecting accuracy and
+speed measurements were achieved by developing and implementing metrics within the simulation
+program used in the study. The metric used to measure the accuracy of a tangent-stiffness matrix was
+the Frobenius norm of the element-wise difference between the tangent-stiffness matrix produced by
+the method being evaluated and the tangent-stiffness matrix produced by the AD based method, given
+that both methods were set upon the same problem. The lower the value of this metric, the closer the
+other method's Jacobian was to the AD Jacobian, and therefore the more accurate the method was. The
+expression for the accuracy metric:\\ $D = \sqrt{\sum_{i=0}^{n-1}\sum_{j=0}^{n-1} (J_{AD}(i,j) -
+J_{M}(i,j))^2}$\\ Where $D$ is distance, n is the number of degrees of freedom, $J_{AD}$ is the
+Jacobian matrix produced by the AD based method, M stands for other \emph{M}ethod, and the root of
+the summed squared element-wise differences between the matrices is taken.  The metric used to
+compare the speed of the methods was the time in seconds taken to resume execution in the calling
+method after instructing the implementation of the method being evaluated to run and complete itself. Another basis of
+comparison used in the study, but not quite a metric, called computational efficiency is based
+on the ratio of calculation time to number of Jacobian matrix elements.
 
 \subsubsection{Materials: Computers, Other Software}
+Mesh generation for the test problems run in the study was done using the software package $Cubit$
+developed at REFERENCE.
 
@@
 \section{Results and Discussion}
+
+\begin{center}
+    \begin{tabular}{c c c c }
+    \hline
+    \multicolumn{4}{c}{Averaged Results, Entire Test Matrix} \\ \hline
+    Cores & Jacobian Els. & Calc. Time (s) & Accuracy Diff. (MPa)\\ 
+    \multicolumn{2}{c}{} & CS, CD, FD, AD & CS, CD, FD \\ \hline 
+	1 & 1.99E6  & 3.5, 2.7, 1.7, 1.6 & 1.92E-10, 1.21E-4, .137 \\ \hline
+	1 & 4.25E6  & 6.2, 4.9, 3.2, 2.9 & 2.28E-10, 9.94E-4, .148 \\ \hline
+	1 & 7.79E6  & 11.2, 8.9, 5.7, 5.2 & 2.38E-10, 1.59E-4, .145\\ \hline
+	1 & 1.54E7  & 26.7, 21.0, 13.4, 12.2 & 2.33E-10, 4.61E-4, .12 \\ \hline
+	1 & 2.00E7  & 28.1, 22.2, 14.4, 13.1 & 2.76E-10, 1.05E-3, .145 \\ \hline
+	1 & 3.14E7  & 47.6, 37.7, 24.2, 21.9 & 2.64E-10, 1.65E-3, .133 \\ \hline
+	1 & 4.01E7  & 55.6, 44.1, 28.4, 25.9 & 3.03E-10, 1.92E-3, .148 \\ \hline
+	1 & 8.39E7  & 138.9, 109.6, 70.2, 64.0 & 3.63E-10, 1.64E-3, .123 \\ \hline
+	1 & 1.65E8  & 277.3, 218.1, 139.4, 126.5 & 3.26E-10, 2.18E-3, .128 \\ \hline
+	32 & 1.67E10  & 336.1, 277.7, 200.6, 233.1 & 6.21E-10, 1.52E-2, .176 \\ \hline
+	64 & 1.67E10  & 169.9, 140.7, 102.0, 119.7 & 6.20E-10, 1.50E-2, .177 \\ \hline
+	96 & 1.67E10  & 114.7, 95.0, 69.1, 79.7 & 6.18E-10, 1.50E-2, .177 \\ \hline
+	128 & 1.67E10  & 86.4, 71.8, 52.4, 58.8 & 6.16E-10, 1.47E-2, .177 \\ \hline
+    \end{tabular}
+\end{center}
+
 \subsection{Speed Data}
@@
 \subsection{Final Thoughts on Complex-Step}
-
-\section{Unfinished table}
-
-\begin{center}
-    \begin{tabular}{  p{1cm}  p{1cm}  p{1cm}  p{1cm}  p{2cm}  p{3cm} p{2cm} }
-    \hline
-    \multicolumn{7}{c}{Full Test Matrix} \\ \hline
-    Cores & Target & Rows & horizon (m) & Nonzero Rows & Calc. Time (s) & Accuracy Diff. (MPa)\\ 
-    \multicolumn{5}{c}{} & CS, CD, FD, AD & CS, CD, FD \\ \hline 
-	1 & 1E3 & 3000 & .3 & 1994454  & 3.5, 2.7, 1.7, 1.6 & 1.92E-10, 1.21E-4, .137 \\ \hline
-	1 & 2E3 & 5400 & .24 & 4252230  & 6.2, 4.9, 3.2, 2.9 & 2.28E-10, 9.94E-4, .148 \\ \hline
-	1 & 3E3 & 8526 & .21 & 7792128  & 11.2, 8.9, 5.7, 5.2 & 2.38E-10, 1.59E-4, .145\\ \hline
-	1 & 4E3 & 12288 & .19 & 15389352  & 26.7, 21.0, 13.4, 12.2 & 2.33E-10, 4.61E-4, .12 \\ \hline
-	1 & 6E3 & 18468 & .16 & 20027466  & 28.1, 22.2, 14.4, 13.1 & 2.76E-10, 1.05E-3, .145 \\ \hline
-	1 & 8E3 & 24000 & .15 & 31388724  & 47.6, 37.7, 24.2, 21.9 & 2.64E-10, 1.65E-3, .133 \\ \hline
-	1 & 1.2E4 & 33396 & .131 & 40060026  & 55.6, 44.1, 28.4, 25.9 & 3.03E-10, 1.92E-3, .148 \\ \hline
-	1 & 1.6E4 & 50700 & .12 & 83778480  & 138.9, 109.6, 70.2, 64.0 & 3.63E-10, 1.64E-3, .123 \\ \hline
-	1 & 3.2E4 & 96768 & .095 & 165140640  & 277.3, 218.1, 139.4, 126.5 & 3.26E-10, 2.18E-3, .128 \\ \hline
-	32 & 1E6 & 3E6 & .03 & 1666968710  & 336.1, 277.7, 200.6, 233.1 & 6.21E-10, 1.52E-2, .176 \\ \hline
-	64 & 1E6 & 3E6 & .03 & 1666962752  & 169.9, 140.7, 102.0, 119.7 & 6.20E-10, 1.50E-2, .177 \\ \hline
-	96 & 1E6 & 3E6 & .03 & 1666760396  & 114.7, 95.0, 69.1, 79.7 & 6.18E-10, 1.50E-2, .177 \\ \hline
-	128 & 1E6 & 3E6 & .03 & 1666892660  & 86.4, 71.8, 52.4, 58.8 & 6.16E-10, 1.47E-2, .177 \\ \hline
-    \end{tabular}
-\end{center}
 

### 2013-09-03 jacobian_calc_CMAME.tex
implementation section, other additions

@@
 \subsubsection{What is a Tangent-stiffness Matrix}
+\label{tsmat}
 
@@
 
-\section{Methods and Materials} 
-\subsection{Material: Perdynamics Code}
+\section{Methods and Materials} \subsection{Material: Perdynamics Code}
 \subsubsection{\emph{Perdigm} and Peridynamics} 
@@
 
+Complex step was implemented in \emph{Perdigm} as follows. First a way of persistently allocating
+memory needed to store imaginary values for nodal position and dilatation was determined. Persistent
+memory allocation is a very basic way to improve performance in general for a computer program. The way
+memory was allocated for similar vector and scalar variables used by the prior existing forward
+difference method was copied and new 'field spec' variables were added into the 'elastic' material
+model to store imaginary position coordinates and imaginary dilatation. Model evaluation computer
+code methods that were used in the course of applying the forward difference method were copied,
+renamed and re-written to follow the CTSE formula, which mainly involved changing the datatype of
+intermediate variables which were method-local and dynamically casting process-local variables to
+complex compatible datatypes. Additional modifications were for the purpose of collecting
+experimental data and are not part of complex-step itself. The mathematical difference between the
+original forward-difference and complex-step code was covered in \ref{tsmat}. The source code for
+implementing complex-step in \emph{Perdigim} can be found on the corresponding author's website. Because
+\emph{Peridigm} was written by computer scientists, many advanced techniques in C++, such as
+multiple virtual inheritance, pointer arithmetic and distributed data structures were used in the
+basic code and by necessity in the modified code. To see complex-step in a more narrow context, it
+is advised that readers also take a look at the examples of complex-step provided on the
+corresponding author's website.
+
 \subsection{The Comparative Study}
 
+The methods were compared by running two sets of test problems where each method would solve a
+problem concurrently. Each test problem simulated a $4 by .25 by .25$ meter block of metal
+undergoing tension along the axis parallel to the long dimension of the block. The test problems
+were set up in \emph{Perdigm} as qausistatic equillibrium problems, where the displacement boundary
+conditions used to apply tension were applied gradually, in 'load steps', and an equilibrium
+solution for each load step was achieved before applying the next load step. The material properties
+of the model were set at $8.0E-9 metric tonne/ mm^3$ density, $1.515E4 MPa$ bulk modulus, and
+$7.813E4 Mpa$ shear modulus. The peridynamic horizon used in a test problem was three times the node
+spacing for the corrisponding model mesh. This number was chosen out of experience to insure that
+every node could have at least one nonlocal connection along the diagonal of the initial node grid.
+
+A series of nine of these test problems were run, with the mesh discretization used in the model
+finer and finer with every test in the series. The aim of increasing the mesh complexity was to see
+if possible differences between the methods in terms of accuracy and speed would be affected. This
+series would comprise the single core runs. The specific parameters of these tests can be found in
+\emph{Peridigm} input xml files included in the archive for this paper which can be found in the
+corresponding author's website. These xml files allow users with appropriately modified versions of
+\emph{Peridigm} to reproduce the results found in this paper.
+
+Another series of four test problems were run, however this time the number of cores used to solve
+the problem was varied from test to test rather than the mesh refinement. While mesh refinement was
+not varied from test to test, it was set at a somewhat high $1E6$ perdidynamic nodes of
+discretization. This allowed the simulation computer to be challenged sufficiently that the effect
+of varying the number of cores used could be more clearly seen. The specific parameters of these
+tests can be found in \emph{Perdigm} input xml files also included in the archive for this paper on
+the corresponding author's website.
+
+What was recorded during the test runs as experimental data is the subject the following
+subsubsection. It is organized to eplain "what", "why" and "how" in that order.
+
 \subsubsection{Justification: Goals, Assumptions, and Metrics} 
+\label{JGAM}
 A goal of the study was to rank complex-step against forward-difference and central difference on
@@
 enough to serve as a standard is supported by AD's implementation as a computerized chain rule as
-explained in \ref{ADsubsection}. For a concrete comparison of each of the methods including AD to a
-known analytical Jacobian, refer to the corresponding author's website, specifically the C++
-source-code example \emph{BeamAndSpringSystem}.
+explained in \ref{ADsubsection}.
 
@@
 Once the selection of what to measure, why to measured it, and if it could be measured was
-determined, the task was to find a way to take measurements that satisfied our goals. The goals of collecting accuracy and
-speed measurements were achieved by developing and implementing metrics within the simulation
-program used in the study. The metric used to measure the accuracy of a tangent-stiffness matrix was
-the Frobenius norm of the element-wise difference between the tangent-stiffness matrix produced by
-the method being evaluated and the tangent-stiffness matrix produced by the AD based method, given
-that both methods were set upon the same problem. The lower the value of this metric, the closer the
-other method's Jacobian was to the AD Jacobian, and therefore the more accurate the method was. The
-expression for the accuracy metric:\\ $D = \sqrt{\sum_{i=0}^{n-1}\sum_{j=0}^{n-1} (J_{AD}(i,j) -
-J_{M}(i,j))^2}$\\ Where $D$ is distance, n is the number of degrees of freedom, $J_{AD}$ is the
-Jacobian matrix produced by the AD based method, M stands for other \emph{M}ethod, and the root of
-the summed squared element-wise differences between the matrices is taken.  The metric used to
-compare the speed of the methods was the time in seconds taken to resume execution in the calling
-method after instructing the implementation of the method being evaluated to run and complete itself. Another basis of
-comparison used in the study, but not quite a metric, called computational efficiency is based
-on the ratio of calculation time to number of Jacobian matrix elements.
+determined, the task was to find a way to take measurements that satisfied our goals. The goals of
+collecting accuracy and speed measurements were achieved by developing and implementing metrics
+within the simulation program used in the study. The metric used to measure the accuracy of a
+tangent-stiffness matrix was the Frobenius norm of the element-wise difference between the
+tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix
+produced by the AD based method, given that both methods were set upon the same problem. The lower
+the value of this metric, the closer the other method's Jacobian was to the AD Jacobian, and
+therefore the more accurate the method was. The expression for the accuracy metric:\\ $D =
+\sqrt{\sum_{i=0}^{n-1}\sum_{j=0}^{n-1} (J_{AD}(i,j) - J_{M}(i,j))^2}$\\ Where $D$ is distance, n is
+the number of degrees of freedom, $J_{AD}$ is the Jacobian matrix produced by the AD based method, M
+stands for other \emph{M}ethod, and the root of the summed squared element-wise differences between
+the matrices is taken.  The metric used to compare the speed of the methods was the time in seconds
+taken to resume execution in the calling method after instructing the implementation of the method
+being evaluated to run and complete itself. Another basis of comparison used in the study, but not
+quite a metric, called computational efficiency is based on the ratio of calculation time to number
+of Jacobian matrix elements.
 
 \subsubsection{Materials: Computers, Other Software}
-Mesh generation for the test problems run in the study was done using the software package $Cubit$
-developed at REFERENCE.
 
-\subsubsection{Methods: Reduction of Data}
+Mesh generation for the test problems run in the study was done using the software package
+\emph{Cubit} developed at REFERENCE. For the test series in which peridynamic node density was
+varied, journal scripts compatible with \emph{Cubit} were used. The rule used to select the
+arangement of nodes within the borders designated by simulated material block's dimensions was to
+manually choose a node spacing, fit the maximum whole number of nodes in a regular grid within the
+block, and then check whether or not this number of nodes matched the number of nodes desired for a
+particular test. The same process was used for the multicore series. An example journal script can
+be found on the corresponding author's website. For future work, it may be beneficial to automate
+and randomize this process for statistic reasons and for time efficiency.
+
+The test problems were run the \emph{Stampede} HPC cluster computer housed at \emph{TACC} (Texas
+Advanced Computing Center) based out of UT Austin. The single core test series was run using the
+normal queue compute nodes, while the multicore test series was run using one to four "large memory"
+nodes each having one terrabyte of local RAM. This large amount of memory was required for reasons
+explained in \ref{JGAM}.
+
+\subsubsection{Methods: Reduction of Data} The output produced by \emph{Perdigm} including the
+additional accuracy and speed measurements it was modified to provide was redirected from the
+console to text files by the resource manager, \emph{SLURM} (Simple Linux User Resource Manager).
+Directories and filenames were chosen so that data would be easy to sort by individual test run.
+Extraneous information was manually deleted from copies versions of the output text files so that
+they would be regular enough to be easily read by scripts written in \emph{Python}. These scripts
+averaged accuracy or speed data for a test corresponding to a particular peridynamic node density
+over all iterations for that run, and then plotted this averaged data as if it were a function of
+peridynamic node density as it varied from test to test in the series.  Similarly, for the multicore
+tests, data was averaged and plotted in much the same way, but as a funciton of number of cores used
+to solve the problem rather than as a function peridynamic node density which was held constant.
+These data reduction and plotting scripts can be found on the corresponding autor's website.
+
+Since samples were not repeated, because there was not resources for it, confidence intervals were
+not calculated. However, if results of iterations within the same test problem were considered
+samples of the same thing, then a confidence interval could be calculated. This assumption was made
+for the purpose of plotting reporting average values, but a confidence interval was not calculated.
+But, a cursory review of data showed no range overlap among the methods for accuracy or speed data
+from any test run of either series.
 
@@
 \section{Conclusions and Further Work}
+
 \subsection{Thoughts on Results}
+
 \subsection{Thoughts on \emph{Perdigm}}
+
+After the bulk of the study had been conducted, it was determined that the implementations of
+forward-difference, central-difference and by inherited code complex-step, in \emph{Peridigm} were
+not optimal. Each of these implementations included extensive looping over noncontiguous addresses
+in memory and unnecessary re-calculation of constant values. Unpublished preliminary results show
+that when calculations for the quantity 'zeta' are re-used and only perturbed and previously
+pertubed functions of independent variables are recomputed throughout the course of a Jacobian
+evaluation, complex-step performs at the speed of forward-difference. Because of this it is
+speculated that forward-difference and central-difference could be modified in nearly the same way to
+perform more quickly than AD. An enhanced performance implementation of forward-difference will
+eventually be made available on the corresponding author's website.
+
 \subsection{Final Thoughts on Complex-Step}
+
+Complex-step was shown to be highly accurate by the standards of this study, but as it is
+implemented in \emph{Peridigm} may not be any better than AD, and it is almost certainly slower for
+a naive implementations. When looking to the future of \emph{Peridigm}, Complex-step holds the
+advantage of relying on byte-copiable datatypes, while AD requires a complicated serialization and
+deserialization in order to function as byte-copiable. This is fine for copying in between MPI
+processes, but deserialization on an accelerator or GPU is very difficult, as this would require
+compiling a version of the AD library being used that were compatible with the accelerator or GPU
+being used. What this means is that if \emph{Peridigm} were to be reformulated to take advantage of
+accelerators like GPU's or similar, which is likely, given their performance and power efficiency
+advantages over traditional host CPU's, complex-step would become more attractive.
 

### 2013-09-04 jacobian_calc_CMAME.tex
Spelling errors

@@
 
-\section{Methods and Materials} \subsection{Material: Perdynamics Code}
+\section{Methods and Materials}
+
+\subsection{Material: Perdynamics Code}
+
+The following subsubsections identify the program modified to run the simulations for the
+comparative study and explain what modifications were necessary.
+
 \subsubsection{\emph{Perdigm} and Peridynamics} 
@@
 What was recorded during the test runs as experimental data is the subject the following
-subsubsection. It is organized to eplain "what", "why" and "how" in that order.
+subsubsection. It is organized to explain "what", "why" and "how" in that order.
 

### 2013-09-04 jacobian_calc_CMAME.tex
Added eqn. nums., added ts cal meth exprns

@@
 The scope of the application component of the study was limited to the Linear Peridynamic Solid
-material model implemented in the computational peridynamics code, \textit{Peridigm}. Identifying a
+material model implemented in the computational peridynamics code, \emph{Peridigm}. Identifying a
 specific application provided a practical framework for implementing the new method and solving
@@
 How is a tangent-stiffness matrix computed? The mathematical formula for a tangent stiffness matrix
-can be expressed in indicial notation as:\\\\ $\frac{\partial F_i}{\partial X_j}\mid_{X_0,...,X_n}$
-\\\\ $F_i$ is the i'th component of the vector valued function, $X_j$ is the j'th component of the
+can be expressed in indicial notation as:
+
+\begin{equation} \frac{\partial F_i}{\partial X_j}\mid_{X_0,...,X_n} \end{equation}
+
+Where $F_i$ is the i'th component of the vector valued function, $X_j$ is the j'th component of the
 vector variable, and a particular value of the vector variable is chosen $X_0,...,X_n$. One then
@@
 
+\subsubsection{Calculating Tangent-stiffness Matrices with Each Method}
 For reference, expressions for calculating tangent stiffness matrices with CTSE, AD,
 central-difference and forward-difference will be shown and explained.
+
+When the using the forward-difference technique the formula for calculating a tangent-stiffness
+matrix is:
+
+\begin{equation} K_{ij} = \frac{\partial F_i^{int}(x + \delta e_j) - F_i^{int}(x)}{\delta}
+\end{equation}
+
+Where $K_{ij}$ is the index notation representation of an element of the tangent-stiffness matrix at
+row $i$, column $j$. The expression says that the partial derivative of the $i$'th component of $F$
+w.r.t the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
+and a denomenator. The first of the 'two terms' is $F_i$ evaluated at $x$ where
+the $j$'th component of $x$ has been perturbed by an very small amount in the direction of the
+positive $x_j$ axis. The second of the 'two terms' is $F_I$ evaluated at an unperturbed $x$. The
+denominator is the magnitude of perturbation called "probe distance" {ADDAGIO MANUAL REF}
+
+When using the central-difference technique the formula for calculating a tangent-stiffness matrix
+is:
+
+\begin{equation} K_{ij} = \frac{\partial F_i^{int}(x + \delta e_j) - F_i^{int}(x - \delta
+e_j)}{2\delta} \end{equation}
+
+The distiction between the central difference-method and the forward-difference method
+tangent-stiffness calculation formula is that the partial difference is now between two perturbed
+function evaluations, rather than between an unperturbed and perturbed function evaluation as in
+forward-difference. Additionally, the minuend involves a perturbation of $x$ in the direction opposite to
+the direction of the $j$'th component of $x$.
+
+When using the comple-step technique, which is based on CTSE, the formula for calculating a
+tangent-stiffness matrix is:
+
+\begin{equation} K_{ij} = \frac{imag(F_i^{int}(x + \delta e_j^{imag}))}{\delta} \end{equation}
+
+Where our model $F$ is treated as function of complex variables and $x$ is treated as a vector of
+complex values. The expressions says that the real valued partial derivative of the $i$'th component of $F$
+w.r.t the $j$'th component of $x$ is equal to the imaginary component of $F_i$ evaluated at $x$
+where $x$ is perturbed by a very small amount in the direciton of the imaginary axis corresponding
+to the $j$'th component of $x$ all divided by the magnitude of the perturbation. There is no partial
+notation because a partial difference is not taken.
+
+When using automatic-differentiation to calculate a tangent-stiffness matrix, one only needs to
+follow the definition of the tangent-stiffness matrix and issue the correct commands to the AD
+system. The formula is just as expected:
+
+\begin{equation} K_{ij} = \frac{\partial F_i^{int}}{\partial x_j} \end{equation}
 
@@
 the value of this metric, the closer the other method's Jacobian was to the AD Jacobian, and
-therefore the more accurate the method was. The expression for the accuracy metric:\\ $D =
-\sqrt{\sum_{i=0}^{n-1}\sum_{j=0}^{n-1} (J_{AD}(i,j) - J_{M}(i,j))^2}$\\ Where $D$ is distance, n is
-the number of degrees of freedom, $J_{AD}$ is the Jacobian matrix produced by the AD based method, M
-stands for other \emph{M}ethod, and the root of the summed squared element-wise differences between
-the matrices is taken.  The metric used to compare the speed of the methods was the time in seconds
-taken to resume execution in the calling method after instructing the implementation of the method
-being evaluated to run and complete itself. Another basis of comparison used in the study, but not
-quite a metric, called computational efficiency is based on the ratio of calculation time to number
-of Jacobian matrix elements.
+therefore the more accurate the method was. The expression for the accuracy metric:
+
+\begin{equation} D = \sqrt{\sum_{i=0}^{n-1}\sum_{j=0}^{n-1} (J_{AD}(i,j) - J_{M}(i,j))^2}
+\end{equation}
+
+Where $D$ is distance, n is the number of degrees of freedom, $J_{AD}$ is the Jacobian matrix
+produced by the AD based method, M stands for other \emph{M}ethod, and the root of the summed
+squared element-wise differences between the matrices is taken.  The metric used to compare the
+speed of the methods was the time in seconds taken to resume execution in the calling method after
+instructing the implementation of the method being evaluated to run and complete itself. Another
+basis of comparison used in the study, but not quite a metric, called computational efficiency is
+based on the ratio of calculation time to number of Jacobian matrix elements.
 

### 2013-09-04 jacobian_calc_CMAME.tex
Semi-major rewrite of Introduction

--- 2013-09-04 jacobian_calc_CMAME.tex
@@
 \author{John T. Foster\corref{cor1}}
+\cortext[cor1]{Corresponding Author: john.foster@utsa.edu}
 \author{Harry R. Millwater}
@@
 %% \ead[url]{home page}
-\cortext[cor1]{Corresponding Author: john.foster@utsa.edu}
 %% \address{Address\fnref{label3}}
@@
 %% main text
-\section{Acknowledgements}
-This work was funded in part by a grant from the United States Airforce Office of Scientific
-Research. This work is affiliated with the Center for Simulation Visualization and Real-Time
-Prediction at the University of Texas at San Antonio.
 
@@
 
-In order to maintain the quadratic convergence properties of the first-order Newton's method in
-quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate,
-algorithmically consistent tangent-stiffness matrices. For an extremely small class of nonlinear
-material models, these consistent tangent-stiffness operators can be derived analytically; however,
-most often in practice, they are found through numerical approximation of derivatives.
+In order to maintain the quadratic convergence properties of the first-order Newton's method in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
+For an extremely small class of nonlinear material models, these consistent tangent-stiffness operators can be derived analytically; however, most often in practice, they are found through numerical approximation of derivatives.
 
-A goal of this comparative study, was to develop and evaluate a new, accurate, and practical method
-for calculating tangent-stiffness matrices against established methods.  This new method is based on
-a complex number Taylor series expansion and refered to as CTSE or the ``complex-step'' method.  The
-distinction of `accurate' is defined by comparison of the new method to popular finite differencing
-techniques used for computing tangent-stiffness matrices and with the exact algorithmically
-consistent derivatives computed with \emph{automatic differentiation}.
+A goal of this study was to develop and evaluate a new, accurate, and practical method for calculating tangent-stiffness matrices against established methods.
+This new method is based on a complex number Taylor series expansion and referred to as CTSE or the ``complex-step'' method.
+The
+distinction of `accurate' is defined by comparison of the new method to popular finite differencing techniques used for computing tangent-stiffness matrices and with the exact algorithmically consistent derivatives computed with \emph{automatic differentiation}.
 
-It was thought that in addition to comparative information regarding the new method, comparisons of
-exclusively finite-difference, central-difference and automatic differentiation could prove valuable
-for developers or analysts not necessarily considering complex-step but needing data to help decide
-whether or not to invest in automatic-differentiation vs. finite-difference.  A comparison and
-discrimination of the methods was to be achieved through application and the measurement of results
-in-situ. The results would largely be treated as deterministic consequences of method choice. There
-was not an attempt made to mathematically prove the suitability of one method over another of the
-methods. The study was retrospective, as explained in section {SECTION} of the instant paper.
+Comparative data regarding the accuracy and compute cycle timing for the complex step, finite-difference, central-difference, and automatic differentiation methods are included to aide developers and analysts in computational mechanics code design who are faced with implementing a general method for tangent stiffness matrix calculation.
+A comparison and discrimination of the methods was achieved through in-situ instrumentation of a massively parallel computational mechanics code.
 
-The scope of the application component of the study was limited to the Linear Peridynamic Solid
-material model implemented in the computational peridynamics code, \textit{Peridigm}. Identifying a
-specific application provided a practical framework for implementing the new method and solving
-engineering problems to generate the data needed to compare the methods. In particular,
-\emph{Peridigm} was chosen cause it combined several helpful characteristics: The reliance on
-Newton's method and tangent-stiffness matrices for solving quasi-static implicit problems, prior
-inclusion of finite-difference, central-difference and automatic-differentiation based methods
-needed for comparison to a new method that would be developed, and being largely based in the Ansi
-standard version of the popular C++ programming language.
+The scope of the application component of the study was limited to a single material model implemented in the computational peridynamics code, \textit{Peridigm} \cite{peridigm}.
+Identifying a specific application provided a practical framework for implementing the new method and solving engineering problems to generate the data needed to compare the methods.
+In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of finite-difference, central-difference, and automatic-differentiation methods needed for comparison to the new complex-step methed was developed, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
 
-The aim of this paper is to introduce new information, but also to compile prior art to serve as a
-general reference for solving non-linear systems. \remove[Michael]{The former is mainly intended for
-an audience of experienced researchers and professionals who are seeking novel tools, the later
-is intended for an audience of undergraduate to graduate level students who would benefit from a
-source detailing basic methodology.  According to this aim, concepts are presented in multiple
-levels of detail to accommodate both a broad and deep audience.} After giving background on the
-\add[Michael]{possibly less familiar} differentiation techniques underlying the methods to be
-discussed here, presented in this paper is: a description of tangent-stiffness matrices and how to
-use them to solve non-linear systems, detailed directions for producing tangent-stiffness matrices
-with each the methods identified above \remove[Michael]{preceded by background as it becomes
-necessary}, a description and justification of a new method, called 'complex-step' for calculating
-tangent-stiffness matrices, a description of implementing complex-step in the software package used
-in the comparative study, a description of the parameters of the comparative study conducted to rank
-the methods, a presentation and analysis of the results of the comparative study, and
-conclusions and thoughts on potential future work. Finally, the reader is referred to the
-corresponding author's website for c++ source-code and data necessary to reproduce the results
-presented here. Additionally included on the website is a C+ template library with classes for
-solving non-linear systems using the techniques dicussed here, as well as two validation and
-verification example problemsi which show how to use the library.
+The aim of this paper is not only to introduce the new complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques and as a general reference for solving non-linear systems with Newton's method.
+After presenting background information on the underlying methods to be discussed, presented in this paper are: a description of tangent-stiffness matrices and how to use them to solve non-linear systems, detailed directions for producing tangent-stiffness matrices with each the methods identified above, a description and justification of a the new complex-step method for calculating tangent-stiffness matrices, a description of implementing complex-step in the \emph{Peridigm} software, a description of the quantities-of-interest used to rank the methods, a presentation and analysis of the results of the comparative study, and conclusions and thoughts on potential future work.
+Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for C++ source-code and data necessary to reproduce the results presented here.
+Additionally included on the website is a C++ template library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
 
@@
 \subsection{Final Thoughts on Complex-Step}
-
 Complex-step was shown to be highly accurate by the standards of this study, but as it is
@@
 %% appendix sections are then done as normal sections
-\appendix
 
-\section{Extra}
-\label{}
+\section{Acknowledgements}
+This work was funded in part by grants from the United States Air Force Office of Scientific
+Research grant number XXXX and National Energy Technology Laboratory grant number XXXXX. 
+\label{sec:ack}
 

### 2013-09-05 jacobian_calc_CMAME.tex
Adding figures
Cleaned up repo
Fix up first figure
Added results figures, finalized some removes

@@
 %% use the graphics package for simple commands
-%% \usepackage{graphics}
+%%\usepackage{graphics}
 %% or use the graphicx package for more complicated commands
-%% \usepackage{graphicx}
+%%\usepackage{graphicx}
 %% or use the epsfig package if you prefer to use the old commands
 %% \usepackage{epsfig}
+%%\usepackage{color}
+%%\usepackage{tkz-base}
+\usepackage{pgfplots}
 
@@
 contemporary to the study, complex-step was also the slowest method among the four for each
-datum for each run. A mathematical definition of \change[Michael]{the methods}{complex-step and
-and automatic differentiation}, description of the computer implementations and justification
+datum for each run. A mathematical definition of complex-step and
+and automatic differentiation, description of the computer implementations and justification
 of the comparative study precede results for clarity. The intended audience of this paper
-includes researchers, professionals and students, \change[Michael]{therefore conceptually explicit language
-appealing to multiple levels of understanding is used}{therefor readers are directed to
+includes researchers, professionals and students, therefor readers are directed to
 selected background sources in the text for well-known concepts while less familiar concepts
 are derived in-paper with an attempt made to lay steps out plainly for those in disparate fields
-or who are new to their studies.}  \end{abstract}
+or who are new to their studies.  \end{abstract}
 
@@
 tangent-stiffness matrix calculation methods compared in the study. Since first-order finite
-difference techniques are considered well known, they are not described where. Instead the
+difference techniques are considered well known, they are not described here. Instead the
 reader is referred to \cite[Chap. 4.1.3]{chapra2010}.
@@
 \subsubsection{Basis of New Method, 'Complex Step', in CTSE}  
-\note[Michael]{TODO: cite lyness and moler, complete section}
+\note[Michael]{Will be replaced by Dr. Millwater's background section}
 
@@
 subtractive-cancellation.
-
-\remove[Michael]{
-\subsubsection{First-order Finite-difference}
-
-The types of first-order finite-difference that will be discussed are forward-difference and
-central-difference. CTSE and these finite-difference techniques are similar in that they have been
-developed by algebraically solving approximated Taylor series expansions for first derivative terms.
-}
 
@@
 \subsubsection{Solve a Nonlinear System with a Tangent-stiffness Matrix} \label{solve_a_system}
-\remove[Michael]{ 
-An accurate tangent-stiffness matrix is a main part of the Newton-Raphson method for iteratively
-solving non-linear systems for equilibrium solutions. The method is derived by a linearisation using
-Taylor expansion of the vector valued function modelling the system in a way reminiscent of how
-forward-difference is derived from the Taylor expansion of a scalar valued function. Here the two
-derivations are shown side by side: The difference between the two is that in the forward difference
-derivation, it is assumed that the derivative term is unknown and can be determined from knowing
-change in location, while in the Newton-Raphson derivation, it is assumed that the change in
-location is unknown and can be determined from knowing the value of the derivative term.
-Additionally we could derive a version of Newton Raphson for CTSE directly: The Newton Raphson
-method is applied in the following manner: It is clear to see how if the tangent-stiffness matrix
-used in the Newton Raphson method is inaccurate, predicted changes in location will be inaccurate.
-However, non-linear functions don't have constant slopes by definition, meaning that a perfectly
-accurate tangent-stiffness matrix at the current guess location is not a guarantee of a one step
-solution path to the solution location since slope changes over the interval guess to updated guess.
-}
 The reader is referred to a succinct explanation of the Newton Raphson method, which shows the
@@
 \subsection{Speed Data}
+
+\begin{figure}[h]
+  \centering
+  \scalebox{.7}{\input{./figs/serial_speed.pgf}}
+  \caption{Serial test series speed measurements.}
+  \label{fig:serial_speed}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \scalebox{.7}{\input{./figs/multi_speed.pgf}}
+  \caption{Multicore test series speed measurements.}
+  \label{fig:multi_speed}
+\end{figure}
+
 \subsection{Accuracy Data}
+
+\begin{figure}[h]
+  \centering
+  \scalebox{.7}{\input{./figs/serial_accuracy.pgf}}
+  \caption{Serial test series speed measurements.}
+  \label{fig:serial_accuracy}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \scalebox{.7}{\input{./figs/multi_accuracy.pgf}}
+  \caption{Multicore test series accuracy measurements.}
+  \label{fig:multi_accuracy}
+\end{figure}
+
 \subsection{Efficiency Data}
+
+\begin{figure}[h]
+  \centering
+  \scalebox{.7}{\input{./figs/serial_speed_rel.pgf}}
+  \caption{Serial test series efficiency measurements.}
+  \label{fig:serial_efficiency}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \scalebox{.7}{\input{./figs/multi_speed_rel.pgf}}
+  \caption{Multicore test series efficiency measurements.}
+  \label{fig:multi_efficiency}
+\end{figure}
+
+%%\includegraphics[width=\textwidth]{serial_speed_rel.png}
+%%\includegraphics[width=\textwidth]{multi_speed_rel.png}
 

### 2013-09-09 jacobian_calc_CMAME.tex
grammar edits, trimming down flourid language

@@
 Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for C++ source-code and data necessary to reproduce the results presented here.
-Additionally included on the website is a C++ template library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
+Additionally included on the website is a C++ library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
 
@@
 a single point in the space of the vector valued variable. A tangent-stiffness matrix is similar to the
-"$m$" in the scalar function $y = mx + b$, except that $y$ and $x$ are vectors while $m$ is
+"$m$" in the function $y = mx + b$, except that $y$, $x$ and $b$ are vectors while $m$ is
 a matrix. \\
 
-In detail, a tangent-stiffness matrix (or operator) can be described as a collection of the first
+In detail, a tangent-stiffness matrix can be described as a collection of the first
 order partial derivatives of a vector valued function w.r.t each degree of freedom of the vector
@@
 this situation, the tangent-stiffness matrix would be the first derivatives of every force component
-variable with respect to every displacement component variable, with the rows ordering the force
-components and the columns ordering the displacement components.  \\
+variable with respect to every displacement component variable, with the rows corresponding to the force
+components and the columns corresponding to the displacement components.  \\
 
@@
 For reference, expressions for calculating tangent stiffness matrices with CTSE, AD,
-central-difference and forward-difference will be shown and explained.
+central-difference and forward-difference will be shown and explained. The expressions for AD,
+central-difference and forward-difference are taken from {ADDAGIO MANUAL REF}, while the expression
+for CTSE is original.
 
@@
 memory needed to store imaginary values for nodal position and dilatation was determined. Persistent
-memory allocation is a very basic way to improve performance in general for a computer program. The way
-memory was allocated for similar vector and scalar variables used by the prior existing forward
+memory allocation is a very basic way to improve performance in general for a computer program. The
+way memory was allocated for similar vector and scalar variables used by the prior existing forward
 difference method was copied and new 'field spec' variables were added into the 'elastic' material
-model to store imaginary position coordinates and imaginary dilatation. Model evaluation computer
-code methods that were used in the course of applying the forward difference method were copied,
-renamed and re-written to follow the CTSE formula, which mainly involved changing the datatype of
-intermediate variables which were method-local and dynamically casting process-local variables to
-complex compatible datatypes. Additional modifications were for the purpose of collecting
-experimental data and are not part of complex-step itself. The mathematical difference between the
-original forward-difference and complex-step code was covered in \ref{tsmat}. The source code for
-implementing complex-step in \emph{Perdigim} can be found on the corresponding author's website. Because
-\emph{Peridigm} was written by computer scientists, many advanced techniques in C++, such as
-multiple virtual inheritance, pointer arithmetic and distributed data structures were used in the
-basic code and by necessity in the modified code. To see complex-step in a more narrow context, it
-is advised that readers also take a look at the examples of complex-step provided on the
-corresponding author's website.
+model to store imaginary position coordinates and imaginary dilatation. Because of \emph{Peridigm}'s
+dependence on the \emph{Epetra} package of the \emph{Trilinos} library, 'field spec' variables could
+not be simply be declared as having a complex datatype. Program methods that were used in the
+course of applying the forward-difference method were copied, renamed and re-written to follow the
+CTSE formula, which involved changing the datatype of intermediate variables which were locally
+scoped and dynamically casting persistent variables, like the 'field spec' variables, to complex
+datatypes where necessary. Additional modifications were for the purpose of collecting experimental
+data and are not part of complex-step itself. The mathematical difference between the original
+forward-difference and complex-step code was covered in \ref{tsmat}. The source code for
+implementing complex-step in \emph{Perdigim} can be found on the corresponding author's website.
+Because \emph{Peridigm} was written by computer scientists rather than mechanical engineers
+primarily, many advanced techniques in C++ such as multiple virtual inheritance, pointer arithmetic
+and distributed data structures were used in the basic code and by necessity in the modified code.
+To see complex-step in a more narrow context, it is advised that readers also take a look at the
+examples of complex-step and the other methods provided on the corresponding author's website.
 
@@
 conditions used to apply tension were applied gradually, in 'load steps', and an equilibrium
-solution for each load step was achieved before applying the next load step. The material properties
-of the model were set at $8.0E-9 metric tonne/ mm^3$ density, $1.515E4 MPa$ bulk modulus, and
-$7.813E4 Mpa$ shear modulus. The peridynamic horizon used in a test problem was three times the node
-spacing for the corrisponding model mesh. This number was chosen out of experience to insure that
-every node could have at least one nonlocal connection along the diagonal of the initial node grid.
+solution for each load step was achieved before applying the next load step. The purpose of applying
+the load in gradual steps is a widely used technique to ensure the Newton-Raphson method is always
+guessing near the actual solution. The material properties of the model were set at $8.0E-9 metric
+tonne/ mm^3$ density, $1.515E4 MPa$ bulk modulus, and $7.813E4 Mpa$ shear modulus. The peridynamic
+horizon used in a test problem was three times the node spacing for the corrisponding model mesh.
+This value for horizon was chosen out of experience to insure that every node could have at least
+one nonlocal connection along the diagonal of the node grid.
 
-A series of nine of these test problems were run, with the mesh discretization used in the model
-finer and finer with every test in the series. The aim of increasing the mesh complexity was to see
-if possible differences between the methods in terms of accuracy and speed would be affected. This
-series would comprise the single core runs. The specific parameters of these tests can be found in
-\emph{Peridigm} input xml files included in the archive for this paper which can be found in the
-corresponding author's website. These xml files allow users with appropriately modified versions of
-\emph{Peridigm} to reproduce the results found in this paper.
+A series of nine of these single core test problems were run, with the mesh discretization used in
+the model finer and finer with every test in the series. The aim of increasing the mesh complexity
+was to see if possible differences between the methods in terms of accuracy and speed would arise
+and trend based on model complexity. This series would comprise the single core runs. The specific
+parameters of these tests can be found in \emph{Peridigm} input xml files included in the archive
+for this paper which can be found in the corresponding author's website. These xml files allow users
+with appropriately modified versions of \emph{Peridigm} to reproduce the results found in this
+paper.
 
@@
 discretization. This allowed the simulation computer to be challenged sufficiently that the effect
-of varying the number of cores used could be more clearly seen. The specific parameters of these
-tests can be found in \emph{Perdigm} input xml files also included in the archive for this paper on
-the corresponding author's website.
+of varying the number of cores on speed and accuracy between the methods used could be more clearly
+seen. The specific parameters of these tests can be found in \emph{Perdigm} input xml files also
+included in the archive for this paper on the corresponding author's website.
 
@@
 
-\subsubsection{Justification: Goals, Assumptions, and Metrics} 
-\label{JGAM}
-A goal of the study was to rank complex-step against forward-difference and central difference on
-the basis of accuracy. AD was ommited from the comparison because it served as the standard of
-accuracy for the other methods in the absence of appropriate analaytical forms for the Jacobian
-associated with the nonlinear system solved in the study. The assumption that AD is accurate
-enough to serve as a standard is supported by AD's implementation as a computerized chain rule as
-explained in \ref{ADsubsection}.
+\subsubsection{Justification: Goals, Assumptions, and Metrics} \label{JGAM} As stated in the
+previous section a goal of the study was to rank complex-step against forward-difference and central
+difference on the basis of accuracy. AD was ommited from the comparison because it served as the
+standard of accuracy for the other methods in the absence of appropriate analaytical forms for the
+Jacobian associated with the nonlinear system solved in the study. The assumption that AD is
+accurate enough to serve as a standard is supported by AD's implementation as a computerized chain
+rule as explained in \ref{ADsubsection}.
 
-Because it makes no sense to compare tangent-stiffness matrices from different problems, load steps,
-or from different iterations, it was necessary to solve one load step and conclude each iteration
-within that load step by updating displacement guess only from the the results of the AD method, and
-to subsequently feed all four methods the same updated guess the following iteration. This decision
-precluded a comparison of convergence rate, since if the methods were allowed to solve a problem at
-their individual pace, differences in accuracy would produce differences in guess updates and
-therefor the number and nature of iterations perfomed. Different guesses from iterations started
-with different previous guesses could not be data for a valid comparison of tangent-stiffness
-accuracy between methods.  However a separate convergence rate comparison study could be done given
-some modifications to the materials for the study described in this paper, and of course, the
-resources to perform that study.   Additionally, having each of the methods physically operate in
-the same process, single or parallel, allowed the comparison of tangent-stiffness matrices as loaded
-from random access memory rather than from the hard-disk.  Loading from RAM gives the benefit of
-vastly greater speed and the simpler programming compared to some other solution speculatively
-involving dynamic file management on files which for one of the components of the study would be on
-the order of $1XE10$ bytes for which "vastly" is not hyperbole. However, the price of running the
-methods together in the same process and avoiding the use of the hard-disk was that at least two
-tangent-stiffness matrices had to be stored in RAM during simulation, which meant that special
-high-memory compute nodes were needed for the $1E6$ peridynamic element problem series.
+Because it makes no sense to compare tangent-stiffness matrices from different problems, load steps
+that started with difference guesses, or from different iterations, it was necessary to solve one
+load step and conclude each iteration within that load step by updating the displacement guess only
+from the results of the AD method and to subsequently feed all four methods the same updated guess
+the following iteration. This decision precluded a comparison of convergence rate, since if the
+methods were allowed to solve a problem at their individual pace, differences in accuracy would
+produce differences in guess updates and therefor the number and nature of iterations perfomed.
+Different guesses from iterations started with different previous guesses could not be data for a
+valid comparison of tangent-stiffness accuracy between methods.  However a separate convergence rate
+comparison study could be done given some modifications to the materials for the study described in
+this paper, and of course, the resources to perform that study.   Additionally, having each of the
+methods physically operate in the same process, single or one of associated parallel, allowed the
+comparison of tangent-stiffness matrices as loaded from random access memory rather than from the
+hard-disk.  Loading from RAM gives the benefit of vastly greater speed and the simpler programming
+compared to some other solution speculatively involving dynamic file management on files which for
+one of the components of the study would be on the order of $1XE10$ bytes. However, the price of
+running the methods together in the same process and avoiding the use of the hard-disk was that at
+least two tangent-stiffness matrices had to be stored in RAM during simulation, which meant that
+special high-memory compute nodes were needed for the $1E6$ peridynamic element problem series.
 
@@
 evaluation order and computer system load due to other user's processes or processes not associated
-with the study. It begs the question that if it is assumed that the value of independent variables
-had no effect on iteration speed, than why couldn't each of the method have been run independently
+with the study. The question is begged: if it is assumed that the value of independent variables
+had no effect on iteration speed, then why couldn't each of the method have been run independently
 so that convergence data were recorded, so long as system being solved was the same? The answer is
@@
 notice how convergence rate decreases when the step size selected for forward-difference or
-complex-step is set to larger numbers such that accuracy relative to AD is decreased. While likely
-not the best possible implementations of these methods, they each were coded with the same skill
-level such that a comparison of their relative strengths is reasonable from the standpoint of
-intuition.
+complex-step is set to larger numbers such that accuracy relative to AD is decreased.
 
-Once the selection of what to measure, why to measured it, and if it could be measured was
-determined, the task was to find a way to take measurements that satisfied our goals. The goals of
-collecting accuracy and speed measurements were achieved by developing and implementing metrics
-within the simulation program used in the study. The metric used to measure the accuracy of a
-tangent-stiffness matrix was the Frobenius norm of the element-wise difference between the
-tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix
-produced by the AD based method, given that both methods were set upon the same problem. The lower
-the value of this metric, the closer the other method's Jacobian was to the AD Jacobian, and
-therefore the more accurate the method was. The expression for the accuracy metric:
+Once the selection of what to measure was determined, the task was to find a way to take measurements
+that satisfied our goals. The goals of collecting accuracy and speed measurements were achieved by
+developing and implementing metrics within the simulation program used in the study. The metric used
+to measure the accuracy of a tangent-stiffness matrix was the Frobenius norm of the element-wise
+difference between the tangent-stiffness matrix produced by the method being evaluated and the
+tangent-stiffness matrix produced by the AD based method, given that both methods were set upon the
+same problem. The lower the value of this metric, the closer the other method's Jacobian was to the
+AD Jacobian, and therefore the more accurate the method was. The expression for the accuracy metric:
 
@@
 After the bulk of the study had been conducted, it was determined that the implementations of
-forward-difference, central-difference and by inherited code complex-step, in \emph{Peridigm} were
+forward-difference, central-difference and by virtue of reusing code, complex-step, in \emph{Peridigm} were
 not optimal. Each of these implementations included extensive looping over noncontiguous addresses
-in memory and unnecessary re-calculation of constant values. Unpublished preliminary results show
+in memory and unnecessary re-calculation of constant values. The approach used does minimize
+temporary memory use, but is very wasteful of cpu and memory bandwidth. Unpublished preliminary results show
 that when calculations for the quantity 'zeta' are re-used and only perturbed and previously
 pertubed functions of independent variables are recomputed throughout the course of a Jacobian
-evaluation, complex-step performs at the speed of forward-difference. Because of this it is
+evaluation, complex-step performs at the speed of unmodified forward-difference. Because of this it is
 speculated that forward-difference and central-difference could be modified in nearly the same way to
@@
 
-\subsection{Final Thoughts on Complex-Step}
-Complex-step was shown to be highly accurate by the standards of this study, but as it is
-implemented in \emph{Peridigm} may not be any better than AD, and it is almost certainly slower for
-a naive implementations. When looking to the future of \emph{Peridigm}, Complex-step holds the
-advantage of relying on byte-copiable datatypes, while AD requires a complicated serialization and
-deserialization in order to function as byte-copiable. This is fine for copying in between MPI
-processes, but deserialization on an accelerator or GPU is very difficult, as this would require
-compiling a version of the AD library being used that were compatible with the accelerator or GPU
-being used. What this means is that if \emph{Peridigm} were to be reformulated to take advantage of
-accelerators like GPU's or similar, which is likely, given their performance and power efficiency
-advantages over traditional host CPU's, complex-step would become more attractive.
+\subsection{Final Thoughts on Complex-Step} Complex-step was shown to be highly accurate by the
+standards of this study, but as it is implemented in \emph{Peridigm} may not be any better than AD,
+and it is almost certainly slower for a naive implementation. When looking to the future of
+\emph{Peridigm}, Complex-step holds the advantage of relying on byte-copiable datatypes, while AD
+requires a complicated serialization and deserialization in order to function as byte-copiable. This
+is fine for copying in between MPI processes, but deserialization on an accelerator or GPU is very
+difficult, as this would require compiling a version of the AD library being used that were
+compatible with the accelerator or GPU being used. What this means is that if \emph{Peridigm} were
+to be reformulated to take advantage of accelerators or GPUs, attractive given their performance and
+power efficiency advantages over traditional CPUs, the flexibilty of complex-step owed to its simple
+implementation may make it more viable fo rthat application.
 

### 2013-09-09 jacobian_calc_CMAME.tex
Results and conslusion sections

@@
 \subsection{The Comparative Study} 
+\label{tcs}
 
@@
 guessing near the actual solution. The material properties of the model were set at $8.0E-9 metric
-tonne/ mm^3$ density, $1.515E4 MPa$ bulk modulus, and $7.813E4 Mpa$ shear modulus. The peridynamic
+tonne/ mm^3$ density, $1.515E4 MPa$ bulk modulus, and $7.813E4 MPa$ shear modulus. The peridynamic
 horizon used in a test problem was three times the node spacing for the corrisponding model mesh.
@@
 
-\begin{center}
+This section contains the averaged results from the study presented in tabular and graphic format.
+Table \ref{tab:results} shows the core count, number of nonzero Jacobian matrix elements, load step
+average calculation time and load step average accuracy. The units for accuracy are derived from the
+expression used to calculate the Jacobian matrix elements and the units used to define bulk material
+properties, mentioned in \ref{tcs}.
+
+\begin{table}
     \begin{tabular}{c c c c }
+        \caption{Averaged Results, Each Test} \label{tab:results}
     \hline
-    \multicolumn{4}{c}{Averaged Results, Entire Test Matrix} \\ \hline
     Cores & Jacobian Els. & Calc. Time (s) & Accuracy Diff. (MPa)\\ 
@@
     \end{tabular}
-\end{center}
+\end{table}
+
+The data presented in \ref{tab:results} will now be shown in graphic form alongside some discussion
+of results and trends.
 
 \subsection{Speed Data}
+Speed measurements were taken according to the methods mentioned in \ref{tcs}. The averaged results
+of these speed measurments for the single core tests can be seen in \ref{fig:serial_speed}.
+Average time in seconds to compute a Jacobian matrix element is plotted against number of nonzero
+Jacobian matrix ekements.
+Complex-step turned out to be the slowest of the four methods, taking the longest average time per
+iteration. The next slowest method was central-difference, followed by forward-difference and
+automatic-differentiation. Because the x axis of \ref{fig:serial_speed} is on a log scale, it
+bears mentioning that while visually it may appear that the trend indicated by the progression of
+datapoints on the plot is non-linear, the relationship would be closer to linear. This bodes well
+for the scalability of each of the methods by themselves in \emph{Peridigm}.
 
@@
 \end{figure}
+
+For the multicore tests, averaged speed results appear in \ref{fig:multi_speed}. For this series,
+rather than varying the number of nonzero Jacobian matrix elements the number of cores used to
+solve the problem is varied, while the number of nonzero Jacobian matrix elements remained constant.
+Complex-step again was the slowest method, followed by central-difference, automatic-differentiation
+and then forward-difference, taking the title of fastests method for this group of tests.
+\ref{fig:multi_speed} indicates that when looking at average in-iteration speed, each of the
+methods scale linearly with the number of cores used for the given conditions of this group of
+tests. Again, these results show that each of the methods as implemented in \emph{Peridigm} scale
+well when additional processors are used. An explanation as to why automatic-differentiation and
+forward-differences swapped places in speed ranking relative to the single core results was not
+determined.
 
@@
 \subsection{Accuracy Data}
+Accuracy measurements where taken according to the methods mentioned in \red{tcs}. The averaged
+results of these accuracy measurements for the single core tests can be seen in
+\ref{fig:serial_accuracy}. Averaged Frobenius norm of the element-wise difference between a given
+Jacobian and the Jacobian produced by the automatic-differentiation method is plotted against number
+of nonzero Jacobian matrix elements. Complex-step was revealed as the most accurate of the methods
+besides automatic-differentiation, complex-step resulting in several orders of magnitude less
+difference that resulted from using the central-difference method and the most inaccurate
+forward-difference method. This accuracy ranking order was mantained for each test in the series. It
+should be pointed out compared to the L2 norms of any of the Jacobian matrices by themselves, the
+magnitude of these differences for any of the methods is relatively small. These values appear in
+the experimental data which is available on the corresponding author's website. The effect of these
+accuracy differences on convergence rate directly was not determined.
 
@@
   \scalebox{.7}{\input{./figs/serial_accuracy.pgf}}
-  \caption{Serial test series speed measurements.}
+  \caption{Serial test series accuracy measurements.}
   \label{fig:serial_accuracy}
 \end{figure}
+
+The averaged resuts of the accuracy measurements taken for the multicore test series can be seen in
+\ref{fig:multi_accuracy}. Y axis units remain unchanged from \ref{fig:serial_accuracy}, but the x
+axis now indicated number of cores used to solve the test problem. Accuracy trends matched those for
+the single core tests.
 
@@
 \subsection{Efficiency Data}
+A derived measurement was also used to compare the methods. The single core results are shown in
+\ref{fig:serial_efficiency}. Efficiency, calculated as the average number of seconds taken per nonzero Jacobian matrix element 
+was ploted against the number nonzero Jacobian matrix elements. The ranking order of the methods
+matched their speed ranking in \ref{fig:serial_speed}. This plot indicates that there is a small,
+yet present decrease in efficiency w.r.t number of nonzero Jacobian elements. But judging by the
+speed results, this trend in efficiency doesn't appear to have great effect in human terms for the
+mesh densities used in this test series.
 
@@
 
+The multicore efficiency results are shown in \ref{fig:multi_efficiency}. Y axis is average number
+of seconds taken per nonzero Jacobian matrix element times number of cores used to solve the
+problem. The x axis in number of cores used to solve the problem. The multicore efficiency plot
+shows that automatic-differentiation was slower than forward-difference because for the given test
+problem it had a lower efficiency. Strangely automatic-differentiation appeared to gain in
+efficiency over the last few datapoints in the plot, looking at the raw data confirmed that this was
+what was measured rather than this being an error during data reduction. The reason behind this
+trend was not determined.
+
 \begin{figure}[h]
@@
 
-%%\includegraphics[width=\textwidth]{serial_speed_rel.png}
-%%\includegraphics[width=\textwidth]{multi_speed_rel.png}
-
 \section{Conclusions and Further Work}
 
-\subsection{Thoughts on Results}
+\subsection{Thoughts on Results} Based on the experimental results, it appears that for users of
+\emph{Peridigm} there is a case to be made for using automatic-differentiation for reasons of
+performance and accuracy. Although, the greater speed of finite-difference for certain multicore
+tests done in this study may give some users pause. While the results showed that complex-step was
+clearly more accurate than central-difference and forward-difference for the parameters of the
+tests, it was not determined whether or not this is a clear advantage of complex-step over
+finite-difference in terms of accuracy of final results and speed of convergence.   For other
+simulation software it is obviously necessary to collect the same types measurements as decribed in
+this paper in order to make a design decision. For new developments it is recommended that combined
+accuracy and speed of iteration mesurments be conduted along with entirely separate convergence
+studies.  For reasons explained in the next subsection, a convergence study in \emph{Peridigm} performed using
+the implementations of the methods used for this paper would be moot.
 
@@
 
-After the bulk of the study had been conducted, it was determined that the implementations of
-forward-difference, central-difference and by virtue of reusing code, complex-step, in \emph{Peridigm} were
-not optimal. Each of these implementations included extensive looping over noncontiguous addresses
-in memory and unnecessary re-calculation of constant values. The approach used does minimize
-temporary memory use, but is very wasteful of cpu and memory bandwidth. Unpublished preliminary results show
-that when calculations for the quantity 'zeta' are re-used and only perturbed and previously
-pertubed functions of independent variables are recomputed throughout the course of a Jacobian
-evaluation, complex-step performs at the speed of unmodified forward-difference. Because of this it is
-speculated that forward-difference and central-difference could be modified in nearly the same way to
-perform more quickly than AD. An enhanced performance implementation of forward-difference will
-eventually be made available on the corresponding author's website.
+After the bulk of the study had been conducted, gained familiarity with \emph{Peridigm} revealed
+that the implementations of forward-difference, central-difference and by virtue of reusing code,
+complex-step, in \emph{Peridigm} were not optimal due to inefficiencies in the implementations of
+the Linear Peridynamic Solid (LPS) model. Each of these implementations included extensive looping
+over noncontiguous addresses in memory and unnecessary re-calculation of constant values.  The
+approach used does minimize temporary memory use, and preserve modularity of the code for non-LPS
+models, but is very wasteful of cpu and memory bandwidth for use with LPS. Unpublished preliminary
+results show that when calculations for the quantity 'zeta' are re-used and only perturbed and
+previously pertubed functions of independent variables are recomputed throughout the course of a
+Jacobian evaluation, complex-step performs at the speed of unmodified forward-difference. Because of
+this it is speculated that forward-difference and central-difference could be modified in nearly the
+same way to perform more quickly than AD. In fact, because forward difference involves double the
+number of model evaluations than that of complex-step, forward-difference stands to gain
+proportionally more efficiency. Although no formal academic paper will be likely written based on
+this, a brief report and an enhanced performance implementation of forward-difference for peridynamic LPS
+in \emph{Peridigm} will eventually be made available on the corresponding author's website and is
+intended to be submitted for official inclusion in future releases of \emph{Peridigm}.
 
@@
 difficult, as this would require compiling a version of the AD library being used that were
-compatible with the accelerator or GPU being used. What this means is that if \emph{Peridigm} were
-to be reformulated to take advantage of accelerators or GPUs, attractive given their performance and
-power efficiency advantages over traditional CPUs, the flexibilty of complex-step owed to its simple
-implementation may make it more viable fo rthat application.
+compatible with the accelerator or GPU being used. What this means is that if \emph{Peridigm} or any
+simulation software were to be reformulated to take advantage of accelerators or GPUs, attractive
+given their performance and power efficiency advantages over traditional CPUs, the flexibilty of
+complex-step owed to its simple implementation may make it more viable fo rthat application.
 

### 2013-09-25 jacobian_calc_CMAME.tex
fixed typos and label syntax for table
editing and additing references

@@
 of the comparative study precede results for clarity. The intended audience of this paper
-includes researchers, professionals and students, therefor readers are directed to
+includes researchers, professionals and students, therefore readers are directed to
 selected background sources in the text for well-known concepts while less familiar concepts
@@
 Identifying a specific application provided a practical framework for implementing the new method and solving engineering problems to generate the data needed to compare the methods.
-In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of finite-difference, central-difference, and automatic-differentiation methods needed for comparison to the new complex-step methed was developed, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
+In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of finite-difference, central-difference, and automatic-differentiation methods needed for comparison to the new complex-step method was developed, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
 
@@
 \cite{ref-sacado-presentation}. For reference, the AD
-implementation used in the study is the "Sacado" package from the "Trilinos" library developed out of Sandia
-National Labs \cite{ref-Sacado}.
+implementation used in the study is part of the \emph{Sacado} package from the \emph{Trilinos} library developed out of Sandia National Labs \cite{ref-Sacado}.
 
@@
 central-difference and forward-difference will be shown and explained. The expressions for AD,
-central-difference and forward-difference are taken from {ADDAGIO MANUAL REF}, while the expression
+central-difference and forward-difference are taken from \cite{ref-Adaggio}, while the expression
 for CTSE is original.
@@
 positive $x_j$ axis. The second of the 'two terms' is $F_I$ evaluated at an unperturbed $x$. The
-denominator is the magnitude of perturbation called "probe distance" {ADDAGIO MANUAL REF}
+denominator is the magnitude of perturbation called "probe distance" \cite{ref-Adaggio}
 
@@
 Mesh generation for the test problems run in the study was done using the software package
-\emph{Cubit} developed at REFERENCE. For the test series in which peridynamic node density was
+\emph{Cubit} developed at Sandia National Labs \cite{ref-Cubit}. For the test series in which peridynamic node density was
 varied, journal scripts compatible with \emph{Cubit} were used. The rule used to select the
@@
 not calculated. However, if results of iterations within the same test problem were considered
-samples of the same thing, then a confidence interval could be calculated. This assumption was made
-for the purpose of plotting reporting average values, but a confidence interval was not calculated.
-But, a cursory review of data showed no range overlap among the methods for accuracy or speed data
-from any test run of either series.
+samples of the same thing, then a confidence interval could be calculated. This assumption was only made
+for the purpose of reporting average values on the plots. But, a cursory review of data showed no 
+range overlap among the methods for accuracy or speed data from any test run of either series.
 
@@
 \begin{table}    
+        \caption{Averaged Results, Each Test} \label{tab:results}   
     \begin{tabular}{c c c c }
-        \caption{Averaged Results, Each Test} \label{tab:results}
-    \hline
     Cores & Jacobian Els. & Calc. Time (s) & Accuracy Diff. (MPa)\\ 
@@
 Complex-step again was the slowest method, followed by central-difference, automatic-differentiation
-and then forward-difference, taking the title of fastests method for this group of tests.
+and then forward-difference, taking the title of fastest method for this group of tests.
 \ref{fig:multi_speed} indicates that when looking at average in-iteration speed, each of the
@@
 \subsection{Accuracy Data}
-Accuracy measurements where taken according to the methods mentioned in \red{tcs}. The averaged
+Accuracy measurements where taken according to the methods mentioned in \ref{tcs}. The averaged
 results of these accuracy measurements for the single core tests can be seen in
@@
 performance and accuracy. Although, the greater speed of finite-difference for certain multicore
-tests done in this study may give some users pause. While the results showed that complex-step was
-clearly more accurate than central-difference and forward-difference for the parameters of the
-tests, it was not determined whether or not this is a clear advantage of complex-step over
-finite-difference in terms of accuracy of final results and speed of convergence.   For other
+tests done in this study may give some users pause. While the results showed that complex-step
+produced more accurate tangent-stiffness matrices than central-difference and forward-difference 
+under the parameters of the tests, it was not determined whether or not this is a clear advantage of complex-step over
+finite-difference in terms of accuracy of final predicted node positions and speed of convergence. For other
 simulation software it is obviously necessary to collect the same types measurements as decribed in
@@
 complex-step, in \emph{Peridigm} were not optimal due to inefficiencies in the implementations of
-the Linear Peridynamic Solid (LPS) model. Each of these implementations included extensive looping
-over noncontiguous addresses in memory and unnecessary re-calculation of constant values.  The
-approach used does minimize temporary memory use, and preserve modularity of the code for non-LPS
-models, but is very wasteful of cpu and memory bandwidth for use with LPS. Unpublished preliminary
-results show that when calculations for the quantity 'zeta' are re-used and only perturbed and
-previously pertubed functions of independent variables are recomputed throughout the course of a
-Jacobian evaluation, complex-step performs at the speed of unmodified forward-difference. Because of
-this it is speculated that forward-difference and central-difference could be modified in nearly the
-same way to perform more quickly than AD. In fact, because forward difference involves double the
-number of model evaluations than that of complex-step, forward-difference stands to gain
-proportionally more efficiency. Although no formal academic paper will be likely written based on
-this, a brief report and an enhanced performance implementation of forward-difference for peridynamic LPS
-in \emph{Peridigm} will eventually be made available on the corresponding author's website and is
-intended to be submitted for official inclusion in future releases of \emph{Peridigm}.
+the Linear Peridynamic Solid (LPS) model. Each of these implementations included a large amount of
+unnecessary re-calculation of constant values.  The approach used does minimize temporary memory
+use, and allow other objects in \emph{Peridigm} to make calls directly to methods in the material
+objects, but is wastefull. Unpublished preliminary results show that when calculations for the
+constant quantity 'zeta' in the LPS implementation are re-used and only perturbed and previously
+pertubed functions of independent variables are recomputed throughout the course of a Jacobian
+evaluation, complex-step performs at the speed of unmodified forward-difference. Because of this it
+is speculated that forward-difference and central-difference could be modified in nearly the same
+way to perform more quickly than AD. In fact, because central-difference involves double the number
+of model evaluations than that of complex-step, central-difference stands to gain proportionally
+more efficiency. Although no formal academic paper will be likely written based on this, a brief
+report and an enhanced performance implementation of finite-difference for peridynamic LPS Jacbian
+calculation  in \emph{Peridigm} will eventually be made available on the corresponding author's
+website and is intended to be submitted for official inclusion in future releases of \emph{Peridigm}.
 
-\subsection{Final Thoughts on Complex-Step} Complex-step was shown to be highly accurate by the
-standards of this study, but as it is implemented in \emph{Peridigm} may not be any better than AD,
-and it is almost certainly slower for a naive implementation. When looking to the future of
-\emph{Peridigm}, Complex-step holds the advantage of relying on byte-copiable datatypes, while AD
-requires a complicated serialization and deserialization in order to function as byte-copiable. This
-is fine for copying in between MPI processes, but deserialization on an accelerator or GPU is very
-difficult, as this would require compiling a version of the AD library being used that were
-compatible with the accelerator or GPU being used. What this means is that if \emph{Peridigm} or any
-simulation software were to be reformulated to take advantage of accelerators or GPUs, attractive
-given their performance and power efficiency advantages over traditional CPUs, the flexibilty of
-complex-step owed to its simple implementation may make it more viable fo rthat application.
+\subsection{Final Thoughts on Complex-Step} Complex-step was shown to be highly accurate in
+calculating tangent-stiffness matrices by the standards of this study, but as it is implemented in
+\emph{Peridigm} may not be any more accurate than AD, and it is almost certainly slower for a naive
+implementation. When looking to the future of \emph{Peridigm}, Complex-step holds the advantage of
+relying on byte-copiable datatypes, while AD requires a complicated serialization and
+deserialization in order to function as byte-copiable. This is fine for copying in between MPI
+processes, but deserialization on an accelerator or GPU is very difficult, as this would require
+compiling a version of the AD library being used that were compatible with the accelerator or GPU
+being used. What this means is that if \emph{Peridigm} or any simulation software were to be
+reformulated to take advantage of accelerators or GPUs, attractive given their performance and power
+efficiency advantages over traditional CPUs, the flexibilty of complex-step owed to its simple
+implementation may make it more viable fo rthat application.
 

### 2013-09-25 jacobian_calc_CMAME.tex
added reference for LPS model

@@
 complex-step, in \emph{Peridigm} were not optimal due to inefficiencies in the implementations of
-the Linear Peridynamic Solid (LPS) model. Each of these implementations included a large amount of
+the Linear Peridynamic Solid (LPS) model \cite[A 1]{ref-LPS}. Each of these implementations included a large amount of
 unnecessary re-calculation of constant values.  The approach used does minimize temporary memory

### 2013-10-09 jacobian_calc_CMAME.tex
removed some speculation

--- 2013-09-25 jacobian_calc_CMAME.tex
@@
 accuracy and speed of iteration mesurments be conduted along with entirely separate convergence
-studies. For reasons explained in the next subsection, a convergence study in \emph{Peridigm} performed using
-the implementations of the methods used for this paper would be moot.
-
-\subsection{Thoughts on \emph{Perdigm}}
-
-After the bulk of the study had been conducted, gained familiarity with \emph{Peridigm} revealed
-that the implementations of forward-difference, central-difference and by virtue of reusing code,
-complex-step, in \emph{Peridigm} were not optimal due to inefficiencies in the implementations of
-the Linear Peridynamic Solid (LPS) model. Each of these implementations included a large amount of
-unnecessary re-calculation of constant values.  The approach used does minimize temporary memory
-use, and allow other objects in \emph{Peridigm} to make calls directly to methods in the material
-objects, but is wastefull. Unpublished preliminary results show that when calculations for the
-constant quantity 'zeta' in the LPS implementation are re-used and only perturbed and previously
-pertubed functions of independent variables are recomputed throughout the course of a Jacobian
-evaluation, complex-step performs at the speed of unmodified forward-difference. Because of this it
-is speculated that forward-difference and central-difference could be modified in nearly the same
-way to perform more quickly than AD. In fact, because central-difference involves double the number
-of model evaluations than that of complex-step, central-difference stands to gain proportionally
-more efficiency. Although no formal academic paper will be likely written based on this, a brief
-report and an enhanced performance implementation of finite-difference for peridynamic LPS Jacbian
-calculation  in \emph{Peridigm} will eventually be made available on the corresponding author's
-website and is intended to be submitted for official inclusion in future releases of \emph{Peridigm}.
+studies.
 

### 2013-10-25 jacobian_calc_CMAME.tex
Lots of changes, making progress

@@
 Identifying a specific application provided a practical framework for implementing the new method and solving engineering problems to generate the data needed to compare the methods.
-In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of finite-difference, central-difference, and automatic-differentiation methods needed for comparison to the new complex-step method was developed, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
+In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of finite-difference, central-difference, and automatic-differentiation methods needed for comparison to the new complex-step method implementation, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
 
 The aim of this paper is not only to introduce the new complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques and as a general reference for solving non-linear systems with Newton's method.
-After presenting background information on the underlying methods to be discussed, presented in this paper are: a description of tangent-stiffness matrices and how to use them to solve non-linear systems, detailed directions for producing tangent-stiffness matrices with each the methods identified above, a description and justification of a the new complex-step method for calculating tangent-stiffness matrices, a description of implementing complex-step in the \emph{Peridigm} software, a description of the quantities-of-interest used to rank the methods, a presentation and analysis of the results of the comparative study, and conclusions and thoughts on potential future work.
+After presenting background information on the underlying methods to be discussed, presented in this paper are: a description of tangent-stiffness matrices and how to use them to solve non-linear systems, detailed directions for producing tangent-stiffness matrices with each the methods identified above, a description and justification of the new complex-step method for calculating tangent-stiffness matrices, a description of implementing complex-step in the \emph{Peridigm} software, a description of the quantities-of-interest used to rank the methods, a presentation and analysis of the results of the comparative study, and conclusions and thoughts on potential future work.
 Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for C++ source-code and data necessary to reproduce the results presented here.
-Additionally included on the website is a C++ library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
+Additionally, included on the website is a C++ library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
 
@@
 
-This subsection contains background information on the differentiation techniques underlying the
-tangent-stiffness matrix calculation methods compared in the study. Since first-order finite
-difference techniques are considered well known, they are not described here. Instead the
-reader is referred to \cite[Chap. 4.1.3]{chapra2010}.
+This section contains background information on the differentiation techniques underlying the tangent-stiffness matrix calculation methods compared in the study.
+Since first-order finite difference techniques are considered well known, they are not described here.
+Instead the reader is referred to \cite[Chap.
+4.1.3]{chapra2010}.
 
-\subsubsection{Basis of New Method, 'Complex Step', in CTSE}  
-\note[Michael]{Will be replaced by Dr. Millwater's background section}
-
-During a literature review in preparation of material distinct from the instant paper, one of the
-instant authors identified a work presenting an application of the prior work of applied
-mathematicians Lyness and Moler to the numerical approximations of the first derivatives of real
-functions. First called Complex Taylor Series Expansion (CTSE) in (Lyness and Moler) and later by
-\cite{voorhees2011complex}, where Dr. Millwater is one of the  instant authors, the use of the term and
-its acronym are adopted out of deference to their  originators. Lyness and Moler described CTSE
-based on  approximating the Taylor Series Expansion of a function F, where $F:R^1 \rightarrow R^1$  about a
-complex point and solving for the first derivative. The application of Lyness and Moler's work
-demonstrated by Squire and Trap showed that there was empirical evidence that a CTSE based numerical
-derivative formula could resist the accuracy reducing phenomenon of subtractive cancellation in
-circumstances where central-difference and forward-difference were  demonstrated to succumb to
-subtractive-cancellation.
+\subsubsection{The ``complex-step'' method.}  
+It is possible to approximate derivatives quite accurately with a technique based on a complex variable Taylor series expansion of a function.
+This method was first described by Lyness and Moler \cite{lyness1967numerical, lyness1968differentiation} and has more recently been rediscovered \cite{squire1998using, voorhees2011complex, al2010complex} for use in engineering analysis.
+The basic idea is that a model parameter can be be made to be complex and expanded in a Taylor series about a small perturbation in the imaginary plane by some arbitrary value $h$ as follows
+%
+\begin{equation}
+f ( \beta_n + i h ) = f (\beta_n) + \frac{\partial f( \beta_n )}{\partial \beta_n} \frac{i h}{1!} + \frac{\partial f^2(\beta_n)}{\partial \beta_n^2} \frac{(i h)^2}{2!} + \frac{\partial f^3(\beta_n)}{\partial \beta_n^3} \frac{(i h)^3}{3!} + \cdots
+\label{eqn:complexTaylor}
+\end{equation}
+%
+Now we can take the imaginary part of both sides of equation (\ref{eqn:complexTaylor}) and solve for the first derivative ignoring terms of $\mathcal{O}\left ( h^2 \right)$ to yield an the estimate
+%
+\begin{equation}
+ \frac{\partial f( \beta_n )}{\partial \beta_n} \approx \frac{\mbox{Im} \left( f (\beta_n + i h) \right)}{h}.
+\label{eqn:complexFirstDeriv}
+\end{equation}
+%
+This shows that an estimate of the first derivative of a function can be made by only utilizing one functional evaluation of a perturbed model parameter up the imaginary axis.
+The step-size $h$ is arbitrary and can be made as small as practical (even to machine precision without the dangers of roundoff error as in finite-differences) to yield an accurate estimate of this derivative.
+The only disadvantage of this technique is the necessity of requiring the functions to accept complex numbers as arguments.
+Appendix~\ref{sec:appendixA} includes a simple example that illustraties the use of the complex-step method for derivative calculation of a function.
 
@@
 
-Automatic-differentiation or 'AD' is a computerized method for computing exact derivatives based the
-chain-rule from calculus. AD takes advantage of the fact that any
-mathematical function executed on a computer, no matter how complicated, is a "composition of simple
-operations" (add, multiply, power, transcendental and the like) each having known analytical derivatives
-\cite{ref-sacado-presentation}. For reference, the AD
-implementation used in the study is part of the \emph{Sacado} package from the \emph{Trilinos} library developed out of Sandia National Labs \cite{ref-Sacado}.
+Automatic-differentiation (AD) is a computerized method for computing exact derivatives based the chain-rule from calculus.
+AD takes advantage of the fact that any mathematical function executed on a computer, no matter how complicated, is a ``composition of simple operations'' (add, multiply, power, transcendental and the like) each having known analytical derivatives \cite{ref-sacado-presentation}.
+For reference, the AD implementation used in the study is part of the \emph{Sacado} package from the \emph{Trilinos} agile component libraries developed at Sandia National Laboratories \cite{ref-Sacado}.
 
-The way an AD system works is by first evaluating the innermost function of the composition, then presenting
-that function's output as input to the next level function until all levels are complete, in a way no
-different from how a normal computer program or human would evaluate composition functions.
-AD departs in that as each nested function is evaluated, the function's
-partial derivative with respect to the designated variables of the given input is also calculated.
-This is possible because the elementary math functions are hard-coded into the
-AD source-code along with their analytical derivatives, and linked by special instructions, so that when
-the elementary math functions are called upon for computation, their partial derivatives may be computed and
-stored in a sequence. The AD system then multiples the final sequence of partial derivatives together to
-produce the exact equivalent to taking a partial derivative of the corresponding composition
-function made up of the elementary functions with respect to a designated variable at a particular
-value. It is obviously, but bears mentioning, that the AD system could simply store one value for partial
-derivatives, modifying it as appropriate for every function evaluation rather than keeping a sequence.
-In the literature, the AD scheme described here is called 'forward automatic-differentiation'. For
-brevity, only forward AD will be covered since it is pertinent to the study, however the reader is referred
-to the introduction section of \cite{ref-AD-methods} and its references list for further information on AD,
-particularly \cite{ref-on-AD} which is foundational.
+The way an AD system works is by first evaluating the innermost function of the composition, then presenting that function's output as input to the next level function until all levels are complete, in a way no different from how a normal computer program or human would evaluate composition functions.
+AD departs in that as each nested function is evaluated, the function's partial derivative with respect to the designated variables of the given input is also calculated.
+This is possible because the elementary math functions are hard-coded into the AD source-code along with their analytical derivatives, and linked by special instructions, so that when the elementary math functions are called upon for computation, their partial derivatives may be computed and stored in a sequence.
+The AD system then multiples the final sequence of partial derivatives together to produce the exact equivalent to taking a partial derivative of the corresponding composition function with respect to a designated variable at a particular value.
+It is obvious, but bears mentioning, that the AD system could simply store one value for partial derivatives, modifying it as appropriate for every function evaluation rather than keeping a sequence.
+In the literature, the AD scheme described here is called \emph{forward automatic-differentiation}.
+Only forward AD will be covered here since it is the implementation used in \emph{Sacado} and therefore this study; however, the reader is referred to the introduction section of \cite{ref-AD-methods} and its citations for further information on AD, particularly \cite{ref-on-AD} which is foundational.
+Appendix~\ref{sec:appendixB} includes an example that walks the reader through the process a computer uses to compute derivates via AD.
+Some things to note about AD are that no approximation of derivatives is being made because the analytical forms of the partials of the elementary math functions are defined alongside them.
+The accuracy of AD is then limited by the precision of the AD system's definition of the elementary math functions and their partial derivatives .
 
-To make the above description complete, this subsection concludes with an example forward AD process, with 
-an emphasis on how general purpose information known to the computer is specialized with additional
-information from the problem and used to carry out AD:
+\subsection{Tangent-stiffness}
 
-\begin{enumerate}
-%Set up example problem. The example is meant to walk the reader through the forward AD algorithm
-%while encouraging them to rely on their understanding of the chain rule
-\item Take an example composition function $f(x) = (sin(cos(x)))^2$\\
+In solid and structural computational mechanics the tangent-stiffness matrix is a linearization operator that describes the stiffness of a system in response to small displacements imposed upon the current configuration of the system.
+Mathematically speaking, the tangent-stiffness represents the gradient of a high-dimensional energy surface the ``points'' in the minimum direction.
+While we've adopted the terminology of solid mechanics, it's important to note that these operators appear in other physical settings and are known by other names, e.g.
+a \emph{transmisability matrix} in the context of a Poisson problem or, more generally, a \emph{Jacobian matrix} in mathematical optimization.
 
-\item Suppose we want to know $\frac{df}{dx} \mid_{x_0}$ where $x_0$ is a particular value of $x$. \\
+While, for simplicity, the examples shown in the sequel refer only to linear problems, tangent-stiffness matrices generally arise in the context of non-linear analysis where Newton's method or quasi-Newton's method's are used in the minimization of some residual function.
+In this context, the tangent-stiffness matrix can be thought of as the linearization of the system about about a particular configuration.
+It has been shown \cite{hughes1978consistent, hughes1978unconditionally} that in order to preserve the quadratic convergence properties of the Newton's methods, it is necessary that this linearization is carried out in a manner that is \emph{consistent} with the algorthmic constraint equations used when computing the internal forces that arise due to deformations.
+An example of these constraint equations would be the Kuhn-Tucker conditions \cite{simo1998} that are used in integration of a flow rule for plasticity modeling.
+If the continuum tangent-moduli are used in place of the \emph{consistent} or \emph{algorithmic tangent moduli} in the solution of a non-linear solid mechanics problem, convergence may still be achieved, but not in the quadratic manner that makes Newton's method so attractive for this class of problems.
+Again, in the setting of non-linear solid mechanics, there is a very small class of material model algorithms in which consistent tangent-stiffness operators can be derived analytically.
+Perfect plasticity and plasticity with isotropic hardening used in combination with a general nearest-point projection or \emph{radial return algorithm} are a few examples.
+Therefore, when these models (and certainly more complex ones) are implemented into a general purpose computational mechanics code, the tangent-stiffness operators are typically defined via a numerical approximation.
 
-\begin{enumerate}
-%This first block of subitems is meant to establish that the values of x that the user cares about
-%about evaluating functions for are included within the set of numbers that functions g,h,k are allowed 
-%to deal with, and that nested functions can take each other as input, where s may not necessarily equal
-%x.
-\item Given that: \\
-\label{given}
-\begin{enumerate}
-\item $S \in R^1$ 
-\item $X \in S$ 
-\item $g, h, k \in H: S \rightarrow S$ 
-\item $\forall s \in S : g(s) = {s}^2$, $h(s) = sin(s)$, $k(s) = cos(s)$ 
-\item
-\label{composition}
-$\forall x \in X : f(x) = g(h(k(x)))$ \\
-\end{enumerate}
+In detail, a tangent-stiffness matrix can be described as a collection of the first-order partial derivatives of a vector valued function with respect to each independent vector component of the function for a given vector.
+A general mathematical formula for a tangent-stiffness matrix, $K_{ij}$,
+can be expressed in indicial notation as 
+%
+\begin{equation} 
+  K_{ij} = \left. \frac{\partial F_i}{\partial X_j}\right|_{\vec{X}_0}
+\end{equation}
+%
+Where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector is chosen to linearize about $\vec{X}_0$.
+One then evaluates the expression for each combination $i, j$ corresponding to an (row, column) location in the tangent-stiffness matrix.
+By inspection, the elements of a tangent-stiffness matrix can be estimated using the any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others.
+\\
 
-\item Given that the computer is programmed with some mathematical definitions:\\
-\begin{enumerate}
-\item
-\begin{enumerate}
-\item
-%The computer has at least definitions for these functions and variables
-$u, v, w \in H: R^3 \rightarrow R^1$ \\
-\item
-$s,a,b,x \in R^1$\\
-\end{enumerate}
-
-\item
-%AD doesn't return symbolic epressions for derivitives that can be re-evaluated.
-%Instead, generic versions of functions and evaluations of their partials are defined together.
-particular values can be identified: \\
-$s = s_0,..,s_n,...s_{\infty} \mid n=[0,\infty)$ \\ 
-and similarly for the other variables $a,b,x$ \\
-\item
-functions $u,v,w$ and their partial derivatives w.r.t $s$ are defined such that: \\
-\begin{tabular}{l | l}
-\multicolumn{2}{c}{for $a, b, s$ equal to $a_n, b_n, s_n$:} \\ \hline
-$u \mid _{a_n, b_n, s_n} = a_n \cdot s_n^{b_n}$ & $\frac{\partial{u}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot s_n^{b_n - 1}$ \\
-$v \mid _{a_n, b_n, s_n} = a_n \cdot sin(b_n \cdot s_n)$ & $\frac{\partial{v}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot cos(b_n \cdot s_n)$\\ 
-$w \mid _{a_n, b_n, s_n} = a_n \cdot cos(b_n \cdot s_n)$ & $\frac{\partial{w}}{\partial{s}} \mid _{a_n, b_n, s_n} = -a_n \cdot b_n \cdot sin(b_n \cdot s_n)$\\
-\end{tabular} 
-\end{enumerate}
-%a comment
-\item 
-\label{abcvalues}
-%General definitions of functions are specialized by arguments. 
-Given that it is possible to describe $f(x) = (sin(cos(x)))^2$ in terms the computer understands by inputting
-$f(x)$ such that the computer stores an equivalent statement $f(x) = u(a, b, s) \mid_{arguments}$,
-iff the arguments of $u,v,w$ are chosen such that $u,v,w$ approximate $g,h,k$ as follows:\\
-\begin{tabular}{l l l l | c}
-	Function & a & b & s & Approximates Function\\ \hline 
-	$u$ & $1$ & $2$ & $v$ & $g$\\ \hline 
-	$v$ & $1$ & $1$ & $w$ & $h$\\ \hline 
-	$w$ & $1$ & $1$ & $x$ & $k$\\ \hline 	
-\end{tabular}
-\end{enumerate}
-
-\item 
-\label{differentiation}
-%AD is eqivalent to the chain rule, when the computer has definitions for functions and partials
-%that can be adapted to match the compostition function.
-It follows from \ref{composition} that we can evaluate $\frac{d}{d x}f(x) \mid_{x_0}$ with the chain rule: \\ \\
-$\frac{d}{d x}f(x) = \frac{d}{d x} \cdot g(h(k(x))) \mid_{x_0}$ \\
-$\frac{d}{d x}f(x) =  \frac{d{g}}{d{h}} \cdot \frac{d{h}}{d{k}}
-\cdot \frac{d{k}}{d{x}}\mid_{x_0}$ \\ \\
-From the rest of \ref{given} it follows that we can approximate  $\frac{d}{d x}f(x) \mid_{x_0}$ by
-specializing the computer's general forms of $u, v, w$ 
-according to \ref{abcvalues}, with parameters $a, b$ chosen for each function and held as constant, and
-evaluating, such that the total derivative our original function w.r.t $x$ where $x = x_0$ is approximated
-by the partial derivative of our equivalent statement, $f(x) = u(a, b, s) \mid_{arguments}$,  w.r.t $s$ 
-where $s = x_0$. \\
-
-\end{enumerate}
-\label{doingthework}
-%This shows how information flows through the nested functions in one expression, like a table of contents
-For completeness we write out the computer's steps to evaluate \ref{differentiation} under the conditions of
-\ref{abcvalues} with a particular value of $s = x_0$, in equation format: \\ \\
-$\frac{\partial}{\partial x}f(x) \mid _{x_0} = \frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot \frac{ds}{dx}$\\ \\
-Repeating above in tabular format: \\ \\
-%This shows the same information but is ordered to show progression, like chapters
-\begin{tabular}{l l l}
-Current Evaluation & $s$ Value & Partial Derivative \\ \hline
-$w(a, b, s)\mid_{1, 1, s}$ & $x_0$ & $\frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
-$v(a, b, s)\mid_{1, 1, s}$ & $w(a, b, s)\mid_{1, 1, x_0}$ & $\frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
-$u(a, b, s)\mid_{1, 2, s}$ & $v(a, b, s)\mid_{1, 1, w(a, b, s)\mid_{1, 1, x_0}}$ & $\frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot 1$ \\
-\end{tabular} \\ \\
-%This explains what the equation and table above mean.
-Because the computer knew the analytical forms of the partial derivatives of each of $u,v,w$ beforehand,
-all it needed to do was:
-\begin{enumerate}
-\item to evaluate each of $u,v,w$ according to \ref{abcvalues}, in order from $w \rightarrow
-v \rightarrow u$, 
-\item to remember the values for $x_0$ and the output of each function evaluation besides  $u$, 
-\item then to use $x_0$ and the output of the function evaluations as input for the corresponding partial
-derivative function  evaluations, and
-\item store the individual partials.
-\end{enumerate}
-Lastly, to compute the partial derivative of the entire composition function, the computer multiplies
-the individual partials together in observance of the chain-rule.
-
-Some things to note about AD are that no approximation of derivatives is being made because the analytical 
-forms of the partials of the elementary math functions are defined alongside them. Accuracy of AD is then limited
-by the precision of the AD system's definition of the elementary math functions and their partials.
-
-\subsection{Tangent-stiffness Matrices}
-
-This subsection gives a working definition of a tangent-stiffness matrix and background on a
-Newton-Raphson method which uses a tangent-stiffness matrix to help solve nonlinear systems of
-equations, similarly to the the solution method used by the software package used in the study.
-
-\subsubsection{What is a Tangent-stiffness Matrix}
-\label{tsmat}
-
-It is important to identify what a tangent-stiffness matrix is and what it is used for to help show
-the motivation for the work discussed here. It is helpful to begin with the thought that a
-tangent-stiffness matrix is a type of slope for vector valued functions of vector variables at
-a single point in the space of the vector valued variable. A tangent-stiffness matrix is similar to the
-"$m$" in the function $y = mx + b$, except that $y$, $x$ and $b$ are vectors while $m$ is
-a matrix. \\
-
-In detail, a tangent-stiffness matrix can be described as a collection of the first
-order partial derivatives of a vector valued function w.r.t each degree of freedom of the vector
-valued function for a given value of the vector variable.  The tangent-stiffness matrix comprises a
-linear operator that can be used to transform a difference in the vector variable into a difference
-in the value of the vector valued function via an inner product between the tangent-stiffness matrix
-and the vector variable.  As an example in a solid mechanics system, suppose some dependent
-variables represented force components of force vectors at nodes of discretization, and independent
-variables represented displacement components for those same nodes. Suppose that as is possible in
-peridynamic nodes each dependent variable is a function of all of the independent variables. For
-this situation, the tangent-stiffness matrix would be the first derivatives of every force component
-variable with respect to every displacement component variable, with the rows corresponding to the force
-components and the columns corresponding to the displacement components.  \\
-
-How is a tangent-stiffness matrix computed? The mathematical formula for a tangent stiffness matrix
-can be expressed in indicial notation as:
-
-\begin{equation} \frac{\partial F_i}{\partial X_j}\mid_{X_0,...,X_n} \end{equation}
-
-Where $F_i$ is the i'th component of the vector valued function, $X_j$ is the j'th component of the
-vector variable, and a particular value of the vector variable is chosen $X_0,...,X_n$. One then
-evaluates the expression for each combination $i, j$ corresponding to an element at $row, column$ in
-the tangent-stiffness matrix. By inspection, the elements of a tangent-stiffness matrix can be
-estimated using the CTSE, AD and finite-difference techniques for functions $F:R^1 \rightarrow R^1$,
-since taking partial derivatives entails holding all but a single independent variable of the
-function constant, and each component of a vector valued function can be evaluated independently of
-the other components. \\
-
-\subsubsection{Calculating Tangent-stiffness Matrices with Each Method}
+\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation with each method.}
 For reference, expressions for calculating tangent stiffness matrices with CTSE, AD,
@@
 \label{sec:ack}
-
 %% References
@@
 %% not want to use elsarticle-num.bst.
+\appendix
+\renewcommand*{\thesection}{\Alph{section}}
+%
+\section{Complex-Step Example}
+\label{sec:appendixA}
+%
+Using $f(x) = \cos(x)$, we have
+%
+\[
+f (x + i h) = \cos(x + i h),
+\]
+%
+or through trigonometric identities equivalently
+%
+\[
+f(x + i h) = \cos(x) \cosh(h) - i \sin(x) \sinh(h),
+\]
+%
+Applying equation (\ref{eqn:complexFirstDeriv})
+\[
+\frac{\partial f(x)}{\partial x} = - \frac{\sin(x) \sinh(h)}{h},
+\]
+Since $h$ is arbitrary we desire it to be as small as possible or
+\[
+\frac{\partial f(x)}{\partial x} = \lim_{h \to 0} - \frac{\sin(x) \sinh(h)}{h},
+\]
+using L'H\^opital's rule
+\[
+\frac{\partial f(x)}{\partial x} = \lim_{h \to 0} - \frac{\sin(x) \cosh(h)}{1},
+\]
+of course, as $h \to 0$ then $\cosh(h) \to 1$ thereby recovering the exact derivative
+\[
+\frac{\partial f(x)}{\partial x} =-\sin(x).
+\]
+
+\section{AD Example}
+\label{sec:appendixB}
+
+\begin{enumerate}
+%Set up example problem. The example is meant to walk the reader through the forward AD algorithm
+%while encouraging them to rely on their understanding of the chain rule
+\item Take an example composition function $f(x) = (sin(cos(x)))^2$\\
+
+\item Suppose we want to know $\frac{df}{dx} \mid_{x_0}$ where $x_0$ is a particular value of $x$. \\
+
+\begin{enumerate}
+%This first block of subitems is meant to establish that the values of x that the user cares about
+%about evaluating functions for are included within the set of numbers that functions g,h,k are allowed 
+%to deal with, and that nested functions can take each other as input, where s may not necessarily equal
+%x.
+\item Given that: \\
+\label{given}
+\begin{enumerate}
+\item $S \in R^1$ 
+\item $X \in S$ 
+\item $g, h, k \in H: S \rightarrow S$ 
+\item $\forall s \in S : g(s) = {s}^2$, $h(s) = sin(s)$, $k(s) = cos(s)$ 
+\item
+\label{composition}
+$\forall x \in X : f(x) = g(h(k(x)))$ \\
+\end{enumerate}
+
+\item Given that the computer is programmed with some mathematical definitions:\\
+\begin{enumerate}
+\item
+\begin{enumerate}
+\item
+%The computer has at least definitions for these functions and variables
+$u, v, w \in H: R^3 \rightarrow R^1$ \\
+\item
+$s,a,b,x \in R^1$\\
+\end{enumerate}
+
+\item
+%AD doesn't return symbolic epressions for derivitives that can be re-evaluated.
+%Instead, generic versions of functions and evaluations of their partials are defined together.
+particular values can be identified: \\
+$s = s_0,..,s_n,...s_{\infty} \mid n=[0,\infty)$ \\ 
+and similarly for the other variables $a,b,x$ \\
+\item
+functions $u,v,w$ and their partial derivatives w.r.t $s$ are defined such that: \\
+\begin{tabular}{l | l}
+\multicolumn{2}{c}{for $a, b, s$ equal to $a_n, b_n, s_n$:} \\ \hline
+$u \mid _{a_n, b_n, s_n} = a_n \cdot s_n^{b_n}$ & $\frac{\partial{u}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot s_n^{b_n - 1}$ \\
+$v \mid _{a_n, b_n, s_n} = a_n \cdot sin(b_n \cdot s_n)$ & $\frac{\partial{v}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot cos(b_n \cdot s_n)$\\ 
+$w \mid _{a_n, b_n, s_n} = a_n \cdot cos(b_n \cdot s_n)$ & $\frac{\partial{w}}{\partial{s}} \mid _{a_n, b_n, s_n} = -a_n \cdot b_n \cdot sin(b_n \cdot s_n)$\\
+\end{tabular} 
+\end{enumerate}
+%a comment
+\item 
+\label{abcvalues}
+%General definitions of functions are specialized by arguments. 
+Given that it is possible to describe $f(x) = (sin(cos(x)))^2$ in terms the computer understands by inputting
+$f(x)$ such that the computer stores an equivalent statement $f(x) = u(a, b, s) \mid_{arguments}$,
+iff the arguments of $u,v,w$ are chosen such that $u,v,w$ approximate $g,h,k$ as follows:\\
+\begin{tabular}{l l l l | c}
+	Function & a & b & s & Approximates Function\\ \hline 
+	$u$ & $1$ & $2$ & $v$ & $g$\\ \hline 
+	$v$ & $1$ & $1$ & $w$ & $h$\\ \hline 
+	$w$ & $1$ & $1$ & $x$ & $k$\\ \hline 	
+\end{tabular}
+\end{enumerate}
+
+\item 
+\label{differentiation}
+%AD is eqivalent to the chain rule, when the computer has definitions for functions and partials
+%that can be adapted to match the compostition function.
+It follows from \ref{composition} that we can evaluate $\frac{d}{d x}f(x) \mid_{x_0}$ with the chain rule: \\ \\
+$\frac{d}{d x}f(x) = \frac{d}{d x} \cdot g(h(k(x))) \mid_{x_0}$ \\
+$\frac{d}{d x}f(x) =  \frac{d{g}}{d{h}} \cdot \frac{d{h}}{d{k}}
+\cdot \frac{d{k}}{d{x}}\mid_{x_0}$ \\ \\
+From the rest of \ref{given} it follows that we can approximate  $\frac{d}{d x}f(x) \mid_{x_0}$ by
+specializing the computer's general forms of $u, v, w$ 
+according to \ref{abcvalues}, with parameters $a, b$ chosen for each function and held as constant, and
+evaluating, such that the total derivative our original function w.r.t $x$ where $x = x_0$ is approximated
+by the partial derivative of our equivalent statement, $f(x) = u(a, b, s) \mid_{arguments}$,  w.r.t $s$ 
+where $s = x_0$. \\
+
+\end{enumerate}
+\label{doingthework}
+%This shows how information flows through the nested functions in one expression, like a table of contents
+For completeness we write out the computer's steps to evaluate \ref{differentiation} under the conditions of
+\ref{abcvalues} with a particular value of $s = x_0$, in equation format: \\ \\
+$\frac{\partial}{\partial x}f(x) \mid _{x_0} = \frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot \frac{ds}{dx}$\\ \\
+Repeating above in tabular format: \\ \\
+%This shows the same information but is ordered to show progression, like chapters
+\begin{tabular}{l l l}
+Current Evaluation & $s$ Value & Partial Derivative \\ \hline
+$w(a, b, s)\mid_{1, 1, s}$ & $x_0$ & $\frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
+$v(a, b, s)\mid_{1, 1, s}$ & $w(a, b, s)\mid_{1, 1, x_0}$ & $\frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
+$u(a, b, s)\mid_{1, 2, s}$ & $v(a, b, s)\mid_{1, 1, w(a, b, s)\mid_{1, 1, x_0}}$ & $\frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot 1$ \\
+\end{tabular} \\ \\
+%This explains what the equation and table above mean.
+Because the computer knew the analytical forms of the partial derivatives of each of $u,v,w$ beforehand,
+all it needed to do was:
+\begin{enumerate}
+\item to evaluate each of $u,v,w$ according to \ref{abcvalues}, in order from $w \rightarrow
+v \rightarrow u$, 
+\item to remember the values for $x_0$ and the output of each function evaluation besides  $u$, 
+\item then to use $x_0$ and the output of the function evaluations as input for the corresponding partial
+derivative function  evaluations, and
+\item store the individual partials.
+\end{enumerate}
+Lastly, to compute the partial derivative of the entire composition function, the computer multiplies
+the individual partials together in observance of the chain-rule.
 

### 2013-10-25 jacobian_calc_CMAME.tex
small fix
spelling and typo cleanup

@@
 This is possible because the elementary math functions are hard-coded into the AD source-code along with their analytical derivatives, and linked by special instructions, so that when the elementary math functions are called upon for computation, their partial derivatives may be computed and stored in a sequence.
-The AD system then multiples the final sequence of partial derivatives together to produce the exact equivalent to taking a partial derivative of the corresponding composition function with respect to a designated variable at a particular value.
+The AD system then multiplies the final sequence of partial derivatives together to produce the exact equivalent to taking a partial derivative of the corresponding composition function with respect to a designated variable at a particular value.
 It is obvious, but bears mentioning, that the AD system could simply store one value for partial derivatives, modifying it as appropriate for every function evaluation rather than keeping a sequence.
@@
 In solid and structural computational mechanics the tangent-stiffness matrix is a linearization operator that describes the stiffness of a system in response to small displacements imposed upon the current configuration of the system.
-Mathematically speaking, the tangent-stiffness represents the gradient of a high-dimensional energy surface the ``points'' in the minimum direction.
+Mathematically speaking, the tangent-stiffness represents the gradient of a high-dimensional energy surface that ``points'' in the minimum direction.
 While we've adopted the terminology of solid mechanics, it's important to note that these operators appear in other physical settings and are known by other names, e.g.
@@
 Where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector is chosen to linearize about $\vec{X}_0$.
-One then evaluates the expression for each combination $i, j$ corresponding to an (row, column) location in the tangent-stiffness matrix.
-By inspection, the elements of a tangent-stiffness matrix can be estimated using the any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others.
+One then evaluates the expression for each combination $i, j$ corresponding to a (row, column) location in the tangent-stiffness matrix.
+By inspection, the elements of a tangent-stiffness matrix can be estimated using any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others.
 \\
@@
 
-When the using the forward-difference technique the formula for calculating a tangent-stiffness
+When using the forward-difference technique the formula for calculating a tangent-stiffness
 matrix is:
@@
 positive $x_j$ axis. The second of the 'two terms' is $F_I$ evaluated at an unperturbed $x$. The
-denominator is the magnitude of perturbation called "probe distance" \cite{ref-Adaggio}
+denominator is the magnitude of perturbation called "probe distance" \cite{ref-Adaggio}.
 
@@
 Where our model $F$ is treated as function of complex variables and $x$ is treated as a vector of
-complex values. The expressions says that the real valued partial derivative of the $i$'th component of $F$
+complex values. The expression says that the real valued partial derivative of the $i$'th component of $F$
 w.r.t the $j$'th component of $x$ is equal to the imaginary component of $F_i$ evaluated at $x$
@@
 
-Complex step was implemented in \emph{Perdigm} as follows. First a way of persistently allocating
-memory needed to store imaginary values for nodal position and dilatation was determined. Persistent
-memory allocation is a very basic way to improve performance in general for a computer program. The
-way memory was allocated for similar vector and scalar variables used by the prior existing forward
-difference method was copied and new 'field spec' variables were added into the 'elastic' material
-model to store imaginary position coordinates and imaginary dilatation. Because of \emph{Peridigm}'s
-dependence on the \emph{Epetra} package of the \emph{Trilinos} library, 'field spec' variables could
-not be simply be declared as having a complex datatype. Program methods that were used in the
-course of applying the forward-difference method were copied, renamed and re-written to follow the
-CTSE formula, which involved changing the datatype of intermediate variables which were locally
-scoped and dynamically casting persistent variables, like the 'field spec' variables, to complex
-datatypes where necessary. Additional modifications were for the purpose of collecting experimental
-data and are not part of complex-step itself. The mathematical difference between the original
-forward-difference and complex-step code was covered in \ref{tsmat}. The source code for
-implementing complex-step in \emph{Perdigim} can be found on the corresponding author's website.
-Because \emph{Peridigm} was written by computer scientists rather than mechanical engineers
-primarily, many advanced techniques in C++ such as multiple virtual inheritance, pointer arithmetic
-and distributed data structures were used in the basic code and by necessity in the modified code.
-To see complex-step in a more narrow context, it is advised that readers also take a look at the
-examples of complex-step and the other methods provided on the corresponding author's website.
+Complex step was implemented in \emph{Perdigm} as follows. First a way of
+persistently allocating memory needed to store imaginary values for nodal
+position and dilatation was determined. Persistent memory allocation is a very
+basic way to improve performance in general for a computer program. The way
+memory was allocated for similar vector and scalar variables used by the prior
+existing forward difference method was copied and new 'field spec' variables
+were added into the 'elastic' material model to store imaginary position
+coordinates and imaginary dilatation. Because of \emph{Peridigm}'s dependence
+on the \emph{Epetra} package of the \emph{Trilinos} library, 'field spec'
+variables could not be simply be declared as having a complex datatype. Program
+methods that were used in the course of applying the forward-difference method
+were copied, renamed and re-written to follow the CTSE formula, which involved
+changing the datatype of intermediate variables which were locally scoped and
+dynamically casting persistent variables, like the 'field spec' variables, to
+complex datatypes where necessary. Additional modifications were for the
+purpose of collecting experimental data and are not part of complex-step
+itself. The source code for implementing complex-step in \emph{Perdigim} can be
+found on the corresponding author's website.  Because \emph{Peridigm} was
+written by computer scientists rather than mechanical engineers primarily, many
+advanced techniques in C++ such as multiple virtual inheritance, pointer
+arithmetic and distributed data structures were used in the basic code and by
+necessity in the modified code.  To see complex-step in a more narrow context,
+it is advised that readers also take a look at the examples of complex-step and
+the other methods provided on the corresponding author's website.
 
@@
 The methods were compared by running two sets of test problems where each method would solve a
-problem concurrently. Each test problem simulated a $4 by .25 by .25$ meter block of metal
+problem concurrently. Each test problem simulated a $4$ by $.25$ by $.25$ meter block of metal
 undergoing tension along the axis parallel to the long dimension of the block. The test problems
-were set up in \emph{Perdigm} as qausistatic equillibrium problems, where the displacement boundary
+were set up in \emph{Perdigm} as quasistatic equillibrium problems, where the displacement boundary
 conditions used to apply tension were applied gradually, in 'load steps', and an equilibrium
@@
 the load in gradual steps is a widely used technique to ensure the Newton-Raphson method is always
-guessing near the actual solution. The material properties of the model were set at $8.0E-9 metric
-tonne/ mm^3$ density, $1.515E4 MPa$ bulk modulus, and $7.813E4 MPa$ shear modulus. The peridynamic
+guessing near the actual solution. The material properties of the model were set at $8.0E-9$ $ metric
+tonne/ mm^3$ density, $1.515E4$ $MPa$ bulk modulus, and $7.813E4$ $MPa$ shear modulus. The peridynamic
 horizon used in a test problem was three times the node spacing for the corrisponding model mesh.
@@
 What was recorded during the test runs as experimental data is the subject the following
-subsubsection. It is organized to explain "what", "why" and "how" in that order.
+subsubsection.
 
@@
 compared to some other solution speculatively involving dynamic file management on files which for
-one of the components of the study would be on the order of $1XE10$ bytes. However, the price of
+one of the components of the study would be on the order of $1E10$ bytes. However, the price of
 running the methods together in the same process and avoiding the use of the hard-disk was that at
@@
 be found on the corresponding author's website. For future work, it may be beneficial to automate
-and randomize this process for statistic reasons and for time efficiency.
+and randomize this process for time efficiency and to assist statistical analysis.
 
@@
 Directories and filenames were chosen so that data would be easy to sort by individual test run.
-Extraneous information was manually deleted from copies versions of the output text files so that
+Extraneous information was manually deleted from copied versions of the output text files so that
 they would be regular enough to be easily read by scripts written in \emph{Python}. These scripts
@@
 besides automatic-differentiation, complex-step resulting in several orders of magnitude less
-difference that resulted from using the central-difference method and the most inaccurate
+difference than resulted from using the central-difference method and the most inaccurate
 forward-difference method. This accuracy ranking order was mantained for each test in the series. It
@@
 efficiency advantages over traditional CPUs, the flexibilty of complex-step owed to its simple
-implementation may make it more viable fo rthat application.
+implementation may make it more viable for that application.
 

### 2013-10-25 jacobian_calc_CMAME.tex
figure aspect and axes
More incremental fixes

--- 2013-10-25 jacobian_calc_CMAME.tex
@@
 Where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector is chosen to linearize about $\vec{X}_0$.
-One then evaluates the expression for each combination $i, j$ corresponding to an (row, column) location in the tangent-stiffness matrix.
+One then evaluates the expression for each combination $i, j$ corresponding to a (row, column) location in the tangent-stiffness matrix.
 By inspection, the elements of a tangent-stiffness matrix can be estimated using the any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others.
@@
 
-\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation with each method.}
-For reference, expressions for calculating tangent stiffness matrices with CTSE, AD,
-central-difference and forward-difference will be shown and explained. The expressions for AD,
-central-difference and forward-difference are taken from \cite{ref-Adaggio}, while the expression
-for CTSE is original.
+\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation.}
+
+For reference, expressions for calculating tangent stiffness matrices with complex-step, automatic differentiation, central-difference and forward-difference will be presented and explained.
 
 When the using the forward-difference technique the formula for calculating a tangent-stiffness
-matrix is:
-
-\begin{equation} K_{ij} = \frac{\partial F_i^{int}(x + \delta e_j) - F_i^{int}(x)}{\delta}
+matrix is
+%
+\begin{equation} 
+  K_{ij} = \frac{\partial F_i^{int}(x + h e_j) - F_i^{int}(x)}{h}
 \end{equation}
-
-Where $K_{ij}$ is the index notation representation of an element of the tangent-stiffness matrix at
-row $i$, column $j$. The expression says that the partial derivative of the $i$'th component of $F$
-w.r.t the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
+%
+Where $K_{ij}$ is the indicial notation representation of an element of the tangent-stiffness matrix at row $i$, column $j$.
+The expression says that the partial derivative of the $i$'th component of $F$ with respect the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
 and a denomenator. The first of the 'two terms' is $F_i$ evaluated at $x$ where

### 2013-10-25 jacobian_calc_CMAME.tex

--- 2013-10-25 jacobian_calc_CMAME.tex
@@
 
-\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation with each method.}
-For reference, expressions for calculating tangent stiffness matrices with CTSE, AD,
-central-difference and forward-difference will be shown and explained. The expressions for AD,
-central-difference and forward-difference are taken from \cite{ref-Adaggio}, while the expression
-for CTSE is original.
+\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation.}
 
@@
 matrix is:
-
-\begin{equation} K_{ij} = \frac{\partial F_i^{int}(x + \delta e_j) - F_i^{int}(x)}{\delta}
+%
+\begin{equation} 
+  K_{ij} = \frac{\partial F_i^{int}(x + h e_j) - F_i^{int}(x)}{h}
 \end{equation}
-
-Where $K_{ij}$ is the index notation representation of an element of the tangent-stiffness matrix at
-row $i$, column $j$. The expression says that the partial derivative of the $i$'th component of $F$
-w.r.t the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
+%
+Where $K_{ij}$ is the indicial notation representation of an element of the tangent-stiffness matrix at row $i$, column $j$.
+The expression says that the partial derivative of the $i$'th component of $F$ with respect the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
 and a denomenator. The first of the 'two terms' is $F_i$ evaluated at $x$ where
@@
   \centering
-  \scalebox{.7}{\input{./figs/serial_speed.pgf}}
+  \scalebox{1.0}{\input{./figs/serial_speed.pgf}}
   \caption{Serial test series speed measurements.}
@@
   \centering
-  \scalebox{.7}{\input{./figs/multi_speed.pgf}}
+  \scalebox{1.0}{\input{./figs/multi_speed.pgf}}
   \caption{Multicore test series speed measurements.}
@@
   \centering
-  \scalebox{.7}{\input{./figs/serial_accuracy.pgf}}
+  \scalebox{1.0}{\input{./figs/serial_accuracy.pgf}}
   \caption{Serial test series accuracy measurements.}
@@
   \centering
-  \scalebox{.7}{\input{./figs/multi_accuracy.pgf}}
+  \scalebox{1.0}{\input{./figs/multi_accuracy.pgf}}
   \caption{Multicore test series accuracy measurements.}
@@
   \centering
-  \scalebox{.7}{\input{./figs/serial_speed_rel.pgf}}
+  \scalebox{1.0}{\input{./figs/serial_speed_rel.pgf}}
   \caption{Serial test series efficiency measurements.}
@@
   \centering
-  \scalebox{.7}{\input{./figs/multi_speed_rel.pgf}}
+  \scalebox{1.0}{\input{./figs/multi_speed_rel.pgf}}
   \caption{Multicore test series efficiency measurements.}

### 2013-10-26 data/multicore/relative_ploter.py
added data to repository, fixed gridlines

@@ -1,0 +1,110 @@
+#!/usr/bin/env python
+import speed_multi	
+import accuracy_multi
+import numpy as np
+
+import matplotlib 
+matplotlib.use('pgf')
+import matplotlib.pyplot as plt
+
+font = {'family':'serif',
+	'serif':['Times'],
+	'size':12}
+
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+
+speed_data = speed_multi.Speed()
+accuracy_data = accuracy_multi.Accuracy()
+
+speed_averages = {}
+speed_per_nonzero_element = {}
+speed_per_mat_element = {}
+accuracy_averages = {}
+
+series_to_nonzero_elements = {}
+series_to_nonzero_elements[32] = 1666968710.0
+series_to_nonzero_elements[64] = 1666962752.0
+series_to_nonzero_elements[96] = 1666760396.0
+series_to_nonzero_elements[128] = 1666892660.0
+
+matrix_elements_for_million_element_problem = (3.0*1000000.0)**2.0
+
+marker = ""
+c =""
+"""
+fig1= plt.figure(1)
+ax1 = fig1.add_subplot(111)
+"""
+
+print speed_data.master_data_map
+for subkey in speed_data.master_data_map[speed_data.master_data_map.keys()[0]]:
+	speed_averages[subkey] = [[],[],[]]
+	speed_per_mat_element[subkey] = []
+	speed_per_nonzero_element[subkey] = []
+	accuracy_averages[subkey] = [[],[]]
+
+print "Master data map is: ", speed_data.master_data_map
+for index,key in enumerate(speed_data.master_data_map.keys()):
+	print "The key is: " + str(key)
+	for subindex, subkey in enumerate(speed_data.master_data_map[key]):
+		print "The subkey is: " + str(subkey)
+		subdata = [float(entry) for entry in speed_data.master_data_map[key][subkey]]
+		speed_averages[subkey][1].append(np.average(np.copy(subdata)))
+		speed_averages[subkey][0].append(series_to_nonzero_elements[float(key.split("_")[1])])
+		speed_averages[subkey][2].append(float(key.split("_")[1]))
+		speed_per_mat_element[subkey].append(float(key.split("_")[1])*speed_averages[subkey][1][-1]/matrix_elements_for_million_element_problem)
[…]

### 2013-10-26 data/multicore/speed_multi_right.py

@@ -1,0 +1,107 @@
+#!/usr/bin/env python
+#This file is called right because there was a particular correction made.
+#rather than taking the running average of the Jacobian calculation speed
+#and taking the average of the running average downstream the instant
+#calculation time is instead collected and then averaged downstream
+
+
+import numpy as np
+import glob
+
+class Speed:
+	def __init__(self):		
+		speedfilenames = glob.glob("./*_speed")
+
+		stripchars = '_.\/abcdefghijklmnopqrstuvwxyzABCDEFGHIJKLMNOPQRSTUVWXYZ()'
+
+		lines = []
+		data = []
+		keys = []
+
+		'''
+		load all records into data, divided by problem size -> iteration -> calculation type
+		'''
+		for filename in speedfilenames:
+			f = open(filename)
+			lines = [[entry.strip('()') for entry in line.split()[-6:]] for line in f.readlines()]
+			data.append([ [filename.strip(stripchars), line[0], line[1]] for line in lines])
+		'''
+		create a dictionary which holds the index of the problem
+		size as found in the data array
+		'''
+		for problemsize in data:
+			keys.append(problemsize[0][0])
+
+		problemsize_dictionary = {}
+		for index, key in zip(range(len(keys)),keys):
+			problemsize_dictionary[key] = index
+
+		#print"The dictionary is: ", problemsize_dictionary
+		'''
+		load the problem size numbers and measurements from the data array into 
+		speedmeasurements
+		'''
+		speedmeasurements = []
+		for problemsize in data:
+			speedmeasurements.append([])
+
+		for problemsize in data:
+			speedmeasurements[problemsize_dictionary[problemsize[0][0]]].append([])
+			speedmeasurements[problemsize_dictionary[problemsize[0][0]]].append([])
+			speedmeasurements[problemsize_dictionary[problemsize[0][0]]].append([])
+			speedmeasurements[problemsize_dictionary[problemsize[0][0]]].append([])
+			for row in range(len(problemsize)):
+				calc_id = (row +1)%4 
+
+				if(calc_id == 0):
+					if row == 3:
+						#print('Forward Difference')
+						speedmeasurements[problemsize_dictionary[problemsize[0][0]]][3].append(float(problemsize[row][1]))
+					else:
[…]

### 2013-10-26 data/singlecore/accuracy.py

@@ -1,0 +1,77 @@
+#!/usr/bin/env python
+
+import numpy as np
+import glob
+
+class Accuracy:
+	def __init__(self):
+		accuracyfilenames = glob.glob("./*_accuracy")
+
+		stripchars = '_.\/abcdefghijklmnopqrstuvwxyzABCDEFGHIJKLMNOPQRSTUVWXYZ()'
+
+		lines = []
+		data = []
+		keys = []
+
+		'''
+		load all records into data, divided by problem size -> iteration -> calculation type
+		'''
+		for filename in accuracyfilenames:
+			f = open(filename)
+			lines = [line.split()[-1] for line in f.readlines()]
+			data.append([ [filename.strip(stripchars), line] for line in lines])
+		'''
+		create a dictionary which holds the index of the problem
+		size as found in the data array
+		'''
+		for problemsize in data:
+			keys.append(problemsize[0][0])
+
+		problemsize_dictionary = {}
+		for index, key in zip(range(len(keys)),keys):
+			problemsize_dictionary[key] = index
+
+		#print("The dictionary is: ", problemsize_dictionary)
+		'''
+		load the problem size numbers and measurements from the data array into 
+		accuracymeasurements
+		'''
+		accuracymeasurements = []
+		for problemsize in data:
+			accuracymeasurements.append([])
+
+		for problemsize in data:
+			accuracymeasurements[problemsize_dictionary[problemsize[0][0]]].append([])
+			accuracymeasurements[problemsize_dictionary[problemsize[0][0]]].append([])
+			accuracymeasurements[problemsize_dictionary[problemsize[0][0]]].append([])
+			for row in range(len(problemsize)):
+				calc_id = (row +1)%3 
+
+				if(calc_id == 0):
+					#print('Central Differnce')
+					accuracymeasurements[problemsize_dictionary[problemsize[0][0]]][2].append(problemsize[row][1])
+
+				elif(calc_id == 1):
+					#print('Complex Step')
+					accuracymeasurements[problemsize_dictionary[problemsize[0][0]]][0].append(problemsize[row][1])
+
+				elif(calc_id == 2):
+					#print('Forward DIfference')
+					accuracymeasurements[problemsize_dictionary[problemsize[0][0]]][1].append(problemsize[row][1])
+
+				else:
+					print("calc_id reached an unexpected value, calc_id was: ", calc_id)
+
+		self.master_data_map = {}
+
+		for key in problemsize_dictionary.keys():
[…]

### 2013-10-26 data/singlecore/accuracy_validity_study.py

@@ -1,0 +1,36 @@
+#!/usr/bin/env python
+
+import numpy as np
+
+level_1 = np.ones((10,10), dtype=np.longdouble)*2.0 
+level_1_norm = np.linalg.norm(level_1)
+print level_1_norm
+
+level_1 - np.ones((1*np.sqrt(10.0)
+
+level_2 = np.ones((100,100), dtype=np.longdouble)*(np.divide(level_1_norm, level_1.size))
+level_2_norm = np.linalg.norm(level_2)
+print level_2_norm
+
+level_3 = np.ones((1000,1000), dtype=np.longdouble)*(np.divide(level_1_norm, np.power(level_1.size,1.5)))
+level_3_norm = np.linalg.norm(level_3)
+print level_3_norm
+
+level_4 = np.ones((10000,10000), dtype=np.longdouble)*(np.divide(level_1_norm, np.power(level_1.size,2.0)))
+level_4_norm = np.linalg.norm(level_4)
+print level_4_norm
+
+level_1 += np.random.normal(scale = .001, size=(level_1.shape[0], level_1.shape[1]))
+level_2 += np.random.normal(scale = .001, size=(level_2.shape[0], level_2.shape[1]))
+level_3 += np.random.normal(scale = .001, size=(level_3.shape[0], level_3.shape[1]))
+level_4 += np.random.normal(scale = .001, size=(level_4.shape[0], level_4.shape[1]))
+
+level_1_norm = np.linalg.norm(level_1)
+level_2_norm = np.linalg.norm(level_2)
+level_3_norm = np.linalg.norm(level_3)
+level_4_norm = np.linalg.norm(level_4)
+
+print level_1_norm
+print level_2_norm
+print level_3_norm
+print level_4_norm

### 2013-10-26 data/singlecore/example.py

@@ -1,0 +1,46 @@
+#!/usr/bin/python
+
+from matplotlib import rc
+## for Palatino and other serif fonts use:
+rc('font',**{'family':'serif','serif':['Times']})
+rc('text', usetex=True)
+
+#Here is the standard example, tex_demo.py:
+
+"""
+Demo of TeX rendering.
+
+You can use TeX to render all of your matplotlib text if the rc
+parameter text.usetex is set.  This works currently on the agg and ps
+backends, and requires that you have tex and the other dependencies
+described at http://matplotlib.sf.net/matplotlib.texmanager.html
+properly installed on your system.  The first time you run a script
+you will see a lot of output from tex and associated tools.  The next
+time, the run may be silent, as a lot of the information is cached in
+~/.tex.cache
+
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+# Example data
+t = np.arange(0.0, 1.0 + 0.01, 0.01)
+s = np.cos(4 * np.pi * t) + 2
+
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+plt.plot(t, s)
+
+plt.xlabel(r'\textbf{time} (s)')
+plt.ylabel(r'\textit{voltage} (mV)',fontsize=12)
+plt.title(r"\TeX\ is Number "
+          r"$\displaystyle\sum_{n=1}^\infty\frac{-e^{i\pi}}{2^n}$!",
+          fontsize=12, color='gray')
+# Make room for the ridiculously large title.
+plt.subplots_adjust(top=0.8)
+
+plt.savefig('tex_demo')
+plt.show()
+
+

### 2013-10-26 data/singlecore/plotter.py

@@ -1,0 +1,145 @@
+#!/usr/bin/env python
+import speed
+import accuracy
+import numpy as np
+
+import matplotlib 
+matplotlib.use('pgf')
+import matplotlib.pyplot as plt
+#import mpl_toolkits.axisartist as axisartist
+
+font = {'family':'serif',
+	'serif':['Times'],
+	'size':12}
+
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+
+#matplotlib.rc('font', **font)
+
+speed_data = speed.Speed()
+accuracy_data = accuracy.Accuracy()
+
+speed_averages = {}
+speed_per_mat_element = {}
+accuracy_averages = {}
+accuracy_per_mat_element = {}
+
+series_to_rowcount = {}
+series_to_rowcount[1000.0] = 3000.0
+series_to_rowcount[2000.0] = 5400.0
+series_to_rowcount[3000.0] = 8526.0
+series_to_rowcount[4000.0] = 12288.0
+series_to_rowcount[6000.0] = 18468.0
+series_to_rowcount[8000.0] = 24000.0
+series_to_rowcount[12000.0] = 33396.0
+series_to_rowcount[16000.0] = 50700.0
+series_to_rowcount[32000.0] = 96768.0
+
+series_to_nonzero_entries = {}
+series_to_nonzero_entries[1000.0] = 1994454.0 
+series_to_nonzero_entries[2000.0] = 4252230.0
+series_to_nonzero_entries[3000.0] = 7792128.0
+series_to_nonzero_entries[4000.0] = 15389352.0
+series_to_nonzero_entries[6000.0] = 20027466.0
+series_to_nonzero_entries[8000.0] = 31388724.0
+series_to_nonzero_entries[12000.0] = 40060026.0
+series_to_nonzero_entries[16000.0] = 83778480.0
+series_to_nonzero_entries[32000.0] = 165140640.0
+
+
+xlimit = 100000.0
+xmin = 2000.0
+
+
+marker = ""
+c =""
+fig1= plt.figure(1, figsize=(4,3.7))
+ax1 = fig1.add_subplot(111)
+
+
+print speed_data.master_data_map
+for subkey in speed_data.master_data_map[speed_data.master_data_map.keys()[0]]:
+	speed_averages[subkey] = [[],[]]
+	speed_per_mat_element[subkey] = []
+	accuracy_averages[subkey] = [[],[]]
+	accuracy_per_mat_element[subkey] = []
+
[…]

### 2013-10-26 data/singlecore/relative_ploter.py

@@ -1,0 +1,144 @@
+#!/usr/bin/env python
+import speed
+import accuracy
+import numpy as np
+import numpy as np
+
+import matplotlib 
+matplotlib.use('pgf')
+import matplotlib.pyplot as plt
+#import mpl_toolkits.axisartist as axisartist
+
+font = {'family':'serif',
+	'serif':['Times'],
+	'size':12}
+
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+
+speed_data = speed.Speed()
+accuracy_data = accuracy.Accuracy()
+
+speed_averages = {}
+speed_per_mat_element = {}
+speed_per_nonzero_element = {}
+accuracy_averages = {}
+accuracy_per_mat_element = {}
+
+series_to_rowcount = {}
+series_to_rowcount[1000.0] = 3000.0
+series_to_rowcount[2000.0] = 5400.0
+series_to_rowcount[3000.0] = 8526.0
+series_to_rowcount[4000.0] = 12288.0
+series_to_rowcount[6000.0] = 18468.0
+series_to_rowcount[8000.0] = 24000.0
+series_to_rowcount[12000.0] = 33396.0
+series_to_rowcount[16000.0] = 50700.0
+series_to_rowcount[32000.0] = 96768.0
+
+series_to_nonzero_entries = {}
+series_to_nonzero_entries[1000.0] = 1994454.0 
+series_to_nonzero_entries[2000.0] = 4252230.0
+series_to_nonzero_entries[3000.0] = 7792128.0
+series_to_nonzero_entries[4000.0] = 15389352.0
+series_to_nonzero_entries[6000.0] = 20027466.0
+series_to_nonzero_entries[8000.0] = 31388724.0
+series_to_nonzero_entries[12000.0] = 40060026.0
+series_to_nonzero_entries[16000.0] = 83778480.0
+series_to_nonzero_entries[32000.0] = 165140640.0
+
+xlimit = 100000.0
+xmin = 2000.0
+
+
+marker = ""
+c =""
+"""
+fig1= plt.figure(1)
+ax1 = fig1.add_subplot(axisartist.Subplot(fig1, "111"))
+"""
+print speed_data.master_data_map
+for subkey in speed_data.master_data_map[speed_data.master_data_map.keys()[0]]:
+	speed_averages[subkey] = [[],[],[]]
+	speed_per_mat_element[subkey] = []
+	speed_per_nonzero_element[subkey] = []
[…]

### 2013-10-27 jacobian_calc_CMAME.tex
More changes thru Sec. 1.2.1

@@
 The only disadvantage of this technique is the necessity of requiring the functions to accept complex numbers as arguments.
-Appendix~\ref{sec:appendixA} includes a simple example that illustraties the use of the complex-step method for derivative calculation of a function.
+Appendix~\ref{sec:appendixA} includes a simple example that illustrates the use of the complex-step method for derivative calculation of a function.
 
@@
 In this context, the tangent-stiffness matrix can be thought of as the linearization of the system about about a particular configuration.
-It has been shown \cite{hughes1978consistent, hughes1978unconditionally} that in order to preserve the quadratic convergence properties of the Newton's methods, it is necessary that this linearization is carried out in a manner that is \emph{consistent} with the algorthmic constraint equations used when computing the internal forces that arise due to deformations.
+It has been shown \cite{hughes1978consistent, hughes1978unconditionally} that in order to preserve the quadratic convergence properties of the Newton's methods, it is necessary that this linearization is carried out in a manner that is \emph{consistent} with the algorithmic constraint equations used when computing the internal forces that arise due to deformations.
 An example of these constraint equations would be the Kuhn-Tucker conditions \cite{simo1998} that are used in integration of a flow rule for plasticity modeling.
@@
 \begin{equation} 
-  K_{ij} = \left. \frac{\partial F_i}{\partial X_j}\right|_{\vec{X}_0}
+  K_{ij} = \left. \frac{\partial F_i}{\partial X_j}\right|_{\vec{X}_0},
 \end{equation}
 %
-Where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector is chosen to linearize about $\vec{X}_0$.
+where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector is chosen to linearize about $\vec{X}_0$.
 One then evaluates the expression for each combination $i, j$ corresponding to a (row, column) location in the tangent-stiffness matrix.
@@
 
-When using the forward-difference technique the formula for calculating a tangent-stiffness
-matrix is:
+The \emph{forward-difference} (FD) formula for calculating a tangent-stiffness matrix in the setting of a solid structural mechanics problem is
 %
 \begin{equation} 
-  K_{ij} = \frac{\partial F_i^{int}(x + h e_j) - F_i^{int}(x)}{h}
+  K_{ij} = \frac{\partial F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u})}{h},
 \end{equation}
@@
 Where $K_{ij}$ is the indicial notation representation of an element of the tangent-stiffness matrix at row $i$, column $j$.
-The expression says that the partial derivative of the $i$'th component of $F$ with respect the $j$'th component of $x$ is equal to the quotient of the partial difference of two terms
-and a denomenator. The first of the 'two terms' is $F_i$ evaluated at $x$ where
-the $j$'th component of $x$ has been perturbed by an very small amount in the direction of the
-positive $x_j$ axis. The second of the 'two terms' is $F_I$ evaluated at an unperturbed $x$. The
-denominator is the magnitude of perturbation called "probe distance" \cite{ref-Adaggio}.
+The first of the two terms in the numerator is the internal force, $F_i^{int}$ evaluated in the current deformed configuration, $\vec{u}$,  plus a small perturbation $h$ in the direction $\hat{e}_j$ where $\hat{e}_j$ represents a unit-vector corresponding to the $j^{\mbox{th}}$ degree-of-freedom.
+The second of the two terms is $F_i^{int}$ evaluated in the current configuration.
+The denominator is the magnitude of perturbation called \emph{probe-distance}.
 
-When using the central-difference technique the formula for calculating a tangent-stiffness matrix
-is:
+The \emph{central-difference} (CD) formula for calculating a tangent-stiffness matrix is
+%
+\begin{equation} 
+  K_{ij} = \frac{\partial F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u} - h \hat{e}_j)}{2 h}.
+\end{equation}
+%
+Together, the FD and CD methods, are sometimes termed \emph{finite-difference probing} techniques for tangent-stiffness calculation in literature \cite{ref-Adaggio}.
 
-\begin{equation} K_{ij} = \frac{\partial F_i^{int}(x + \delta e_j) - F_i^{int}(x - \delta
-e_j)}{2\delta} \end{equation}
-
-The distiction between the central difference-method and the forward-difference method
-tangent-stiffness calculation formula is that the partial difference is now between two perturbed
-function evaluations, rather than between an unperturbed and perturbed function evaluation as in
-forward-difference. Additionally, the minuend involves a perturbation of $x$ in the direction opposite to
-the direction of the $j$'th component of $x$.
-
-When using the comple-step technique, which is based on CTSE, the formula for calculating a
-tangent-stiffness matrix is:
-
-\begin{equation} K_{ij} = \frac{imag(F_i^{int}(x + \delta e_j^{imag}))}{\delta} \end{equation}
+The complex-step (CS) formula for calculating a tangent-stiffness matrix is
+%
+\begin{equation} 
+  K_{ij} = \frac{\mbox{Im}(F_i^{int}(\vec{u} + i h \hat{e}_j))}{h}
+\end{equation}
 

### 2013-10-27 jacobian_calc_CMAME.tex
Finished Sec. 1.2.1

@@
 \subsubsection{The ``complex-step'' method.}  
+\label{sec:CSmethod}
 It is possible to approximate derivatives quite accurately with a technique based on a complex variable Taylor series expansion of a function.
@@
 where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector is chosen to linearize about $\vec{X}_0$.
-One then evaluates the expression for each combination $i, j$ corresponding to a (row, column) location in the tangent-stiffness matrix.
+One then evaluates the expression for each combination $(i, j)$ corresponding to a (row, column) location in the tangent-stiffness matrix.
 By inspection, the elements of a tangent-stiffness matrix can be estimated using any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others.
@@
 \begin{equation} 
-  K_{ij} = \frac{\partial F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u})}{h},
+  K_{ij} = \frac{F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u})}{h},
 \end{equation}
@@
 \begin{equation} 
-  K_{ij} = \frac{\partial F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u} - h \hat{e}_j)}{2 h}.
+  K_{ij} = \frac{F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u} - h \hat{e}_j)}{2 h}.
 \end{equation}
@@
 \begin{equation} 
-  K_{ij} = \frac{\mbox{Im}(F_i^{int}(\vec{u} + i h \hat{e}_j))}{h}
+  K_{ij} = \frac{\mbox{Im}(F_i^{int}(\vec{u} + i h \hat{e}_j))}{h},
 \end{equation}
+%
+where the internal force, $F^{int}$ is now treated as function of complex vectors,  $\vec{u}$.
+The small perturbation, $h$, is now carried out on the imaginary axis and only the real coefficient to the imaginary part of the vector returned by $F^{int}$ is kept, hence the $\mbox{Im}(\cdot)$ operation.
+As highlighted in Section~\ref{sec:CSmethod}, notice there is no subtraction operation taking place; therefore $h$ can be made very small to yield highly accurate derivatives.
 
-Where our model $F$ is treated as function of complex variables and $x$ is treated as a vector of
-complex values. The expression says that the real valued partial derivative of the $i$'th component of $F$
-w.r.t the $j$'th component of $x$ is equal to the imaginary component of $F_i$ evaluated at $x$
-where $x$ is perturbed by a very small amount in the direciton of the imaginary axis corresponding
-to the $j$'th component of $x$ all divided by the magnitude of the perturbation. There is no partial
-notation because a partial difference is not taken.
+When using automatic-differentiation to calculate a tangent-stiffness matrix, one only needs to follow the definition of the tangent-stiffness matrix and issue the correct commands to the AD system.
+The formula is just as expected:
 
-When using automatic-differentiation to calculate a tangent-stiffness matrix, one only needs to
-follow the definition of the tangent-stiffness matrix and issue the correct commands to the AD
-system. The formula is just as expected:
-
-\begin{equation} K_{ij} = \frac{\partial F_i^{int}}{\partial x_j} \end{equation}
+\begin{equation} K_{ij} = \frac{\partial F_i^{int}(\vec{u})}{\partial u_j} \end{equation}
 

### 2013-10-27 jacobian_calc_CMAME.tex
Through with Peridigm section

@@
 
-In order to maintain the quadratic convergence properties of the first-order Newton's method in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
+In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
 For an extremely small class of nonlinear material models, these consistent tangent-stiffness operators can be derived analytically; however, most often in practice, they are found through numerical approximation of derivatives.
@@
 This section contains background information on the differentiation techniques underlying the tangent-stiffness matrix calculation methods compared in the study.
-Since first-order finite difference techniques are considered well known, they are not described here.
+Since first-order finite-difference techniques are considered well known, they are not described here.
 Instead the reader is referred to \cite[Chap.
@@
 
-\subsubsection{Differentiation Technique: Automatic-differentiation} 
+\subsubsection{Automatic-differentiation} 
 \label{ADsubsection}
@@
 When using automatic-differentiation to calculate a tangent-stiffness matrix, one only needs to follow the definition of the tangent-stiffness matrix and issue the correct commands to the AD system.
-The formula is just as expected:
+The formula is the same as the continuum formula
+%
+\begin{equation} 
+  K_{ij} = \frac{\partial F_i^{int}(\vec{u})}{\partial u_j},
+\end{equation}
+%
+but, with the caveat that $F^{int}$ is evaluated in a way that is algorithmically consistent, and therefore should yield quadratic convergence with Newton's method.
 
-\begin{equation} K_{ij} = \frac{\partial F_i^{int}(\vec{u})}{\partial u_j} \end{equation}
+\section{Description of analysis tools and approach.} 
+%
+\subsection{Peridigm and  peridynamics.}
+%
+Comparisons of the different tangent-stiffness calculation techniques were carried out through code-development and \emph{in situ} instrumentation of the computational peridynamics code \emph{Peridigm} \cite{peridigm}.
+\emph{Peridigm} is distributed by Sandia National Laboratories as its primary open-source computational peridynamics code.
+It is a massively-parallel simulation code for implicit and explicit multi-physics simulations primarily used for solid mechanics and material failure.
+\emph{Peridigm} is a C++ code utilizing agile software components from Sandia's  \emph{Trilinos} project \cite{trilinos}.
+It's important to note that there is no specialization of the tangent-stiffness calculation methods described previously to this particular code or more generally to a computational peridynamics approach and the analysis carried out in this study should provide insight into the speed and accruacy of these methods when implemented into other computational mechanics analysis tools as well.
+\emph{Peridigm} was chosen because of the author's relationship with the code development project and primarily because the FD, CD, and AD methods for tangent-stiffness calculation where areleady implemented in the code, leaving only the CS method to for development.
+However, because peridynamics is a nonlocal theory, the tangent-stiffness matrices have a much higher bandwidth (i.e.
+less sparsity) than traditional finite-element tangent-stiffness computations and the time required to construct the tangent-stiffness is a majority of the total computation time in any quasi-static or implicit dynamics analysis; therefore, the \emph{Peridigm} development team has an interest in profiling the performance of the tangent-stiffness computation methods.
 
-\subsubsection{Solve a Nonlinear System with a Tangent-stiffness Matrix} \label{solve_a_system}
-The reader is referred to a succinct explanation of the Newton Raphson method, which shows the
-role of a tangent-stiffness matrix or Jacobian in that solution method \cite[chap.
-13]{young2009}.
+Briefly, peridynamics \cite{silling2000ret, silling:psa, silling2010peridynamic} is a nonlocal reformulation of the partial-differential equations that provide the statement of momentum balance in classical continuum mechanics.
+Its primary goal is the avoid the use of spatial derivatives in the balance equations or constitutive laws such that discontinuous displacements, i.e.
+cracks, are mathematically consistent with the governing equations.
+It has demonstrated great promise in modeling problems with massive pervasive failure \cite{littlewood2010}, interesting phenomenon such as dynamic crack branching \cite{ha2010sod}, and is being extended to multiphyics phenomenon \cite{bobaru2011peridynamic,katiyar2013} and coupling to molecular simulations \cite{seleson2009peridynamics}.
 
-\section{Methods and Materials}
+\subsubsection{Implementing the complex-step method in Peridigm} 
+%
+The CS method was implemented in \emph{Perdigm} in a manner that follows closely the implementation of the FD technique.
+However, because of \emph{Peridigm}'s reliance on \emph{Trilinos} and specifically the \emph{Epetra} package, special steps had to be followed to allow the use of complex number data types.
+This is because \emph{Epetra} vectors, which can be thought of as distributed memory parallel cousins to standard C++ STL vectors, are hard coded to be of type {\tt double} and can not be simply be declared as having a complex datatype.
+Therefore, program methods that were used in the course of applying the FD method were copied, renamed and re-written to follow the CS formula, which involved changing the datatype of intermediate variables which were locally scoped and dynamically casting persistent memory variables to complex datatypes where necessary.
+We made no atttempt to ``tune'' the CS code in any way for performance, we attempted to implement the algorithm is such as way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurt its overall performance and this issue could be alleviated by using the next generation templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any datatype including complex.
+There where additional modifications for the purpose instrumenting the code to collect comparitive data and are not specifically part of CS or other methods.
+The source code patch to \emph{Peridigm} for implementing CS can be found on the corresponding author's website.
 
-\subsection{Material: Perdynamics Code}
-
-The following subsubsections identify the program modified to run the simulations for the
-comparative study and explain what modifications were necessary.
-
-\subsubsection{\emph{Perdigm} and Peridynamics} 
-\subsubsection{Implementing Complex-Step in \emph{Peridigm}}
-
-Complex step was implemented in \emph{Perdigm} as follows. First a way of
-persistently allocating memory needed to store imaginary values for nodal
-position and dilatation was determined. Persistent memory allocation is a very
-basic way to improve performance in general for a computer program. The way
-memory was allocated for similar vector and scalar variables used by the prior
-existing forward difference method was copied and new 'field spec' variables
-were added into the 'elastic' material model to store imaginary position
-coordinates and imaginary dilatation. Because of \emph{Peridigm}'s dependence
-on the \emph{Epetra} package of the \emph{Trilinos} library, 'field spec'
-variables could not be simply be declared as having a complex datatype. Program
-methods that were used in the course of applying the forward-difference method
-were copied, renamed and re-written to follow the CTSE formula, which involved
-changing the datatype of intermediate variables which were locally scoped and
-dynamically casting persistent variables, like the 'field spec' variables, to
-complex datatypes where necessary. Additional modifications were for the
-purpose of collecting experimental data and are not part of complex-step
-itself. The source code for implementing complex-step in \emph{Perdigim} can be
-found on the corresponding author's website.  Because \emph{Peridigm} was
-written by computer scientists rather than mechanical engineers primarily, many
-advanced techniques in C++ such as multiple virtual inheritance, pointer
-arithmetic and distributed data structures were used in the basic code and by
-necessity in the modified code.  To see complex-step in a more narrow context,
-it is advised that readers also take a look at the examples of complex-step and
-the other methods provided on the corresponding author's website.
+\emph{Peridigm} was primarily written by expert computational scientists; therefore, many advanced techniques in C++ such as multiple virtual inheritance, pointer arithmetic, and distributed data structures were used in the basic code and by necessity in the modified code.
+To see CS and other methods in a more narrow context, it is advised that readers also take a look at the simple examples provided on the corresponding author's website.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[tbp]
   \centering

### 2013-10-27 jacobian_calc_CMAME.tex
Fix author block

@@
 \usepackage{pgfplots}
+\usepackage{siunitx}
 
@@
 \author{John T. Foster\corref{cor1}}
-\cortext[cor1]{Corresponding Author: john.foster@utsa.edu}
-\author{Harry R. Millwater}
+\ead{john.foster@utsa.edu}
+\cortext[cor1]{Corresponding Author}
+
+\author{Harry R. Millwater\corref{}}
 \address{Mechanical Engineering Department, The University of Texas at San Antonio}
@@
 are derived in-paper with an attempt made to lay steps out plainly for those in disparate fields
-or who are new to their studies.  \end{abstract}
+or who are new to their studies.  
+\end{abstract}
 
@@
 %% or \MSC[2008] code \sep code (2000 is the default)
-\MSC [2010] 65D25 
+%\MSC [2010] 65D25 
 \end{keyword}
@@
 
-The methods were compared by running two sets of test problems where each method would solve a
-problem concurrently. Each test problem simulated a $4$ by $.25$ by $.25$ meter block of metal
-undergoing tension along the axis parallel to the long dimension of the block. The test problems
-were set up in \emph{Perdigm} as quasistatic equillibrium problems, where the displacement boundary
-conditions used to apply tension were applied gradually, in 'load steps', and an equilibrium
-solution for each load step was achieved before applying the next load step. The purpose of applying
-the load in gradual steps is a widely used technique to ensure the Newton-Raphson method is always
-guessing near the actual solution. The material properties of the model were set at $8.0E-9$ $ metric
-tonne/ mm^3$ density, $1.515E4$ $MPa$ bulk modulus, and $7.813E4$ $MPa$ shear modulus. The peridynamic
-horizon used in a test problem was three times the node spacing for the corrisponding model mesh.
-This value for horizon was chosen out of experience to insure that every node could have at least
-one nonlocal connection along the diagonal of the node grid.
+The methods were compared by running two sets of test problems where each method would solve a problem concurrently.
+Each test problem simulated a $\SI{4}{\meter}$ by $\SI{0.25}{\meter}$ by $\SI{0.25}{\meter}$ meter block of a material undergoing tension along the axis parallel to the long dimension of the block.
+The test problems were set up in \emph{Perdigm} as quasistatic equillibrium problems, where the displacement boundary conditions used to apply tension were applied gradually, in `load steps', and an equilibrium solution for each load step was achieved before applying the next load step.
+The purpose of applying the load in gradual steps is a widely used technique to ensure the Newton's method is always initialized near the actual solution and therefore likely to converge.
+The material properties of the model were meant to replicate a typical metal and had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus.
+The peridynamic horizon, a length scale that effectively sets the bandwidth of tangent-stiffness in the computation, used in the test problems was always set to three times the nominal discretization size (i.e.
+node spacing in the particle discretiztion scheme).
+This results in a minimum tangent-stiffness matrix bandwidth of 7 (for an opitimal node numbering scheme).
 

### 2013-10-27 jacobian_calc_CMAME.tex
Change name of Metics section to qoi's

@@
 \begin{equation}
-f ( \beta_n + i h ) = f (\beta_n) + \frac{\partial f( \beta_n )}{\partial \beta_n} \frac{i h}{1!} + \frac{\partial f^2(\beta_n)}{\partial \beta_n^2} \frac{(i h)^2}{2!} + \frac{\partial f^3(\beta_n)}{\partial \beta_n^3} \frac{(i h)^3}{3!} + \cdots
+f ( x + i h ) = f (x) + \frac{\partial f(x)}{\partial x} \frac{i h}{1!} + \frac{\partial f^2(x)}{\partial x^2} \frac{(i h)^2}{2!} + \frac{\partial f^3(x)}{\partial x_n^3} \frac{(i h)^3}{3!} + \cdots
 \label{eqn:complexTaylor}
@@
 \begin{equation}
- \frac{\partial f( \beta_n )}{\partial \beta_n} \approx \frac{\mbox{Im} \left( f (\beta_n + i h) \right)}{h}.
+ \frac{\partial f( x )}{\partial x} \approx \frac{\mbox{Im} \left( f (x + i h) \right)}{h}.
 \label{eqn:complexFirstDeriv}
@@
 The purpose of applying the load in gradual steps is a widely used technique to ensure the Newton's method is always initialized near the actual solution and therefore likely to converge.
-The material properties of the model were meant to replicate a typical metal and had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus.
+The material properties of the model were meant to replicate a typical metal and had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus\note[JTF]{Micheal, are these numbers correct?}.
 The peridynamic horizon, a length scale that effectively sets the bandwidth of tangent-stiffness in the computation, used in the test problems was always set to three times the nominal discretization size (i.e.
@@
 
-A series of nine of these single core test problems were run, with the mesh discretization used in
-the model finer and finer with every test in the series. The aim of increasing the mesh complexity
-was to see if possible differences between the methods in terms of accuracy and speed would arise
-and trend based on model complexity. This series would comprise the single core runs. The specific
-parameters of these tests can be found in \emph{Peridigm} input xml files included in the archive
-for this paper which can be found in the corresponding author's website. These xml files allow users
-with appropriately modified versions of \emph{Peridigm} to reproduce the results found in this
-paper.
+A series of nine single compute core (i.e.
+serial) test problems were run with the discretization level being refined with every test in the series.
+The aim of increasing the refinement was to see if possible differences between the methods in terms of accuracy and speed would arise.
+This series would comprise the single core runs.
+The specific parameters of these tests can be found in \emph{Peridigm} input {\tt xml} files included in the archive for this paper which can be found in the corresponding author's website.
+These {\tt xml} files allow users with appropriately modified versions of \emph{Peridigm} to reproduce the results found in this paper.
 
-Another series of four test problems were run, however this time the number of cores used to solve
-the problem was varied from test to test rather than the mesh refinement. While mesh refinement was
-not varied from test to test, it was set at a somewhat high $1E6$ perdidynamic nodes of
-discretization. This allowed the simulation computer to be challenged sufficiently that the effect
-of varying the number of cores on speed and accuracy between the methods used could be more clearly
-seen. The specific parameters of these tests can be found in \emph{Perdigm} input xml files also
-included in the archive for this paper on the corresponding author's website.
+Another series of four test problems were simulated, however this time the number of compute cores used to solve the problem was increased from test to test while the discretization level was held constant at 1 million computational nodes (3 million degrees of freedom).
+This allowed for sufficient computational nodes per core even at the highest level of parallelization such that message passing compunication did not overwhelm the simulation.
+The specific parameters of these tests can be found in \emph{Perdigm} input {\tt xml} files also included in the archive for this paper on the corresponding author's website.
 
-What was recorded during the test runs as experimental data is the subject the following
-subsubsection.
+\subsubsection{Quantities of interested} 
+\label{JGAM} 
+As stated in the Introduction a goal of the study was to rank CS against FD and CD on the basis of accuracy.
+AD was ommited from the comparison because it served as the standard of accuracy for the other methods in the absence of appropriate analaytical forms for the Jacobian associated with the system solved in the study.
+The assumption that AD is accurate enough to serve as a standard is supported by ADs implementation as a computerized chain rule as explained in Section~\ref{ADsubsection}.
 
-\subsubsection{Justification: Goals, Assumptions, and Metrics} \label{JGAM} As stated in the
-previous section a goal of the study was to rank complex-step against forward-difference and central
-difference on the basis of accuracy. AD was ommited from the comparison because it served as the
-standard of accuracy for the other methods in the absence of appropriate analaytical forms for the
-Jacobian associated with the nonlinear system solved in the study. The assumption that AD is
-accurate enough to serve as a standard is supported by AD's implementation as a computerized chain
-rule as explained in \ref{ADsubsection}.
-
-Because it makes no sense to compare tangent-stiffness matrices from different problems, load steps
-that started with difference guesses, or from different iterations, it was necessary to solve one
+Because it makes no sense to compare tangent-stiffness matrices from different problems, load steps that started with difference guesses, or from different iterations, it was necessary to solve one
 load step and conclude each iteration within that load step by updating the displacement guess only

### 2013-10-28 jacobian_calc_CMAME.tex
More changes to qoi section

@@
 
-\subsubsection{Quantities of interested} 
+\subsubsection{Quantities of interest} 
 \label{JGAM} 
-As stated in the Introduction a goal of the study was to rank CS against FD and CD on the basis of accuracy.
-AD was ommited from the comparison because it served as the standard of accuracy for the other methods in the absence of appropriate analaytical forms for the Jacobian associated with the system solved in the study.
+As stated in the Introduction, a goal of the study was to compare CS, FD, and CD on the basis of accuracy.
+AD was ommited from the comparison because it served as the standard of accuracy for the other methods in the absence of appropriate analaytical forms for the tangent-stiffness matrix associated with the system solved in the study.
 The assumption that AD is accurate enough to serve as a standard is supported by ADs implementation as a computerized chain rule as explained in Section~\ref{ADsubsection}.
 
-Because it makes no sense to compare tangent-stiffness matrices from different problems, load steps that started with difference guesses, or from different iterations, it was necessary to solve one
-load step and conclude each iteration within that load step by updating the displacement guess only
-from the results of the AD method and to subsequently feed all four methods the same updated guess
-the following iteration. This decision precluded a comparison of convergence rate, since if the
-methods were allowed to solve a problem at their individual pace, differences in accuracy would
-produce differences in guess updates and therefor the number and nature of iterations perfomed.
-Different guesses from iterations started with different previous guesses could not be data for a
-valid comparison of tangent-stiffness accuracy between methods.  However a separate convergence rate
-comparison study could be done given some modifications to the materials for the study described in
-this paper, and of course, the resources to perform that study.   Additionally, having each of the
-methods physically operate in the same process, single or one of associated parallel, allowed the
-comparison of tangent-stiffness matrices as loaded from random access memory rather than from the
-hard-disk.  Loading from RAM gives the benefit of vastly greater speed and the simpler programming
-compared to some other solution speculatively involving dynamic file management on files which for
-one of the components of the study would be on the order of $1E10$ bytes. However, the price of
-running the methods together in the same process and avoiding the use of the hard-disk was that at
-least two tangent-stiffness matrices had to be stored in RAM during simulation, which meant that
-special high-memory compute nodes were needed for the $1E6$ peridynamic element problem series.
+It would be a poor comparative study to compare tangent-stiffness matrices from different problems, load steps that start with different current configurations, or from different iterations; because of this, it was necessary to solve one
+load step and conclude each Newton iteration within that load step by updating the displacement iterate with only the results of the AD method and to subsequently feed all four methods the same updated displacement in the following iteration.
+This decision precluded a comparison of Newton iteration convergence rate, since if the methods were allowed to solve a problem at their individual pace, differences in accuracy would produce differences in guess updates and therefore the number and nature of Newton iterations performed (e.g.
+lower accuracy predictions of the algorithmically correct tangent-stiffness could cause a loss of quadratic convergence).
+Different guesses from iterations started with different previous guesses could not be data for a valid comparison of tangent-stiffness accuracy between methods.
+Additionally, having each of the methods physically operate in the same process, serial or one of associated parallel, allowed the comparison of tangent-stiffness matric calculation time as cached from random access memory rather than from the hard-disk.
+Caching from RAM gives the benefit of vastly greater speed and the simpler programming compared to some other solution speculatively involving dynamic file management on files which for one of the components of the study would be on the order of 10 terabytes in size.
+However, the price of running the methods together in the same process and avoiding the use of the hard-disk was that at least two tangent-stiffness matrices had to be stored in RAM during the simulation, which meant that special high-memory compute nodes were needed for the 1 million peridynamic node test series.
 
-The other goal was to rank the four methods, that is including AD, on the basis of speed. Instead of
-convergence rate being measured, speed of iteration was measured because it could be so done at the
-same time as accuracy was being measured given a single tangent-stiffness calculation.  It was
-assumed that the order that each method was evaluated was unimportant, that evaluating each method
-sucessively within each solver iteration did not affect their performance individually and that
-speed of computation did not change with time. These assumptions allowed the test program to run the
-same problem with each of the methods at effectively the same time and generate an equal volume of
-data from each method. It was also assumed that the Jacobian matrix calculation time for each of the
-methods did not vary based on what values the independent variables held, as they do from iteration
-to iteration as the guessed equilibrium solution is updated.  This assumption allowed calculation
-time measurements to be averaged over all iterations within a solution attempt. The purpose of
-averaging calculation time measurements was to informally address the extraneous variables of
-evaluation order and computer system load due to other user's processes or processes not associated
-with the study. The question is begged: if it is assumed that the value of independent variables
-had no effect on iteration speed, then why couldn't each of the method have been run independently
-so that convergence data were recorded, so long as system being solved was the same? The answer is
-that this would have made accuracy comparisons invalid if measured concurrently with convergence
-rate, while it would cost additional time and resources if convergence rate were measured in
-serparate runs. Instead, while not part of the study, the example programs available on the
+The other goal was to compare the four methods, that is including AD, on the basis of speed.
+Instead of convergence rate being measured, speed of iteration was measured because it could be so done at the same time as accuracy was being measured given a single tangent-stiffness calculation.
+It was assumed that the order that each method was evaluated was unimportant, that evaluating each method sucessively within each solver iteration did not affect their performance individually and that speed of computation did not change with time.
+These assumptions allowed the test program to run the
+same problem with each of the methods at effectively the same time and generate an equal volume of data from each method.
+It was also assumed that the tangent-stiffness matrix calculation time for each of the methods did not vary based on what values the independent variables held, as they do from iteration to iteration as the Newton solution iterate is updated.
+This assumption allowed calculation time measurements to be averaged over all iterations within a solution attempt.
+The purpose of averaging calculation time measurements was to informally address the extraneous variables of evaluation order and computer system load due to system processes not associated with the study.
+The question is begged: if it is assumed that the value of independent variables had no effect on iteration speed, then why couldn't each of the method have been run independently so that convergence data were recorded, so long as system being solved was the same?
+The answer is that this would have made accuracy comparisons invalid if measured concurrently with convergence
+rate, while it would cost additional time and resources if convergence rate were measured in serparate runs. Instead, while not part of the study, the example programs available on the
 corresponding author's website allow the reader to make a comparison of the methods for themselves

### 2013-10-28 jacobian_calc_CMAME.tex
More qoi section fixes

@@
 same problem with each of the methods at effectively the same time and generate an equal volume of data from each method.
-It was also assumed that the tangent-stiffness matrix calculation time for each of the methods did not vary based on what values the independent variables held, as they do from iteration to iteration as the Newton solution iterate is updated.
+It was also assumed that the tangent-stiffness matrix calculation time for each of the methods did not vary based on what values the independent variables held, as they will from iteration to iteration as the Newton solution iterate is updated.
 This assumption allowed calculation time measurements to be averaged over all iterations within a solution attempt.
-The purpose of averaging calculation time measurements was to informally address the extraneous variables of evaluation order and computer system load due to system processes not associated with the study.
-The question is begged: if it is assumed that the value of independent variables had no effect on iteration speed, then why couldn't each of the method have been run independently so that convergence data were recorded, so long as system being solved was the same?
-The answer is that this would have made accuracy comparisons invalid if measured concurrently with convergence
-rate, while it would cost additional time and resources if convergence rate were measured in serparate runs. Instead, while not part of the study, the example programs available on the
-corresponding author's website allow the reader to make a comparison of the methods for themselves
-on the basis of accuracy and convergence rate for two example nonlinear systems. In the examples,
-notice how convergence rate decreases when the step size selected for forward-difference or
-complex-step is set to larger numbers such that accuracy relative to AD is decreased.
+The purpose of averaging calculation time measurements was to informally address the extraneous variables of parallel evaluation order and computer system load due to system processes not associated with the study.
+While not part of the study, the example programs available on the corresponding author's website allow the reader to make a comparison of the methods for themselves on the basis of accuracy and convergence rate for two example nonlinear systems.
+In the examples, one will notice that the forward-difference method will lose quadratic convergence when the step size selected too large such that accuracy relative to AD is decreased.
 

### 2013-10-28 jacobian_calc_CMAME.tex
Start work on Results section

@@
 The other goal was to compare the four methods, that is including AD, on the basis of speed.
-Instead of convergence rate being measured, speed of iteration was measured because it could be so done at the same time as accuracy was being measured given a single tangent-stiffness calculation.
-It was assumed that the order that each method was evaluated was unimportant, that evaluating each method sucessively within each solver iteration did not affect their performance individually and that speed of computation did not change with time.
-These assumptions allowed the test program to run the
-same problem with each of the methods at effectively the same time and generate an equal volume of data from each method.
-It was also assumed that the tangent-stiffness matrix calculation time for each of the methods did not vary based on what values the independent variables held, as they will from iteration to iteration as the Newton solution iterate is updated.
-This assumption allowed calculation time measurements to be averaged over all iterations within a solution attempt.
+Speed is defined as the total compute-time of one iteration of a Newton step which includes the tangent-stiffness calculation and parallel assembly.
+The speed of iteration was measured because it could be so done at the same time as accuracy was being measured given a single tangent-stiffness calculation.
+It was assumed that the order that each method was evaluated in parallel was unimportant, that evaluating each method sucessively within each solver iteration did not affect their performance individually and that speed of computation did not change with time.
+These assumptions allowed the test program to run the same problem with each of the methods at effectively the same time and generate an equal volume of data from each method.
+It was also assumed that the tangent-stiffness matrix calculation time for each of the methods did not vary based on the current configuration, as it changes slightly from iteration to iteration as the Newton solution iterate is updated.
+This assumption allowed calculation time measurements to be averaged over all iterations within a single load step.
 The purpose of averaging calculation time measurements was to informally address the extraneous variables of parallel evaluation order and computer system load due to system processes not associated with the study.
 While not part of the study, the example programs available on the corresponding author's website allow the reader to make a comparison of the methods for themselves on the basis of accuracy and convergence rate for two example nonlinear systems.
-In the examples, one will notice that the forward-difference method will lose quadratic convergence when the step size selected too large such that accuracy relative to AD is decreased.
+In these examples, one will notice that the finite difference methods will lose quadratic convergence when the step size selected too large such that accuracy relative to AD is decreased.
 
-Once the selection of what to measure was determined, the task was to find a way to take measurements
-that satisfied our goals. The goals of collecting accuracy and speed measurements were achieved by
-developing and implementing metrics within the simulation program used in the study. The metric used
-to measure the accuracy of a tangent-stiffness matrix was the Frobenius norm of the element-wise
-difference between the tangent-stiffness matrix produced by the method being evaluated and the
-tangent-stiffness matrix produced by the AD based method, given that both methods were set upon the
-same problem. The lower the value of this metric, the closer the other method's Jacobian was to the
-AD Jacobian, and therefore the more accurate the method was. The expression for the accuracy metric:
+Once the quantities-of-interest (QoIs) for comparison were determined, the task was to find a way to take measurements that satisfied our goals.
+The goals of collecting accuracy and speed measurements were achieved by developing and implementing metrics within the simulation program used in the study.
+The metric used to measure the accuracy of a tangent-stiffness matrix was the Frobenius norm (analgous to an $l^2$-norm except defined on a matrix) of the element-wise difference between the tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix produced by the AD based method (i.e., the \emph{exact} derivitives), given that both methods were set upon the same problem.
+The lower the value of this metric, the closer the tangent-stiffness of the method being evaluted (i.e, CS, AD, FD)  was to the AD Jacobian, the more accurate the method was.
+The expression for the accuracy metric is
+%
+\begin{equation} 
+  D = \sqrt{\sum_i \sum_j(K^{AD}_{ij} - K^{M}_{ij})^2}
+\end{equation}
+%
+Where $D$ is distance, $K^{AD}$ is the tangent-stiffness matrix produced by the AD based method, $M$ is replaced with either CS, FD, or CD depending on the method being compared.
+The metric used to compare the speed of the different methods was the time in seconds required to calculate the tangent-stiffness matrix\note[JTF]{Verify the correctness of the this statement.
+What was here previously was too wordy}.
+A final basis of comparison used in the study called \emph{computational efficiency} is based on the ratio of calculation time to the total number of tangent-stiffness matrix elements.
 
-\begin{equation} D = \sqrt{\sum_{i=0}^{n-1}\sum_{j=0}^{n-1} (J_{AD}(i,j) - J_{M}(i,j))^2}
-\end{equation}
+\subsubsection{Computers and other software} 
+%
+Discritization for the test problems run in the study was done using the software package \emph{Cubit} developed at Sandia National Laboratories \cite{ref-Cubit}.
+\emph{Cubit} generates a finite element mesh that is internally converted to a ``particle'' discritization internally when calling \emph{Peridigm}, its important to node that the number of computational nodes in \emph{Peridigm} correspond to the number of finite elements in the discritization, not the number of finite element nodes.
+An example journal script can be found on the corresponding author's website.
 
-Where $D$ is distance, n is the number of degrees of freedom, $J_{AD}$ is the Jacobian matrix
-produced by the AD based method, M stands for other \emph{M}ethod, and the root of the summed
-squared element-wise differences between the matrices is taken.  The metric used to compare the
-speed of the methods was the time in seconds taken to resume execution in the calling method after
-instructing the implementation of the method being evaluated to run and complete itself. Another
-basis of comparison used in the study, but not quite a metric, called computational efficiency is
-based on the ratio of calculation time to number of Jacobian matrix elements.
+The test problems were run the \emph{Stampede} HPC cluster computer housed at \emph{TACC} (Texas Advanced Computing Center) at The University of Texas at Austin.
+The single core test series was run using the {\tt normal} queue compute nodes, while the parallel test series was run using one to four {\tt large memory} nodes each having one terrabyte of local RAM.
+This large amount of memory was required for reasons explained in Section~\ref{JGAM}.
 
-\subsubsection{Materials: Computers, Other Software}
-
-Mesh generation for the test problems run in the study was done using the software package
-\emph{Cubit} developed at Sandia National Labs \cite{ref-Cubit}. For the test series in which peridynamic node density was
-varied, journal scripts compatible with \emph{Cubit} were used. The rule used to select the
-arangement of nodes within the borders designated by simulated material block's dimensions was to
-manually choose a node spacing, fit the maximum whole number of nodes in a regular grid within the
-block, and then check whether or not this number of nodes matched the number of nodes desired for a
-particular test. The same process was used for the multicore series. An example journal script can
-be found on the corresponding author's website. For future work, it may be beneficial to automate
-and randomize this process for time efficiency and to assist statistical analysis.
-
-The test problems were run the \emph{Stampede} HPC cluster computer housed at \emph{TACC} (Texas
-Advanced Computing Center) based out of UT Austin. The single core test series was run using the
-normal queue compute nodes, while the multicore test series was run using one to four "large memory"
-nodes each having one terrabyte of local RAM. This large amount of memory was required for reasons
-explained in \ref{JGAM}.
-
-\subsubsection{Methods: Reduction of Data} The output produced by \emph{Perdigm} including the
-additional accuracy and speed measurements it was modified to provide was redirected from the
-console to text files by the resource manager, \emph{SLURM} (Simple Linux User Resource Manager).
+\subsubsection{Data Reduction} 
+The output produced by \emph{Perdigm} including the additional accuracy and speed measurements was redirected from the console to text files by the resource manager, \emph{SLURM} (Simple Linux User Resource Manager).
 Directories and filenames were chosen so that data would be easy to sort by individual test run.
-Extraneous information was manually deleted from copied versions of the output text files so that
-they would be regular enough to be easily read by scripts written in \emph{Python}. These scripts
-averaged accuracy or speed data for a test corresponding to a particular peridynamic node density
-over all iterations for that run, and then plotted this averaged data as if it were a function of
-peridynamic node density as it varied from test to test in the series.  Similarly, for the multicore
-tests, data was averaged and plotted in much the same way, but as a funciton of number of cores used
-to solve the problem rather than as a function peridynamic node density which was held constant.
+Extraneous information was manually deleted from copied versions of the output text files so that they would be regular enough to be easily read by scripts written in \emph{Python}.
+These scripts averaged accuracy or speed data for a tests corresponding to a particular peridynamic node density over all iterations for that run, and then plotted this averaged data  as a function of peridynamic node density as it varied from test to test in the series.
+Similarly, for the parallel tests, the data was averaged and plotted in much the same way, but as a function of number of compute cores used to solve the problem rather than as a function peridynamic node density which was held constant.
 These data reduction and plotting scripts can be found on the corresponding autor's website.
 
-Since samples were not repeated, because there was not resources for it, confidence intervals were
-not calculated. However, if results of iterations within the same test problem were considered
-samples of the same thing, then a confidence interval could be calculated. This assumption was only made
-for the purpose of reporting average values on the plots. But, a cursory review of data showed no 
-range overlap among the methods for accuracy or speed data from any test run of either series.
+Since samples were not repeated, because there was not resources for it, confidence intervals were not calculated.
+However, if results of iterations within the same test problem were considered samples of the same thing, then a confidence interval could be calculated.
+This assumption was only made for the purpose of reporting average values on the plots.
+But, a cursory review of data showed no range overlap among the methods for accuracy or speed data from any test run of either series.
+\note[JTF]{Revise this paragraph.}
 
 \section{Results and Discussion}
-
-This section contains the averaged results from the study presented in tabular and graphic format.
-Table \ref{tab:results} shows the core count, number of nonzero Jacobian matrix elements, load step
-average calculation time and load step average accuracy. The units for accuracy are derived from the
-expression used to calculate the Jacobian matrix elements and the units used to define bulk material
-properties, mentioned in \ref{tcs}.
-
+%
+This section contains the average results from the study presented in tabular and graphic format.
+Table~\ref{tab:results} shows the compute core count, number of nonzero tangent-stiffness (TS) non-zero matrix elements, load step average calculation time and load step average accuracy.
+The units for accuracy are derived from the
+expression used to calculate the TS matrix elements and the units used to define bulk material properties, mentioned in Section~\ref{tcs}.
+%
+\note[JTF]{Michael, work on cleaning this table up for legibility}
 \begin{table}    
@@
     \begin{tabular}{c c c c }
-    Cores & Jacobian Els. & Calc. Time (s) & Accuracy Diff. (MPa)\\ 
+      Cores & TS Elements & Calc. Time (s) & Accuracy Diff. ($\si{\mega\pascal}$)\\ 
     \multicolumn{2}{c}{} & CS, CD, FD, AD & CS, CD, FD \\ \hline 
-	1 & 1.99E6  & 3.5, 2.7, 1.7, 1.6 & 1.92E-10, 1.21E-4, .137 \\ \hline
+    1 & $\num{1.99E6}$  & 3.5, 2.7, 1.7, 1.6 & 1.92E-10, 1.21E-4, .137 \\ \hline
 	1 & 4.25E6  & 6.2, 4.9, 3.2, 2.9 & 2.28E-10, 9.94E-4, .148 \\ \hline
@@
 
-The data presented in \ref{tab:results} will now be shown in graphic form alongside some discussion
-of results and trends.
-
 \subsection{Speed Data}
-Speed measurements were taken according to the methods mentioned in \ref{tcs}. The averaged results
-of these speed measurments for the single core tests can be seen in \ref{fig:serial_speed}.
-Average time in seconds to compute a Jacobian matrix element is plotted against number of nonzero
-Jacobian matrix ekements.
-Complex-step turned out to be the slowest of the four methods, taking the longest average time per
-iteration. The next slowest method was central-difference, followed by forward-difference and
-automatic-differentiation. Because the x axis of \ref{fig:serial_speed} is on a log scale, it
-bears mentioning that while visually it may appear that the trend indicated by the progression of
-datapoints on the plot is non-linear, the relationship would be closer to linear. This bodes well
-for the scalability of each of the methods by themselves in \emph{Peridigm}.
-
+Speed measurements were taken according to the methods mentioned in Section~\ref{tcs}.
+The averaged results of these speed measurments for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
+Average time in seconds to compute a TS element is plotted against number of nonzero TS elements.
+CS turned out to be the slowest of the four methods, taking the longest average time per iteration.
+The next slowest method was CD, followed by FD and AD.
+Because the abcissa of Fig.~\ref{fig:serial_speed} is on a log scale, it bears mentioning that while visually it may appear that the trend indicated by the progression of datapoints on the plot is non-linear, the relationship would be closer to linear \note[JTF]{???
+Revise}.
+This bodes well for the scalability of each of the methods by themselves in \emph{Peridigm}.
+%
 \begin{figure}[tbp]
@@
 \end{figure}
-
-For the multicore tests, averaged speed results appear in \ref{fig:multi_speed}. For this series,
-rather than varying the number of nonzero Jacobian matrix elements the number of cores used to
-solve the problem is varied, while the number of nonzero Jacobian matrix elements remained constant.
-Complex-step again was the slowest method, followed by central-difference, automatic-differentiation
-and then forward-difference, taking the title of fastest method for this group of tests.
-\ref{fig:multi_speed} indicates that when looking at average in-iteration speed, each of the
-methods scale linearly with the number of cores used for the given conditions of this group of
+%
+For the parallel simulations, averaged speed results appear in Fig.~\ref{fig:multi_speed}.
+For this series, rather than varying the number of nonzero TS matrix elements the number of compute cores used to solve the problem is increased, while the number of nonzero TS matrix elements remained constant.
+CS again was the slowest method, followed by central-difference, automatic-differentiation and then forward-difference.
+There is a somewhat surprisong refersal of FD and AD for these simulations.
+Figure~\ref{fig:multi_speed} indicates that when looking at average iteration time, each of the methods scale linearly with the number of cores used for the given conditions of this group of
 tests. Again, these results show that each of the methods as implemented in \emph{Peridigm} scale

### 2013-10-28 jacobian_calc_CMAME.tex
Finished first draft edits

@@
 \usepackage{siunitx}
+%\usepackage{hyperref}
 
@@
 \begin{abstract} %% Text of abstract Shown is a retrospective comparative study of tangent-stiffness
-matrix calculation methods including a newly developed Complex Taylor Series Expansion
-(CTSE)-based ``complex-step'' method alongside established methods: forward-difference,
-central-difference and automatic differentiation.  Tangent-stiffness matrix is a term from
-computational mechanics which here refers to the Jacobian matrix as used in first-order
-Newton-Raphson methods for the solution of non-linear algebraic systems of equations developed
-to describe physical or theoretical systems studied in math, science, engineering and business.
-To perform the comparative study, the above tangent-stiffness calculation methods were applied
-in a massively parallel computational peridynamics code developed at Sandia National Labs,
-called \emph{Peridigm}. In the comparative study, for each datum for each run, complex-step was
-multiple orders of magnitude more accurate the finite-difference methods, according to a common
-comparison made to the automatic-differentiation method datum. However, for the implementation
-contemporary to the study, complex-step was also the slowest method among the four for each
-datum for each run. A mathematical definition of complex-step and
-and automatic differentiation, description of the computer implementations and justification
-of the comparative study precede results for clarity. The intended audience of this paper
-includes researchers, professionals and students, therefore readers are directed to
-selected background sources in the text for well-known concepts while less familiar concepts
-are derived in-paper with an attempt made to lay steps out plainly for those in disparate fields
-or who are new to their studies.  
+In order to maintain the quadratic convergence properties of the first-order Newton's method in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
+For an extremely small class of nonlinear material models, these consistent tangent-stiffness operators can be derived analytically; however, most often in practice, they are found through numerical approximation of derivatives.
+This paper compares four methods of numerical derivative calculation: automatic differentiation, complex-step, forward finite difference, and central finite difference, in the context of tangent-stiffness matrix calculation in a massively parallel computational peridynamics code.
+The methods are compared through \emph{in situ} profiling of the code for accuracy, speed, efficiency, and parallel scalability.
+The research provides data that can serve as practical guide for code developers and analysts faced with choseing which method best suit the needs of their application code.
+Additionally, motivated by the reproducable research movement, all the of the code, examples, and workflow to regenerate the data and figures in this paper are provided open-source.
 \end{abstract}
@@
 The averaged results of these speed measurments for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
-Average time in seconds to compute a TS element is plotted against number of nonzero TS elements.
-CS turned out to be the slowest of the four methods, taking the longest average time per iteration.
-The next slowest method was CD, followed by FD and AD.
+Average time in seconds to compute a TS element is plotted as a function of the number of nonzero TS elements.
+CS turned out to be the slowest of the four methods, taking the longest average time per iteration,  followed by CD, FD and AD in that order.
 Because the abcissa of Fig.~\ref{fig:serial_speed} is on a log scale, it bears mentioning that while visually it may appear that the trend indicated by the progression of datapoints on the plot is non-linear, the relationship would be closer to linear \note[JTF]{???
@@
 CS again was the slowest method, followed by central-difference, automatic-differentiation and then forward-difference.
-There is a somewhat surprisong refersal of FD and AD for these simulations.
-Figure~\ref{fig:multi_speed} indicates that when looking at average iteration time, each of the methods scale linearly with the number of cores used for the given conditions of this group of
-tests. Again, these results show that each of the methods as implemented in \emph{Peridigm} scale
-well when additional processors are used. An explanation as to why automatic-differentiation and
-forward-differences swapped places in speed ranking relative to the single core results was not
-determined.
-
+There is a somewhat surprise refersal of FD and AD for these parallel simulations.
+Figure~\ref{fig:multi_speed} indicates that when looking at average iteration time, each of the methods scale linearly with the number of compute cores used for the given conditions of this group of tests.
+Again, these results show that each of the methods as implemented in \emph{Peridigm} scale well when additional processors are used.
+%
 \begin{figure}[tbp]
@@
 \end{figure}
+%
+A speculative explaination as to why FD and AD have reversed places in the speed tests with respect to the serial results is that the AD data-structures carry a list of partial derivatives of the functional evaluations along side the functional evaluations themselves only to be processed at the end of the tangent-stiffness calculation when the derivates are requested.
+It is likely that the communication cost associated with message passing these extra terms is what causes the AD method to be slower that FD.
 
 \subsection{Accuracy Data}
-Accuracy measurements where taken according to the methods mentioned in \ref{tcs}. The averaged
-results of these accuracy measurements for the single core tests can be seen in
-\ref{fig:serial_accuracy}. Averaged Frobenius norm of the element-wise difference between a given
-Jacobian and the Jacobian produced by the automatic-differentiation method is plotted against number
-of nonzero Jacobian matrix elements. Complex-step was revealed as the most accurate of the methods
-besides automatic-differentiation, complex-step resulting in several orders of magnitude less
-difference than resulted from using the central-difference method and the most inaccurate
-forward-difference method. This accuracy ranking order was mantained for each test in the series. It
-should be pointed out compared to the L2 norms of any of the Jacobian matrices by themselves, the
-magnitude of these differences for any of the methods is relatively small. These values appear in
-the experimental data which is available on the corresponding author's website. The effect of these
-accuracy differences on convergence rate directly was not determined.
-
+Accuracy measurements where taken according to the methods mentioned in Section~\ref{tcs}\note[JTF]{Change L2 to $l^2$ in the figure label}.
+The averaged results of these accuracy measurements for the serial tests can be seen in Fig.~\ref{fig:serial_accuracy}.
+Averaged Frobenius norm (refered to as $l^2$-norm in the figure labels for conciseness) of the element-wise difference between a given TS matrix and the TS matrix produced by the AD method is plotted as a function of the number of nonzero TS matrix elements.
+CS was shown to be the most accurate of the methods when compared to AD, CS resulting in several orders of magnitude less difference than resulted from using the CD method and as expected FD method was the least accurate.
+This accuracy ranking order was mantained for each test in the series.
+It should be pointed out compared to the $l^2$-norms of any of the TS matrices by themselves, the magnitude of these differences for any of the methods is relatively small.
+These values appear in the simulation data which is available on the corresponding author's website.
+The effect of these accuracy differences on convergence rate of the Newton solver was not determined.
+Additionaly, we must point out that the accuracy of FD and CD could be improved by using a smaller probe distance, $h$; however, if $h$ is too small, then the methods can suffer from severe effects, including solution divergence, of the round-off error associated with subtracting two numbers that are very close to one another.
+The value of $h$ used in this study is default value used in \emph{Peridigm} which is hueristically derived based on computational node spacing.
+One of the beauties of the CS method is that there is no mystery as how to choose, $h$, it cannot be too small, therefore it can be set to machine epsilon and the analysis performed without the worry of issues associated with round-off error.
+%
 \begin{figure}[tbp]
@@
 \end{figure}
-
-The averaged resuts of the accuracy measurements taken for the multicore test series can be seen in
-\ref{fig:multi_accuracy}. Y axis units remain unchanged from \ref{fig:serial_accuracy}, but the x
-axis now indicated number of cores used to solve the test problem. Accuracy trends matched those for
-the single core tests.
-
+%
+The averaged results of the accuracy measurements taken for the parallel test series can be seen in Fig.~\ref{fig:multi_accuracy}\note[JTF]{Change L2 to $l^2$ and make the ordinate axis the same scale as Figure 3}.
+The ordinate axis units remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the absicca now indicats number of compute cores used to solve the test problem.
+Accuracy trends matched those for the single core tests.
+%
 \begin{figure}[tbp]
@@
 \subsection{Efficiency Data}
-A derived measurement was also used to compare the methods. The single core results are shown in
-\ref{fig:serial_efficiency}. Efficiency, calculated as the average number of seconds taken per nonzero Jacobian matrix element 
-was ploted against the number nonzero Jacobian matrix elements. The ranking order of the methods
-matched their speed ranking in \ref{fig:serial_speed}. This plot indicates that there is a small,
-yet present decrease in efficiency w.r.t number of nonzero Jacobian elements. But judging by the
-speed results, this trend in efficiency doesn't appear to have great effect in human terms for the
-mesh densities used in this test series.
-
+A derived measurement was also used to compare the methods.
+The single core results are shown in Fig.~\ref{fig:serial_efficiency}.
+Efficiency, calculated as the average number of seconds taken per nonzero Jacobian matrix element was plotted as a function of the number nonzero TS matrix elements.
+The ranking order of the methods
+matched their speed ranking in Fig.~\ref{fig:serial_speed}.
+This plot indicates that there is a small, yet present decrease in efficiency with respect to the number of nonzero TS elements.
+This trend in efficiency doesn't appear to have great effect in practical terms for the discritizations used in this test series.
+%
 \begin{figure}[tbp]
@@
 \end{figure}
-
-The multicore efficiency results are shown in \ref{fig:multi_efficiency}. Y axis is average number
-of seconds taken per nonzero Jacobian matrix element times number of cores used to solve the
-problem. The x axis in number of cores used to solve the problem. The multicore efficiency plot
-shows that automatic-differentiation was slower than forward-difference because for the given test
-problem it had a lower efficiency. Strangely automatic-differentiation appeared to gain in
-efficiency over the last few datapoints in the plot, looking at the raw data confirmed that this was
-what was measured rather than this being an error during data reduction. The reason behind this
-trend was not determined.
-
+%
+The parallel efficiency results are shown in Fig.~\ref{fig:multi_efficiency}.
+The parallel efficiency plot shows that AD was slower than FD because for the given test problem it had a lower efficiency.
+Strangely AD appears to gain in efficiency over the last few datapoints in the plot, looking at the raw data confirmed that this was what was measured rather than this being an error during data reduction.
+The reason behind this trend was not determined.
+%
 \begin{figure}[tbp]
@@
 
-\section{Conclusions and Further Work}
+\section{Conclusions}
+%
+Based on the experimental results, it appears that for users of \emph{Peridigm} there is a case to be made for using AD for reasons of performance and accuracy.
+Although, for certain parallel tests done in this study small perfamance gains over AD where shown with the FD method.
+While the results showed that CS produced more accurate tangent-stiffness matrices than CD and FD under the parameters of the tests, it was not determined whether or not this is a clear advantage of CS over FD in terms of accuracy of final predicted displacements and speed of convergence.
+For other simulation software it is obviously necessary to collect the same types measurements as decribed in this paper in order to make a design decision.
+For new developments it is recommended that combined accuracy and speed of iteration mesurments be conduted along with entirely separate convergence studies.
 
-\subsection{Thoughts on Results} Based on the experimental results, it appears that for users of
-\emph{Peridigm} there is a case to be made for using automatic-differentiation for reasons of
-performance and accuracy. Although, the greater speed of finite-difference for certain multicore
-tests done in this study may give some users pause. While the results showed that complex-step
-produced more accurate tangent-stiffness matrices than central-difference and forward-difference 
-under the parameters of the tests, it was not determined whether or not this is a clear advantage of complex-step over
-finite-difference in terms of accuracy of final predicted node positions and speed of convergence. For other
-simulation software it is obviously necessary to collect the same types measurements as decribed in
-this paper in order to make a design decision. For new developments it is recommended that combined
-accuracy and speed of iteration mesurments be conduted along with entirely separate convergence
-studies.
-
-\subsection{Final Thoughts on Complex-Step} Complex-step was shown to be highly accurate in
-calculating tangent-stiffness matrices by the standards of this study, but as it is implemented in
-\emph{Peridigm} may not be any more accurate than AD, and it is almost certainly slower for a naive
-implementation. When looking to the future of \emph{Peridigm}, Complex-step holds the advantage of
-relying on byte-copiable datatypes, while AD requires a complicated serialization and
-deserialization in order to function as byte-copiable. This is fine for copying in between MPI
-processes, but deserialization on an accelerator or GPU is very difficult, as this would require
-compiling a version of the AD library being used that were compatible with the accelerator or GPU
-being used. What this means is that if \emph{Peridigm} or any simulation software were to be
-reformulated to take advantage of accelerators or GPUs, attractive given their performance and power
-efficiency advantages over traditional CPUs, the flexibilty of complex-step owed to its simple
-implementation may make it more viable for that application.
+The CS method was shown to be highly accurate in calculating tangent-stiffness matrices by the standards of this study, and it is slower than other methods for a naive implementation.
+In the context of future use of CS in \emph{Peridigm}, CS holds the advantage of relying on byte-copiable datatypes, while AD requires a complicated serialization and deserialization in order to function as byte-copiable.
+This is fine for copying in between MPI processes, but deserialization on an accelerator or GPU is very difficult, as this would require compiling a version of the AD library being used that were compatible with the accelerator or GPU being used.
+What this means is that if \emph{Peridigm} or any simulation software were to be reformulated to take advantage of accelerators or GPUs, attractive given their performance and power efficiency advantages over traditional CPUs, the flexibilty of complex-step owed to its simple implementation may make it more viable for that application given its high accuracy.
+Additional performance gains may be achieved through the use of the templated \emph{Tpetra} package in \emph{Trilinos} which would allow the use of complex data types natively, and alievate the expense of the dynamic-casts that were utilized in this study.
+As a final comment, it should be noted that from a code development, or implementation, perspective the CS method may be the easiest of the methods to implement.
+It only requires one functional evaluation, there is no hueristic choice of probe distance, and it there is no dependence on an external AD library.
+From, this perspective, the method may still be an attractive method for TS matrix evaluation.
 
@@
 \section{Acknowledgements}
-This work was funded in part by grants from the United States Air Force Office of Scientific
-Research grant number XXXX and National Energy Technology Laboratory grant number XXXXX. 
 \label{sec:ack}
+This work was funded in part by grants from the United States Air Force Office of Scientific Research grant number XXXX \note[JTF]{Harry: Your grant number here} and National Energy Technology Laboratory grant number XXXX.
+The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.
+URL: http://www.tacc.utexas.edu
 %% References

### 2013-10-28 jacobian_calc_CMAME.tex
Spell check

@@
 The methods are compared through \emph{in situ} profiling of the code for accuracy, speed, efficiency, and parallel scalability.
-The research provides data that can serve as practical guide for code developers and analysts faced with choseing which method best suit the needs of their application code.
-Additionally, motivated by the reproducable research movement, all the of the code, examples, and workflow to regenerate the data and figures in this paper are provided open-source.
+The research provides data that can serve as practical guide for code developers and analysts faced with choosing which method best suit the needs of their application code.
+Additionally, motivated by the reproducible research movement, all the of the code, examples, and workflow to regenerate the data and figures in this paper are provided open-source.
 \end{abstract}
@@
 Only forward AD will be covered here since it is the implementation used in \emph{Sacado} and therefore this study; however, the reader is referred to the introduction section of \cite{ref-AD-methods} and its citations for further information on AD, particularly \cite{ref-on-AD} which is foundational.
-Appendix~\ref{sec:appendixB} includes an example that walks the reader through the process a computer uses to compute derivates via AD.
+Appendix~\ref{sec:appendixB} includes an example that walks the reader through the process a computer uses to compute derivatives via AD.
 Some things to note about AD are that no approximation of derivatives is being made because the analytical forms of the partials of the elementary math functions are defined alongside them.
@@
 While we've adopted the terminology of solid mechanics, it's important to note that these operators appear in other physical settings and are known by other names, e.g.
-a \emph{transmisability matrix} in the context of a Poisson problem or, more generally, a \emph{Jacobian matrix} in mathematical optimization.
+a \emph{transmissibility matrix} in the context of a Poisson problem or, more generally, a \emph{Jacobian matrix} in mathematical optimization.
 
@@
 \emph{Peridigm} is a C++ code utilizing agile software components from Sandia's  \emph{Trilinos} project \cite{trilinos}.
-It's important to note that there is no specialization of the tangent-stiffness calculation methods described previously to this particular code or more generally to a computational peridynamics approach and the analysis carried out in this study should provide insight into the speed and accruacy of these methods when implemented into other computational mechanics analysis tools as well.
-\emph{Peridigm} was chosen because of the author's relationship with the code development project and primarily because the FD, CD, and AD methods for tangent-stiffness calculation where areleady implemented in the code, leaving only the CS method to for development.
+It's important to note that there is no specialization of the tangent-stiffness calculation methods described previously to this particular code or more generally to a computational peridynamics approach and the analysis carried out in this study should provide insight into the speed and accuracy of these methods when implemented into other computational mechanics analysis tools as well.
+\emph{Peridigm} was chosen because of the author's relationship with the code development project and primarily because the FD, CD, and AD methods for tangent-stiffness calculation where already implemented in the code, leaving only the CS method to for development.
 However, because peridynamics is a nonlocal theory, the tangent-stiffness matrices have a much higher bandwidth (i.e.
@@
 However, because of \emph{Peridigm}'s reliance on \emph{Trilinos} and specifically the \emph{Epetra} package, special steps had to be followed to allow the use of complex number data types.
-This is because \emph{Epetra} vectors, which can be thought of as distributed memory parallel cousins to standard C++ STL vectors, are hard coded to be of type {\tt double} and can not be simply be declared as having a complex datatype.
-Therefore, program methods that were used in the course of applying the FD method were copied, renamed and re-written to follow the CS formula, which involved changing the datatype of intermediate variables which were locally scoped and dynamically casting persistent memory variables to complex datatypes where necessary.
-We made no atttempt to ``tune'' the CS code in any way for performance, we attempted to implement the algorithm is such as way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurt its overall performance and this issue could be alleviated by using the next generation templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any datatype including complex.
-There where additional modifications for the purpose instrumenting the code to collect comparitive data and are not specifically part of CS or other methods.
+This is because \emph{Epetra} vectors, which can be thought of as distributed memory parallel cousins to standard C++ STL vectors, are hard coded to be of type {\tt double} and can not be simply be declared as having a complex data type.
+Therefore, program methods that were used in the course of applying the FD method were copied, renamed and re-written to follow the CS formula, which involved changing the data type of intermediate variables which were locally scoped and dynamically casting persistent memory variables to complex data types where necessary.
+We made no attempt to ``tune'' the CS code in any way for performance, we attempted to implement the algorithm is such as way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurt its overall performance and this issue could be alleviated by using the next generation templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any data type including complex.
+There where additional modifications for the purpose instrumenting the code to collect comparative data and are not specifically part of CS or other methods.
 The source code patch to \emph{Peridigm} for implementing CS can be found on the corresponding author's website.
@@
 Each test problem simulated a $\SI{4}{\meter}$ by $\SI{0.25}{\meter}$ by $\SI{0.25}{\meter}$ meter block of a material undergoing tension along the axis parallel to the long dimension of the block.
-The test problems were set up in \emph{Perdigm} as quasistatic equillibrium problems, where the displacement boundary conditions used to apply tension were applied gradually, in `load steps', and an equilibrium solution for each load step was achieved before applying the next load step.
+The test problems were set up in \emph{Perdigm} as quasistatic equilibrium problems, where the displacement boundary conditions used to apply tension were applied gradually, in `load steps', and an equilibrium solution for each load step was achieved before applying the next load step.
 The purpose of applying the load in gradual steps is a widely used technique to ensure the Newton's method is always initialized near the actual solution and therefore likely to converge.
-The material properties of the model were meant to replicate a typical metal and had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus\note[JTF]{Micheal, are these numbers correct?}.
+The material properties of the model were meant to replicate a typical metal and had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus\note[JTF]{Michael, are these numbers correct?}.
 The peridynamic horizon, a length scale that effectively sets the bandwidth of tangent-stiffness in the computation, used in the test problems was always set to three times the nominal discretization size (i.e.
-node spacing in the particle discretiztion scheme).
-This results in a minimum tangent-stiffness matrix bandwidth of 7 (for an opitimal node numbering scheme).
+node spacing in the particle discretization scheme).
+This results in a minimum tangent-stiffness matrix bandwidth of 7 (for an optimal node numbering scheme).
 
@@
 Another series of four test problems were simulated, however this time the number of compute cores used to solve the problem was increased from test to test while the discretization level was held constant at 1 million computational nodes (3 million degrees of freedom).
-This allowed for sufficient computational nodes per core even at the highest level of parallelization such that message passing compunication did not overwhelm the simulation.
+This allowed for sufficient computational nodes per core even at the highest level of parallelization such that message passing computation did not overwhelm the simulation.
 The specific parameters of these tests can be found in \emph{Perdigm} input {\tt xml} files also included in the archive for this paper on the corresponding author's website.
@@
 As stated in the Introduction, a goal of the study was to compare CS, FD, and CD on the basis of accuracy.
-AD was ommited from the comparison because it served as the standard of accuracy for the other methods in the absence of appropriate analaytical forms for the tangent-stiffness matrix associated with the system solved in the study.
+AD was omitted from the comparison because it served as the standard of accuracy for the other methods in the absence of appropriate analytical forms for the tangent-stiffness matrix associated with the system solved in the study.
 The assumption that AD is accurate enough to serve as a standard is supported by ADs implementation as a computerized chain rule as explained in Section~\ref{ADsubsection}.
@@
 Different guesses from iterations started with different previous guesses could not be data for a valid comparison of tangent-stiffness accuracy between methods.
-Additionally, having each of the methods physically operate in the same process, serial or one of associated parallel, allowed the comparison of tangent-stiffness matric calculation time as cached from random access memory rather than from the hard-disk.
+Additionally, having each of the methods physically operate in the same process, serial or one of associated parallel, allowed the comparison of tangent-stiffness matrix calculation time as cached from random access memory rather than from the hard-disk.
 Caching from RAM gives the benefit of vastly greater speed and the simpler programming compared to some other solution speculatively involving dynamic file management on files which for one of the components of the study would be on the order of 10 terabytes in size.
@@
 The speed of iteration was measured because it could be so done at the same time as accuracy was being measured given a single tangent-stiffness calculation.
-It was assumed that the order that each method was evaluated in parallel was unimportant, that evaluating each method sucessively within each solver iteration did not affect their performance individually and that speed of computation did not change with time.
+It was assumed that the order that each method was evaluated in parallel was unimportant, that evaluating each method successively within each solver iteration did not affect their performance individually and that speed of computation did not change with time.
 These assumptions allowed the test program to run the same problem with each of the methods at effectively the same time and generate an equal volume of data from each method.
@@
 The goals of collecting accuracy and speed measurements were achieved by developing and implementing metrics within the simulation program used in the study.
-The metric used to measure the accuracy of a tangent-stiffness matrix was the Frobenius norm (analgous to an $l^2$-norm except defined on a matrix) of the element-wise difference between the tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix produced by the AD based method (i.e., the \emph{exact} derivitives), given that both methods were set upon the same problem.
-The lower the value of this metric, the closer the tangent-stiffness of the method being evaluted (i.e, CS, AD, FD)  was to the AD Jacobian, the more accurate the method was.
+The metric used to measure the accuracy of a tangent-stiffness matrix was the Frobenius norm (analogous to an $l^2$-norm except defined on a matrix) of the element-wise difference between the tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix produced by the AD based method (i.e., the \emph{exact} derivatives), given that both methods were set upon the same problem.
+The lower the value of this metric, the closer the tangent-stiffness of the method being evaluated (i.e, CS, AD, FD)  was to the AD Jacobian, the more accurate the method was.
 The expression for the accuracy metric is
@@
 %
-Discritization for the test problems run in the study was done using the software package \emph{Cubit} developed at Sandia National Laboratories \cite{ref-Cubit}.
-\emph{Cubit} generates a finite element mesh that is internally converted to a ``particle'' discritization internally when calling \emph{Peridigm}, its important to node that the number of computational nodes in \emph{Peridigm} correspond to the number of finite elements in the discritization, not the number of finite element nodes.
+Discretization for the test problems run in the study was done using the software package \emph{Cubit} developed at Sandia National Laboratories \cite{ref-Cubit}.
+\emph{Cubit} generates a finite element mesh that is internally converted to a ``particle'' discretization internally when calling \emph{Peridigm}, its important to node that the number of computational nodes in \emph{Peridigm} correspond to the number of finite elements in the discretization, not the number of finite element nodes.
 An example journal script can be found on the corresponding author's website.
@@
 The test problems were run the \emph{Stampede} HPC cluster computer housed at \emph{TACC} (Texas Advanced Computing Center) at The University of Texas at Austin.
-The single core test series was run using the {\tt normal} queue compute nodes, while the parallel test series was run using one to four {\tt large memory} nodes each having one terrabyte of local RAM.
+The single core test series was run using the {\tt normal} queue compute nodes, while the parallel test series was run using one to four {\tt large memory} nodes each having one terabyte of local RAM.
 This large amount of memory was required for reasons explained in Section~\ref{JGAM}.
@@
 Similarly, for the parallel tests, the data was averaged and plotted in much the same way, but as a function of number of compute cores used to solve the problem rather than as a function peridynamic node density which was held constant.
-These data reduction and plotting scripts can be found on the corresponding autor's website.
+These data reduction and plotting scripts can be found on the corresponding author's website.
 
@@
 Speed measurements were taken according to the methods mentioned in Section~\ref{tcs}.
-The averaged results of these speed measurments for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
+The averaged results of these speed measurements for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
 Average time in seconds to compute a TS element is plotted as a function of the number of nonzero TS elements.
 CS turned out to be the slowest of the four methods, taking the longest average time per iteration,  followed by CD, FD and AD in that order.
-Because the abcissa of Fig.~\ref{fig:serial_speed} is on a log scale, it bears mentioning that while visually it may appear that the trend indicated by the progression of datapoints on the plot is non-linear, the relationship would be closer to linear \note[JTF]{???
+Because the abscissa of Fig.~\ref{fig:serial_speed} is on a log scale, it bears mentioning that while visually it may appear that the trend indicated by the progression of data points on the plot is non-linear, the relationship would be closer to linear \note[JTF]{???
 Revise}.
@@
 CS again was the slowest method, followed by central-difference, automatic-differentiation and then forward-difference.
-There is a somewhat surprise refersal of FD and AD for these parallel simulations.
+There is a somewhat surprise reversal of FD and AD for these parallel simulations.
 Figure~\ref{fig:multi_speed} indicates that when looking at average iteration time, each of the methods scale linearly with the number of compute cores used for the given conditions of this group of tests.
@@
 %
-A speculative explaination as to why FD and AD have reversed places in the speed tests with respect to the serial results is that the AD data-structures carry a list of partial derivatives of the functional evaluations along side the functional evaluations themselves only to be processed at the end of the tangent-stiffness calculation when the derivates are requested.
+A speculative explanation as to why FD and AD have reversed places in the speed tests with respect to the serial results is that the AD data-structures carry a list of partial derivatives of the functional evaluations along side the functional evaluations themselves only to be processed at the end of the tangent-stiffness calculation when the derivatives are requested.
 It is likely that the communication cost associated with message passing these extra terms is what causes the AD method to be slower that FD.
@@
 The averaged results of these accuracy measurements for the serial tests can be seen in Fig.~\ref{fig:serial_accuracy}.
-Averaged Frobenius norm (refered to as $l^2$-norm in the figure labels for conciseness) of the element-wise difference between a given TS matrix and the TS matrix produced by the AD method is plotted as a function of the number of nonzero TS matrix elements.
+Averaged Frobenius norm (referred to as $l^2$-norm in the figure labels for conciseness) of the element-wise difference between a given TS matrix and the TS matrix produced by the AD method is plotted as a function of the number of nonzero TS matrix elements.
 CS was shown to be the most accurate of the methods when compared to AD, CS resulting in several orders of magnitude less difference than resulted from using the CD method and as expected FD method was the least accurate.
-This accuracy ranking order was mantained for each test in the series.
+This accuracy ranking order was maintained for each test in the series.
 It should be pointed out compared to the $l^2$-norms of any of the TS matrices by themselves, the magnitude of these differences for any of the methods is relatively small.
@@
 The effect of these accuracy differences on convergence rate of the Newton solver was not determined.
-Additionaly, we must point out that the accuracy of FD and CD could be improved by using a smaller probe distance, $h$; however, if $h$ is too small, then the methods can suffer from severe effects, including solution divergence, of the round-off error associated with subtracting two numbers that are very close to one another.
-The value of $h$ used in this study is default value used in \emph{Peridigm} which is hueristically derived based on computational node spacing.
+Additionally, we must point out that the accuracy of FD and CD could be improved by using a smaller probe distance, $h$; however, if $h$ is too small, then the methods can suffer from severe effects, including solution divergence, of the round-off error associated with subtracting two numbers that are very close to one another.
+The value of $h$ used in this study is default value used in \emph{Peridigm} which is heuristically derived based on computational node spacing.
 One of the beauties of the CS method is that there is no mystery as how to choose, $h$, it cannot be too small, therefore it can be set to machine epsilon and the analysis performed without the worry of issues associated with round-off error.
@@
 The averaged results of the accuracy measurements taken for the parallel test series can be seen in Fig.~\ref{fig:multi_accuracy}\note[JTF]{Change L2 to $l^2$ and make the ordinate axis the same scale as Figure 3}.
-The ordinate axis units remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the absicca now indicats number of compute cores used to solve the test problem.
+The ordinate axis units remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the absicca now indicates number of compute cores used to solve the test problem.
 Accuracy trends matched those for the single core tests.
@@
 This plot indicates that there is a small, yet present decrease in efficiency with respect to the number of nonzero TS elements.
-This trend in efficiency doesn't appear to have great effect in practical terms for the discritizations used in this test series.
+This trend in efficiency doesn't appear to have great effect in practical terms for the discretization used in this test series.
 %
@@
 The parallel efficiency plot shows that AD was slower than FD because for the given test problem it had a lower efficiency.
-Strangely AD appears to gain in efficiency over the last few datapoints in the plot, looking at the raw data confirmed that this was what was measured rather than this being an error during data reduction.
+Strangely AD appears to gain in efficiency over the last few data points in the plot, looking at the raw data confirmed that this was what was measured rather than this being an error during data reduction.
 The reason behind this trend was not determined.
@@
 Based on the experimental results, it appears that for users of \emph{Peridigm} there is a case to be made for using AD for reasons of performance and accuracy.
-Although, for certain parallel tests done in this study small perfamance gains over AD where shown with the FD method.
+Although, for certain parallel tests done in this study small performance gains over AD where shown with the FD method.
 While the results showed that CS produced more accurate tangent-stiffness matrices than CD and FD under the parameters of the tests, it was not determined whether or not this is a clear advantage of CS over FD in terms of accuracy of final predicted displacements and speed of convergence.
-For other simulation software it is obviously necessary to collect the same types measurements as decribed in this paper in order to make a design decision.
-For new developments it is recommended that combined accuracy and speed of iteration mesurments be conduted along with entirely separate convergence studies.
+For other simulation software it is obviously necessary to collect the same types measurements as described in this paper in order to make a design decision.
+For new developments it is recommended that combined accuracy and speed of iteration measurements be conducted along with entirely separate convergence studies.
 
 The CS method was shown to be highly accurate in calculating tangent-stiffness matrices by the standards of this study, and it is slower than other methods for a naive implementation.
-In the context of future use of CS in \emph{Peridigm}, CS holds the advantage of relying on byte-copiable datatypes, while AD requires a complicated serialization and deserialization in order to function as byte-copiable.
+In the context of future use of CS in \emph{Peridigm}, CS holds the advantage of relying on byte-copyable data types, while AD requires a complicated serialization and deserialization in order to function as byte-copyable.
 This is fine for copying in between MPI processes, but deserialization on an accelerator or GPU is very difficult, as this would require compiling a version of the AD library being used that were compatible with the accelerator or GPU being used.
-What this means is that if \emph{Peridigm} or any simulation software were to be reformulated to take advantage of accelerators or GPUs, attractive given their performance and power efficiency advantages over traditional CPUs, the flexibilty of complex-step owed to its simple implementation may make it more viable for that application given its high accuracy.
-Additional performance gains may be achieved through the use of the templated \emph{Tpetra} package in \emph{Trilinos} which would allow the use of complex data types natively, and alievate the expense of the dynamic-casts that were utilized in this study.
+What this means is that if \emph{Peridigm} or any simulation software were to be reformulated to take advantage of accelerators or GPUs, attractive given their performance and power efficiency advantages over traditional CPUs, the flexibility of complex-step owed to its simple implementation may make it more viable for that application given its high accuracy.
+Additional performance gains may be achieved through the use of the templated \emph{Tpetra} package in \emph{Trilinos} which would allow the use of complex data types natively, and alieviate the expense of the dynamic-casts that were utilized in this study.
 As a final comment, it should be noted that from a code development, or implementation, perspective the CS method may be the easiest of the methods to implement.
-It only requires one functional evaluation, there is no hueristic choice of probe distance, and it there is no dependence on an external AD library.
+It only requires one functional evaluation, there is no heuristic choice of probe distance, and it there is no dependence on an external AD library.
 From, this perspective, the method may still be an attractive method for TS matrix evaluation.

### 2013-10-28 jacobian_calc_CMAME.tex
Revised/updated unclear language, cleaned up table, made accuracy plots easier to compare, added new table column type

@@
 %\usepackage{hyperref}
+
+\usepackage{array}
 
@@
 The purpose of applying the load in gradual steps is a widely used technique to ensure the Newton's method is always initialized near the actual solution and therefore likely to converge.
-The material properties of the model were meant to replicate a typical metal and had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus\note[JTF]{Michael, are these numbers correct?}.
+The fictional material represented in the model had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus.
 The peridynamic horizon, a length scale that effectively sets the bandwidth of tangent-stiffness in the computation, used in the test problems was always set to three times the nominal discretization size (i.e.
@@
 Where $D$ is distance, $K^{AD}$ is the tangent-stiffness matrix produced by the AD based method, $M$ is replaced with either CS, FD, or CD depending on the method being compared.
-The metric used to compare the speed of the different methods was the time in seconds required to calculate the tangent-stiffness matrix\note[JTF]{Verify the correctness of the this statement.
-What was here previously was too wordy}.
+The metric used to compare the speed of the different methods was the time in seconds required to calculate the tangent-stiffness matrix.
 A final basis of comparison used in the study called \emph{computational efficiency} is based on the ratio of calculation time to the total number of tangent-stiffness matrix elements.
@@
 Similarly, for the parallel tests, the data was averaged and plotted in much the same way, but as a function of number of compute cores used to solve the problem rather than as a function peridynamic node density which was held constant.
+Confidence intervals were not calculated on the measurements.
+The main reason is that network traffic from other users makes running tests on the HPC cluster a time varying process.
 These data reduction and plotting scripts can be found on the corresponding author's website.
-
-Since samples were not repeated, because there was not resources for it, confidence intervals were not calculated.
-However, if results of iterations within the same test problem were considered samples of the same thing, then a confidence interval could be calculated.
-This assumption was only made for the purpose of reporting average values on the plots.
-But, a cursory review of data showed no range overlap among the methods for accuracy or speed data from any test run of either series.
-\note[JTF]{Revise this paragraph.}
 
@@
 \note[JTF]{Michael, work on cleaning this table up for legibility}
+\note[Michael]{Columns are now separated out and can be fine tuned}
+
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
 \begin{table}    
         \caption{Averaged Results, Each Test} \label{tab:results}   
-    \begin{tabular}{c c c c }
-      Cores & TS Elements & Calc. Time (s) & Accuracy Diff. ($\si{\mega\pascal}$)\\ 
-    \multicolumn{2}{c}{} & CS, CD, FD, AD & CS, CD, FD \\ \hline 
-    1 & $\num{1.99E6}$  & 3.5, 2.7, 1.7, 1.6 & 1.92E-10, 1.21E-4, .137 \\ \hline
-	1 & 4.25E6  & 6.2, 4.9, 3.2, 2.9 & 2.28E-10, 9.94E-4, .148 \\ \hline
-	1 & 7.79E6  & 11.2, 8.9, 5.7, 5.2 & 2.38E-10, 1.59E-4, .145\\ \hline
-	1 & 1.54E7  & 26.7, 21.0, 13.4, 12.2 & 2.33E-10, 4.61E-4, .12 \\ \hline
-	1 & 2.00E7  & 28.1, 22.2, 14.4, 13.1 & 2.76E-10, 1.05E-3, .145 \\ \hline
-	1 & 3.14E7  & 47.6, 37.7, 24.2, 21.9 & 2.64E-10, 1.65E-3, .133 \\ \hline
-	1 & 4.01E7  & 55.6, 44.1, 28.4, 25.9 & 3.03E-10, 1.92E-3, .148 \\ \hline
-	1 & 8.39E7  & 138.9, 109.6, 70.2, 64.0 & 3.63E-10, 1.64E-3, .123 \\ \hline
-	1 & 1.65E8  & 277.3, 218.1, 139.4, 126.5 & 3.26E-10, 2.18E-3, .128 \\ \hline
-	32 & 1.67E10  & 336.1, 277.7, 200.6, 233.1 & 6.21E-10, 1.52E-2, .176 \\ \hline
-	64 & 1.67E10  & 169.9, 140.7, 102.0, 119.7 & 6.20E-10, 1.50E-2, .177 \\ \hline
-	96 & 1.67E10  & 114.7, 95.0, 69.1, 79.7 & 6.18E-10, 1.50E-2, .177 \\ \hline
-	128 & 1.67E10  & 86.4, 71.8, 52.4, 58.8 & 6.16E-10, 1.47E-2, .177 \\ \hline
+       \hskip-2.0cm\begin{tabular}{C{1.5cm} C{1.88cm} C{1.5cm} C{1.5cm} C{1.5cm} C{1.5cm} C{1.88cm} C{1.88cm} C{1.5cm}}
+            Cores & TS Elements & Calc. Time (s) & & & & Accuracy Diff. ($\si{\mega\pascal}$) & & \\ 
+        \multicolumn{2}{l}{} & CS & CD & FD & AD & CS & CD & FD \\ \hline 
+        1 & $\num{1.99E6}$  & 3.5 & 2.7 & 1.7 & 1.6 & $\num{1.92E-10}$ & $\num{1.21E-4}$ & .137 \\ \hline
+        1 & $\num{4.25E6}$  & 6.2& 4.9& 3.2& 2.9 & $\num{2.28E-10}$& $\num{9.94E-4}$ & .148 \\ \hline
+        1 & $\num{7.79E6}$  & 11.2& 8.9& 5.7& 5.2 & $\num{2.38E-10}$& $\num{1.59E-4}$ & .145\\ \hline
+        1 & $\num{1.54E7}$  & 26.7& 21.0& 13.4& 12.2 & $\num{2.33E-10}$ & $\num{4.61E-4}$ & .12 \\ \hline
+        1 & $\num{2.00E7}$  & 28.1& 22.2& 14.4& 13.1 & $\num{2.76E-10}$ & $\num{1.05E-3}$ & .145 \\ \hline
+        1 & $\num{3.14E7}$  & 47.6& 37.7& 24.2& 21.9 & $\num{2.64E-10}$ & $\num{1.65E-3}$ & .133 \\ \hline
+        1 & $\num{4.01E7}$  & 55.6& 44.1& 28.4& 25.9 & $\num{3.03E-10}$ & $\num{1.92E-3}$ & .148 \\ \hline
+        1 & $\num{8.39E7}$  & 138.9& 109.6& 70.2& 64.0 & $\num{3.63E-10}$ & $\num{1.64E-3}$ & .123 \\ \hline
+        1 & $\num{1.65E8}$  & 277.3& 218.1& 139.4& 126.5 & $\num{3.26E-10}$ & $\num{2.18E-3}$ & .128 \\ \hline
+        32 & $\num{1.67E10}$  & 336.1& 277.7& 200.6& 233.1 & $\num{6.21E-10}$ & $\num{1.52E-2}$ & .176 \\ \hline
+        64 & $\num{1.67E10}$  & 169.9& 140.7& 102.0& 119.7 & $\num{6.20E-10}$ & $\num{1.50E-2}$ & .177 \\ \hline
+        96 & $\num{1.67E10}$  & 114.7& 95.0& 69.1 & 79.7 & $\num{6.18E-10}$ & $\num{1.50E-2}$ & .177 \\ \hline
+        128 & $\num{1.67E10}$  & 86.4& 71.8& 52.4 &58.8 & $\num{6.16E-10}$ & $\num{1.47E-2}$ & .177 \\ \hline
     \end{tabular}
@@
 CS turned out to be the slowest of the four methods, taking the longest average time per iteration,  followed by CD, FD and AD in that order.
-Because the abscissa of Fig.~\ref{fig:serial_speed} is on a log scale, it bears mentioning that while visually it may appear that the trend indicated by the progression of data points on the plot is non-linear, the relationship would be closer to linear \note[JTF]{???
-Revise}.
+Calculation time appears to have a linear relationship with problem size, judging by the appearance of the plot shown in Fig.~\ref{fig:serial_speed}.
 This bodes well for the scalability of each of the methods by themselves in \emph{Peridigm}.
@@
 \subsection{Accuracy Data}
-Accuracy measurements where taken according to the methods mentioned in Section~\ref{tcs}\note[JTF]{Change L2 to $l^2$ in the figure label}.
+Accuracy measurements where taken according to the methods mentioned in Section~\ref{tcs}.
 The averaged results of these accuracy measurements for the serial tests can be seen in Fig.~\ref{fig:serial_accuracy}.
@@
 %
-The averaged results of the accuracy measurements taken for the parallel test series can be seen in Fig.~\ref{fig:multi_accuracy}\note[JTF]{Change L2 to $l^2$ and make the ordinate axis the same scale as Figure 3}.
+The averaged results of the accuracy measurements taken for the parallel test series can be seen in Fig.~\ref{fig:multi_accuracy}.
 The ordinate axis units remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the absicca now indicates number of compute cores used to solve the test problem.

### 2013-10-28 data/multicore/plotter.py

@@ -75,17 +75,17 @@ cs2= ax2.scatter(accuracy_averages['CS'][0][:], accuracy_averages['CS'][1][:], m
 cd2= ax2.scatter(accuracy_averages['CD'][0][:], accuracy_averages['CD'][1][:], marker='x', c='r', s=64 )
 fd2= ax2.scatter(accuracy_averages['FD'][0][:], accuracy_averages['FD'][1][:], marker='s', c='b', s=64) 
 plt.xlim(24.0, 160.0)
-plt.ylim(1.e-10, 1.0)
+#plt.ylim(1.e-10, 1.0)
 plt.yscale('log')
 plt.xlabel("Number of cores @ 1TB RAM per 32 cores")
-plt.ylabel("L2 norm of difference to AD TS (MPa)")
+plt.ylabel("$l^2$ norm of difference to AD TS (MPa)")
 #plt.title("Frobenius norm of difference to A.D. Jacobian vs. number of cores")
 plt.legend([cs2, cd2, fd2],['CS', 'CD', 'FD', 'AD'], scatterpoints = 1, loc='center left',
                   fancybox=True, shadow=True, ncol=1)
 
 
 plt.xticks(np.arange(0.0, 160.0, 32))
-plt.yticks([1.E-10, 1.E-8, 1.E-6, 1.E-4, 1.E-2, 1.0], [1.E-10, 1.E-8, 1.E-6, 1.E-4, 1.E-2, 1.0])
+plt.yticks([1.E-10, 1.E-7, 1.E-4, 1.E-1, 1.E2], [1.E-10, 1.E-7, 1.E-4, 1.E-1, 1.E2])
 plt.grid(True)
 ax2.grid(which='major', linestyle='--', color='grey')
 

### 2013-10-28 data/singlecore/plotter.py

@@ -118,7 +118,7 @@ fd2= ax2.scatter(accuracy_averages['FD'][0][:], accuracy_averages['FD'][1][:], m
 plt.xlim(1.e5, 1.e9)
 #plt.xticks(np.arange(1.e-10, 1.0, .25))
 plt.xlabel("Number of nonzero elements in TS")
-plt.ylabel("L2 norm of difference to AD TS (MPa)")
+plt.ylabel("$l^2$ norm of difference to AD TS (MPa)")
 #plt.title("Frobenius norm of difference to AD TS vs.\nnumber of nonzero TS elements")
 plt.legend([cs2, cd2, fd2],['CS', 'CD', 'FD'], scatterpoints = 1, loc='upper left', bbox_to_anchor=(0., .38),
                   fancybox=True, shadow=True, ncol=1)

### 2013-10-29 jacobian_calc_CMAME.tex
journal files say modeled object is 4x.5X.5 rather than what I stated

@@
 The methods were compared by running two sets of test problems where each method would solve a problem concurrently.
-Each test problem simulated a $\SI{4}{\meter}$ by $\SI{0.25}{\meter}$ by $\SI{0.25}{\meter}$ meter block of a material undergoing tension along the axis parallel to the long dimension of the block.
+Each test problem simulated a $\SI{4}{\meter}$ by $\SI{0.5}{\meter}$ by $\SI{0.5}{\meter}$ meter block of a material undergoing tension along the axis parallel to the long dimension of the block.
 The test problems were set up in \emph{Perdigm} as quasistatic equilibrium problems, where the displacement boundary conditions used to apply tension were applied gradually, in `load steps', and an equilibrium solution for each load step was achieved before applying the next load step.

### 2013-11-25 transition.tex
Made springer edits CMAME comptible

--- 2013-10-29 jacobian_calc_CMAME.tex
@@
 \usepackage{siunitx}
-%\usepackage{hyperref}
-
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{booktabs}
+\usepackage{siunitx}
 \usepackage{array}
-
 %% The amssymb package provides various useful mathematical symbols
@@
 \begin{abstract} %% Text of abstract Shown is a retrospective comparative study of tangent-stiffness
-In order to maintain the quadratic convergence properties of the first-order Newton's method in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
-For an extremely small class of nonlinear material models, these consistent tangent-stiffness operators can be derived analytically; however, most often in practice, they are found through numerical approximation of derivatives.
-This paper compares four methods of numerical derivative calculation: automatic differentiation, complex-step, forward finite difference, and central finite difference, in the context of tangent-stiffness matrix calculation in a massively parallel computational peridynamics code.
-The methods are compared through \emph{in situ} profiling of the code for accuracy, speed, efficiency, and parallel scalability.
-The research provides data that can serve as practical guide for code developers and analysts faced with choosing which method best suit the needs of their application code.
-Additionally, motivated by the reproducible research movement, all the of the code, examples, and workflow to regenerate the data and figures in this paper are provided open-source.
+In order to maintain the quadratic convergence properties of Newton's method in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
+A goal of the study described in this paper was to establish the suitability of an under-explored method for numerical computation of tangent-stiffness operators, referred to as ``complex-step'', and compare the new method with other techniques for numerical derivative calculation: automatic differentiation, forward finite-difference, and central finite-difference.
+The complex-step method was newly implemented in a massively parallel computational peridynamics code for the purpose of this comparison.
+The methods were compared through in situ profiling of the code for accuracy, speed, efficiency, and parallel scalability.
+The research provides data that can serve as practical guide for code developers and analysts faced with choosing which method best suits the needs of their application code.
+Additionally, motivated by the reproducible research movement, all the of the code, examples, and workflow to regenerate the data and figures in this paper are provided as open source.
+
 \end{abstract}
@@
 \label{sec:intro} 
-
 In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
@@
 The
-distinction of `accurate' is defined by comparison of the new method to popular finite differencing techniques used for computing tangent-stiffness matrices and with the exact algorithmically consistent derivatives computed with \emph{automatic differentiation}.
+distinction of ``accurate'' is defined by comparison of the new method to popular finite differencing techniques used for computing tangent-stiffness matrices and with the exact to machine precision algorithmically consistent derivatives computed with \emph{automatic differentiation}.
 
@@
 
-The scope of the application component of the study was limited to a single material model implemented in the computational peridynamics code, \textit{Peridigm} \cite{peridigm}.
+The scope of the application component of the study was limited to a single material model, an elastic peridynamic solid, implemented in the computational peridynamics code, \textit{Peridigm} \cite{peridigm}.
 Identifying a specific application provided a practical framework for implementing the new method and solving engineering problems to generate the data needed to compare the methods.
 In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of finite-difference, central-difference, and automatic-differentiation methods needed for comparison to the new complex-step method implementation, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
+Agile-components is a software development term meaning to structure a software project for continuous development to respond to evolving specifications by making much use of techniques from object-oriented programming and reusing proven pre-existing code to reduce development time and increase reliability.
+The modularity of \emph{Peridigm} allowed the adding of new features, such as required by the study.
 
-The aim of this paper is not only to introduce the new complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques and as a general reference for solving non-linear systems with Newton's method.
-After presenting background information on the underlying methods to be discussed, presented in this paper are: a description of tangent-stiffness matrices and how to use them to solve non-linear systems, detailed directions for producing tangent-stiffness matrices with each the methods identified above, a description and justification of the new complex-step method for calculating tangent-stiffness matrices, a description of implementing complex-step in the \emph{Peridigm} software, a description of the quantities-of-interest used to rank the methods, a presentation and analysis of the results of the comparative study, and conclusions and thoughts on potential future work.
-Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for C++ source-code and data necessary to reproduce the results presented here.
-Additionally, included on the website is a C++ library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
+The aim of this paper is not only to introduce the new complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques useful for solving non-linear systems with Newton's method.
+After presenting background information on the underlying methods to be discussed, presented in this paper are: A) a description of tangent-stiffness matrices, B) detailed directions for producing tangent-stiffness matrices with each the methods identified above, C) a description and justification of the new complex-step method for calculating tangent-stiffness matrices, D) a description of implementing complex-step in the \emph{Peridigm} software, E) a description of the quantities-of-interest used to rank the methods, F) a presentation and analysis of the results of the comparative study, and G) conclusions and thoughts on potential future work.
+Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for \emph{Peridigm} C++ source-code and data necessary to reproduce the results presented here.
+Additionally, included on the website is an example serial C++ library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
+The purpose of the serial C++ library is to demonstrate in a less complex manner than in the \emph{Peridigm} code how one would use automatic-differentiation or complex-step to better encourage their use and discussion.
+This open source resource will be referred to as the \emph{paper repository} in what follows.
 
@@
 This section contains background information on the differentiation techniques underlying the tangent-stiffness matrix calculation methods compared in the study.
-Since first-order finite-difference techniques are considered well known, they are not described here.
+Since the derivation of first-order finite-difference techniques from Taylor series expansion is considered well known, it is not described here.
 Instead the reader is referred to \cite[Chap.
@@
 It is possible to approximate derivatives quite accurately with a technique based on a complex variable Taylor series expansion of a function.
-This method was first described by Lyness and Moler \cite{lyness1967numerical, lyness1968differentiation} and has more recently been rediscovered \cite{squire1998using, voorhees2011complex, al2010complex} for use in engineering analysis.
-The basic idea is that a model parameter can be be made to be complex and expanded in a Taylor series about a small perturbation in the imaginary plane by some arbitrary value $h$ as follows
+This method was first described by Lyness and Moler \cite{lyness1967numerical,lyness1968differentiation} and has more recently been rediscovered \cite{squire1998using,voorhees2011complex,al2010complex,martins2000automated} for use in engineering analysis.
+The basic idea is that a model parameter can be be made complex and expanded in a Taylor series about a small perturbation along the imaginary axis by some arbitrary value $h$ as follows
 %
-\begin{equation}
-f ( x + i h ) = f (x) + \frac{\partial f(x)}{\partial x} \frac{i h}{1!} + \frac{\partial f^2(x)}{\partial x^2} \frac{(i h)^2}{2!} + \frac{\partial f^3(x)}{\partial x_n^3} \frac{(i h)^3}{3!} + \cdots
+\begin{align}
+f ( x + i h ) =& f(x) + \frac{\partial f(x)}{\partial x} \frac{i h}{1!}  \notag \\
+               &+ \frac{\partial f^2(x)}{\partial x^2} \frac{(i h)^2}{2!} + \frac{\partial f^3(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots,
 \label{eqn:complexTaylor}
-\end{equation}
+\end{align}
 %
-Now we can take the imaginary part of both sides of equation (\ref{eqn:complexTaylor}) and solve for the first derivative ignoring terms of $\mathcal{O}\left ( h^2 \right)$ to yield an the estimate
+Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor}) and solving for the first derivative and ignoring terms of $\mathcal{O}\left ( h^2 \right)$ yields an estimate
 %
@@
 %
-This shows that an estimate of the first derivative of a function can be made by only utilizing one functional evaluation of a perturbed model parameter up the imaginary axis.
+Thus, an estimate of the first derivative of a function can be made by only utilizing one functional evaluation of a perturbed model parameter along the imaginary axis.
 The step-size $h$ is arbitrary and can be made as small as practical (even to machine precision without the dangers of roundoff error as in finite-differences) to yield an accurate estimate of this derivative.
 The only disadvantage of this technique is the necessity of requiring the functions to accept complex numbers as arguments.
-Appendix~\ref{sec:appendixA} includes a simple example that illustrates the use of the complex-step method for derivative calculation of a function.
+Appendix~\ref{sec:appendixA} includes a simple example that illustrates the use of the complex-step method for the derivative calculation of a function.
 
@@
 AD takes advantage of the fact that any mathematical function executed on a computer, no matter how complicated, is a ``composition of simple operations'' (add, multiply, power, transcendental and the like) each having known analytical derivatives \cite{ref-sacado-presentation}.
-For reference, the AD implementation used in the study is part of the \emph{Sacado} package from the \emph{Trilinos} agile component libraries developed at Sandia National Laboratories \cite{ref-Sacado}.
+For reference, the AD implementation used in the study is part of the \emph{Sacado} package from the \emph{Trilinos} agile-component libraries developed at Sandia National Laboratories \cite{ref-Sacado}.
 
-The way an AD system works is by first evaluating the innermost function of the composition, then presenting that function's output as input to the next level function until all levels are complete, in a way no different from how a normal computer program or human would evaluate composition functions.
-AD departs in that as each nested function is evaluated, the function's partial derivative with respect to the designated variables of the given input is also calculated.
+An AD system evaluates composition functions in the expected manner: it works by first evaluating the innermost function of the composition, then presenting that function's output as input to the next level function until all levels are complete, observing commonly expected order of operations.
+However, an AD system does additional work during a function evaluation in that as each nested function is evaluated, the function's partial derivative with respect to the designated variables of the given input is also calculated.
 This is possible because the elementary math functions are hard-coded into the AD source-code along with their analytical derivatives, and linked by special instructions, so that when the elementary math functions are called upon for computation, their partial derivatives may be computed and stored in a sequence.
@@
 It is obvious, but bears mentioning, that the AD system could simply store one value for partial derivatives, modifying it as appropriate for every function evaluation rather than keeping a sequence.
+This is significant because modifying a single value rather than keeping a list of values corresponding to
+every level of a composition function, which itself may not be the sole one needing to be evaluated, represents a savings in memory usage, which at large scales is an active concern.
 In the literature, the AD scheme described here is called \emph{forward automatic-differentiation}.
-Only forward AD will be covered here since it is the implementation used in \emph{Sacado} and therefore this study; however, the reader is referred to the introduction section of \cite{ref-AD-methods} and its citations for further information on AD, particularly \cite{ref-on-AD} which is foundational.
+Only forward AD will be covered here since it is the implementation used in \emph{Sacado} and therefore in this study; however, the reader is referred to the introduction section of \cite{ref-AD-methods} and its citations for further information on AD, and particularly \cite{ref-on-AD} which is foundational.
 Appendix~\ref{sec:appendixB} includes an example that walks the reader through the process a computer uses to compute derivatives via AD.
 Some things to note about AD are that no approximation of derivatives is being made because the analytical forms of the partials of the elementary math functions are defined alongside them.
-The accuracy of AD is then limited by the precision of the AD system's definition of the elementary math functions and their partial derivatives .
+The accuracy of AD is then limited by the precision of the AD system's definition of the elementary math functions and their partial derivatives.
 
@@
 Mathematically speaking, the tangent-stiffness represents the gradient of a high-dimensional energy surface that ``points'' in the minimum direction.
-While we've adopted the terminology of solid mechanics, it's important to note that these operators appear in other physical settings and are known by other names, e.g.
+While the terminology of solid mechanics is used here, it's important to note that these operators appear in other physical settings and are known by other names, e.g.
 a \emph{transmissibility matrix} in the context of a Poisson problem or, more generally, a \emph{Jacobian matrix} in mathematical optimization.
 
-While, for simplicity, the examples shown in the sequel refer only to linear problems, tangent-stiffness matrices generally arise in the context of non-linear analysis where Newton's method or quasi-Newton's method's are used in the minimization of some residual function.
+While, for simplicity, the examples shown in the following refer only to linear problems, tangent-stiffness matrices generally arise in the context of non-linear analysis where Newton's method or quasi-Newton's method's are used in the minimization of a given residual function.
 In this context, the tangent-stiffness matrix can be thought of as the linearization of the system about about a particular configuration.
-It has been shown \cite{hughes1978consistent, hughes1978unconditionally} that in order to preserve the quadratic convergence properties of the Newton's methods, it is necessary that this linearization is carried out in a manner that is \emph{consistent} with the algorithmic constraint equations used when computing the internal forces that arise due to deformations.
+It has been shown \cite{hughes1978consistent,hughes1978unconditionally} that in order to preserve the quadratic convergence properties of the Newton's methods, it is necessary that this linearization is carried out in a manner that is \emph{consistent} with the algorithmic constraint equations used when computing the internal forces that arise due to deformations.
 An example of these constraint equations would be the Kuhn-Tucker conditions \cite{simo1998} that are used in integration of a flow rule for plasticity modeling.
-If the continuum tangent-moduli are used in place of the \emph{consistent} or \emph{algorithmic tangent moduli} in the solution of a non-linear solid mechanics problem, convergence may still be achieved, but not in the quadratic manner that makes Newton's method so attractive for this class of problems.
-Again, in the setting of non-linear solid mechanics, there is a very small class of material model algorithms in which consistent tangent-stiffness operators can be derived analytically.
+If the continuum tangent-moduli are used in place of the \emph{consistent} or \emph{algorithmic tangent moduli} in the solution of a non-linear solid mechanics problem, convergence may still be achieved, but not in the quadratic manner that makes Newton's method attractive for this class of problems.
+In the setting of non-linear solid mechanics, there is a very small class of material model algorithms in which consistent tangent-stiffness operators can be derived analytically.
 Perfect plasticity and plasticity with isotropic hardening used in combination with a general nearest-point projection or \emph{radial return algorithm} are a few examples.
-Therefore, when these models (and certainly more complex ones) are implemented into a general purpose computational mechanics code, the tangent-stiffness operators are typically defined via a numerical approximation.
+Therefore, when general models (and certainly more complex ones) are implemented into a general purpose computational mechanics code, the tangent-stiffness operators are typically defined via a numerical approximation.
 
-In detail, a tangent-stiffness matrix can be described as a collection of the first-order partial derivatives of a vector valued function with respect to each independent vector component of the function for a given vector.
-A general mathematical formula for a tangent-stiffness matrix, $K_{ij}$,
-can be expressed in indicial notation as 
+A general mathematical formula for a tangent-stiffness matrix, $K_{ij}$, can be expressed in indicial notation as 
 %
@@
 %
-where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector is chosen to linearize about $\vec{X}_0$.
+where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector, $\vec{X}_0$ is chosen as the linearization point.
 One then evaluates the expression for each combination $(i, j)$ corresponding to a (row, column) location in the tangent-stiffness matrix.
-By inspection, the elements of a tangent-stiffness matrix can be estimated using any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others.
+The elements of a tangent-stiffness matrix can be estimated using any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others.
 \\
@@
 The second of the two terms is $F_i^{int}$ evaluated in the current configuration.
-The denominator is the magnitude of perturbation called \emph{probe-distance}.
+The denominator is the magnitude of perturbation called the \emph{probe-distance}.
 
@@
 Together, the FD and CD methods, are sometimes termed \emph{finite-difference probing} techniques for tangent-stiffness calculation in literature \cite{ref-Adaggio}.
+From the well-known 1-dimensional formulas for FD and CD, it is understood that the accuracy of these expressions is dependent theoretically upon selecting a small step-size $h$, however experiments shown in \cite{squire1998using} demonstrate that too small an $h$ can lead to inaccuracy from ``subtractive-cancellation''.
+Therefore it is fair to say that a general method for selecting $h$ in practice is unobtainable.
 
@@
 The small perturbation, $h$, is now carried out on the imaginary axis and only the real coefficient to the imaginary part of the vector returned by $F^{int}$ is kept, hence the $\mbox{Im}(\cdot)$ operation.
-As highlighted in Section~\ref{sec:CSmethod}, notice there is no subtraction operation taking place; therefore $h$ can be made very small to yield highly accurate derivatives.
+As highlighted in Section~\ref{sec:CSmethod}, notice that there is no subtraction operation taking place; therefore $h$ can be made very small to yield highly accurate derivatives.
+It should be mentioned that using complex-step to compute tangent-stiffness matrices was done in \cite{perez2000numerical,perez2012numerical}, but those references did not investigate the performance of complex-step in a massively parallel setting nor make comparisons to AD.
 
@@
 %
-\subsection{Peridigm and  peridynamics.}
+\subsection{\emph{Peridigm} and  peridynamics.}
 %
@@
 \emph{Peridigm} is distributed by Sandia National Laboratories as its primary open-source computational peridynamics code.
-It is a massively-parallel simulation code for implicit and explicit multi-physics simulations primarily used for solid mechanics and material failure.
+It is a massively parallel simulation code for implicit and explicit multi-physics simulations primarily used for solid mechanics and material failure.
 \emph{Peridigm} is a C++ code utilizing agile software components from Sandia's  \emph{Trilinos} project \cite{trilinos}.
 It's important to note that there is no specialization of the tangent-stiffness calculation methods described previously to this particular code or more generally to a computational peridynamics approach and the analysis carried out in this study should provide insight into the speed and accuracy of these methods when implemented into other computational mechanics analysis tools as well.
-\emph{Peridigm} was chosen because of the author's relationship with the code development project and primarily because the FD, CD, and AD methods for tangent-stiffness calculation where already implemented in the code, leaving only the CS method to for development.
-However, because peridynamics is a nonlocal theory, the tangent-stiffness matrices have a much higher bandwidth (i.e.
+\emph{Peridigm} was chosen primarily because the FD, CD, and AD methods for tangent-stiffness calculation where already implemented in the code, leaving only the CS method for development.
+Also, because peridynamics is a nonlocal theory, the tangent-stiffness matrices have a much higher bandwidth (i.e.
 less sparsity) than traditional finite-element tangent-stiffness computations and the time required to construct the tangent-stiffness is a majority of the total computation time in any quasi-static or implicit dynamics analysis; therefore, the \emph{Peridigm} development team has an interest in profiling the performance of the tangent-stiffness computation methods.
 
-Briefly, peridynamics \cite{silling2000ret, silling:psa, silling2010peridynamic} is a nonlocal reformulation of the partial-differential equations that provide the statement of momentum balance in classical continuum mechanics.
-Its primary goal is the avoid the use of spatial derivatives in the balance equations or constitutive laws such that discontinuous displacements, i.e.
+Briefly, peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic} is a nonlocal reformulation of the partial-differential equations that provide the statement of momentum balance in classical continuum mechanics.
+Its primary goal is to avoid the use of spatial derivatives in the balance equations or constitutive laws such that discontinuous displacements, i.e.
 cracks, are mathematically consistent with the governing equations.
-It has demonstrated great promise in modeling problems with massive pervasive failure \cite{littlewood2010}, interesting phenomenon such as dynamic crack branching \cite{ha2010sod}, and is being extended to multiphyics phenomenon \cite{bobaru2011peridynamic,katiyar2013} and coupling to molecular simulations \cite{seleson2009peridynamics}.
+It has demonstrated great promise in modeling problems with pervasive failure \cite{littlewood2010}, interesting phenomenon such as dynamic crack branching \cite{ha2010sod}, and is being extended to multiphyics phenomena \cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale coupling to molecular simulations \cite{seleson2009peridynamics}.
 
@@
 This is because \emph{Epetra} vectors, which can be thought of as distributed memory parallel cousins to standard C++ STL vectors, are hard coded to be of type {\tt double} and can not be simply be declared as having a complex data type.
-Therefore, program methods that were used in the course of applying the FD method were copied, renamed and re-written to follow the CS formula, which involved changing the data type of intermediate variables which were locally scoped and dynamically casting persistent memory variables to complex data types where necessary.
-We made no attempt to ``tune'' the CS code in any way for performance, we attempted to implement the algorithm is such as way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurt its overall performance and this issue could be alleviated by using the next generation templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any data type including complex.
-There where additional modifications for the purpose instrumenting the code to collect comparative data and are not specifically part of CS or other methods.
-The source code patch to \emph{Peridigm} for implementing CS can be found on the corresponding author's website.
+Therefore, program methods that were used in the course of applying the FD method were copied, renamed and re-written to follow the CS formula.
+This involved changing the data type of intermediate variables which were locally scoped, and dynamically casting persistent memory variables to complex data types where necessary.
+No attempt was made to ``tune'' the CS code in any way for performance.
+The algorithm  was implemented in such as way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurts its overall performance and this issue could be alleviated by using the next generation of templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any data type including {\tt complex}.
+The version of \emph{Peridigm} modified to implement CS can be found in the paper repository.
 
-\emph{Peridigm} was primarily written by expert computational scientists; therefore, many advanced techniques in C++ such as multiple virtual inheritance, pointer arithmetic, and distributed data structures were used in the basic code and by necessity in the modified code.
+\emph{Peridigm} was primarily written by computational scientists; therefore, many advanced techniques in C++ such as multiple virtual inheritance, pointer arithmetic, and distributed data structures were used in the basic code and by necessity in the modified code.
 To see CS and other methods in a more narrow context, it is advised that readers also take a look at the simple examples provided on the corresponding author's website.
@@
 The methods were compared by running two sets of test problems where each method would solve a problem concurrently.
+In this context, concurrently means successively, drawing from the same independent variables, yet within the same computational process; this definition is characterized in further detail in Section~\ref{JGAM}.
 Each test problem simulated a $\SI{4}{\meter}$ by $\SI{0.5}{\meter}$ by $\SI{0.5}{\meter}$ meter block of a material undergoing tension along the axis parallel to the long dimension of the block.
-The test problems were set up in \emph{Perdigm} as quasistatic equilibrium problems, where the displacement boundary conditions used to apply tension were applied gradually, in `load steps', and an equilibrium solution for each load step was achieved before applying the next load step.
+The test problems were set up in \emph{Perdigm} as quasistatic equilibrium problems, where the displacement boundary conditions used to apply tension were applied gradually in ``load steps'', and an equilibrium solution for each load step was achieved before applying the next load step.
 The purpose of applying the load in gradual steps is a widely used technique to ensure the Newton's method is always initialized near the actual solution and therefore likely to converge.
@@
 A series of nine single compute core (i.e.
-serial) test problems were run with the discretization level being refined with every test in the series.
-The aim of increasing the refinement was to see if possible differences between the methods in terms of accuracy and speed would arise.
+serial) test problems were run with the discretization size being refined with every test in the series.
+The aim of increasing the refinement was to examine differences between the methods in terms of accuracy and speed.
 This series would comprise the single core runs.
-The specific parameters of these tests can be found in \emph{Peridigm} input {\tt xml} files included in the archive for this paper which can be found in the corresponding author's website.
-These {\tt xml} files allow users with appropriately modified versions of \emph{Peridigm} to reproduce the results found in this paper.
+The specific parameters of these tests can be found in \emph{Peridigm} input {\tt xml} files included in the paper repository.
+These {\tt xml} files allow users with the appropriately modified version of \emph{Peridigm} to reproduce the results found in this paper.
 
@@
 This allowed for sufficient computational nodes per core even at the highest level of parallelization such that message passing computation did not overwhelm the simulation.
-The specific parameters of these tests can be found in \emph{Perdigm} input {\tt xml} files also included in the archive for this paper on the corresponding author's website.
+The specific parameters of these tests can be found in \emph{Perdigm} input {\tt xml} files also included in the paper repository.
 
@@
 Additionally, having each of the methods physically operate in the same process, serial or one of associated parallel, allowed the comparison of tangent-stiffness matrix calculation time as cached from random access memory rather than from the hard-disk.
-Caching from RAM gives the benefit of vastly greater speed and the simpler programming compared to some other solution speculatively involving dynamic file management on files which for one of the components of the study would be on the order of 10 terabytes in size.
+Caching from RAM gives the benefit of vastly greater speed and simpler programming compared to some other solution speculatively involving dynamic file management on files which for one of the components of the study would be on the order of a terabyte in size.
 However, the price of running the methods together in the same process and avoiding the use of the hard-disk was that at least two tangent-stiffness matrices had to be stored in RAM during the simulation, which meant that special high-memory compute nodes were needed for the 1 million peridynamic node test series.
 
-The other goal was to compare the four methods, that is including AD, on the basis of speed.
+The other goal was to compare the four methods, including AD, on the basis of speed.
 Speed is defined as the total compute-time of one iteration of a Newton step which includes the tangent-stiffness calculation and parallel assembly.
@@
 The purpose of averaging calculation time measurements was to informally address the extraneous variables of parallel evaluation order and computer system load due to system processes not associated with the study.
-While not part of the study, the example programs available on the corresponding author's website allow the reader to make a comparison of the methods for themselves on the basis of accuracy and convergence rate for two example nonlinear systems.
-In these examples, one will notice that the finite difference methods will lose quadratic convergence when the step size selected too large such that accuracy relative to AD is decreased.
+While not part of the study, the example programs available in the paper repository allow the reader to make a comparison of the methods for themselves on the basis of accuracy and convergence rate for two example nonlinear systems.
+In these examples, one will notice that the finite difference methods will lose quadratic convergence when the step size selected is too large such that accuracy relative to AD is decreased.
 
-Once the quantities-of-interest (QoIs) for comparison were determined, the task was to find a way to take measurements that satisfied our goals.
 The goals of collecting accuracy and speed measurements were achieved by developing and implementing metrics within the simulation program used in the study.
 The metric used to measure the accuracy of a tangent-stiffness matrix was the Frobenius norm (analogous to an $l^2$-norm except defined on a matrix) of the element-wise difference between the tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix produced by the AD based method (i.e., the \emph{exact} derivatives), given that both methods were set upon the same problem.
-The lower the value of this metric, the closer the tangent-stiffness of the method being evaluated (i.e, CS, AD, FD)  was to the AD Jacobian, the more accurate the method was.
-The expression for the accuracy metric is
+The lower the value of this metric, the more accurate the method.
+The expression for the accuracy metric was
 %
@@
   D = \sqrt{\sum_i \sum_j(K^{AD}_{ij} - K^{M}_{ij})^2}
+  \label{eqn:accuracy}
 \end{equation}
@@
 The metric used to compare the speed of the different methods was the time in seconds required to calculate the tangent-stiffness matrix.
-A final basis of comparison used in the study called \emph{computational efficiency} is based on the ratio of calculation time to the total number of tangent-stiffness matrix elements.
+A final basis of comparison used in the study called \emph{computational efficiency} is based on specific calculation time per tangent-stiffness matrix element.
 
@@
 Discretization for the test problems run in the study was done using the software package \emph{Cubit} developed at Sandia National Laboratories \cite{ref-Cubit}.
-\emph{Cubit} generates a finite element mesh that is internally converted to a ``particle'' discretization internally when calling \emph{Peridigm}, its important to node that the number of computational nodes in \emph{Peridigm} correspond to the number of finite elements in the discretization, not the number of finite element nodes.
+\emph{Cubit} generates a finite element mesh that is internally converted to a ``particle'' discretization internally when calling \emph{Peridigm}.
+The number of computational nodes in \emph{Peridigm} correspond to the number of finite elements in the discretization, not the number of finite element nodes.
 An example journal script can be found on the corresponding author's website.
 
-The test problems were run the \emph{Stampede} HPC cluster computer housed at \emph{TACC} (Texas Advanced Computing Center) at The University of Texas at Austin.
+The test problems were run on the \emph{Stampede} HPC cluster computer housed at \emph{TACC} (Texas Advanced Computing Center) at The University of Texas at Austin.
 The single core test series was run using the {\tt normal} queue compute nodes, while the parallel test series was run using one to four {\tt large memory} nodes each having one terabyte of local RAM.
@@
 Directories and filenames were chosen so that data would be easy to sort by individual test run.
-Extraneous information was manually deleted from copied versions of the output text files so that they would be regular enough to be easily read by scripts written in \emph{Python}.
+\emph{Python} scripts were then used to post-process and plot the data.
 These scripts averaged accuracy or speed data for a tests corresponding to a particular peridynamic node density over all iterations for that run, and then plotted this averaged data  as a function of peridynamic node density as it varied from test to test in the series.
-Similarly, for the parallel tests, the data was averaged and plotted in much the same way, but as a function of number of compute cores used to solve the problem rather than as a function peridynamic node density which was held constant.
+Similarly, for the parallel tests, the data was averaged and plotted in much the same way, but as a function of number of compute cores used to solve the problem.
 Confidence intervals were not calculated on the measurements.
@@
 %
-This section contains the average results from the study presented in tabular and graphic format.
 Table~\ref{tab:results} shows the compute core count, number of nonzero tangent-stiffness (TS) non-zero matrix elements, load step average calculation time and load step average accuracy.
-The units for accuracy are derived from the
-expression used to calculate the TS matrix elements and the units used to define bulk material properties, mentioned in Section~\ref{tcs}.
+The units for accuracy are derived equation~(\ref{eqn:accuracy}) and the units used to define bulk material properties, mentioned in Section~\ref{tcs}.
 %
-\note[JTF]{Michael, work on cleaning this table up for legibility}
-\note[Michael]{Columns are now separated out and can be fine tuned}
-
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-\begin{table}    
+\begin{table*}[!ht]    
+  \centering
         \caption{Averaged Results, Each Test} \label{tab:results}   
-       \hskip-2.0cm\begin{tabular}{C{1.5cm} C{1.88cm} C{1.5cm} C{1.5cm} C{1.5cm} C{1.5cm} C{1.88cm} C{1.88cm} C{1.5cm}}
-            Cores & TS Elements & Calc. Time (s) & & & & Accuracy Diff. ($\si{\mega\pascal}$) & & \\ 
-        \multicolumn{2}{l}{} & CS & CD & FD & AD & CS & CD & FD \\ \hline 
-        1 & $\num{1.99E6}$  & 3.5 & 2.7 & 1.7 & 1.6 & $\num{1.92E-10}$ & $\num{1.21E-4}$ & .137 \\ \hline
-        1 & $\num{4.25E6}$  & 6.2& 4.9& 3.2& 2.9 & $\num{2.28E-10}$& $\num{9.94E-4}$ & .148 \\ \hline
-        1 & $\num{7.79E6}$  & 11.2& 8.9& 5.7& 5.2 & $\num{2.38E-10}$& $\num{1.59E-4}$ & .145\\ \hline
-        1 & $\num{1.54E7}$  & 26.7& 21.0& 13.4& 12.2 & $\num{2.33E-10}$ & $\num{4.61E-4}$ & .12 \\ \hline
-        1 & $\num{2.00E7}$  & 28.1& 22.2& 14.4& 13.1 & $\num{2.76E-10}$ & $\num{1.05E-3}$ & .145 \\ \hline
-        1 & $\num{3.14E7}$  & 47.6& 37.7& 24.2& 21.9 & $\num{2.64E-10}$ & $\num{1.65E-3}$ & .133 \\ \hline
-        1 & $\num{4.01E7}$  & 55.6& 44.1& 28.4& 25.9 & $\num{3.03E-10}$ & $\num{1.92E-3}$ & .148 \\ \hline
-        1 & $\num{8.39E7}$  & 138.9& 109.6& 70.2& 64.0 & $\num{3.63E-10}$ & $\num{1.64E-3}$ & .123 \\ \hline
-        1 & $\num{1.65E8}$  & 277.3& 218.1& 139.4& 126.5 & $\num{3.26E-10}$ & $\num{2.18E-3}$ & .128 \\ \hline
-        32 & $\num{1.67E10}$  & 336.1& 277.7& 200.6& 233.1 & $\num{6.21E-10}$ & $\num{1.52E-2}$ & .176 \\ \hline
-        64 & $\num{1.67E10}$  & 169.9& 140.7& 102.0& 119.7 & $\num{6.20E-10}$ & $\num{1.50E-2}$ & .177 \\ \hline
-        96 & $\num{1.67E10}$  & 114.7& 95.0& 69.1 & 79.7 & $\num{6.18E-10}$ & $\num{1.50E-2}$ & .177 \\ \hline
-        128 & $\num{1.67E10}$  & 86.4& 71.8& 52.4 &58.8 & $\num{6.16E-10}$ & $\num{1.47E-2}$ & .177 \\ \hline
+       \begin{tabular}{c c c c c c c c c}
+         \toprule
+         Cores & TS Elements & \multicolumn{4}{c}{Calculation Time ($\si{\second}$)} & \multicolumn{3}{c}{Accuracy Difference ($\si{\mega\pascal}$)} \\ 
+         & $\times 10^6$ & CS & CD & FD & AD & CS & CD & FD \\
+        \midrule
+        1 & $\num{1.99}$  & 3.5 & 2.7 & 1.7 & 1.6 & $\num{1.92E-10}$ & $\num{1.21E-4}$ & .137 \\
+        1 & $\num{4.25}$  & 6.2& 4.9& 3.2& 2.9 & $\num{2.28E-10}$& $\num{9.94E-4}$ & .148 \\
+        1 & $\num{7.79}$  & 11.2& 8.9& 5.7& 5.2 & $\num{2.38E-10}$& $\num{1.59E-4}$ & .145\\
+        1 & $\num{15.4}$  & 26.7& 21.0& 13.4& 12.2 & $\num{2.33E-10}$ & $\num{4.61E-4}$ & .12 \\
+        1 & $\num{20.0}$  & 28.1& 22.2& 14.4& 13.1 & $\num{2.76E-10}$ & $\num{1.05E-3}$ & .145 \\
+        1 & $\num{31.4}$  & 47.6& 37.7& 24.2& 21.9 & $\num{2.64E-10}$ & $\num{1.65E-3}$ & .133 \\
+        1 & $\num{40.1}$  & 55.6& 44.1& 28.4& 25.9 & $\num{3.03E-10}$ & $\num{1.92E-3}$ & .148 \\
+        1 & $\num{83.9}$  & 138.9& 109.6& 70.2& 64.0 & $\num{3.63E-10}$ & $\num{1.64E-3}$ & .123 \\
+        1 & $\num{165.0}$  & 277.3& 218.1& 139.4& 126.5 & $\num{3.26E-10}$ & $\num{2.18E-3}$ & .128 \\
+        32 & $\num{1670}$  & 336.1& 277.7& 200.6& 233.1 & $\num{6.21E-10}$ & $\num{1.52E-2}$ & .176 \\
+        64 & $\num{1670}$  & 169.9& 140.7& 102.0& 119.7 & $\num{6.20E-10}$ & $\num{1.50E-2}$ & .177 \\
+        96 & $\num{1670}$  & 114.7& 95.0& 69.1 & 79.7 & $\num{6.18E-10}$ & $\num{1.50E-2}$ & .177 \\
+        128 & $\num{1670}$  & 86.4& 71.8& 52.4 &58.8 & $\num{6.16E-10}$ & $\num{1.47E-2}$ & .177 \\
+        \bottomrule
     \end{tabular}
-\end{table}
+\end{table*}
 
@@
 The averaged results of these speed measurements for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
-Average time in seconds to compute a TS element is plotted as a function of the number of nonzero TS elements.
-CS turned out to be the slowest of the four methods, taking the longest average time per iteration,  followed by CD, FD and AD in that order.
-Calculation time appears to have a linear relationship with problem size, judging by the appearance of the plot shown in Fig.~\ref{fig:serial_speed}.
+Average time in seconds to compute a TS matrix is plotted as a function of the number of nonzero TS elements.
+CS was the slowest of the four methods, taking the longest average time per iteration,  followed by CD, FD and AD in that order.
+Calculation time has a power-law relationship with problem size, with a power-law index of roughly $\num{1.0E-6}$ for each of the methods.
 This bodes well for the scalability of each of the methods by themselves in \emph{Peridigm}.
@@
 For the parallel simulations, averaged speed results appear in Fig.~\ref{fig:multi_speed}.
-For this series, rather than varying the number of nonzero TS matrix elements the number of compute cores used to solve the problem is increased, while the number of nonzero TS matrix elements remained constant.
-CS again was the slowest method, followed by central-difference, automatic-differentiation and then forward-difference.
+For this series, rather than varying the number of nonzero TS matrix elements, the number of compute cores used to solve the problem is increased, while the number of nonzero TS matrix elements remained constant at $\num{1.67E9}$ .
+CS again was the slowest method, followed by CD, AD and then FD.
 There is a somewhat surprise reversal of FD and AD for these parallel simulations.
-Figure~\ref{fig:multi_speed} indicates that when looking at average iteration time, each of the methods scale linearly with the number of compute cores used for the given conditions of this group of tests.
+Figure~\ref{fig:multi_speed} indicates that when looking at average iteration time, the factor of speed improvement scales linearly with the number of compute cores used for the given conditions of this group of tests.
+Looking at Fig.~\ref{fig:multi_speed}, the reader can see that when core count doubles, compute time halves, for each method.
 Again, these results show that each of the methods as implemented in \emph{Peridigm} scale well when additional processors are used.
@@
 A speculative explanation as to why FD and AD have reversed places in the speed tests with respect to the serial results is that the AD data-structures carry a list of partial derivatives of the functional evaluations along side the functional evaluations themselves only to be processed at the end of the tangent-stiffness calculation when the derivatives are requested.
-It is likely that the communication cost associated with message passing these extra terms is what causes the AD method to be slower that FD.
+This represents additional load on the hardware memory subsystem, such that may exceed that of FD yet still fall under a supposed memory bandwidth threshold during serial operation.
+Cores within a multicore processor use the same shared bus to access memory, such that cores compete for memory access.
+Supposing that exceeding the memory bandwidth threshold penalizes additional requested bandwidth, it is likely that the decrease in available memory bandwidth per core caused by parallel operation coupled with AD's additional resource requirements causes the AD method to be slower than FD.
 
 \subsection{Accuracy Data}
-Accuracy measurements where taken according to the methods mentioned in Section~\ref{tcs}.
+Data for accuracy measurements were taken according to the methods mentioned in Section~\ref{tcs}, and reduced using equation~(\ref{eqn:accuracy}).
 The averaged results of these accuracy measurements for the serial tests can be seen in Fig.~\ref{fig:serial_accuracy}.
 Averaged Frobenius norm (referred to as $l^2$-norm in the figure labels for conciseness) of the element-wise difference between a given TS matrix and the TS matrix produced by the AD method is plotted as a function of the number of nonzero TS matrix elements.
-CS was shown to be the most accurate of the methods when compared to AD, CS resulting in several orders of magnitude less difference than resulted from using the CD method and as expected FD method was the least accurate.
+CS was shown to be the most accurate of the methods when compared to AD and was roughly \emph{$6$ orders of magnitude more accurate than CD} and \emph{$9$ orders of magnitude more accurate than FD}.
 This accuracy ranking order was maintained for each test in the series.
-It should be pointed out compared to the $l^2$-norms of any of the TS matrices by themselves, the magnitude of these differences for any of the methods is relatively small.
-These values appear in the simulation data which is available on the corresponding author's website.
+It should be pointed out that compared to the $l^2$-norms of any of the TS matrices by themselves, the magnitude of the accuracy metric calculated with Eqn.~\ref{eqn:accuracy} for any of the methods is relatively small.
+These values for the $l^2$-norms of the TS matrices appear in the simulation data which is available in the paper repository.
 The effect of these accuracy differences on convergence rate of the Newton solver was not determined.
-Additionally, we must point out that the accuracy of FD and CD could be improved by using a smaller probe distance, $h$; however, if $h$ is too small, then the methods can suffer from severe effects, including solution divergence, of the round-off error associated with subtracting two numbers that are very close to one another.
-The value of $h$ used in this study is default value used in \emph{Peridigm} which is heuristically derived based on computational node spacing.
-One of the beauties of the CS method is that there is no mystery as how to choose, $h$, it cannot be too small, therefore it can be set to machine epsilon and the analysis performed without the worry of issues associated with round-off error.
+%
+The accuracy of FD and CD could be improved by using a smaller probe distance, $h$; however, if $h$ is too small, then the methods can suffer from severe effects, including solution divergence, of the round-off error associated with subtracting two numbers that are very close to one another.
+The value of $h$ used in this study is the default value used in \emph{Peridigm} which is heuristically derived based on computational node spacing.
+One of the beauties of the CS method is that since dependence on $h$ decreases for small $h$, which a numerical example in \cite[Table 1]{squire1998using} demonstrates, $h$, cannot be too small, therefore it can be set to machine epsilon and the analysis performed without the issues associated with round-off error.
+The value chosen for $h$ for CS used in this study was $\num{1.E-100}$ to demonstrate this.
 %
@@
 The averaged results of the accuracy measurements taken for the parallel test series can be seen in Fig.~\ref{fig:multi_accuracy}.
-The ordinate axis units remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the absicca now indicates number of compute cores used to solve the test problem.
+The ordinate axis units remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the abscissa now indicates number of compute cores used to solve the test problem.
 Accuracy trends matched those for the single core tests.
@@
 \subsection{Efficiency Data}
-A derived measurement was also used to compare the methods.
+Efficiency, calculated as the average number of seconds taken per nonzero Jacobian matrix element, was plotted as a function of the number nonzero TS matrix elements.
 The single core results are shown in Fig.~\ref{fig:serial_efficiency}.
-Efficiency, calculated as the average number of seconds taken per nonzero Jacobian matrix element was plotted as a function of the number nonzero TS matrix elements.
 The ranking order of the methods
-matched their speed ranking in Fig.~\ref{fig:serial_speed}.
-This plot indicates that there is a small, yet present decrease in efficiency with respect to the number of nonzero TS elements.
-This trend in efficiency doesn't appear to have great effect in practical terms for the discretization used in this test series.
+matched their speed ranking in Fig.~\ref{fig:serial_speed}. This plot indicates that efficiency is insensitive to the the number of nonzero TS elements.  
 %
@@
 The parallel efficiency plot shows that AD was slower than FD because for the given test problem it had a lower efficiency.
-Strangely AD appears to gain in efficiency over the last few data points in the plot, looking at the raw data confirmed that this was what was measured rather than this being an error during data reduction.
+Strangely AD appears to gain in efficiency over the last few data points in the plot.
 The reason behind this trend was not determined.
@@
 Based on the experimental results, it appears that for users of \emph{Peridigm} there is a case to be made for using AD for reasons of performance and accuracy.
-Although, for certain parallel tests done in this study small performance gains over AD where shown with the FD method.
+However, certain parallel tests done in this study showed a small performance gain over AD with the FD method.
 While the results showed that CS produced more accurate tangent-stiffness matrices than CD and FD under the parameters of the tests, it was not determined whether or not this is a clear advantage of CS over FD in terms of accuracy of final predicted displacements and speed of convergence.
-For other simulation software it is obviously necessary to collect the same types measurements as described in this paper in order to make a design decision.
+The reader is reminded that the time required to complete an iteration must be considered along with the total number of iterations taken order to measure speed of convergence.
+The goal of the comparative study was to benchmark single iteration performance and not convergence rate.
 For new developments it is recommended that combined accuracy and speed of iteration measurements be conducted along with entirely separate convergence studies.
+It should be mentioned that a related work, \cite{perez2012numerical}, compares the speed and accuracy of complex-step to finite difference for producing tangent-stiffness matrices in solid mechanics simulations using complex inelastic material model, where the researchers similarly found CS to be accurate yet relatively computationally expensive compared to FD and CD.
+In that work, full convergence studies were also done and indicated that quadratic convergence could be achieved with complex-step all of the methods \cite[p.28]{perez2012numerical}.
 
@@
 This is fine for copying in between MPI processes, but deserialization on an accelerator or GPU is very difficult, as this would require compiling a version of the AD library being used that were compatible with the accelerator or GPU being used.
-What this means is that if \emph{Peridigm} or any simulation software were to be reformulated to take advantage of accelerators or GPUs, attractive given their performance and power efficiency advantages over traditional CPUs, the flexibility of complex-step owed to its simple implementation may make it more viable for that application given its high accuracy.
-Additional performance gains may be achieved through the use of the templated \emph{Tpetra} package in \emph{Trilinos} which would allow the use of complex data types natively, and alieviate the expense of the dynamic-casts that were utilized in this study.
-As a final comment, it should be noted that from a code development, or implementation, perspective the CS method may be the easiest of the methods to implement.
-It only requires one functional evaluation, there is no heuristic choice of probe distance, and it there is no dependence on an external AD library.
-From, this perspective, the method may still be an attractive method for TS matrix evaluation.
-
-%% The Appendices part is started with the command \appendix;
-%% appendix sections are then done as normal sections
+What this means is that if \emph{Peridigm} or any simulation software were to be reformulated to take advantage of accelerators or GPUs, the flexibility of complex-step owed to its simple implementation may make it more viable for that application given its high accuracy.
+Additional performance gains may be achieved through the use of the templated \emph{Tpetra} package in \emph{Trilinos} which would allow the use of complex data types natively, and alleviate the expense of the dynamic-casts that were utilized in this study.
+As a final comment, it should be noted that from a code development perspective the CS method may be the easiest of the methods to implement; It only requires one functional evaluation, there is no heuristic choice of probe distance, and there is no dependence on an external AD library.
+From this perspective, the method may still be an attractive method for TS matrix evaluation.
 
@@
 \label{sec:ack}
-This work was funded in part by grants from the United States Air Force Office of Scientific Research grant number XXXX \note[JTF]{Harry: Your grant number here} and National Energy Technology Laboratory grant number XXXX.
+This work was funded in part by grants from the United States Air Force Office of Scientific Research grant number W911NF-11-1-0208 and National Energy Technology Laboratory grant number DE-FE0010808.
 The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.
@@
 
-%% Authors are advised to submit their bibtex database files. They are
-%% requested to list a bibtex style file in the manuscript if they do
-%% not want to use elsarticle-num.bst.
 \appendix
@@
 \item
-%AD doesn't return symbolic epressions for derivitives that can be re-evaluated.
+%AD doesn't return symbolic expressions for derivatives that can be re-evaluated.
 %Instead, generic versions of functions and evaluations of their partials are defined together.
@@
 functions $u,v,w$ and their partial derivatives w.r.t $s$ are defined such that: \\
-\begin{tabular}{l | l}
-\multicolumn{2}{c}{for $a, b, s$ equal to $a_n, b_n, s_n$:} \\ \hline
-$u \mid _{a_n, b_n, s_n} = a_n \cdot s_n^{b_n}$ & $\frac{\partial{u}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot s_n^{b_n - 1}$ \\
-$v \mid _{a_n, b_n, s_n} = a_n \cdot sin(b_n \cdot s_n)$ & $\frac{\partial{v}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot cos(b_n \cdot s_n)$\\ 
-$w \mid _{a_n, b_n, s_n} = a_n \cdot cos(b_n \cdot s_n)$ & $\frac{\partial{w}}{\partial{s}} \mid _{a_n, b_n, s_n} = -a_n \cdot b_n \cdot sin(b_n \cdot s_n)$\\
+\begin{tabular}{l}
+%\multicolumn{2}{c}{for $a, b, s$ equal to $a_n, b_n, s_n$:} \\ \hline
+for $a, b, s$ equal to $a_n, b_n, s_n$: \\ \hline
+$u \mid _{a_n, b_n, s_n} = a_n \cdot s_n^{b_n}$ \\
+$\frac{\partial{u}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot s_n^{b_n - 1}$ \\
+\\
+$v \mid _{a_n, b_n, s_n} = a_n \cdot sin(b_n \cdot s_n)$ \\
+$\frac{\partial{v}}{\partial{s}} \mid _{a_n, b_n, s_n} = a_n \cdot b_n \cdot cos(b_n \cdot s_n)$ \\ 
+\\
+$w \mid _{a_n, b_n, s_n} = a_n \cdot cos(b_n \cdot s_n)$ \\
+$\frac{\partial{w}}{\partial{s}} \mid _{a_n, b_n, s_n} = -a_n \cdot b_n \cdot sin(b_n \cdot s_n)$\\
 \end{tabular} 
@@
 \label{differentiation}
-%AD is eqivalent to the chain rule, when the computer has definitions for functions and partials
-%that can be adapted to match the compostition function.
+%AD is equivalent to the chain rule, when the computer has definitions for functions and partials
+%that can be adapted to match the composition function.
 It follows from \ref{composition} that we can evaluate $\frac{d}{d x}f(x) \mid_{x_0}$ with the chain rule: \\ \\
@@
 according to \ref{abcvalues}, with parameters $a, b$ chosen for each function and held as constant, and
-evaluating, such that the total derivative our original function w.r.t $x$ where $x = x_0$ is approximated
+evaluating, such that the total derivative of our original function w.r.t $x$ where $x = x_0$ is approximated
 by the partial derivative of our equivalent statement, $f(x) = u(a, b, s) \mid_{arguments}$,  w.r.t $s$ 
@@
 $\frac{\partial}{\partial x}f(x) \mid _{x_0} = \frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot \frac{ds}{dx}$\\ \\
-Repeating above in tabular format: \\ \\
+%Repeating above in tabular format: \\ \\
 %This shows the same information but is ordered to show progression, like chapters
-\begin{tabular}{l l l}
-Current Evaluation & $s$ Value & Partial Derivative \\ \hline
-$w(a, b, s)\mid_{1, 1, s}$ & $x_0$ & $\frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
-$v(a, b, s)\mid_{1, 1, s}$ & $w(a, b, s)\mid_{1, 1, x_0}$ & $\frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
-$u(a, b, s)\mid_{1, 2, s}$ & $v(a, b, s)\mid_{1, 1, w(a, b, s)\mid_{1, 1, x_0}}$ & $\frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot 1$ \\
-\end{tabular} \\ \\
+%\begin{tabular}{l l l}
+%Current Evaluation & $s$ Value & Partial Derivative \\ \hline
+%$w(a, b, s)\mid_{1, 1, s}$ & $x_0$ & $\frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
+%$v(a, b, s)\mid_{1, 1, s}$ & $w(a, b, s)\mid_{1, 1, x_0}$ & $\frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0}$ \\
+%$u(a, b, s)\mid_{1, 2, s}$ & $v(a, b, s)\mid_{1, 1, w(a, b, s)\mid_{1, 1, x_0}}$ & $\frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot 1$ \\
+%\end{tabular} \\ \\
 %This explains what the equation and table above mean.
@@
 \end{document}
-%%
-%% End of file `elsarticle-template-num.tex'.

### 2013-11-25 data/multicore/plotter.py

@@ -7,9 +7,12 @@ import matplotlib
 matplotlib.use('pgf')
 import matplotlib.pyplot as plt
 
+font_size = 8
+fig_width = 3.0
+
 font = {'family':'serif',
 	'serif':['Times'],
-	'size':12}
+	'size':font_size}
 
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif')
@@ -23,7 +26,7 @@ accuracy_per_mat_element = {}
 
 marker = ""
 c =""
-fig1= plt.figure(1, figsize=(4,3.7))
+fig1= plt.figure(1, figsize=(fig_width*3.5/3.0,fig_width))
 ax1 = fig1.add_subplot(111)
 
 print speed_data.master_data_map
@@ -50,42 +53,42 @@ for index, key in enumerate(accuracy_data.master_data_map.keys()):
 
 print speed_averages
 print accuracy_averages
-cs = ax1.scatter(speed_averages['CS'][0][:], speed_averages['CS'][1][:], marker='^', c='g', s=64)
-cd = ax1.scatter(speed_averages['CD'][0][:], speed_averages['CD'][1][:], marker='x', c='r', s=64 )
-fd = ax1.scatter(speed_averages['FD'][0][:], speed_averages['FD'][1][:], marker='s', c='b', s=64) 
-ad = ax1.scatter(speed_averages['AD'][0][:], speed_averages['AD'][1][:], marker='o', c='y', s=64)
+cs = ax1.scatter(speed_averages['CS'][0][:], speed_averages['CS'][1][:], marker='^', c='g', s=24)
+cd = ax1.scatter(speed_averages['CD'][0][:], speed_averages['CD'][1][:], marker='x', c='r', s=24 )
+fd = ax1.scatter(speed_averages['FD'][0][:], speed_averages['FD'][1][:], marker='s', c='b', s=24) 
+ad = ax1.scatter(speed_averages['AD'][0][:], speed_averages['AD'][1][:], marker='o', c='y', s=24)
 #ax1.set_yscale('log')
 plt.xlim(24.0, 160.0)
 plt.ylim(0.0, 350.0)
-plt.xlabel("Number of cores @ 1TB RAM per 32 cores")
-plt.ylabel("Avg. time to compute TS (s)")
+plt.xlabel("Number of cores at 1TB RAM per 32 cores", fontsize=font_size)
[…]

### 2013-11-25 data/multicore/relative_ploter.py

@@ -7,9 +7,12 @@ import matplotlib
 matplotlib.use('pgf')
 import matplotlib.pyplot as plt
 
+font_size = 8
+fig_width = 3.0
+
 font = {'family':'serif',
 	'serif':['Times'],
-	'size':12}
+	'size':font_size }
 
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif')
@@ -69,10 +72,10 @@ for index, key in enumerate(accuracy_data.master_data_map.keys()):
 		accuracy_averages[subkey][1].append(np.average(np.copy(subdata)))
 		accuracy_averages[subkey][0].append(float(key.split("_")[1]))
 """
-cs = ax1.scatter(speed_averages['CS'][2][:], speed_per_mat_element['CS'][:], marker='^', c='g', s=64)
-cd = ax1.scatter(speed_averages['CD'][2][:], speed_per_mat_element['CD'][:], marker='x', c='r', s=64 )
-fd = ax1.scatter(speed_averages['FD'][2][:], speed_per_mat_element['FD'][:], marker='s', c='b', s=64) 
-ad = ax1.scatter(speed_averages['AD'][2][:], speed_per_mat_element['AD'][:], marker='o', c='y', s=64)
+cs = ax1.scatter(speed_averages['CS'][2][:], speed_per_mat_element['CS'][:], marker='^', c='g', s=24)
+cd = ax1.scatter(speed_averages['CD'][2][:], speed_per_mat_element['CD'][:], marker='x', c='r', s=24 )
+fd = ax1.scatter(speed_averages['FD'][2][:], speed_per_mat_element['FD'][:], marker='s', c='b', s=24) 
+ad = ax1.scatter(speed_averages['AD'][2][:], speed_per_mat_element['AD'][:], marker='o', c='y', s=24)
 #ax1.set_yscale('log')
 plt.xlim(24.0, 136.0)
 plt.ylim(5.e-10, 1.3e-9)
@@ -86,22 +89,22 @@ plt.grid(True)
 """
 
 
-fig2= plt.figure(2, figsize=(4,3.7))
+fig2= plt.figure(2, figsize=(fig_width*3.5/3.0,fig_width))
 ax2 = fig2.add_subplot(111)
-cs2= ax2.scatter(speed_averages['CS'][2][:], speed_per_nonzero_element['CS'][:], marker='^', c='g', s=64)
[…]

### 2013-11-25 data/singlecore/plotter.py

@@ -8,12 +8,16 @@ matplotlib.use('pgf')
 import matplotlib.pyplot as plt
 #import mpl_toolkits.axisartist as axisartist
 
+font_size = 8
+fig_width = 3.0
+
 font = {'family':'serif',
 	'serif':['Times'],
-	'size':12}
+	'size':font_size}
 
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif')
+plt.rcParams['text.latex.preamble'] = r'\usepackage{siunitx}'
 
 #matplotlib.rc('font', **font)
 
@@ -54,7 +58,7 @@ xmin = 2000.0
 
 marker = ""
 c =""
-fig1= plt.figure(1, figsize=(4,3.7))
+fig1= plt.figure(1, figsize=(fig_width*3.5/3.0,fig_width))
 ax1 = fig1.add_subplot(111)
 
 
@@ -87,16 +91,17 @@ for index, key in enumerate(accuracy_data.master_data_map.keys()):
 
 print "The speed averages are:", speed_averages
 print "The accuracy averages are:", accuracy_averages
-cs = ax1.scatter(speed_averages['CS'][0][:], speed_averages['CS'][1][:], marker='^', c='g', s=64)
-cd = ax1.scatter(speed_averages['CD'][0][:], speed_averages['CD'][1][:], marker='x', c='r', s=64 )
-fd = ax1.scatter(speed_averages['FD'][0][:], speed_averages['FD'][1][:], marker='s', c='b', s=64) 
-ad = ax1.scatter(speed_averages['AD'][0][:], speed_averages['AD'][1][:], marker='o', c='y', s=64)
-plt.xlim(1.e5, 1.e9)
-plt.xlabel("Number of nonzero elements in TS")
-plt.ylabel("Average time to compute TS (s)")
-#plt.title("Average time to compute a Jacobian matrix vs.\nnumber of nonzero Jacobian matrix elements")
+cs = ax1.scatter(speed_averages['CS'][0][:], speed_averages['CS'][1][:], marker='^', c='g', s=24)
+cd = ax1.scatter(speed_averages['CD'][0][:], speed_averages['CD'][1][:], marker='x', c='r', s=24 )
[…]

### 2013-11-25 data/singlecore/relative_ploter.py

@@ -9,9 +9,12 @@ matplotlib.use('pgf')
 import matplotlib.pyplot as plt
 #import mpl_toolkits.axisartist as axisartist
 
+font_size = 8
+fig_width = 3.0
+
 font = {'family':'serif',
 	'serif':['Times'],
-	'size':12}
+	'size':font_size}
 
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif')
@@ -118,24 +121,25 @@ ax1.grid(which='major', linestyle='--', color='grey')
 """
 
 
-fig2= plt.figure(2, figsize=(4,3.7))
+fig2= plt.figure(2, figsize=(fig_width*3.5/3.0,fig_width))
 ax2 = fig2.add_subplot(111, aspect=4)
 
-cs2= ax2.scatter(speed_averages['CS'][0][:], speed_per_nonzero_element['CS'][:], marker='^', c='g', s=64)
-cd2= ax2.scatter(speed_averages['CD'][0][:], speed_per_nonzero_element['CD'][:], marker='x', c='r', s=64 )
-fd2= ax2.scatter(speed_averages['FD'][0][:], speed_per_nonzero_element['FD'][:], marker='s', c='b', s=64) 
-ad2= ax2.scatter(speed_averages['AD'][0][:], speed_per_nonzero_element['AD'][:], marker='o', c='y', s=64) 
+cs2= ax2.scatter(speed_averages['CS'][0][:], speed_per_nonzero_element['CS'][:], marker='^', c='g', s=24)
+cd2= ax2.scatter(speed_averages['CD'][0][:], speed_per_nonzero_element['CD'][:], marker='x', c='r', s=24 )
+fd2= ax2.scatter(speed_averages['FD'][0][:], speed_per_nonzero_element['FD'][:], marker='s', c='b', s=24) 
+ad2= ax2.scatter(speed_averages['AD'][0][:], speed_per_nonzero_element['AD'][:], marker='o', c='y', s=24) 
 plt.xlim(1.e5, 1.e9)
 plt.ylim(0.0, 2.E-6)
-plt.xlabel("Number of nonzero elements in TS")
-plt.ylabel("Avg. comp time of TS per nonzero element")
+plt.xlabel("Number of TS elements", fontsize=font_size)
+plt.ylabel("Compute time of TS per element ($\\times 10^{-6}$)", fontsize=font_size)
[…]

### 2014-02-05 response/response.tex
Added response folder and trackchanges.sty

\documentclass{article}
\usepackage[left=1.0in,right=1.0in,top=1.0in,bottom=1.0in]{geometry}
\usepackage{fancyhdr}
\usepackage{color}
\usepackage{amsmath}
\definecolor{light-gray}{gray}{0.5}

\begin{document}

\begin{center}
    {\LARGE \bf Point-by-Point Response to Reviewer's Comments} \\
     Author's Responses in {\color{red} Red}
\end{center}

%%%% First Reviewer %%%%
\section{Reviewer}
%
\subsection*{General Statement}
In general, I am not convinced of the benefits of particle methods when compared to traditional finite element analysis. Although they eliminate the sometimes troublesome problem of mesh generation, they create new problems of their own. In particular, they are hard to debug (because the patch test often doesn’t work), it is typically difficult to pin down an order of accuracy, and it is typically difficult to apply boundary conditions. Nonetheless, I recognize that they are part of the toolbox and therefore are within the scope of IJNME.
%
{\color{red}
  \begin{itemize}
    \item We have added a paragraph at the end of Section 1 that highlights the difference between peridynamics and previous “particle methods.” The main difference is that in peridynamics, we numerically approximate a different set of continuum equations.
  \end{itemize}}

This particular paper, however, has insufficient novelty to merit publication in its present form. In particular,

\begin{enumerate}
  \item The peridynamic framework, including the method to compute the deformation gradient, is already established in the previous literature.
  {\color{red}
    \begin{itemize}
      \item We have added statements in the abstract and introduction that highlight what is new about the work being reported. Specifically, this represents the first successful implementation of a rate-dependent material model within peridynamics. We also present the first full-scale 3D solution of a realistic, reasonably complex application. This work therefore demonstrates to the community, for the first time, the ability of the peridynamic method to solve practical problems involving complex materials resulting in good agreement with experimental data.
    \end{itemize}}
  %
  \item The viscoplastic formulation and constitive model follows fairly classical nonlinear solid mechanics, and therefore does not appear to be novel. 
  {\color{red}
    \begin{itemize}
      \item Viscoplasticity has never been used within peridynamics.  See previous responses.
    \end{itemize}}
  %
  \item The computational studies are quite limited, by the authors’ own admission (“while failure can be modeled with peridynamics, it is beyond the scope of this paper; therefore, these experiments were not attempted in numerical simulation”, p. 19). 
  {\color{red}
  \begin{itemize}
    \item The focus of this paper was to show that peridynamics could be used with conventional (stress-strain) constitutive models.   Modeling failure is the main advantage of the peridynamic approach, which has been demonstrated in the cited papers and will be further developed in a future paper with the combination of a viscoplastic constitutive model and a novel failure mechanism applied by breaking ``bonds''.
  \end{itemize}}
  %
\end{enumerate}

\subsection*{Specific Comments}

\begin{enumerate}
%
   \item p. 3, It has been proved that Equation 1 reduces to the classical continuum partial differential equation\dots Which one? Momentum?
   {\color{red}
     \begin{itemize}
        \item Yes, linear momentum.  This has been made more clear in the rewritten introduction.
     \end{itemize}}
%
  \item p. 5, should be $F = I + \nabla_x u$ (standard notation)
   {\color{red}
     \begin{itemize}
        \item The notation I used is identical to the notation used by Lawrence E. Malvern in his classic book \underline{Introduction to the Mechanics of a Continuous Medium} (p.\ 162 in the version I have).  This is further clarified by showing the incidental notation in the revision.
     \end{itemize}}
%
  \item Eq. (30) is an additive decomposition of the deviator of the referential tensor $\mathbf{d}$ based on the additive decomposition of the rate of deformation $\mathbf{D}$. A reference and/or some background should be included.
  {\color{red}
     \begin{itemize}
        \item Eq.\ (30) is an additive decomposition of the deviatoric part of $\mathbf{d}$ which is referred to as the {\em unrotated rate of deformation tensor}.  $\mathbf{d}$ is found from Eq.\ (17) by transforming $\mathbf{D}$, the {\em rate of deformation tensor}, with the rotation tensor $\mathbf{R}$.  This entire process is taken from the paper by Flanagan and Taylor ``An accurate numerical algorithm for stress integration with finite rotations'', which I believe to be a standard algorithm for developing kinematics in large displacement problems.
     \end{itemize}}
%
     \item Notation: Usually lower case letters are used for spatial quantities and upper case letters are used for referential quantities, e.g., $\mathbf{d}$ is standard for the rate of deformation tensor.
  {\color{red}
     \begin{itemize}
        \item All of the notation used in this paper follows very closely the notation used by Malvern and/or Flanagan and Taylor.
     \end{itemize}}
%
     \item I don't understand why eq.\ (35) is presented. As far as I know, the use of previous values of strain in order to extrapolate the strain rate is never used in the literature, and all previous authors obtain the strain by solving an equation like (36). So unless there is some good reason to believe that (35) might be plausible (e.g., an instance where it is successfully used in the previous literature), I would drop it from the paper.
  {\color{red}
     \begin{itemize}
        \item For the reviewer's reference a paper from Brown University by Peirce, Shih, and Needleman titled ``A tangent modulus method for rate dependent solids'' uses a method similar to what is presented in Eq.\ (35), a fully explicit method.
        \end{itemize}}
%
        \item Mistaken forward reference to eq. 59 on p. 14.
  {\color{red}
     \begin{itemize}
        \item Corrected in revision
     \end{itemize}}
%
      \item The paper mixes up modeling with discretization. For example, the model apparently doesn't discuss how the pressure component of stress is handled, and yet this pops up in the discussion of the numerics (eq.\ (51)).
  {\color{red}
     \begin{itemize}
        \item I do not fully understand this comment, von Mises plasticity only deals with the deviatoric part of stress to define a yield surface, but the pressure component must be added back in after the deviatoric stress has been updated. Nevertheless, a mistake was found in Eq.\ (51) and corrected in the revision.
      \end{itemize}}
%
        \item The total strain is apparently defined on a particle-by-particle basis using (61) and (12). I suppose the plastic strain is a nodal history variable?
  {\color{red}
     \begin{itemize}
        \item All stress or strain quantities including the internal state variable, {\em equivalent plastic strain}, are calculated for each particle at each timestep. 
     \end{itemize}}
%
        \item The number of nodes required seems to be huge in both simulations in Section 10. Is there any compensating benefit (versus finite elements)?
  {\color{red}
     \begin{itemize}
        \item The number on nodes was large, but these simulations ran in only a few hours on 4 processors.  The largest compensating benefit is the ability to model failure (outside the scope of this paper, but discussed in others) with the same set of fundamental equations without the need for separate governing equations for crack initiation, propagation, etc. (e.g., cohesive zones, element death).
     \end{itemize}}
 %
        \item The literature review includes previous papers (mainly by the same group of authors) on peridynamics but does not review the larger universe of particle methods for solid mechanics, and I think some references in this regard would be helpful. In particular, I think some of the previous particle methods are quite similar to peridynamics. Unfortunately, I am not an expert in this, but I know of at least three previous papers proposing particle methods for solid mechanics.
       \begin{enumerate}
         \item Rabczuk, Belytschko and Xiao, Stable particle methods based on Lagrangian kernels, CMAME 193 (2004) 1035-1063.
         \item Delaplace and Ibrahimbegovic, Performance of time-stepping schemes for discrete models in fracture dynamics, IJNME 65 (2005) 1527-1544.
         \item Gao and Klein, Numerical simulation of crack growht in an isotropic solid with randomized internal cohesive bonds, JMPS 46 (1998) 187-218.
       \end{enumerate}
%
     {\color{red}
     \begin{itemize}
        \item The main reason other particle methods were not reviewed was because all of these other methods are based on theory which solves partial differential equations and are therefore fundamentally different from the nonlocal integral equations solved in peridynamics.  However, additional discussion was added in the revision which points out these differences and references the reviewers suggested papers where appropriate. 
         \end{itemize}}
        %
        \item Inconsistent verb tense in second paragraph of summary (is, was.)

p. 2, Where $\rho \rightarrow$ where $\rho$

Van der Walls $\rightarrow$ Van der Waals

any pair of particles interact $\rightarrow$ interacts

singe $\rightarrow$ single

spacial $\rightarrow$ spatial

principle $\rightarrow$ principal

et. al $\rightarrow$ et al. (italicize)

%
     {\color{red}
     \begin{itemize}
       \item Corrected in revision
     \end{itemize}}
%
\end{enumerate}

\section{Reviewer}
\subsection*{General Statement}
I recommend that the authors prepare a revised version of their paper to be considered for publication in the International Journal of Numerical Methods in Engineering.  Specifically, the authors should consider adding discussion related to the following eleven comments (1)-(11).
%
\subsection*{Specific Comments}

\begin{enumerate}

  \item This is the first time that I have read about the peridynamic theory.  In many ways it seems to be similar to the notions used in SPH (smooth particle hydrodynamics).  It is know that the SPH theories have difficulties handling material strength in a consistent stable manner.  I would expect the same problems with the peridynamic theory. It would be helpful to compare these theories and explain important differences and potential advantages. 
%
     {\color{red}
     \begin{itemize}
        \item Peridynamics is fundamentally different from SPH because SPH is still based on theory that solves partial differential equations.  In SPH, properties of each particle are smeared over a smoothing length, but the motion of the particles is still governed by PDE's.  In peridynamics the conservation of momentum equation is a nonlocal integral equation capable of handling even discontinuous displacements fields that would cause problems in SPH where the PDE's could not be evaluated.  SPH also suffers from a tensile instability that is not problematic for peridynamics.  Additional discussion has been added to the introduction to address these issues.
     \end{itemize}}
%
  \item Introducing nonlocality through the integral over the sphere of radius $\delta$ in the reference configuration violates the notion of a finite wave speed in the material.  Obviously the same problem occurs when finite elements are used to average deformation information over a finite region.  Some comments about the effect of this in this peridynamic theory would be helpful
     {\color{red}
     \begin{itemize}
        \item As the reviewer points out this is a similar issue in finite elements.  Wave speeds have been analyzed in peridynamics and have been shown to not have an infinite wave speed, but the wave speeds are a function of wave length.  When analyzed it turns out that the bulk modulus is directly related to $\delta$ and therefore, becomes an integral part of the defination of wave speed, i.e. $\delta$ is a material property.
     \end{itemize}}
%
  \item On page 2 it is stated that:
  \begin{quotation}
  \noindent ``In this implementation, $\omega$ is always either 0 or 1, which corresponds to whether the bond is `broken' or not.'' 
 \end{quotation}
 It seems to me that $\delta$ defines the radius of influence of material points and $\omega$ defines the strength of the influence.  In particular, it is surprising to me that a uniform distribution of $\omega$ with a sharp cut-off would not cause serious problems with wave propagation.  It would be helpful to comment on this in a revised version of the paper. 
     {\color{red}
     \begin{itemize}
        \item The reviewer is correct in his recognition of $\omega$ as a strength of influence.  So far only uniform distributions of $\omega$ have been evaluated and they do not show any numerical instablity.  Additional discussion was added to the revision.
     \end{itemize}}
%
  \item It would be helpful to provide a discussion of how displacement and traction boundary conditions can be implemented using this theory.
     {\color{red}
     \begin{itemize}
        \item Displacements, velocities, and body force densities can be applied to regions of finite volume.  Information on this was added to the revised article in the discussion of the numerical simulations.  See Silling ``Reformulation of elasticity theory for discontinuities and long-range forces'' cited in article for further reading.
     \end{itemize}}
%
  \item It would be helpful to discuss the specific approximations that are used when the sphere of influence (associated with $\delta$) extends outside of the boundaries of the body. 
     {\color{red}
     \begin{itemize}
        \item There are no issues when $\delta$ extends outside the boundaries of the body.  A proof has been included which shows that the deformation gradient can be correctly calculated for any assumed deformation state in the revised article.
     \end{itemize}}
 %
  \item It is my understanding that this method tracts the motion of material points. Since the sphere of influence (associated with $\delta$) is defined relative to the reference configuration, the same material points remain in communication even if the material is highly stretched or highly compressed. It would seem to me that for the highly stretched case the method would yield unphysical dispersion of waves since the region of influence can be quite large.
     {\color{red}
     \begin{itemize}
        \item The reviewer's assumptions are correct, but the same thing could happen in a finite element simulation as well when elements become severely stretched or distorted.  It would be possible to redefine the $\delta$ at every time step where material points far away would be removed from this horizon and others added; however, this adaptivity feature does not currenlty exist in the code used here, Emu.
     \end{itemize}}
 %
  \item For an elastic solid, which has a unique unstressed reference shape, it makes sense to refer the constitutive response to a reference configuration.  In contrast, the material response of an elastic-viscoplastic material, which has no unique unstressed configuration or shape, should be formulated in terms of constitutive equations which are explicitly independent of the arbitrary choice of the reference configuration. Moreover, since the main motivation for developing meshless methods is to handle large deformations it is unfortunate that the elastic-viscoplastic constitutive equations have been formulated with respect to the reference configuration using the rotation tensor $\mathbf{R}$ in (15).  Also, the integral operators are specified by considering regions defined in the reference configuration.  It is possible to develop this method in an Eulerian framework more similar to SPH methods? 
     {\color{red}
     \begin{itemize}
       \item The method of using $\mathbf{R}$ to provide an unrotated configuration in which to apply the constitutive model such that the stress rates will be objective is used in many finite element codes, including the production code at Sandia National Laboratories, PRESTO.  This method is described in detail in the Flanagan and Taylor paper ``An accurate numerical algorithm for stress integration with finite rotations.''  Futhermore, Eulerian frame constitutive equations can be used in peridynamics as long as we define $\mathbf{F}$ and $\mathbf{R}$.
     \end{itemize}}
     %
  \item The constitutive equations used for determining viscoplastic response are not clear. It is my understanding that (28) is a yield function which limits the magnitude of the deviatoric stress during loading
  \begin{equation}
    S = | \mathbf{S} | = \sqrt{\frac{2}{3}} G \quad \mbox{during loading.}
    \tag{28}
  \end{equation}
  Then, using the additive decomposition (29), the elastic constitutive equation (31), the flow rule (32) and (33), and the expression (38) it can be shown that 
  \begin{equation}
    \dot{S} = 2 \mu \left ( \mathbf{Q} : \dot{\mathbf{e}} - \sqrt{\frac{3}{2}} \dot{\varepsilon}_p \right ) = \sqrt{\frac{2}{3}} \dot{G}
    \tag{A1}
  \end{equation}
  However, viscoplasticity is introduced in by taking the value of $G$ to be a function of the equivalent plastic strain $\varepsilon_p$ and plastic strain rate $\dot{\varepsilon}_p$
  \begin{equation}
    G = f \left ( \varepsilon_p ,  \dot{\varepsilon}_p \right ) .
    \tag{34}
  \end{equation}
  Thus, substituting (34) into (A1) yields
  \begin{equation}
    2 \mu \left ( \mathbf{Q} : \dot{\mathbf{e}} - \sqrt{\frac{3}{2}} \dot{\varepsilon_p} \right ) = \left ( \frac{\partial f}{\partial \varepsilon_p} \right ) \dot{\varepsilon}_p + \left ( \frac{\partial f}{\partial \dot{\varepsilon}_p} \right ) \ddot{\varepsilon}_p
    \tag{A2}
  \end{equation}
  which is a second order differential equation for the equivalent plastic strain.  It is not clear that the proposed procedure described in sections 6 and 7 actually solve this second order equation for $\varepsilon_p$.
     {\color{red}
     \begin{itemize}
       \item What is derived in the paper is the finite differencing scheme for solving for an increment of equivalent plastic strain.  There is no need to consider the continous derivatives when we do this.  Nevertheless I will start with equation (A1) and show that it equals the final result from the paper.  Taking the right-hand equality from (A1) we can substitute (39)\footnote{\color{red} Equation numbers refer to equation numbers in original paper, not the revision} for $\mathbf{Q:\dot{e}}$. Noting that because $\mathbf{Q}$ and $\mathbf{\dot{e}}$ are symmetric tensors $\mathbf{Q:\dot{e}} = \mathbf{\dot{e}:Q}$.
        \begin{equation}
	  2 \mu \left ( \dot{e} - \sqrt{\frac{3}{2}} \dot{\varepsilon}_p \right ) = \sqrt{\frac{2}{3}} \dot{G}
	    \tag{B1}
	  \end{equation}
	  Now substitute (38) and (41)
	  \begin{equation}
	    2 \mu \dot{e}^e = \sqrt{\frac{2}{3}} \dot{G}
	    \tag{B2}
	  \end{equation}
	  Convert from continous derivatives to rates.
	  \begin{equation}
	    2 \mu \frac{\Delta e_{t-\Delta t/2}^e}{\Delta t} = \sqrt{\frac{2}{3}} \frac{\Delta G_{t-\Delta t/2}}{\Delta t}
	    \tag{B3}
	  \end{equation}
	  Multiply by $\Delta t$ and expand the RHS of (B3)
	  \begin{equation}
	    2 \mu {\Delta e_{t-\Delta t/2}^e} = \sqrt{\frac{2}{3}} \left ( G_{t}- G_{t-\Delta t} \right )
	    \tag{B4}
	  \end{equation}
	  and of course $G_t = \sqrt{3/2} S_t$ and similary for $G_{t-\Delta t}$, using these in (B4)
	  \begin{equation}
	    2 \mu {\Delta e_{t-\Delta t/2}^e} = S_t - S_{t-\Delta t}
	    \tag{B5}
	  \end{equation}
	  finally substituting the rate form of (41) and rearranging we have:
	  \begin{equation}
	    \Delta e_{t-\Delta t/2} - \frac{S_t - S_{t-\Delta t}}{2 \mu} - \Delta \lambda_{t-\Delta t/2} = 0
	    \tag{B6}
	  \end{equation}
	  (B6) is exactly (44).  The rest of derivation follows the paper from (44).
     \end{itemize}}
 %
  \item Line 7 after (8).  $\mathbf{K}$  is a second order moment not a first order moment since $\boldsymbol{\xi} \otimes \boldsymbol{\xi}$ is a quadratic term.
     {\color{red}
     \begin{itemize}
        \item The description of $\mathbf{K}$ as a moment has been removed in the revision.
     \end{itemize}}
     %
  \item The value of $\delta$ used for the simulations should be specified.  In this regard, it would be good to also state how many material points used in the simulation fit into this sphere of radius $\delta$. 
     {\color{red}
     \begin{itemize}
        \item Suggestion was included in revision along with additional discussion about how changing $\delta$ makes insignificant changes in the solution of the Taylor impact problem.
     \end{itemize}}
        %
  \item Figure 2 refers to true strain but the measure $\mathbf{E}$ in (61) is the Lagranian strain.  Please provide a definition of the strain measure being used for true strain in Fig. 2. 
     {\color{red}
     \begin{itemize}
        \item Suggestion was included in revision.
     \end{itemize}}
\end{enumerate}

\end{document}

### 2014-02-05 jacobian_calc_CMAME.tex
Added response folder and trackchanges.sty

@@
 
-\begin{document}
+\begin{document}\sloppy
 
@@
 %
-This section contains the average results from the study presented in tabular and graphic format.
 Table~\ref{tab:results} shows the compute core count, number of nonzero tangent-stiffness (TS) non-zero matrix elements, load step average calculation time and load step average accuracy.
-The units for accuracy are derived from the
-expression used to calculate the TS matrix elements and the units used to define bulk material properties, mentioned in Section~\ref{tcs}.
+The units for accuracy are derived equation~(\ref{eqn:accuracy}) and the units used to define bulk material properties, mentioned in Section~\ref{tcs}.
 %
-\note[JTF]{Michael, work on cleaning this table up for legibility}
-\note[Michael]{Columns are now separated out and can be fine tuned}
-
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-
-\begin{table}    
+\begin{table*}[!ht]    
+  \scriptsize
+  \centering
         \caption{Averaged Results, Each Test} \label{tab:results}   
-       \hskip-2.0cm\begin{tabular}{C{1.5cm} C{1.88cm} C{1.5cm} C{1.5cm} C{1.5cm} C{1.5cm} C{1.88cm} C{1.88cm} C{1.5cm}}
-            Cores & TS Elements & Calc. Time (s) & & & & Accuracy Diff. ($\si{\mega\pascal}$) & & \\ 
-        \multicolumn{2}{l}{} & CS & CD & FD & AD & CS & CD & FD \\ \hline 
-        1 & $\num{1.99E6}$  & 3.5 & 2.7 & 1.7 & 1.6 & $\num{1.92E-10}$ & $\num{1.21E-4}$ & .137 \\ \hline
-        1 & $\num{4.25E6}$  & 6.2& 4.9& 3.2& 2.9 & $\num{2.28E-10}$& $\num{9.94E-4}$ & .148 \\ \hline
-        1 & $\num{7.79E6}$  & 11.2& 8.9& 5.7& 5.2 & $\num{2.38E-10}$& $\num{1.59E-4}$ & .145\\ \hline
-        1 & $\num{1.54E7}$  & 26.7& 21.0& 13.4& 12.2 & $\num{2.33E-10}$ & $\num{4.61E-4}$ & .12 \\ \hline
-        1 & $\num{2.00E7}$  & 28.1& 22.2& 14.4& 13.1 & $\num{2.76E-10}$ & $\num{1.05E-3}$ & .145 \\ \hline
-        1 & $\num{3.14E7}$  & 47.6& 37.7& 24.2& 21.9 & $\num{2.64E-10}$ & $\num{1.65E-3}$ & .133 \\ \hline
-        1 & $\num{4.01E7}$  & 55.6& 44.1& 28.4& 25.9 & $\num{3.03E-10}$ & $\num{1.92E-3}$ & .148 \\ \hline
-        1 & $\num{8.39E7}$  & 138.9& 109.6& 70.2& 64.0 & $\num{3.63E-10}$ & $\num{1.64E-3}$ & .123 \\ \hline
-        1 & $\num{1.65E8}$  & 277.3& 218.1& 139.4& 126.5 & $\num{3.26E-10}$ & $\num{2.18E-3}$ & .128 \\ \hline
-        32 & $\num{1.67E10}$  & 336.1& 277.7& 200.6& 233.1 & $\num{6.21E-10}$ & $\num{1.52E-2}$ & .176 \\ \hline
-        64 & $\num{1.67E10}$  & 169.9& 140.7& 102.0& 119.7 & $\num{6.20E-10}$ & $\num{1.50E-2}$ & .177 \\ \hline
-        96 & $\num{1.67E10}$  & 114.7& 95.0& 69.1 & 79.7 & $\num{6.18E-10}$ & $\num{1.50E-2}$ & .177 \\ \hline
-        128 & $\num{1.67E10}$  & 86.4& 71.8& 52.4 &58.8 & $\num{6.16E-10}$ & $\num{1.47E-2}$ & .177 \\ \hline
+       \begin{tabular}{c c c c c c c c c}
+         \toprule
+         Cores & TS Elements & \multicolumn{4}{c}{Calculation Time ($\si{\second}$)} & \multicolumn{3}{c}{Accuracy Difference ($\si{\mega\pascal}$)} \\ 
+         & $\times 10^6$ & CS & CD & FD & AD & CS & CD & FD \\
+        \midrule
+        1 & $\num{1.99}$  & 3.5 & 2.7 & 1.7 & 1.6 & $\num{1.92E-10}$ & $\num{1.21E-4}$ & .137 \\
+        1 & $\num{4.25}$  & 6.2& 4.9& 3.2& 2.9 & $\num{2.28E-10}$& $\num{9.94E-4}$ & .148 \\
+        1 & $\num{7.79}$  & 11.2& 8.9& 5.7& 5.2 & $\num{2.38E-10}$& $\num{1.59E-4}$ & .145\\
+        1 & $\num{15.4}$  & 26.7& 21.0& 13.4& 12.2 & $\num{2.33E-10}$ & $\num{4.61E-4}$ & .12 \\
+        1 & $\num{20.0}$  & 28.1& 22.2& 14.4& 13.1 & $\num{2.76E-10}$ & $\num{1.05E-3}$ & .145 \\
+        1 & $\num{31.4}$  & 47.6& 37.7& 24.2& 21.9 & $\num{2.64E-10}$ & $\num{1.65E-3}$ & .133 \\
+        1 & $\num{40.1}$  & 55.6& 44.1& 28.4& 25.9 & $\num{3.03E-10}$ & $\num{1.92E-3}$ & .148 \\
+        1 & $\num{83.9}$  & 138.9& 109.6& 70.2& 64.0 & $\num{3.63E-10}$ & $\num{1.64E-3}$ & .123 \\
+        1 & $\num{165.0}$  & 277.3& 218.1& 139.4& 126.5 & $\num{3.26E-10}$ & $\num{2.18E-3}$ & .128 \\
+        32 & $\num{1670}$  & 336.1& 277.7& 200.6& 233.1 & $\num{6.21E-10}$ & $\num{1.52E-2}$ & .176 \\
+        64 & $\num{1670}$  & 169.9& 140.7& 102.0& 119.7 & $\num{6.20E-10}$ & $\num{1.50E-2}$ & .177 \\
+        96 & $\num{1670}$  & 114.7& 95.0& 69.1 & 79.7 & $\num{6.18E-10}$ & $\num{1.50E-2}$ & .177 \\
+        128 & $\num{1670}$  & 86.4& 71.8& 52.4 &58.8 & $\num{6.16E-10}$ & $\num{1.47E-2}$ & .177 \\
+        \bottomrule
     \end{tabular}
-\end{table}
+\end{table*}
 

### 2014-02-08 response/response.tex
response to I.1 started

@@
 %%%% First Reviewer %%%%
-\section{Reviewer}
+\section{Information to the authors}
 %
-\subsection*{General Statement}
-In general, I am not convinced of the benefits of particle methods when compared to traditional finite element analysis.
-Although they eliminate the sometimes troublesome problem of mesh generation, they create new problems of their own.
-In particular, they are hard to debug (because the patch test often doesn’t work), it is typically difficult to pin down an order of accuracy, and it is typically difficult to apply boundary conditions.
-Nonetheless, I recognize that they are part of the toolbox and therefore are within the scope of IJNME.
+\subsection*{Larger issues}
 %
+\begin{enumerate}
+    \item
+         Some parts of the manuscript suggest that the authors introduced a new method,
+         i.e., the complex-step (CS) method. For example, in the introduction, they say
+         “A goal of this study was to develop and evaluate a new, accurate, and practical
+         method ...” Another example appears in page 3, where the authors say “The aim
+         of this paper is not only to introduce the new complex-step method in the context
+         of evaluating tangent-stiffness matrices ..”. However, the CS method has already
+         appeared in many publications since 1998 [1]. Please revise the way you refer to
+         the CS method along the manuscript
+
 {\color{red}
   \begin{itemize}
-    \item We have added a paragraph at the end of Section 1 that highlights the difference between peridynamics and previous “particle methods.” The main difference is that in peridynamics, we numerically approximate a different set of continuum equations.
+    \item We have added a paragraph at the end of Section 1 that highlights the difference between peridynamics and previous Òparticle methods.Ó The main difference is that in peridynamics, we numerically approximate a different set of continuum equations.
   \end{itemize}}
 
-This particular paper, however, has insufficient novelty to merit publication in its present form. In particular,
-
-\begin{enumerate}
-  \item The peridynamic framework, including the method to compute the deformation gradient, is already established in the previous literature.
-  {\color{red}
-    \begin{itemize}
-      \item We have added statements in the abstract and introduction that highlight what is new about the work being reported.
-Specifically, this represents the first successful implementation of a rate-dependent material model within peridynamics.
-We also present the first full-scale 3D solution of a realistic, reasonably complex application.
-This work therefore demonstrates to the community, for the first time, the ability of the peridynamic method to solve practical problems involving complex materials resulting in good agreement with experimental data.
-    \end{itemize}}
-  %
-  \item The viscoplastic formulation and constitive model follows fairly classical nonlinear solid mechanics, and therefore does not appear to be novel. 
-  {\color{red}
-    \begin{itemize}
-      \item Viscoplasticity has never been used within peridynamics.  See previous responses.
-    \end{itemize}}
-  %
-  \item The computational studies are quite limited, by the authors’ own admission (“while failure can be modeled with peridynamics, it is beyond the scope of this paper; therefore, these experiments were not attempted in numerical simulation”, p.
-19).
-  {\color{red}
-  \begin{itemize}
-    \item The focus of this paper was to show that peridynamics could be used with conventional (stress-strain) constitutive models.
-Modeling failure is the main advantage of the peridynamic approach, which has been demonstrated in the cited papers and will be further developed in a future paper with the combination of a viscoplastic constitutive model and a novel failure mechanism applied by breaking ``bonds''.
-  \end{itemize}}
-  %
-\end{enumerate}
-
-\subsection*{Specific Comments}
-
-\begin{enumerate}
-%
-   \item p. 3, It has been proved that Equation 1 reduces to the classical continuum partial differential equation\dots Which one? Momentum?
-   {\color{red}
-     \begin{itemize}
-        \item Yes, linear momentum.  This has been made more clear in the rewritten introduction.
-     \end{itemize}}
-%
-  \item p. 5, should be $F = I + \nabla_x u$ (standard notation)
-   {\color{red}
-     \begin{itemize}
-        \item The notation I used is identical to the notation used by Lawrence E.
-Malvern in his classic book \underline{Introduction to the Mechanics of a Continuous Medium} (p.\ 162 in the version I have).
-This is further clarified by showing the incidental notation in the revision.
-     \end{itemize}}
-%
-  \item Eq.
-(30) is an additive decomposition of the deviator of the referential tensor $\mathbf{d}$ based on the additive decomposition of the rate of deformation $\mathbf{D}$.
-A reference and/or some background should be included.
-  {\color{red}
-     \begin{itemize}
-        \item Eq.\ (30) is an additive decomposition of the deviatoric part of $\mathbf{d}$ which is referred to as the {\em unrotated rate of deformation tensor}.
-$\mathbf{d}$ is found from Eq.\ (17) by transforming $\mathbf{D}$, the {\em rate of deformation tensor}, with the rotation tensor $\mathbf{R}$.
-This entire process is taken from the paper by Flanagan and Taylor ``An accurate numerical algorithm for stress integration with finite rotations'', which I believe to be a standard algorithm for developing kinematics in large displacement problems.
-     \end{itemize}}
-%
-     \item Notation: Usually lower case letters are used for spatial quantities and upper case letters are used for referential quantities, e.g., $\mathbf{d}$ is standard for the rate of deformation tensor.
-  {\color{red}
-     \begin{itemize}
-        \item All of the notation used in this paper follows very closely the notation used by Malvern and/or Flanagan and Taylor.
-     \end{itemize}}
-%
-     \item I don't understand why eq.\ (35) is presented.
-As far as I know, the use of previous values of strain in order to extrapolate the strain rate is never used in the literature, and all previous authors obtain the strain by solving an equation like (36).
-So unless there is some good reason to believe that (35) might be plausible (e.g., an instance where it is successfully used in the previous literature), I would drop it from the paper.
-  {\color{red}
-     \begin{itemize}
-        \item For the reviewer's reference a paper from Brown University by Peirce, Shih, and Needleman titled ``A tangent modulus method for rate dependent solids'' uses a method similar to what is presented in Eq.\ (35), a fully explicit method.
-        \end{itemize}}
-%
-        \item Mistaken forward reference to eq. 59 on p. 14.
-  {\color{red}
-     \begin{itemize}
-        \item Corrected in revision
-     \end{itemize}}
-%
-      \item The paper mixes up modeling with discretization.
-For example, the model apparently doesn't discuss how the pressure component of stress is handled, and yet this pops up in the discussion of the numerics (eq.\ (51)).
-  {\color{red}
-     \begin{itemize}
-        \item I do not fully understand this comment, von Mises plasticity only deals with the deviatoric part of stress to define a yield surface, but the pressure component must be added back in after the deviatoric stress has been updated.
-Nevertheless, a mistake was found in Eq.\ (51) and corrected in the revision.
-      \end{itemize}}
-%
-        \item The total strain is apparently defined on a particle-by-particle basis using (61) and (12). I suppose the plastic strain is a nodal history variable?
-  {\color{red}
-     \begin{itemize}
-        \item All stress or strain quantities including the internal state variable, {\em equivalent plastic strain}, are calculated for each particle at each timestep. 
-     \end{itemize}}
-%
-        \item The number of nodes required seems to be huge in both simulations in Section 10. Is there any compensating benefit (versus finite elements)?
-  {\color{red}
-     \begin{itemize}
-        \item The number on nodes was large, but these simulations ran in only a few hours on 4 processors.
-The largest compensating benefit is the ability to model failure (outside the scope of this paper, but discussed in others) with the same set of fundamental equations without the need for separate governing equations for crack initiation, propagation, etc.
-(e.g., cohesive zones, element death).
-     \end{itemize}}
- %
-        \item The literature review includes previous papers (mainly by the same group of authors) on peridynamics but does not review the larger universe of particle methods for solid mechanics, and I think some references in this regard would be helpful.
-In particular, I think some of the previous particle methods are quite similar to peridynamics.
-Unfortunately, I am not an expert in this, but I know of at least three previous papers proposing particle methods for solid mechanics.
-       \begin{enumerate}
-         \item Rabczuk, Belytschko and Xiao, Stable particle methods based on Lagrangian kernels, CMAME 193 (2004) 1035-1063.
-         \item Delaplace and Ibrahimbegovic, Performance of time-stepping schemes for discrete models in fracture dynamics, IJNME 65 (2005) 1527-1544.
-         \item Gao and Klein, Numerical simulation of crack growht in an isotropic solid with randomized internal cohesive bonds, JMPS 46 (1998) 187-218.
-       \end{enumerate}
-%
-     {\color{red}
-     \begin{itemize}
-        \item The main reason other particle methods were not reviewed was because all of these other methods are based on theory which solves partial differential equations and are therefore fundamentally different from the nonlocal integral equations solved in peridynamics.
-However, additional discussion was added in the revision which points out these differences and references the reviewers suggested papers where appropriate.
-         \end{itemize}}
-        %
-        \item Inconsistent verb tense in second paragraph of summary (is, was.)
-
-p. 2, Where $\rho \rightarrow$ where $\rho$
-
-Van der Walls $\rightarrow$ Van der Waals
-
-any pair of particles interact $\rightarrow$ interacts
-
-singe $\rightarrow$ single
-
-spacial $\rightarrow$ spatial
-
-principle $\rightarrow$ principal
-
-et. al $\rightarrow$ et al. (italicize)
-
-%
-     {\color{red}
-     \begin{itemize}
-       \item Corrected in revision
-     \end{itemize}}
-%
-\end{enumerate}
-
-\section{Reviewer}
-\subsection*{General Statement}
-I recommend that the authors prepare a revised version of their paper to be considered for publication in the International Journal of Numerical Methods in Engineering.
-Specifically, the authors should consider adding discussion related to the following eleven comments (1)-(11).
-%
-\subsection*{Specific Comments}
-
-\begin{enumerate}
-
-  \item This is the first time that I have read about the peridynamic theory.
-In many ways it seems to be similar to the notions used in SPH (smooth particle hydrodynamics).
-It is know that the SPH theories have difficulties handling material strength in a consistent stable manner.
-I would expect the same problems with the peridynamic theory.
-It would be helpful to compare these theories and explain important differences and potential advantages.
-%
-     {\color{red}
-     \begin{itemize}
-        \item Peridynamics is fundamentally different from SPH because SPH is still based on theory that solves partial differential equations.
-In SPH, properties of each particle are smeared over a smoothing length, but the motion of the particles is still governed by PDE's.
-In peridynamics the conservation of momentum equation is a nonlocal integral equation capable of handling even discontinuous displacements fields that would cause problems in SPH where the PDE's could not be evaluated.
-SPH also suffers from a tensile instability that is not problematic for peridynamics.
-Additional discussion has been added to the introduction to address these issues.
-     \end{itemize}}
-%
-  \item Introducing nonlocality through the integral over the sphere of radius $\delta$ in the reference configuration violates the notion of a finite wave speed in the material.
-Obviously the same problem occurs when finite elements are used to average deformation information over a finite region.
-Some comments about the effect of this in this peridynamic theory would be helpful
-     {\color{red}
-     \begin{itemize}
-        \item As the reviewer points out this is a similar issue in finite elements.
-Wave speeds have been analyzed in peridynamics and have been shown to not have an infinite wave speed, but the wave speeds are a function of wave length.
-When analyzed it turns out that the bulk modulus is directly related to $\delta$ and therefore, becomes an integral part of the defination of wave speed, i.e.
-$\delta$ is a material property.
-     \end{itemize}}
-%
-  \item On page 2 it is stated that:
-  \begin{quotation}
-  \noindent ``In this implementation, $\omega$ is always either 0 or 1, which corresponds to whether the bond is `broken' or not.'' 
- \end{quotation}
- It seems to me that $\delta$ defines the radius of influence of material points and $\omega$ defines the strength of the influence.
-In particular, it is surprising to me that a uniform distribution of $\omega$ with a sharp cut-off would not cause serious problems with wave propagation.
-It would be helpful to comment on this in a revised version of the paper.
-     {\color{red}
-     \begin{itemize}
-        \item The reviewer is correct in his recognition of $\omega$ as a strength of influence.
-So far only uniform distributions of $\omega$ have been evaluated and they do not show any numerical instablity.
-Additional discussion was added to the revision.
-     \end{itemize}}
-%
-  \item It would be helpful to provide a discussion of how displacement and traction boundary conditions can be implemented using this theory.
-     {\color{red}
-     \begin{itemize}
-        \item Displacements, velocities, and body force densities can be applied to regions of finite volume.
-Information on this was added to the revised article in the discussion of the numerical simulations.
-See Silling ``Reformulation of elasticity theory for discontinuities and long-range forces'' cited in article for further reading.
-     \end{itemize}}
-%
-  \item It would be helpful to discuss the specific approximations that are used when the sphere of influence (associated with $\delta$) extends outside of the boundaries of the body. 
-     {\color{red}
-     \begin{itemize}
-        \item There are no issues when $\delta$ extends outside the boundaries of the body.
-A proof has been included which shows that the deformation gradient can be correctly calculated for any assumed deformation state in the revised article.
-     \end{itemize}}
- %
-  \item It is my understanding that this method tracts the motion of material points.
-Since the sphere of influence (associated with $\delta$) is defined relative to the reference configuration, the same material points remain in communication even if the material is highly stretched or highly compressed.
-It would seem to me that for the highly stretched case the method would yield unphysical dispersion of waves since the region of influence can be quite large.
-     {\color{red}
-     \begin{itemize}
-        \item The reviewer's assumptions are correct, but the same thing could happen in a finite element simulation as well when elements become severely stretched or distorted.
-It would be possible to redefine the $\delta$ at every time step where material points far away would be removed from this horizon and others added; however, this adaptivity feature does not currenlty exist in the code used here, Emu.
-     \end{itemize}}
- %
-  \item For an elastic solid, which has a unique unstressed reference shape, it makes sense to refer the constitutive response to a reference configuration.
-In contrast, the material response of an elastic-viscoplastic material, which has no unique unstressed configuration or shape, should be formulated in terms of constitutive equations which are explicitly independent of the arbitrary choice of the reference configuration.
-Moreover, since the main motivation for developing meshless methods is to handle large deformations it is unfortunate that the elastic-viscoplastic constitutive equations have been formulated with respect to the reference configuration using the rotation tensor $\mathbf{R}$ in (15).
-Also, the integral operators are specified by considering regions defined in the reference configuration.
-It is possible to develop this method in an Eulerian framework more similar to SPH methods?
-     {\color{red}
-     \begin{itemize}
-       \item The method of using $\mathbf{R}$ to provide an unrotated configuration in which to apply the constitutive model such that the stress rates will be objective is used in many finite element codes, including the production code at Sandia National Laboratories, PRESTO.
-This method is described in detail in the Flanagan and Taylor paper ``An accurate numerical algorithm for stress integration with finite rotations.''  Futhermore, Eulerian frame constitutive equations can be used in peridynamics as long as we define $\mathbf{F}$ and $\mathbf{R}$.
-     \end{itemize}}
-     %
-  \item The constitutive equations used for determining viscoplastic response are not clear.
-It is my understanding that (28) is a yield function which limits the magnitude of the deviatoric stress during loading
-  \begin{equation}
-    S = | \mathbf{S} | = \sqrt{\frac{2}{3}} G \quad \mbox{during loading.}
-    \tag{28}
-  \end{equation}
-  Then, using the additive decomposition (29), the elastic constitutive equation (31), the flow rule (32) and (33), and the expression (38) it can be shown that 
-  \begin{equation}
-    \dot{S} = 2 \mu \left ( \mathbf{Q} : \dot{\mathbf{e}} - \sqrt{\frac{3}{2}} \dot{\varepsilon}_p \right ) = \sqrt{\frac{2}{3}} \dot{G}
-    \tag{A1}
-  \end{equation}
-  However, viscoplasticity is introduced in by taking the value of $G$ to be a function of the equivalent plastic strain $\varepsilon_p$ and plastic strain rate $\dot{\varepsilon}_p$
-  \begin{equation}
-    G = f \left ( \varepsilon_p ,  \dot{\varepsilon}_p \right ) .
-    \tag{34}
-  \end{equation}
-  Thus, substituting (34) into (A1) yields
-  \begin{equation}
-    2 \mu \left ( \mathbf{Q} : \dot{\mathbf{e}} - \sqrt{\frac{3}{2}} \dot{\varepsilon_p} \right ) = \left ( \frac{\partial f}{\partial \varepsilon_p} \right ) \dot{\varepsilon}_p + \left ( \frac{\partial f}{\partial \dot{\varepsilon}_p} \right ) \ddot{\varepsilon}_p
-    \tag{A2}
-  \end{equation}
-  which is a second order differential equation for the equivalent plastic strain.
-It is not clear that the proposed procedure described in sections 6 and 7 actually solve this second order equation for $\varepsilon_p$.
-     {\color{red}
-     \begin{itemize}
-       \item What is derived in the paper is the finite differencing scheme for solving for an increment of equivalent plastic strain.
-There is no need to consider the continous derivatives when we do this.
-Nevertheless I will start with equation (A1) and show that it equals the final result from the paper.
-Taking the right-hand equality from (A1) we can substitute (39)\footnote{\color{red} Equation numbers refer to equation numbers in original paper, not the revision} for $\mathbf{Q:\dot{e}}$.
-Noting that because $\mathbf{Q}$ and $\mathbf{\dot{e}}$ are symmetric tensors $\mathbf{Q:\dot{e}} = \mathbf{\dot{e}:Q}$.
-        \begin{equation}
-	  2 \mu \left ( \dot{e} - \sqrt{\frac{3}{2}} \dot{\varepsilon}_p \right ) = \sqrt{\frac{2}{3}} \dot{G}
-	    \tag{B1}
-	  \end{equation}
-	  Now substitute (38) and (41)
-	  \begin{equation}
-	    2 \mu \dot{e}^e = \sqrt{\frac{2}{3}} \dot{G}
-	    \tag{B2}
-	  \end{equation}
-	  Convert from continous derivatives to rates.
-	  \begin{equation}
-	    2 \mu \frac{\Delta e_{t-\Delta t/2}^e}{\Delta t} = \sqrt{\frac{2}{3}} \frac{\Delta G_{t-\Delta t/2}}{\Delta t}
-	    \tag{B3}
-	  \end{equation}
-	  Multiply by $\Delta t$ and expand the RHS of (B3)
-	  \begin{equation}
-	    2 \mu {\Delta e_{t-\Delta t/2}^e} = \sqrt{\frac{2}{3}} \left ( G_{t}- G_{t-\Delta t} \right )
-	    \tag{B4}
-	  \end{equation}
-	  and of course $G_t = \sqrt{3/2} S_t$ and similary for $G_{t-\Delta t}$, using these in (B4)
-	  \begin{equation}
-	    2 \mu {\Delta e_{t-\Delta t/2}^e} = S_t - S_{t-\Delta t}
-	    \tag{B5}
-	  \end{equation}
-	  finally substituting the rate form of (41) and rearranging we have:
-	  \begin{equation}
-	    \Delta e_{t-\Delta t/2} - \frac{S_t - S_{t-\Delta t}}{2 \mu} - \Delta \lambda_{t-\Delta t/2} = 0
-	    \tag{B6}
-	  \end{equation}
-	  (B6) is exactly (44).  The rest of derivation follows the paper from (44).
-     \end{itemize}}
- %
-  \item Line 7 after (8).  $\mathbf{K}$  is a second order moment not a first order moment since $\boldsymbol{\xi} \otimes \boldsymbol{\xi}$ is a quadratic term.
-     {\color{red}
-     \begin{itemize}
-        \item The description of $\mathbf{K}$ as a moment has been removed in the revision.
-     \end{itemize}}
-     %
-  \item The value of $\delta$ used for the simulations should be specified.
-In this regard, it would be good to also state how many material points used in the simulation fit into this sphere of radius $\delta$.
-     {\color{red}
-     \begin{itemize}
-        \item Suggestion was included in revision along with additional discussion about how changing $\delta$ makes insignificant changes in the solution of the Taylor impact problem.
-     \end{itemize}}
-        %
-  \item Figure 2 refers to true strain but the measure $\mathbf{E}$ in (61) is the Lagranian strain.  Please provide a definition of the strain measure being used for true strain in Fig. 2. 
-     {\color{red}
-     \begin{itemize}
-        \item Suggestion was included in revision.
-     \end{itemize}}
-\end{enumerate}
-
 \end{document}

### 2014-02-08 jacobian_calc_CMAME.tex
response to I.1 started

@@
 \begin{abstract} %% Text of abstract Shown is a retrospective comparative study of tangent-stiffness
-In order to maintain the quadratic convergence properties of Newton's method in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
-A goal of the study described in this paper was to establish the suitability of an under-explored method for numerical computation of tangent-stiffness operators, referred to as ``complex-step'', and compare the new method with other techniques for numerical derivative calculation: automatic differentiation, forward finite-difference, and central finite-difference.
-The complex-step method was newly implemented in a massively parallel computational peridynamics code for the purpose of this comparison.
-The methods were compared through in situ profiling of the code for accuracy, speed, efficiency, and parallel scalability.
-The research provides data that can serve as practical guide for code developers and analysts faced with choosing which method best suits the needs of their application code.
-Additionally, motivated by the reproducible research movement, all the of the code, examples, and workflow to regenerate the data and figures in this paper are provided as open source.
+
+In order to maintain the quadratic convergence properties of Newton's method in
+quasi-static nonlinear analysis of solid structures it is crucial to obtain
+accurate, algorithmically consistent tangent-stiffness matrices. A goal of the
+study described in this paper was to establish the suitability of an
+under-explored method for numerical computation of tangent-stiffness operators,
+referred to as ``complex-step'', and compare the \note[Michael]{re: I.1}
+\remove[Michael]{new} method with other techniques for numerical derivative
+calculation: automatic differentiation, forward finite-difference, and central
+finite-difference. The complex-step method was \note[Michael]{re: I.1, could
+suggest claim of novelty} \remove[Michael]{newly} implemented in a massively
+parallel computational peridynamics code for the purpose of this comparison.
+The methods were compared through in situ profiling of the code for accuracy,
+speed, efficiency, and parallel scalability. The research provides data that
+can serve as practical guide for code developers and analysts faced with
+choosing which method best suits the needs of their application code.
+Additionally, motivated by the reproducible research movement, all the of the
+code, examples, and workflow to regenerate the data and figures in this paper
+are provided as open source.
 
@@
 \label{sec:intro}
-In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
-For an extremely small class of nonlinear material models, these consistent tangent-stiffness operators can be derived analytically; however, most often in practice, they are found through numerical approximation of derivatives.
 
-A goal of this study was to develop and evaluate a new, accurate, and practical method for calculating tangent-stiffness matrices against established methods.
-This new method is based on a complex number Taylor series expansion and referred to as CTSE or the ``complex-step'' method.
-The
-distinction of ``accurate'' is defined by comparison of the new method to popular finite differencing techniques used for computing tangent-stiffness matrices and with the exact to machine precision algorithmically consistent derivatives computed with \emph{automatic differentiation}.
+In order to maintain the quadratic convergence properties of the first-order
+Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in
+quasi-static nonlinear analysis of solid structures it is crucial to obtain
+accurate, algorithmically consistent tangent-stiffness matrices. For an
+extremely small class of nonlinear material models, these consistent
+tangent-stiffness operators can be derived analytically; however, most often in
+practice, they are found through numerical approximation of derivatives.
 
-Comparative data regarding the accuracy and compute cycle timing for the complex step, finite-difference, central-difference, and automatic differentiation methods are included to aide developers and analysts in computational mechanics code design who are faced with implementing a general method for tangent stiffness matrix calculation.
-A comparison and discrimination of the methods was achieved through in-situ instrumentation of a massively parallel computational mechanics code.
+A goal of this study was to develop and evaluate a \note[Michael]{re: I.1}
+\change[Michael]{new}{relatively underexplored}, accurate, and practical method
+for calculating tangent-stiffness matrices against established methods.  This
+new method is based on a complex number Taylor series expansion and referred to
+as CTSE or the ``complex-step'' method.  The distinction of ``accurate'' is
+defined by comparison of the \note[Michael]{re: I.1}\remove[Michael]{new} method to popular finite differencing
+techniques used for computing tangent-stiffness matrices and with the exact to
+machine precision algorithmically consistent derivatives computed with
+\emph{automatic differentiation}.
 
-The scope of the application component of the study was limited to a single material model, an elastic peridynamic solid, implemented in the computational peridynamics code, \textit{Peridigm} \cite{peridigm}.
-Identifying a specific application provided a practical framework for implementing the new method and solving engineering problems to generate the data needed to compare the methods.
-In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of finite-difference, central-difference, and automatic-differentiation methods needed for comparison to the new complex-step method implementation, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
-Agile-components is a software development term meaning to structure a software project for continuous development to respond to evolving specifications by making much use of techniques from object-oriented programming and reusing proven pre-existing code to reduce development time and increase reliability.
-The modularity of \emph{Peridigm} allowed the adding of new features, such as required by the study.
+Comparative data regarding the accuracy and compute cycle timing for the
+complex step, finite-difference, central-difference, and automatic
+differentiation methods are included to aide developers and analysts in
+computational mechanics code design who are faced with implementing a general
+method for tangent stiffness matrix calculation.  A comparison and
+discrimination of the methods was achieved through in-situ instrumentation of a
+massively parallel computational mechanics code.
 
-The aim of this paper is not only to introduce the new complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques useful for solving non-linear systems with Newton's method.
-After presenting background information on the underlying methods to be discussed, presented in this paper are: A) a description of tangent-stiffness matrices, B) detailed directions for producing tangent-stiffness matrices with each the methods identified above, C) a description and justification of the new complex-step method for calculating tangent-stiffness matrices, D) a description of implementing complex-step in the \emph{Peridigm} software, E) a description of the quantities-of-interest used to rank the methods, F) a presentation and analysis of the results of the comparative study, and G) conclusions and thoughts on potential future work.
-Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for \emph{Peridigm} C++ source-code and data necessary to reproduce the results presented here.
-Additionally, included on the website is an example serial C++ library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
-The purpose of the serial C++ library is to demonstrate in a less complex manner than in the \emph{Peridigm} code how one would use automatic-differentiation or complex-step to better encourage their use and discussion.
-This open source resource will be referred to as the \emph{paper repository} in what follows.
+The scope of the application component of the study was limited to a single
+material model, an elastic peridynamic solid, implemented in the computational
+peridynamics code, \textit{Peridigm} \cite{peridigm}. Identifying a specific
+application provided a practical framework for implementing the
+\note[Michael]{re: I.1}\remove[Michael]{new} method and solving engineering
+problems to generate the data needed to compare the methods. In particular,
+\emph{Peridigm} was chosen because it combined several helpful
+characteristics: the utilization of Newton's method and tangent-stiffness
+matrices for solving nonlinear quasi-static problems, prior inclusion of
+finite-difference, central-difference, and automatic-differentiation
+methods needed for comparison to the new complex-step method
+implementation, and being an agile-components code that makes use of
+distributed computing data structures via Trilinos \cite{trilinos} for
+efficient parallelization on large clusters. Agile-components is a software
+development term meaning to structure a software project for continuous
+development to respond to evolving specifications by making much use of
+techniques from object-oriented programming and reusing proven pre-existing
+code to reduce development time and increase reliability. The modularity of
+\emph{Peridigm} allowed the adding of new features, such as required by the
+study.
+
+The aim of this paper is not only to \note[Michael]{re: I.1}\change[Michael]{introduce the new}{study the} complex-step method \change[Michael]{in
+the context of}{of} evaluating tangent-stiffness matrices, but to serve as a review
+for other differentiation techniques useful for solving non-linear systems with
+Newton's method. After presenting background information on the underlying
+methods to be discussed, presented in this paper are: A) a description of
+tangent-stiffness matrices, B) detailed directions for producing
+tangent-stiffness matrices with each of the methods identified above, C) a
+description and justification of the \note[Michael]{re: I.1}\remove[Michael]{new}complex-step method for calculating
+tangent-stiffness matrices, D) a description of implementing complex-step in
+the \emph{Peridigm} software, E) a description of the quantities-of-interest
+used to rank the methods, F) a presentation and analysis of the results of the
+comparative study, and G) conclusions and thoughts on potential future work.
+Finally, in the interest of replicable research, the reader is referred to the
+corresponding author's website for \emph{Peridigm} C++ source-code and data
+necessary to reproduce the results presented here. Additionally, included on
+the website is an example serial C++ library with classes for solving
+non-linear systems using the techniques discussed here, as well as two
+validation and verification example problems which show how to use the library.
+The purpose of the serial C++ library is to demonstrate in a less complex
+manner than in the \emph{Peridigm} code how one would use
+automatic-differentiation or complex-step to better encourage their use and
+discussion. This open source resource will be referred to as the \emph{paper
+repository} in what follows.
 
@@
 
-\subsubsection{The ``complex-step'' method.}  
-\label{sec:CSmethod}
-It is possible to approximate derivatives quite accurately with a technique based on a complex variable Taylor series expansion of a function.
-This method was first described by Lyness and Moler \cite{lyness1967numerical,lyness1968differentiation} and has more recently been rediscovered \cite{squire1998using,voorhees2011complex,al2010complex,martins2000automated} for use in engineering analysis.
-The basic idea is that a model parameter can be be made complex and expanded in a Taylor series about a small perturbation along the imaginary axis by some arbitrary value $h$ as follows
+\subsubsection{The ``complex-step'' method.}  \label{sec:CSmethod} It is
+possible to approximate derivatives quite accurately with a technique based on
+a complex variable Taylor series expansion of a function.  This method was
+first described by Lyness and Moler
+\cite{lyness1967numerical,lyness1968differentiation} and has more recently been
+rediscovered
+\cite{squire1998using,voorhees2011complex,al2010complex,martins2000automated}
+for use in engineering analysis.  The basic idea is that a model parameter can
+be be made complex and expanded in a Taylor series about a small perturbation
+along the imaginary axis by some arbitrary value $h$ as follows
 %
-\begin{align}
-f ( x + i h ) =& f(x) + \frac{\partial f(x)}{\partial x} \frac{i h}{1!}  \notag \\
-               &+ \frac{\partial f^2(x)}{\partial x^2} \frac{(i h)^2}{2!} + \frac{\partial f^3(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots,
-\label{eqn:complexTaylor}
-\end{align}
+\begin{align} f ( x + i h ) =& f(x) + \frac{\partial f(x)}{\partial x} \frac{i
+h}{1!}  \notag \\ &+ \frac{\partial f^2(x)}{\partial x^2} \frac{(i h)^2}{2!} +
+\frac{\partial f^3(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots,
+\label{eqn:complexTaylor} \end{align}
 %
-Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor}) and solving for the first derivative and ignoring terms of $\mathcal{O}\left ( h^2 \right)$ yields an estimate
+Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor})
+and solving for the first derivative and ignoring terms of $\mathcal{O}\left (
+h^2 \right)$ yields an estimate
 %
-\begin{equation}
- \frac{\partial f( x )}{\partial x} \approx \frac{\mbox{Im} \left( f (x + i h) \right)}{h}.
-\label{eqn:complexFirstDeriv}
-\end{equation}
+\begin{equation} \frac{\partial f( x )}{\partial x} \approx \frac{\mbox{Im}
+\left( f (x + i h) \right)}{h}.  \label{eqn:complexFirstDeriv} \end{equation}
 %
-Thus, an estimate of the first derivative of a function can be made by only utilizing one functional evaluation of a perturbed model parameter along the imaginary axis.
-The step-size $h$ is arbitrary and can be made as small as practical (even to machine precision without the dangers of roundoff error as in finite-differences) to yield an accurate estimate of this derivative.
-The only disadvantage of this technique is the necessity of requiring the functions to accept complex numbers as arguments.
-Appendix~\ref{sec:appendixA} includes a simple example that illustrates the use of the complex-step method for the derivative calculation of a function.
+Thus, an estimate of the first derivative of a function can be made by only
+utilizing one functional evaluation of a perturbed model parameter along the
+imaginary axis.  The step-size $h$ is arbitrary and can be made as small as
+practical (even to machine precision without the dangers of roundoff error as
+in finite-differences) to yield an accurate estimate of this derivative.  The
+only disadvantage of this technique is the necessity of requiring the functions
+to accept complex numbers as arguments. Appendix~\ref{sec:appendixA} includes a
+simple example that illustrates the use of the complex-step method for the
+derivative calculation of a function.
 
@@
 
-Automatic-differentiation (AD) is a computerized method for computing exact derivatives based the chain-rule from calculus.
-AD takes advantage of the fact that any mathematical function executed on a computer, no matter how complicated, is a ``composition of simple operations'' (add, multiply, power, transcendental and the like) each having known analytical derivatives \cite{ref-sacado-presentation}.
-For reference, the AD implementation used in the study is part of the \emph{Sacado} package from the \emph{Trilinos} agile-component libraries developed at Sandia National Laboratories \cite{ref-Sacado}.
+Automatic-differentiation (AD) is a computerized method for computing exact
+derivatives based the chain-rule from calculus. AD takes advantage of the fact
+that any mathematical function executed on a computer, no matter how
+complicated, is a ``composition of simple operations'' (add, multiply, power,
+transcendental and the like) each having known analytical derivatives
+\cite{ref-sacado-presentation}. For reference, the AD implementation used in
+the study is part of the \emph{Sacado} package from the \emph{Trilinos}
+agile-component libraries developed at Sandia National Laboratories
+\cite{ref-Sacado}.
 
-An AD system evaluates composition functions in the expected manner: it works by first evaluating the innermost function of the composition, then presenting that function's output as input to the next level function until all levels are complete, observing commonly expected order of operations.
-However, an AD system does additional work during a function evaluation in that as each nested function is evaluated, the function's partial derivative with respect to the designated variables of the given input is also calculated.
-This is possible because the elementary math functions are hard-coded into the AD source-code along with their analytical derivatives, and linked by special instructions, so that when the elementary math functions are called upon for computation, their partial derivatives may be computed and stored in a sequence.
-The AD system then multiplies the final sequence of partial derivatives together to produce the exact equivalent to taking a partial derivative of the corresponding composition function with respect to a designated variable at a particular value.
-It is obvious, but bears mentioning, that the AD system could simply store one value for partial derivatives, modifying it as appropriate for every function evaluation rather than keeping a sequence.
-This is significant because modifying a single value rather than keeping a list of values corresponding to
-every level of a composition function, which itself may not be the sole one needing to be evaluated, represents a savings in memory usage, which at large scales is an active concern.
-In the literature, the AD scheme described here is called \emph{forward automatic-differentiation}.
-Only forward AD will be covered here since it is the implementation used in \emph{Sacado} and therefore in this study; however, the reader is referred to the introduction section of \cite{ref-AD-methods} and its citations for further information on AD, and particularly \cite{ref-on-AD} which is foundational.
-Appendix~\ref{sec:appendixB} includes an example that walks the reader through the process a computer uses to compute derivatives via AD.
-Some things to note about AD are that no approximation of derivatives is being made because the analytical forms of the partials of the elementary math functions are defined alongside them.
-The accuracy of AD is then limited by the precision of the AD system's definition of the elementary math functions and their partial derivatives.
+An AD system evaluates composition functions in the expected manner: it works
+by first evaluating the innermost function of the composition, then presenting
+that function's output as input to the next level function until all levels are
+complete, observing commonly expected order of operations. However, an AD
+system does additional work during a function evaluation in that as each nested
+function is evaluated, the function's partial derivative with respect to the
+designated variables of the given input is also calculated.  This is possible
+because the elementary math functions are hard-coded into the AD source-code
+along with their analytical derivatives, and linked by special instructions, so
+that when the elementary math functions are called upon for computation, their
+partial derivatives may be computed and stored in a sequence. The AD system
+then multiplies the final sequence of partial derivatives together to produce
+the exact equivalent to taking a partial derivative of the corresponding
+composition function with respect to a designated variable at a particular
+value. It is obvious, but bears mentioning, that the AD system could simply
+store one value for partial derivatives, modifying it as appropriate for every
+function evaluation rather than keeping a sequence. This is significant because
+modifying a single value rather than keeping a list of values corresponding to
+every level of a composition function, which itself may not be the sole one
+needing to be evaluated, represents a savings in memory usage, which at large
+scales is an active concern. In the literature, the AD scheme described here is
+called \emph{forward automatic-differentiation}. Only forward AD will be
+covered here since it is the implementation used in \emph{Sacado} and therefore
+in this study; however, the reader is referred to the introduction section of
+\cite{ref-AD-methods} and its citations for further information on AD, and
+particularly \cite{ref-on-AD} which is foundational.
+Appendix~\ref{sec:appendixB} includes an example that walks the reader through
+the process a computer uses to compute derivatives via AD.  Some things to note
+about AD are that no approximation of derivatives is being made because the
+analytical forms of the partials of the elementary math functions are defined
+alongside them. The accuracy of AD is then limited by the precision of the AD
+system's definition of the elementary math functions and their partial
+derivatives.
 
@@
 
-In solid and structural computational mechanics the tangent-stiffness matrix is a linearization operator that describes the stiffness of a system in response to small displacements imposed upon the current configuration of the system.
-Mathematically speaking, the tangent-stiffness represents the gradient of a high-dimensional energy surface that ``points'' in the minimum direction.
-While the terminology of solid mechanics is used here, it's important to note that these operators appear in other physical settings and are known by other names, e.g.
-a \emph{transmissibility matrix} in the context of a Poisson problem or, more generally, a \emph{Jacobian matrix} in mathematical optimization.
+In solid and structural computational mechanics the tangent-stiffness matrix is
+a linearization operator that describes the stiffness of a system in response
+to small displacements imposed upon the current configuration of the system.
+Mathematically speaking, the tangent-stiffness represents the gradient of a
+high-dimensional energy surface that ``points'' in the minimum direction. While
+the terminology of solid mechanics is used here, it's important to note that
+these operators appear in other physical settings and are known by other names,
+e.g. a \emph{transmissibility matrix} in the context of a Poisson problem or,
+more generally, a \emph{Jacobian matrix} in mathematical optimization.
 
-While, for simplicity, the examples shown in the following refer only to linear problems, tangent-stiffness matrices generally arise in the context of non-linear analysis where Newton's method or quasi-Newton's method's are used in the minimization of a given residual function.
-In this context, the tangent-stiffness matrix can be thought of as the linearization of the system about about a particular configuration.
-It has been shown \cite{hughes1978consistent,hughes1978unconditionally} that in order to preserve the quadratic convergence properties of the Newton's methods, it is necessary that this linearization is carried out in a manner that is \emph{consistent} with the algorithmic constraint equations used when computing the internal forces that arise due to deformations.
-An example of these constraint equations would be the Kuhn-Tucker conditions \cite{simo1998} that are used in integration of a flow rule for plasticity modeling.
-If the continuum tangent-moduli are used in place of the \emph{consistent} or \emph{algorithmic tangent moduli} in the solution of a non-linear solid mechanics problem, convergence may still be achieved, but not in the quadratic manner that makes Newton's method attractive for this class of problems.
-In the setting of non-linear solid mechanics, there is a very small class of material model algorithms in which consistent tangent-stiffness operators can be derived analytically.
-Perfect plasticity and plasticity with isotropic hardening used in combination with a general nearest-point projection or \emph{radial return algorithm} are a few examples.
-Therefore, when general models (and certainly more complex ones) are implemented into a general purpose computational mechanics code, the tangent-stiffness operators are typically defined via a numerical approximation.
+While, for simplicity, the examples shown in the following refer only to linear
+problems, tangent-stiffness matrices generally arise in the context of
+non-linear analysis where Newton's method or quasi-Newton's method's are used
+in the minimization of a given residual function.  In this context, the
+tangent-stiffness matrix can be thought of as the linearization of the system
+about about a particular configuration.  It has been shown
+\cite{hughes1978consistent,hughes1978unconditionally} that in order to preserve
+the quadratic convergence properties of the Newton's methods, it is necessary
+that this linearization is carried out in a manner that is \emph{consistent}
+with the algorithmic constraint equations used when computing the internal
+forces that arise due to deformations.  An example of these constraint
+equations would be the Kuhn-Tucker conditions \cite{simo1998} that are used in
+integration of a flow rule for plasticity modeling.  If the continuum
+tangent-moduli are used in place of the \emph{consistent} or \emph{algorithmic
+tangent moduli} in the solution of a non-linear solid mechanics problem,
+convergence may still be achieved, but not in the quadratic manner that makes
+Newton's method attractive for this class of problems.  In the setting of
+non-linear solid mechanics, there is a very small class of material model
+algorithms in which consistent tangent-stiffness operators can be derived
+analytically.  Perfect plasticity and plasticity with isotropic hardening used
+in combination with a general nearest-point projection or \emph{radial return
+algorithm} are a few examples.  Therefore, when general models (and certainly
+more complex ones) are implemented into a general purpose computational
+mechanics code, the tangent-stiffness operators are typically defined via a
+numerical approximation.
 
-A general mathematical formula for a tangent-stiffness matrix, $K_{ij}$, can be expressed in indicial notation as 
+A general mathematical formula for a tangent-stiffness matrix, $K_{ij}$, can be
+expressed in indicial notation as 
 %
-\begin{equation} 
-  K_{ij} = \left. \frac{\partial F_i}{\partial X_j}\right|_{\vec{X}_0},
-\end{equation}
+\begin{equation} K_{ij} = \left. \frac{\partial F_i}{\partial
+X_j}\right|_{\vec{X}_0}, \end{equation}
 %
-where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a particular value of the vector, $\vec{X}_0$ is chosen as the linearization point.
-One then evaluates the expression for each combination $(i, j)$ corresponding to a (row, column) location in the tangent-stiffness matrix.
-The elements of a tangent-stiffness matrix can be estimated using any of the complex-step, AD, or common finite-difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the function constant, and each element of the function can be evaluated independently of the others.
-\\
+where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function,
+$X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a
+particular value of the vector, $\vec{X}_0$ is chosen as the linearization
+point. One then evaluates the expression for each combination $(i, j)$
+corresponding to a (row, column) location in the tangent-stiffness matrix. The
+elements of a tangent-stiffness matrix can be estimated using any of the
+complex-step, AD, or common finite-difference techniques for functions $F:R^1
+\rightarrow R^1$, since taking partial derivatives entails holding all but a
+single independent vector component of the function constant, and each element
+of the function can be evaluated independently of the others. \\
 
@@
 %
-\begin{equation} 
-  K_{ij} = \frac{\mbox{Im}(F_i^{int}(\vec{u} + i h \hat{e}_j))}{h},
+\begin{equation} K_{ij} = \frac{\mbox{Im}(F_i^{int}(\vec{u} + i h
+\hat{e}_j))}{h}, \end{equation}
+%
+where the internal force, $F^{int}$ is now treated as function of complex
+vectors,  $\vec{u}$. The small perturbation, $h$, is now carried out on the
+imaginary axis and only the real coefficient to the imaginary part of the
+vector returned by $F^{int}$ is kept, hence the $\mbox{Im}(\cdot)$ operation.
+As highlighted in Section~\ref{sec:CSmethod}, notice that there is no
+subtraction operation taking place; therefore $h$ can be made very small to
+yield highly accurate derivatives. It should be mentioned that using
+complex-step to compute tangent-stiffness matrices was done in
+\cite{perez2000numerical,perez2012numerical}, but those references did not
+investigate the performance of complex-step in a massively parallel setting nor
+make comparisons to AD.
+
+When using automatic-differentiation to calculate a tangent-stiffness matrix,
+one only needs to follow the definition of the tangent-stiffness matrix and
+issue the correct commands to the AD system. The formula is the same as the
+continuum formula
+%
+\begin{equation} K_{ij} = \frac{\partial F_i^{int}(\vec{u})}{\partial u_j},
 \end{equation}
 %
-where the internal force, $F^{int}$ is now treated as function of complex vectors,  $\vec{u}$.
-The small perturbation, $h$, is now carried out on the imaginary axis and only the real coefficient to the imaginary part of the vector returned by $F^{int}$ is kept, hence the $\mbox{Im}(\cdot)$ operation.
-As highlighted in Section~\ref{sec:CSmethod}, notice that there is no subtraction operation taking place; therefore $h$ can be made very small to yield highly accurate derivatives.
-It should be mentioned that using complex-step to compute tangent-stiffness matrices was done in \cite{perez2000numerical,perez2012numerical}, but those references did not investigate the performance of complex-step in a massively parallel setting nor make comparisons to AD.
-
-When using automatic-differentiation to calculate a tangent-stiffness matrix, one only needs to follow the definition of the tangent-stiffness matrix and issue the correct commands to the AD system.
-The formula is the same as the continuum formula
-%
-\begin{equation} 
-  K_{ij} = \frac{\partial F_i^{int}(\vec{u})}{\partial u_j},
-\end{equation}
-%
-but, with the caveat that $F^{int}$ is evaluated in a way that is algorithmically consistent, and therefore should yield quadratic convergence with Newton's method.
+but, with the caveat that $F^{int}$ is evaluated in a way that is
+algorithmically consistent, and therefore should yield quadratic convergence
+with Newton's method.
 
@@
 %
-Comparisons of the different tangent-stiffness calculation techniques were carried out through code-development and \emph{in situ} instrumentation of the computational peridynamics code \emph{Peridigm} \cite{peridigm}.
-\emph{Peridigm} is distributed by Sandia National Laboratories as its primary open-source computational peridynamics code.
-It is a massively parallel simulation code for implicit and explicit multi-physics simulations primarily used for solid mechanics and material failure.
-\emph{Peridigm} is a C++ code utilizing agile software components from Sandia's  \emph{Trilinos} project \cite{trilinos}.
-It's important to note that there is no specialization of the tangent-stiffness calculation methods described previously to this particular code or more generally to a computational peridynamics approach and the analysis carried out in this study should provide insight into the speed and accuracy of these methods when implemented into other computational mechanics analysis tools as well.
-\emph{Peridigm} was chosen primarily because the FD, CD, and AD methods for tangent-stiffness calculation where already implemented in the code, leaving only the CS method for development.
-Also, because peridynamics is a nonlocal theory, the tangent-stiffness matrices have a much higher bandwidth (i.e.
-less sparsity) than traditional finite-element tangent-stiffness computations and the time required to construct the tangent-stiffness is a majority of the total computation time in any quasi-static or implicit dynamics analysis; therefore, the \emph{Peridigm} development team has an interest in profiling the performance of the tangent-stiffness computation methods.
+Comparisons of the different tangent-stiffness calculation techniques were
+carried out through code-development and \emph{in situ} instrumentation of the
+computational peridynamics code \emph{Peridigm} \cite{peridigm}.
+\emph{Peridigm} is distributed by Sandia National Laboratories as its primary
+open-source computational peridynamics code. It is a massively parallel
+simulation code for implicit and explicit multi-physics simulations primarily
+used for solid mechanics and material failure. \emph{Peridigm} is a C++ code
+utilizing agile software components from Sandia's  \emph{Trilinos} project
+\cite{trilinos}. It's important to note that there is no specialization of the
+tangent-stiffness calculation methods described previously to this particular
+code or more generally to a computational peridynamics approach and the
+analysis carried out in this study should provide insight into the speed and
+accuracy of these methods when implemented into other computational mechanics
+analysis tools as well.  \emph{Peridigm} was chosen primarily because the FD,
+CD, and AD methods for tangent-stiffness calculation where already implemented
+in the code, leaving only the CS method for development. Also, because
+peridynamics is a nonlocal theory, the tangent-stiffness matrices have a much
+higher bandwidth (i.e. less sparsity) than traditional finite-element
+tangent-stiffness computations and the time required to construct the
+tangent-stiffness is a majority of the total computation time in any
+quasi-static or implicit dynamics analysis; therefore, the \emph{Peridigm}
+development team has an interest in profiling the performance of the
+tangent-stiffness computation methods.
 
-Briefly, peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic} is a nonlocal reformulation of the partial-differential equations that provide the statement of momentum balance in classical continuum mechanics.
-Its primary goal is to avoid the use of spatial derivatives in the balance equations or constitutive laws such that discontinuous displacements, i.e.
-cracks, are mathematically consistent with the governing equations.
-It has demonstrated great promise in modeling problems with pervasive failure \cite{littlewood2010}, interesting phenomenon such as dynamic crack branching \cite{ha2010sod}, and is being extended to multiphyics phenomena \cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale coupling to molecular simulations \cite{seleson2009peridynamics}.
+Briefly, peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
+is a nonlocal reformulation of the partial-differential equations that provide
+the statement of momentum balance in classical continuum mechanics. Its primary
+goal is to avoid the use of spatial derivatives in the balance equations or
+constitutive laws such that discontinuous displacements, i.e. cracks, are
+mathematically consistent with the governing equations. It has demonstrated
+great promise in modeling problems with pervasive failure
+\cite{littlewood2010}, interesting phenomenon such as dynamic crack branching
+\cite{ha2010sod}, and is being extended to multiphyics phenomena
+\cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale
+coupling to molecular simulations \cite{seleson2009peridynamics}.
 
@@
 %
-The CS method was implemented in \emph{Perdigm} in a manner that follows closely the implementation of the FD technique.
-However, because of \emph{Peridigm}'s reliance on \emph{Trilinos} and specifically the \emph{Epetra} package, special steps had to be followed to allow the use of complex number data types.
-This is because \emph{Epetra} vectors, which can be thought of as distributed memory parallel cousins to standard C++ STL vectors, are hard coded to be of type {\tt double} and can not be simply be declared as having a complex data type.
-Therefore, program methods that were used in the course of applying the FD method were copied, renamed and re-written to follow the CS formula.
-This involved changing the data type of intermediate variables which were locally scoped, and dynamically casting persistent memory variables to complex data types where necessary.
-No attempt was made to ``tune'' the CS code in any way for performance.
-The algorithm  was implemented in such as way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurts its overall performance and this issue could be alleviated by using the next generation of templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any data type including {\tt complex}.
-The version of \emph{Peridigm} modified to implement CS can be found in the paper repository.
+The CS method was implemented in \emph{Perdigm} in a manner that follows
+closely the implementation of the FD technique.  However, because of
+\emph{Peridigm}'s reliance on \emph{Trilinos} and specifically the
+\emph{Epetra} package, special steps had to be followed to allow the use of
+complex number data types.  This is because \emph{Epetra} vectors, which can be
+thought of as distributed memory parallel cousins to standard C++ STL vectors,
+are hard coded to be of type {\tt double} and can not be simply be declared as
+having a complex data type. Therefore, program methods that were used in the
+course of applying the FD method were copied, renamed and re-written to follow
+the CS formula. This involved changing the data type of intermediate variables
+which were locally scoped, and dynamically casting persistent memory variables
+to complex data types where necessary. No attempt was made to ``tune'' the CS
+code in any way for performance. The algorithm  was implemented in such as way
+that it was in one-to-one correspondence with the FD and CD methods; however,
+it must be noted that the overhead associated with the dynamic casts in the CS
+method likely hurts its overall performance and this issue could be alleviated
+by using the next generation of templated \emph{TPetra} vectors available in
+\emph{Trilinos} that are capable of taking an explicit instantiation of any
+data type including {\tt complex}. The version of \emph{Peridigm} modified to
+implement CS can be found in the paper repository.
 
-\emph{Peridigm} was primarily written by computational scientists; therefore, many advanced techniques in C++ such as multiple virtual inheritance, pointer arithmetic, and distributed data structures were used in the basic code and by necessity in the modified code.
-To see CS and other methods in a more narrow context, it is advised that readers also take a look at the simple examples provided on the corresponding author's website.
+\emph{Peridigm} was primarily written by computational scientists; therefore,
+many advanced techniques in C++ such as multiple virtual inheritance, pointer
+arithmetic, and distributed data structures were used in the basic code and by
+necessity in the modified code.  To see CS and other methods in a more narrow
+context, it is advised that readers also take a look at the simple examples
+provided on the corresponding author's website.
 
@@
 
-The methods were compared by running two sets of test problems where each method would solve a problem concurrently.
-In this context, concurrently means successively, drawing from the same independent variables, yet within the same computational process; this definition is characterized in further detail in Section~\ref{JGAM}.
-Each test problem simulated a $\SI{4}{\meter}$ by $\SI{0.5}{\meter}$ by $\SI{0.5}{\meter}$ meter block of a material undergoing tension along the axis parallel to the long dimension of the block.
-The test problems were set up in \emph{Perdigm} as quasistatic equilibrium problems, where the displacement boundary conditions used to apply tension were applied gradually in ``load steps'', and an equilibrium solution for each load step was achieved before applying the next load step.
-The purpose of applying the load in gradual steps is a widely used technique to ensure the Newton's method is always initialized near the actual solution and therefore likely to converge.
-The fictional material represented in the model had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus.
-The peridynamic horizon, a length scale that effectively sets the bandwidth of tangent-stiffness in the computation, used in the test problems was always set to three times the nominal discretization size (i.e.
-node spacing in the particle discretization scheme).
-This results in a minimum tangent-stiffness matrix bandwidth of 7 (for an optimal node numbering scheme).
+The methods were compared by running two sets of test problems where each
+method would solve a problem concurrently. In this context, concurrently means
+successively, drawing from the same independent variables, yet within the same
+computational process; this definition is characterized in further detail in
+Section~\ref{JGAM}. Each test problem simulated a $\SI{4}{\meter}$ by
+$\SI{0.5}{\meter}$ by $\SI{0.5}{\meter}$ meter block of a material undergoing
+tension along the axis parallel to the long dimension of the block. The test
+problems were set up in \emph{Perdigm} as quasistatic equilibrium problems,
+where the displacement boundary conditions used to apply tension were applied
+gradually in ``load steps'', and an equilibrium solution for each load step was
+achieved before applying the next load step. The purpose of applying the load
+in gradual steps is a widely used technique to ensure the Newton's method is
+always initialized near the actual solution and therefore likely to converge.
+The fictional material represented in the model had values of
+$\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$
+for shear modulus. The peridynamic horizon, a length scale that effectively
+sets the bandwidth of tangent-stiffness in the computation, used in the test
+problems was always set to three times the nominal discretization size (i.e.
+node spacing in the particle discretization scheme). This results in a minimum
+tangent-stiffness matrix bandwidth of 7 (for an optimal node numbering scheme).
 
-A series of nine single compute core (i.e.
-serial) test problems were run with the discretization size being refined with every test in the series.
-The aim of increasing the refinement was to examine differences between the methods in terms of accuracy and speed.
-This series would comprise the single core runs.
-The specific parameters of these tests can be found in \emph{Peridigm} input {\tt xml} files included in the paper repository.
-These {\tt xml} files allow users with the appropriately modified version of \emph{Peridigm} to reproduce the results found in this paper.
+A series of nine single compute core (i.e. serial) test problems were run with
+the discretization size being refined with every test in the series. The aim of
+increasing the refinement was to examine differences between the methods in
+terms of accuracy and speed. This series would comprise the single core runs.
+The specific parameters of these tests can be found in \emph{Peridigm} input
+{\tt xml} files included in the paper repository. These {\tt xml} files allow
+users with the appropriately modified version of \emph{Peridigm} to reproduce
+the results found in this paper.
 
-Another series of four test problems were simulated, however this time the number of compute cores used to solve the problem was increased from test to test while the discretization level was held constant at 1 million computational nodes (3 million degrees of freedom).
-This allowed for sufficient computational nodes per core even at the highest level of parallelization such that message passing computation did not overwhelm the simulation.
-The specific parameters of these tests can be found in \emph{Perdigm} input {\tt xml} files also included in the paper repository.
+Another series of four test problems were simulated, however this time the
+number of compute cores used to solve the problem was increased from test to
+test while the discretization level was held constant at 1 million
+computational nodes (3 million degrees of freedom). This allowed for sufficient
+computational nodes per core even at the highest level of parallelization such
+that message passing computation did not overwhelm the simulation.   The
+specific parameters of these tests can be found in \emph{Perdigm} input {\tt
+xml} files also included in the paper repository.
 
-\subsubsection{Quantities of interest} 
-\label{JGAM} 
-As stated in the Introduction, a goal of the study was to compare CS, FD, and CD on the basis of accuracy.
-AD was omitted from the comparison because it served as the standard of accuracy for the other methods in the absence of appropriate analytical forms for the tangent-stiffness matrix associated with the system solved in the study.
-The assumption that AD is accurate enough to serve as a standard is supported by ADs implementation as a computerized chain rule as explained in Section~\ref{ADsubsection}.
+\subsubsection{Quantities of interest} \label{JGAM}
 
-It would be a poor comparative study to compare tangent-stiffness matrices from different problems, load steps that start with different current configurations, or from different iterations; because of this, it was necessary to solve one
-load step and conclude each Newton iteration within that load step by updating the displacement iterate with only the results of the AD method and to subsequently feed all four methods the same updated displacement in the following iteration.
-This decision precluded a comparison of Newton iteration convergence rate, since if the methods were allowed to solve a problem at their individual pace, differences in accuracy would produce differences in guess updates and therefore the number and nature of Newton iterations performed (e.g.
-lower accuracy predictions of the algorithmically correct tangent-stiffness could cause a loss of quadratic convergence).
-Different guesses from iterations started with different previous guesses could not be data for a valid comparison of tangent-stiffness accuracy between methods.
-Additionally, having each of the methods physically operate in the same process, serial or one of associated parallel, allowed the comparison of tangent-stiffness matrix calculation time as cached from random access memory rather than from the hard-disk.
-Caching from RAM gives the benefit of vastly greater speed and simpler programming compared to some other solution speculatively involving dynamic file management on files which for one of the components of the study would be on the order of a terabyte in size.
-However, the price of running the methods together in the same process and avoiding the use of the hard-disk was that at least two tangent-stiffness matrices had to be stored in RAM during the simulation, which meant that special high-memory compute nodes were needed for the 1 million peridynamic node test series.
+As stated in the Introduction, a goal of the study was to compare CS, FD, and
+CD on the basis of accuracy. AD was omitted from the comparison because it
+served as the standard of accuracy for the other methods in the absence of
+appropriate analytical forms for the tangent-stiffness matrix associated with
+the system solved in the study. The assumption that AD is accurate enough to
+serve as a standard is supported by ADs implementation as a computerized chain
+rule as explained in Section~\ref{ADsubsection}.
 
-The other goal was to compare the four methods, including AD, on the basis of speed.
-Speed is defined as the total compute-time of one iteration of a Newton step which includes the tangent-stiffness calculation and parallel assembly.
-The speed of iteration was measured because it could be so done at the same time as accuracy was being measured given a single tangent-stiffness calculation.
-It was assumed that the order that each method was evaluated in parallel was unimportant, that evaluating each method successively within each solver iteration did not affect their performance individually and that speed of computation did not change with time.
-These assumptions allowed the test program to run the same problem with each of the methods at effectively the same time and generate an equal volume of data from each method.
-It was also assumed that the tangent-stiffness matrix calculation time for each of the methods did not vary based on the current configuration, as it changes slightly from iteration to iteration as the Newton solution iterate is updated.
-This assumption allowed calculation time measurements to be averaged over all iterations within a single load step.
-The purpose of averaging calculation time measurements was to informally address the extraneous variables of parallel evaluation order and computer system load due to system processes not associated with the study.
-While not part of the study, the example programs available in the paper repository allow the reader to make a comparison of the methods for themselves on the basis of accuracy and convergence rate for two example nonlinear systems.
-In these examples, one will notice that the finite difference methods will lose quadratic convergence when the step size selected is too large such that accuracy relative to AD is decreased.
+It would be a poor comparative study to compare tangent-stiffness matrices from
+different problems, load steps that start with different current
+configurations, or from different iterations; because of this, it was necessary
+to solve one load step and conclude each Newton iteration within that load step
+by updating the displacement iterate with only the results of the AD method and
+to subsequently feed all four methods the same updated displacement in the
+following iteration. This decision precluded a comparison of Newton iteration
+convergence rate, since if the methods were allowed to solve a problem at their
+individual pace, differences in accuracy would produce differences in guess
+updates and therefore the number and nature of Newton iterations performed
+(e.g. lower accuracy predictions of the algorithmically correct
+tangent-stiffness could cause a loss of quadratic convergence).  Different
+guesses from iterations started with different previous guesses could not be
+data for a valid comparison of tangent-stiffness accuracy between methods.
+Additionally, having each of the methods physically operate in the same
+process, serial or one of associated parallel, allowed the comparison of
+tangent-stiffness matrix calculation time as cached from random access memory
+rather than from the hard-disk.  Caching from RAM gives the benefit of vastly
+greater speed and simpler programming compared to some other solution
+speculatively involving dynamic file management on files which for one of the
+components of the study would be on the order of a terabyte in size. However,
+the price of running the methods together in the same process and avoiding the
+use of the hard-disk was that at least two tangent-stiffness matrices had to be
+stored in RAM during the simulation, which meant that special high-memory
+compute nodes were needed for the 1 million peridynamic node test series.
 
-The goals of collecting accuracy and speed measurements were achieved by developing and implementing metrics within the simulation program used in the study.
-The metric used to measure the accuracy of a tangent-stiffness matrix was the Frobenius norm (analogous to an $l^2$-norm except defined on a matrix) of the element-wise difference between the tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix produced by the AD based method (i.e., the \emph{exact} derivatives), given that both methods were set upon the same problem.
-The lower the value of this metric, the more accurate the method.
-The expression for the accuracy metric was
+The other goal was to compare the four methods, including AD, on the basis of
+speed. Speed is defined as the total compute-time of one iteration of a Newton
+step which includes the tangent-stiffness calculation and parallel assembly.
+The speed of iteration was measured because it could be so done at the same
+time as accuracy was being measured given a single tangent-stiffness
+calculation.  It was assumed that the order that each method was evaluated in
+parallel was unimportant, that evaluating each method successively within each
+solver iteration did not affect their performance individually and that speed
+of computation did not change with time. These assumptions allowed the test
+program to run the same problem with each of the methods at effectively the
+same time and generate an equal volume of data from each method. It was also
+assumed that the tangent-stiffness matrix calculation time for each of the
+methods did not vary based on the current configuration, as it changes slightly
+from iteration to iteration as the Newton solution iterate is updated.  This
+assumption allowed calculation time measurements to be averaged over all
+iterations within a single load step. The purpose of averaging calculation time
+measurements was to informally address the extraneous variables of parallel
+evaluation order and computer system load due to system processes not
+associated with the study. While not part of the study, the example programs
+available in the paper repository allow the reader to make a comparison of the
+methods for themselves on the basis of accuracy and convergence rate for two
+example nonlinear systems. In these examples, one will notice that the finite
+difference methods will lose quadratic convergence when the step size selected
+is too large such that accuracy relative to AD is decreased.
+
+The goals of collecting accuracy and speed measurements were achieved by
+developing and implementing metrics within the simulation program used in the
+study. The metric used to measure the accuracy of a tangent-stiffness matrix
+was the Frobenius norm (analogous to an $l^2$-norm except defined on a matrix)
+of the element-wise difference between the tangent-stiffness matrix produced by
+the method being evaluated and the tangent-stiffness matrix produced by the AD
+based method (i.e., the \emph{exact} derivatives), given that both methods were
+set upon the same problem. The lower the value of this metric, the more
+accurate the method. The expression for the accuracy metric was
 %
-\begin{equation} 
-  D = \sqrt{\sum_i \sum_j(K^{AD}_{ij} - K^{M}_{ij})^2}
-  \label{eqn:accuracy}
-\end{equation}
+\begin{equation} D = \sqrt{\sum_i \sum_j(K^{AD}_{ij} - K^{M}_{ij})^2}
+\label{eqn:accuracy} \end{equation}
 %
-Where $D$ is distance, $K^{AD}$ is the tangent-stiffness matrix produced by the AD based method, $M$ is replaced with either CS, FD, or CD depending on the method being compared.
-The metric used to compare the speed of the different methods was the time in seconds required to calculate the tangent-stiffness matrix.
-A final basis of comparison used in the study called \emph{computational efficiency} is based on specific calculation time per tangent-stiffness matrix element.
+Where $D$ is distance, $K^{AD}$ is the tangent-stiffness matrix produced by the
+AD based method, $M$ is replaced with either CS, FD, or CD depending on the
+method being compared.  The metric used to compare the speed of the different
+methods was the time in seconds required to calculate the tangent-stiffness
+matrix. A final basis of comparison used in the study called
+\emph{computational efficiency} is based on specific calculation time per
+tangent-stiffness matrix element.
 
@@
 %
-Discretization for the test problems run in the study was done using the software package \emph{Cubit} developed at Sandia National Laboratories \cite{ref-Cubit}.
-\emph{Cubit} generates a finite element mesh that is internally converted to a ``particle'' discretization internally when calling \emph{Peridigm}.
-The number of computational nodes in \emph{Peridigm} correspond to the number of finite elements in the discretization, not the number of finite element nodes.
-An example journal script can be found on the corresponding author's website.
+Discretization for the test problems run in the study was done using the
+software package \emph{Cubit} developed at Sandia National Laboratories
+\cite{ref-Cubit}. \emph{Cubit} generates a finite element mesh that is
+internally converted to a ``particle'' discretization internally when calling
+\emph{Peridigm}. The number of computational nodes in \emph{Peridigm}
+correspond to the number of finite elements in the discretization, not the
+number of finite element nodes.  An example journal script can be found on the
+corresponding author's website.
 
-The test problems were run on the \emph{Stampede} HPC cluster computer housed at \emph{TACC} (Texas Advanced Computing Center) at The University of Texas at Austin.
-The single core test series was run using the {\tt normal} queue compute nodes, while the parallel test series was run using one to four {\tt large memory} nodes each having one terabyte of local RAM.
-This large amount of memory was required for reasons explained in Section~\ref{JGAM}.
+The test problems were run on the \emph{Stampede} HPC cluster computer housed
+at \emph{TACC} (Texas Advanced Computing Center) at The University of Texas at
+Austin. The single core test series was run using the {\tt normal} queue
+compute nodes, while the parallel test series was run using one to four {\tt
+large memory} nodes each having one terabyte of local RAM. This large amount of
+memory was required for reasons explained in Section~\ref{JGAM}.
 
 \subsubsection{Data Reduction}
-The output produced by \emph{Perdigm} including the additional accuracy and speed measurements was redirected from the console to text files by the resource manager, \emph{SLURM} (Simple Linux User Resource Manager).
-Directories and filenames were chosen so that data would be easy to sort by individual test run.
-\emph{Python} scripts were then used to post-process and plot the data.
-These scripts averaged accuracy or speed data for a tests corresponding to a particular peridynamic node density over all iterations for that run, and then plotted this averaged data  as a function of peridynamic node density as it varied from test to test in the series.
-Similarly, for the parallel tests, the data was averaged and plotted in much the same way, but as a function of number of compute cores used to solve the problem.
-Confidence intervals were not calculated on the measurements.
-The main reason is that network traffic from other users makes running tests on the HPC cluster a time varying process.
-These data reduction and plotting scripts can be found on the corresponding author's website.
+
+The output produced by \emph{Perdigm} including the additional accuracy and
+speed measurements was redirected from the console to text files by the
+resource manager, \emph{SLURM} (Simple Linux User Resource Manager).
+Directories and filenames were chosen so that data would be easy to sort by
+individual test run.  \emph{Python} scripts were then used to post-process and
+plot the data. These scripts averaged accuracy or speed data for a tests
+corresponding to a particular peridynamic node density over all iterations for
+that run, and then plotted this averaged data  as a function of peridynamic
+node density as it varied from test to test in the series.  Similarly, for the
+parallel tests, the data was averaged and plotted in much the same way, but as
+a function of number of compute cores used to solve the problem. Confidence
+intervals were not calculated on the measurements. The main reason is that
+network traffic from other users makes running tests on the HPC cluster a time
+varying process. These data reduction and plotting scripts can be found on the
+corresponding author's website.
 
@@
 
-\subsection{Speed Data}
-Speed measurements were taken according to the methods mentioned in Section~\ref{tcs}.
-The averaged results of these speed measurements for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
-Average time in seconds to compute a TS matrix is plotted as a function of the number of nonzero TS elements.
-CS was the slowest of the four methods, taking the longest average time per iteration,  followed by CD, FD and AD in that order.
-Calculation time has a power-law relationship with problem size, with a power-law index of roughly $\num{1.0E-6}$ for each of the methods.
-This bodes well for the scalability of each of the methods by themselves in \emph{Peridigm}.
+\subsection{Speed Data} Speed measurements were taken according to the methods
+mentioned in Section~\ref{tcs}. The averaged results of these speed
+measurements for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
+Average time in seconds to compute a TS matrix is plotted as a function of the
+number of nonzero TS elements.  CS was the slowest of the four methods, taking
+the longest average time per iteration,  followed by CD, FD and AD in that
+order. Calculation time has a power-law relationship with problem size, with a
+power-law index of roughly $\num{1.0E-6}$ for each of the methods. This bodes
+well for the scalability of each of the methods by themselves in
+\emph{Peridigm}.
 %
-\begin{figure}[tbp]
-  \centering
-  \scalebox{1.0}{\input{./figs/serial_speed.pgf}}
-  \caption{Serial test series speed measurements.}
-  \label{fig:serial_speed}
+\begin{figure}[tbp] \centering \scalebox{1.0}{\input{./figs/serial_speed.pgf}}
+\caption{Serial test series speed measurements.} \label{fig:serial_speed}
 \end{figure}
 %
-For the parallel simulations, averaged speed results appear in Fig.~\ref{fig:multi_speed}.
-For this series, rather than varying the number of nonzero TS matrix elements, the number of compute cores used to solve the problem is increased, while the number of nonzero TS matrix elements remained constant at $\num{1.67E9}$ .
-CS again was the slowest method, followed by CD, AD and then FD.
-There is a somewhat surprise reversal of FD and AD for these parallel simulations.
-Figure~\ref{fig:multi_speed} indicates that when looking at average iteration time, the factor of speed improvement scales linearly with the number of compute cores used for the given conditions of this group of tests.
-Looking at Fig.~\ref{fig:multi_speed}, the reader can see that when core count doubles, compute time halves, for each method.
-Again, these results show that each of the methods as implemented in \emph{Peridigm} scale well when additional processors are used.
+For the parallel simulations, averaged speed results appear in
+Fig.~\ref{fig:multi_speed}. For this series, rather than varying the number of
+nonzero TS matrix elements, the number of compute cores used to solve the
+problem is increased, while the number of nonzero TS matrix elements remained
+constant at $\num{1.67E9}$ .  CS again was the slowest method, followed by CD,
+AD and then FD.  There is a somewhat surprise reversal of FD and AD for these
+parallel simulations.  Figure~\ref{fig:multi_speed} indicates that when looking
+at average iteration time, the factor of speed improvement scales linearly with
+the number of compute cores used for the given conditions of this group of
+tests. Looking at Fig.~\ref{fig:multi_speed}, the reader can see that when core
+count doubles, compute time halves, for each method. Again, these results show
+that each of the methods as implemented in \emph{Peridigm} scale well when
+additional processors are used. 
 %
-\begin{figure}[tbp]
-  \centering
-  \scalebox{1.0}{\input{./figs/multi_speed.pgf}}
-  \caption{Multicore test series speed measurements.}
-  \label{fig:multi_speed}
+\begin{figure}[tbp] \centering \scalebox{1.0}{\input{./figs/multi_speed.pgf}}
+\caption{Multicore test series speed measurements.} \label{fig:multi_speed}
 \end{figure}
 %
-A speculative explanation as to why FD and AD have reversed places in the speed tests with respect to the serial results is that the AD data-structures carry a list of partial derivatives of the functional evaluations along side the functional evaluations themselves only to be processed at the end of the tangent-stiffness calculation when the derivatives are requested.
-This represents additional load on the hardware memory subsystem, such that may exceed that of FD yet still fall under a supposed memory bandwidth threshold during serial operation.
-Cores within a multicore processor use the same shared bus to access memory, such that cores compete for memory access.
-Supposing that exceeding the memory bandwidth threshold penalizes additional requested bandwidth, it is likely that the decrease in available memory bandwidth per core caused by parallel operation coupled with AD's additional resource requirements causes the AD method to be slower than FD.
+A speculative explanation as to why FD and AD have reversed places in the speed
+tests with respect to the serial results is that the AD data-structures carry a
+list of partial derivatives of the functional evaluations along side the
+functional evaluations themselves only to be processed at the end of the
+tangent-stiffness calculation when the derivatives are requested. This
+represents additional load on the hardware memory subsystem, such that may
+exceed that of FD yet still fall under a supposed memory bandwidth threshold
+during serial operation. Cores within a multicore processor use the same shared
+bus to access memory, such that cores compete for memory access. Supposing that
+exceeding the memory bandwidth threshold penalizes additional requested
+bandwidth, it is likely that the decrease in available memory bandwidth per
+core caused by parallel operation coupled with AD's additional resource
+requirements causes the AD method to be slower than FD.
 
 \subsection{Accuracy Data}
-Data for accuracy measurements were taken according to the methods mentioned in Section~\ref{tcs}, and reduced using equation~(\ref{eqn:accuracy}).
-The averaged results of these accuracy measurements for the serial tests can be seen in Fig.~\ref{fig:serial_accuracy}.
-Averaged Frobenius norm (referred to as $l^2$-norm in the figure labels for conciseness) of the element-wise difference between a given TS matrix and the TS matrix produced by the AD method is plotted as a function of the number of nonzero TS matrix elements.
-CS was shown to be the most accurate of the methods when compared to AD and was roughly \emph{$6$ orders of magnitude more accurate than CD} and \emph{$9$ orders of magnitude more accurate than FD}.
-This accuracy ranking order was maintained for each test in the series.
-It should be pointed out that compared to the $l^2$-norms of any of the TS matrices by themselves, the magnitude of the accuracy metric calculated with Eqn.~\ref{eqn:accuracy} for any of the methods is relatively small.
-These values for the $l^2$-norms of the TS matrices appear in the simulation data which is available in the paper repository.
-The effect of these accuracy differences on convergence rate of the Newton solver was not determined.
+
+Data for accuracy measurements were taken according to the methods mentioned in
+Section~\ref{tcs}, and reduced using equation~(\ref{eqn:accuracy}). The
+averaged results of these accuracy measurements for the serial tests can be
+seen in Fig.~\ref{fig:serial_accuracy}. Averaged Frobenius norm (referred to as
+$l^2$-norm in the figure labels for conciseness) of the element-wise difference
+between a given TS matrix and the TS matrix produced by the AD method is
+plotted as a function of the number of nonzero TS matrix elements. CS was shown
+to be the most accurate of the methods when compared to AD and was roughly
+\emph{$6$ orders of magnitude more accurate than CD} and \emph{$9$ orders of
+magnitude more accurate than FD}. This accuracy ranking order was maintained
+for each test in the series. It should be pointed out that compared to the
+$l^2$-norms of any of the TS matrices by themselves, the magnitude of the
+accuracy metric calculated with Eqn.~\ref{eqn:accuracy} for any of the methods
+is relatively small. These values for the $l^2$-norms of the TS matrices appear
+in the simulation data which is available in the paper repository. The effect
+of these accuracy differences on convergence rate of the Newton solver was not
+determined.  
 %
-The accuracy of FD and CD could be improved by using a smaller probe distance, $h$; however, if $h$ is too small, then the methods can suffer from severe effects, including solution divergence, of the round-off error associated with subtracting two numbers that are very close to one another.
-The value of $h$ used in this study is the default value used in \emph{Peridigm} which is heuristically derived based on computational node spacing.
-One of the beauties of the CS method is that since dependence on $h$ decreases for small $h$, which a numerical example in \cite[Table 1]{squire1998using} demonstrates, $h$, cannot be too small, therefore it can be set to machine epsilon and the analysis performed without the issues associated with round-off error.
-The value chosen for $h$ for CS used in this study was $\num{1.E-100}$ to demonstrate this.
+The accuracy of FD and CD could be improved by using a smaller probe distance,
+$h$; however, if $h$ is too small, then the methods can suffer from severe
+effects, including solution divergence, of the round-off error associated with
+subtracting two numbers that are very close to one another.  The value of $h$
+used in this study is the default value used in \emph{Peridigm} which is
+heuristically derived based on computational node spacing.  One of the beauties
+of the CS method is that since dependence on $h$ decreases for small $h$, which
+a numerical example in \cite[Table 1]{squire1998using} demonstrates, $h$,
+cannot be too small, therefore it can be set to machine epsilon and the
+analysis performed without the issues associated with round-off error. The
+value chosen for $h$ for CS used in this study was $\num{1.E-100}$ to
+demonstrate this.
 %
-\begin{figure}[tbp]
-  \centering
-  \scalebox{1.0}{\input{./figs/serial_accuracy.pgf}}
-  \caption{Serial test series accuracy measurements.}
-  \label{fig:serial_accuracy}
-\end{figure}
+\begin{figure}[tbp] \centering
+\scalebox{1.0}{\input{./figs/serial_accuracy.pgf}} \caption{Serial test series
+accuracy measurements.} \label{fig:serial_accuracy} \end{figure}
 %
-The averaged results of the accuracy measurements taken for the parallel test series can be seen in Fig.~\ref{fig:multi_accuracy}.
-The ordinate axis units remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the abscissa now indicates number of compute cores used to solve the test problem.
-Accuracy trends matched those for the single core tests.
+The averaged results of the accuracy measurements taken for the parallel test
+series can be seen in Fig.~\ref{fig:multi_accuracy}. The ordinate axis units
+remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the abscissa now
+indicates number of compute cores used to solve the test problem. Accuracy
+trends matched those for the single core tests. 
 %
-\begin{figure}[tbp]
-  \centering
-  \scalebox{1.0}{\input{./figs/multi_accuracy.pgf}}
-  \caption{Multicore test series accuracy measurements.}
-  \label{fig:multi_accuracy}
-\end{figure}
+\begin{figure}[tbp] \centering
+\scalebox{1.0}{\input{./figs/multi_accuracy.pgf}} \caption{Multicore test
+series accuracy measurements.} \label{fig:multi_accuracy} \end{figure}
 
@@
 %
-Based on the experimental results, it appears that for users of \emph{Peridigm} there is a case to be made for using AD for reasons of performance and accuracy.
-However, certain parallel tests done in this study showed a small performance gain over AD with the FD method.
-While the results showed that CS produced more accurate tangent-stiffness matrices than CD and FD under the parameters of the tests, it was not determined whether or not this is a clear advantage of CS over FD in terms of accuracy of final predicted displacements and speed of convergence.
-The reader is reminded that the time required to complete an iteration must be considered along with the total number of iterations taken order to measure speed of convergence.
-The goal of the comparative study was to benchmark single iteration performance and not convergence rate.
-For new developments it is recommended that combined accuracy and speed of iteration measurements be conducted along with entirely separate convergence studies.
-It should be mentioned that a related work, \cite{perez2012numerical}, compares the speed and accuracy of complex-step to finite difference for producing tangent-stiffness matrices in solid mechanics simulations using complex inelastic material model, where the researchers similarly found CS to be accurate yet relatively computationally expensive compared to FD and CD.
-In that work, full convergence studies were also done and indicated that quadratic convergence could be achieved with complex-step all of the methods \cite[p.28]{perez2012numerical}.
+Based on the experimental results, it appears that for users of \emph{Peridigm}
+there is a case to be made for using AD for reasons of performance and
+accuracy. However, certain parallel tests done in this study showed a small
+performance gain over AD with the FD method.  While the results showed that CS
+produced more accurate tangent-stiffness matrices than CD and FD under the
+parameters of the tests, it was not determined whether or not this is a clear
+advantage of CS over FD in terms of accuracy of final predicted displacements
+and speed of convergence. The reader is reminded that the time required to
+complete an iteration must be considered along with the total number of
+iterations taken order to measure speed of convergence. The goal of the
+comparative study was to benchmark single iteration performance and not
+convergence rate. For new developments it is recommended that combined accuracy
+and speed of iteration measurements be conducted along with entirely separate
+convergence studies. It should be mentioned that a related work,
+\cite{perez2012numerical}, compares the speed and accuracy of complex-step to
+finite difference for producing tangent-stiffness matrices in solid mechanics
+simulations using a complex inelastic material model, where the researchers
+similarly found CS to be accurate yet relatively computationally expensive
+compared to FD and CD. In that work, full convergence studies were also done
+and indicated that quadratic convergence could be achieved with complex-step
+\remove[Michael]{all of the methods} \cite[p.28]{perez2012numerical}.
 
-The CS method was shown to be highly accurate in calculating tangent-stiffness matrices by the standards of this study, and it is slower than other methods for a naive implementation.
-In the context of future use of CS in \emph{Peridigm}, CS holds the advantage of relying on byte-copyable data types, while AD requires a complicated serialization and deserialization in order to function as byte-copyable.
-This is fine for copying in between MPI processes, but deserialization on an accelerator or GPU is very difficult, as this would require compiling a version of the AD library being used that were compatible with the accelerator or GPU being used.
-What this means is that if \emph{Peridigm} or any simulation software were to be reformulated to take advantage of accelerators or GPUs, the flexibility of complex-step owed to its simple implementation may make it more viable for that application given its high accuracy.
-Additional performance gains may be achieved through the use of the templated \emph{Tpetra} package in \emph{Trilinos} which would allow the use of complex data types natively, and alleviate the expense of the dynamic-casts that were utilized in this study.
-As a final comment, it should be noted that from a code development perspective the CS method may be the easiest of the methods to implement; It only requires one functional evaluation, there is no heuristic choice of probe distance, and there is no dependence on an external AD library.
-From this perspective, the method may still be an attractive method for TS matrix evaluation.
+The CS method was shown to be highly accurate in calculating tangent-stiffness
+matrices by the standards of this study, and it is slower than other methods
+for a naive implementation. In the context of future use of CS in
+\emph{Peridigm}, CS holds the advantage of relying on byte-copyable data types,
+while AD requires a complicated serialization and deserialization in order to
+function as byte-copyable. This is fine for copying in between MPI processes,
+but deserialization on an accelerator or GPU is very difficult, as this would
+require compiling a version of the AD library being used that were compatible
+with the accelerator or GPU being used. What this means is that if
+\emph{Peridigm} or any simulation software were to be reformulated to take
+advantage of accelerators or GPUs, the flexibility of complex-step owed to its
+simple implementation may make it more viable for that application given its
+high accuracy. Additional performance gains may be achieved through the use of
+the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
+use of complex data types natively, and alleviate the expense of the
+dynamic-casts that were utilized in this study.  As a final comment, it should
+be noted that from a code development perspective the CS method may be the
+easiest of the methods to implement; It only requires one functional
+evaluation, there is no heuristic choice of probe distance, and there is no
+dependence on an external AD library.  From this perspective, the method may
+still be an attractive method for TS matrix evaluation.
 

### 2014-02-08 response/RefereeReport.tex
response to I.1 started

Page 1

Computer Methods in Applied Mechanics and Engineering (CMAME) \\Referee
Repor t \\

Title: ``A comparison of different methods for calculating
tangent-stiffness matrices in a massively \\parallel computational
peridynamics code'' \\

Authors: M. D. Brothers, J. T. Foster, and H. R. Millwater \\

Manuscript number: CMAME-D-13-01004 \\

A. Summary \\

This manuscript presents comparisons between different numerical
differentiation methods \\for the computation of entries of
tangent-stiffness matrices, used in Newton's method for \\quasi-static
nonlinear analysis of solid structures. The methods include: automatic
differen- \\tiation (AD), forward ﬁnite difference (FD), centered ﬁnite
difference (CD) and complex-step \\(CS). The authors perform the
comparisons within the computational peridynamics code \\Peridigm, where
the authors implemented the CS method. Numerical results include com-
\\parisons of computational times between all the methods, as well as
comparisons of accu- \\racy between FD, CD, and CS, relative to AD. The
results show that AD is the fastest method \\in serial and FD in
parallel, whereas CS is the slowest method. However, the accuracy of
\\CS is several orders of magnitude higher than the ﬁnite difference
methods (FD and CD). \\Consequently, although AD seems to be the best
method in terms of combined speed and \\accuracy, the results suggest
that CS may represent a viable alternative for cases where AD
\\implementations are not practical. \\B. Evaluation \\

This manuscript focuses on the CS method. This method and its comparison
to other ﬁnite- \\difference methods have been previously presented in
the literature. For example, the CS \\method has been described in
{[}1{]}, with a comparison between FD and CS. Applications of \\the CS
method to compute tangent matrices appear in {[}2, 3{]}, with
comparisons between \\FD, CD, and CS. In addition, the CS method has
been used for sensitivity analysis in {[}5, 6{]}, \\and in {[}4{]}, a
detailed description of the CS method and comparisons to FD, CD, and AD
are \\presented. Fur ther uses of the complex-step method appear in
{[}7, 8, 9{]}, and extensions of \\the CS method are discussed in {[}10,
11{]}. \\However, this manuscript presents some aspects of the CS method
which seem novel. \\These are: \\

1. comparisons between FD, CD, CS, and AD in terms of computational
time, \\2. implementation of the CS method in parallel, \\3.
implementation of the CS method within Peridigm. \\

In par ticular, the implementation in Peridigm is of strong interest to
researchers working with \\the peridynamic theory of solid mechanics.
\\I believe this manuscript is of value to both the computational
mechanics community in gen- \\eral, and the peridynamic community in par
ticular. \\I suggest it to be considered for pub- \\lication in the
Computer Methods in Applied Mechanics and Engineering journal, after the
\\issues described below have been properly addressed. \\

1 \\

Page 2

C. Information to the authors \\

I. Larger issues \\

1. Some par ts of the manuscript suggest that the authors introduced a
new method, \\i.e., the complex-step (CS) method. For example, in the
introduction, they say \\``A goal of this study was to develop and
evaluate a new, accurate, and practical \\method \ldots{}'' Another
example appears in page 3, where the authors say ``The aim \\of this
paper is not only to introduce the new complex-step method in the
context \\of evaluating tangent-stiffness matrices ..''. However, the CS
method has already \\appeared in many publications since 1998 {[}1{]}.
Please revise the way you refer to \\the CS method along the manuscript.
\\2. The work presented focuses on a parallel computational peridynamic
code, and \\uses peridynamic models for mechanics. However, no detailed
explanation of the \\equations involved is presented. Although the
tangent-stiffness calculations are \\not specialized to Peridigm, the
authors should include a shor t section introducing \\the peridynamic
theory and the speciﬁc constitutive model used, i.e., the elastic
\\peridynamic solid. You may add this to the already included very brief
description \\of peridynamics in page 9 (lines 15-26). This will also
help to explain the concept \\of the peridynamic horizon in page 10.
\\3. The manuscript seems to be written in a language more familiar to
computer scien- \\tists. I would encourage the authors to revise cer
tain par ts of the manuscript to be \\presented in a language more
familiar to the computational mechanics community. \\4. In Section
1.1.1, please include a more comprehensive literature review for the CS
\\method, including fur ther references (some of them are listed below).
\\In par ticu- \\lar, specify what aspects of the CS method have been
investigated in each refer- \\ence. Then, you need to clearly mention
your contribution within the CS method, \\in contrast to the other
references (some comments in this regard appear in the \\manuscript in
page 8). \\5. The comparisons of the different algorithms are only
performed with respect to the \\construction of tangent-stiffness
matrices. In par ticular, the authors always use the \\solution of AD at
each time step to compute the entries of the tangent-stiffness
\\matrices, even for the FD, CD, and CS methods. However, the authors
need to \\demonstrate the performance of the CS method, in comparison to
the other ones, \\with respect to the solution of real, ideally
non-linear problems. In other words, how \\the solutions compare in
terms of accuracy and speed between the methods when \\solving given
problems? Please include two examples of interest. The authors \\suggest
that two examples involving non-linear systems are already included in
the \\paper repository. \\The authors say in the Conclusion: ``While the
results showed that CS produced \\more accurate tangent-stiffness
matrices than CD and FD under the parameters \\of the tests, it was not
determined whether or not this is a clear advantage of CS \\over FD in
terms of accuracy of ﬁnal predicted displacements and speed of conver-
\\gence.'' This needs to be studied and results in this regard added to
this manuscript. \\II. Smaller issues \\

1. Pg. 2: In line 14, explain what does the term ``algorithmically
consistent'' mean. \\2. Pg. 2: In line 36, what do you mean by
``discrimination''? \\3. Pg. 2: In line 37, explain what do you mean by
``in-situ instrumentation''. \\4. Pg. 3: In lines 19-29, specify the
sections where each par t of the manuscript is \\included. \\

2 \\

Page 3

5. Pg. 3: You may consider including the exact web address of the paper
repository. \\6. Pg. 3: Please include the forward and centered ﬁnite
difference expressions for \\easier comparison with (2); include the
truncation error as well. You do not have to \\derive these expressions
though. \\7. Pg. 4: In Eq. (1), `` ∂ f 2 \\∂x2 '' should be `` ∂ 2 f
\\∂x2 ''; similarly, for the third derivative. \\8. Pg. 4: You may
consider writing Eq. (2) with the truncation error, i.e., \\

+ O(h2 ). \\

(2) \\

Im(f (x + ih)) \\h \\

∂ f (x) \\= \\∂x \\In this case, one replaces ``≈'' by ``=''. \\9. Pg.
6: In lines 9-11, please add references to the ``Poisson problem'' and
to the \\``mathematical optimization''. \\10. Pg. 7: Should Eqs. (4),
(5), and (6) use ``≈'' instead of ``=''? \\11. Pg. 7: In line 28, note
that (cid:126)u is not the current conﬁguration, but the displacement
\\ﬁeld. Please revise the sentence. \\12. Pg. 7: Please provide an
expression for F int \\in the model used. \\i \\13. Pg. 9: In line 26,
please add the reference {[}12{]} below to ``coupling to molecular
\\simulations {[}29{]}.'' \\14. Pg. 10: Please add an illustration of
the block undergoing tension in Section 2.2, \\for clarity. Also,
clarify the displacement boundary conditions; are they applied in a
\\volumetric layer? \\15. Pg. 10: In line 40, you mentioned a bandwidth
of 7. This seems to be clear in 1D. \\Does this hold in higher
dimensions? Please clarify. \\16. Pg. 10: In lines 47-49, you say ``The
speciﬁc parameters \ldots{} can be found \ldots{} in the \\paper
repository.'' It seems a good idea to have the paper repository with
ﬁles that a \\user can eventually use. However, the content of the paper
should be independent \\of the repository. Please include a table with
the parameters used. Similarly, in \\page 11 (lines 11-13). \\17. Pg.
14: In line 55, is there any conclusion you can draw from the power law
index \\1.0 × 10−6? \\18. Pgs. 14-: Please clarify the value of the
meshsize used on each of the experiments. \\19. Pg. 16: Would it be
possible to rescale Fig. 2, so that the dependence of the time \\on the
number of cores will look indeed linear? \\20. Pg. 16: In lines 54-56,
you say ``It should be pointed out that compared to the (cid:96)2 -
\\norms of any of the TS matrices by themselves, the magnitude of the
accuracy \\metric \ldots{} is relatively small.'' Please include in the
manuscript the explicit informa- \\tion on the relative error. \\It
seems that you already have such information in the \\paper repository.
\\21. Pg. 17: Include the values of the probe distances h used for each
method. \\22. Pg. 17: In lines 53-54, you say ``Accuracy trends matched
those for the single core \\tests.'' Indeed, comparing Figs. 3 and 4,
the results for FD seem to be quite the \\same in serial and in
parallel. For CS the accuracy seems higher in serial. For CD, \\the
accuracy is also higher in serial and, in serial, it has a ﬂuctuating
behavior. Can \\you comment on these observations? \\23. Pgs. 17-18:
What are the units of K ? If F int in Eq. (4) etc. is internal force,
then K \\should have units of force/length. However, in Figs. 3 and 4
you use the unit of \\MPa, which is force/length2 . Please clarify.
\\24. Pg. 19: Are the units of the ver tical axis in Fig. 5 ``seconds'',
as in Figs. 1 and 2? \\Please indicate that. Also, is the label of the
ver tical axis in Fig. 6 correct? \\

3 \\

Page 4

25. Pg. 20: In lines 32-35, clarify what ``byte-copyable'' mean and why
is that impor tant. \\26. Pg. 21: Please include a section title for the
references. \\27. Pg. 24: In line 23, you may add the truncation error
O(h2 ) to the expression. Oth- \\erwise, you may need to replace the
``='' sign by the ``≈'' sign. However, adding \\the error term would
make more clear the limiting expression when h → 0. Also, \\replace the
comma by a period. \\28. Pg. 24: Why do we need S and X in Appendix B?
\\29. Pgs. 25-26: Are the functions in the second column of the table in
2(c) (page 25) \\approximations? Why do you label the column
``Approximates Function''? Simi- \\larly, in line 16 (page 26) you say
``approximated by the par tial derivative''; is it an \\approximation?
\\30. Pg. 26: Please clarify whether storage is needed in automatic
differentiation. For \\example, do we need to store w, v , and u? \\III.
Typos/editing suggestions \\

is \\

1. Pg. 1: In line 43, remove ``the'' in ``all the of the code''. \\2.
Pg. 1: In line 46, replace ``numeric differentiation'' by ``numerical
differentiation''. \\3. Pg. 2: In line 25, the sentence ``The
distinction of ``accurate'' is deﬁned by \ldots{}'' \\awkward. Please
revise it. \\4. Pg. 2: In lines 32 and 48, ``ﬁnite-difference'' should
be ``forward-difference''. \\5. Pg. 2: In lines 33, I believe ``aide''
should be ``aid''. \\6. Pg. 3: In line 12, I would replace ``such as
required'' by ``such as those required''. \\7. Pg. 3: In line 22,
replace ``with each the methods'' by ``with each of the methods''. \\8.
Pg. 3: In line 25, replace ``implementing complex-step'' by
``implementing the complex- \\step method''. \\9. Pg. 3: I would replace
the title of Section 1.1, ``Differentiation Techniques'' by ``Dif-
\\ferentiation techniques'' for consistency, i.e., only capitalize the
ﬁrst letter of the ﬁrst \\word. Similarly, for the rest of the section
titles. \\10. Pg. 4: In line 10, ``be be made complex'' should be ``be
made complex''. \\11. Pg. 4: In Eq. (1), replace comma by period. \\12.
Pg. 4: In line 22, you have an extra white space in ``O (h2 )''. \\13.
Pg. 4: I would check the proper use of hyphens along the manuscript. For
example, \\in the title of Section 1.1.2, I would replace
``Automatic-differentiation'' by ``Automatic \\differentiation'' (i.e.,
no hyphen). \\14. Pg. 4: In line 42, replace ``based the chain-rule'' by
``based on the chain rule''. \\15. Pg. 4: In lines 45-46, ``(add,
multiply, power, transcendental and the like)'' should \\probably be
``(addition, multiplication, exponentiation, and the like)'' or similar.
\\16. Pg. 4: In line 47, I would replace ``used in the study'' by ``used
in this study''. Similarly, \\in other instances of the manuscript.
\\17. Pg. 5: In line 39, you may replace ``the par tials'' by ``the par
tial derivatives''. Similarly \\in page 26. \\18. Pg. 5: I would replace
the title of Section 1.2 ``Tangent-stiffness'' by ``Tangent-stiffness
\\matrix''. Similarly, in line 50. \\19. Pg. 6: In line 14,
``quasi-Newton's method's'' should probably be ``quasi-Newton
\\methods''. \\20. Pg. 6: In line 18, ``about about a par ticular''
should be ``about a par ticular''. \\21. Pg. 6: In line 51, you may
replace ``argument to F '' by ``argument of F ''. \\22. Pg. 7: In line
11, you may replace ``single independent vector component of the
\\function'' by ``single independent vector component of the argument of
the function''. \\

4 \\

Page 5

23. Pg. 7: In line 24, replace ``Where Kij is'' by ``where Kij is''.
\\24. Pg. 7: In line 40, replace ``in literature {[}19{]}'' by ``in the
literature {[}19{]}''. \\25. Pg. 8: Remove the period at the end of the
titles of Sections 2 and 2.1, for consis- \\tency. Similarly, for the
rest of the manuscript. Also, use italic font for ``Peridigm'' in \\the
title of §2.1. \\26. Pg. 9: In line 29, replace ``Perdigm'' by
``Peridigm''. Similarly in page 13 (line 44). \\27. Pg. 9: In line 37,
replace ``be simply be declared'' by ``be simply declared''. \\28. Pg.
9: In line 55, I would replace ``computational scientists'' by
``computer scientists''. \\29. Pg. 10: In line 30, I would replace ``The
purpose of applying the load'' by ``Applying \\the load''. \\30. Pg. 13:
I would add a comma at the end of Eq. (8), and replace ``Where D is
\\distance'' by ``where D is distance'' in line 14. \\31. Pg. 13: In
line 15, replace ``based method, M is'' by ``based method, and M is''.
\\32. Pg. 13: In line 31, replace ``correspond to'' by ``corresponds
to''. \\33. Pg. 14: In lines 19-21, you have twice the word ``nonzero''
in ``number of nonzero \\tangent-stiffness (TS) non-zero matrix
elements''. \\34. Pg. 14: In lines 22, I would replace ``The units for
accuracy are derived equation (8) \\and the units used to deﬁne bulk
material proper ties, mentioned in Section 2.2.'' by \\``The accuracy
measure is given by equation (8) and the choice of elastic moduli is
\\mentioned in Section 2.2.'' \\35. Pg. 15: In line 54, it seems that
``yet still fall'' should be ``yet still falls''. \\36. Pg. 17: In line
45 you say ``h, cannot be too small,''. \\I understand that you mean
\\that you can take h as small as possible. However, the way written may
have the \\opposite connotation. I would revise it for clarity. \\37.
Pg. 18: In lines 35-36, replace ``the number nonzero'' by ``the number
of nonzero''. \\38. Pg. 18: In line 39, replace ``the the number of
nonzero'' by ``the number of nonzero''. \\39. Pg. 18: In line 50, I
would replace ``experimental results,'' by ``computational experi-
\\ments,''. Otherwise, it may give the impression of physical
experiments. \\40. Pg. 20: In line 13, is the use of the word ``order''
in ``number of iterations taken order \\to measure'' correct? \\41. Pg.
20: In lines 26-27, replace ``complex-step all of the methods'' by
``complex-step \\in all of the methods''. \\42. Pg. 20: In line 50, you
may either replace ``to implement; It only'' by ``to implement; \\it
only''; or replace the semicolon by a period. \\43. Pgs. 21-23: You may
capitalize the ﬁrst letters on the journal names in your refer- \\ences
{[}9, 20{]} \\44. Pg. 22: In your reference {[}10{]}, you may replace
``fr ´echet'' by ``Fr ´echet''. \\45. Pg. 23: Update your reference
{[}21{]} (see the reference {[}3{]} in this repor t). Also, \\update
your reference {[}27{]}. \\46. Pg. 24: In line 18, replace the comma by
a period. \\47. Pg. 24: I would replace the comma by a period in line
28, and replace ``using \\L'H ˆopital's rule'' by ``Using l'H ˆopital's
rule'' in line 31. \\48. Pg. 24: I would replace the comma by a period
in line 34, and replace ``of course, \\as h → 0'' by ``As h → 0'' in
line 37. \\49. Pg. 24: I would replace the title of Appendix B ``AD
Example'' by ``Automatic differ- \\entiation example''. \\50. Pg. 24: In
line 45, you may replace ``sin(cos(x))'' by ``sin(cos(x))''. For that,
use \textbackslash{}sin \\instead of sin, etc. (similarly in the bottom
of the page and in page 25). Also, add \\a period at the end of the
expression. \\

5 \\

Page 6

51. Pg. 25: In line 55, you may remove the ``·'' symbol in `` d \\dx ·
g(h(k(x)))''. \\52. Pgs. 25 and 26: I would number the table in 2(c)
(page 25). Let's assume the \\number is ``2''. Then, in line 13 (page
26), I would replace ``chosen for each function \\and held as constant''
by ``chosen for each function as in Table 2 and held constant''. \\53.
Pg. 26: In line 24, add a period at the end of the expression. \\

References \\

{[}1{]} W. Squire and G. Trapp, Using complex variables to estimate
derivatives of real functions, \\SIAM Rev. 40 (1998), pp. 110-112. \\

{[}2{]} A. P ´erez-Foguet, A. Rodr´ıguez-Ferran, A. Huer ta, Numerical
differentiation for local and \\global tangent operators in
computational plasticity, Comput. Methods Appl. Mech. Engrg. \\189
(2000), pp. 277-296. \\

{[}3{]} A. P ´erez-Foguet, A. Rodr´ıguez-Ferran, A. Huer ta, Numerical
differentiation for non-trivial con- \\sistent tangent matrices: an
application to the MRS-Lade model, Int. J. Numer. Meth. Engng \\48
(2000), pp. 159-184. \\

{[}4{]} J. Mar tins, P. Sturdza, J. Alonso, The complex-step derivative
approximation, ACM Transac- \\tions on Mathematical Software (TOMS),
ISSN 0098-3500, 09/2003, Volume 29, Issue 3, pp. \\245-262. \\

{[}5{]} W. K. Anderson, J. C. Newman, D. L. Whitﬁeld, and E. J. Nielsen,
Sensitivity analysis for \\Navier--Stokes equations on unstructured
meshes using complex variables, AIAA Journal, \\ISSN 0001-1452, 01/2001,
Volume 39, Issue 1, pp. 56-63. \\

{[}6{]} J. C. Newman, W. K. Anderson, and L. D. Whitﬁeld,
Multidisciplinary sensitivity derivatives \\using complex variables,
Tech. Rep. MSSU-COE-ERC-98-08 (July), Computational Fluid Dy- \\namics
Laboratory. \\

{[}7{]} A. H. Al-Mohy and N. J. Higham, The complex step approximation
to the Fr ´echet derivative \\of a matrix function, Numer. Algor. 53
(2010), pp.133-148. \\

{[}8{]} W. Jin, B. H. Dennis, and B. P. Wang, Improved sensitivity
analysis using a complex variable \\semi-analytical method, Struct.
Multidisc. Optim. 41 (2010), pp. 433-439. \\

{[}9{]} A. Voorhees, H. Millwater, and R. Bagley, Finite Elements in
Analysis and Design 47 (2011), \\pp. 1146-1156. \\

{[}10{]} K.-L. Lai and J. L. Crassidis, Extensions of the ﬁrst and
second complex-step derivative ap- \\proximations, Journal of
Computational and Applied Mathematics 219 (2008), pp. 276-293. \\

{[}11{]} R. Abreu, D. Stich, and J. Morales, On the generalization of
the complex step method, Journal \\of Computational and Applied
Mathematics 241 (2013), pp. 84-102. \\

{[}12{]} P. Seleson, M. L. Parks, and M. Gunzburger, Peridynamic
state-based models and the \\embedded-atom model. Communications in
Computational Physics, 15(1) (2014), pp. 179- \\205. \\

6 \\

Page: \hyperref[1]{1}, \hyperref[2]{2}, \hyperref[3]{3},
\hyperref[4]{4}, \hyperref[5]{5}, \hyperref[6]{6}

### 2014-02-08 response/response.tex
re,I.2 ts mat. in terms of perd. eqn. mot.

@@
 %%%% First Reviewer %%%%
-\section{Information to the authors}
+\section{Re: Information to the authors}
 %
@@
   \begin{itemize}
-    \item We have added a paragraph at the end of Section 1 that highlights the difference between peridynamics and previous Òparticle methods.Ó The main difference is that in peridynamics, we numerically approximate a different set of continuum equations.
+      \item
+        The way the CS method is referred to has been revised throughout,
+        specifically: 
+    \item In the abstract, "... the new method ..." has been
+        replaced by "... the method ...", "... newly implemented ..." has been
+        replaced by "... implemented ..."; 
+    \item in the introduction, paragraph two,
+        "... a new, accurate ..." has been replaced by "... a relatively
+        underexplored, accurate ...", in paragraph two and four "... the new
+        method ..." has been replaced by "... the method ...", in paragraph
+        five, "... introduce the new complex-step method in the context of ..."
+        has been replaced by "... study the complex-step method of ...", "...
+        the new complex-step ..." has been replaced by "... the complex-step
+        ...".
+    \item 
+        It is the authors' opinion that the above changes correct the suggestion
+        that the authors introduce a new method.
+
+  \end{itemize}}
+
+    \item
+        The work presented focuses on a parallel computational peridynamic code, and
+        uses peridynamic models for mechanics. However, no detailed explanation of the
+        equations involved is presented. Although the tangent-stiffness calculations are
+        not specialized to Peridigm, the authors should include a short section introducing
+        the peridynamic theory and the speciﬁc constitutive model used, i.e., the elastic
+        peridynamic solid. You may add this to the already included very brief description
+        of peridynamics in page 9 (lines 15-26). This will also help to explain the concept
+        of the peridynamic horizon in page 10.
+
+{\color{red}
+  \begin{itemize}
+    \item 
+        We have revised the brief desciption of peridynamics in page 9 to be more complete:
+    \item
+        Material from silling2007 is interpreted including: the definition of vector states,
+        the derivation of conservation laws in terms of state-based peridynamics, the definition
+        of the isotropic elastic solid AKA the Linear elastic Peridynamic Solid.
+    \item
+        It is the authors' opinion that the additional material mentioned above explains the equations involved
+        in the peridynamic models for mechanics, used in relation to the peridynamics code
+        studied in the work.
+
   \end{itemize}}

### 2014-02-08 jacobian_calc_CMAME.tex
re,I.2 ts mat. in terms of perd. eqn. mot.

@@
 
-Briefly, peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
+\note[Michael]{re: I.2}\change[Michael]{Briefly, peridynamics}Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
 is a nonlocal reformulation of the partial-differential equations that provide
@@
 \cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale
-coupling to molecular simulations \cite{seleson2009peridynamics}.
+coupling to molecular simulations \cite{seleson2009peridynamics}. The related mathematical
+definitions and parts of peridynamic theory especially relevant to the present
+work are explained in the sequel.
+
+\subsubsection{Vector states for state-based peridynamic theory.}
+
+\subsubsection{The state-based peridynamic equation of motion.}
+
+\subsubsection{State-based consitutive models.}
+
+\subsubsection{The linear peridynamic solid.}
+
+\subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simualtion.}
+The tangent stiffness matrix is the frechet derivative of the residual wrt the
+deformation state.  For the purposes of this work, the residual represents the unbalanced forces in the
+equilibrium version of the peridynamic equation of motion. The internal force
+density is computed in the peridynamic equation of motion by computing the
+volume integral over the whole body of the Frechet derivative of the 
+strain energy state wrt to the deformed configuration inner product with the deformation state, per node,
+for each of the three directions in the original reference frame.
+
+For the finite-difference type methods in \emph{Peridigm}, the perturbation method is employed
+by probing the deformed configuration one degree of freedom at a time and so allowing the
+sensitivity of the force density state wrt the deformation state for all nodes to be computed.
 

### 2014-02-08 jacobian_calc_CMAME.tex
intro to response for I.2

@@
 work are explained in the sequel.
+
+The type of material model used in this study is referred to as a linear
+peridynamic solid, having what is called a state-based constitutive model. In
+order for a definition of this material to be useful here, background material
+needs to be laid out to enable the discussion. To begin with, some essential
+mathematical definitions will be listed and peridynamic theory summarized. Then
+state-based constitutive models will be summarized, followed by a description of
+the linear peridynamic solid model used in this work. After the peridynamic
+equation of motion and consitutive model in use have been discussed, it is then
+possible to give a more precise definition of the tangent stiffness matrix for
+the problem being discussed in this work. The following discussion is largely
+taken from \cite{silling:psa}, but reproduced here fo clarity.
 

### 2014-02-14 jacobian_calc_CMAME.tex
added reason for changes as notes

@@
 
-\note[Michael]{re: I.2}\change[Michael]{Briefly, peridynamics}Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
+\note[Michael]{re: I.2, John I think you better handle expanding this section,
+I was going to essentially copy silling:psa.}\change[Michael]{Briefly, peridynamics}Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
 is a nonlocal reformulation of the partial-differential equations that provide
@@
 
+\note[Michael]{re: I.2. John, this was to be the skeleton for the new sections on peridynamics.}
 \subsubsection{Vector states for state-based peridynamic theory.}
@@
 
+\note[Michael]{re: I.2} \add[Michael]{
 \subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simualtion.}
@@
 density is computed in the peridynamic equation of motion by computing the
-volume integral over the whole body of the Frechet derivative of the 
+volume integral over, a horizon, of the Frechet derivative of the 
 strain energy state wrt to the deformed configuration inner product with the deformation state, per node,
@@
 sensitivity of the force density state wrt the deformation state for all nodes to be computed.
-
+}
 \subsubsection{Implementing the complex-step method in Peridigm} 

### 2014-02-14 jacobian_calc_CMAME.tex
I.3 addressed, I.2 partially addressed

@@
 
-\note[Michael]{re: I.2, John I think you better handle expanding this section,
+\note[Michael]{re: I.2, John I think you would better handle expanding this section,
 I was going to essentially copy silling:psa.}\change[Michael]{Briefly, peridynamics}Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
@@
 
-For the finite-difference type methods in \emph{Peridigm}, the perturbation method is employed
-by probing the deformed configuration one degree of freedom at a time and so allowing the
-sensitivity of the force density state wrt the deformation state for all nodes to be computed.
-}
+For the finite-difference type methods in \emph{Peridigm}, the perturbation
+method is employed by probing the deformed configuration one degree of freedom
+at a time and so allowing the sensitivity of the force density state wrt the
+deformation state for all nodes to be computed.  }
+
 \subsubsection{Implementing the complex-step method in Peridigm} 
@@
 complex number data types.  This is because \emph{Epetra} vectors, which can be
-thought of as distributed memory parallel cousins to standard C++ STL vectors,
-are hard coded to be of type {\tt double} and can not be simply be declared as
-having a complex data type. Therefore, program methods that were used in the
-course of applying the FD method were copied, renamed and re-written to follow
-the CS formula. This involved changing the data type of intermediate variables
-which were locally scoped, and dynamically casting persistent memory variables
-to complex data types where necessary. No attempt was made to ``tune'' the CS
-code in any way for performance. The algorithm  was implemented in such as way
-that it was in one-to-one correspondence with the FD and CD methods; however,
-it must be noted that the overhead associated with the dynamic casts in the CS
-method likely hurts its overall performance and this issue could be alleviated
-by using the next generation of templated \emph{TPetra} vectors available in
-\emph{Trilinos} that are capable of taking an explicit instantiation of any
-data type including {\tt complex}. The version of \emph{Peridigm} modified to
-implement CS can be found in the paper repository.
+thought of as distributed memory parallel cousins to standard C++
+\change[Michael]{STL}{Standard Template Library} vectors, are hard coded to be
+of type {\tt double} and can not be simply be declared as having a complex data
+type. \note[Michael]{re: I.3}\add[Michael]{For the reader's convenience,
+    \emph{Epetra} vectors and STL vectors are what are known as 'container
+    classes', whose purpose in object oriented computer programming is to
+    automate the creation, accessing, modification and destruction of the
+    underlying contained data. Templated container classes allow for user
+    choice in what properties that underlying data exihibits.  For example, in
+    this work, complex arithmetic cannot be performed on the data contained
+    within an \emph{Epetra} vector because that data, of type {\tt double} does
+    not have the property of following the rules of complex arithmetic. The
+    reason the non-templated \emph{Epetra} vector could not be dispensed with
+    for the purposes of this work is because the \emph{Epetra} vector was
+tightly integrated with \emph{Peridigm's} parallel communication strategy and
+linear algebra code.} Program methods that were used in the course of applying
+the FD method were copied, renamed and re-written to follow the CS formula.
+Because of restrictions imposed by the container class being used, this
+involved changing the data type of intermediate variables which were locally
+scoped, and dynamically casting persistent memory variables to complex data
+types where necessary. \note[Michael]{re: I.3}\add[Michael]{For the reader's
+convenience, 'locally scoped' in the C++ programming language refers to
+data which is created within a context more specific than the context of
+the entire program, or 'translation unit'.  Most data created this way is
+destroyed when that specific context is left. The topic of contexts in C++
+is very broad, so for the purposes of this work context refers to a program
+method or function that begins, concludes and yields a result, also
+destroying locally scoped data. Dynamic casting for the purposes of this
+work refers to copying the value of data contained within an \emph{Epetra}
+vector into locally scoped data that has the property of following the rules of
+complex arithmetic, where a mapping called 'type casting' allows the
+translation of data between different types.} No attempt was made to ``tune''
+the CS code in any way for performance. The algorithm  was implemented in such
+as way that it was in one-to-one correspondence with the FD and CD methods;
+however, it must be noted that the overhead associated with the dynamic casts
+in the CS method likely hurts its overall performance and this issue could be
+alleviated by using the next generation of templated \emph{TPetra} vectors
+available in \emph{Trilinos} that are capable of taking an explicit
+instantiation of any data type including {\tt complex}. \note[Michael]{re:
+I.3}\add[Michael]{For the purposes of this work, an explicit instantiation
+refers to requesting the creation of an object, or instance, of a templated
+class or container class that holds data which has the property of
+following desired mathematic or logic rules.} The version of \emph{Peridigm}
+modified to implement CS can be found in the paper repository.
 
@@
 many advanced techniques in C++ such as multiple virtual inheritance, pointer
-arithmetic, and distributed data structures were used in the basic code and by
-necessity in the modified code.  To see CS and other methods in a more narrow
-context, it is advised that readers also take a look at the simple examples
-provided on the corresponding author's website.
+arithmetic, and distributed data structures, \note[Michael]{re:
+I.3}\add[Michael]{for example \emph{Epetra} vectors, were used in the basic
+    code and by necessity in the modified code. These concepts are not vital to
+understanding the complex step method iteself, but may be necessary to understand complex
+step in \emph{Peridigm}.} To see CS and other methods in a more narrow context,
+it is advised that readers also take a look at the simple examples provided on
+the corresponding author's website.
 
@@
 specific parameters of these tests can be found in \emph{Perdigm} input {\tt
-xml} files also included in the paper repository.
+xml} files also included in the paper repository. \note[Michael]{re:
+I.3}\add[Michael]{For the purposes of this work message passing is what
+synchronizes multiple computers that each are working on a separate section
+of the problem domain. This operation involves the time-consuming use of
+the data network. To reduce the proportion of simulation time spent on
+message passing, good practice is to maximize the amount of information
+contained within a subdomain solely managed by a single computer, relative
+to the amount of information shared accross multiple subdomains and
+therefor multiple computers. A useful experiment to illustrate this idea
+concretely would be to implement a code to solve a 2D Laplace equation for
+simulating a simple heat condution problem on two processors. Use the
+textbook finite difference method, which in this case does not refer just
+to the differentiation technique.  First decompose the domain into two
+contiguous chunks, having one shared border, and distribute these chunks,
+or subdomains to the two processors. Perfom the simulation a multitude of
+times to calculate an average runtime.  Next implement the solution a
+different way, instead partitioning the problem domain into many narrow,
+alternating strips, and distribute those to again two processors, where adjacent
+strips in the graph of the domain should be sent to different processors. This way
+of decomposing the domain will cause much overlap between the subdomains
+managed by the two processors and much time will be spent synchronzing the
+subdomain overlap or 'ghost nodes' between the processors. Run the modified
+simulation a multitude of times and take the average runtime. The larger
+the unpartitioned domain, the worse the discrepancy should be.}
 
@@
 large memory} nodes each having one terabyte of local RAM. This large amount of
-memory was required for reasons explained in Section~\ref{JGAM}.
+memory was required for reasons explained in Section~\ref{JGAM}, second paragraph.
 
@@
 require compiling a version of the AD library being used that were compatible
-with the accelerator or GPU being used. What this means is that if
-\emph{Peridigm} or any simulation software were to be reformulated to take
-advantage of accelerators or GPUs, the flexibility of complex-step owed to its
-simple implementation may make it more viable for that application given its
-high accuracy. Additional performance gains may be achieved through the use of
-the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
+with the accelerator or GPU being used. \note[Michael]{re:
+I.3}\add[Michael]{For the reader's convenience, a collection of contiguously
+stored data elements are said to be 'byte-copiable' when the the interval
+in bytes of memory spanned by a data element in the collection is constant
+throughout the collection. Suppose 'contiguously stored' means, stored in
+memory with a minimum of unrelated infomation stored in between data
+elements in the collection. Complicated objects, such as AD datatype
+elements, do not have the same size in memory as other objects even of that
+same class, for the reason that they themselves contain an uncertain
+ammount of state data according to their design and work done in the
+program. This means that a collection of such objects, even if they are
+contiguously stored in memory, will not be alligned according to a constant
+interval, such that elements within a collection cannot be located by
+coordinates consisting of the location of the beginning of the first
+element in the collection and an interval multiplied by a natural number.
+This method of locating is integral to the serialized communication
+implemented by MPI the libraries as well as CUDA, OpenCL, and others,
+which allow only the transmission of collections of byte-copiable datatypes
+(that is 'primitive datatypes'). A common method of bypassing this
+limitation for communication purposes is to write the complicated data to a
+'character buffer', and transmit the byte-copiable character buffer,
+however this is useless unless the recipient device of the communication
+can recover the original data and process it, which is impossible if the
+recipient device only can work with a subset of the programming language
+that the sending device must use, such as in a communication between a CPU
+running C++ and a GPU running C (such as in CUDA programs), where the gap in
+sophistication contains the features of the "complicated" datatype.} What this means is
+that if \emph{Peridigm} or any simulation software were to be reformulated to
+take advantage of accelerators or GPUs, the flexibilty of complex-step owed to
+its simple implementation may make it more viable for that application given
+its high accuracy. Additional performance gains may be achieved through the use
+of the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
 use of complex data types natively, and alleviate the expense of the

### 2014-02-17 response/response.tex
updated response for I.3

@@
         not specialized to Peridigm, the authors should include a short section introducing
-        the peridynamic theory and the speciﬁc constitutive model used, i.e., the elastic
+        the peridynamic theory and the specific constitutive model used, i.e., the elastic
         peridynamic solid. You may add this to the already included very brief description
@@
     \item 
-        We have revised the brief desciption of peridynamics in page 9 to be more complete:
+        We have revised the brief description of peridynamics in page 9 to be more complete:
     \item
@@
   \end{itemize}}
+    \item 
+        The manuscript seems to be written in a language more familiar to computer scientists. 
+        I would encourage the authors to revise certain parts of the manuscript to be
+        presented in a language more familiar to the computational mechanics community.
+
+{\color{red}
+  \begin{itemize}
+    \item 
+        We have revised certain parts of the manuscript for clarity, adding additional information for the reader's
+        convenience:
+    \item
+        In subsection 'Implementing the complex-step method in Peridigm', explanation of what is meant by datatype,
+        container class, template, scope and instantiation are included since these concepts may be unfamiliar to
+        a computational mechanics audience at least as denoted by these terms. 
+    \item 
+        In subsection 'The Comparative Study', the way message passing computations can slow runtime is explained
+        by an analogy to an experiment given as an exercise for the reader. 
+    \item 
+        In section 'Conclusions', the term byte-copiable is explained. 
+    \item 
+        It is the authors' opinion that the above revisions to the language of the manuscript allow the intended audience
+        to follow the discussion where concepts from computer science are relevant.
+  \end{itemize}}
 

### 2014-02-20 jacobian_calc_CMAME.tex
added solution accuracy results and convergence studies

@@
 with Newton's method.
+
+\section{Working examples of CS, AD and FD for Newton's method}
+\label{sec:WorkingExamples}
+In order to illustrate the complex-step technique in a simpler form than it
+appears in the \emph{Peridigm} code written for this study, an example
+header-only library and two example problem definition programs that use it
+have been created. All the materials described here are available in the
+associated project repository.
+
+\subsection{Common code} 
+\label{subsec:CommonCode}
+
+The source code for the header-only library is entitled
+\emph{NewtonRaphson.hpp}. Note that the library is an interface to functions and
+data structures made available in the \emph{Trilinos} library, mentioned
+earlier, such that it is required in order for the examples described here
+to work. \emph{Trilinos} is freely available for download as of this
+writing at: \emph{trilinos.sandia.gov}.  Generally the header-only library
+works by taking in the users initial guessed equilibrium solution, target
+dependent variable values, system constants, and a function pointer to a
+model evaluation function. Solver parameters including an update scaling
+factor, tolerance, maximum iterations and probe length are chosen after the
+solver is instantiated as a new object. The behavior of the solver is
+specialized through a coding feature called inheritance to perform each of
+CS, AD and FD.  But, in common, each version of the solver works by
+computing the system Jacobian, solving a linear system as in that step of
+the Newton Raphson method, computing the residual and checking for
+convergence or overrun. Here the linear solution method is LU
+decomposition, while the chief convergence criterion is change in residual
+between iterations.  Additionally the problem can be partially respecified
+to allow for running a series of load steps having progressing target
+values, which is a typical mode also seen in \emph{Peridigm}. Applying
+boundary conditions incrementally in load steps assists convergence
+\cite{rezaiee2010dynamic}.
+
+\subsection{First example}
+\label{subsec:ProgOne}
+The first example program, \emph{NRexamples.cpp}, solves a simple nonlinear
+system describing the intersection of three infinite paraboloids in threespace.
+It works by defining force computation functions for each of the models, and
+then passing an address called a function pointer corresponding to each of
+these methods to a solver object which coordinates solving the problem and
+reporting output for the user based on their convergence criteria and other
+preferences like step-size and maximum number of iterations. CS, AD and FD are
+each represented and the user is free to adjust input settings by specifying
+command line arguments as the compiled program instructs, or perhaps create a
+shell script that varies parameters for each of the methods in order to perform
+a comparative study.
+
+\subsection{Second example and related studies} 
+\label{subsec:Validation}
+The second program,a \emph{NRComparison.cpp}, follows a case study that appears in
+\cite{rezaiee2010dynamic}. The example program departs from the cited
+literature by applying each of CS, AD and FD and verifying them against an
+analytical model of a nonlinear mechanical system involving a connected beam
+and spring. Fig. \ref{fig:TrussSchematic} is a schematic of the beam and spring
+system.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=8cm]{./figs/truss.png}
+  \caption{Beam and spring simulated in example program.}
+  \label{fig:TrussSchematic}
+\end{figure}
+
+The equations describing this system and results for verification were obtained
+in \cite{rezaiee2010dynamic}. They include a force function, Eqn.
+\ref{eqn:TrussForce}, and an analytical force derivative or stiffness, Eqn.
+\ref{eqn:TrussStiffness}.
+
+\begin{equation} 
+    \label{eqn:TrussForce}
+    f(D) = .5AE(cos^{2}\phi)(\frac{D}{L_{0}})^{2}[\frac{D}{L_{0}}cos^{2}\phi - 3sin\phi] + k_{s}D + (AE\frac{D}{L_{0}})sin^{2}\phi,
+\end{equation}
+
+\begin{equation} 
+    \label{eqn:TrussStiffness}
+    S_{T}(D) = 1.5AE(cos^{2}\phi)[\frac{D}{L_{0}}cos^{2}\phi - 2sin\phi](\frac{D}{L_{0}^{2}}) + k_{s} + \frac{AEsin^{2}\phi}{L_{0}}.
+\end{equation}
+
+Where $\phi$ is a constant, beam angle, $AE$ is rigidity, $D$ is displacement,
+$L_{0}$ is original beam length, and $k_{s}$ is the Hooke's constant for the
+coil spring attached to the beam. For a verification of each of the methods as
+well as a convergence rate study, this truss and spring model was analyzed in
+the second program. $f(D)$ is restorative force to balance downward force
+loading at the connection point of the beam and the spring, parallel to the
+wall. As in the reference, equilibrium solutions for 12 incremental load steps
+of 4.448 Newtons were sought. Differing from the reference, Newton's method was
+used rather than 'Dynamic Relaxation'. Solver parameters common to all tests
+were a limit of 100 iterations for convergence, a residual change tolerance of
+$1\times{10^{-9}} \si{\pico\meter}$ and an update multiplier of 1.0. Data
+collected included number of iterations to converge summed over all load-steps
+and final displacement, D, after the final load-step. Additionally
+finite-difference step-size was varied in order to measure the impact on each
+of the Jacobian calculation methods. Naturally, varying step size does not
+affect AD or the analytically derived Jacobian. The accuracy results for
+verifying the Jacobian calculation methods are summarized in Table
+\ref{tab:ConvergenceStudy}.
+
+\begin{table*}[!ht]    
+  \centering
+        \caption{Displacement, D, after load-step 12} \label{tab:Verification}   
+        \begin{tabular}{c c c c c}
+         \toprule
+         Step exponent & Actual $\si{\centi\meter}$ & AD $\si{\centi\meter}$ & CS $\si{\centi\meter}$ & FD $\si{\centi\meter}$\\ 
+        \midrule
+        0 & 5.07996 & " & " & "\\
+        -4 & 5.07996 & " & " & "\\
+        -8 & 5.07996 & " & " & "\\
+        -12 & 5.07996 & " & " & " \\
+        -16 & 5.07996 & " & " & failed \\
+        -20 & 5.07996 & " & " & failed \\
+        \bottomrule
+    \end{tabular}
+\end{table*}
+
+Table \ref{tab:Verification} shows that each of the methods solved the problem
+as accurately as one another within the precision displayed on the table. The
+exception to this is that the finite difference method failed due to an
+arithmetic error in association with extremely small step size, suggesting
+subtractive cancellation. However, since there is no accuracy reason for
+choosing an extremely small step size, each of the methods can be used to
+produce acceptable solutions to the nonlinear truss system. Table
+\ref{tab:ConvergenceStudy} shows the total number of iterations taken over all
+load-steps for each of the methods.
+
+\begin{table*}[!ht]    
+  \centering
+        \caption{Summed iterations over all load steps} \label{tab:ConvergenceStudy}   
+        \begin{tabular}{c c c c c}
+         \toprule
+         Step exponent & Actual $\si{\centi\meter}$ & AD $\si{\centi\meter}$ & CS $\si{\centi\meter}$ & FD $\si{\centi\meter}$\\ 
+        \midrule
+        0 & 63 & " & 213 & 417\\
+        -4 & 63 & " & " & "\\
+        -8 & 63& " & " & "\\
+        -12 &63& " & " &  65\\
+        -16 &63& " & " & fail\\
+        -20 &63& " & " & fail\\
+        \bottomrule
+    \end{tabular}
+\end{table*}
+
+Table \ref{tab:ConvergenceStudy} indicates that when using AD or the 'actual'
+analytically derived Jacobian, a minimum number of iterations is achieved
+corresponding to a maximum convergence rate. CS and FD are similarly affected
+by a grossly large step-size, while subtractive cancellation invalidates the
+results for FD for very small step-sizes. Inaccuracy for large step-sizes is is
+predicted by the FD and CS formulas, which are contingent upon a small
+step-size to allow the neglect of higher order 12 terms. The results of this
+test show that step-size selection is not unimportant when using FD, but when
+properly configured any of the methods may return the same maximum convergence
+rate for this problem.  Overall conclusions regarding the methods that may be
+kept in mind for problems besides this example are that the accuracy of the
+Jacobian should not be confused with the accuracy of the equi- librium
+solution, since each of the methods discussed here may be equally capable given
+enough iterations. Additionally, while the accuracy of the Jacobian may
+strongly affect convergence rates, this relationship was not quantified here.
 

### 2014-02-20 jacobian_calc_CMAME.tex
beginning to address I.5

@@
 
+\note[Michael]{re: I.5, this is taken from the thesis.}
+\add[Michael]{
 \section{Working examples of CS, AD and FD for Newton's method}
@@
 of the Jacobian calculation methods. Naturally, varying step size does not
-affect AD or the analytically derived Jacobian. The accuracy results for
+affect AD or the analytically derived Jacobian since those are not functions
+of step size. The accuracy results for
 verifying the Jacobian calculation methods are summarized in Table
@@
 results for FD for very small step-sizes. Inaccuracy for large step-sizes is is
-predicted by the FD and CS formulas, which are contingent upon a small
+predicted by the FD and CS formulas, which require a small
 step-size to allow the neglect of higher order 12 terms. The results of this
@@
 properly configured any of the methods may return the same maximum convergence
-rate for this problem.  Overall conclusions regarding the methods that may be
-kept in mind for problems besides this example are that the accuracy of the
-Jacobian should not be confused with the accuracy of the equi- librium
+rate for this problem.  Overall conclusions regarding the methods that the accuracy of the
+Jacobian should not be confused with the accuracy of the equilibrium
 solution, since each of the methods discussed here may be equally capable given
-enough iterations. Additionally, while the accuracy of the Jacobian may
-strongly affect convergence rates, this relationship was not quantified here.
+enough iterations. Numerical measurements of accuracy besides number of iterations taken
+are not possible here because there is not basis for comparison, as the methods are solving
+the problem separately and producing difference Newton iterates. Additionally, while the inaccuracy of the Jacobian may
+strongly affect convergence rates, this relationship was not precisely quantified here.
+}
 
@@
 
+\note[Michael]{Re: I.5}
+\add[Michael]{
+\section{Convergence and displacement accuracy studies} 
+\label{sec:PeridigmConvergenceStudy}
+%
+While the previous sections profiled the per-iteration performance of each of
+the methods, it was necessary to conduct convergence rate studies in order to
+get a complete picture. It is necessary for an analyist to know if the
+accuracy differences measured in the per-iteration tests would amount to
+differences in convergence rates among the methods if they were allowed to solve problems
+normally. The question is posed, given that step-size $h$ is heuristically
+selected by \emph{Perdigm} for the finite difference methods and presumable
+makes a good choice, do the single-iteration Jacobian accuracy results predict
+the ranking of the methods in terms of convergence rate for a given study? For
+brevity, convergence rate was measured for each of the methods in solving just
+the largest problem of the single-core test series. As was seen in the per-iteration results,
+Jacobian accuracy did not vary greatly from test to test, such that any pair of tests
+from the two series could have been considered reasonably representative. As in
+the validation problem shown in ~\ref{subsec:Validation}, would each of the methods return an
+equal convergence rate so long as $h$ were optimally selected?
+
+The results shown in Table ~\ref{tab:ConvergenceStudy2} indicate that this was the case. 
+The difference in Jacobian accuracy between the methods seen in the per-iteration tests
+was not sufficient to lead to differences in the convergence rate of the methods.
+
+\begin{table*}[!ht]   
+    \centering \caption{Average iterations/ load-step} 
+    \label{tab:ConvergenceStudy2}   
+    \begin{tabular}{c c c c c}
+    \toprule Test type & AD & CS & CD & FD\\
+        \midrule 
+        single-core & 63 & " & " & "\\ 
+        multi-core  & 63 & " & " & "\\ 
+\bottomrule \end{tabular} \end{table*}
+
+The end displacement results shown in ~\ref{subsec:Validation} indicated that up to a certain precision the
+solutions produced by each of the methods were identical. It was examined if this was the case for the 
+solutions produced in \emph{Peridigm}. Here each of the solutions produced by each of the 
+methods were compared to the AD produced solutions, under the assumption that the AD produced solutions
+would be closest to the solution produced by using an 'exact' Jacobian. Table ~\ref{tab:PeridigmSolutionAccuracy} shows that the methods
+did indeed produce differing solutions in terms of nodal force density, although only in decimal places at least $9$ orders 
+of magnitude smaller than the force density value at the node where the difference was detected.
+
+\begin{table*}[!ht]   
+    \centering \caption{Nodal force density maximum difference} 
+    \label{tab:PeridigmSolutionAccuracy}   
+    \begin{tabular}{c c c c}
+    \toprule  & CS & CD & FD\\
+        \midrule  Order of magnitude & $\num{1.E-8}$ & $\num{1.E-8}$ & $\num{1.E-5}$ \\ 
+\bottomrule \end{tabular} \end{table*}
+
+The solution accuracy results do follow the same ranking seen in the Jacobian accuracy results, however,
+again as in the validation problem the results differ only slightly. Speculatively this degree of error
+may be important to those running probabilistic simulations such that the propagation of error may be 
+a concern. Also it is conceivable that if this error is a fixed quantity, insensitive to material 
+properties or loading conditions, it may have a greater impact on simulations involving very low
+force densities such as for small displacements or very lax materials. In this case it could be recommended
+that forward-difference be avoided.
+
 \section{Conclusions}
 %
+\subsection{For further studies of Jacobian accuracy in Newton's method}
+\label{subsec:FurtherStudies}
+The question is raised; what is the significance of Jacobian estimation
+accuracy to the convergence of Newton's method?  The per-iteration and even
+convergence studies shown here indicate the effect of Jacobian accuracy for
+only a single datum, a very accurate Jacobian estimate, while the validation
+problem indicates the effect of using a grossly inaccurate Jacobian. The results here show that
+for very small Jacobian estimate inaccuracies, convergence rate is not
+affected, while for grossly inaccurate Jacobians Newton's method fails. 
+Is it worthwhile then to investigate the effect of using moderately inaccurate Jacobians, 
+produced deliberately or otherwise, on the convergence rate of Newton's method?
+
+The answer may be no at least unless a more highly non-linear class of problem
+is selected for study or a controllable method of producing a moderately
+inaccurate Jacobian is found along with a reason for using it. While it is possible
+to incorrectly apply finite-differencing and CS by choosing an inappropriate
+step-size, there is no reason that this should be done. The experiments show
+that step-size $h$ can be properly selected and that the methods are capable of
+producing accurate Jacobians. It should be noted that the model used to perform
+the simulation could always be substituted for a more highly non-linear one and
+perhaps this would somewhat increase the importance of Jacobian accuracy. The
+study presented here failed to pose the question of how wrong a correctly
+implemented Jacobian calculation scheme could go for plastic or otherwise
+non-linear material models, here instead non-linearity was imposed by the
+nonlocal interaction between material points essential to peridynamic models,
+like the linear model used in the study.
+
+Nevertheless, Newton's method is not defined for using an inaccurate or 'wrong'
+Jacobian, so using an inaccurate Jacobian is failing to perform Newton's method
+by definition. But it is also true that it should not be expected that every
+differentiation technique performs flawlessly under all situations. The
+pragmatic strategy to study this may be to compare a prospective method's
+Jacobian accuracy to that produced by AD and develop a rule of thumb based on
+observed convergence rates from prior simulations as they relate to the percent
+difference of the prospective and AD produced Jacobians. Also it is possible to
+calculate a sensitivity of Jacobian error to step-size alongside a sensitivity
+of convergence rate to step-size, and then to relate these two quantities as a
+map of Jacobian error to convergence rate change from ideal. Additionally it is
+still possible to look at the convergence conditions for Newton's method and
+suggest a relationship between Jacobian accuracy and convergence rate in a
+theoretical manner. To this point, there are a set of convergence conditions derived in
+\cite[Chap. 2, p.41]{burkeLectures} which relate quality of Jacobian inverse approximations
+to convergence rate classification, here reproduced,
+%
+\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq K,
+\end{equation}
+
+for worse than linear convergence,
+
+\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq \Theta_{1}^{k}K; \Theta_{1} \in (0,1),
+\end{equation}
+
+for linear convergence,
+
+\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq min{M_{2} \mid \mid x_{k} - x_{k-1} \mid \mid, K}; M_{2} > 0
+\end{equation}
+
+for two step quadratic,
+
+\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq min{M_{3} \mid \mid r(x_{k}) \mid \mid, K}; M_{3} > 0
+\end{equation}
+
+and lastly for quadratic convergence. Here $r \prime$ is the true system
+Jacobian and $J$ is the approximation. Essentially the conditions measure how
+well the difference of the true Jacobian inverse and approximation works as
+a contraction map within the closure of a radius of convergence about the true
+solution. This property is necessary for the solution method using the
+approximation to converge rather than to diverge, as is true of the exact Newton
+method.
+
+It may be interesting to artificially perturb the Jacobian calculations with
+random error in order to better control experimentally the onset of inaccuracy
+to develop a sensitivity of convergence rate to this type of error. Perhaps an
+analyst could use this knowledge to devise an adaptive differentiation
+technique selection method which economizes on accuracy until a significant
+Jacobian error is measured during a special doubly calculated iteration.
+
+Additionally, there are Quasi-Newton methods such as Broyden's method which
+uses an approximate inverse Jacobian and updates it with information from the 
+iterate and residual \cite{dennis1971convergence}. A method called
+Incomplete Jacobian Newton's method may the closest to the idea of deliberately
+making use of an inaccurate Jacobian in the sense that arbitrary elements of
+the full Jacobian are set to zero to reduce computational effort
+\cite{liu2008incomplete}. The Quasi-Newton methods represent a compromise of
+Newton's method for the promise of faster iterations in exchange for a lower
+convergence rate. For both of these methods, the iterate and residual are
+computed as in Newton's method. As to the question of algorithmic consistency,
+these numerical methods evaluate the same linearized constitutive law and
+governing equation as for the exact Newton's method which in the case of a
+plastic simulation are involved in iteratively determining algorithmic tangent moduli. 
+In fact, a secant equation based method is appropriate for Newton's method with algorithmic
+tangent moduli since the Jacobian ought only depend on known converged
+values such that the Newton iterate could not artificially drive the plasticity
+of the model \cite[Chap. 6]{belytschko1999nonlinear}. It should be noted that
+constraints or a line search algorithm to minimize the residual are necessary
+whereas the pure secant method does not return unique solutions in
+multi-dimensions.
+%
+
+\subsection{Grain of salt for speed results}
+The per-iteration speed results shown here indicate that the AD method is
+faster than CD or FD.  While this may be the case for the implementation in the
+stock version of \emph{Peridigm} this is not reflective of any limitation of
+these methods. In a side experiment, the finite-difference code in
+\emph{Peridigm} was re-written to take advantage of the benefits of storing the
+results of duplicated intermediate calculations. These modifications were 
+general, low memory footprint and could be adopted in any numerical simulation
+simulation software as good practice. An informal performance test was done
+using a personal AMD Phenom-2 X4 workstation using GCC 4.7 compilers rather than a
+Intel Xeon E5 node with Intel Cluster Studio 13.0 compilers as provided by TACC. The problem
+being run was the largest single core test from the main study, instead run on four 
+compute cores. Table ~\ref{tab:PerformanceMod} shows the timing
+results for AD as reference along with an unmodified CD, and then modified CD
+and FD.
+
+\begin{table*}[!ht]   
+\centering \caption{Total time spent computing Jacobian} 
+\label{tab:PerformanceMod}   
+\begin{tabular}{c c c c c}
+\toprule & AD & CD & Modded CD & Modded FD\\
+\midrule  Total seconds & $~430$ & $~570$ & $~430$ &$~320$ \\ 
+\bottomrule \end{tabular} \end{table*}
+
+The results indicated that in exchange for a modest programming effort, a
+performance increase of around a quarter was available for the
+finite-difference methods and that FD could eclipse AD. What allowed the performance modifications to
+enhance the finite-difference methods was their lack of sophistication
+compared to AD in that their explicitly laid out and
+repetitive code could be readily optimized.
+
+\subsection{Remarks}
 Based on the experimental results, it appears that for users of \emph{Peridigm}
-there is a case to be made for using AD for reasons of performance and
-accuracy. However, certain parallel tests done in this study showed a small
-performance gain over AD with the FD method.  While the results showed that CS
-produced more accurate tangent-stiffness matrices than CD and FD under the
-parameters of the tests, it was not determined whether or not this is a clear
-advantage of CS over FD in terms of accuracy of final predicted displacements
-and speed of convergence. The reader is reminded that the time required to
-complete an iteration must be considered along with the total number of
-iterations taken order to measure speed of convergence. The goal of the
-comparative study was to benchmark single iteration performance and not
-convergence rate. For new developments it is recommended that combined accuracy
-and speed of iteration measurements be conducted along with entirely separate
-convergence studies. It should be mentioned that a related work,
-\cite{perez2012numerical}, compares the speed and accuracy of complex-step to
-finite difference for producing tangent-stiffness matrices in solid mechanics
-simulations using a complex inelastic material model, where the researchers
-similarly found CS to be accurate yet relatively computationally expensive
-compared to FD and CD. In that work, full convergence studies were also done
-and indicated that quadratic convergence could be achieved with complex-step
-\remove[Michael]{all of the methods} \cite[p.28]{perez2012numerical}.
-
+there is a case to be made for using AD for reasons of speed. However, certain
+parallel tests done in this study showed a small performance gain over AD with
+the FD method.  The results showed that CS produced more accurate
+tangent-stiffness matrices than CD and FD under the parameters of the tests,
+however it was determined that this difference in accuracy was not sufficient
+to alter convergence rate for a test problem. In order for Jacobian accuracy to
+have an impact on convergence rate, and therefor serve as a predictor of
+convergence rate in per-iteration results, the amount of error detected between
+a supposed perfectly accurate Jacobian and the test Jacobian needs to be
+greater than allowed by the conditions seen in \cite{burkeLectures}, 
+reproduced in \ref{sec:FurtherStudies}.
+
+It should be mentioned that another work, \cite{perez2012numerical}, compares
+the speed and accuracy of complex-step to finite difference for producing
+tangent-stiffness matrices in solid mechanics simulations using the 'MRS-Lade'
+material model, where the researchers similarly found CS to be accurate yet
+relatively computationally expensive compared to FD and CD. In that work, full
+convergence studies were also done and indicated that quadratic convergence was
+achieved with the methods \cite[p.28]{perez2012numerical}.
+
+However, at least for the models producing the results shown here, each of the
+methods returned quadratic convergence, indicating each was sufficiently
+accurate. 'Marching orders' for the analyst would be to select the swiftest
+method, while it would seem accuracy above what is adequate does not enhance
+the convergence rate or iteration speed until either classical Newton's convergence
+conditions or the approximate case conditions are violated. 
+}
 The CS method was shown to be highly accurate in calculating tangent-stiffness
@@
 dependence on an external AD library.  From this perspective, the method may
-still be an attractive method for TS matrix evaluation.
+still be an attractive method for TS matrix evaluation. \add[Michael]{However, at least for 
+the models producing the results shown here, each of the
+methods returned quadratic convergence, indicating each was sufficiently
+accurate. 'Marching orders' for the analyst would be to select the swiftest
+method, while it would seem accuracy above what is adequate does not enhance
+the convergence rate or iteration speed until either classical Newton's convergence
+conditions or the approximate case conditions are violated.}
 

### 2014-02-20 jacobian_calc_CMAME.tex
made additions render

@@
-%%
-%% Copyright 2007, 2008, 2009 Elsevier Ltd
-%%
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%%
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%%
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
-
-%% Template article for Elsevier's document class `elsarticle'
-%% with numbered style bibliographic references
-%% SP 2008/03/01
-%%
-%%
-%%
-%% $Id: elsarticle-template-num.tex 4 2009-10-24 08:22:58Z rishi $
-%%
-%%
 \documentclass[preprint,12pt]{elsarticle}
@@
 
-\note[Michael]{re: I.5, this is taken from the thesis.}
-\add[Michael]{
+\note[Michael]{re: I.5, next section is taken from the thesis.}
 \section{Working examples of CS, AD and FD for Newton's method}
@@
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=8cm]{./figs/truss.png}
-  \caption{Beam and spring simulated in example program.}
-  \label{fig:TrussSchematic}
+\begin{figure}[tbp] \centering \includegraphics{./figs/truss.png}
+\caption{Beam and spring simulated in example program.} \label{fig:TrussSchematic}
 \end{figure}
@@
 strongly affect convergence rates, this relationship was not precisely quantified here.
-}
 
@@
 
-\note[Michael]{Re: I.5}
-\add[Michael]{
+\note[Michael]{Re: I.5, everything after this line is changed}
 \section{Convergence and displacement accuracy studies} 
@@
 greater than allowed by the conditions seen in \cite{burkeLectures}, 
-reproduced in \ref{sec:FurtherStudies}.
+reproduced in Subsection~\ref{subsec:FurtherStudies}.
 
@@
 conditions or the approximate case conditions are violated.
-}
+
 The CS method was shown to be highly accurate in calculating tangent-stiffness
@@
 specializing the computer's general forms of $u, v, w$ 
-according to \ref{abcvalues}, with parameters $a, b$ chosen for each function and held as constant, and
+according to ~\ref{abcvalues}, with parameters $a, b$ chosen for each function and held as constant, and
 evaluating, such that the total derivative of our original function w.r.t $x$ where $x = x_0$ is approximated

### 2014-02-20 response/response.tex
finished addressing I.5

@@
 
+  \item
+      The comparisons of the different algorithms are only performed with respect to the
+      construction of tangent-stiffness matrices. In par ticular, the authors always use the
+      solution of AD at each time step to compute the entries of the tangent-stiffness
+      matrices, even for the FD, CD, and CS methods. However, the authors need to
+      demonstrate the performance of the CS method, in comparison to the other ones,
+      with respect to the solution of real, ideally non-linear problems. 
+{\color{red}
+  \begin{itemize}
+    \item 
+        To address the reviewers comment, material from work done after the
+        initial submission of the manuscript has been included. New sections
+        include 'Working examples of CS, AD and FD for Newton's method' with
+        new subsections. To existing sections were appended new subsections
+        including 'Convergence and displacement accuracy studies', 'For further
+        studies of Jacobian accuracy in Newton's method', 'Grain of salt for
+        speed results' and 'Remarks'. Specifically, the added material addresses
+        the reviewers comments:
+    \end{itemize}}
+
+\item 
+      In other words, how the solutions compare in terms of accuracy and speed between the methods when
+      solving given problems? Please include two examples of interest. The authors
+      suggest that two examples involving non-linear systems are already included in the
+      paper repository.
+
+{\color{red}
+  \begin{itemize}
+    \item 
+        The added section 'Working examples of CS, AD and FD for Newton's method' includes
+        the two examples alluded to in the manuscript and requested by the reviewer. 
+        % 
+        To summarize, in this section is an example of interest: a displacement accuracy study and 
+        a convergence rate study on a non-linear beam problem, where each of the methods
+        discussed in the manuscript are used individually to solve the problem. 
+ \end{itemize}}
+
+\item
+      The authors say in the Conclusion: “While the results showed that CS
+      produced more accurate tangent-stiffness matrices than CD and FD under
+      the parameters of the tests, it was not determined whether or not this is
+      a clear advantage of CS over FD in terms of accuracy of final predicted
+      displacements and speed of convergence.” This needs to be studied and
+      results in this regard added to this manuscript.
+
+{\color{red}
+  \begin{itemize}
+    \item 
+      The added subsection 'Convergence and displacement accuracy studies'
+      includes convergence rate and displacement accuracy studies similar to
+      those shown in 'Working examples ...', but with the models and tools of
+      the main study. This subsection provides an example of interest. 
+  \item 
+      To summarize, these studies are done in \emph{Peridigm} using the most refined mesh
+      from the 'serial' test runs from the original manuscript. Measured was the performance of the methods in terms of speed
+      (convergence rate) and accuracy (displacement accuracy).  Iteration speed
+      was reported in the original manuscript.
+  \item
+      It was determined that the accuracy of CS over FD resulted in no appreciable
+      advantage in terms of displacement accuracy or convergence rate for the problems
+      solved in this study. 
+    \item
+      Additional comment on the significance of the accuracy of the Jacobian
+      to Newton's method is included in the subsections 'For further studies of
+      Jacobian accuracy in Newton's method', 'Grain of salt for speed results'
+      and 'Remarks'. 
+  \item 
+      To summarize, Newton's method is insensitive to the accuracy differences
+      shown by the methods, for the problems shown, when the methods are not
+      deliberately misused with inappropriate step sizes. For completeness,
+      formulas describing a relationship between Jacobian accuracy and
+      convergence rate order are included. 
+  \item
+      It is the authors' opinion that the above revisions offer the discussion of relative problem solving
+      performance of the methods that the reviewer has requested, including consideration
+      of speed and accuracy in solve real, non-linear problems.
+
+  \end{itemize}}
+
 \end{document}

### 2014-02-20 jacobian_calc_CMAME.tex
finished addressing I.5

@@
 brevity, convergence rate was measured for each of the methods in solving just
-the largest problem of the single-core test series. As was seen in the per-iteration results,
+the largest problem of the single-core test series. That problem was solved with each of the methods
+both on a single core and all sixteen cores of a 'normal' node on \emph{Stampede}. As was seen in the per-iteration results,
 Jacobian accuracy did not vary greatly from test to test, such that any pair of tests
@@
 affected, while for grossly inaccurate Jacobians Newton's method fails. 
+Additional confounding factors include that a Jacobian is only first
+order approximation of an often curved energy surface and that a non-naive implementation of Newton's
+method includes a line search phase that helps in choosing an optimal iterate given a Jacobian.
 Is it worthwhile then to investigate the effect of using moderately inaccurate Jacobians, 

### 2014-02-20 response/response.tex
gathering info for lit review section

@@
       with respect to the solution of real, ideally non-linear problems.
+
 {\color{red}
@@
       performance of the methods that the reviewer has requested, including consideration
-      of speed and accuracy in solve real, non-linear problems.
+      of speed and accuracy in solving real, non-linear problems.
+  \end{itemize}}
 
+  \item
+     In Section 1.1.1, please include a more comprehensive literature review for
+     the CS method, including fur ther references (some of them are listed below).
+     In particular, specify what aspects of the CS method have been investigated in
+     each reference. Then, you need to clearly mention your contribution within
+     the CS method, in contrast to the other references (some comments in this
+     regard appear in the manuscript in page 8).
+
+{\color{red}
+  \begin{itemize}
   \end{itemize}}

### 2014-02-20 response/candc.tex
gathering info for lit review section

\documentclass[10pt,letterpaper,draft]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\author{Michael D. Brothers}
\title{Comparing and contrasting with prior work in the field}
\begin{document}
\section{Abstract of the reviewed work}

In order to maintain the quadratic convergence properties of Newton's method in
quasi-static non-linear analysis of solid structures it is crucial to obtain
accurate, algorithmically consistent tangent-stiffness matrices. A goal of the
study described in this paper was to establish the suitability of an
under-explored method for numerical computation of tangent-stiffness operators,
referred to as ``complex-step'', and compare the method with other techniques
for numerical derivative calculation: automatic differentiation, forward
finite-difference, and central finite-difference. The complex-step method was
implemented in a massively parallel computational peridynamics code for the
purpose of this comparison.  The methods were compared through invasive
profiling of the code for accuracy, speed, efficiency, and parallel
scalability. The research provides data that can serve as practical guide for
code developers and analysts faced with choosing which method best suits the
needs of their application code.  Additionally, motivated by the reproducible
research movement, all the of the code, examples, and workflow to regenerate
the data and figures in this paper are provided as open source.

\section{Re: W. Squire and G. Trapp (1998)}
\subsection{Abstract of prior work in field}

A method to approximate derivatives of real functions using complex variables
which avoids the subtractive cancellation errors inherent in the classical
derivative approximations is described. Numerical examples illustrating the
power of the approximation are presented.

\subsection{Aspects of CS investigated}

Establishing the accuracy of the method for computing first derivatives of
real functions through empirical tests, versus forward and central difference 
methods.

\subsection{Similarities with review work}

The reviewed work uses the numerical differentiation method as shown in the
prior work. The accuracy of CS demonstrated in the prior work served as encouragement
for investigating CS further.

\subsection{Differences with review work}

In the reviewed work the CS method is used to compute derivatives of a
non-linear vector function of vector variables wrt the components of the vector
variables, an extension of the method shown in the prior work.
Additionally, automatic differentiation is compared to CS. Beyond that, not
only is the accuracy of the method measured, but its impact as part of a Newton
Raphson solution scheme for a Peridynamics simulation run on a parallel
computer is investigated.

\subsection{Novelty of reviewed work}
The reviewed work examines applications of the CS method beyond taking derivatives alone,
whereas the prior work does not.

\section{Re: A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta \emph{n.d. for local...} (2000)}
\subsection{Abstract of prior work in field}

In this paper, numerical differentiation is applied to integrate plastic
constitutive laws and to compute the corresponding consistent tangent
operators. The derivatives of the constitutive equations are approximated by
means of difference schemes. These derivatives are needed to achieve quadratic
convergence in the integration at Gauss-point level and in the solution of the
boundary value problem.  Numerical differentiation is shown to be a simple,
robust and competitive alternative to analytical derivatives. Quadratic
convergence is maintained, provided that adequate schemes and stepsizes are
chosen. This point is illustrated by means of some numerical examples.

\subsection{Aspects of CS investigated}

The prior work applies CS in computing derivatives needed for integrating
plastic stress, obtaining algorithmically consistent elastoplastic modulii, and
obtaining the Jacobian of the residual, for FEA simulations of two plastic
material models.

\subsection{Similarities with review work}

First order CS is being used to compute the Jacobian of the residual in a solid mechanics.

\subsection{Differences with review work}

The prior work demonstrates CS on a serial computer, and with a classical solid mechanics
formulation. Automatic differentiation is not investigated, and speed measurements are
not taken. Scalability studies with a variety of equivalent, progressively more refined meshes
are not done. Scalability as a function of distributed system size is not investigated.

\subsection{Novelty of reviewed work}

While the prior work does apply CS to computing Jacobians of the residual, and further applies CS
to computing elastoplastic modulii, the reviewed work is novel by contrast in offering discussion
of CS as applied to Peridynamics on a parallel computer, a larger range of problem parameters, 
CPU time data, and comparison to automatic differentiation as a competing method.

\section{Re: A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta \emph{n.d. for nontrivial...}(2000)}
\subsection{Abstract of prior work in field}

In Reference [1] the authors have shown that numerical differentiation is a
competitive alternative to analytical derivatives for the computation of
consistent tangent matrices. Relatively simple models were treated in that
reference. The approach is extended here to a complex model: the MRS-Lade model
[2,3]. This plastic model has a cone-cap yield surface and exhibits strong
coupling between the flow vector and the hardening moduli. Because of this,
derivating these quantities with respect to stresses and internal variables
—the crucial step in obtaining consistent tangent matrices— is rather involved.
Numerical differentiation is used here to approximate these derivatives. The
approximated derivatives are then used 1) to compute consistent tangent
matrices (global problem) and 2) to integrate the constitutive equation at each
Gauss point (local problem) with the Newton-Raphson method. The choice of the
stepsize (i.e. the perturbation in the approximation schemes), based on the
concept of relative stepsize, poses no difficulties. In contrast to previous
approaches for the MRS-Lade model, quadratic convergence is achieved, for both
the local and the global problems. The computational efficiency (CPU time) and
robustness of the proposed approach is illustrated by means of several
numerical examples, where the major relevant topics are discussed in detail.

\subsection{Aspects of CS investigated}

The prior work applies CS in computing derivatives needed for integrating
plastic stress, obtaining algorithmically consistent elastoplastic modulii, and
obtaining the Jacobian of the residual, for FEA simulations of the MRS-Lade material
model.

\subsection{Similarities with review work}

CS is used to compute the Jacobian of the residual. CPU runtime is measured.

\subsection{Differences with review work}

The prior work demonstrates CS on a serial computer, and with a classical solid
mechanics formulation. Automatic differentiation is not investigated.
Scalability studies with a variety of equivalent, progressively more refined
meshes are not done. Scalability as a function of distributed system size is
not investigated.

\subsection{Novelty of reviewed work}

While the prior work does apply CS to computing Jacobians of the residual, and further applies CS
to computing elastoplastic modulii, the reviewed work is novel by contrast in offering discussion
of CS as applied to Peridynamics on a parallel computer, a larger range of problem parameters, 
and comparison to automatic differentiation as a competing method.

\section{Re: J. Martins, P. Sturdza, J. Alonso (2003)}
\subsection{Abstract of prior work in field}

The complex-step derivative approximation and its application to numerical
algorithms are presented. Improvements to the basic method are suggested that
further increase its accuracy and robustness and unveil the connection to
algorithmic differentiation theory. A general procedure for the implementation
of the complex-step method is described in detail and a script is developed
that automates its implementation. Automatic implementations of the
complex-step method for Fortran and C/C++ are presented and compared to
existing algorithmic differentiation tools. The complex-step method is tested
in two large multidisciplinary solvers and the resulting sensitivities are
compared to results given by finite differences. The resulting sensitivities
are shown to be as accurate as the analyses. Accuracy, robustness, ease of
implementation and maintainability make these complex-step derivative
approximation tools very attractive options for sensitivity analysis.

\subsection{Aspects of CS investigated}

The prior work demonstrates that the CS method can be performed by using complex
datatypes for the variables of functions comprising an analysis code. The analysis code
can comprise a simulation or similar, and derivatives of response variables called
sensitivities can be computed. The accuracy of this approach is compared to automatic 
differentiation.

\subsection{Similarities with review work}

CS is used to compute derivatives.

\subsection{Differences with review work}

The prior work mentions employing CS within an iterative solver code for a sensitivity analysis, but
does not mention if CS is also used by the iterative solver itself.

\subsection{Novelty of reviewed work}

\section{Re: W. K. Anderson, J. C. Newman, D. l. Whitfield and E. J. Nielsen (2001)}
\subsection{Abstract of prior work in field}

The use of complex variables for determining sensitivity derivatives for
turbulent flows is examined. Although a step size parameter is required, the
numerical derivatives are not subject to subtractive cancellation errors and,
therefore, exhibit true second-order accuracy as the step size is reduced. As a
result, this technique guarantees two additional digits of accuracy each time
the step size is reduced one order of magnitude. This behavior is in contrast
to the use of finite differences, which suffer from inaccuracies due to
subtractive cancellation errors. In addition, the complex-variable procedure is
easily implemented into existing codes.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{A. H. Al-Mohy and N. J. Higham (2010)}
\subsection{Abstract of prior work in field}

We show that the Fréchet derivative of a matrix function f at A in the
direction E, where A and E are real matrices, can be approximated by Im f (A +
ihE)/ h for some suitably small h. This approximation, requiring a single
function evaluation at a complex argument, generalizes the complex step
approximation known in the scalar case. The approximation is proved to be of
second order in h for analytic functions f and also for the matrix sign
function. It is shown that it does not suffer the inherent cancellation that
limits the accuracy of finite difference approximations in floating point
arithmetic. However, cancellation does nevertheless vitiate the approximation
when the underlying method for evaluating f employs complex arithmetic.  The
ease of implementation of the approximation, and its superiority over finite
differences, make it attractive when specialized methods for evaluating the
Fréchet derivative are not available, and in particular for condition number
estimation when used in conjunction with a block 1-norm estimation algorithm.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{W. Jin, B. H. Dennis, and B. P. Wang (2010)}
\subsection{Abstract of prior work in field}

The semi-analytical method (SAM) is a computationally efficient and easy to
implement approach often used for the sensitivity analysis of finite element
models.  However, it is known to exhibit serious inaccuracy for shape
sensitivity analysis for structures modeled by beam, frame, plate, or shell
elements. In the present paper, we use a semi-analytical approach based on
complex variables (SACVM) to compute the sensitivity of finite element mod- els
composed of beam and plate elements. The SACVM combines the complex variable
method (CVM) with the semi-analytical method (SAM) to obtain the response
sensitivity accurately and efficiently. The current approach maintains the
computational efficiency of the semi-analytical method but with higher
accuracy. In addition, the current approach is insensitive to the choice of
step size, a feature that simplifies its use in practical problems. The method
is applicable to any structural elements including beam, frame, plate, or shell
elements and only requires minor modifications to existing finite element
codes.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{A. Voorhees, H. Millwater, and R. Bagley (2011)}
\subsection{Abstract of prior work in field}

Shape sensitivity analysis of finite element models is useful for structural
optimization and design modifications. Complex variable methods for shape
sensitivity analysis have some potential advantages over other methods. In
particular, for first order sensitivities using the complex Taylor series
expansion method (CTSE), the implementation is straightforward, only requiring
a perturbation of the finite element mesh along the imaginary axis. That is,
the real valued coordinates of the mesh are unaltered and no other
modifications to the software are required. Fourier differentiation (FD)
provides higher order sensitivities by conducting an FFT analysis of multiple
complex variable analyses around a sampling radius in the complex plane.
Implementation of complex variable sensitivity methods requires complex
variable finite element software such that complex nodal coordinates can be
used to implement a perturbation in the shape of interest in the complex
domain. All resulting finite element outputs such as displacements, strains and
stresses become complex and accurate derivatives of all finite element outputs
with respect to the shape parameter of interest are available. The
methodologies are demonstrated using two-dimensional finite element models of
linear elasticity problems with known analytical solutions. It is found that
the error in the sensitivities is primarily defined by the error in the finite
element solution not the error in the sensitivity method. Hence, more accurate
sensitivities can be obtained through mesh refinement.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{K.-L. Lai and J. L. Crassidis (2008)}
\subsection{Abstract of prior work in field}

A general framework for the first and second complex-step derivative
approximation to compute numerical derivatives is presented.  For first
derivatives the complex-step approach does not suffer roundoff errors as in
standard numerical finite-difference approaches.  Therefore, since an
arbitrarily small step size can be chosen, the complex-step approach can
achieve near analytical accuracy.  However, for second derivatives straight
implementation of the complex-step approach does suffer from roundoff errors.
Therefore, an arbitrarily small step size cannot be chosen. In this paper the
standard complex-step approach is expanded by using general complex-step sizes
to provide a wider range of accuracy for both the first- and second-derivative
approximations. Even higher accuracy formulations are obtained by repetitively
applying Richardson extrapolations. The new extensions can allow the use of one
step size to provide optimal accuracy for both derivative approximations.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{R. Abreu, D. Stich, and J. Morales (2013)}
\subsection{Abstract of prior work in field}

We generalize the well known Complex Step Method for computing derivatives by
introducing a complex step in a strict sense. Exploring different combinations
of terms, we derive 52 approximations for computing the first order derivatives
and 43 for the second order derivatives. For an appropriate combination of
terms and appropriate choice of the step size in the real and imaginary
directions, fourth order accuracy can be achieved in a very simple and
efficient scheme on a compact stencil. New different ways of computing second
order derivatives in one single step are shown. Many of the first order
derivative approximations avoid the problem of subtractive cancellation
inherent to the classic finite difference approximations for real valued steps,
and the superior accuracy and stability of the generalized complex step
approximations are demonstrated for an analytic test function.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{J. N. Lyness and C. B. Moler (1967)}
\subsection{Abstract of prior work in field}

Many algebraic computer languages now include facili- ties for the evaluation
of functions of a complex variable. Such facilities can be effectively used for
numerical differentiation. The method we de- scribe is appropriate for
computing the derivatives f(n) (x) of any analytic function which can be
evaluated at points in the complex plane near x.  Such a function might be a
complicated rational combination of elementary functions of x which is
difficult to differentiate analytically.  In this paper we derive several
formulas, any of which is suitable for evaluating the nth derivative of a
complex analytic function at a point in terms of function evaluations at
neighborimg points. For simplicity we assume throughout this paper that the
derivatives are to be evaluated at the origin. In this Introduction we state
one formula; in succeeding sections we prove this and related formulas, discuss
their degree and the error in their application, and give an example.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\section{J. N. Lyness (1968)}
\subsection{Abstract of prior work in field}

In a previous paper (Lyness and Moler [1]), several closely related formulas of
use for obtaining a derivative of an analytic function numerically are derived.
Each of these formulas consists of a convergent series, each term being a sum
of function evaluations in the complex plane.  In this paper we introduce a
simple generalization of the previous methods; we investigate the "truncation
error" associated with truncating the infinite series.  Finally we recommend a
particular differentiation rule, not given in the previous paper.

\subsection{Similarities with review work}
\subsection{Differences with review work}
\subsection{Novelty of reviewed work}

\end{document}

### 2014-02-21 response/candc.tex
first draft of lit review done

@@
 
-\subsection{Similarities with review work}
+\subsection{Similarities with reviewed work}
 
@@
 
-\subsection{Differences with review work}
+\subsection{Differences with reviewed work}
 
@@
 
-\subsection{Differences with review work}
+\subsection{Differences with reviewed work}
 
-The prior work mentions employing CS within an iterative solver code for a sensitivity analysis, but
+The prior work mentions employing CS within an iterative solver code within a sensitivity analysis, but
 does not mention if CS is also used by the iterative solver itself.
@@
 \subsection{Novelty of reviewed work}
+
+The reviewed work is about applying CS to solving Peridynamics problems using Newton's method, and is therefor about an
+application of the derivatives produced by CS. The prior work is about the application of CS itself to produce derivatives
+which themselves are the object of interest. The reviewed work is novel by comparison since it examines aspects of CS
+different to those examined in the prior work.
 
@@
 
-\subsection{Similarities with review work}
-\subsection{Differences with review work}
+\subsection{Aspects of CS investigated}
+Based on the accuracy of CS, the method is used to compute sensitivities in CFD
+simulations. For the application of computing sensitivities, CS is compared to 
+automatic differentiation.
+
+\subsection{Similarities with reviewed work}
+CS is used to compute derivatives and CS is compared to AD.
+
+\subsection{Differences with reviewed work}
+The prior work does not discus using the derivatives produced by CS to compute
+the residual Jacobian, nor does it discuss the same for AD. The field is
+fluid dynamics rather than solid mechanics, and the use of a prallel computer
+is not discussed.
+
 \subsection{Novelty of reviewed work}
+
+The reviewed work offers discussion on the use of CS and AD to compute
+residual Jcobians and the effect of doing so on the performance of Newton's
+method in Peridynamics simulations run on a parallel computer, where the prior
+work discusses the derivitives produced by CS and AD themselves as the objects 
+of interest.
 
@@
 
-\subsection{Similarities with review work}
-\subsection{Differences with review work}
+\subsection{Aspect of CS discussed}
+The prior work investigates the use of complex step in computing Frechet derivatives 
+of matix to matrix functions.
+
+\subsection{Similarities with reviewed work}
+To compute the residual Jacobian in the peridynamics simulation dicussed in the reviewed
+work is essentially computing a Frechet derivative of a higher dimension function of higher
+dimension variables.
+
+\subsection{Differences with reviewed work}
+Where the application of the CS derivatives in the prior work is limited to a particular 
+class of abstract functions to determine the method's suitability for use in condition number
+estimation, the reviewed work applies CS to solving a mechanics problem.
+
 \subsection{Novelty of reviewed work}
+Because the reviewed work is not about just computing Frechet derivatives with CS but using
+them in a numerical method to solve a set of equations describing a physical system, the reviewed
+work is largely different from the prior work.
 
@@
 
-\subsection{Similarities with review work}
-\subsection{Differences with review work}
+\subsection{Aspects of CS discussed}
+The use of CS to compute sensitivities is disucssed in the contect of a solid mechanics
+FEA analysis.
+
+\subsection{Similarities with reviewed work}
+CS is used to compute derivitives.
+
+\subsection{Differences with reviewed work}
+In the prior work the derivitives produced by CS themselves are the object of interest, where
+the reviewed work uses CS in a numerical method for solving a non-linear system.
+
 \subsection{Novelty of reviewed work}
+The prior work and the reviewed work discuss different aspects of the CS method. The
+reviewed work discusses using the CS in a numerical method to solve a non-linear system,
+where the prior work discusses the derivatves produced by CS as an object of interest.
 
@@
 
-\subsection{Similarities with review work}
-\subsection{Differences with review work}
+\subsection{Aspect of CS discussed}
+The prior work dicusses the background of the CS method and examines it as a subset
+of Fourier Differentiation. CS and Fourier Differentiation are then applied to 
+computing sensitivities within a solid mechanics FEA.
+
+\subsection{Similarities with reviewed work}
+CS is used to compute derivatives.
+
+\subsection{Differences with reviewed work}
+In the prior work, CS is discussed and related to a broader class of numerical methods. CS is then used
+to compute sensitivities. In the reviewed work, complex step is used to compute residual Jacobians
+which are used in Newton's method for solving non-linear systems.
+
 \subsection{Novelty of reviewed work}
+The reviewed work and prior work discuss different aspects of CS in that they examine different
+applications of the method.
 
@@
 
-\subsection{Similarities with review work}
-\subsection{Differences with review work}
+\subsection{Aspects of CS discussed}
+First and second order CS methods for computing derivatives. Choosing optimal
+step size.
+
+\subsection{Similarities with reviewed work}
+Derivatives are computed with CS.
+
+\subsection{Differences with reviewed work}
+The prior work discusses computing first and second order accurate derivatives with CS and choosing optimal step size.
+The reviewed work discussing using the derviatives computed with the CS method in a numerical method for solving 
+a non-linear system.
+
 \subsection{Novelty of reviewed work}
+The reviewed work is offers discusion of the CS method as used in a numerical method for solving a non-linear system, where 
+the prior work dicusses the derivatives produced by CS themselves as an object of interest.
 
@@
 
-\subsection{Similarities with review work}
-\subsection{Differences with review work}
+\subsection{Aspects of CS discussed}
+Computing derivatives with CS is discussed. Combinations of real and imaginary step components
+are studied to determine CS methods with extended accuracy order.
+
+\subsection{Similarities with reviewed work}
+Derivatives are computed with CS.
+
+\subsection{Differences with reviewed work}
+The prior work discusses using the CS method to compute derivatives and attempts to extend the
+methods for that application. The reviewed work us the derivatives computed with CS in a numerical
+method for solving a non-linear system.
+
 \subsection{Novelty of reviewed work}
+The prior work discusses the derviatives computed by the CS method as an object
+of interest, while the reviewed work discusses using the derivatives computed
+by the CS method in a numerical method for solving a non-linear system.
+Because the the reviewed work and the prior work discuss different aspects of
+CS and different applications, the reviewed work is novel.
 
@@
 
-\subsection{Similarities with review work}
-\subsection{Differences with review work}
+\subsection{Aspect of CS discussed}
+The CS method is identified in an exact form.
+
+\subsection{Similarities with reviewed work}
+Derivatives are computed with the CS method.
+
+\subsection{Differences with reviewed work}
+The prior work develops the CS method, while the reviewed work
+is about an application of the CS method.
+
 \subsection{Novelty of reviewed work}
+Because the reviewed work discusses an application the material presented in the
+prior work not discussed in the prior work, it introduces novel material relative to the
+prior work.
 
@@
 
-\subsection{Similarities with review work}
-\subsection{Differences with review work}
+\subsection{Aspects of CS discussed}
+The series representation derived CS method is shown in its truncated form.
+
+\subsection{Similarities with reviewed work}
+Derivatives are computed with the CS method.
+
+\subsection{Differences with reviewed work}
+The prior work develops the CS method, while the reviewed work
+is about an application of the CS method.
+
 \subsection{Novelty of reviewed work}
+Because the reviewed work discusses an application the material presented in the
+prior work not discussed in the prior work, it introduces novel material relative to the
+prior work.
 

### 2014-02-22 jacobian_calc_CMAME.tex
added lit review to paper

@@
 
-This section contains background information on the differentiation techniques underlying the tangent-stiffness matrix calculation methods compared in the study.
-Since the derivation of first-order finite-difference techniques from Taylor series expansion is considered well known, it is not described here.
-Instead the reader is referred to \cite[Chap.
-4.1.3]{chapra2010}.
+This section contains background information on the differentiation techniques
+underlying the tangent-stiffness matrix calculation methods compared in the
+study. Since the derivation of first-order finite-difference techniques from
+Taylor series expansion is considered well known, it is not described here.
+Instead the reader is referred to \cite[Chap. 4.1.3]{chapra2010}.
 
@@
 \cite{lyness1967numerical,lyness1968differentiation} and has more recently been
-rediscovered
-\cite{squire1998using,voorhees2011complex,al2010complex,martins2000automated}
-for use in engineering analysis.  The basic idea is that a model parameter can
+rediscovered for use in engineering analysis.  The basic idea is that a model parameter can
 be be made complex and expanded in a Taylor series about a small perturbation
@@
 derivative calculation of a function.
+
+\note[Michael]{re: I.4}
+It is appropriate to detail the prior work discussing aspects of the
+complex-step method in order to indicate what is new in the instant work.  The
+CS method was introduced in a work by J. N. Lyness and C. B. Moler
+\cite{lyness1967numerical}.  In this prior work, the CS method is identified in
+an exact form and the comptuing of analytical derivatives is shown. This prior
+work develops the CS method, while the instant work is about an application of
+the CS method.
+
+A follow-on work by J. N. Lyness demonstrates a 'truncated' approximate version
+of the original CS method to make it suitable for use on digital computers \cite{lyness1968differentiation}.
+This form of the CS method is used in the instant work.
+The prior work develops the CS method, while the instant work is about an application of the CS method.
+
+A work by W. Squire and G. Trapp showns through empirical data that CS offers superior accuracy
+to FD and CD \cite{squire1998using}. The instant work uses the CS method as shown in the
+prior work and was motivated by the prior work's suggestion that CS is highly accurate. 
+The instant work examines applications of the CS method beyond taking derivatives alone,
+whereas the prior work does not.
+
+A work by J. C. Newman, W. K. Anderson and L. D. Whitfield, discusses computing
+sensitivities in a CFD analysis using complex step and compares the results to
+central difference produced results \cite{newman1998}. The work is also notable
+for identifying automatic differentiation as a competing method to CS and
+suggesting that complex step could be used to compute Jacobians of residuals
+for use in Newton-Kyrlov schemes for solving non-linear systems. The instant
+work is novel with respect to the prior work because the instant work
+investigates CS as applied to a numerical method for solving non-linear
+systems, rather than as in the prior work investigating the derivatives
+themselves produced with the CS method.
+
+In a work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta, CS is used to
+compute Jacobians for both integrating plastic stress to determine
+elastoplastic modulii, but also for computing the Jacobian of the global
+problem, both of which are an application of the CS method to the solution of
+non-linear systems \cite{perez2000numerical}. These elastoplastic modulii are
+used in computing the global Jacobian  and said to be 'algorithmically
+consistent' in that they are consistent with the algorithm used to return the
+predicted stress to the yield surface. Examples of interest include FEM
+simulations on two plastic material models with complex hardening rules.
+Comparing the prior work with the instant work, the prior work demonstrates CS
+on a serial computer and discusses a classical solid mechanics formulation.
+Automatic differentiation is not investigated, and CPU time measurements are not
+taken. Scalability studies with a variety of equivalent, progressively more
+refined meshes are not done. Scalability as a function of parallel computer
+size is not investigated. The instant work is novel with respect to the prior work in
+offering a discussion of CS as applied to peridynamics simulation on a parallel
+computer, CPU time data, and comparison to automatic differentiation as a competing method.
+
+Another work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta The prior work
+applies CS in computing derivatives needed for integrating stress, obtaining
+algorithmically consistent elastoplastic modulii, and obtaining the Jacobian of
+the global residual, for FEM simulations using the the MRS-Lade material model
+\cite{perez2012numerical}. In this prior work, CPU runtime is measured.
+Comparing the prior work to the instant work, the prior work demonstrates CS on
+a serial computer and discusses a classical solid mechanics formulation.
+Automatic differentiation is not investigated.  Scalability studies with a
+variety of equivalent, progressively more refined meshes are not done.
+Scalability as a function of distributed system size is not investigated.  The
+instant work is novel with respect to the prior work in offering a discussion
+of CS as applied to peridynamics simulation on a parallel computer, and
+comparison to automatic differentiation as a competing method.
+
+A work by Joaquim R. R. A. Martins, Ilan M. Kroo and Juan J. Alonso
+investigates using CS for computing aerodynamic sensitivities
+\cite{martins2000automated}. Additionally higher order accuracy versions of CS
+are discussed, as is the task of implementing complex-step in most widely used
+general purpose programming languages. Also dicussed is the relative ease of
+implementing a CS scheme versus an AD based scheme for computing sensitivities
+in an established code. The instant work is novel with respect to the prior
+work because the instant work investigates CS as applied to a numerical method
+for solving non-linear systems, rather than as in the prior work investigating
+the using the CS method to obtain sensitivities.
+
+A work by W. K. Anderson, J. C. Newman, D. l. Whitfield and E. J. Nielsen discusses
+the CS applied to computing sensitivities in CFD simulations \cite{anderson2001sensitivity}. CS is
+compared to AD both in Fortran and C++ applications. In this prior work it is suggested that forward mode
+AD is mathematically equivalent to first order CS. Forware mode AD isThe instant work is novel with respect to the prior
+work because the instant work investigates CS as applied to a numerical method
+for solving non-linear systems, rather than as in the prior work investigating
+the derivatives themselves produced with the CS method.
+
+A work by J. Martins, P. Sturdza, J. Alonso demonstrates that the CS method can
+be performed by using complex datatypes for the variables of functions
+comprising an analysis code. The analysis code can comprise a simulation and
+derivatives of response variables called sensitivities can be computed. The
+accuracy of this approach is compared to automatic differentiation
+\cite{martins2003complex}. The instant work is novel with respect to the prior
+work because the instant work investigates CS as applied to a numerical method
+for solving non-linear systems, rather than as in the prior work investigating
+the derivatives themselves produced with the CS method.
+
+A work by  K. L. Lai and J. L. Crassidis discusses first and second order CS
+methods for computing derivatives. Choosing optimal step size
+\cite{lai2008extensions}. This prior work discusses computing first and second
+order accurate derivatives with CS and choosing optimal step size.  The intant
+work discusses using the derviatives computed with the CS method in a numerical
+method for solving a non-linear system.
+
+A work by A. H. Al-Mohy and N. J. Higham investigates the use of complex step
+in computing Frechet derivatives of matix to matrix functions \cite{al2010complex}. To
+compute the residual Jacobian in the peridynamics simulation discussed in the
+instant work is analogous because it is computing a Frechet derivative of a
+higher dimension function of higher dimension variables. Where the application
+of the CS derivatives in the prior work is limited to a particular class of
+abstract functions to determine the CS' suitability for use in condition number
+estimation, the instant work applies CS to solving a mechanics problem. The
+instant work is not about computing Frechet derivatives with CS but moreover using them
+in a numerical method to solve a set of equations describing a physical system.
+
+A work by W. Jin, B. H. Dennis, and B. P. Wang discusses using CS to compute
+sensitivities in a solid mechanics FEM analysis \cite{jin2010improved}. In the
+prior work the derivitives produced by CS themselves are the object of
+interest, where the instant work uses CS in a numerical method for solving a
+non-linear system.
+
+A work by A. Voorhees, H. Millwater, and R. Bagley, where H. Millwater is an author of
+the instant work, dicusses the background of the CS method and frames it as a subset
+of Fourier Differentiation. CS and Fourier Differentiation are then applied to 
+computing sensitivities within a solid mechanics FEA \cite{voorhees2011complex}. A notable
+technique demonstrated by this prior work is the representation of complex valued
+nodal components as additional degrees of freedom in the FE model, to be interpreted
+by a matrix representation of multicomplex numbers during computations, yet without storing
+the entire matrix. In the prior work, CS is discussed and related to a broader class of differentiation methods and used
+to compute sensitivities. In the instant work, CS is used in a numerical method for solving non-linear systems.
+
+In another work by H. Millwater, C. Quintana, J. Garza and David Wagner the
+concept of incorporating complex values into the degrees of freedom of finite
+elements is labelled ZFEM \cite{millwater2013application}. Notable in this work
+is the discussion of creating user defined element types for use with the
+popular \emph{ABAQUS} software suite which enable the use of ZFEM in that
+software, \emph{ABAQUS} is a trademark of Dassault Systemes
+\cite{systemes2012abaqus}. In this prior work the goal of using CS through the
+ZFEM technique is to compute shape sensitivities, while in the instant work CS
+is used in a numerical method for solving non-linear systems.
+
+A work by R. Abreu, D. Stich, and J. Morales discusses computing derivatives
+with a re-interpreted version of the CS method which features the combined use
+of real and imaginary valued step components \cite{abreu2013generalization}.
+This prior work suggests that the modified CS method can achieve extended
+approximation accuracy up to fourth order.  The prior work discusses using the
+CS method to compute derivatives and attempts to extend the method for that
+application, where the derivatives themselves are the object of interest. In
+contrast, the instant work discusses using the derivatives computed with CS in
+a numerical method for solving a non-linear system.
+
+The body of prior literature has introduced the application of CS to
+single dimension and multidimensional derivatives, sensitivity analysis,
+non-linear solution schemes, the extension of CS for higher accuracy, 
+and implementation strategies for CS in new and existing codes.
+With respect to the prior work dicussing the CS method, the primary novel
+aspects of the instant work are then its dicussion of the application of CS to
+peridynamics simulations run in \emph{Peridigm}, a parallel implementation of
+CS and a compilation of CPU time data, where each of CS, AD, FD and CD are
+compared.
 
@@
 
-\begin{figure}[tbp] \centering \includegraphics{./figs/truss.png}
+\begin{figure}[tbp] \centering \includegraphics[width=0.7\textwidth, height=30px]{./figs/truss.png}
 \caption{Beam and spring simulated in example program.} \label{fig:TrussSchematic}

### 2014-02-22 response/response.tex
I.4 addressed

@@
   \item
+     In Section 1.1.1, please include a more comprehensive literature review for
+     the CS method, including fur ther references (some of them are listed below).
+     In particular, specify what aspects of the CS method have been investigated in
+     each reference. Then, you need to clearly mention your contribution within
+     the CS method, in contrast to the other references (some comments in this
+     regard appear in the manuscript in page 8).
+
+{\color{red}
+  \begin{itemize}
+     \item 
+         The authors have appended a more comprehensive literature review for the CS method to section 1.1.1.
+         The aspects of CS investigated by the prior work, as well as any particularly interesting
+         developments introduced by those works, have been listed. To argue for the novelty of the instant work,
+         each prior work is compared to the instant work in terms of the the use of the CS method and the similarity of the fields of research of the works.
+     \item
+         The literature review concludes with an argument for the novelty of the instant
+         work where the prior works are taken as a whole and compared to the instant work. It is the authors' opinion
+         that the instant work is novel for at least the reasons set out in the literature review material, where
+         the contribution to the CS method is in the comparison to other methods in terms of computational
+         time,  parallel implementation, and implementation within \emph{Peridigm} .
+  \end{itemize}}
+
+  \item
       The comparisons of the different algorithms are only performed with respect to the
-      construction of tangent-stiffness matrices. In par ticular, the authors always use the
+      construction of tangent-stiffness matrices. In particular, the authors always use the
       solution of AD at each time step to compute the entries of the tangent-stiffness
@@
 
-  \item
-     In Section 1.1.1, please include a more comprehensive literature review for
-     the CS method, including fur ther references (some of them are listed below).
-     In particular, specify what aspects of the CS method have been investigated in
-     each reference. Then, you need to clearly mention your contribution within
-     the CS method, in contrast to the other references (some comments in this
-     regard appear in the manuscript in page 8).
-
-{\color{red}
-  \begin{itemize}
-  \end{itemize}}
-
 \end{document}

### 2014-02-22 jacobian_calc_CMAME.tex
spelling mistakes in manuscript fixed

@@
 \cite{lyness1967numerical}.  In this prior work, the CS method is identified in
-an exact form and the comptuing of analytical derivatives is shown. This prior
+an series form and the computing of analytical derivatives is shown. This prior
 work develops the CS method, while the instant work is about an application of
@@
 
-A work by W. Squire and G. Trapp showns through empirical data that CS offers superior accuracy
+A work by W. Squire and G. Trapp shows through empirical data that CS offers superior accuracy
 to FD and CD \cite{squire1998using}. The instant work uses the CS method as shown in the
@@
 suggesting that complex step could be used to compute Jacobians of residuals
-for use in Newton-Kyrlov schemes for solving non-linear systems. The instant
+for use in Newton-Krylov schemes for solving non-linear systems. The instant
 work is novel with respect to the prior work because the instant work
@@
 algorithmically consistent elastoplastic modulii, and obtaining the Jacobian of
-the global residual, for FEM simulations using the the MRS-Lade material model
+the global residual, for FEM simulations using the MRS-Lade material model
 \cite{perez2012numerical}. In this prior work, CPU runtime is measured.
@@
 are discussed, as is the task of implementing complex-step in most widely used
-general purpose programming languages. Also dicussed is the relative ease of
+general purpose programming languages. Also discussed is the relative ease of
 implementing a CS scheme versus an AD based scheme for computing sensitivities
@@
 compared to AD both in Fortran and C++ applications. In this prior work it is suggested that forward mode
-AD is mathematically equivalent to first order CS. Forware mode AD isThe instant work is novel with respect to the prior
+AD is mathematically equivalent to first order CS. The instant work is novel with respect to the prior
 work because the instant work investigates CS as applied to a numerical method
@@
 \cite{lai2008extensions}. This prior work discusses computing first and second
-order accurate derivatives with CS and choosing optimal step size.  The intant
-work discusses using the derviatives computed with the CS method in a numerical
+order accurate derivatives with CS and choosing optimal step size.  The instant
+work discusses using the derivatives computed with the CS method in a numerical
 method for solving a non-linear system.
@@
 sensitivities in a solid mechanics FEM analysis \cite{jin2010improved}. In the
-prior work the derivitives produced by CS themselves are the object of
+prior work the derivatives produced by CS themselves are the object of
 interest, where the instant work uses CS in a numerical method for solving a
@@
 A work by A. Voorhees, H. Millwater, and R. Bagley, where H. Millwater is an author of
-the instant work, dicusses the background of the CS method and frames it as a subset
+the instant work, discusses the background of the CS method and frames it as a subset
 of Fourier Differentiation. CS and Fourier Differentiation are then applied to 
@@
 nodal components as additional degrees of freedom in the FE model, to be interpreted
-by a matrix representation of multicomplex numbers during computations, yet without storing
+by a matrix representation of complex numbers during computations, yet without storing
 the entire matrix. In the prior work, CS is discussed and related to a broader class of differentiation methods and used
@@
 and implementation strategies for CS in new and existing codes.
-With respect to the prior work dicussing the CS method, the primary novel
-aspects of the instant work are then its dicussion of the application of CS to
+With respect to the prior work discussing the CS method, the primary novel
+aspects of the instant work are then its discussion of the application of CS to
 peridynamics simulations run in \emph{Peridigm}, a parallel implementation of
@@
 The ranking order of the methods
-matched their speed ranking in Fig.~\ref{fig:serial_speed}. This plot indicates that efficiency is insensitive to the the number of nonzero TS elements.  
+matched their speed ranking in Fig.~\ref{fig:serial_speed}. This plot indicates that efficiency is insensitive to the number of nonzero TS elements.  
 %
@@
 the methods, it was necessary to conduct convergence rate studies in order to
-get a complete picture. It is necessary for an analyist to know if the
+get a complete picture. It is necessary for an analyst to know if the
 accuracy differences measured in the per-iteration tests would amount to
@@
 I.3}\add[Michael]{For the reader's convenience, a collection of contiguously
-stored data elements are said to be 'byte-copiable' when the the interval
+stored data elements are said to be 'byte-copiable' when the interval
 in bytes of memory spanned by a data element in the collection is constant
 throughout the collection. Suppose 'contiguously stored' means, stored in
-memory with a minimum of unrelated infomation stored in between data
+memory with a minimum of unrelated information stored in between data
 elements in the collection. Complicated objects, such as AD datatype
@@
 same class, for the reason that they themselves contain an uncertain
-ammount of state data according to their design and work done in the
+amount of state data according to their design and work done in the
 program. This means that a collection of such objects, even if they are
@@
 that if \emph{Peridigm} or any simulation software were to be reformulated to
-take advantage of accelerators or GPUs, the flexibilty of complex-step owed to
+take advantage of accelerators or GPUs, the flexibility of complex-step owed to
 its simple implementation may make it more viable for that application given

### 2014-02-22 response/response.tex
fixed spelling errors in response

@@
      In Section 1.1.1, please include a more comprehensive literature review for
-     the CS method, including fur ther references (some of them are listed below).
+     the CS method, including further references (some of them are listed below).
      In particular, specify what aspects of the CS method have been investigated in
@@
          developments introduced by those works, have been listed. To argue for the novelty of the instant work,
-         each prior work is compared to the instant work in terms of the the use of the CS method and the similarity of the fields of research of the works.
+         each prior work is compared to the instant work in terms of the use of the CS method and the similarity of the fields of research of the works.
      \item

### 2014-02-22 jacobian_calc_CMAME.tex
I.4 response section corrected for grammar and style

@@
 \cite{lyness1967numerical}.  In this prior work, the CS method is identified in
-an series form and the computing of analytical derivatives is shown. This prior
+an infinite series form and the computing of analytical derivatives is shown. This prior
 work develops the CS method, while the instant work is about an application of
@@
 
-A work by J. C. Newman, W. K. Anderson and L. D. Whitfield, discusses computing
+A work by J. C. Newman, W. K. Anderson and L. D. Whitfield discusses computing
 sensitivities in a CFD analysis using complex step and compares the results to
@@
 In a work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta, CS is used to
-compute Jacobians for both integrating plastic stress to determine
-elastoplastic modulii, but also for computing the Jacobian of the global
+compute Jacobians for integrating plastic stress to determine
+elastoplastic modulii and for computing the Jacobian of the residual for the global
 problem, both of which are an application of the CS method to the solution of
 non-linear systems \cite{perez2000numerical}. These elastoplastic modulii are
-used in computing the global Jacobian  and said to be 'algorithmically
+used in computing the global Jacobian  and are said to be 'algorithmically
 consistent' in that they are consistent with the algorithm used to return the
 predicted stress to the yield surface. Examples of interest include FEM
-simulations on two plastic material models with complex hardening rules.
+simulations on two plastic material models with complicated hardening rules.
 Comparing the prior work with the instant work, the prior work demonstrates CS
 on a serial computer and discusses a classical solid mechanics formulation.
-Automatic differentiation is not investigated, and CPU time measurements are not
+Automatic differentiation is not investigated and CPU time measurements are not
 taken. Scalability studies with a variety of equivalent, progressively more
@@
 offering a discussion of CS as applied to peridynamics simulation on a parallel
-computer, CPU time data, and comparison to automatic differentiation as a competing method.
-
-Another work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta The prior work
-applies CS in computing derivatives needed for integrating stress, obtaining
-algorithmically consistent elastoplastic modulii, and obtaining the Jacobian of
-the global residual, for FEM simulations using the MRS-Lade material model
+computer, CPU time data, and comparison of CS to automatic differentiation as a competing method.
+
+Another work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta applies CS in
+computing derivatives needed for integrating stress, obtaining algorithmically
+consistent elastoplastic modulii, and obtaining the Jacobian of the global
+residual, for FEM simulations using the MRS-Lade material model
 \cite{perez2012numerical}. In this prior work, CPU runtime is measured.
@@
 of CS as applied to peridynamics simulation on a parallel computer, and
-comparison to automatic differentiation as a competing method.
+comparison of CS to automatic differentiation as a competing method.
 
@@
 investigates using CS for computing aerodynamic sensitivities
-\cite{martins2000automated}. Additionally higher order accuracy versions of CS
+\cite{martins2000automated}. Additionally, higher order accuracy versions of CS
 are discussed, as is the task of implementing complex-step in most widely used
@@
 for solving non-linear systems, rather than as in the prior work investigating
-the using the CS method to obtain sensitivities.
-
-A work by W. K. Anderson, J. C. Newman, D. l. Whitfield and E. J. Nielsen discusses
-the CS applied to computing sensitivities in CFD simulations \cite{anderson2001sensitivity}. CS is
-compared to AD both in Fortran and C++ applications. In this prior work it is suggested that forward mode
-AD is mathematically equivalent to first order CS. The instant work is novel with respect to the prior
+the use of the CS method to obtain sensitivities.
+
+A work by W. K. Anderson, J. C. Newman, D. l. Whitfield and E. J. Nielsen
+discusses the CS method as applied to computing sensitivities in CFD simulations
+\cite{anderson2001sensitivity}. CS is compared to AD both in Fortran and C++
+applications. In this prior work it is suggested that forward mode AD is
+mathematically equivalent to first order CS. The instant work is novel with
+respect to the prior work because the instant work investigates CS as applied
+to a numerical method for solving non-linear systems, rather than as in the
+prior work investigating the derivatives themselves produced with the CS
+method.
+
+A work by J. Martins, P. Sturdza, J. Alonso demonstrates that the CS method can
+be performed by using complex datatypes for the variables of functions
+comprising an analysis code \cite{martins2003complex}. The analysis code can
+comprise a simulation and derivatives of response variables called
+sensitivities can be computed. The accuracy of this approach is compared to
+automatic differentiation. The instant work is novel with respect to the prior
 work because the instant work investigates CS as applied to a numerical method
@@
 
-A work by J. Martins, P. Sturdza, J. Alonso demonstrates that the CS method can
-be performed by using complex datatypes for the variables of functions
-comprising an analysis code. The analysis code can comprise a simulation and
-derivatives of response variables called sensitivities can be computed. The
-accuracy of this approach is compared to automatic differentiation
-\cite{martins2003complex}. The instant work is novel with respect to the prior
-work because the instant work investigates CS as applied to a numerical method
-for solving non-linear systems, rather than as in the prior work investigating
-the derivatives themselves produced with the CS method.
-
 A work by  K. L. Lai and J. L. Crassidis discusses first and second order CS
-methods for computing derivatives. Choosing optimal step size
-\cite{lai2008extensions}. This prior work discusses computing first and second
-order accurate derivatives with CS and choosing optimal step size.  The instant
-work discusses using the derivatives computed with the CS method in a numerical
-method for solving a non-linear system.
+methods for computing derivatives and choosing optimal step size
+\cite{lai2008extensions}. The instant work discusses using the derivatives
+computed with the CS method in a numerical method for solving a non-linear
+system, while in the prior work the derivatives obtained from the CS method 
+are the object of interest.
 
 A work by A. H. Al-Mohy and N. J. Higham investigates the use of complex step
-in computing Frechet derivatives of matix to matrix functions \cite{al2010complex}. To
-compute the residual Jacobian in the peridynamics simulation discussed in the
-instant work is analogous because it is computing a Frechet derivative of a
-higher dimension function of higher dimension variables. Where the application
-of the CS derivatives in the prior work is limited to a particular class of
-abstract functions to determine the CS' suitability for use in condition number
-estimation, the instant work applies CS to solving a mechanics problem. The
-instant work is not about computing Frechet derivatives with CS but moreover using them
-in a numerical method to solve a set of equations describing a physical system.
+in computing Frechet derivatives of matix to matrix functions
+\cite{al2010complex}. To compute the Jacobian of the residual in the
+peridynamics simulation as discussed in the instant work is analogous is
+computing a Frechet derivative. Where the application of the CS derivatives in
+the prior work is limited to a particular class of abstract functions to
+determine the CS method's suitability for use in condition number estimation, the
+instant work applies CS to solving a solid mechanics problem. The instant work is not
+about computing Frechet derivatives with CS but moreover using them in a
+numerical method to solve a set of equations describing a physical system.
 
@@
 sensitivities in a solid mechanics FEM analysis \cite{jin2010improved}. In the
-prior work the derivatives produced by CS themselves are the object of
-interest, where the instant work uses CS in a numerical method for solving a
+prior work the derivatives produced by CS are the object of
+interest, whereas the instant work uses CS in a numerical method for solving a
 non-linear system.
 
-A work by A. Voorhees, H. Millwater, and R. Bagley, where H. Millwater is an author of
-the instant work, discusses the background of the CS method and frames it as a subset
-of Fourier Differentiation. CS and Fourier Differentiation are then applied to 
-computing sensitivities within a solid mechanics FEA \cite{voorhees2011complex}. A notable
-technique demonstrated by this prior work is the representation of complex valued
-nodal components as additional degrees of freedom in the FE model, to be interpreted
-by a matrix representation of complex numbers during computations, yet without storing
-the entire matrix. In the prior work, CS is discussed and related to a broader class of differentiation methods and used
-to compute sensitivities. In the instant work, CS is used in a numerical method for solving non-linear systems.
+A work by A. Voorhees, H. Millwater, and R. Bagley discusses the background of
+the CS method and frames it as a subset of Fourier Differentiation. CS and
+Fourier Differentiation are then applied to computing sensitivities within a
+solid mechanics FEA \cite{voorhees2011complex}. A notable technique
+demonstrated by this prior work is the representation of complex valued nodal
+coordinates with additional degrees of freedom in the FE model, to be
+interpreted by a matrix representation of complex numbers during computations.
+In the prior work, CS is discussed and related to a broader class of
+differentiation methods and used to compute sensitivities. In the instant work,
+CS is used in a numerical method for solving non-linear systems.
 
@@
 concept of incorporating complex values into the degrees of freedom of finite
-elements is labelled ZFEM \cite{millwater2013application}. Notable in this work
+elements is called ZFEM \cite{millwater2013application}. Notable in this work
 is the discussion of creating user defined element types for use with the
-popular \emph{ABAQUS} software suite which enable the use of ZFEM in that
-software, \emph{ABAQUS} is a trademark of Dassault Systemes
+popular \emph{ABAQUS} software package which enables the use of ZFEM in that
+software. \emph{ABAQUS} is a trademark of Dassault Systemes
 \cite{systemes2012abaqus}. In this prior work the goal of using CS through the
@@
 A work by R. Abreu, D. Stich, and J. Morales discusses computing derivatives
-with a re-interpreted version of the CS method which features the combined use
+with a version of the CS method which features the combined use
 of real and imaginary valued step components \cite{abreu2013generalization}.
@@
 and implementation strategies for CS in new and existing codes.
-With respect to the prior work discussing the CS method, the primary novel
+With respect to the prior work discussing the CS method, the novel
 aspects of the instant work are then its discussion of the application of CS to

### 2014-02-22 jacobian_calc_CMAME.tex
removed treatise on byte copiable data I had crammed into the conclusion, as excess

@@
 require compiling a version of the AD library being used that were compatible
-with the accelerator or GPU being used. \note[Michael]{re:
-I.3}\add[Michael]{For the reader's convenience, a collection of contiguously
-stored data elements are said to be 'byte-copiable' when the interval
-in bytes of memory spanned by a data element in the collection is constant
-throughout the collection. Suppose 'contiguously stored' means, stored in
-memory with a minimum of unrelated information stored in between data
-elements in the collection. Complicated objects, such as AD datatype
-elements, do not have the same size in memory as other objects even of that
-same class, for the reason that they themselves contain an uncertain
-amount of state data according to their design and work done in the
-program. This means that a collection of such objects, even if they are
-contiguously stored in memory, will not be alligned according to a constant
-interval, such that elements within a collection cannot be located by
-coordinates consisting of the location of the beginning of the first
-element in the collection and an interval multiplied by a natural number.
-This method of locating is integral to the serialized communication
-implemented by MPI the libraries as well as CUDA, OpenCL, and others,
-which allow only the transmission of collections of byte-copiable datatypes
-(that is 'primitive datatypes'). A common method of bypassing this
-limitation for communication purposes is to write the complicated data to a
-'character buffer', and transmit the byte-copiable character buffer,
-however this is useless unless the recipient device of the communication
-can recover the original data and process it, which is impossible if the
-recipient device only can work with a subset of the programming language
-that the sending device must use, such as in a communication between a CPU
-running C++ and a GPU running C (such as in CUDA programs), where the gap in
-sophistication contains the features of the "complicated" datatype.} What this means is
+with the accelerator or GPU being used.  What this means is
 that if \emph{Peridigm} or any simulation software were to be reformulated to

### 2014-02-22 jacobian_calc_CMAME.tex
supported the assertion of byte-copyability being significant with evidence that should matter to stampede users

@@
 require compiling a version of the AD library being used that were compatible
-with the accelerator or GPU being used.  What this means is
-that if \emph{Peridigm} or any simulation software were to be reformulated to
-take advantage of accelerators or GPUs, the flexibility of complex-step owed to
-its simple implementation may make it more viable for that application given
-its high accuracy. Additional performance gains may be achieved through the use
-of the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
+with the accelerator or GPU being used. A working definition of byte-copyable
+is that the collection of data to transfer is made of identically sized pieces.
+Byte-copyability is significant for example because it is a requirement for the
+basic offload mode provided for Intel MIC accelerator boards which are
+installed in \emph{Stampede}, where the alternative mode requires "more
+extensive" effort \cite{intel_byte_copyable}. What this means is that if
+\emph{Peridigm} or any simulation software were to be reformulated to take
+advantage of accelerators or GPUs, the flexibility of complex-step owed to its
+simple implementation may make it more viable for that application given its
+high accuracy. Additional performance gains may be achieved through the use of
+the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
 use of complex data types natively, and alleviate the expense of the
@@
 dependence on an external AD library.  From this perspective, the method may
-still be an attractive method for TS matrix evaluation. \add[Michael]{However, at least for 
-the models producing the results shown here, each of the
+still be an attractive method for TS matrix evaluation. \add[Michael]{However,
+at least for the models producing the results shown here, each of the
 methods returned quadratic convergence, indicating each was sufficiently
 accurate. 'Marching orders' for the analyst would be to select the swiftest
-method, while it would seem accuracy above what is adequate does not enhance
-the convergence rate or iteration speed until either classical Newton's convergence
-conditions or the approximate case conditions are violated.}
+method, while it would seem accuracy above what is adequate does not
+enhance the convergence rate or iteration speed until either classical
+Newton's convergence conditions or the approximate case conditions are
+violated.}
 

### 2014-02-22 jacobian_calc_CMAME.tex
working on peridynamics section, considering citing appendix from seleson paper and reproducing that verbatim in the appendix of this work

@@
 
-\begin{figure}[tbp] \centering \includegraphics[width=0.7\textwidth, height=30px]{./figs/truss.png}
+\begin{figure}[tbp] \centering \includegraphics[width=0.7\textwidth]{./figs/truss.png}
 \caption{Beam and spring simulated in example program.} \label{fig:TrussSchematic}
@@
 peridynamic solid, having what is called a state-based constitutive model. In
-order for a definition of this material to be useful here, background material
-needs to be laid out to enable the discussion. To begin with, some essential
+order for a definition of this material to be useful her. To begin with, some essential
 mathematical definitions will be listed and peridynamic theory summarized. Then
@@
 the problem being discussed in this work. The following discussion is largely
-taken from \cite{silling:psa}, but reproduced here fo clarity.
+taken from \cite{silling:psa} but reproduced here for clarity.
 
@@
 
-\subsubsection{State-based consitutive models.}
+In \cite{silling:psa}, the primary reference for state based peridynamics, the
+equation of motion for state based peridynamic theory is derived. No
+development of the theory is claimed; the equation of motion is reproduced here for the reader's convenience.
+%
+The general equation is written as it appears in \cite{silling:psa}:
+\begin{equation} 
+    \label{eqn:PDMotion}
+    \rho(\boldsymbol{x})\boldsymbol{\ddot{u}}(\boldsymbol{x}, t) = \int_{\mathcal{H}_{\boldsymbol{x}}} \{\boldsymbol{\underline{T}}[\boldsymbol{x}, t]\langle \boldsymbol{x}^{\prime} - \boldsymbol{x} \rangle - \boldsymbol{\underline{T}}[\boldsymbol{x}^{\prime}, t]\langle \boldsymbol{x} - \boldsymbol{x}^{\prime} \rangle \}dV_{\boldsymbol{x}^{\prime}} + \boldsymbol{b}(\boldsymbol{x}, t),
+\end{equation}
+
+where "$\mathcal{H}_{\boldsymbol{x}}$ is a spherical neighborhood" with a
+radius, centered at the original coordinates of a peridynamic node,
+$\boldsymbol{x}$, and \boldsymbol{\underline{T}} is the "force vector state
+field". The force vector state field maps vectors in the original coordinate system 
+measured from the peridynamic node at the center of the neighborhood or 'horizon' to
+neighboring peridynamic nodes that fall within the radius of the horizon. $\boldsymbol{b}$ denotes a body force density. Equation~\ref{eqn:PDMotion} may be simplified based on the assumption that
+\boldsymbol{\underline{T}} is Riemann integrable and that integral is uniformly
+convergent \cite{silling:psa} as,
+
+\begin{equation} 
+    \label{eqn:PDMotion_simple}
+    \rho\boldsymbol{\ddot{u}} = \int_{\mathcal{H}_{\boldsymbol{x}}} \{\boldsymbol{\underline{T}}\langle \boldsymbol{x}^{\prime} - \boldsymbol{x} \rangle - \boldsymbol{\underline{T}}^{\prime}\langle \boldsymbol{x} - \boldsymbol{x}^{\prime} \rangle \}dV_{\boldsymbol{x}^{\prime}} + \boldsymbol{b},
+\end{equation}
+
+where,
+
+\begin{equation} 
+    \label{eqn:PDMotion_simple}
+    \boldsymbol{\underline{T}} = \boldsymbol{\underline{T}}[\boldsymbol{x}, t], 
+    \boldsymbol{\underline{T}}^{\prime} = \boldsymbol{\underline{T}}[\boldsymbol{x}^{\prime}, t]. 
+\end{equation}
 
@@
 \subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simualtion.}
-The tangent stiffness matrix is the frechet derivative of the residual wrt the
+The tangent stiffness matrix is the Frechet derivative of the residual wrt the
 deformation state.  For the purposes of this work, the residual represents the unbalanced forces in the
-equilibrium version of the peridynamic equation of motion. The internal force
-density is computed in the peridynamic equation of motion by computing the
-volume integral over, a horizon, of the Frechet derivative of the 
-strain energy state wrt to the deformed configuration inner product with the deformation state, per node,
+static peridynamic momentum equation. The internal force
+density is computed in the peridynamic momentum equation by computing the
+volume integral over a horizon of the Frechet derivative of the 
+strain energy density state wrt to the deformed configuration inner product with the deformation state, per node,
 for each of the three directions in the original reference frame.
@@
 method is employed by probing the deformed configuration one degree of freedom
-at a time and so allowing the sensitivity of the force density state wrt the
-deformation state for all nodes to be computed.  }
+at a time and so allowing the derivative of the force density state wrt the
+deformation state for all nodes to be computed. 
+}
 

### 2014-02-23 jacobian_calc_CMAME.tex
continuing I.2

@@
 \cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale
-coupling to molecular simulations \cite{seleson2009peridynamics}. The related mathematical
-definitions and parts of peridynamic theory especially relevant to the present
-work are explained in the sequel.
+coupling to molecular simulations \cite{seleson2009peridynamics,seleson2014atom}.
 
 The type of material model used in this study is referred to as a linear
-peridynamic solid, having what is called a state-based constitutive model. In
-order for a definition of this material to be useful her. To begin with, some essential
-mathematical definitions will be listed and peridynamic theory summarized. Then
-state-based constitutive models will be summarized, followed by a description of
-the linear peridynamic solid model used in this work. After the peridynamic
-equation of motion and consitutive model in use have been discussed, it is then
-possible to give a more precise definition of the tangent stiffness matrix for
-the problem being discussed in this work. The following discussion is largely
+peridynamic solid, having what is called a state based constitutive model. For
+the convenience of the reader, vector states will be introduced. Then the state
+based peridynamic equation of motion will be shown, followed by a description
+of the linear peridynamic solid model used in this work. After the peridynamic
+equation of motion and a consitutive model have been discussed, a definition of
+the tangent stiffness matrix is given. The following discussion is largely
 taken from \cite{silling:psa} but reproduced here for clarity.
@@
 \note[Michael]{re: I.2. John, this was to be the skeleton for the new sections on peridynamics.}
-\subsubsection{Vector states for state-based peridynamic theory.}
-
-\subsubsection{The state-based peridynamic equation of motion.}
+\subsubsection{Vector states for state based peridynamic theory.}
+Because the vector state, which is fair to say an uncommon notation, is used in the literature discussing state based peridynamic theory
+it will be enumerated here for the sake of continuity. 
+%
+Paraphrasing, let a spherical region, centered about a
+location in threespace, called a horizon or neighborhood with a certain radius
+be defined. A vector state of order $m$ is a mapping of a vector contained
+within the horizon to a tensor of order $m$ \cite{silling:psa}. Vector states
+have have their own mathematical properties and definitions for sums, composition,
+point product, reference state, identity scalar state, null reference state, null scalar state,
+dot product, magnitude state, unit state, and direction state which are detailed in \cite{silling:psa}. 
+It is important to think of vector states as non-linear, not necessarily continuous maps of vectors to real Euclidian space, where second order tensors are linear maps of three dimensional vectors to threespace.
+%
+One important tool mechanics gains from the concept of vector states is a way
+to represent complicated deformations within a body in the form of a state
+function. Whereas a symmetric second order tensor maps a spherical region in a
+body in the reference configuration to an ellipsoid in the deformed
+configuration, a vector state can map a spherical region in a body into a
+complicated shape \cite{silling:psa}.
+
+(PLACE PERIDYNAMIC POTATO DIAGRAM HERE)
+
+%
+The properties and definitions of vector states allow interaction with vectors
+and second order tensors to be defined which includes a tensor vector product
+in terms of a state function called an expansion, the tensor product of two states in terms of the
+volume integral over a horizon of the dyadic product of those states times a
+scalar function, the shape tensor as the tensor product of the reference vector
+state with itself, and a tensor reduced form of a state as the conventionally
+defined tensor product of the state tensor product of the state and the
+reference state with the inverse of the shape tensor, and the inheritance of
+the properties of a proper orthogonal second order tensor in its state
+expansion. These ideas are detailed in \cite{silling:psa}. 
+%
+Because of its importance in this work, the definition of a derivative of a vector state
+will be given a more complete explanation.
+
+\subsubsection{The state based peridynamic equation of motion.}
 

### 2014-02-27 response/response.tex
addressed smaller issues 1 through 4

@@
 
+  \subsection*{Smaller issues}
+  \begin{enumerate}
+  \item
+	1. Pg. 2: In line 14, explain what does the term “algorithmically consistent” mean.
+
+{\color{red}
+  \begin{itemize}
+     \item
+      A brief definition for the convenience of the reader was added in the manuscript.
+But, algorithmically consistent is a fairly common term in plasticity/mechanics.
+This same meaning is used for 'consistent' refering to 'tangent matrices' in \cite[perez2012consistent, perez2010].
+It means consistent with integrating the plasticity model under the Khun-Tucker constraints.
+Integrating the plasticity model means, given a strain increment, determining a stress that does not exceed the yield stress, and by doing so computing a linear slope, called a consistent tangent modulus, that relates the strain increment to a correct stress increment.
+An alogorithmically consistent tangent stiffness matrix is then one that uses the consistent tangent modulii, making it a more accurate approximation of the stiffness of the system.
+  \end{itemize}}
+
+  \item
+	2. Pg. 2: In line 36, what do you mean by “discrimination”?
+
+{\color{red}  
+\begin{itemize}
+     \item
+      The word, 'discrimination' has been removed from the manuscript.
+'Discrimination' was meant to mean 'measuring differences and making judgements on relative usefulness'.
+Since this meaning is probably understood from the word 'comparison' alone, the word 'discrimination' has been removed from the manuscript for clarity.
+  \end{itemize}}
+
+` \item	
+	3. Pg. 2: In line 37, explain what do you mean by “in-situ instrumentation”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      The words, 'in-situ' have been removed from the manuscript.
+'In-situ' was meant to suggest that timing measurements were taken from within and by the program being analyzed itself, rather than by an external program, but it has been removed for clarity.
+  \end{itemize}}
+
+  \item 
+	4. Pg. 3: In lines 19-29, specify the sections where each part of the manuscript is
+included.
+{\color{red}  
+\begin{itemize}
+     \item
+      The correction is made in the manunscript.
+  \end{itemize}}
+
 \end{document}

### 2014-02-27 jacobian_calc_CMAME.tex
addressed smaller issues 1 through 4

@@
 parallel computational peridynamics code for the purpose of this comparison.
-The methods were compared through in situ profiling of the code for accuracy,
+The methods were compared through profiling of the code for accuracy,
 speed, efficiency, and parallel scalability. The research provides data that
@@
 as CTSE or the ``complex-step'' method.  The distinction of ``accurate'' is
-defined by comparison of the \note[Michael]{re: I.1}\remove[Michael]{new} method to popular finite differencing
-techniques used for computing tangent-stiffness matrices and with the exact to
-machine precision algorithmically consistent derivatives computed with
-\emph{automatic differentiation}.
+defined by comparison of the \note[Michael]{re: I.1}\remove[Michael]{new}
+method to popular finite differencing techniques used for computing
+tangent-stiffness matrices and with the exact to machine precision
+algorithmically consistent derivatives computed with \emph{automatic
+differentiation}. For the reader's convenience, from plasticity theory, the algorithmic consistency of
+the tangent-stiffness matrix refers to consistency with the integration of a
+plasticity model under the Khun-Tucker constraints or similar where the tangent-stiffness matrix is
+computed using the elastoplastic modulii that are determined from the integration \cite{simo1998}.
 
@@
 computational mechanics code design who are faced with implementing a general
-method for tangent stiffness matrix calculation.  A comparison and
-discrimination of the methods was achieved through in-situ instrumentation of a
+method for tangent stiffness matrix calculation.  A comparison of the methods was achieved through instrumentation of a
 massively parallel computational mechanics code.
@@
 Newton's method. After presenting background information on the underlying
-methods to be discussed, presented in this paper are: A) a description of
-tangent-stiffness matrices, B) detailed directions for producing
-tangent-stiffness matrices with each of the methods identified above, C) a
+methods to be discussed, presented in this paper are: In subsection~\ref{subsec:TS} a description of
+tangent-stiffness matrices, detailed directions for producing
+tangent-stiffness matrices with each of the methods identified above, and a
 description and justification of the \note[Michael]{re: I.1}\remove[Michael]{new}complex-step method for calculating
-tangent-stiffness matrices, D) a description of implementing complex-step in
-the \emph{Peridigm} software, E) a description of the quantities-of-interest
-used to rank the methods, F) a presentation and analysis of the results of the
-comparative study, and G) conclusions and thoughts on potential future work.
+tangent-stiffness matrices, in sub-subsection~\ref{subsubsec:Impl} a description of implementing complex-step in
+the \emph{Peridigm} software, in section~\ref{JGAM} a description of the quantities-of-interest
+used to rank the methods, in section~\ref{sec:Res} a presentation and analysis of the results of the
+comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
 Finally, in the interest of replicable research, the reader is referred to the
@@
 \subsection{Tangent-stiffness} 
-
+\label{subsec:TS}
 In solid and structural computational mechanics the tangent-stiffness matrix is
@@
 \subsection{\emph{Peridigm} and peridynamics.}
+\label{subsec:PaP}
 %
 Comparisons of the different tangent-stiffness calculation techniques were
-carried out through code-development and \emph{in situ} instrumentation of the
+carried out through code-development and instrumentation of the
 computational peridynamics code \emph{Peridigm} \cite{peridigm}.
@@
 \subsubsection{Implementing the complex-step method in Peridigm} 
+\label{subsubsec:Impl}
 %
@@
 \section{Results and Discussion}
+\label{sec:Res}
 %
@@
 \section{Conclusions}
+\label{sec:Conc}
 %

### 2014-02-27 response/response.tex
more progress on II

@@
   \end{itemize}}
+  \end{enumerate}
 
@@
   \begin{enumerate}
+
   \item
-	1. Pg. 2: In line 14, explain what does the term “algorithmically consistent” mean.
+	Pg. 2: In line 14, explain what does the term “algorithmically consistent” mean.
 
@@
   \item
-	2. Pg. 2: In line 36, what do you mean by “discrimination”?
+	Pg. 2: In line 36, what do you mean by “discrimination”?
 
@@
 ` \item	
-	3. Pg. 2: In line 37, explain what do you mean by “in-situ instrumentation”.
+	Pg. 2: In line 37, explain what do you mean by “in-situ instrumentation”.
 
@@
   \item 
-	4. Pg. 3: In lines 19-29, specify the sections where each part of the manuscript is
+	Pg. 3: In lines 19-29, specify the sections where each part of the manuscript is
 included.
+
 {\color{red}  
@@
      \item
-      The correction is made in the manunscript.
+      The correction is made in the manuscript.
   \end{itemize}}
 
+  \item 
+Pg. 3: You may consider including the exact web address of the paper repository.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      The correction is made in the manuscript. An exact web address for the paper repository has been added.
+  \end{itemize}}
+
+  \item 
+Pg. 3: Please include the forward and centered ﬁnite difference expressions for
+easier comparison with (2); include the truncation error as well. You do not have to
+derive these expressions though.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      The correction is made in the manuscript. Expressions for forward and centered difference have been included for faster comparison with (2).  
+  \end{itemize}}
+
+  \item 
+Pg. 4: In Eq. (1), “$\partial f^2$” should be “$\frac{\partial^2 f}{\partial x^2}$”; similarly, for the third derivative.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      The correction is made in the manuscript.
+  \end{itemize}}
+
+  \item 
+Pg. 4: You may consider writing Eq. (2) with the truncation error
+
+{\color{red}  
+\begin{itemize}
+     \item
+      The correction is made in the manuscript.
+  \end{itemize}}
+
+9. Pg. 6: In lines 9-11, please add references to the “Poisson problem” and to the
+“mathematical optimization”.
+
+10. Pg. 7: Should Eqs. (4), (5), and (6) use “$\approx$” instead of “$=$”?
+
+11. Pg. 7: In line 28, note that (cid:126)u is not the current conﬁguration, but the displacement
+ﬁeld. Please revise the sentence.
+
+12. Pg. 7: Please provide an expression for ${F^int}_i$
+in the model used.
+
+13. Pg. 9: In line 26, please add the reference [12] below to “coupling to molecular
+simulations [29].”
+
+14. Pg. 10: Please add an illustration of the block undergoing tension in Section 2.2,
+for clarity. Also, clarify the displacement boundary conditions; are they applied in a
+volumetric layer?
+
+15. Pg. 10: In line 40, you mentioned a bandwidth of 7. This seems to be clear in 1D.
+Does this hold in higher dimensions? Please clarify.
+
+16. Pg. 10: In lines 47-49, you say “The speciﬁc parameters ... can be found ... in the
+paper repository.” It seems a good idea to have the paper repository with ﬁles that a
+user can eventually use. However, the content of the paper should be independent
+of the repository. Please include a table with the parameters used. Similarly, in
+page 11 (lines 11-13).
+
+17. Pg. 14: In line 55, is there any conclusion you can draw from the power law index
+$1.0 * 10^-6$?
+{\color{red}  
+\begin{itemize}
+     \item
+      Manuscript updated.
+More significant than the power law index is that the methods have the same power law relationship with computation time at all, meaning that there is some, but weak evidence that the methods have equal algorithm complexity.
+  \end{itemize}}
+
+18. Pgs. 14-: Please clarify the value of the meshsize used on each of the experiments.
+{\color{red}  
+\begin{itemize}
+     \item
+      [TODO]
+  \end{itemize}}
+
+19. Pg. 16: Would it be possible to rescale Fig. 2, so that the dependence of the time
+on the number of cores will look indeed linear?
+{\color{red}  
+\begin{itemize}
+     \item
+     Manuscript is in error and has been corrected. An accurate statement is that CPU time and number of cores are inversely related.
+  \end{itemize}}
+
+20. Pg. 16: In lines 54-56, you say “It should be pointed out that compared to the (cid:96)2 -
+norms of any of the TS matrices by themselves, the magnitude of the accuracy
+metric ... is relatively small.” Please include in the manuscript the explicit informa-
+tion on the relative error. It seems that you already have such information in the
+paper repository.
+{\color{red}  
+\begin{itemize}
+     \item
+[TODO]
+  \end{itemize}}
+
+21. Pg. 17: Include the values of the probe distances h used for each method.
+{\color{red}  
+\begin{itemize}
+     \item
+[TODO]
+     The probe distances used in Peridigm are automatically determined with a heuristic algorithm and are not set by the user.
+They are 1.E-6 times the node spacing [ADAGIO REF] for FD and CD which is determined to limit the onset of subtractive cancellation.
+1.E-100 was hard-coded for CS, while AD had no probe distance.
+  \end{itemize}}
+
+22. Pg. 17: In lines 53-54, you say “Accuracy trends matched those for the single core
+tests.” Indeed, comparing Figs. 3 and 4, the results for FD seem to be quite the
+same in serial and in parallel. For CS the accuracy seems higher in serial. For CD,
+the accuracy is also higher in serial and, in serial, it has a ﬂuctuating behavior. Can
+you comment on these observations?
+{\color{red}  
+\begin{itemize}
+     \item
+[TODO]
+	The fluctuation seems to be a function of the amount of nonzero TS matrix elements relative to total TS matrix elements.
+Information supporting this has been added to the manuscript.
+This may be related to the fact that neighborhood radius in terms of number of Peridynamic nodes was kept constant and may indicate that connectivity in the mesh oscillates when the mesh is refined incrementally.
+  \end{itemize}}
+
+23. Pgs. 17-18: What are the units of K ? If F int in Eq. (4) etc. is internal force, then K
+should have units of force/length. However, in Figs. 3 and 4 you use the unit of
+MPa, which is force/length2 . Please clarify.
+
+24. Pg. 19: Are the units of the vertical axis in Fig. 5 “seconds”, as in Figs. 1 and 2?
+Please indicate that. Also, is the label of the vertical axis in Fig. 6 correct?
+
+25. Pg. 20: In lines 32-35, clarify what “byte-copyable” mean and why is that important.
+{\color{red}  
+\begin{itemize}
+     \item
+     Clarification made in manuscript, new reference added.
+'Byte-copiable' means that members of an array of data elements are all of the same size in memory so that if one knew the position of the first element in the array, the position of all other elements in the array could be inferred with the additional information of the length of the array.
+This is important, in particular, to the designers and users of the thousands of MIC accelerator cards installed in the \emph{Stampede} computer system.
+MIC accelerator cards permit only the transfer of byte-copiable data from the host computer system for use with the '#pragma offload' mode, which is to say most code written for the MIC.
+This is even more important when it is realized that the great majority of \emph{Stampede}'s theoretical processing power is counted from these MIC accelerator cards.
+Meaning that if an AD user wanted to use any of that power, they would be out of luck since not even the full time personel at TACC could manage to compile the AD library Trilinos::Sacado for the MIC.
+A more 'casual' computational scientist would find the task impossible.
+  \end{itemize}}
+
+26. Pg. 21: Please include a section title for the references.
+
+27. Pg. 24: In line 23, you may add the truncation error O(h2 ) to the expression. Oth-
+erwise, you may need to replace the “=” sign by the “≈” sign. However, adding
+the error term would make more clear the limiting expression when h → 0. Also,
+replace the comma by a period.
+
+28. Pg. 24: Why do we need S and X in Appendix B?
+{\color{red}  
+\begin{itemize}
+     \item
+     All the members of set X need to be members of set S otherwise the approximations don't work.
+X are numbers from the 'real' world and S are numbers from the computer world.
+For example: "http://en.wikipedia.org/wiki/IEEE_floating_point" shows that computers and languages following IEEE standards don't represent very large numbers with very fine precision at the same time due to memory constraints, and there are upper and lower bounds on the magnitudes of nonzero numbers that can be represented.
+Obviously, approximation error is expected, but the fact that a user doesn't actually get 'all' the numbers is something that is not obvious.
+  \end{itemize}}
+
+29. Pgs. 25-26: Are the functions in the second column of the table in 2(c) (page 25)
+approximations? Why do you label the column “Approximates Function”? Simi-
+larly, in line 16 (page 26) you say “approximated by the partial derivative”; is it an
+approximation?
+{\color{red}  
+\begin{itemize}
+     \item
+      A computer can only provide finite precision approximations, of approxmations, of functions.
+Approximations of derivatives of functions are also approximations, therefor the label is correct in both cases.
+For example: "https://sourceware.org/git/?p=glibc.git;a=blob;f=sysdeps/x86_64/fpu/e_expl.S" is the implementation of the $exp(x)$ function on 80-bit extended precision floating point numbers in the popular GCC compiler.
+This implementation is the most accurate that GCC provides for the $e$ function on real numbers, but ultimately is still a discrete numerical algorithm.
+  \end{itemize}}
+
+30. Pg. 26: Please clarify whether storage is needed in automatic differentiation. For
+example, do we need to store w, v , and u?
+
+{\color{red}  
+\begin{itemize}
+     \item
+      Clarification is added in the manuscript.
+For an actual AD system like Trilinos::Sacado or ADIFOR, w, v and u (stand-ins for simple or elementary math operations) need to be written in code differently from the built-in implementation to replace all math operations as part of a complete AD datatype.
+What doesn't absolutely have to be stored are the evaluations of w, v and u.
+Not storing the evaluations happens in 'forward' AD, storing the evaluations happens in 'backward' or 'reverse accumulation' AD.
+  \end{itemize}}
+
+  \end{enumerate}
 \end{document}

### 2014-02-27 jacobian_calc_CMAME.tex
more progress on II

@@
 automatic-differentiation or complex-step to better encourage their use and
-discussion. This open source resource will be referred to as the \emph{paper
-repository} in what follows.
+discussion. This open source resource is available at the correposinding author's website: http://idl.utsa.edu/jtfoster/paper-repositories/
+ or directly at https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive.
 
@@
 \begin{align} f ( x + i h ) =& f(x) + \frac{\partial f(x)}{\partial x} \frac{i
-h}{1!}  \notag \\ &+ \frac{\partial f^2(x)}{\partial x^2} \frac{(i h)^2}{2!} +
-\frac{\partial f^3(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots,
+h}{1!}  \notag \\ &+ \frac{\partial^2 f(x)}{\partial x^2} \frac{(i h)^2}{2!} +
+\frac{\partial^3 f(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots,
 \label{eqn:complexTaylor} \end{align}
@@
 %
-\begin{equation} \frac{\partial f( x )}{\partial x} \approx \frac{\mbox{Im}
-\left( f (x + i h) \right)}{h}.  \label{eqn:complexFirstDeriv} \end{equation}
+\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{\mbox{Im}
+\left( f (x + i h) \right)}{h} +\mathcal{O}\left ( h^2 \right).  \label{eqn:complexFirstDeriv} \end{equation}
 %
@@
 derivative calculation of a function.
+
+For the reader's convenience the foward difference equivalent to~\ref{eqn:complexFirstDeriv} is
+%
+\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x)}{h} +\mathcal{O}\left ( h \right). \label{eqn:forwardFirstDeriv} \end{equation},
+%
+while the central difference equivalent to~\ref{eqn:complexFirstDeriv} is,
+%
+\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x - h)}{2h} +\mathcal{O}\left ( h^2 \right). \label{eqn:centeredFirstDeriv} \end{equation} \cite[Chap. 4.1.3]{chapra2010}.
+%
+    Note that the order of the truncation error in~\ref{eqn:forwardFirstDeriv} indicates that for a given $h$ forward difference is less accurate an approximation than central difference or complex step.
 
@@
 composition function with respect to a designated variable at a particular
-value. It is obvious, but bears mentioning, that the AD system could simply
+value. 
+%
+It is obvious, but bears mentioning, that the AD system could simply
 store one value for partial derivatives, modifying it as appropriate for every
@@
 order. Calculation time has a power-law relationship with problem size, with a
-power-law index of roughly $\num{1.0E-6}$ for each of the methods. This bodes
-well for the scalability of each of the methods by themselves in
-\emph{Peridigm}.
+power-law index of roughly $\num{1.0E-6}$ for each of the methods. This bodes well
+for the scalability of each of the methods by themselves in \emph{Peridigm}. If we
+assumed every individual operation, generally speaking, took the same amount of
+time on average such that compute time were a measure of number of operations,
+this suggests that the methods as in \emph{Peridigm} have the same order
+algorithm complexity when considering computing a TS matrix alone, that is
+computation time is $\mathcal{O}(10^{(10^{-6} \times \log_{10}n)})$. This would
+suggest, to the algorithm design way of thinking, that the methods are equally
+well suited to the task of computing a TS matrix, if it were further assumed
+that the TS matrices produced by each method were identical. This is not the case
+as seen in the following subsection.
 %
@@
 parallel simulations.  Figure~\ref{fig:multi_speed} indicates that when looking
-at average iteration time, the factor of speed improvement scales linearly with
+at average iteration time, the factor of speed improvement scales inversely with
 the number of compute cores used for the given conditions of this group of

### 2014-02-28 response/response.tex
readthrough and cleanup

@@
      \item
-      A brief definition for the convenience of the reader was added in the manuscript.
-But, algorithmically consistent is a fairly common term in plasticity/mechanics.
-This same meaning is used for 'consistent' refering to 'tangent matrices' in \cite[perez2012consistent, perez2010].
-It means consistent with integrating the plasticity model under the Khun-Tucker constraints.
-Integrating the plasticity model means, given a strain increment, determining a stress that does not exceed the yield stress, and by doing so computing a linear slope, called a consistent tangent modulus, that relates the strain increment to a correct stress increment.
-An alogorithmically consistent tangent stiffness matrix is then one that uses the consistent tangent modulii, making it a more accurate approximation of the stiffness of the system.
+
+      A brief definition for the convenience of the reader was added in the
+      manuscript. But, algorithmically consistent is a fairly common term in
+      plasticity/mechanics. This same meaning is used for 'consistent' refering
+      to 'tangent matrices' in \cite[perez2012consistent, perez2010]. It means
+      consistent with integrating the plasticity model under the Khun-Tucker
+      constraints. Integrating the plasticity model means, given a strain
+      increment, determining a stress that does not exceed the yield stress,
+      and by doing so computing a linear slope, called a consistent tangent
+      modulus, that relates the strain increment to a correct stress increment.
+      An alogorithmically consistent tangent stiffness matrix is then one that
+      uses the consistent tangent modulii, making it a more accurate
+      approximation of the stiffness of the system.
+
   \end{itemize}}
@@
      \item
+
       The word, 'discrimination' has been removed from the manuscript.
-'Discrimination' was meant to mean 'measuring differences and making judgements on relative usefulness'.
-Since this meaning is probably understood from the word 'comparison' alone, the word 'discrimination' has been removed from the manuscript for clarity.
+      'Discrimination' was meant to mean 'measuring differences and making
+      judgements on relative usefulness'. Since this meaning is probably
+      understood from the word 'comparison' alone, the word 'discrimination'
+      has been removed from the manuscript for clarity.
+
   \end{itemize}}
@@
      \item
-      The words, 'in-situ' have been removed from the manuscript.
-'In-situ' was meant to suggest that timing measurements were taken from within and by the program being analyzed itself, rather than by an external program, but it has been removed for clarity.
+
+      The words, 'in-situ' have been removed from the manuscript. 'In-situ' was
+      meant to suggest that timing measurements were taken from within and by
+      the program being analyzed itself, rather than by an external program,
+      but it has been removed for clarity.
+
   \end{itemize}}
@@
     Pg. 3: Please include the forward and centered ﬁnite difference expressions for
-easier comparison with (2); include the truncation error as well. You do not have to
-derive these expressions though.
+    easier comparison with (2); include the truncation error as well. You do not
+    have to derive these expressions though.
 
@@
 
-9. Pg. 6: In lines 9-11, please add references to the “Poisson problem” and to the
+\item
+Pg. 6: In lines 9-11, please add references to the “Poisson problem” and to the
 “mathematical optimization”.
 
-10. Pg. 7: Should Eqs. (4), (5), and (6) use “$\approx$” instead of “$=$”?
+{\color{red}  
+\begin{itemize}
+     \item
+         [TODO]
+         No changes made. 
+  \end{itemize}}
 
-11. Pg. 7: In line 28, note that (cid:126)u is not the current conﬁguration, but the displacement
+\item
+Pg. 7: Should Eqs. (4), (5), and (6) use “$\approx$” instead of “$=$”?
+
+{\color{red}  
+\begin{itemize}
+     \item
+         [TODO]
+         Manuscript corrected. These expressions are approximations.
+  \end{itemize}}
+
+\item
+Pg. 7: In line 28, note that (cid:126)u is not the current conﬁguration, but the displacement
 ﬁeld. Please revise the sentence.
 
-12. Pg. 7: Please provide an expression for ${F^int}_i$
-in the model used.
+{\color{red}  
+\begin{itemize}
+    \item
+        [TODO]
+        Manuscript corrected. 'u' has been removed from the sentence since while the equations are functions of displacement, 'u' is not a configuration. 
+        \end{itemize}}
 
-13. Pg. 9: In line 26, please add the reference [12] below to “coupling to molecular
+\item
+Pg. 7: Please provide an expression for ${F^int}_i$ in the model used.
+
+{\color{red}  
+\begin{itemize}
+    \item
+        [TODO]
+        Manuscript corrected.
+In the section 'Peridigm and peridynamics' a constitutive model for the LPS is shown.
+On the page that was originally page 7, the force density corresponding to this constitutive model is expressed in descrete form.
+  \end{itemize}}
+
+\item
+Pg. 9: In line 26, please add the reference [12] below to “coupling to molecular
 simulations [29].”
 
-14. Pg. 10: Please add an illustration of the block undergoing tension in Section 2.2,
+{\color{red}  
+\begin{itemize}
+    \item
+        The manuscript has been corrected.
+  \end{itemize}}
+
+\item
+Pg. 10: Please add an illustration of the block undergoing tension in Section 2.2,
 for clarity. Also, clarify the displacement boundary conditions; are they applied in a
@@
 
-15. Pg. 10: In line 40, you mentioned a bandwidth of 7. This seems to be clear in 1D.
+{\color{red}  
+\begin{itemize}
+    \item
+        [TODO]
+        An conceptual illustration of the block undergoing tensions has been included.
+Additionally is a close-up illustration and some exposition explaining how displacement boundary conditions are achieved in peridynamics.
+  \end{itemize}}
+
+\item
+Pg. 10: In line 40, you mentioned a bandwidth of 7. This seems to be clear in 1D.
 Does this hold in higher dimensions? Please clarify.
 
-16. Pg. 10: In lines 47-49, you say “The speciﬁc parameters ... can be found ... in the
+{\color{red}  
+\begin{itemize}
+    \item
+        [TODO]
+        Bandwidth was referring to connections between nodes rather than components and degrees of freedom.
+It would be more clear to say, a bandwidth of 21 for the 3D case.
+The manuscript has been corrected with this change.
+  \end{itemize}}
+
+    \item
+        In lines 47-49, you say “The speciﬁc parameters ... can be found ... in the
         paper repository.” It seems a good idea to have the paper repository with ﬁles that a
@@
 
-17. Pg. 14: In line 55, is there any conclusion you can draw from the power law index
+{\color{red}  
+\begin{itemize}
+    \item
+        [TODO]
+        Correction made in the manuscript. The test parameters not already reported throughout the manuscript include: material density
+        and loading rates and information relevant to loading rates.
+  \end{itemize}}
+
+  \item
+    Pg. 14: In line 55, is there any conclusion you can draw from the power law index
     $1.0 * 10^-6$?
+
 {\color{red}  
@@
       Manuscript updated.
-More significant than the power law index is that the methods have the same power law relationship with computation time at all, meaning that there is some, but weak evidence that the methods have equal algorithm complexity.
+More significant than the power law index is that the methods have the same power law relationship with computation time at all, meaning that there is some weak evidence that the methods have equal algorithm complexity.
   \end{itemize}}
 
-18. Pgs. 14-: Please clarify the value of the meshsize used on each of the experiments.
+  \item
+    Pgs. 14-: Please clarify the value of the meshsize used on each of the experiments.
+
 {\color{red}  
@@
      \item
-      [TODO]
+         No correction made. Justification follows.
+         %
+         Mesh sizes from smallest to largest were 1000, 2000, 4000, 8000,
+         16000, 32000, 1000000 peridynamic nodes, but this is not relevant to
+         the study, since the experiments were about the fullness of the
+         tangent stiffness matrix and mesh size only places bounds on the
+         mimimum and maximum fullness of the tangent stiffness matrix but does
+         not determine it.  Furthermore, other users should not take their
+         meshes of similar node count and make comparisons to the work based on that, since
+         differing connectivity between nodes could make a 2000 node mesh more
+         computationally challenging than a 1000000 node mesh.
+         %
+         The most extreme, but illustrative example of this is a stretched rope shaped mesh versus a cannonball shaped mesh of nearly overlapping nodes.
+         %
+         Instead, a better measure of problem size for this study was number of nonzero jacobian matrix elements, which was reported for the experiments in the work.
+Additionally, in \emph{Peridigm}, only the nonzero elements are stored in memory at all.
+This measure is a function of model dimensions, node spacing and horizon and not usefully related to number of nodes unless the meshes are at maximum or minimum connectivity.
       \end{itemize}}
 
-19. Pg. 16: Would it be possible to rescale Fig. 2, so that the dependence of the time
+  \item
+    Pg. 16: Would it be possible to rescale Fig. 2, so that the dependence of the time
     on the number of cores will look indeed linear?
+
 {\color{red}  
@@
 
-20. Pg. 16: In lines 54-56, you say “It should be pointed out that compared to the (cid:96)2 -
+  \item
+    Pg. 16: In lines 54-56, you say “It should be pointed out that compared to the (cid:96)2 -
     norms of any of the TS matrices by themselves, the magnitude of the accuracy
@@
     paper repository.
+
 {\color{red}  
@@
 [TODO]
+    The manuscript has been corrected. 
+    %
+    The Frobenius norm of the TS matrices for each of the experiments was around 150 thousand, while the discrepancy between the matrices produced by the different methods, using the same metric, was on the order of 10 E-7.
+The relative error is on the order of 10 E-14.
     \end{itemize}}
 
-21. Pg. 17: Include the values of the probe distances h used for each method.
+  \item
+    Pg. 17: Include the values of the probe distances h used for each method.
+
 {\color{red}  
@@
 [TODO]
-     The probe distances used in Peridigm are automatically determined with a heuristic algorithm and are not set by the user.
-They are 1.E-6 times the node spacing [ADAGIO REF] for FD and CD which is determined to limit the onset of subtractive cancellation.
-1.E-100 was hard-coded for CS, while AD had no probe distance.
+    Correction made in the manuscript, values used are collected and the reference to the material explaining these values was made more clear.
+    %
+    The probe distances used in Peridigm are automatically determined with a
+    heuristic algorithm and are not set by the user. The algorithm for this in
+    \emph{Peridigm} is the used for another Sandia software project called
+    \emph{Adagio}. The probe distances are 1.E-6 times the node spacing [ADAGIO
+    REF] for FD and CD which is indended to prevent the onset of roundoff error.
+    1.E-100 was hard-coded for CS into the CS method implementation, while AD had no probe distance.
   \end{itemize}}
 
-22. Pg. 17: In lines 53-54, you say “Accuracy trends matched those for the single core
+  \item
+    Pg. 17: In lines 53-54, you say “Accuracy trends matched those for the single core
     tests.” Indeed, comparing Figs. 3 and 4, the results for FD seem to be quite the
@@
     you comment on these observations?
+
 {\color{red}  
@@
 
-23. Pgs. 17-18: What are the units of K ? If F int in Eq. (4) etc. is internal force, then K
+  \item
+    Pgs. 17-18: What are the units of K ? If F int in Eq. (4) etc. is internal force, then K
     should have units of force/length. However, in Figs. 3 and 4 you use the unit of
@@
 
-24. Pg. 19: Are the units of the vertical axis in Fig. 5 “seconds”, as in Figs. 1 and 2?
+{\color{red}  
+\begin{itemize}
+     \item
+[TODO]
+    The manuscript has been corrected to clarify that in Peridynamics, there is only force density rather than internal force.
+The units of K are actually force / length4, and this is corrected in the manuscript.
+  \end{itemize}}
+
+    \item
+        Pg. 19: Are the units of the vertical axis in Fig. 5 “seconds”, as in Figs. 1 and 2?
         Please indicate that. Also, is the label of the vertical axis in Fig. 6 correct?
 
-25. Pg. 20: In lines 32-35, clarify what “byte-copyable” mean and why is that important.
+{\color{red}  
+\begin{itemize}
+     \item
+[TODO]
+        The manuscript has been corrected. The labels on the vertical axis in Fig. 5 were incorrect. 
+  \end{itemize}}
+
+Pg. 20: In lines 32-35, clarify what “byte-copyable” mean and why is that important.
+
 {\color{red}  
@@
      Clarification made in manuscript, new reference added. 
-'Byte-copiable' means that members of an array of data elements are all of the same size in memory so that if one knew the position of the first element in the array, the position of all other elements in the array could be inferred with the additional information of the length of the array.
-This is important, in particular, to the designers and users of the thousands of MIC accelerator cards installed in the \emph{Stampede} computer system.
-MIC accelerator cards permit only the transfer of byte-copiable data from the host computer system for use with the '#pragma offload' mode, which is to say most code written for the MIC.
-This is even more important when it is realized that the great majority of \emph{Stampede}'s theoretical processing power is counted from these MIC accelerator cards.
-Meaning that if an AD user wanted to use any of that power, they would be out of luck since not even the full time personel at TACC could manage to compile the AD library Trilinos::Sacado for the MIC.
+     %
+     'Byte-copiable' means that members of an array of data elements are all of
+     the same size in memory so that if one knew the position of the first
+     element in the array, the position of all other elements in the array
+     could be inferred with the additional information of the length of the
+     array. This is important, in particular, to the designers and users of the
+     thousands of MIC accelerator cards installed in the \emph{Stampede}
+     computer system. MIC accelerator cards permit only the transfer of
+     byte-copiable data from the host computer system for use with the '#pragma
+     offload' mode, which is to say most code written for the MIC. This is even
+     more important when it is realized that the great majority of
+     \emph{Stampede}'s theoretical processing power, the figure used to advertise to potential clients, is counted from these MIC
+     accelerator cards. Meaning that if an AD user wanted to use any of that
+     power, they would be out of luck since not even the full time personel at
+     TACC could manage to compile the AD library Trilinos::Sacado for the MIC.
      A more 'casual' computational scientist would find the task impossible.
@@
 
-26. Pg. 21: Please include a section title for the references.
+ \item
+    Pg. 21: Please include a section title for the references.
 
-27. Pg. 24: In line 23, you may add the truncation error O(h2 ) to the expression. Oth-
+{\color{red}  
+\begin{itemize}
+     \item
+         [TODO]
+         Manuscript corrected.
+  \end{itemize}}
+
+ \item
+    Pg. 24: In line 23, you may add the truncation error O(h2 ) to the expression. Oth-
     erwise, you may need to replace the “=” sign by the “≈” sign. However, adding
@@
 
-28. Pg. 24: Why do we need S and X in Appendix B?
+{\color{red}  
+\begin{itemize}
+     \item
+         [TODO]
+         Manuscript corrected.
+  \end{itemize}}
+
+\item    
+Pg. 24: Why do we need S and X in Appendix B?
+
 {\color{red}  
@@
 
-29. Pgs. 25-26: Are the functions in the second column of the table in 2(c) (page 25)
+\item
+    Pgs. 25-26: Are the functions in the second column of the table in 2(c) (page 25)
     approximations? Why do you label the column “Approximates Function”? Simi-
@@
     approximation?
+
 {\color{red}  
@@
 
-30. Pg. 26: Please clarify whether storage is needed in automatic differentiation. For
+  \item
+Pg. 26: Please clarify whether storage is needed in automatic differentiation. For
 example, do we need to store w, v , and u?
@@
      \item
+         [TODO]
       Clarification is added in the manuscript.

### 2014-03-02 response/response.tex
added mesh diagrams, fixed spelling throughout

@@
          [TODO]
-         No changes made. 
+
   \end{itemize}}
@@
      \item
-         [TODO]
          Manuscript corrected. These expressions are approximations.
@@
     \item
-        [TODO]
-        Manuscript corrected. 'u' has been removed from the sentence since while the equations are functions of displacement, 'u' is not a configuration. 
+        Manuscript corrected.
+Sentence was revised to mean that F, evaluated in the current configuration, is a function of the displacement rather than to say that the displacement is the current configuration.
         \end{itemize}}
@@
     \item
-        [TODO]
-        An conceptual illustration of the block undergoing tensions has been included.
-Additionally is a close-up illustration and some exposition explaining how displacement boundary conditions are achieved in peridynamics.
+        The manuscript has been corrected.
+Viewgraphs of the 2000 node and 1000000 node blocks undergoing tension has been included.
+The images depict colorized meshes in a way that is used in the revised manuscript to help explain what boundary conditions or 'volume constraints' are like in peridynamics.
+Displacement boundary conditions are typically applied to a volumetric layer, and in the study shown in the manuscript that is they way they were applied.
   \end{itemize}}
@@
     \item
-        [TODO]
+        The manuscript has been corrected.
 Bandwidth was referring to connections between nodes rather than components and degrees of freedom.
-It would be more clear to say, a bandwidth of 21 for the 3D case.
-The manuscript has been corrected with this change.
+It is more clear to say a bandwidth of 21 for the 3D case.
+The formula is number of nodes in a neighborhood times degrees of freedom.
   \end{itemize}}
@@
         [TODO]
-        Correction made in the manuscript. The test parameters not already reported throughout the manuscript include: material density
-        and loading rates and information relevant to loading rates.
+        Correction made in the manuscript. 
   \end{itemize}}
@@
      \item
-         No correction made. Justification follows.
-         %
-         Mesh sizes from smallest to largest were 1000, 2000, 4000, 8000,
-         16000, 32000, 1000000 peridynamic nodes, but this is not relevant to
-         the study, since the experiments were about the fullness of the
-         tangent stiffness matrix and mesh size only places bounds on the
-         mimimum and maximum fullness of the tangent stiffness matrix but does
-         not determine it.  Furthermore, other users should not take their
-         meshes of similar node count and make comparisons to the work based on that, since
-         differing connectivity between nodes could make a 2000 node mesh more
-         computationally challenging than a 1000000 node mesh.
-         %
-         The most extreme, but illustrative example of this is a stretched rope shaped mesh versus a cannonball shaped mesh of nearly overlapping nodes.
-         %
-         Instead, a better measure of problem size for this study was number of nonzero jacobian matrix elements, which was reported for the experiments in the work.
-Additionally, in \emph{Peridigm}, only the nonzero elements are stored in memory at all.
-This measure is a function of model dimensions, node spacing and horizon and not usefully related to number of nodes unless the meshes are at maximum or minimum connectivity.
+         Manuscript is corrected. Mesh sizes from smallest to largest were 1000, 2000, 3000, 4000, 6000, 8000, 12000,
+         16000, 32000, 1000000 peridynamic nodes.
       \end{itemize}}
@@
 [TODO]
-    The manuscript has been corrected. 
-    %
-    The Frobenius norm of the TS matrices for each of the experiments was around 150 thousand, while the discrepancy between the matrices produced by the different methods, using the same metric, was on the order of 10 E-7.
-The relative error is on the order of 10 E-14.
+    The manuscript has been corrected. These measurements are included in the output from the experiments.
     \end{itemize}}
@@
 
+\item
 Pg. 20: In lines 32-35, clarify what “byte-copyable” mean and why is that important.

### 2014-03-02 jacobian_calc_CMAME.tex
added mesh diagrams, fixed spelling throughout

@@
 automatic-differentiation or complex-step to better encourage their use and
-discussion. This open source resource is available at the correposinding author's website: http://idl.utsa.edu/jtfoster/paper-repositories/
+discussion. This open source resource is available at the corresponding author's website: http://idl.utsa.edu/jtfoster/paper-repositories/
  or directly at https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive.
@@
 
-For the reader's convenience the foward difference equivalent to~\ref{eqn:complexFirstDeriv} is
+For the reader's convenience the forward difference equivalent to~\ref{eqn:complexFirstDeriv} is
 %
@@
 A work by A. H. Al-Mohy and N. J. Higham investigates the use of complex step
-in computing Frechet derivatives of matix to matrix functions
+in computing Frechet derivatives of matrix to matrix functions
 \cite{al2010complex}. To compute the Jacobian of the residual in the
@@
 \begin{equation} 
-  K_{ij} = \frac{F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u})}{h},
+  K_{ij} \approx \frac{F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u})}{h},
 \end{equation}
 %
-Where $K_{ij}$ is the indicial notation representation of an element of the tangent-stiffness matrix at row $i$, column $j$.
-The first of the two terms in the numerator is the internal force, $F_i^{int}$ evaluated in the current deformed configuration, $\vec{u}$,  plus a small perturbation $h$ in the direction $\hat{e}_j$ where $\hat{e}_j$ represents a unit-vector corresponding to the $j^{\mbox{th}}$ degree-of-freedom.
-The second of the two terms is $F_i^{int}$ evaluated in the current configuration.
-The denominator is the magnitude of perturbation called the \emph{probe-distance}.
+
+Where $K_{ij}$ is the indicial notation representation of an element of the
+tangent-stiffness matrix at row $i$, column $j$. The first of the two terms in
+the numerator is the internal force, $F_i^{int}$ evaluated in the current
+deformed configuration, as a function of the displacement, $\vec{u}$,  plus a small perturbation $h$ in the
+direction $\hat{e}_j$ where $\hat{e}_j$ represents a unit-vector corresponding
+to the $j^{\mbox{th}}$ degree-of-freedom. The second of the two terms is
+$F_i^{int}$ evaluated in the current configuration. The denominator is the
+magnitude of perturbation called the \emph{probe-distance}.
 
@@
 \begin{equation} 
-  K_{ij} = \frac{F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u} - h \hat{e}_j)}{2 h}.
+  K_{ij} \approx \frac{F_i^{int}(\vec{u} + h \hat{e}_j) - F_i^{int}(\vec{u} - h \hat{e}_j)}{2 h}.
 \end{equation}
@@
 %
-\begin{equation} K_{ij} = \frac{\mbox{Im}(F_i^{int}(\vec{u} + i h
+\begin{equation} K_{ij} \approx \frac{\mbox{Im}(F_i^{int}(\vec{u} + i h
 \hat{e}_j))}{h}, \end{equation}
@@
 %
-\begin{equation} K_{ij} = \frac{\partial F_i^{int}(\vec{u})}{\partial u_j},
+\begin{equation} K_{ij} \approx \frac{\partial F_i^{int}(\vec{u})}{\partial u_j},
 \end{equation}
@@
 of the linear peridynamic solid model used in this work. After the peridynamic
-equation of motion and a consitutive model have been discussed, a definition of
+equation of motion and a constitutive model have been discussed, a definition of
 the tangent stiffness matrix is given. The following discussion is largely
@@
 \note[Michael]{re: I.2} \add[Michael]{
-\subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simualtion.}
+\subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simulation.}
 The tangent stiffness matrix is the Frechet derivative of the residual wrt the
@@
     underlying contained data. Templated container classes allow for user
-    choice in what properties that underlying data exihibits.  For example, in
+    choice in what properties that underlying data exhibits.  For example, in
     this work, complex arithmetic cannot be performed on the data contained
@@
     code and by necessity in the modified code. These concepts are not vital to
-understanding the complex step method iteself, but may be necessary to understand complex
+understanding the complex step method itself, but may be necessary to understand complex
 step in \emph{Peridigm}.} To see CS and other methods in a more narrow context,
@@
 node spacing in the particle discretization scheme). This results in a minimum
-tangent-stiffness matrix bandwidth of 7 (for an optimal node numbering scheme).
+tangent-stiffness matrix bandwidth of 21 for a 3D problem (for an optimal node
+numbering scheme). Pictured in Fig.~\ref{fig:SCMesh} is a viewgraph for one of
+the peridynamic meshes used in this study. The coloration in the image
+indicates node numbering, where a smooth color transition means that the nodes
+have close ID numbers. The abrupt color changes in the center and at the ends
+of the body show that numbering is not contiguous and is assigned with respect to four regions,
+called \emph{material blocks} in \emph{Peridigm}. The purpose of designating
+these material blocks, for this study, is to register the blocks on the ends of
+the mesh as volumetric regions over which to apply displacement boundary
+conditions called \emph{volume constraints} in peridynamics.
+
+\begin{figure}[tbp] \centering \includegraphics[width=0.7\textwidth]{./figs/mesh2000.png}
+\caption{2000 node mesh} \label{fig:SCMesh}
+\end{figure}
 
@@
 The specific parameters of these tests can be found in \emph{Peridigm} input
-{\tt xml} files included in the paper repository. These {\tt xml} files allow
+{\tt xml} files included in the paper repository, however the parameters
+are reproduced in Table~\ref{tab:TestParams}. These {\tt xml} files allow
 users with the appropriately modified version of \emph{Peridigm} to reproduce
 the results found in this paper.
+
+\begin{table*}[!ht]    
+  \scriptsize
+  \centering
+        \caption{Test Parameters} \label{tab:TestParams}   
+       \begin{tabular}{c c c c c c c}
+         \toprule
+         Peri. nodes & Nozero TS elems. & Nonzero Fraction & Horizon & Load rate & # Load steps & Abs. tolerance \\ 
+                     & $\times 10^6$    &                  &\si{\meter}& \si{\centimeter\per\second} &   &     \\
+        \midrule
+        1000 & $\num{1.99}$             &                   & \num{.3}   & 2.0 & 2 & \num{10^-8}\\
+        2000 & $\num{4.25}$             &                   &            &     &   &            \\
+        3000 & $\num{7.79}$             &                   &            &     &   &            \\
+        4000 & $\num{15.4}$             &                   &            &     &   &            \\
+        6000 & $\num{20.0}$             &                   &            &     &   &            \\
+        8000 & $\num{31.4}$             &                   &            &     &   &            \\
+        12000 & $\num{40.1}$            &                   &            &     &   &            \\
+        16000 & $\num{83.9}$            &                   &            &     &   &            \\
+        32000 & $\num{165.0}$           &                   &            &     &   &            \\
+        1000000 & $\num{1670}$          &                   & \num{.03}  & 2.0 & 2 & \num{10^-8}\\
+        \bottomrule
+    \end{tabular}
+\end{table*}
 
@@
 that message passing computation did not overwhelm the simulation.  The
-specific parameters of these tests can be found in \emph{Perdigm} input {\tt
-xml} files also included in the paper repository. \note[Michael]{re:
-I.3}\add[Michael]{For the purposes of this work message passing is what
+specific parameters of these tests can be found in \emph{Peridigm} input {\tt
+xml} files also included in the paper repository, but a summary of the most relevant
+ones appears in Table~\ref{tab:TestParams}. A viewgraph of the 1 million node mesh is
+shown in Fig.~\ref{fig:MCMesh}. The nodes in the image, although actually discrete points
+in space are dense enough to appear as a continuous body at a distance.
+
+\begin{figure}[tbp] \centering \includegraphics[width=0.7\textwidth]{./figs/mesh1000000.png}
+\caption{1000000 node mesh} \label{fig:MCMesh}
+\end{figure}
+
+\note[Michael]{re: I.3}\add[Michael]{For the purposes of this work message passing is what
 synchronizes multiple computers that each are working on a separate section
@@
 contained within a subdomain solely managed by a single computer, relative
-to the amount of information shared accross multiple subdomains and
+to the amount of information shared across multiple subdomains and
 therefor multiple computers. A useful experiment to illustrate this idea
 concretely would be to implement a code to solve a 2D Laplace equation for
-simulating a simple heat condution problem on two processors. Use the
+simulating a simple heat conduction problem on two processors. Use the
 textbook finite difference method, which in this case does not refer just
@@
 contiguous chunks, having one shared border, and distribute these chunks,
-or subdomains to the two processors. Perfom the simulation a multitude of
+or subdomains to the two processors. Perform the simulation a multitude of
 times to calculate an average runtime.  Next implement the solution a
@@
 of decomposing the domain will cause much overlap between the subdomains
-managed by the two processors and much time will be spent synchronzing the
+managed by the two processors and much time will be spent synchronizing the
 subdomain overlap or 'ghost nodes' between the processors. Run the modified
@@
 the system solved in the study. The assumption that AD is accurate enough to
-serve as a standard is supported by ADs implementation as a computerized chain
+serve as a standard is supported by AD's implementation as a computerized chain
 rule as explained in Section~\ref{ADsubsection}.
@@
 Table~\ref{tab:results} shows the compute core count, number of nonzero tangent-stiffness (TS) non-zero matrix elements, load step average calculation time and load step average accuracy.
-The units for accuracy are derived equation~(\ref{eqn:accuracy}) and the units used to define bulk material properties, mentioned in Section~\ref{tcs}.
+The units for accuracy are derived Eqn.~(\ref{eqn:accuracy}) and the units used to define bulk material properties, mentioned in Section~\ref{tcs}.
 %
@@
 Data for accuracy measurements were taken according to the methods mentioned in
-Section~\ref{tcs}, and reduced using equation~(\ref{eqn:accuracy}). The
+Section~\ref{tcs}, and reduced using Eqn.~(\ref{eqn:accuracy}). The
 averaged results of these accuracy measurements for the serial tests can be
@@
 in the simulation data which is available in the paper repository. The effect
-of these accuracy differences on convergence rate of the Newton solver was not
-determined.  
+of these accuracy differences on convergence rate of the Newton solver was studied and the
+results are reported in section~\ref{sec:PeridigmConvergenceStudy}.  
 %
@@
 normally. The question is posed, given that step-size $h$ is heuristically
-selected by \emph{Perdigm} for the finite difference methods and presumable
+selected by \emph{Peridigm} for the finite difference methods and presumable
 makes a good choice, do the single-iteration Jacobian accuracy results predict
@@
 greater than allowed by the conditions seen in \cite{burkeLectures}, 
-reproduced in Subsection~\ref{subsec:FurtherStudies}.
+reproduced in subsection~\ref{subsec:FurtherStudies}.
 

### 2014-03-03 response/response.tex
completed first draft of small issues section

@@
       plasticity/mechanics. This same meaning is used for 'consistent' refering
-      to 'tangent matrices' in \cite[perez2012consistent, perez2010]. It means
+      to 'tangent matrices' in \cite{perez2012numerical, perez2000numerical}. It means
       consistent with integrating the plasticity model under the Khun-Tucker
@@
     \item
-        [TODO]
         Correction made in the manuscript. 
@@
      \item
-[TODO]
-    The manuscript has been corrected. These measurements are included in the output from the experiments.
+    The manuscript has been corrected. 
     \end{itemize}}
@@
      \item
-[TODO]
     Correction made in the manuscript, values used are collected and the reference to the material explaining these values was made more clear.
@@
     \emph{Peridigm} is the used for another Sandia software project called
-    \emph{Adagio}. The probe distances are 1.E-6 times the node spacing [ADAGIO
-    REF] for FD and CD which is indended to prevent the onset of roundoff error.
+    \emph{Adagio} \cite{ref-Adaggio}. The probe distances are 1.E-6 times the node spacing for FD and CD which is indended to prevent the onset of roundoff error.
     1.E-100 was hard-coded for CS into the CS method implementation, while AD had no probe distance.
@@
      \item
-[TODO]
-	The fluctuation seems to be a function of the amount of nonzero TS matrix elements relative to total TS matrix elements.
-Information supporting this has been added to the manuscript.
-This may be related to the fact that neighborhood radius in terms of number of Peridynamic nodes was kept constant and may indicate that connectivity in the mesh oscillates when the mesh is refined incrementally.
+     	Manuscript has been updatded.
+When accuracy is looked at in terms of relative error, where relative error is the accuracy metric previously plotted in the figures divided by the Frobenius norm of the AD TS matrix alone times 100, there is actually little fluctuation, between tests or between serial and parallel tests.
+The order of magnitude for this measure remains steady for all the methods.
   \end{itemize}}
@@
      \item
-[TODO]
-    The manuscript has been corrected to clarify that in Peridynamics, there is only force density rather than internal force.
-The units of K are actually force / length4, and this is corrected in the manuscript.
+    The manuscript has been corrected.
+The units of K are actually force / length4, however there is no intuitive relationship between the norm of the TS matrix and some physical characteristic of the system being modeled.
   \end{itemize}}
@@
      \item
-[TODO]
         The manuscript has been corrected. The labels on the vertical axis in Fig. 5 were incorrect. 
@@
      \item
-         [TODO]
          Manuscript corrected.
@@
      \item
-     All the members of set X need to be members of set S otherwise the approximations don't work.
-X are numbers from the 'real' world and S are numbers from the computer world.
-For example: "http://en.wikipedia.org/wiki/IEEE_floating_point" shows that computers and languages following IEEE standards don't represent very large numbers with very fine precision at the same time due to memory constraints, and there are upper and lower bounds on the magnitudes of nonzero numbers that can be represented.
-Obviously, approximation error is expected, but the fact that a user doesn't actually get 'all' the numbers is something that is not obvious.
+
+     All the members of set X need to be members of set S otherwise the
+     approximations don't work. X are numbers from the 'real' world and S are
+     numbers from the computer world. For example:
+
+     \begin{verbatim}
+     "http://en.wikipedia.org/wiki/IEEE_floating_point" 
+     \end{verbatim}
+
+     shows that computers
+     and languages following IEEE standards don't represent very large numbers
+     with very fine precision at the same time due to memory constraints, and
+     there are upper and lower bounds on the magnitudes of nonzero numbers that
+     can be represented. Obviously, approximation error is expected, but the
+     fact that a user doesn't actually get 'all' the numbers is something that
+     is not obvious.
+
   \end{itemize}}
@@
      \item
-      A computer can only provide finite precision approximations, of approxmations, of functions.
-Approximations of derivatives of functions are also approximations, therefor the label is correct in both cases.
-For example: "https://sourceware.org/git/?p=glibc.git;a=blob;f=sysdeps/x86_64/fpu/e_expl.S" is the implementation of the $exp(x)$ function on 80-bit extended precision floating point numbers in the popular GCC compiler.
-This implementation is the most accurate that GCC provides for the $e$ function on real numbers, but ultimately is still a discrete numerical algorithm.
+
+      Strictly speaking, a computer can only provide finite precision
+      approximations, of approxmations, of functions. Approximations of
+      derivatives of functions are also approximations, therefor the label is
+      correct in both cases. For example:
+
+      \begin{verbatim}
+      "https://sourceware.org/git/?p=glibc.git;a=blob;f=sysdeps/x86_64/fpu/e_expl.S"
+      \end{verbatim}
+
+      is the implementation of the $exp(x)$ function on 80-bit extended
+      precision floating point numbers in the popular GCC compiler. This
+      implementation is the most accurate that GCC provides for the $e$
+      function on real numbers, but ultimately is still a discrete numerical
+      algorithm.
+
   \end{itemize}}
@@
      \item
-         [TODO]
       Clarification is added in the manuscript.  
-For an actual AD system like Trilinos::Sacado or ADIFOR, w, v and u (stand-ins for simple or elementary math operations) need to be written in code differently from the built-in implementation to replace all math operations as part of a complete AD datatype.
-What doesn't absolutely have to be stored are the evaluations of w, v and u.
-Not storing the evaluations happens in 'forward' AD, storing the evaluations happens in 'backward' or 'reverse accumulation' AD.
   \end{itemize}}
@@
   \end{enumerate}
+
+\bibliographystyle{elsarticle-num}
+\bibliography{all}
 \end{document}

### 2014-03-03 jacobian_calc_CMAME.tex
completed first draft of small issues section

@@
 have close ID numbers. The abrupt color changes in the center and at the ends
-of the body show that numbering is not contiguous and is assigned with respect to four regions,
-called \emph{material blocks} in \emph{Peridigm}. The purpose of designating
-these material blocks, for this study, is to register the blocks on the ends of
-the mesh as volumetric regions over which to apply displacement boundary
+of the body show that numbering is not contiguous and is assigned with respect to four regioins. The purpose of designating
+these regions, for this study, was to apply displacement boundary
 conditions called \emph{volume constraints} in peridynamics.
@@
         \caption{Test Parameters} \label{tab:TestParams}   
-       \begin{tabular}{c c c c c c c}
+       \begin{tabular}{c c c c c c}
          \toprule
-         Peri. nodes & Nozero TS elems. & Nonzero Fraction & Horizon & Load rate & # Load steps & Abs. tolerance \\ 
-                     & $\times 10^6$    &                  &\si{\meter}& \si{\centimeter\per\second} &   &     \\
+         Peri. nodes & Nozero TS elems. & Horizon & Load rate & # Load steps & Abs. tolerance \\ 
+                     & $\times 10^6$    & \si{\meter}& \si{\centimeter\per\second} &   &     \\
         \midrule
-        1000 & $\num{1.99}$             &                   & \num{.3}   & 2.0 & 2 & \num{10^-8}\\
-        2000 & $\num{4.25}$             &                   &            &     &   &            \\
-        3000 & $\num{7.79}$             &                   &            &     &   &            \\
-        4000 & $\num{15.4}$             &                   &            &     &   &            \\
-        6000 & $\num{20.0}$             &                   &            &     &   &            \\
-        8000 & $\num{31.4}$             &                   &            &     &   &            \\
-        12000 & $\num{40.1}$            &                   &            &     &   &            \\
-        16000 & $\num{83.9}$            &                   &            &     &   &            \\
-        32000 & $\num{165.0}$           &                   &            &     &   &            \\
-        1000000 & $\num{1670}$          &                   & \num{.03}  & 2.0 & 2 & \num{10^-8}\\
+        1000 & 1.99             & .3   & 2.0 & 2 & 10^-8\\
+        2000 & 4.25             & .24  & 2.0 & 2 & 10^-8\\
+        3000 & 7.79             & .21  & 2.0 & 2 & 10^-8\\
+        4000 & 15.4             & .19  & 2.0 & 2 & 10^-8\\
+        6000 & 20.0             & .16  & 2.0 & 2 & 10^-8\\
+        8000 & 31.4             & .15  & 2.0 & 2 & 10^-8\\
+        12000 & 40.1            & .131 & 2.0 & 2 & 10^-8\\
+        16000 & 83.9            & .12  & 2.0 & 2 & 10^-8\\
+        32000 & 165.0           & .095 & 2.0 & 2 & 10^-8\\
+        1000000 & 1670          & .03  & 2.0 & 2 & 10^-8\\
         \bottomrule
@@
 accuracy metric calculated with Eqn.~\ref{eqn:accuracy} for any of the methods
-is relatively small. These values for the $l^2$-norms of the TS matrices appear
+is relatively small. Relative error was computed from these measurements and is reported in
+Table~\ref{tab:RelAcc}. The values for the $l^2$-norms of the TS matrices appear
 in the simulation data which is available in the paper repository. The effect
@@
 used in this study is the default value used in \emph{Peridigm} which is
-heuristically derived based on computational node spacing.  One of the beauties
+heuristically derived based on computational node spacing using the method explained in \cite[pp. 90]{ref-Adaggio}.  
+\add[Michael]{To summarize, a spacing of $1.E-6$ times grid spacing is used to provide acceptable accuracy but prevent roundoff error, 
+as is not a user determined setting in \emph{Peridigm}.} One of the beauties
 of the CS method is that since dependence on $h$ decreases for small $h$, which
@@
 series accuracy measurements.} \label{fig:multi_accuracy} \end{figure}
+%
+It can be seen in Figs.~\ref{fig:serial_accuracy} and ~\ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods.
+Those figures are depicting absolute error.
+If relative error is computed from the absolute error measurements by dividing them by the Frobenius norm of the appropriate AD produced TS matrix, it can be seen as in Table~\ref{tab:RelAcc}, that there is little variation in the accuracy of the methods between tests or between serial and parallel test groups.
+
+\begin{table*}[!ht]    
+  \scriptsize
+  \centering
+        \caption{Relative accuracy results} \label{tab:RelAcc}   
+       \begin{tabular}{c c c c c}
+         \toprule
+         Cores & TS Elements & \multicolumn{3}{c}{Percent error} \\ 
+         & $\times 10^6$ & CD & CS & FD \\
+        \midrule
+        1 & $1.99$  & 1.28x10^-8 & 2.29x10^-15 & 1.46x10^-5\\
+        1 & $4.25$  & 8.86x10^-8 & 2.25x10^-14 & 1.34x10^-5\\
+        1 & $7.79$  & 1.37x10^-8 & 2.27x10^-14 & 1.25x10^-5\\
+        1 & $15.4$  & 4.33x10^-8 & 2.52x10^-14 & 1.16x10^-5\\
+        1 & $20.0$  & 7.97x10^-8 & 2.18x10^-15 & 1.13x10^-5\\
+        1 & $31.4$  & 1.44x10^-8 & 2.29x10^-15 & 1.09x10^-5\\
+        1 & $40.1$  & 1.36x10^-7 & 2.24x10^-15 & 1.05x10^-5\\
+        1 & $83.9$  & 1.30x10^-7 & 2.44x10^-14 & 9.63x10^-6\\
+        1 & $165.0$  &1.30x10^-7 & 2.44x10^-14 & 9.63x10^-6\\
+        32 & $1670$  &6.00x10^-7 & 2.29x10^-14 & 6.93x10^-6\\
+        64 & $1670$  & 6.01x10^-7 & 2.31x10^-14 & 7.02x10^-6\\
+        96 & $1670$  & 6.03x10^-7 & 2.33x10^-14 & 7.10x10^-6\\
+        128 & $1670$  & 5.97x10^-7 & 2.34x10^-14 & 7.19x10^-6\\
+        \bottomrule
+    \end{tabular}
+\end{table*}
 
@@
 
+\section{References}
+
 \bibliographystyle{elsarticle-num}
@@
 \[
-\frac{\partial f(x)}{\partial x} = - \frac{\sin(x) \sinh(h)}{h},
+    \frac{\partial f(x)}{\partial x} = - \frac{\sin(x) \sinh(h)}{h} + \mathcal{O}(h^2).
 \]
@@
 %This explains what the equation and table above mean.
-Because the computer knew the analytical forms of the partial derivatives of each of $u,v,w$ beforehand,
-all it needed to do was:
+Because the computer knew the analytical forms of the partial derivatives of each of $u,v,w$ beforehand, which of course needed to be specified using an AD data structure from a library like Trilinos::Sacado \cite{trilinos} or ADIFOR \cite{bischof1995adifor},
+all it needed to do (for 'backward' AD) was:
 \begin{enumerate}

### 2014-03-03 response/response.tex
changes denoted more plainly

@@
       Manuscript updated.
-More significant than the power law index is that the methods have the same power law relationship with computation time at all, meaning that there is some weak evidence that the methods have equal algorithm complexity.
+More significant than the power law index is that the methods have the same power law relationship with computation time at all, meaning that there is some weak evidence that the methods have the same complexity class.
   \end{itemize}}

### 2014-03-03 jacobian_calc_CMAME.tex
changes denoted more plainly

@@
 algorithmically consistent derivatives computed with \emph{automatic
-differentiation}. For the reader's convenience, from plasticity theory, the algorithmic consistency of
+differentiation}. \note[Michael]{re: II.1}For the reader's convenience, 
+from plasticity theory, the algorithmic consistency of
 the tangent-stiffness matrix refers to consistency with the integration of a
@@
 automatic-differentiation or complex-step to better encourage their use and
-discussion. This open source resource is available at the corresponding author's website: http://idl.utsa.edu/jtfoster/paper-repositories/
+discussion. \note[Michael]{re: II.5}This open source resource is available at the corresponding author's website: http://idl.utsa.edu/jtfoster/paper-repositories/
 or directly at https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive.
@@
 derivative calculation of a function. 
-
+%
+\note[Michael]{re: II.6}
 For the reader's convenience the forward difference equivalent to~\ref{eqn:complexFirstDeriv} is
@@
 
-\note[Michael]{re: I.4}
+\note[Michael]{re: I.4, John all the rest of this section is new}
 It is appropriate to detail the prior work discussing aspects of the
@@
 terms of accuracy and speed. This series would comprise the single core runs. 
+\note[Michael]{re: II.16, specific test parameters were asked for}
 The specific parameters of these tests can be found in \emph{Peridigm} input
@@
         \midrule
-        1000 & 1.99             & .3   & 2.0 & 2 & 10^-8\\
-        2000 & 4.25             & .24  & 2.0 & 2 & 10^-8\\
-        3000 & 7.79             & .21  & 2.0 & 2 & 10^-8\\
-        4000 & 15.4             & .19  & 2.0 & 2 & 10^-8\\
-        6000 & 20.0             & .16  & 2.0 & 2 & 10^-8\\
-        8000 & 31.4             & .15  & 2.0 & 2 & 10^-8\\
-        12000 & 40.1            & .131 & 2.0 & 2 & 10^-8\\
-        16000 & 83.9            & .12  & 2.0 & 2 & 10^-8\\
-        32000 & 165.0           & .095 & 2.0 & 2 & 10^-8\\
-        1000000 & 1670          & .03  & 2.0 & 2 & 10^-8\\
+        1000 & $1.99$             & $.3$   & $2.0$ & $2$ & $\num{10E-8}$ \\
+        2000 & $4.25$             & $.24$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        3000 & $7.79$             & $.21$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        4000 & $15.4$             & $.19$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        6000 & $20.0$             & $.16$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        8000 & $31.4$             & $.15$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        12000 & $40.1$            & $.131$ & $2.0$ & $2$ & $\num{10E-8}$ \\
+        16000 & $83.9$            & $.12$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        32000 & $165.0$           & $.095$ & $2.0$ & $2$ & $\num{10E-8}$ \\
+        1000000 & $1670$          & $.03$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        1000000 & $1670$          & $.03$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        1000000 & $1670$          & $.03$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        1000000 & $1670$          & $.03$  & $2.0$ & $2$ & $\num{10E-8}$ \\
         \bottomrule
@@
         1 & $\num{1.99}$  & 3.5 & 2.7 & 1.7 & 1.6 & $\num{1.92E-10}$ & $\num{1.21E-4}$ & .137 \\
-        1 & $\num{4.25}$  & 6.2& 4.9& 3.2& 2.9 & $\num{2.28E-10}$& $\num{9.94E-4}$ & .148 \\
-        1 & $\num{7.79}$  & 11.2& 8.9& 5.7& 5.2 & $\num{2.38E-10}$& $\num{1.59E-4}$ & .145\\
+        1 & $\num{4.25}$  & 6.2& 4.9& 3.2& 2.9 & $\num{2.28E-10}$ & $\num{9.94E-4}$ & .148 \\
+        1 & $\num{7.79}$  & 11.2& 8.9& 5.7& 5.2 & $\num{2.38E-10}$ & $\num{1.59E-4}$ & .145\\
         1 & $\num{15.4}$  & 26.7& 21.0& 13.4& 12.2 & $\num{2.33E-10}$ & $\num{4.61E-4}$ & .12 \\
@@
 time on average such that compute time were a measure of number of operations,
-this suggests that the methods as in \emph{Peridigm} have the same order
-algorithm complexity when considering computing a TS matrix alone, that is
-computation time is $\mathcal{O}(10^{(10^{-6} \times \log_{10}n)})$. This would
-suggest, to the algorithm design way of thinking, that the methods are equally
-well suited to the task of computing a TS matrix, if it were further assumed
+this suggests that the methods as in \emph{Peridigm} have the same complexity class
+when considering computing a TS matrix alone, that is
+computation time is $\mathcal{O}(10^{(10^{-6} \times \log_{10}n)})$. The definition of 
+complexity class used here is the informal one given in \cite[pp. 1059]{cormen2009introduction}. This would
+suggest that the methods are equally well suited to the task of computing a TS matrix, if it were further assumed
 that the TS matrices produced by each method were identical. This is not the case
@@
 %
+\note[Michael]{re: II.22, relative accuracy figures were asked for, these were taken from the archive data}
 It can be seen in Figs.~\ref{fig:serial_accuracy} and ~\ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods.
@@
         \midrule
-        1 & $1.99$  & 1.28x10^-8 & 2.29x10^-15 & 1.46x10^-5\\
-        1 & $4.25$  & 8.86x10^-8 & 2.25x10^-14 & 1.34x10^-5\\
-        1 & $7.79$  & 1.37x10^-8 & 2.27x10^-14 & 1.25x10^-5\\
-        1 & $15.4$  & 4.33x10^-8 & 2.52x10^-14 & 1.16x10^-5\\
-        1 & $20.0$  & 7.97x10^-8 & 2.18x10^-15 & 1.13x10^-5\\
-        1 & $31.4$  & 1.44x10^-8 & 2.29x10^-15 & 1.09x10^-5\\
-        1 & $40.1$  & 1.36x10^-7 & 2.24x10^-15 & 1.05x10^-5\\
-        1 & $83.9$  & 1.30x10^-7 & 2.44x10^-14 & 9.63x10^-6\\
-        1 & $165.0$  &1.30x10^-7 & 2.44x10^-14 & 9.63x10^-6\\
-        32 & $1670$  &6.00x10^-7 & 2.29x10^-14 & 6.93x10^-6\\
-        64 & $1670$  & 6.01x10^-7 & 2.31x10^-14 & 7.02x10^-6\\
-        96 & $1670$  & 6.03x10^-7 & 2.33x10^-14 & 7.10x10^-6\\
-        128 & $1670$  & 5.97x10^-7 & 2.34x10^-14 & 7.19x10^-6\\
+        1 & $\num{1.99}$ &  $\num{1.28E-8}$ &  $\num{2.29E-15}$ &  $\num{1.46E-5}$ \\
+        1 & $\num{4.25}$ &  $\num{8.86E-8}$ &  $\num{2.25E-14}$ &  $\num{1.34E-5}$ \\
+        1 & $\num{7.79}$ &  $\num{1.37E-8}$ &  $\num{2.27E-14}$ &  $\num{1.25E-5}$ \\
+        1 & $\num{15.4}$ &  $\num{4.33E-8}$ &  $\num{2.52E-14}$ &  $\num{1.16E-5}$ \\
+        1 & $\num{20.0}$ &  $\num{7.97E-8}$ &  $\num{2.18E-15}$ &  $\num{1.13E-5}$ \\
+        1 & $\num{31.4}$ &  $\num{1.44E-8}$ &  $\num{2.29E-15}$ &  $\num{1.09E-5}$ \\
+        1 & $\num{40.1}$ &  $\num{1.36E-7}$ &  $\num{2.24E-15}$ &  $\num{1.05E-5}$ \\
+        1 & $\num{83.9}$ &  $\num{1.30E-7}$ &  $\num{2.44E-14}$ &  $\num{9.63E-6}$ \\
+        1 & $\num{165.0}$ &  $\num{1.30E-7}$ &  $\num{2.44E-14}$ &  $\num{9.63E-6}$ \\
+        32 & $\num{1670}$ &  $\num{6.00E-7}$ &  $\num{2.29E-14}$ &  $\num{6.93E-6}$ \\
+        64 & $\num{1670}$ &  $\num{6.01E-7}$ &  $\num{2.31E-14}$ &  $\num{7.02E-6}$ \\
+        96 & $\num{1670}$ &  $\num{6.03E-7}$ &  $\num{2.33E-14}$ &  $\num{7.10E-6}$ \\
+        128 & $\num{1670}$ &  $\num{5.97E-7}$ &  $\num{2.34E-14}$ &  $\num{7.19E-6}$ \\
         \bottomrule
@@
 
+\note[Michael]{re: I.5, this and other material was added in from the thesis to discuss accuracy}
+\note[Michael]{re: I.5, the entire conclusion section was replaced with the thesis version}
 \section{Conclusions}
@@
 require compiling a version of the AD library being used that were compatible
-with the accelerator or GPU being used. A working definition of byte-copyable
+with the accelerator or GPU being used. \add[Michael]{A working definition of byte-copyable
 is that the collection of data to transfer is made of identically sized pieces.
@@
 installed in \emph{Stampede}, where the alternative mode requires "more
-extensive" effort \cite{intel_byte_copyable}. What this means is that if
+extensive" effort} \cite{intel_byte_copyable}. What this means is that if
 \emph{Peridigm} or any simulation software were to be reformulated to take

### 2014-03-04 response/response.tex
setup stand-in text for next section of response

@@
 
+\subsection*{Typos/editing suggestions}
+\begin{enumerate}
+
+\item Pg. 1: In line 43, remove “the” in “all the of the code”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 1: In line 46, replace “numeric differentiation” by “numerical differentiation”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 2: In line 25, the sentence “The distinction of “accurate” is deﬁned by ...”
+awkward. Please revise it.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 2: In lines 32 and 48, “ﬁnite-difference” should be “forward-difference”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 2: In lines 33, I believe “aide” should be “aid”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 3: In line 12, I would replace “such as required” by “such as those required”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 3: In line 22, replace “with each the methods” by “with each of the methods”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 3: In line 25, replace “implementing complex-step” by “implementing the complex-
+step method”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 3: I would replace the title of Section 1.1, “Differentiation Techniques” by “Dif-
+ferentiation techniques” for consistency, i.e., only capitalize the ﬁrst letter of the ﬁrst
+word. Similarly, for the rest of the section titles.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 4: In line 10, “be be made complex” should be “be made complex”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 4: In Eq. (1), replace comma by period.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 4: In line 22, you have an extra white space in “O (h2 )”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 4: I would check the proper use of hyphens along the manuscript. For example,
+in the title of Section 1.1.2, I would replace “Automatic-differentiation” by “Automatic
+differentiation” (i.e., no hyphen).
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 4: In line 42, replace “based the chain-rule” by “based on the chain rule”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 4: In lines 45-46, “(add, multiply, power, transcendental and the like)” should
+probably be “(addition, multiplication, exponentiation, and the like)” or similar.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 4: In line 47, I would replace “used in the study” by “used in this study”. Similarly,
+in other instances of the manuscript.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 5: In line 39, you may replace “the par tials” by “the par tial derivatives”. Similarly
+in page 26.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 5: I would replace the title of Section 1.2 “Tangent-stiffness” by “Tangent-stiffness
+matrix”. Similarly, in line 50.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 6: In line 14, “quasi-Newton’s method’s” should probably be “quasi-Newton
+methods”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 6: In line 18, “about about a par ticular” should be “about a par ticular”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 6: In line 51, you may replace “argument to F ” by “argument of F ”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 7: In line 11, you may replace “single independent vector component of the
+function” by “single independent vector component of the argument of the function”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 7: In line 24, replace “Where Kij is” by “where Kij is”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 7: In line 40, replace “in literature [19]” by “in the literature [19]”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 8: Remove the period at the end of the titles of Sections 2 and 2.1, for consis-
+tency. Similarly, for the rest of the manuscript. Also, use italic font for “Peridigm” in
+the title of §2.1.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 9: In line 29, replace “Perdigm” by “Peridigm”. Similarly in page 13 (line 44).
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 9: In line 37, replace “be simply be declared” by “be simply declared”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 9: In line 55, I would replace “computational scientists” by “computer scientists”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 10: In line 30, I would replace “The purpose of applying the load” by “Applying
+the load”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 13: I would add a comma at the end of Eq. (8), and replace “Where D is
+distance” by “where D is distance” in line 14.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 13: In line 15, replace “based method, M is” by “based method, and M is”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 13: In line 31, replace “correspond to” by “corresponds to”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 14: In lines 19-21, you have twice the word “nonzero” in “number of nonzero
+tangent-stiffness (TS) non-zero matrix elements”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 14: In lines 22, I would replace “The units for accuracy are derived equation (8)
+and the units used to deﬁne bulk material proper ties, mentioned in Section 2.2.” by
+“The accuracy measure is given by equation (8) and the choice of elastic moduli is
+mentioned in Section 2.2.”
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 15: In line 54, it seems that “yet still fall” should be “yet still falls”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 17: In line 45 you say “h, cannot be too small,”.
+I understand that you mean
+that you can take h as small as possible. However, the way written may have the
+opposite connotation. I would revise it for clarity.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 18: In lines 35-36, replace “the number nonzero” by “the number of nonzero”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 18: In line 39, replace “the the number of nonzero” by “the number of nonzero”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 18: In line 50, I would replace “experimental results,” by “computational experi-
+ments,”. Otherwise, it may give the impression of physical experiments.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 20: In line 13, is the use of the word “order” in “number of iterations taken order
+to measure” correct?
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 20: In lines 26-27, replace “complex-step all of the methods” by “complex-step
+in all of the methods”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 20: In line 50, you may either replace “to implement; It only” by “to implement;
+it only”; or replace the semicolon by a period.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pgs. 21-23: You may capitalize the ﬁrst letters on the journal names in your refer-
+ences [9, 20]
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 22: In your reference [10], you may replace “fr ´echet” by “Fr ´echet”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 23: Update your reference [21] (see the reference [3] in this repor t). Also,
+update your reference [27].
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 24: In line 18, replace the comma by a period.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 24: I would replace the comma by a period in line 28, and replace “using
+L’H ˆopital’s rule” by “Using l’H ˆopital’s rule” in line 31.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 24: I would replace the comma by a period in line 34, and replace “of course,
+as h → 0” by “As h → 0” in line 37.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 24: I would replace the title of Appendix B “AD Example” by “Automatic differ-
+entiation example”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 24: In line 45, you may replace “sin(cos(x))” by “sin(cos(x))”. For that, use \sin
+instead of sin, etc. (similarly in the bottom of the page and in page 25). Also, add
+a period at the end of the expression.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pg. 25: In line 55, you may remove the “·” symbol in “ d
+dx · g(h(k(x)))”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+ \item Pgs. 25 and 26: I would number the table in 2(c) (page 25). Let’s assume the
+number is “2”. Then, in line 13 (page 26), I would replace “chosen for each function
+and held as constant” by “chosen for each function as in Table 2 and held constant”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
+\item Pg. 26: In line 24, add a period at the end of the expression.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      apples
+  \end{itemize}}
+
   \end{enumerate}

### 2014-03-04 response/response.tex
typos

@@
      \item
-      apples
+      Manuscript corrected.
   \end{itemize}}
@@
      \item
-      apples
+      Manuscript corrected. It is made clear that CS is being compared with FD, CD where AD is the standard used for the comparison.
   \end{itemize}}
@@
      \item
-      apples
+    Manuscript corrected.
   \end{itemize}}
@@
      \item
-      apples
+ Manuscript corrected. 
   \end{itemize}}
@@
      \item
-      apples
+ Manuscript corrected. Section and subsection heading changed as well.
   \end{itemize}}
@@
      \item
-      apples
-  \end{itemize}}
-
- \item Pg. 5: In line 39, you may replace “the par tials” by “the par tial derivatives”. Similarly
+ Manuscript corrected. 
+  \end{itemize}}
+
+ \item Pg. 5: In line 39, you may replace “the partials” by “the partial derivatives”. Similarly
 in page 26.
@@
      \item
-      apples
-  \end{itemize}}
-
- \item Pg. 6: In line 18, “about about a par ticular” should be “about a par ticular”.
-
-{\color{red}  
-\begin{itemize}
-     \item
-      apples
+ Manuscript corrected. 
+  \end{itemize}}
+
+ \item Pg. 6: In line 18, “about about a particular” should be “about a particular”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+ Manuscript corrected. 
   \end{itemize}}
@@
  \item Pg. 17: In line 45 you say “h, cannot be too small,”.
-I understand that you mean
-that you can take h as small as possible. However, the way written may have the
+I understand that you mean that you can take h as small as possible. However, the way written may have the
 opposite connotation. I would revise it for clarity.
@@
      \item
-      apples
+ Manuscript corrected.
   \end{itemize}}

### 2014-03-04 jacobian_calc_CMAME.tex
typos

@@
 study described in this paper was to establish the suitability of an
-under-explored method for numerical computation of tangent-stiffness operators,
+underexplored method for numerical computation of tangent-stiffness operators,
 referred to as ``complex-step'', and compare the \note[Michael]{re: I.1}
 \remove[Michael]{new} method with other techniques for numerical derivative
-calculation: automatic differentiation, forward finite-difference, and central
-finite-difference. The complex-step method was \note[Michael]{re: I.1, could
+calculation: automatic differentiation, forward finite difference, and central
+finite difference. The complex-step method was \note[Michael]{re: I.1, could
 suggest claim of novelty} \remove[Michael]{newly} implemented in a massively
@@
 choosing which method best suits the needs of their application code.
-Additionally, motivated by the reproducible research movement, all the of the
+Additionally, motivated by the reproducible research movement, all of the
 code, examples, and workflow to regenerate the data and figures in this paper
@@
 %% keywords here, in the form: keyword \sep keyword
-Newton's method \sep Newton-Raphson \sep numeric differentiation \sep complex-step \sep finite-difference \sep automatic-differentiation \sep Jacobian \sep tangent-stiffness
+Newton's method \sep Newton-Raphson \sep numerical differentiation \sep complex-step \sep finite difference \sep automatic differentiation \sep Jacobian \sep tangent-stiffness
 %% MSC codes here, in the form: \MSC code \sep code
@@
 new method is based on a complex number Taylor series expansion and referred to
-as CTSE or the ``complex-step'' method.  The distinction of ``accurate'' is
-defined by comparison of the \note[Michael]{re: I.1}\remove[Michael]{new}
-method to popular finite differencing techniques used for computing
-tangent-stiffness matrices and with the exact to machine precision
-algorithmically consistent derivatives computed with \emph{automatic
-differentiation}. \note[Michael]{re: II.1}For the reader's convenience, 
-from plasticity theory, the algorithmic consistency of
-the tangent-stiffness matrix refers to consistency with the integration of a
-plasticity model under the Khun-Tucker constraints or similar where the tangent-stiffness matrix is
-computed using the elastoplastic modulii that are determined from the integration \cite{simo1998}.
+as CTSE or the ``complex-step'' method. The practical accuracy of CTSE is
+determined by comparing its measured accuracy\note[Michael]{re:
+I.1}\remove[Michael]{new} relative to the measured accuracy of popular finite
+differencing techniques used for computing tangent-stiffness matrices. For this
+comparison the exact to machine precision derivatives computed with
+\emph{automatic differentiation} serve as the standard. \note[Michael]{re:
+II.1}For the reader's convenience, from plasticity theory, the algorithmic
+consistency of the tangent-stiffness matrix refers to consistency with the
+integration of a plasticity model under the Khun-Tucker constraints or similar
+where the tangent-stiffness matrix is computed using the elastoplastic modulii
+that are determined from the integration \cite{simo1998}.
 
 Comparative data regarding the accuracy and compute cycle timing for the
-complex step, finite-difference, central-difference, and automatic
-differentiation methods are included to aide developers and analysts in
+complex step, forward difference, central difference, and automatic
+differentiation methods are included to aid developers and analysts in
 computational mechanics code design who are faced with implementing a general
-method for tangent stiffness matrix calculation.  A comparison of the methods was achieved through instrumentation of a
+method for tangent-stiffness matrix calculation.  A comparison of the methods was achieved through instrumentation of a
 massively parallel computational mechanics code.
 
-The scope of the application component of the study was limited to a single
+The scope of the application component of this study was limited to a single
 material model, an elastic peridynamic solid, implemented in the computational
@@
 matrices for solving nonlinear quasi-static problems, prior inclusion of
-finite-difference, central-difference, and automatic-differentiation
+forward difference, central difference, and automatic differentiation
 methods needed for comparison to the new complex-step method
@@
 code to reduce development time and increase reliability. The modularity of
-\emph{Peridigm} allowed the adding of new features, such as required by the
+\emph{Peridigm} allowed the adding of new features, such as those required by the
 study.
@@
 description and justification of the \note[Michael]{re: I.1}\remove[Michael]{new}complex-step method for calculating
-tangent-stiffness matrices, in sub-subsection~\ref{subsubsec:Impl} a description of implementing complex-step in
+tangent-stiffness matrices, in sub-subsection~\ref{subsubsec:Impl} a description of implementing the complex-step method in
 the \emph{Peridigm} software, in section~\ref{JGAM} a description of the quantities-of-interest
@@
 manner than in the \emph{Peridigm} code how one would use
-automatic-differentiation or complex-step to better encourage their use and
+automatic differentiation or complex-step to better encourage their use and
 discussion. \note[Michael]{re: II.5}This open source resource is available at the corresponding author's website: http://idl.utsa.edu/jtfoster/paper-repositories/
@@
 
-\subsection{Differentiation Techniques}
+\subsection{Differentiation techniques}
 
@@
 underlying the tangent-stiffness matrix calculation methods compared in the
-study. Since the derivation of first-order finite-difference techniques from
+study. Since the derivation of first-order finite difference techniques from
 Taylor series expansion is considered well known, it is not described here.
@@
 
-\subsubsection{The ``complex-step'' method.}  \label{sec:CSmethod} It is
+\subsubsection{The ``complex-step'' method}  
+\label{sec:CSmethod}
+
+It is
 possible to approximate derivatives quite accurately with a technique based on
@@
 rediscovered for use in engineering analysis.  The basic idea is that a model parameter can
-be be made complex and expanded in a Taylor series about a small perturbation
+be made complex and expanded in a Taylor series about a small perturbation
 along the imaginary axis by some arbitrary value $h$ as follows
@@
 h}{1!}  \notag \\ &+ \frac{\partial^2 f(x)}{\partial x^2} \frac{(i h)^2}{2!} +
-\frac{\partial^3 f(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots,
+\frac{\partial^3 f(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots.
 \label{eqn:complexTaylor} \end{align}
@@
 Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor})
-and solving for the first derivative and ignoring terms of $\mathcal{O}\left (
+and solving for the first derivative and ignoring terms of $\mathcal{O}\left(
 h^2 \right)$ yields an estimate
@@
 \begin{equation} \frac{\partial f( x )}{\partial x} = \frac{\mbox{Im}
-\left( f (x + i h) \right)}{h} +\mathcal{O}\left ( h^2 \right).  \label{eqn:complexFirstDeriv} \end{equation}
+\left( f (x + i h) \right)}{h} +\mathcal{O}\left( h^2 \right).  \label{eqn:complexFirstDeriv} \end{equation}
 %
@@
 practical (even to machine precision without the dangers of roundoff error as
-in finite-differences) to yield an accurate estimate of this derivative.  The
+in finite differences) to yield an accurate estimate of this derivative.  The
 only disadvantage of this technique is the necessity of requiring the functions
@@
 %
-\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x)}{h} +\mathcal{O}\left ( h \right). \label{eqn:forwardFirstDeriv} \end{equation},
+\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x)}{h} +\mathcal{O}\left( h \right). \label{eqn:forwardFirstDeriv} \end{equation},
 %
@@
 %
-\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x - h)}{2h} +\mathcal{O}\left ( h^2 \right). \label{eqn:centeredFirstDeriv} \end{equation} \cite[Chap. 4.1.3]{chapra2010}.
+\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x - h)}{2h} +\mathcal{O}\left( h^2 \right). \label{eqn:centeredFirstDeriv} \end{equation} \cite[Chap. 4.1.3]{chapra2010}.
 %
@@
 
-\subsubsection{Automatic-differentiation} 
+\subsubsection{Automatic differentiation} 
 \label{ADsubsection}
 
-Automatic-differentiation (AD) is a computerized method for computing exact
-derivatives based the chain-rule from calculus. AD takes advantage of the fact
+Automatic differentiation (AD) is a computerized method for computing exact
+derivatives based on the chain-rule from calculus. AD takes advantage of the fact
 that any mathematical function executed on a computer, no matter how
-complicated, is a ``composition of simple operations'' (add, multiply, power,
-transcendental and the like) each having known analytical derivatives
+complicated, is a ``composition of simple operations'' (addition, multiplication, exponentiation and the like) each having known analytical derivatives
 \cite{ref-sacado-presentation}. For reference, the AD implementation used in
-the study is part of the \emph{Sacado} package from the \emph{Trilinos}
+this study is part of the \emph{Sacado} package from the \emph{Trilinos}
 agile-component libraries developed at Sandia National Laboratories
@@
 scales is an active concern. In the literature, the AD scheme described here is
-called \emph{forward automatic-differentiation}. Only forward AD will be
+called \emph{forward automatic differentiation}. Only forward AD will be
 covered here since it is the implementation used in \emph{Sacado} and therefore
@@
 about AD are that no approximation of derivatives is being made because the
-analytical forms of the partials of the elementary math functions are defined
+analytical forms of the partial derivatives of the elementary math functions are defined
 alongside them. The accuracy of AD is then limited by the precision of the AD
@@
 \label{subsec:TS}
+
 In solid and structural computational mechanics the tangent-stiffness matrix is
@@
 problems, tangent-stiffness matrices generally arise in the context of
-non-linear analysis where Newton's method or quasi-Newton's method's are used
+non-linear analysis where Newton's method or quasi-Newton methods are used
 in the minimization of a given residual function.  In this context, the
 tangent-stiffness matrix can be thought of as the linearization of the system
-about about a particular configuration.  It has been shown
+about a particular configuration.  It has been shown
 \cite{hughes1978consistent,hughes1978unconditionally} that in order to preserve
@@
 where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function,
-$X_j$ is the $j^{\mbox{th}}$ component of the vector argument to $F$, and a
+$X_j$ is the $j^{\mbox{th}}$ component of the vector argument of $F$, and a
 particular value of the vector, $\vec{X}_0$ is chosen as the linearization
@@
 elements of a tangent-stiffness matrix can be estimated using any of the
-complex-step, AD, or common finite-difference techniques for functions $F:R^1
+complex-step, AD, or common finite difference techniques for functions $F:R^1
 \rightarrow R^1$, since taking partial derivatives entails holding all but a
-single independent vector component of the function constant, and each element
+single independent vector component of the argument of the function constant, and each element
 of the function can be evaluated independently of the others. \\
 
-\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation.}
-
-The \emph{forward-difference} (FD) formula for calculating a tangent-stiffness matrix in the setting of a solid structural mechanics problem is 
+\subsubsection{Differencing formulas for tangent-stiffness matrix evaluation}
+
+The \emph{forward difference} (FD) formula for calculating a tangent-stiffness matrix in the setting of a solid structural mechanics problem is 
 %
@@
 
-Where $K_{ij}$ is the indicial notation representation of an element of the
+where $K_{ij}$ is the indicial notation representation of an element of the
 tangent-stiffness matrix at row $i$, column $j$. The first of the two terms in
@@
 
-The \emph{central-difference} (CD) formula for calculating a tangent-stiffness matrix is
+The \emph{central difference} (CD) formula for calculating a tangent-stiffness matrix is
 %
@@
 %
-Together, the FD and CD methods, are sometimes termed \emph{finite-difference probing} techniques for tangent-stiffness calculation in literature \cite{ref-Adaggio}.
+Together, the FD and CD methods, are sometimes termed \emph{finite difference probing} techniques for tangent-stiffness matrix calculation in the literature \cite{ref-Adaggio}.
 From the well-known 1-dimensional formulas for FD and CD, it is understood that the accuracy of these expressions is dependent theoretically upon selecting a small step-size $h$, however experiments shown in \cite{squire1998using} demonstrate that too small an $h$ can lead to inaccuracy from ``subtractive-cancellation''.
@@
 
-When using automatic-differentiation to calculate a tangent-stiffness matrix,
+When using automatic differentiation to calculate a tangent-stiffness matrix,
 one only needs to follow the definition of the tangent-stiffness matrix and
@@
 \label{sec:WorkingExamples}
+
 In order to illustrate the complex-step technique in a simpler form than it
@@
 and final displacement, D, after the final load-step. Additionally
-finite-difference step-size was varied in order to measure the impact on each
+finite difference step-size was varied in order to measure the impact on each
 of the Jacobian calculation methods. Naturally, varying step size does not
@@
 
-\section{Description of analysis tools and approach.} 
-%
-\subsection{\emph{Peridigm} and peridynamics.}
+\section{Description of analysis tools and approach} 
+%
+\subsection{\emph{Peridigm} and peridynamics}
 \label{subsec:PaP}
 %
-Comparisons of the different tangent-stiffness calculation techniques were
+Comparisons of the different tangent-stiffness matrix calculation techniques were
 carried out through code-development and instrumentation of the
@@
 \cite{trilinos}. It's important to note that there is no specialization of the
-tangent-stiffness calculation methods described previously to this particular
+tangent-stiffness matrix calculation methods described previously to this particular
 code or more generally to a computational peridynamics approach and the
@@
 analysis tools as well.  \emph{Peridigm} was chosen primarily because the FD,
-CD, and AD methods for tangent-stiffness calculation where already implemented
+CD, and AD methods for tangent-stiffness matrix calculation where already implemented
 in the code, leaving only the CS method for development. Also, because
@@
 higher bandwidth (i.e. less sparsity) than traditional finite-element
-tangent-stiffness computations and the time required to construct the
-tangent-stiffness is a majority of the total computation time in any
+tangent-stiffness matrix computations and the time required to construct the
+tangent-stiffness matrix is a majority of the total computation time in any
 quasi-static or implicit dynamics analysis; therefore, the \emph{Peridigm}
 development team has an interest in profiling the performance of the
-tangent-stiffness computation methods.
+tangent-stiffness matrix computation methods.
 
@@
 \note[Michael]{re: I.2. John, this was to be the skeleton for the new sections on peridynamics.}
-\subsubsection{Vector states for state based peridynamic theory.}
+\subsubsection{Vector states for state based peridynamic theory}
 Because the vector state, which is fair to say an uncommon notation, is used in the literature discussing state based peridynamic theory
@@
 
-\subsubsection{The state based peridynamic equation of motion.}
+\subsubsection{The state based peridynamic equation of motion}
 
@@
 
-\subsubsection{The linear peridynamic solid.}
+\subsubsection{The linear peridynamic solid}
 
 \note[Michael]{re: I.2} \add[Michael]{
-\subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simulation.}
-The tangent stiffness matrix is the Frechet derivative of the residual wrt the
+\subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simulation}
+The tangent-stiffness matrix is the Frechet derivative of the residual wrt the
 deformation state.  For the purposes of this work, the residual represents the unbalanced forces in the
@@
 
-For the finite-difference type methods in \emph{Peridigm}, the perturbation
+For the finite difference type methods in \emph{Peridigm}, the perturbation
 method is employed by probing the deformed configuration one degree of freedom
@@
 %
-The CS method was implemented in \emph{Perdigm} in a manner that follows
+The CS method was implemented in \emph{Peridigm} in a manner that follows
 closely the implementation of the FD technique.  However, because of
@@
 \change[Michael]{STL}{Standard Template Library} vectors, are hard coded to be
-of type {\tt double} and can not be simply be declared as having a complex data
+of type {\tt double} and can not be simply declared as having a complex data
 type. \note[Michael]{re: I.3}\add[Michael]{For the reader's convenience,
@@
 
-\emph{Peridigm} was primarily written by computational scientists; therefore,
+\emph{Peridigm} was primarily written by computer scientists; therefore,
 many advanced techniques in C++ such as multiple virtual inheritance, pointer
@@
 
-\subsection{The Comparative Study} 
+\subsection{Comparative study} 
 \label{tcs}
@@
 tension along the axis parallel to the long dimension of the block. The test
-problems were set up in \emph{Perdigm} as quasistatic equilibrium problems,
+problems were set up in \emph{Peridigm} as quasistatic equilibrium problems,
 where the displacement boundary conditions used to apply tension were applied
 gradually in ``load steps'', and an equilibrium solution for each load step was
-achieved before applying the next load step. The purpose of applying the load
+achieved before applying the next load step. Applying the load
 in gradual steps is a widely used technique to ensure the Newton's method is
@@
 for shear modulus. The peridynamic horizon, a length scale that effectively
-sets the bandwidth of tangent-stiffness in the computation, used in the test
+sets the bandwidth of tangent-stiffness matrix in the computation, used in the test
 problems was always set to three times the nominal discretization size (i.e.
@@
 
-As stated in the Introduction, a goal of the study was to compare CS, FD, and
+As stated in the Introduction, a goal of this study was to compare CS, FD, and
 CD on the basis of accuracy. AD was omitted from the comparison because it
@@
 appropriate analytical forms for the tangent-stiffness matrix associated with
-the system solved in the study. The assumption that AD is accurate enough to
+the system solved in this study. The assumption that AD is accurate enough to
 serve as a standard is supported by AD's implementation as a computerized chain
@@
 (e.g. lower accuracy predictions of the algorithmically correct
-tangent-stiffness could cause a loss of quadratic convergence).  Different
+tangent-stiffness matrix could cause a loss of quadratic convergence).  Different
 guesses from iterations started with different previous guesses could not be
-data for a valid comparison of tangent-stiffness accuracy between methods.
+data for a valid comparison of tangent-stiffness matrix accuracy between methods.
 Additionally, having each of the methods physically operate in the same
@@
 speculatively involving dynamic file management on files which for one of the
-components of the study would be on the order of a terabyte in size. However,
+components of this study would be on the order of a terabyte in size. However,
 the price of running the methods together in the same process and avoiding the
@@
 speed. Speed is defined as the total compute-time of one iteration of a Newton
-step which includes the tangent-stiffness calculation and parallel assembly.
+step which includes the tangent-stiffness matrix calculation and parallel assembly.
 The speed of iteration was measured because it could be so done at the same
@@
 evaluation order and computer system load due to system processes not
-associated with the study. While not part of the study, the example programs
+associated with this study. While not part of this study, the example programs
 available in the paper repository allow the reader to make a comparison of the
@@
 %
-\begin{equation} D = \sqrt{\sum_i \sum_j(K^{AD}_{ij} - K^{M}_{ij})^2}
+\begin{equation} D = \sqrt{\sum_i \sum_j(K^{AD}_{ij} - K^{M}_{ij})^2},
 \label{eqn:accuracy} \end{equation}
 %
-Where $D$ is distance, $K^{AD}$ is the tangent-stiffness matrix produced by the
-AD based method, $M$ is replaced with either CS, FD, or CD depending on the
+where $D$ is distance, $K^{AD}$ is the tangent-stiffness matrix produced by the
+AD based method, and $M$ is replaced with either CS, FD, or CD depending on the
 method being compared.  The metric used to compare the speed of the different
 methods was the time in seconds required to calculate the tangent-stiffness
-matrix. A final basis of comparison used in the study called
+matrix. A final basis of comparison used in this study called
 \emph{computational efficiency} is based on specific calculation time per
@@
 %
-Discretization for the test problems run in the study was done using the
+Discretization for the test problems run in this study was done using the
 software package \emph{Cubit} developed at Sandia National Laboratories
@@
 \emph{Peridigm}. The number of computational nodes in \emph{Peridigm}
-correspond to the number of finite elements in the discretization, not the
+corresponds to the number of finite elements in the discretization, not the
 number of finite element nodes.  An example journal script can be found on the
@@
 
-\subsubsection{Data Reduction}
-
-The output produced by \emph{Perdigm} including the additional accuracy and
+\subsubsection{Data reduction}
+
+The output produced by \emph{Peridigm} including the additional accuracy and
 speed measurements was redirected from the console to text files by the
@@
 
-\section{Results and Discussion}
+\section{Results and discussion}
 \label{sec:Res}
 %
-Table~\ref{tab:results} shows the compute core count, number of nonzero tangent-stiffness (TS) non-zero matrix elements, load step average calculation time and load step average accuracy.
-The units for accuracy are derived Eqn.~(\ref{eqn:accuracy}) and the units used to define bulk material properties, mentioned in Section~\ref{tcs}.
+Table~\ref{tab:results} shows the compute core count, number of nonzero tangent-stiffness (TS) matrix elements, load step average calculation time and load step average accuracy.
+The accuracy measure is given by Eqn.~(\ref{eqn:accuracy}) and the choice of elastic moduli mentioned in Section~\ref{tcs}.
 %
@@
 
-\subsection{Speed Data} Speed measurements were taken according to the methods
+\subsection{Speed data}
+
+Speed measurements were taken according to the methods
 mentioned in Section~\ref{tcs}. The averaged results of these speed
@@
 functional evaluations themselves only to be processed at the end of the
-tangent-stiffness calculation when the derivatives are requested. This
+tangent-stiffness matrix calculation when the derivatives are requested. This
 represents additional load on the hardware memory subsystem, such that may
-exceed that of FD yet still fall under a supposed memory bandwidth threshold
+exceed that of FD yet still falls under a supposed memory bandwidth threshold
 during serial operation. Cores within a multicore processor use the same shared
@@
 
-\subsection{Accuracy Data}
+\subsection{Accuracy data}
 
@@
 a numerical example in \cite[Table 1]{squire1998using} demonstrates, $h$,
-cannot be too small, therefore it can be set to machine epsilon and the
+has no lower limit for effective values, therefore it can be set to machine epsilon and the
 analysis performed without the issues associated with round-off error. The
@@
 
-\subsection{Efficiency Data}
-Efficiency, calculated as the average number of seconds taken per nonzero Jacobian matrix element, was plotted as a function of the number nonzero TS matrix elements.
+\subsection{Efficiency data}
+
+Efficiency, calculated as the average number of seconds taken per nonzero Jacobian matrix element, was plotted as a function of the number of nonzero TS matrix elements.
 The single core results are shown in Fig.~\ref{fig:serial_efficiency}.
@@
 force densities such as for small displacements or very lax materials. In this case it could be recommended
-that forward-difference be avoided.
+that forward difference be avoided.
 
@@
 inaccurate Jacobian is found along with a reason for using it. While it is possible
-to incorrectly apply finite-differencing and CS by choosing an inappropriate
+to incorrectly apply finite differencing and CS by choosing an inappropriate
 step-size, there is no reason that this should be done. The experiments show
@@
 nonlocal interaction between material points essential to peridynamic models,
-like the linear model used in the study.
+like the linear model used in this study.
 
@@
 stock version of \emph{Peridigm} this is not reflective of any limitation of
-these methods. In a side experiment, the finite-difference code in
+these methods. In a side experiment, the finite difference code in
 \emph{Peridigm} was re-written to take advantage of the benefits of storing the
@@
 performance increase of around a quarter was available for the
-finite-difference methods and that FD could eclipse AD. What allowed the performance modifications to
-enhance the finite-difference methods was their lack of sophistication
+finite difference methods and that FD could eclipse AD. What allowed the performance modifications to
+enhance the finite difference methods was their lack of sophistication
 compared to AD in that their explicitly laid out and
@@
 \subsection{Remarks}
-Based on the experimental results, it appears that for users of \emph{Peridigm}
+Based on the computational experiments, it appears that for users of \emph{Peridigm}
 there is a case to be made for using AD for reasons of speed. However, certain
@@
 be noted that from a code development perspective the CS method may be the
-easiest of the methods to implement; It only requires one functional
+easiest of the methods to implement; it only requires one functional
 evaluation, there is no heuristic choice of probe distance, and there is no

### 2014-03-05 response/response.tex
all revisions addressed except peridynamics section

@@
 
- \item Pg. 22: In your reference [10], you may replace “fr ´echet” by “Fr ´echet”.
-
-{\color{red}  
-\begin{itemize}
-     \item
-      apples
-  \end{itemize}}
-
- \item Pg. 23: Update your reference [21] (see the reference [3] in this repor t). Also,
+ \item Pg. 22: In your reference [10], you may replace “fr{\'e}chet” by “Fr{\'e}chet”.
+
+{\color{red}  
+\begin{itemize}
+     \item
+ Manuscript corrected.
+  \end{itemize}}
+
+ \item Pg. 23: Update your reference [21] (see the reference [3] in this report). Also,
 update your reference [27].
@@
  \item Pg. 24: I would replace the comma by a period in line 28, and replace “using
-L’H ˆopital’s rule” by “Using l’H ˆopital’s rule” in line 31.
-
-{\color{red}  
-\begin{itemize}
-     \item
-      apples
+L’H{\ˆo}pital’s rule” by “Using l’H{\ˆo}pital’s rule” in line 31.
+
+{\color{red}  
+\begin{itemize}
+     \item
+ Manuscript corrected.
   \end{itemize}}

### 2014-03-05 jacobian_calc_CMAME.tex
all revisions addressed except peridynamics section

@@
 A work by A. H. Al-Mohy and N. J. Higham investigates the use of complex step
-in computing Frechet derivatives of matrix to matrix functions
+in computing Fr\'{e}chet derivatives of matrix to matrix functions
 \cite{al2010complex}. To compute the Jacobian of the residual in the
 peridynamics simulation as discussed in the instant work is analogous is
-computing a Frechet derivative. Where the application of the CS derivatives in
+computing a Fr\'{e}chet derivative. Where the application of the CS derivatives in
 the prior work is limited to a particular class of abstract functions to
@@
 instant work applies CS to solving a solid mechanics problem. The instant work is not
-about computing Frechet derivatives with CS but moreover using them in a
+about computing Fr\'{e}chet derivatives with CS but moreover using them in a
 numerical method to solve a set of equations describing a physical system.
@@
     \label{eqn:TrussForce}
-    f(D) = .5AE(cos^{2}\phi)(\frac{D}{L_{0}})^{2}[\frac{D}{L_{0}}cos^{2}\phi - 3sin\phi] + k_{s}D + (AE\frac{D}{L_{0}})sin^{2}\phi,
+    f(D) = .5AE(\cos^{2}\phi)(\frac{D}{L_{0}})^{2}[\frac{D}{L_{0}}\cos^{2}\phi - 3\sin\phi] + k_{s}D + (AE\frac{D}{L_{0}})\sin^{2}\phi,
 \end{equation}
@@
     \label{eqn:TrussStiffness}
-    S_{T}(D) = 1.5AE(cos^{2}\phi)[\frac{D}{L_{0}}cos^{2}\phi - 2sin\phi](\frac{D}{L_{0}^{2}}) + k_{s} + \frac{AEsin^{2}\phi}{L_{0}}.
+    S_{T}(D) = 1.5AE(\cos^{2}\phi)[\frac{D}{L_{0}}\cos^{2}\phi - 2sin\phi](\frac{D}{L_{0}^{2}}) + k_{s} + \frac{AE\sin^{2}\phi}{L_{0}}.
 \end{equation}
@@
 \subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simulation}
-The tangent-stiffness matrix is the Frechet derivative of the residual wrt the
+The tangent-stiffness matrix is the Fr\'{e}chet derivative of the residual wrt the
 deformation state.  For the purposes of this work, the residual represents the unbalanced forces in the
@@
 density is computed in the peridynamic momentum equation by computing the
-volume integral over a horizon of the Frechet derivative of the 
+volume integral over a horizon of the Fr\'{e}chet derivative of the 
 strain energy density state wrt to the deformed configuration inner product with the deformation state, per node,
@@
 \[
-f(x + i h) = \cos(x) \cosh(h) - i \sin(x) \sinh(h),
+f(x + i h) = \cos(x) \cosh(h) - i \sin(x) \sinh(h).
 \]
@@
 \]
-using L'H\^opital's rule
+using l'H\^opital's rule
 \[
-\frac{\partial f(x)}{\partial x} = \lim_{h \to 0} - \frac{\sin(x) \cosh(h)}{1},
+\frac{\partial f(x)}{\partial x} = \lim_{h \to 0} - \frac{\sin(x) \cosh(h)}{1}.
 \]
-of course, as $h \to 0$ then $\cosh(h) \to 1$ thereby recovering the exact derivative
+As $h \to 0$ then $\cosh(h) \to 1$ thereby recovering the exact derivative
 \[
@@
 
-\section{AD Example}
+\section{Automatic differentiation example}
 \label{sec:appendixB}
@@
 %while encouraging them to rely on their understanding of the chain rule
-\item Take an example composition function $f(x) = (sin(cos(x)))^2$\\
+\item Take an example composition function $f(x) = (\sin(\cos(x)))^2$\\
 
@@
 \item $g, h, k \in H: S \rightarrow S$ 
-\item $\forall s \in S : g(s) = {s}^2$, $h(s) = sin(s)$, $k(s) = cos(s)$ 
+\item $\forall s \in S : g(s) = {s}^2$, $h(s) = \sin(s)$, $k(s) = \cos(s)$ 
 \item
@@
 %General definitions of functions are specialized by arguments. 
-Given that it is possible to describe $f(x) = (sin(cos(x)))^2$ in terms the computer understands by inputting
+Given that it is possible to describe $f(x) = (\sin(\cos(x)))^2$ in terms the computer understands by inputting
 $f(x)$ such that the computer stores an equivalent statement $f(x) = u(a, b, s) \mid_{arguments}$,
@@
 It follows from \ref{composition} that we can evaluate $\frac{d}{d x}f(x) \mid_{x_0}$ with the chain rule: \\ \\
-$\frac{d}{d x}f(x) = \frac{d}{d x} \cdot g(h(k(x))) \mid_{x_0}$ \\
+$\frac{d}{d x}f(x) = \frac{d}{d x} g(h(k(x))) \mid_{x_0}$ \\
 $\frac{d}{d x}f(x) =  \frac{d{g}}{d{h}} \cdot \frac{d{h}}{d{k}}
@@
 specializing the computer's general forms of $u, v, w$ 
-according to ~\ref{abcvalues}, with parameters $a, b$ chosen for each function and held as constant, and
+according to~\ref{abcvalues}, with parameters $a, b$ chosen for each function as in~\ref{abcvalues} and held constant, and
 evaluating, such that the total derivative of our original function w.r.t $x$ where $x = x_0$ is approximated
@@
 \ref{abcvalues} with a particular value of $s = x_0$, in equation format: \\ \\
-$\frac{\partial}{\partial x}f(x) \mid _{x_0} = \frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot \frac{ds}{dx}$\\ \\
+$\frac{\partial}{\partial x}f(x) \mid _{x_0} = \frac{\partial{u}}{\partial{v}} \mid _{1, 2, v \mid _{1, 1, w \mid _{ 1, 1, x_0}}} \cdot \frac{\partial{v}}{\partial{w}} \mid _{1,1, w \mid _{1, 1, x_0}} \cdot \frac{\partial{w}}{\partial{s}} \mid _{1, 1, x_0} \cdot \frac{ds}{dx}.$\\ \\
 %Repeating above in tabular format: \\ \\

### 2014-03-05 jacobian_calc_CMAME.tex
spellcheck

@@
 integration of a flow rule for plasticity modeling.  If the continuum
-tangent-moduli are used in place of the \emph{consistent} or \emph{algorithmic
+tangent-modulii are used in place of the \emph{consistent} or \emph{algorithmic
 tangent moduli} in the solution of a non-linear solid mechanics problem,
@@
 
-\subsubsection{Implementing the complex-step method in Peridigm} 
+\subsubsection{Implementing the complex-step method in \emph{Peridigm}} 
 \label{subsubsec:Impl}
@@
 tension along the axis parallel to the long dimension of the block. The test
-problems were set up in \emph{Peridigm} as quasistatic equilibrium problems,
+problems were set up in \emph{Peridigm} as quasi-static equilibrium problems,
 where the displacement boundary conditions used to apply tension were applied
@@
 have close ID numbers. The abrupt color changes in the center and at the ends
-of the body show that numbering is not contiguous and is assigned with respect to four regioins. The purpose of designating
+of the body show that numbering is not contiguous and is assigned with respect to four regions. The purpose of designating
 these regions, for this study, was to apply displacement boundary
@@
 Table~\ref{tab:results} shows the compute core count, number of nonzero tangent-stiffness (TS) matrix elements, load step average calculation time and load step average accuracy.
-The accuracy measure is given by Eqn.~(\ref{eqn:accuracy}) and the choice of elastic moduli mentioned in Section~\ref{tcs}.
+The accuracy measure is given by Eqn.~(\ref{eqn:accuracy}) and the choice of elastic modulii mentioned in Section~\ref{tcs}.
 %
@@
 In fact, a secant equation based method is appropriate for Newton's method with algorithmic
-tangent moduli since the Jacobian ought only depend on known converged
+tangent modulii since the Jacobian ought only depend on known converged
 values such that the Newton iterate could not artificially drive the plasticity

### 2014-03-28 jacobian_calc_CMAME.tex
Fixed a few build errors

@@
 \usepackage{array}
+\usepackage{hyperref}
 %% The amssymb package provides various useful mathematical symbols
@@
 automatic differentiation or complex-step to better encourage their use and
-discussion. \note[Michael]{re: II.5}This open source resource is available at the corresponding author's website: http://idl.utsa.edu/jtfoster/paper-repositories/
-or directly at https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive.
+discussion.
+\note[Michael]{re: II.5}This open source resource is available at the corresponding author's website or at \href{https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive}{https://github.com/utsa-idl/Complex\_Step\_Jacobian\_Project\_Archive}.
 
@@
 radius, centered at the original coordinates of a peridynamic node,
-$\boldsymbol{x}$, and \boldsymbol{\underline{T}} is the "force vector state
+$\boldsymbol{x}$, and $\boldsymbol{\underline{T}}$ is the "force vector state
 field". The force vector state field maps vectors in the original coordinate system 
@@
 neighboring peridynamic nodes that fall within the radius of the horizon. $\boldsymbol{b}$ denotes a body force density. Equation~\ref{eqn:PDMotion} may be simplified based on the assumption that
-\boldsymbol{\underline{T}} is Riemann integrable and that integral is uniformly
+$\boldsymbol{\underline{T}}$ is Riemann integrable and that integral is uniformly
 convergent \cite{silling:psa} as,

### 2014-03-31 response/response.tex
fixed issue with counter too large error

@@
   \end{itemize}}
+
   \end{enumerate}
@@
      computer system. MIC accelerator cards permit only the transfer of
-     byte-copiable data from the host computer system for use with the '#pragma
+     byte-copiable data from the host computer system for use with the 'pragma
      offload' mode, which is to say most code written for the MIC. This is even
@@
 
+\end{enumerate}
+
 \subsection*{Typos/editing suggestions}
@@
 
- \item Pg. 24: In line 45, you may replace “sin(cos(x))” by “sin(cos(x))”. For that, use \sin
+ \item Pg. 24: In line 45, you may replace “sin(cos(x))” by “sin(cos(x))”. For that, use $\sin$
 instead of sin, etc. (similarly in the bottom of the page and in page 25). Also, add

### 2014-04-08 response/response.tex
fixed accent marks
Update Lit. Rev.

@@
 %%%% First Reviewer %%%%
-\section{Re: Information to the authors}
+\section*{Re: Information to the authors}
 %
-\subsection*{Larger issues}
+\subsection*{I. Larger issues}
 %
@@
     \item
-         Some parts of the manuscript suggest that the authors introduced a new method,
-         i.e., the complex-step (CS) method. For example, in the introduction, they say
-         “A goal of this study was to develop and evaluate a new, accurate, and practical
-         method ...” Another example appears in page 3, where the authors say “The aim
-         of this paper is not only to introduce the new complex-step method in the context
-         of evaluating tangent-stiffness matrices ..”. However, the CS method has already
-         appeared in many publications since 1998 [1]. Please revise the way you refer to
-         the CS method along the manuscript
-
-{\color{red}
-  \begin{itemize}
-      \item
-        The way the CS method is referred to has been revised throughout,
-        specifically: 
-    \item In the abstract, "... the new method ..." has been
-        replaced by "... the method ...", "... newly implemented ..." has been
-        replaced by "... implemented ..."; 
-    \item in the introduction, paragraph two,
-        "... a new, accurate ..." has been replaced by "... a relatively
-        underexplored, accurate ...", in paragraph two and four "... the new
-        method ..." has been replaced by "... the method ...", in paragraph
-        five, "... introduce the new complex-step method in the context of ..."
-        has been replaced by "... study the complex-step method of ...", "...
-        the new complex-step ..." has been replaced by "... the complex-step
-        ...".
-    \item 
-        It is the authors' opinion that the above changes correct the suggestion
-        that the authors introduce a new method.
-
-  \end{itemize}}
-
+         Some parts of the manuscript suggest that the authors introduced a new method, i.e., the complex-step (CS) method.
+For example, in the introduction, they say ``A goal of this study was to develop and evaluate a new, accurate, and practical method ...'' Another example appears in page 3, where the authors say ``The aim of this paper is not only to introduce the new complex-step method in the context of evaluating tangent-stiffness matrices ...'' However, the CS method has already appeared in many publications since 1998 [1].
+Please revise the way you refer to the CS method along the manuscript
+
+{\color{red}
+  \begin{itemize}
+      \item In the cases of ambiguity pointed out by the reviewer, the wording has been changed such that it should be clear we are only evaluating an underutilized method, not introducing it.
+  \end{itemize}
+}
     \item
@@
 
-  \subsection*{Smaller issues}
+  \subsection*{II. Smaller issues}
   \begin{enumerate}
@@
   \item
-	Pg. 2: In line 14, explain what does the term “algorithmically consistent” mean.
-
-{\color{red}
-  \begin{itemize}
-     \item
-
-      A brief definition for the convenience of the reader was added in the
-      manuscript. But, algorithmically consistent is a fairly common term in
-      plasticity/mechanics. This same meaning is used for 'consistent' refering
-      to 'tangent matrices' in \cite{perez2012numerical, perez2000numerical}. It means
-      consistent with integrating the plasticity model under the Khun-Tucker
-      constraints. Integrating the plasticity model means, given a strain
-      increment, determining a stress that does not exceed the yield stress,
-      and by doing so computing a linear slope, called a consistent tangent
-      modulus, that relates the strain increment to a correct stress increment.
-      An alogorithmically consistent tangent stiffness matrix is then one that
-      uses the consistent tangent modulii, making it a more accurate
-      approximation of the stiffness of the system.
-
+	Pg. 2: In line 14, explain what does the term ``algorithmically consistent'' mean.
+
+{\color{red}
+  \begin{itemize}
+      \item Footnote 1 has been added to manuscript to address this.
   \end{itemize}}
@@
      \begin{verbatim}
-     "http://en.wikipedia.org/wiki/IEEE_floating_point" 
+    ``http://en.wikipedia.org/wiki/IEEE_floating_point''
      \end{verbatim}
@@
       \begin{verbatim}
-      "https://sourceware.org/git/?p=glibc.git;a=blob;f=sysdeps/x86_64/fpu/e_expl.S"
+     ``https://sourceware.org/git/?p=glibc.git;a=blob;f=sysdeps/x86_64/fpu/e_expl.S"
       \end{verbatim}
@@
  \item Pg. 24: I would replace the comma by a period in line 28, and replace “using
-     L’H{\ˆo}pital’s rule” by “Using l’H{\ˆo}pital’s rule” in line 31.
+     L’H\^opital’s rule” by “Using l’H\^opital’s rule” in line 31.
 

### 2014-04-08 jacobian_calc_CMAME.tex
fixed some errors in a table
Update Lit. Rev.

@@
 \addeditor{JTF}
-\addeditor{Michael}
+\addeditor{MDB}
 %% The amsthm package provides extended theorem environments
@@
 
-In order to maintain the quadratic convergence properties of Newton's method in
-quasi-static nonlinear analysis of solid structures it is crucial to obtain
-accurate, algorithmically consistent tangent-stiffness matrices. A goal of the
-study described in this paper was to establish the suitability of an
-underexplored method for numerical computation of tangent-stiffness operators,
-referred to as ``complex-step'', and compare the \note[Michael]{re: I.1}
-\remove[Michael]{new} method with other techniques for numerical derivative
-calculation: automatic differentiation, forward finite difference, and central
-finite difference. The complex-step method was \note[Michael]{re: I.1, could
-suggest claim of novelty} \remove[Michael]{newly} implemented in a massively
-parallel computational peridynamics code for the purpose of this comparison.
-The methods were compared through profiling of the code for accuracy,
-speed, efficiency, and parallel scalability. The research provides data that
-can serve as practical guide for code developers and analysts faced with
-choosing which method best suits the needs of their application code.
-Additionally, motivated by the reproducible research movement, all of the
-code, examples, and workflow to regenerate the data and figures in this paper
-are provided as open source.
+In order to maintain the quadratic convergence properties of Newton's method in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
+A goal of the study described in this paper was to establish the suitability of an underexplored method for numerical computation of tangent-stiffness operators, referred to as ``complex-step'', and compare the \note{Re: I.1} \remove{new} method with other techniques for numerical derivative calculation: automatic differentiation, forward finite difference, and central finite difference.
+The complex-step method was \note{Re: I.1} \remove{newly} implemented in a massively parallel computational peridynamics code for the purpose of this comparison.
+The methods were compared through profiling of the code for accuracy, speed, efficiency, and parallel scalability.
+The research provides data that can serve as practical guide for code developers and analysts faced with choosing which method best suits the needs of their application code.
+Additionally, motivated by the reproducible research movement, all of the code, examples, and workflow to regenerate the data and figures in this paper are provided as open source.
 
@@
 
-In order to maintain the quadratic convergence properties of the first-order
-Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in
-quasi-static nonlinear analysis of solid structures it is crucial to obtain
-accurate, algorithmically consistent tangent-stiffness matrices. For an
-extremely small class of nonlinear material models, these consistent
-tangent-stiffness operators can be derived analytically; however, most often in
-practice, they are found through numerical approximation of derivatives.
-
-A goal of this study was to develop and evaluate a \note[Michael]{re: I.1}
-\change[Michael]{new}{relatively underexplored}, accurate, and practical method
-for calculating tangent-stiffness matrices against established methods.  This
-new method is based on a complex number Taylor series expansion and referred to
-as CTSE or the ``complex-step'' method. The practical accuracy of CTSE is
-determined by comparing its measured accuracy\note[Michael]{re:
-I.1}\remove[Michael]{new} relative to the measured accuracy of popular finite
-differencing techniques used for computing tangent-stiffness matrices. For this
-comparison the exact to machine precision derivatives computed with
-\emph{automatic differentiation} serve as the standard. \note[Michael]{re:
-II.1}For the reader's convenience, from plasticity theory, the algorithmic
-consistency of the tangent-stiffness matrix refers to consistency with the
-integration of a plasticity model under the Khun-Tucker constraints or similar
-where the tangent-stiffness matrix is computed using the elastoplastic modulii
-that are determined from the integration \cite{simo1998}.
+In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent\footnote{\note{Re: II.1} In the context of plasticity thoery, the algorithmic consistency of the tangent-stiffness matrix refers to consistency with the integration of a plasticity model under the Khun-Tucker constraints or similar where the tangent-stiffness matrix is computed using the elastoplastic modulii that are determined from the integration \cite{simo1998}.} tangent-stiffness matrices.
+For an extremely small class of nonlinear material models, these consistent tangent-stiffness operators can be derived analytically; however, most often in practice, they are found through numerical approximation of derivatives.
+
+A goal of this study was to develop and evaluate a \note{Re: I.1} \change{new }{relatively unexplored}, accurate, and practical method for calculating tangent-stiffness matrices against established methods.
+This \note{Re: I.1} \remove{new} method is based on a complex number Taylor series expansion and referred to as CTSE or the ``complex-step'' method.
+The practical accuracy of CTSE is determined by comparing its measured accuracy relative to the measured accuracy of popular finite differencing techniques used for computing tangent-stiffness matrices.
+For this comparison the,  exact to machine precision, derivatives computed with \emph{automatic differentiation} serve as the standard.
 
 Comparative data regarding the accuracy and compute cycle timing for the
-complex step, forward difference, central difference, and automatic
-differentiation methods are included to aid developers and analysts in
-computational mechanics code design who are faced with implementing a general
-method for tangent-stiffness matrix calculation.  A comparison of the methods was achieved through instrumentation of a
-massively parallel computational mechanics code.
-
-The scope of the application component of this study was limited to a single
-material model, an elastic peridynamic solid, implemented in the computational
-peridynamics code, \textit{Peridigm} \cite{peridigm}. Identifying a specific
-application provided a practical framework for implementing the
-\note[Michael]{re: I.1}\remove[Michael]{new} method and solving engineering
-problems to generate the data needed to compare the methods. In particular,
-\emph{Peridigm} was chosen because it combined several helpful
-characteristics: the utilization of Newton's method and tangent-stiffness
-matrices for solving nonlinear quasi-static problems, prior inclusion of
-forward difference, central difference, and automatic differentiation
-methods needed for comparison to the new complex-step method
-implementation, and being an agile-components code that makes use of
-distributed computing data structures via Trilinos \cite{trilinos} for
-efficient parallelization on large clusters. Agile-components is a software
-development term meaning to structure a software project for continuous
-development to respond to evolving specifications by making much use of
-techniques from object-oriented programming and reusing proven pre-existing
-code to reduce development time and increase reliability. The modularity of
-\emph{Peridigm} allowed the adding of new features, such as those required by the
-study.
-
-The aim of this paper is not only to \note[Michael]{re: I.1}\change[Michael]{introduce the new}{study the} complex-step method \change[Michael]{in
-the context of}{of} evaluating tangent-stiffness matrices, but to serve as a review
-for other differentiation techniques useful for solving non-linear systems with
-Newton's method. After presenting background information on the underlying
-methods to be discussed, presented in this paper are: In subsection~\ref{subsec:TS} a description of
-tangent-stiffness matrices, detailed directions for producing
-tangent-stiffness matrices with each of the methods identified above, and a
-description and justification of the \note[Michael]{re: I.1}\remove[Michael]{new}complex-step method for calculating
-tangent-stiffness matrices, in sub-subsection~\ref{subsubsec:Impl} a description of implementing the complex-step method in
-the \emph{Peridigm} software, in section~\ref{JGAM} a description of the quantities-of-interest
-used to rank the methods, in section~\ref{sec:Res} a presentation and analysis of the results of the
-comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
-Finally, in the interest of replicable research, the reader is referred to the
-corresponding author's website for \emph{Peridigm} C++ source-code and data
-necessary to reproduce the results presented here. Additionally, included on
-the website is an example serial C++ library with classes for solving
-non-linear systems using the techniques discussed here, as well as two
-validation and verification example problems which show how to use the library.
-The purpose of the serial C++ library is to demonstrate in a less complex
-manner than in the \emph{Peridigm} code how one would use
-automatic differentiation or complex-step to better encourage their use and
-discussion.
-\note[Michael]{re: II.5}This open source resource is available at the corresponding author's website or at \href{https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive}{https://github.com/utsa-idl/Complex\_Step\_Jacobian\_Project\_Archive}.
+complex step, forward difference, central difference, and automatic differentiation methods are included to aid developers and analysts in computational mechanics code design who are faced with implementing a general
+method for tangent-stiffness matrix calculation.  A comparison of the methods was achieved through instrumentation of a massively-parallel computational mechanics code.
+
+The scope of the application component of this study was limited to a single material model, an elastic peridynamic solid, implemented in the computational peridynamics code, \textit{Peridigm} \cite{peridigm}.
+Identifying a specific application provided a practical framework for implementing the \note{Re: I.1} \change{new }{complex-step} method and solving engineering problems to generate the data needed to compare the methods.
+In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of forward-difference, central-difference, and automatic-differentiation methods needed for comparison to the complex-step method implementation, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
+Agile-components is a software development term meaning to structure a software project for continuous development to respond to evolving specifications by making much use of techniques from object-oriented programming and reusing proven pre-existing code to reduce development time and increase reliability.
+The modularity of \emph{Peridigm} allowed the adding of new features, such as those required by the study.
+
+The aim of this paper is not only to \note{Re: I.1} \change{introduce the new}{study the} complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques useful for solving non-linear systems with Newton's method.
+After presenting background information on the underlying methods to be discussed, presented in this paper are: In Section~\ref{subsec:TS} a description of tangent-stiffness matrices, detailed directions for producing tangent-stiffness matrices with each of the methods identified above, and a description and justification of the \note{Re: I.1} \remove{new} complex-step method for calculating tangent-stiffness matrices, in Section~\ref{subsubsec:Impl} a description of implementing the complex-step method in the \emph{Peridigm} software, in Section~\ref{JGAM} a description of the quantities-of-interest used to rank the methods, in Section~\ref{sec:Res} a presentation and analysis of the results of the comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
+Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for \emph{Peridigm} C++ source-code and data necessary to reproduce the results presented here.
+Additionally, included on the website is an example serial C++ library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
+The purpose of the serial C++ library is to demonstrate in a less complex manner than in the \emph{Peridigm} code how one would use automatic-differentiation or complex-step to better encourage their use and discussion.
+\note{Re: II.5} \add{This open source resource is available at the corresponding author's website or at} \href{https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive}{https://github.com/utsa-idl/Complex\_Step\_Jacobian\_Project\_Archive}.
 
@@
 
-This section contains background information on the differentiation techniques
-underlying the tangent-stiffness matrix calculation methods compared in the
-study. Since the derivation of first-order finite difference techniques from
-Taylor series expansion is considered well known, it is not described here.
-Instead the reader is referred to \cite[Chap. 4.1.3]{chapra2010}.
+This section contains background information on the differentiation techniques underlying the tangent-stiffness matrix calculation methods compared in the study.
+Since the derivation of first-order finite difference techniques from Taylor series expansion is considered well-known, it is not described here.
+Instead the reader is referred to \cite[Chap.
+4.1.3]{chapra2010}.
 
@@
 
-It is
-possible to approximate derivatives quite accurately with a technique based on
-a complex variable Taylor series expansion of a function.  This method was
-first described by Lyness and Moler
-\cite{lyness1967numerical,lyness1968differentiation} and has more recently been
-rediscovered for use in engineering analysis.  The basic idea is that a model parameter can
-be made complex and expanded in a Taylor series about a small perturbation
-along the imaginary axis by some arbitrary value $h$ as follows
-%
-\begin{align} f ( x + i h ) =& f(x) + \frac{\partial f(x)}{\partial x} \frac{i
-h}{1!}  \notag \\ &+ \frac{\partial^2 f(x)}{\partial x^2} \frac{(i h)^2}{2!} +
-\frac{\partial^3 f(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots.
-\label{eqn:complexTaylor} \end{align}
-%
-Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor})
-and solving for the first derivative and ignoring terms of $\mathcal{O}\left(
-h^2 \right)$ yields an estimate
-%
-\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{\mbox{Im}
-\left( f (x + i h) \right)}{h} +\mathcal{O}\left( h^2 \right).  \label{eqn:complexFirstDeriv} \end{equation}
-%
-Thus, an estimate of the first derivative of a function can be made by only
-utilizing one functional evaluation of a perturbed model parameter along the
-imaginary axis.  The step-size $h$ is arbitrary and can be made as small as
-practical (even to machine precision without the dangers of roundoff error as
-in finite differences) to yield an accurate estimate of this derivative.  The
-only disadvantage of this technique is the necessity of requiring the functions
-to accept complex numbers as arguments. Appendix~\ref{sec:appendixA} includes a
-simple example that illustrates the use of the complex-step method for the
-derivative calculation of a function. 
-%
-\note[Michael]{re: II.6}
-For the reader's convenience the forward difference equivalent to~\ref{eqn:complexFirstDeriv} is
-%
-\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x)}{h} +\mathcal{O}\left( h \right). \label{eqn:forwardFirstDeriv} \end{equation},
-%
-while the central difference equivalent to~\ref{eqn:complexFirstDeriv} is,
-%
-\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x - h)}{2h} +\mathcal{O}\left( h^2 \right). \label{eqn:centeredFirstDeriv} \end{equation} \cite[Chap. 4.1.3]{chapra2010}.
-%
-Note that the order of the truncation error in~\ref{eqn:forwardFirstDeriv} indicates that for a given $h$ forward difference is less accurate an approximation than central difference or complex step.
-
-\note[Michael]{re: I.4, John all the rest of this section is new}
-It is appropriate to detail the prior work discussing aspects of the
-complex-step method in order to indicate what is new in the instant work.  The
-CS method was introduced in a work by J. N. Lyness and C. B. Moler
-\cite{lyness1967numerical}.  In this prior work, the CS method is identified in
-an infinite series form and the computing of analytical derivatives is shown. This prior
-work develops the CS method, while the instant work is about an application of
-the CS method.
-
-A follow-on work by J. N. Lyness demonstrates a 'truncated' approximate version
-of the original CS method to make it suitable for use on digital computers \cite{lyness1968differentiation}.
-This form of the CS method is used in the instant work.
-The prior work develops the CS method, while the instant work is about an application of the CS method.
-
-A work by W. Squire and G. Trapp shows through empirical data that CS offers superior accuracy
-to FD and CD \cite{squire1998using}. The instant work uses the CS method as shown in the
-prior work and was motivated by the prior work's suggestion that CS is highly accurate. 
-The instant work examines applications of the CS method beyond taking derivatives alone,
-whereas the prior work does not.
-
-A work by J. C. Newman, W. K. Anderson and L. D. Whitfield discusses computing
-sensitivities in a CFD analysis using complex step and compares the results to
-central difference produced results \cite{newman1998}. The work is also notable
-for identifying automatic differentiation as a competing method to CS and
-suggesting that complex step could be used to compute Jacobians of residuals
-for use in Newton-Krylov schemes for solving non-linear systems. The instant
-work is novel with respect to the prior work because the instant work
-investigates CS as applied to a numerical method for solving non-linear
-systems, rather than as in the prior work investigating the derivatives
-themselves produced with the CS method.
-
-In a work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta, CS is used to
-compute Jacobians for integrating plastic stress to determine
-elastoplastic modulii and for computing the Jacobian of the residual for the global
-problem, both of which are an application of the CS method to the solution of
-non-linear systems \cite{perez2000numerical}. These elastoplastic modulii are
-used in computing the global Jacobian  and are said to be 'algorithmically
-consistent' in that they are consistent with the algorithm used to return the
-predicted stress to the yield surface. Examples of interest include FEM
-simulations on two plastic material models with complicated hardening rules.
-Comparing the prior work with the instant work, the prior work demonstrates CS
-on a serial computer and discusses a classical solid mechanics formulation.
-Automatic differentiation is not investigated and CPU time measurements are not
-taken. Scalability studies with a variety of equivalent, progressively more
-refined meshes are not done. Scalability as a function of parallel computer
-size is not investigated. The instant work is novel with respect to the prior work in
-offering a discussion of CS as applied to peridynamics simulation on a parallel
-computer, CPU time data, and comparison of CS to automatic differentiation as a competing method.
-
-Another work by A. Perez-Foguet, A. Rodriguez-Ferran, A. Huerta applies CS in
-computing derivatives needed for integrating stress, obtaining algorithmically
-consistent elastoplastic modulii, and obtaining the Jacobian of the global
-residual, for FEM simulations using the MRS-Lade material model
-\cite{perez2012numerical}. In this prior work, CPU runtime is measured.
-Comparing the prior work to the instant work, the prior work demonstrates CS on
-a serial computer and discusses a classical solid mechanics formulation.
-Automatic differentiation is not investigated.  Scalability studies with a
-variety of equivalent, progressively more refined meshes are not done.
-Scalability as a function of distributed system size is not investigated.  The
-instant work is novel with respect to the prior work in offering a discussion
-of CS as applied to peridynamics simulation on a parallel computer, and
-comparison of CS to automatic differentiation as a competing method.
-
-A work by Joaquim R. R. A. Martins, Ilan M. Kroo and Juan J. Alonso
-investigates using CS for computing aerodynamic sensitivities
-\cite{martins2000automated}. Additionally, higher order accuracy versions of CS
-are discussed, as is the task of implementing complex-step in most widely used
-general purpose programming languages. Also discussed is the relative ease of
-implementing a CS scheme versus an AD based scheme for computing sensitivities
-in an established code. The instant work is novel with respect to the prior
-work because the instant work investigates CS as applied to a numerical method
-for solving non-linear systems, rather than as in the prior work investigating
-the use of the CS method to obtain sensitivities.
-
-A work by W. K. Anderson, J. C. Newman, D. l. Whitfield and E. J. Nielsen
-discusses the CS method as applied to computing sensitivities in CFD simulations
-\cite{anderson2001sensitivity}. CS is compared to AD both in Fortran and C++
-applications. In this prior work it is suggested that forward mode AD is
-mathematically equivalent to first order CS. The instant work is novel with
-respect to the prior work because the instant work investigates CS as applied
-to a numerical method for solving non-linear systems, rather than as in the
-prior work investigating the derivatives themselves produced with the CS
-method.
-
-A work by J. Martins, P. Sturdza, J. Alonso demonstrates that the CS method can
-be performed by using complex datatypes for the variables of functions
-comprising an analysis code \cite{martins2003complex}. The analysis code can
-comprise a simulation and derivatives of response variables called
-sensitivities can be computed. The accuracy of this approach is compared to
-automatic differentiation. The instant work is novel with respect to the prior
-work because the instant work investigates CS as applied to a numerical method
-for solving non-linear systems, rather than as in the prior work investigating
-the derivatives themselves produced with the CS method.
-
-A work by  K. L. Lai and J. L. Crassidis discusses first and second order CS
-methods for computing derivatives and choosing optimal step size
-\cite{lai2008extensions}. The instant work discusses using the derivatives
-computed with the CS method in a numerical method for solving a non-linear
-system, while in the prior work the derivatives obtained from the CS method 
-are the object of interest.
-
-A work by A. H. Al-Mohy and N. J. Higham investigates the use of complex step
-in computing Fr\'{e}chet derivatives of matrix to matrix functions
-\cite{al2010complex}. To compute the Jacobian of the residual in the
-peridynamics simulation as discussed in the instant work is analogous is
-computing a Fr\'{e}chet derivative. Where the application of the CS derivatives in
-the prior work is limited to a particular class of abstract functions to
-determine the CS method's suitability for use in condition number estimation, the
-instant work applies CS to solving a solid mechanics problem. The instant work is not
-about computing Fr\'{e}chet derivatives with CS but moreover using them in a
-numerical method to solve a set of equations describing a physical system.
-
-A work by W. Jin, B. H. Dennis, and B. P. Wang discusses using CS to compute
-sensitivities in a solid mechanics FEM analysis \cite{jin2010improved}. In the
-prior work the derivatives produced by CS are the object of
-interest, whereas the instant work uses CS in a numerical method for solving a
-non-linear system.
-
-A work by A. Voorhees, H. Millwater, and R. Bagley discusses the background of
-the CS method and frames it as a subset of Fourier Differentiation. CS and
-Fourier Differentiation are then applied to computing sensitivities within a
-solid mechanics FEA \cite{voorhees2011complex}. A notable technique
-demonstrated by this prior work is the representation of complex valued nodal
-coordinates with additional degrees of freedom in the FE model, to be
-interpreted by a matrix representation of complex numbers during computations.
-In the prior work, CS is discussed and related to a broader class of
-differentiation methods and used to compute sensitivities. In the instant work,
-CS is used in a numerical method for solving non-linear systems.
-
-In another work by H. Millwater, C. Quintana, J. Garza and David Wagner the
-concept of incorporating complex values into the degrees of freedom of finite
-elements is called ZFEM \cite{millwater2013application}. Notable in this work
-is the discussion of creating user defined element types for use with the
-popular \emph{ABAQUS} software package which enables the use of ZFEM in that
-software. \emph{ABAQUS} is a trademark of Dassault Systemes
-\cite{systemes2012abaqus}. In this prior work the goal of using CS through the
-ZFEM technique is to compute shape sensitivities, while in the instant work CS
-is used in a numerical method for solving non-linear systems.
-
-A work by R. Abreu, D. Stich, and J. Morales discusses computing derivatives
-with a version of the CS method which features the combined use
-of real and imaginary valued step components \cite{abreu2013generalization}.
-This prior work suggests that the modified CS method can achieve extended
-approximation accuracy up to fourth order.  The prior work discusses using the
-CS method to compute derivatives and attempts to extend the method for that
-application, where the derivatives themselves are the object of interest. In
-contrast, the instant work discusses using the derivatives computed with CS in
-a numerical method for solving a non-linear system.
-
-The body of prior literature has introduced the application of CS to
-single dimension and multidimensional derivatives, sensitivity analysis,
-non-linear solution schemes, the extension of CS for higher accuracy, 
-and implementation strategies for CS in new and existing codes.
-With respect to the prior work discussing the CS method, the novel
-aspects of the instant work are then its discussion of the application of CS to
-peridynamics simulations run in \emph{Peridigm}, a parallel implementation of
-CS and a compilation of CPU time data, where each of CS, AD, FD and CD are
-compared.
+It is possible to approximate derivatives quite accurately with a technique based on a complex-variable Taylor series expansion of a function.
+This method was first described by Lyness and Moler \cite{lyness1967numerical,lyness1968differentiation} and has more recently been rediscovered for use in engineering analysis.
+The basic idea is that a model parameter can be made complex and expanded in a Taylor series about a small perturbation along the imaginary axis by some arbitrary value $h$ as follows
+%
+\begin{align} 
+    f ( x + i h ) =& f(x) + \frac{\partial f(x)}{\partial x} \frac{i h}{1!}  \notag \\ &+ \frac{\partial^2 f(x)}{\partial x^2} \frac{(i h)^2}{2!} + \frac{\partial^3 f(x)}{\partial x^3} \frac{(i h)^3}{3!} + \cdots.
+\label{eqn:complexTaylor} 
+\end{align}
+%
+Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor}) and solving for the first derivative and ignoring terms of $\mathcal{O}\left(h^2 \right)$ yields an estimate
+%
+\begin{equation} \frac{\partial f( x )}{\partial x} = \frac{\mbox{Im} \left( f (x + i h) \right)}{h} +\mathcal{O}\left( h^2 \right).  
+\label{eqn:complexFirstDeriv} 
+\end{equation}
+%
+Thus, an estimate of the first derivative of a function can be made by only utilizing one functional evaluation of a perturbed model parameter along the imaginary axis.
+The step-size $h$ is arbitrary and can be made as small as practical (even to machine precision without the dangers of roundoff error as in finite differences) to yield an accurate estimate of this derivative.
+The only disadvantage of this technique is the necessity of requiring the functions to accept complex numbers as arguments.
+Appendix~\ref{sec:appendixA} includes a simple example that illustrates the use of the complex-step method for the derivative calculation of a function.
+%
+\note{Re: II.6} \add{For completeness, the forward- and central-difference formulas corresponding to equation}~(\ref{eqn:complexFirstDeriv}) \add{are} \cite[Chap. 4.1.3]{chapra2010}
+%
+\begin{equation} 
+\frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x)}{h} +\mathcal{O}\left( h \right), 
+\label{eqn:forwardFirstDeriv} 
+\end{equation}
+%
+\begin{equation} 
+\frac{\partial f( x )}{\partial x} = \frac{f (x + h) - f (x - h)}{2h} +\mathcal{O}\left( h^2 \right). 
+\label{eqn:centeredFirstDeriv} 
+\end{equation} 
+%
+\add{The order of the truncation error in the forward-difference formula (equation}~(\ref{eqn:forwardFirstDeriv})\add{) indicates that for a given $h$ this method is less accurate an approximation than either central-difference or complex-step which have errors of $\mathcal{O}(h^2)$.}
+
+\note{Re: I.4} \add{As mentioned earlier, the CS method was originally introduced in a paper by Lyness and Moler} \cite{lyness1967numerical}\add{.
+A follow-on contribution by Lyness demonstrated a ``truncated'' approximate version of the original CS method to make it suitable for use on digital computers} \cite{lyness1968differentiation}\add{.
+After these original papers, little can be found in the literature regarding the method until Squire and Trapp showed through empirical data that CS offers superior accuracy to FD and CD} \cite{squire1998using} \add{in computing numerical derivatives.
+Papers by Newman et al.\ } \cite{newman1998} \add{and Anderson et al.\ } \cite{anderson2001sensitivity} \add{discuss computing sensitivities in a CFD analysis using CS and compares the results to sensitivities produced with central-difference derivatives.
+The work is also notable for identifying automatic-differentiation as a competing method to CS and suggesting that complex-step could be used to compute Jacobians of residuals for use in Newton-Krylov schemes for solving non-linear systems.
+Perez-Foguet et al.\ } \cite{perez2000numerical, perez2012numerical} \add{used the CS method to compute algorithmically consistent elastoplastic moduli in non-linear solid mechanics problems.
+A paper by Joaquim et al.\ } \cite{martins2000automated} \add{investigates using CS for computing aerodynamic sensitivities.
+Additionally, higher-order accuracy versions of CS are discussed, as is the task of implementation in most widely used general purpose programming languages.
+Discussed is the relative ease of implementing a CS scheme versus an automatic-differentiation based scheme for computing sensitivities in an established code.
+Martins et al.\ } \cite{martins2003complex}  \add{demonstrates that the CS method can be performed by using complex datatypes for the variables of functions comprising an analysis code.
+The accuracy of this approach, with respect to the derivatives themselves, is compared to automatic-differentiation in a serial analysis code.
+Lai and Crassidis} \cite{lai2008extensions} \add{discuss first- and second-order CS methods for computing derivatives and choosing optimal step sizes.
+Al-Mohy and Higham} \cite{al2010complex} \add{investigate the use of CS in computing Fr\'{e}chet derivatives of matrix functions.
+Jin et al.\ } \cite{jin2010improved} \add{discuss using CS to compute sensitivities in a solid mechanics FEM analysis.
+Voorhees et al.\ } \cite{voorhees2011complex} \add{discusses the background of the CS method and frames it as a subset of Fourier Differentiation.
+CS and Fourier Differentiation are then applied to computing sensitivities within a solid mechanics finite element analysis.
+They implemented the complex-values as additional degrees of freedom in the finite element stiffness matrices using a real-valued matrix representation of the complex numbers where required.
+In a follow-on work,  Millwater et al.\ } \cite{millwater2013application} \add{extended this concept, calling it ZFEM, and implemented the method into user defined element types for use with the popular ABAQUS} \cite{systemes2012abaqus} \add{FEA software package.
+Finally, Abreu et al.\ } \cite{abreu2013generalization} \add{discusses computing derivatives with a version of the CS method which features the combined use of real and imaginary valued step components, as in Fourier differentiation.
+They suggests that the modified CS method can achieve extended approximation accuracy up to fourth order.
+The body of prior literature has introduced the application of CS to single dimension and multidimensional derivatives, sensitivity analysis, non-linear solution schemes, the extension of CS for higher accuracy, and implementation strategies for CS in new and existing codes.
+With respect to the prior work discussing the CS method, the novel aspects of the instant work are then its discussion of the application of CS to peridynamics simulations run in \emph{Peridigm}, a parallel implementation of CS and a compilation of CPU time data, where each of CS, AD, FD and CD are compared.}
 
@@
 
-Automatic differentiation (AD) is a computerized method for computing exact
-derivatives based on the chain-rule from calculus. AD takes advantage of the fact
-that any mathematical function executed on a computer, no matter how
-complicated, is a ``composition of simple operations'' (addition, multiplication, exponentiation and the like) each having known analytical derivatives
-\cite{ref-sacado-presentation}. For reference, the AD implementation used in
-this study is part of the \emph{Sacado} package from the \emph{Trilinos}
-agile-component libraries developed at Sandia National Laboratories
-\cite{ref-Sacado}.
-
-An AD system evaluates composition functions in the expected manner: it works
-by first evaluating the innermost function of the composition, then presenting
-that function's output as input to the next level function until all levels are
-complete, observing commonly expected order of operations. However, an AD
-system does additional work during a function evaluation in that as each nested
-function is evaluated, the function's partial derivative with respect to the
-designated variables of the given input is also calculated.  This is possible
-because the elementary math functions are hard-coded into the AD source-code
-along with their analytical derivatives, and linked by special instructions, so
-that when the elementary math functions are called upon for computation, their
-partial derivatives may be computed and stored in a sequence. The AD system
-then multiplies the final sequence of partial derivatives together to produce
-the exact equivalent to taking a partial derivative of the corresponding
-composition function with respect to a designated variable at a particular
-value. 
-%
-It is obvious, but bears mentioning, that the AD system could simply
-store one value for partial derivatives, modifying it as appropriate for every
-function evaluation rather than keeping a sequence. This is significant because
-modifying a single value rather than keeping a list of values corresponding to
-every level of a composition function, which itself may not be the sole one
-needing to be evaluated, represents a savings in memory usage, which at large
-scales is an active concern. In the literature, the AD scheme described here is
-called \emph{forward automatic differentiation}. Only forward AD will be
-covered here since it is the implementation used in \emph{Sacado} and therefore
-in this study; however, the reader is referred to the introduction section of
-\cite{ref-AD-methods} and its citations for further information on AD, and
+Automatic differentiation (AD) is a computerized method for computing exact derivatives based on the chain-rule from calculus.
+AD takes advantage of the fact that any mathematical function executed on a computer, no matter how complicated, is a ``composition of simple operations'' (addition, multiplication, exponentiation and the like) each having known analytical derivatives \cite{ref-sacado-presentation}.
+For reference, the AD implementation used in this study is part of the \emph{Sacado} package from the \emph{Trilinos} agile-component libraries developed at Sandia National Laboratories \cite{ref-Sacado}.
+
+An AD system evaluates composition functions in the expected manner: it works by first evaluating the innermost function of the composition, then presenting that function's output as input to the next level function until all levels are complete, observing commonly expected order of operations.
+However, an AD system does additional work during a function evaluation in that as each nested function is evaluated, the function's partial derivative with respect to the designated variables of the given input is also calculated.
+This is possible because the elementary math functions are hard-coded into the AD source-code along with their analytical derivatives, and linked by special instructions, so that when the elementary math functions are called upon for computation, their
+partial derivatives may be computed and stored in a sequence.
+The AD system then multiplies the final sequence of partial derivatives together to produce the exact equivalent to taking a partial derivative of the corresponding composition function with respect to a designated variable at a particular value.
+%
+It is obvious, but bears mentioning, that the AD system could simply store one value for partial derivatives, modifying it as appropriate for every function evaluation rather than keeping a sequence.
+This is significant because modifying a single value rather than keeping a list of values corresponding to every level of a composition function, which itself may not be the sole one needing to be evaluated, represents a savings in memory usage, which at large scales is an active concern.
+In the literature, the AD scheme described here is called \emph{forward automatic differentiation}.
+Only forward AD will be covered here since it is the implementation used in \emph{Sacado} and therefore in this study; however, the reader is referred to the introduction section of \cite{ref-AD-methods} and its citations for further information on AD, and
 particularly \cite{ref-on-AD} which is foundational.
-Appendix~\ref{sec:appendixB} includes an example that walks the reader through
-the process a computer uses to compute derivatives via AD.  Some things to note
-about AD are that no approximation of derivatives is being made because the
-analytical forms of the partial derivatives of the elementary math functions are defined
-alongside them. The accuracy of AD is then limited by the precision of the AD
-system's definition of the elementary math functions and their partial
-derivatives.
+
+Appendix~\ref{sec:appendixB} includes an example that walks the reader through the process a computer uses to compute derivatives via AD.
+Some things to note about AD are that no approximation of derivatives is being made because the analytical forms of the partial derivatives of the elementary math functions are defined alongside them.
+The accuracy of AD is then limited by the precision of the AD system's definition of the elementary math functions and their partial derivatives.
 
@@
 
-\note[Michael]{re: I.5, next section is taken from the thesis.}
+\note[]{Re: I.5, next section is taken from the thesis.}
 \section{Working examples of CS, AD and FD for Newton's method}
@@
 
-\note[Michael]{re: I.2, John I think you would better handle expanding this section,
-I was going to essentially copy silling:psa.}\change[Michael]{Briefly, peridynamics}Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
+\note[]{Re: I.2, John I think you would better handle expanding this section,
+I was going to essentially copy silling:psa.}\change[]{Briefly, peridynamics}Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
 is a nonlocal reformulation of the partial-differential equations that provide
@@
 
-\note[Michael]{re: I.2. John, this was to be the skeleton for the new sections on peridynamics.}
+\note[]{Re: I.2. John, this was to be the skeleton for the new sections on peridynamics.}
 \subsubsection{Vector states for state based peridynamic theory}
@@
 
-\note[Michael]{re: I.2} \add[Michael]{
+\note[]{Re: I.2} \add[]{
 \subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simulation}
@@
 thought of as distributed memory parallel cousins to standard C++
-\change[Michael]{STL}{Standard Template Library} vectors, are hard coded to be
+\change[]{STL}{Standard Template Library} vectors, are hard coded to be
 of type {\tt double} and can not be simply declared as having a complex data
-type. \note[Michael]{re: I.3}\add[Michael]{For the reader's convenience,
+type. \note[]{Re: I.3}\add[]{For the reader's convenience,
     \emph{Epetra} vectors and STL vectors are what are known as 'container
@@
 scoped, and dynamically casting persistent memory variables to complex data
-types where necessary. \note[Michael]{re: I.3}\add[Michael]{For the reader's
+types where necessary. \note[]{Re: I.3}\add[]{For the reader's
 convenience, 'locally scoped' in the C++ programming language refers to
@@
 available in \emph{Trilinos} that are capable of taking an explicit
-instantiation of any data type including {\tt complex}. \note[Michael]{re:
-I.3}\add[Michael]{For the purposes of this work, an explicit instantiation
+instantiation of any data type including {\tt complex}. \note[]{Re:
+I.3}\add[]{For the purposes of this work, an explicit instantiation
 refers to requesting the creation of an object, or instance, of a templated
@@
 many advanced techniques in C++ such as multiple virtual inheritance, pointer
-arithmetic, and distributed data structures, \note[Michael]{re:
-I.3}\add[Michael]{for example \emph{Epetra} vectors, were used in the basic
+arithmetic, and distributed data structures, \note[]{Re:
+I.3}\add[]{for example \emph{Epetra} vectors, were used in the basic
     code and by necessity in the modified code. These concepts are not vital to
@@
 terms of accuracy and speed. This series would comprise the single core runs. 
-\note[Michael]{re: II.16, specific test parameters were asked for}
+\note[]{Re: II.16, specific test parameters were asked for}
 The specific parameters of these tests can be found in \emph{Peridigm} input
@@
          \toprule
-         Peri. nodes & Nozero TS elems. & Horizon & Load rate & # Load steps & Abs. tolerance \\ 
-                     & $\times 10^6$    & \si{\meter}& \si{\centimeter\per\second} &   &     \\
+         Peri. nodes & Nozero TS elems. & Horizon & Load rate & Number Load steps & Abs. tolerance \\ 
+                     & $\times 10^6$    & \si{\meter}& \si{\centi\meter\per\second} &   &     \\
         \midrule
@@
 
-\note[Michael]{re: I.3}\add[Michael]{For the purposes of this work message passing is what
+\note[]{Re: I.3}\add[]{For the purposes of this work message passing is what
 synchronizes multiple computers that each are working on a separate section
@@
 heuristically derived based on computational node spacing using the method explained in \cite[pp. 90]{ref-Adaggio}.  
-\add[Michael]{To summarize, a spacing of $1.E-6$ times grid spacing is used to provide acceptable accuracy but prevent roundoff error, 
+\add[]{To summarize, a spacing of $1.E-6$ times grid spacing is used to provide acceptable accuracy but prevent roundoff error, 
 as is not a user determined setting in \emph{Peridigm}.} One of the beauties
@@
 %
-\note[Michael]{re: II.22, relative accuracy figures were asked for, these were taken from the archive data}
+\note[]{Re: II.22, relative accuracy figures were asked for, these were taken from the archive data}
 It can be seen in Figs.~\ref{fig:serial_accuracy} and ~\ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods.
@@
 
-\note[Michael]{Re: I.5, everything after this line is changed}
+\note[]{Re: I.5, everything after this line is changed}
 \section{Convergence and displacement accuracy studies} 
@@
 
-\note[Michael]{re: I.5, this and other material was added in from the thesis to discuss accuracy}
-\note[Michael]{re: I.5, the entire conclusion section was replaced with the thesis version}
+\note[]{Re: I.5, this and other material was added in from the thesis to discuss accuracy}
+\note[]{Re: I.5, the entire conclusion section was replaced with the thesis version}
 \section{Conclusions}
@@
 require compiling a version of the AD library being used that were compatible
-with the accelerator or GPU being used. \add[Michael]{A working definition of byte-copyable
+with the accelerator or GPU being used. \add[]{A working definition of byte-copyable
 is that the collection of data to transfer is made of identically sized pieces.
@@
 dependence on an external AD library.  From this perspective, the method may
-still be an attractive method for TS matrix evaluation. \add[Michael]{However,
+still be an attractive method for TS matrix evaluation. \add[]{However,
 at least for the models producing the results shown here, each of the

### 2014-04-11 jacobian_calc_CMAME.tex
Updates to MDB changes

@@
 A follow-on contribution by Lyness demonstrated a ``truncated'' approximate version of the original CS method to make it suitable for use on digital computers} \cite{lyness1968differentiation}\add{.
-After these original papers, little can be found in the literature regarding the method until Squire and Trapp showed through empirical data that CS offers superior accuracy to FD and CD} \cite{squire1998using} \add{in computing numerical derivatives.
+After these original papers, little can be found in the literature regarding the method until Squire and Trapp showed through numerical experiments that CS offers superior accuracy to FD and CD} \cite{squire1998using} \add{in computing numerical derivatives.
 Papers by Newman et al.\ } \cite{newman1998} \add{and Anderson et al.\ } \cite{anderson2001sensitivity} \add{discuss computing sensitivities in a CFD analysis using CS and compares the results to sensitivities produced with central-difference derivatives.
@@
 CS and Fourier Differentiation are then applied to computing sensitivities within a solid mechanics finite element analysis.
-They implemented the complex-values as additional degrees of freedom in the finite element stiffness matrices using a real-valued matrix representation of the complex numbers where required.
+They implemented the complex-values as additional degrees of freedom in the finite element stiffness matrices using a real-valued matrix representation of the complex-numbers where required.
 In a follow-on work,  Millwater et al.\ } \cite{millwater2013application} \add{extended this concept, calling it ZFEM, and implemented the method into user defined element types for use with the popular ABAQUS} \cite{systemes2012abaqus} \add{FEA software package.
+Nishida et al.\ } \cite{nishida2013} \add{used the complex-step method to compute Jacobian matrices for implementation in a Levenberg-Marquardt optimization routine for design-of-experiments purposes.
 Finally, Abreu et al.\ } \cite{abreu2013generalization} \add{discusses computing derivatives with a version of the CS method which features the combined use of real and imaginary valued step components, as in Fourier differentiation.
 They suggests that the modified CS method can achieve extended approximation accuracy up to fourth order.
-The body of prior literature has introduced the application of CS to single dimension and multidimensional derivatives, sensitivity analysis, non-linear solution schemes, the extension of CS for higher accuracy, and implementation strategies for CS in new and existing codes.
-With respect to the prior work discussing the CS method, the novel aspects of the instant work are then its discussion of the application of CS to peridynamics simulations run in \emph{Peridigm}, a parallel implementation of CS and a compilation of CPU time data, where each of CS, AD, FD and CD are compared.}
+This recent literature has introduced the application of CS to single-dimension and multidimensional derivatives, sensitivity analysis, non-linear solution schemes, optimization, the extension of CS for higher accuracy, and implementation strategies for CS in new and existing codes.
+The current paper presents a comparison of the complex-step method to the other methods for numerical differentiation in as close to a one-to-one implementation as possible for direct comparison purposes in a large-scale engineering analysis code.
+To the author's knowledge it is the first application of the complex-step method to solve peridynamics mechanics problems, and the first implementation and comparison of the method in large-scale distributed memory parallel setting.
+}
 
@@
 However, an AD system does additional work during a function evaluation in that as each nested function is evaluated, the function's partial derivative with respect to the designated variables of the given input is also calculated.
-This is possible because the elementary math functions are hard-coded into the AD source-code along with their analytical derivatives, and linked by special instructions, so that when the elementary math functions are called upon for computation, their
-partial derivatives may be computed and stored in a sequence.
+This is possible because the elementary math functions are hard-coded into the AD source-code along with their analytical derivatives, and linked by special instructions, so that when the elementary math functions are called upon for computation, their partial derivatives may be computed and stored in a sequence.
 The AD system then multiplies the final sequence of partial derivatives together to produce the exact equivalent to taking a partial derivative of the corresponding composition function with respect to a designated variable at a particular value.
@@
 
-In solid and structural computational mechanics the tangent-stiffness matrix is
-a linearization operator that describes the stiffness of a system in response
-to small displacements imposed upon the current configuration of the system.
-Mathematically speaking, the tangent-stiffness represents the gradient of a
-high-dimensional energy surface that ``points'' in the minimum direction. While
-the terminology of solid mechanics is used here, it's important to note that
-these operators appear in other physical settings and are known by other names,
-e.g. a \emph{transmissibility matrix} in the context of a Poisson problem or,
-more generally, a \emph{Jacobian matrix} in mathematical optimization.
-
-While, for simplicity, the examples shown in the following refer only to linear
-problems, tangent-stiffness matrices generally arise in the context of
-non-linear analysis where Newton's method or quasi-Newton methods are used
-in the minimization of a given residual function.  In this context, the
-tangent-stiffness matrix can be thought of as the linearization of the system
-about a particular configuration.  It has been shown
-\cite{hughes1978consistent,hughes1978unconditionally} that in order to preserve
-the quadratic convergence properties of the Newton's methods, it is necessary
-that this linearization is carried out in a manner that is \emph{consistent}
-with the algorithmic constraint equations used when computing the internal
-forces that arise due to deformations.  An example of these constraint
-equations would be the Kuhn-Tucker conditions \cite{simo1998} that are used in
-integration of a flow rule for plasticity modeling.  If the continuum
-tangent-modulii are used in place of the \emph{consistent} or \emph{algorithmic
-tangent moduli} in the solution of a non-linear solid mechanics problem,
-convergence may still be achieved, but not in the quadratic manner that makes
-Newton's method attractive for this class of problems.  In the setting of
-non-linear solid mechanics, there is a very small class of material model
-algorithms in which consistent tangent-stiffness operators can be derived
-analytically.  Perfect plasticity and plasticity with isotropic hardening used
-in combination with a general nearest-point projection or \emph{radial return
-algorithm} are a few examples.  Therefore, when general models (and certainly
-more complex ones) are implemented into a general purpose computational
-mechanics code, the tangent-stiffness operators are typically defined via a
-numerical approximation.
-
-A general mathematical formula for a tangent-stiffness matrix, $K_{ij}$, can be
-expressed in indicial notation as 
+In solid and structural computational mechanics the tangent-stiffness matrix is a linearization operator that describes the stiffness of a system in response to small displacements imposed upon the current configuration of the system.
+Mathematically speaking, the tangent-stiffness represents the gradient of a high-dimensional energy surface that ``points'' in the minimum direction.
+While the terminology of solid mechanics is used here, it's important to note that
+these operators appear in other physical settings and are known by other names, e.g.
+a \emph{transmissibility matrix} in the context of a Poisson problem or, more generally, a \emph{Jacobian matrix} in mathematical optimization.
+
+While, for simplicity, the examples shown in the following refer only to linear problems, tangent-stiffness matrices generally arise in the context of non-linear analysis where Newton's method or quasi-Newton methods are used in the minimization of a given residual function.
+In this context, the tangent-stiffness matrix can be thought of as the linearization of the system
+about a particular configuration.
+It has been shown \cite{hughes1978consistent,hughes1978unconditionally} that in order to preserve the quadratic convergence properties of the Newton's methods, it is necessary that this linearization is carried out in a manner that is \emph{consistent} with the algorithmic constraint equations used when computing the internal forces that arise due to deformations.
+An example of these constraint equations would be the Kuhn-Tucker conditions \cite{simo1998} that are used in integration of a flow rule for plasticity modeling.
+If the continuum tangent-modulii are used in place of the \emph{consistent} or \emph{algorithmic tangent moduli} in the solution of a non-linear solid mechanics problem, convergence may still be achieved, but not in the quadratic manner that makes Newton's method attractive for this class of problems.
+In the setting of non-linear solid mechanics, there is a very small class of material model algorithms in which consistent tangent-stiffness operators can be derived analytically.
+Perfect plasticity and plasticity with isotropic hardening used in combination with a general nearest-point projection or \emph{radial return algorithm} are a few examples.
+Therefore, when general models (and certainly more complex ones) are implemented into a general purpose computational mechanics code, the tangent-stiffness operators are typically defined via a numerical approximation.
+
+A general mathematical formula for a tangent-stiffness matrix, $K_{ij}$, can be expressed in indicial notation as 
 %
@@
 %
-where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function,
-$X_j$ is the $j^{\mbox{th}}$ component of the vector argument of $F$, and a
-particular value of the vector, $\vec{X}_0$ is chosen as the linearization
-point. One then evaluates the expression for each combination $(i, j)$
-corresponding to a (row, column) location in the tangent-stiffness matrix. The
-elements of a tangent-stiffness matrix can be estimated using any of the
-complex-step, AD, or common finite difference techniques for functions $F:R^1
-\rightarrow R^1$, since taking partial derivatives entails holding all but a
-single independent vector component of the argument of the function constant, and each element
-of the function can be evaluated independently of the others. \\
+where $F_i$ is the $i^{\mbox{th}}$ component of the vector valued function, $X_j$ is the $j^{\mbox{th}}$ component of the vector argument of $F$, and a particular value of the vector, $\vec{X}_0$ is chosen as the linearization point.
+One then evaluates the expression for each combination $(i, j)$ corresponding to a (row, column) location in the tangent-stiffness matrix.
+The
+elements of a tangent-stiffness matrix can be estimated using any of the complex-step, AD, or common finite difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the argument of the function constant, and each element of the function can be evaluated independently of the others.
+\\
 
@@
 %
-
-where $K_{ij}$ is the indicial notation representation of an element of the
-tangent-stiffness matrix at row $i$, column $j$. The first of the two terms in
-the numerator is the internal force, $F_i^{int}$ evaluated in the current
-deformed configuration, as a function of the displacement, $\vec{u}$,  plus a small perturbation $h$ in the
-direction $\hat{e}_j$ where $\hat{e}_j$ represents a unit-vector corresponding
-to the $j^{\mbox{th}}$ degree-of-freedom. The second of the two terms is
-$F_i^{int}$ evaluated in the current configuration. The denominator is the
+where $K_{ij}$ is the indicial notation representation of an element of the tangent-stiffness matrix at row $i$, column $j$.
+The first of the two terms in the numerator is the internal force, $F_i^{int}$ evaluated in the current deformed configuration, as a function of the displacement, $\vec{u}$,  plus a small perturbation $h$ in the direction $\hat{e}_j$ where $\hat{e}_j$ represents a unit-vector corresponding to the $j^{\mbox{th}}$ degree-of-freedom.
+The second of the two terms is $F_i^{int}$ evaluated in the current configuration.
+The denominator is the
 magnitude of perturbation called the \emph{probe-distance}.
@@
 %
-where the internal force, $F^{int}$ is now treated as function of complex
-vectors,  $\vec{u}$. The small perturbation, $h$, is now carried out on the
-imaginary axis and only the real coefficient to the imaginary part of the
-vector returned by $F^{int}$ is kept, hence the $\mbox{Im}(\cdot)$ operation.
-As highlighted in Section~\ref{sec:CSmethod}, notice that there is no
-subtraction operation taking place; therefore $h$ can be made very small to
-yield highly accurate derivatives. It should be mentioned that using
-complex-step to compute tangent-stiffness matrices was done in
-\cite{perez2000numerical,perez2012numerical}, but those references did not
-investigate the performance of complex-step in a massively parallel setting nor
-make comparisons to AD.
-
-When using automatic differentiation to calculate a tangent-stiffness matrix,
-one only needs to follow the definition of the tangent-stiffness matrix and
-issue the correct commands to the AD system. The formula is the same as the
-continuum formula
+where the internal force, $F^{int}$ is now treated as function of complex vectors,  $\vec{u}$.
+The small perturbation, $h$, is now carried out on the imaginary axis and only the real coefficient to the imaginary part of the vector returned by $F^{int}$ is kept, hence the $\mbox{Im}(\cdot)$ operation.
+As highlighted in Section~\ref{sec:CSmethod}, notice that there is no subtraction operation taking place; therefore $h$ can be made very small to yield highly accurate derivatives.
+It should be mentioned that using complex-step to compute tangent-stiffness matrices was done in \cite{perez2000numerical,perez2012numerical}, but those references did not
+investigate the performance of complex-step in a massively parallel setting nor make comparisons to AD.
+
+When using automatic differentiation to calculate a tangent-stiffness matrix, one only needs to follow the definition of the tangent-stiffness matrix and issue the correct commands to the AD system.
+The formula is the same as the continuum formula
 %
@@
 %
-but, with the caveat that $F^{int}$ is evaluated in a way that is
-algorithmically consistent, and therefore should yield quadratic convergence
-with Newton's method.
+but, with the caveat that $F^{int}$ is evaluated in a way that is algorithmically consistent, and therefore should yield quadratic convergence with Newton's method.
 
 \note[]{Re: I.5, next section is taken from the thesis.}
-\section{Working examples of CS, AD and FD for Newton's method}
+\section{Simple example of CS, AD and FD for solution of nonlinear problems.}
 \label{sec:WorkingExamples}
 
-In order to illustrate the complex-step technique in a simpler form than it
-appears in the \emph{Peridigm} code written for this study, an example
-header-only library and two example problem definition programs that use it
-have been created. All the materials described here are available in the
-associated project repository.
+In order to illustrate a comparison of the numerical differentiation techniques in a simpler form than it appears in the large-scale analysis code covered in the following sections, a simple example borrowed from \cite{rezaiee2010dynamic} was implemented with each of the methods, CS, AD, and FD.
+All the source code to reproduce the results described here are available in the associated paper repository.
 
@@
 
-The source code for the header-only library is entitled
-\emph{NewtonRaphson.hpp}. Note that the library is an interface to functions and
-data structures made available in the \emph{Trilinos} library, mentioned
-earlier, such that it is required in order for the examples described here
-to work. \emph{Trilinos} is freely available for download as of this
-writing at: \emph{trilinos.sandia.gov}.  Generally the header-only library
-works by taking in the users initial guessed equilibrium solution, target
-dependent variable values, system constants, and a function pointer to a
-model evaluation function. Solver parameters including an update scaling
-factor, tolerance, maximum iterations and probe length are chosen after the
-solver is instantiated as a new object. The behavior of the solver is
-specialized through a coding feature called inheritance to perform each of
-CS, AD and FD.  But, in common, each version of the solver works by
-computing the system Jacobian, solving a linear system as in that step of
-the Newton Raphson method, computing the residual and checking for
-convergence or overrun. Here the linear solution method is LU
-decomposition, while the chief convergence criterion is change in residual
-between iterations.  Additionally the problem can be partially respecified
-to allow for running a series of load steps having progressing target
-values, which is a typical mode also seen in \emph{Peridigm}. Applying
-boundary conditions incrementally in load steps assists convergence
-\cite{rezaiee2010dynamic}.
-
-\subsection{First example}
-\label{subsec:ProgOne}
-The first example program, \emph{NRexamples.cpp}, solves a simple nonlinear
-system describing the intersection of three infinite paraboloids in threespace.
-It works by defining force computation functions for each of the models, and
-then passing an address called a function pointer corresponding to each of
-these methods to a solver object which coordinates solving the problem and
-reporting output for the user based on their convergence criteria and other
-preferences like step-size and maximum number of iterations. CS, AD and FD are
-each represented and the user is free to adjust input settings by specifying
-command line arguments as the compiled program instructs, or perhaps create a
-shell script that varies parameters for each of the methods in order to perform
-a comparative study.
+The source code for the header-only library is entitled \emph{NewtonRaphson.hpp}.
+Note that the library is an interface to functions and data structures made available in the \emph{Trilinos} library, mentioned earlier, such that it is required in order for the examples described here to work.
+\emph{Trilinos} is freely available for download as of this writing at: \emph{trilinos.sandia.gov}.
+Generally the header-only library works by taking in the users initial guessed equilibrium solution, target dependent variable values, system constants, and a function pointer to a model evaluation function.
+Solver parameters including an update scaling factor, tolerance, maximum iterations and probe length are chosen after the solver is instantiated as a new object.
+The behavior of the solver is specialized through a coding feature called inheritance to perform each of CS, AD and FD.
+But, in common, each version of the solver works by computing the system Jacobian, solving a linear system as in that step of the Newton Raphson method, computing the residual and checking for convergence or overrun.
+Here the linear solution method is LU decomposition, while the chief convergence criterion is change in residual between iterations.
+Additionally the problem can be partially respecified to allow for running a series of load steps having progressing target values, which is a typical mode also seen in \emph{Peridigm}.
+Applying boundary conditions incrementally in load steps assists convergence.
 
@@
 \label{subsec:Validation}
-The second program,a \emph{NRComparison.cpp}, follows a case study that appears in
-\cite{rezaiee2010dynamic}. The example program departs from the cited
-literature by applying each of CS, AD and FD and verifying them against an
-analytical model of a nonlinear mechanical system involving a connected beam
-and spring. Fig. \ref{fig:TrussSchematic} is a schematic of the beam and spring
-system.
-
+The second program,a \emph{NRComparison.cpp}, follows a case study that appears in \cite{rezaiee2010dynamic}.
+The example program departs from the cited literature by applying each of CS, AD and FD and verifying them against an analytical model of a nonlinear mechanical system involving a connected beam and spring.
+Fig.
+\ref{fig:TrussSchematic} is a schematic of the beam and spring system.
+%
 \begin{figure}[tbp] \centering \includegraphics[width=0.7\textwidth]{./figs/truss.png}
@@
 \end{figure}
-
+%
 The equations describing this system and results for verification were obtained

### 2014-04-18 jacobian_calc_CMAME.tex
recolored mesh figures

@@
 
-\begin{figure}[tbp] \centering \includegraphics[width=0.7\textwidth]{./figs/mesh2000.png}
+\begin{figure}[tbp] \centering \includegraphics[width=1.0\textwidth]{./figs/mesh2000_replacement.png}
 \caption{2000 node mesh} \label{fig:SCMesh}
@@
 
-\begin{figure}[tbp] \centering \includegraphics[width=0.7\textwidth]{./figs/mesh1000000.png}
+\begin{figure}[tbp] \centering \includegraphics[width=1.0\textwidth]{./figs/mesh1000000_replacement.png}
 \caption{1000000 node mesh} \label{fig:MCMesh}

### 2014-04-19 jacobian_calc_CMAME.tex
made own drawing of truss

@@
 %
-\begin{figure}[tbp] \centering \includegraphics[width=0.7\textwidth]{./figs/truss.png}
+\begin{figure}[tbp] 
+\centering %%\includegraphics[width=0.7\textwidth]{./figs/truss.png}
+
+\begin{tikzpicture}
+\usetikzlibrary{shapes.geometric}
+\usetikzlibrary{decorations.pathmorphing}
+\usetikzlibrary{calc}
+\usetikzlibrary{patterns}
+
+\tikzstyle{wall}=[fill,pattern=north east lines,draw=none,minimum width=0.2cm,minimum height=5.1cm];
+\tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=6.1cm,minimum height=.2cm];
+
+\coordinate (bar_1) at (0,0);
+\coordinate (bar_1_support) at (0,-.5);
+
+\coordinate (bar_2) at (5,3);
+\coordinate (bar_2_pin) at (5.4,3.0);
+\coordinate (spring_1_support) at (5,-.5);
+\coordinate (spring_1_pin) at (5,0);
+
+\coordinate (rhm) at (5.4,1.5);
+\coordinate (rhml) at (4.5,1.5);
+\coordinate (mf) at (2.5, -0.8);
+\coordinate (lm) at (1.0, 0.0);
+\coordinate (mm) at (1.2, 0.0);
+\coordinate (rht) at (5.0, 4.0);
+\coordinate (rhtt) at (5.0, 4.2);
+
+\node (wall_slider) at (bar_2_pin) [circular sector, draw, rotate=180] {};
+\node (ground_triangle_1) at (bar_1_support) [isosceles triangle, draw, rotate=90] {};
+\node (ground_triangle_2) at (spring_1_support) [isosceles triangle, draw, rotate=90] {};
+
+\node (wall) at ($ (rhm) !.3cm!(10, 3) $)  [rectangle, draw, style=wall] {};
+\node (ground) at (mf)  [rectangle, draw, style=ground] {};
+
+\node (spring_label) at (rhml) [] {$k_s$};
+\node (load_label) at (rhtt) [] {$P$};
+\node (angle_label) at ($ (mm)!.4cm!(bar_2) $) [] {$\phi$};
+
+\path (bar_1) -- (bar_2) [draw, very thick];
+\path (bar_1) circle (.10) [fill=black];
+\path (bar_2) circle (.10) [fill=black];
+\path (spring_1_pin) circle (.10) [fill=black];
+\path [decorate, decoration={coil, segment length = 4, amplitude=4}] (bar_2) -- (spring_1_pin) [draw, thin, stroke=black];
+\path (rht) -- ($ (bar_2)!.2cm!(rht) $) [draw, -latex, ultra thick, color=black];
+\path (lm) arc(0:16:2) (lm) [draw];
+\path (bar_1) -- (mm) [draw, -latex];
+--cycle
+\end{tikzpicture}
+
 \caption{Beam and spring simulated in example program.} \label{fig:TrussSchematic}

### 2014-04-21 jacobian_calc_CMAME.tex
Switched to using latexdiff to track changes, makes things much cleaner

@@
 \usepackage{pgfplots}
+\pgfplotsset{compat=1.8}
 \usepackage{siunitx}
@@
 %Package for adding notes/comments/and tracking changes
-\usepackage[inline]{trackchanges}
-\addeditor{JTF}
-\addeditor{MDB}
+\usepackage[]{changes}
+\usepackage[]{todonotes}
+
 %% The amsthm package provides extended theorem environments
@@
 %% use the tnotetext command for the associated footnote;
-%% use the fnref command within \author or \address for footnotes;
+%% use the fnref command within \author or \addedress for footnotes;
 %% use the fntext command for the associated footnote;
@@
 \cortext[cor1]{Corresponding Author} 
-
 \author{Harry R. Millwater\corref{}}
@@
 %% \ead[url]{home page}
-%% \address{Address\fnref{label3}}
+%% \addedress{Address\fnref{label3}}
 %% \fntext[label3]{}
@@
 %% \author[label1,label2]{<author name>}
-%% \address[label1]{<address>}
-%% \address[label2]{<address>}
+%% \addedress[label1]{<address>}
+%% \addedress[label2]{<address>}
 
@@
     In order to maintain the quadratic convergence properties of Newton's method in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
-A goal of the study described in this paper was to establish the suitability of an underexplored method for numerical computation of tangent-stiffness operators, referred to as ``complex-step'', and compare the \note{Re: I.1} \remove{new} method with other techniques for numerical derivative calculation: automatic differentiation, forward finite difference, and central finite difference.
-The complex-step method was \note{Re: I.1} \remove{newly} implemented in a massively parallel computational peridynamics code for the purpose of this comparison.
+A goal of the study described in this paper was to establish the suitability of an underexplored method for numerical computation of tangent-stiffness operators, referred to as ``complex-step'', and compare the \todo[inline]{Re: I.1} method with other techniques for numerical derivative calculation: automatic differentiation, forward finite difference, and central finite difference.
+The complex-step method was \todo[inline]{Re: I.1} implemented in a massively parallel computational peridynamics code for the purpose of this comparison.
 The methods were compared through profiling of the code for accuracy, speed, efficiency, and parallel scalability.
@@
 
-In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent\footnote{\note{Re: II.1} In the context of plasticity thoery, the algorithmic consistency of the tangent-stiffness matrix refers to consistency with the integration of a plasticity model under the Khun-Tucker constraints or similar where the tangent-stiffness matrix is computed using the elastoplastic modulii that are determined from the integration \cite{simo1998}.} tangent-stiffness matrices.
+In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent\todo[disable]{Re: II.1}\footnote{In the context of plasticity thoery, the algorithmic consistency of the tangent-stiffness matrix refers to consistency with the integration of a plasticity model under the Khun-Tucker constraints or similar where the tangent-stiffness matrix is computed using the elastoplastic modulii that are determined from the integration \cite{simo1998}.} tangent-stiffness matrices.
 For an extremely small class of nonlinear material models, these consistent tangent-stiffness operators can be derived analytically; however, most often in practice, they are found through numerical approximation of derivatives.
 
-A goal of this study was to develop and evaluate a \note{Re: I.1} \change{new }{relatively unexplored}, accurate, and practical method for calculating tangent-stiffness matrices against established methods.
-This \note{Re: I.1} \remove{new} method is based on a complex number Taylor series expansion and referred to as CTSE or the ``complex-step'' method.
+A goal of this study was to develop and evaluate a \todo[disable]{Re: I.1} relatively unexplored, accurate, and practical method for calculating tangent-stiffness matrices against established methods.
+This  \todo[disable]{Re: I.1} method is based on a complex number Taylor series expansion and referred to as CTSE or the ``complex-step'' method.
 The practical accuracy of CTSE is determined by comparing its measured accuracy relative to the measured accuracy of popular finite differencing techniques used for computing tangent-stiffness matrices.
@@
 The scope of the application component of this study was limited to a single material model, an elastic peridynamic solid, implemented in the computational peridynamics code, \textit{Peridigm} \cite{peridigm}.
-Identifying a specific application provided a practical framework for implementing the \note{Re: I.1} \change{new }{complex-step} method and solving engineering problems to generate the data needed to compare the methods.
+Identifying a specific application provided a practical framework for implementing the  \todo[disable]{Re: I.1} complex-step method and solving engineering problems to generate the data needed to compare the methods.
 In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of forward-difference, central-difference, and automatic-differentiation methods needed for comparison to the complex-step method implementation, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
@@
 
-The aim of this paper is not only to \note{Re: I.1} \change{introduce the new}{study the} complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques useful for solving non-linear systems with Newton's method.
-After presenting background information on the underlying methods to be discussed, presented in this paper are: In Section~\ref{subsec:TS} a description of tangent-stiffness matrices, detailed directions for producing tangent-stiffness matrices with each of the methods identified above, and a description and justification of the \note{Re: I.1} \remove{new} complex-step method for calculating tangent-stiffness matrices, in Section~\ref{subsubsec:Impl} a description of implementing the complex-step method in the \emph{Peridigm} software, in Section~\ref{JGAM} a description of the quantities-of-interest used to rank the methods, in Section~\ref{sec:Res} a presentation and analysis of the results of the comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
+The aim of this paper is not only to \todo[disable]{Re: I.1} study the complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques useful for solving non-linear systems with Newton's method.
+After presenting background information on the underlying methods to be discussed, presented in this paper are: In Section~\ref{subsec:TS} a description of tangent-stiffness matrices, detailed directions for producing tangent-stiffness matrices with each of the methods identified above, and a description and justification of the  \todo[disable]{Re: I.1} complex-step method for calculating tangent-stiffness matrices, in Section~\ref{subsubsec:Impl} a description of implementing the complex-step method in the \emph{Peridigm} software, in Section~\ref{JGAM} a description of the quantities-of-interest used to rank the methods, in Section~\ref{sec:Res} a presentation and analysis of the results of the comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
 Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for \emph{Peridigm} C++ source-code and data necessary to reproduce the results presented here.
@@
 The purpose of the serial C++ library is to demonstrate in a less complex manner than in the \emph{Peridigm} code how one would use automatic-differentiation or complex-step to better encourage their use and discussion.
-\note{Re: II.5} \add{This open source resource is available at the corresponding author's website or at} \href{https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive}{https://github.com/utsa-idl/Complex\_Step\_Jacobian\_Project\_Archive}.
+\todo[disable]{Re: II.5}This open source resource is available at the corresponding author's website or at \href{https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive}{https://github.com/utsa-idl/Complex\_Step\_Jacobian\_Project\_Archive}.
 
@@
 %
-\note{Re: II.6} \add{For completeness, the forward- and central-difference formulas corresponding to equation}~(\ref{eqn:complexFirstDeriv}) \add{are} \cite[Chap. 4.1.3]{chapra2010}
+\todo[disable]{Re: II.6}For completeness, the forward- and central-difference formulas corresponding to equation~(\ref{eqn:complexFirstDeriv}) are \cite[Chap. 4.1.3]{chapra2010}
 %
@@
 %
-\add{The order of the truncation error in the forward-difference formula (equation}~(\ref{eqn:forwardFirstDeriv})\add{) indicates that for a given $h$ this method is less accurate an approximation than either central-difference or complex-step which have errors of $\mathcal{O}(h^2)$.}
-
-\note{Re: I.4} \add{As mentioned earlier, the CS method was originally introduced in a paper by Lyness and Moler} \cite{lyness1967numerical}\add{.
-A follow-on contribution by Lyness demonstrated a ``truncated'' approximate version of the original CS method to make it suitable for use on digital computers} \cite{lyness1968differentiation}\add{.
-After these original papers, little can be found in the literature regarding the method until Squire and Trapp showed through numerical experiments that CS offers superior accuracy to FD and CD} \cite{squire1998using} \add{in computing numerical derivatives.
-Papers by Newman et al.\ } \cite{newman1998} \add{and Anderson et al.\ } \cite{anderson2001sensitivity} \add{discuss computing sensitivities in a CFD analysis using CS and compares the results to sensitivities produced with central-difference derivatives.
+The order of the truncation error in the forward-difference formula (equation~(\ref{eqn:forwardFirstDeriv})) indicates that for a given $h$ this method is less accurate an approximation than either central-difference or complex-step which have errors of $\mathcal{O}(h^2)$.
+
+\todo[disable]{Re: I.4}As mentioned earlier, the CS method was originally introduced in a paper by Lyness and Moler \cite{lyness1967numerical}.
+A follow-on contribution by Lyness demonstrated a ``truncated'' approximate version of the original CS method to make it suitable for use on digital computers \cite{lyness1968differentiation}.
+After these original papers, little can be found in the literature regarding the method until Squire and Trapp showed through numerical experiments that CS offers superior accuracy to FD and CD \cite{squire1998using} in computing numerical derivatives.
+Papers by Newman et al.\ \cite{newman1998} and Anderson et al.\ \cite{anderson2001sensitivity} discuss computing sensitivities in a CFD analysis using CS and compares the results to sensitivities produced with central-difference derivatives.
 The work is also notable for identifying automatic-differentiation as a competing method to CS and suggesting that complex-step could be used to compute Jacobians of residuals for use in Newton-Krylov schemes for solving non-linear systems.
-Perez-Foguet et al.\ } \cite{perez2000numerical, perez2012numerical} \add{used the CS method to compute algorithmically consistent elastoplastic moduli in non-linear solid mechanics problems.
-A paper by Joaquim et al.\ } \cite{martins2000automated} \add{investigates using CS for computing aerodynamic sensitivities.
+Perez-Foguet et al.\ \cite{perez2000numerical, perez2012numerical} used the CS method to compute algorithmically consistent elastoplastic moduli in non-linear solid mechanics problems.
+A paper by Joaquim et al.\ \cite{martins2000automated} investigates using CS for computing aerodynamic sensitivities.
 Additionally, higher-order accuracy versions of CS are discussed, as is the task of implementation in most widely used general purpose programming languages.
 Discussed is the relative ease of implementing a CS scheme versus an automatic-differentiation based scheme for computing sensitivities in an established code.
-Martins et al.\ } \cite{martins2003complex}  \add{demonstrates that the CS method can be performed by using complex datatypes for the variables of functions comprising an analysis code.
+
+Martins et al.\ \cite{martins2003complex} demonstrates that the CS method can be performed by using complex datatypes for the variables of functions comprising an analysis code.
 The accuracy of this approach, with respect to the derivatives themselves, is compared to automatic-differentiation in a serial analysis code.
-Lai and Crassidis} \cite{lai2008extensions} \add{discuss first- and second-order CS methods for computing derivatives and choosing optimal step sizes.
-Al-Mohy and Higham} \cite{al2010complex} \add{investigate the use of CS in computing Fr\'{e}chet derivatives of matrix functions.
-Jin et al.\ } \cite{jin2010improved} \add{discuss using CS to compute sensitivities in a solid mechanics FEM analysis.
-Voorhees et al.\ } \cite{voorhees2011complex} \add{discusses the background of the CS method and frames it as a subset of Fourier Differentiation.
+Lai and Crassidis \cite{lai2008extensions} discuss first- and second-order CS methods for computing derivatives and choosing optimal step sizes.
+Al-Mohy and Higham \cite{al2010complex} investigate the use of CS in computing Fr\'{e}chet derivatives of matrix functions.
+Jin et al.\ \cite{jin2010improved} discuss using CS to compute sensitivities in a solid mechanics FEM analysis.
+Voorhees et al.\ \cite{voorhees2011complex} discusses the background of the CS method and frames it as a subset of Fourier Differentiation.
 CS and Fourier Differentiation are then applied to computing sensitivities within a solid mechanics finite element analysis.
 They implemented the complex-values as additional degrees of freedom in the finite element stiffness matrices using a real-valued matrix representation of the complex-numbers where required.
-In a follow-on work,  Millwater et al.\ } \cite{millwater2013application} \add{extended this concept, calling it ZFEM, and implemented the method into user defined element types for use with the popular ABAQUS} \cite{systemes2012abaqus} \add{FEA software package.
-Nishida et al.\ } \cite{nishida2013} \add{used the complex-step method to compute Jacobian matrices for implementation in a Levenberg-Marquardt optimization routine for design-of-experiments purposes.
-Finally, Abreu et al.\ } \cite{abreu2013generalization} \add{discusses computing derivatives with a version of the CS method which features the combined use of real and imaginary valued step components, as in Fourier differentiation.
+In a follow-on work,  Millwater et al.\  \cite{millwater2013application} extended this concept, calling it ZFEM, and implemented the method into user defined element types for use with the popular ABAQUS \cite{systemes2012abaqus} FEA software package.
+Nishida et al.\ \cite{nishida2013} used the complex-step method to compute Jacobian matrices for implementation in a Levenberg-Marquardt optimization routine for design-of-experiments purposes.
+Finally, Abreu et al.\ \cite{abreu2013generalization} discusses computing derivatives with a version of the CS method which features the combined use of real and imaginary valued step components, as in Fourier differentiation.
 They suggests that the modified CS method can achieve extended approximation accuracy up to fourth order.
@@
 To the author's knowledge it is the first application of the complex-step method to solve peridynamics mechanics problems, and the first implementation and comparison of the method in large-scale distributed memory parallel setting.
-}
 
@@
 
-\note[]{Re: I.5, next section is taken from the thesis.}
+\todo[disable][]{Re: I.5}
 \section{Simple example of CS, AD and FD for solution of nonlinear problems.}
@@
 
-\note[]{Re: I.2, John I think you would better handle expanding this section,
-I was going to essentially copy silling:psa.}\change[]{Briefly, peridynamics}Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
+\todo[disable]{Re: I.2} Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
 is a nonlocal reformulation of the partial-differential equations that provide
@@
 
-\note[]{Re: I.2. John, this was to be the skeleton for the new sections on peridynamics.}
 \subsubsection{Vector states for state based peridynamic theory}
@@
 field". The force vector state field maps vectors in the original coordinate system 
-measured from the peridynamic node at the center of the neighborhood or 'horizon' to
-neighboring peridynamic nodes that fall within the radius of the horizon. $\boldsymbol{b}$ denotes a body force density. Equation~\ref{eqn:PDMotion} may be simplified based on the assumption that
-$\boldsymbol{\underline{T}}$ is Riemann integrable and that integral is uniformly
+measured from the peridynamic node at the center of the neighborhood or `horizon' to
+neighboring peridynamic nodes that fall within the radius of the horizon.
+$\boldsymbol{b}$ denotes a body force density.
+Equation~(\ref{eqn:PDMotion}) may be simplified based on the assumption that $\boldsymbol{\underline{T}}$ is Riemann integrable and that integral is uniformly
 convergent \cite{silling:psa} as,
@@
 
-\note[]{Re: I.2} \add[]{
+\todo[disable]{Re: I.2}
 \subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simulation}
@@
 deformation state for all nodes to be computed.
-}
 
@@
 thought of as distributed memory parallel cousins to standard C++
-\change[]{STL}{Standard Template Library} vectors, are hard coded to be
+Standard Template Library vectors, are hard coded to be
 of type {\tt double} and can not be simply declared as having a complex data
-type. \note[]{Re: I.3}\add[]{For the reader's convenience,
-    \emph{Epetra} vectors and STL vectors are what are known as 'container
-    classes', whose purpose in object oriented computer programming is to
+type. \todo[disable]{Re: I.3}For the reader's convenience, \emph{Epetra} vectors and STL vectors are what are known as 'container classes', whose purpose in object oriented computer programming is to
     automate the creation, accessing, modification and destruction of the
@@
 tightly integrated with \emph{Peridigm's} parallel communication strategy and
-linear algebra code.} Program methods that were used in the course of applying
+linear algebra code. Program methods that were used in the course of applying
 the FD method were copied, renamed and re-written to follow the CS formula.
@@
 scoped, and dynamically casting persistent memory variables to complex data
-types where necessary. \note[]{Re: I.3}\add[]{For the reader's
+types where necessary. \todo[disable]{Re: I.3}For the reader's
 convenience, 'locally scoped' in the C++ programming language refers to
@@
 complex arithmetic, where a mapping called 'type casting' allows the
-translation of data between different types.} No attempt was made to ``tune''
+translation of data between different types. No attempt was made to ``tune''
 the CS code in any way for performance. The algorithm  was implemented in such
@@
 available in \emph{Trilinos} that are capable of taking an explicit
-instantiation of any data type including {\tt complex}. \note[]{Re:
-I.3}\add[]{For the purposes of this work, an explicit instantiation
+instantiation of any data type including {\tt complex}. \todo[disable][]{Re:
+I.3}For the purposes of this work, an explicit instantiation
 refers to requesting the creation of an object, or instance, of a templated
 class or container class that holds data which has the property of
-following desired mathematic or logic rules.} The version of \emph{Peridigm}
+following desired mathematic or logic rules. The version of \emph{Peridigm}
 modified to implement CS can be found in the paper repository.
@@
 many advanced techniques in C++ such as multiple virtual inheritance, pointer
-arithmetic, and distributed data structures, \note[]{Re:
-I.3}\add[]{for example \emph{Epetra} vectors, were used in the basic
+arithmetic, and distributed data structures, \todo[disable]{Re: I.3}for example \emph{Epetra} vectors, were used in the basic
     code and by necessity in the modified code. These concepts are not vital to
 understanding the complex step method itself, but may be necessary to understand complex
-step in \emph{Peridigm}.} To see CS and other methods in a more narrow context,
+step in \emph{Peridigm}. To see CS and other methods in a more narrow context,
 it is advised that readers also take a look at the simple examples provided on
@@
 
-A series of nine single compute core (i.e. serial) test problems were run with
+A series of nine single compute core (i.e.\ serial) test problems were run with
 the discretization size being refined with every test in the series. The aim of
@@
 terms of accuracy and speed. This series would comprise the single core runs. 
-\note[]{Re: II.16, specific test parameters were asked for}
+\todo[disable]{Re: II.16}
 The specific parameters of these tests can be found in \emph{Peridigm} input
@@
 
-\note[]{Re: I.3}\add[]{For the purposes of this work message passing is what
+\todo[disable]{Re: I.3}For the purposes of this work message passing is what
 synchronizes multiple computers that each are working on a separate section
@@
 simulation a multitude of times and take the average runtime. The larger
-the unpartitioned domain, the worse the discrepancy should be.}
+the unpartitioned domain, the worse the discrepancy should be.
 
@@
 heuristically derived based on computational node spacing using the method explained in \cite[pp. 90]{ref-Adaggio}.  
-\add[]{To summarize, a spacing of $1.E-6$ times grid spacing is used to provide acceptable accuracy but prevent roundoff error, 
-as is not a user determined setting in \emph{Peridigm}.} One of the beauties
+To summarize, a spacing of $1.E-6$ times grid spacing is used to provide acceptable accuracy but prevent roundoff error, 
+as is not a user determined setting in \emph{Peridigm}. One of the beauties
 of the CS method is that since dependence on $h$ decreases for small $h$, which
@@
 %
-\note[]{Re: II.22, relative accuracy figures were asked for, these were taken from the archive data}
-It can be seen in Figs.~\ref{fig:serial_accuracy} and ~\ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods.
+\todo[disable][]{Re: II.22}
+It can be seen in Figs.~\ref{fig:serial_accuracy} and \ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods.
 Those figures are depicting absolute error.
@@
 
-\note[]{Re: I.5, everything after this line is changed}
+\todo[disable][]{Re: I.5}
+
 \section{Convergence and displacement accuracy studies} 
@@
 
-\note[]{Re: I.5, this and other material was added in from the thesis to discuss accuracy}
-\note[]{Re: I.5, the entire conclusion section was replaced with the thesis version}
+\todo[disable]{Re: I.5}
 \section{Conclusions}
@@
 require compiling a version of the AD library being used that were compatible
-with the accelerator or GPU being used. \add[]{A working definition of byte-copyable
+with the accelerator or GPU being used. A working definition of byte-copyable
 is that the collection of data to transfer is made of identically sized pieces.
@@
 installed in \emph{Stampede}, where the alternative mode requires "more
-extensive" effort} \cite{intel_byte_copyable}. What this means is that if
+extensive" effort \cite{intel_byte_copyable}. What this means is that if
 \emph{Peridigm} or any simulation software were to be reformulated to take
@@
 dependence on an external AD library.  From this perspective, the method may
-still be an attractive method for TS matrix evaluation. \add[]{However,
+still be an attractive method for TS matrix evaluation. However,
 at least for the models producing the results shown here, each of the
@@
 Newton's convergence conditions or the approximate case conditions are
-violated.}
+violated.
 

### 2014-04-24 jacobian_calc_CMAME.tex
More edits

@@
 \usepackage{hyperref}
+\usepackage{subcaption}
+\usepackage[T1]{fontenc}
+\usepackage{upquote}
 %% The amssymb package provides various useful mathematical symbols
@@
 %Package for adding notes/comments/and tracking changes
-\usepackage[]{changes}
-\usepackage[]{todonotes}
+\usepackage[color=red!40, textsize=scriptsize]{todonotes}
 
@@
     In order to maintain the quadratic convergence properties of Newton's method in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent tangent-stiffness matrices.
-A goal of the study described in this paper was to establish the suitability of an underexplored method for numerical computation of tangent-stiffness operators, referred to as ``complex-step'', and compare the \todo[inline]{Re: I.1} method with other techniques for numerical derivative calculation: automatic differentiation, forward finite difference, and central finite difference.
-The complex-step method was \todo[inline]{Re: I.1} implemented in a massively parallel computational peridynamics code for the purpose of this comparison.
+A goal of the study described in this paper was to establish the suitability of an underexplored method for numerical computation of tangent-stiffness operators, referred to as ``complex-step'', and compare the  method with other techniques for numerical derivative calculation: automatic differentiation, forward finite difference, and central finite difference.
+The complex-step method was implemented in a massively parallel computational peridynamics code for the purpose of this comparison.
 The methods were compared through profiling of the code for accuracy, speed, efficiency, and parallel scalability.
@@
 
-In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent\todo[disable]{Re: II.1}\footnote{In the context of plasticity thoery, the algorithmic consistency of the tangent-stiffness matrix refers to consistency with the integration of a plasticity model under the Khun-Tucker constraints or similar where the tangent-stiffness matrix is computed using the elastoplastic modulii that are determined from the integration \cite{simo1998}.} tangent-stiffness matrices.
+In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent\todo[]{Re: II.1}\footnote{In the context of plasticity thoery, the algorithmic consistency of the tangent-stiffness matrix refers to consistency with the integration of a plasticity model under the Khun-Tucker constraints or similar where the tangent-stiffness matrix is computed using the elastoplastic modulii that are determined from the integration \cite{simo1998}.} tangent-stiffness matrices.
 For an extremely small class of nonlinear material models, these consistent tangent-stiffness operators can be derived analytically; however, most often in practice, they are found through numerical approximation of derivatives.
 
-A goal of this study was to develop and evaluate a \todo[disable]{Re: I.1} relatively unexplored, accurate, and practical method for calculating tangent-stiffness matrices against established methods.
-This  \todo[disable]{Re: I.1} method is based on a complex number Taylor series expansion and referred to as CTSE or the ``complex-step'' method.
+A goal of this study was to develop and evaluate a \todo[]{Re: I.1} relatively unexplored, accurate, and practical method for calculating tangent-stiffness matrices against established methods.
+This  \todo[]{Re: I.1} method is based on a complex number Taylor series expansion and referred to as CTSE or the ``complex-step'' method.
 The practical accuracy of CTSE is determined by comparing its measured accuracy relative to the measured accuracy of popular finite differencing techniques used for computing tangent-stiffness matrices.
@@
 The scope of the application component of this study was limited to a single material model, an elastic peridynamic solid, implemented in the computational peridynamics code, \textit{Peridigm} \cite{peridigm}.
-Identifying a specific application provided a practical framework for implementing the  \todo[disable]{Re: I.1} complex-step method and solving engineering problems to generate the data needed to compare the methods.
-In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of forward-difference, central-difference, and automatic-differentiation methods needed for comparison to the complex-step method implementation, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
+Identifying a specific application provided a practical framework for implementing the  \todo[]{Re: I.1} complex-step method and solving engineering problems to generate the data needed to compare the methods.
+In particular, \emph{Peridigm} was chosen because it combined several helpful characteristics: the utilization of Newton's method and tangent-stiffness matrices for solving nonlinear quasi-static problems, prior inclusion of forward difference, central difference, and automatic-differentiation methods needed for comparison to the complex-step method implementation, and being an agile-components code that makes use of distributed computing data structures via Trilinos \cite{trilinos} for efficient parallelization on large clusters.
 Agile-components is a software development term meaning to structure a software project for continuous development to respond to evolving specifications by making much use of techniques from object-oriented programming and reusing proven pre-existing code to reduce development time and increase reliability.
@@
 
-The aim of this paper is not only to \todo[disable]{Re: I.1} study the complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques useful for solving non-linear systems with Newton's method.
-After presenting background information on the underlying methods to be discussed, presented in this paper are: In Section~\ref{subsec:TS} a description of tangent-stiffness matrices, detailed directions for producing tangent-stiffness matrices with each of the methods identified above, and a description and justification of the  \todo[disable]{Re: I.1} complex-step method for calculating tangent-stiffness matrices, in Section~\ref{subsubsec:Impl} a description of implementing the complex-step method in the \emph{Peridigm} software, in Section~\ref{JGAM} a description of the quantities-of-interest used to rank the methods, in Section~\ref{sec:Res} a presentation and analysis of the results of the comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
+The aim of this paper is not only to \todo[]{Re: I.1} study the complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques useful for solving non-linear systems with Newton's method.
+After presenting background information on the underlying methods to be discussed, presented in this paper are: In Section~\ref{subsec:TS} a description of tangent-stiffness matrices, detailed directions for producing tangent-stiffness matrices with each of the methods identified above, and a description and justification of the  \todo[]{Re: I.1} complex-step method for calculating tangent-stiffness matrices, in Section~\ref{subsubsec:Impl} a description of implementing the complex-step method in the \emph{Peridigm} software, in Section~\ref{JGAM} a description of the quantities-of-interest used to rank the methods, in Section~\ref{sec:Res} a presentation and analysis of the results of the comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
 Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for \emph{Peridigm} C++ source-code and data necessary to reproduce the results presented here.
@@
 The purpose of the serial C++ library is to demonstrate in a less complex manner than in the \emph{Peridigm} code how one would use automatic-differentiation or complex-step to better encourage their use and discussion.
-\todo[disable]{Re: II.5}This open source resource is available at the corresponding author's website or at \href{https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive}{https://github.com/utsa-idl/Complex\_Step\_Jacobian\_Project\_Archive}.
+\todo[]{Re: II.5}This open source resource is available at the corresponding author's website or at \href{https://github.com/utsa-idl/Complex_Step_Jacobian_Project_Archive}{https://github.com/utsa-idl/Complex\_Step\_Jacobian\_Project\_Archive}.
 
@@
 %
-\todo[disable]{Re: II.6}For completeness, the forward- and central-difference formulas corresponding to equation~(\ref{eqn:complexFirstDeriv}) are \cite[Chap. 4.1.3]{chapra2010}
+\todo[]{Re: II.6}For completeness, the forward and central difference formulas corresponding to equation~(\ref{eqn:complexFirstDeriv}) are \cite[Chap. 4.1.3]{chapra2010}
 %
@@
 %
-The order of the truncation error in the forward-difference formula (equation~(\ref{eqn:forwardFirstDeriv})) indicates that for a given $h$ this method is less accurate an approximation than either central-difference or complex-step which have errors of $\mathcal{O}(h^2)$.
-
-\todo[disable]{Re: I.4}As mentioned earlier, the CS method was originally introduced in a paper by Lyness and Moler \cite{lyness1967numerical}.
+The order of the truncation error in the forward difference formula (equation~(\ref{eqn:forwardFirstDeriv})) indicates that for a given $h$ this method is less accurate an approximation than either central difference or complex-step which have errors of $\mathcal{O}(h^2)$.
+
+\todo[]{Re: I.4}As mentioned earlier, the CS method was originally introduced in a paper by Lyness and Moler \cite{lyness1967numerical}.
 A follow-on contribution by Lyness demonstrated a ``truncated'' approximate version of the original CS method to make it suitable for use on digital computers \cite{lyness1968differentiation}.
 After these original papers, little can be found in the literature regarding the method until Squire and Trapp showed through numerical experiments that CS offers superior accuracy to FD and CD \cite{squire1998using} in computing numerical derivatives.
-Papers by Newman et al.\ \cite{newman1998} and Anderson et al.\ \cite{anderson2001sensitivity} discuss computing sensitivities in a CFD analysis using CS and compares the results to sensitivities produced with central-difference derivatives.
+Papers by Newman et al.\ \cite{newman1998} and Anderson et al.\ \cite{anderson2001sensitivity} discuss computing sensitivities in a CFD analysis using CS and compares the results to sensitivities produced with central difference derivatives.
 The work is also notable for identifying automatic-differentiation as a competing method to CS and suggesting that complex-step could be used to compute Jacobians of residuals for use in Newton-Krylov schemes for solving non-linear systems.
@@
 
-\todo[disable][]{Re: I.5}
-\section{Simple example of CS, AD and FD for solution of nonlinear problems.}
+\section{\todo[]{Re: I.5}Simple example of CS, AD and FD for solution of nonlinear problems.}
 \label{sec:WorkingExamples}
@@
 All the source code to reproduce the results described here are available in the associated paper repository.
-
-\subsection{Common code} 
-\label{subsec:CommonCode}
-
-The source code for the header-only library is entitled \emph{NewtonRaphson.hpp}.
-Note that the library is an interface to functions and data structures made available in the \emph{Trilinos} library, mentioned earlier, such that it is required in order for the examples described here to work.
-\emph{Trilinos} is freely available for download as of this writing at: \emph{trilinos.sandia.gov}.
-Generally the header-only library works by taking in the users initial guessed equilibrium solution, target dependent variable values, system constants, and a function pointer to a model evaluation function.
-Solver parameters including an update scaling factor, tolerance, maximum iterations and probe length are chosen after the solver is instantiated as a new object.
-The behavior of the solver is specialized through a coding feature called inheritance to perform each of CS, AD and FD.
-But, in common, each version of the solver works by computing the system Jacobian, solving a linear system as in that step of the Newton Raphson method, computing the residual and checking for convergence or overrun.
-Here the linear solution method is LU decomposition, while the chief convergence criterion is change in residual between iterations.
-Additionally the problem can be partially respecified to allow for running a series of load steps having progressing target values, which is a typical mode also seen in \emph{Peridigm}.
-Applying boundary conditions incrementally in load steps assists convergence.
-
-\subsection{Second example and related studies} 
-\label{subsec:Validation}
-The second program,a \emph{NRComparison.cpp}, follows a case study that appears in \cite{rezaiee2010dynamic}.
-The example program departs from the cited literature by applying each of CS, AD and FD and verifying them against an analytical model of a nonlinear mechanical system involving a connected beam and spring.
-Fig.
-\ref{fig:TrussSchematic} is a schematic of the beam and spring system.
+This source describes the implemenation and interface to the \emph{Trilinos Secado library} for automatic differentiation.
+
+The example problem is a single degree-of-freedom nonlinear mechanical system involving a connected truss member and spring as illustrated in Fig. \ref{fig:TrussSchematic}. 
 %
@@
 \usetikzlibrary{patterns}
-
 \tikzstyle{wall}=[fill,pattern=north east lines,draw=none,minimum width=0.2cm,minimum height=5.1cm];
 \tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=6.1cm,minimum height=.2cm];
-
 \coordinate (bar_1) at (0,0);
 \coordinate (bar_1_support) at (0,-.5);
-
 \coordinate (bar_2) at (5,3);
@@
 \coordinate (spring_1_pin) at (5,0);
-
 \coordinate (rhm) at (5.4,1.5);
@@
 \coordinate (rhtt) at (5.0, 4.2);
-
 \node (wall_slider) at (bar_2_pin) [circular sector, draw, rotate=180] {};
@@
 \node (ground_triangle_2) at (spring_1_support) [isosceles triangle, draw, rotate=90] {};
-
 \node (wall) at ($ (rhm) !.3cm!(10, 3) $)  [rectangle, draw, style=wall] {};
 \node (ground) at (mf)  [rectangle, draw, style=ground] {};
-
 \node (spring_label) at (rhml) [] {$k_s$};
@@
 \node (angle_label) at ($ (mm)!.4cm!(bar_2) $) [] {$\phi$};
-
 \path (bar_1) -- (bar_2) [draw, very thick];
@@
 \end{tikzpicture}
-
-\caption{Beam and spring simulated in example program.} \label{fig:TrussSchematic}
+\caption{Beam and spring simulated in example program.} 
+\label{fig:TrussSchematic}
 \end{figure}
 %
-The equations describing this system and results for verification were obtained
-in \cite{rezaiee2010dynamic}. They include a force function, Eqn.
-\ref{eqn:TrussForce}, and an analytical force derivative or stiffness, Eqn.
-\ref{eqn:TrussStiffness}.
-
+For this system, the equalibrium reaction force associated with a displacement $D$ of the spring is shown in equation~(\ref{eqn:TrussForce}) 
+%
 \begin{equation} 
     \label{eqn:TrussForce}
-    f(D) = .5AE(\cos^{2}\phi)(\frac{D}{L_{0}})^{2}[\frac{D}{L_{0}}\cos^{2}\phi - 3\sin\phi] + k_{s}D + (AE\frac{D}{L_{0}})\sin^{2}\phi,
+    f(D) = \frac 1 2 AE(\cos^{2}\phi)(\frac{D}{L_{0}})^{2}[\frac{D}{L_{0}}\cos^{2}\phi - 3\sin\phi] + k_{s}D + (AE\frac{D}{L_{0}})\sin^{2}\phi,
 \end{equation} 
-
+%
+and the analytical tangent-stiffness is given by equation~(\ref{eqn:TrussStiffness})
+%
 \begin{equation} 
     \label{eqn:TrussStiffness}
-    S_{T}(D) = 1.5AE(\cos^{2}\phi)[\frac{D}{L_{0}}\cos^{2}\phi - 2sin\phi](\frac{D}{L_{0}^{2}}) + k_{s} + \frac{AE\sin^{2}\phi}{L_{0}}.
+    S_{T}(D) = \frac 3 2 AE(\cos^{2}\phi)[\frac{D}{L_{0}}\cos^{2}\phi - 2sin\phi](\frac{D}{L_{0}^{2}}) + k_{s} + \frac{AE\sin^{2}\phi}{L_{0}}.
 \end{equation}
-
-Where $\phi$ is a constant, beam angle, $AE$ is rigidity, $D$ is displacement,
-$L_{0}$ is original beam length, and $k_{s}$ is the Hooke's constant for the
-coil spring attached to the beam. For a verification of each of the methods as
-well as a convergence rate study, this truss and spring model was analyzed in
-the second program. $f(D)$ is restorative force to balance downward force
-loading at the connection point of the beam and the spring, parallel to the
-wall. As in the reference, equilibrium solutions for 12 incremental load steps
-of 4.448 Newtons were sought. Differing from the reference, Newton's method was
-used rather than 'Dynamic Relaxation'. Solver parameters common to all tests
-were a limit of 100 iterations for convergence, a residual change tolerance of
-$1\times{10^{-9}} \si{\pico\meter}$ and an update multiplier of 1.0. Data
-collected included number of iterations to converge summed over all load-steps
-and final displacement, D, after the final load-step. Additionally
-finite difference step-size was varied in order to measure the impact on each
-of the Jacobian calculation methods. Naturally, varying step size does not
-affect AD or the analytically derived Jacobian since those are not functions
-of step size. The accuracy results for
-verifying the Jacobian calculation methods are summarized in Table
-\ref{tab:ConvergenceStudy}.
-
-\begin{table*}[!ht]    
+%
+where $\phi$ is a constant truss angle, $AE$ is the axial rigidity of the truss, $L_{0}$ is original truss member length, and $k_{s}$ is the Hooke's constant for the spring attached to the truss member.
+For a demonstration and verification of each of the methods as well as a convergence rate study, this truss and spring model was analyzed with each of the methods and compared with the analytic solution.
+As in the reference \cite{rezaiee2010dynamic}, equilibrium solutions for 12 incremental load steps of $\Delta P=\SI{4.448}{\newton}$ were found.
+Solver parameters common to all tests were a maximum limit of 100 iterations for convergence, a residual change tolerance of $\SI{1e-9}{\meter}$.
+Data collected included number of iterations to converge summed over all load-steps and final displacement $D$, after the final load-step.
+Additionally finite difference step-size was varied in order to measure the impact on each of the approximate tangent-stiffness calculation methods.
+Naturally, varying step size does not affect AD or the analytically derived tangent-stiffness since those are not functions of step size.
+The accuracy results for verifying the tangent-stiffness calculation methods are summarized in Table~\ref{tab:ConvergenceStudy}.
+%
+\begin{table}[tbp]    
   \centering
@@
     \end{tabular}
-\end{table*}
-
-Table \ref{tab:Verification} shows that each of the methods solved the problem
-as accurately as one another within the precision displayed on the table. The
-exception to this is that the finite difference method failed due to an
-arithmetic error in association with extremely small step size, suggesting
-subtractive cancellation. However, since there is no accuracy reason for
-choosing an extremely small step size, each of the methods can be used to
-produce acceptable solutions to the nonlinear truss system. Table
-\ref{tab:ConvergenceStudy} shows the total number of iterations taken over all
-load-steps for each of the methods.
-
-\begin{table*}[!ht]    
+\end{table}
+%
+Table.~\ref{tab:Verification} shows that each of the methods solved the problem as accurately as one another within the precision reported.
+The exception to this is that the finite difference method failed due to an arithmetic error in association with extremely small step size, suggesting subtractive cancellation.
+However, since there is no accuracy reason for choosing an extremely small step size, each of the methods can be used to produce acceptable solutions to the nonlinear truss system.
+Table~\ref{tab:ConvergenceStudy} shows the total number of iterations taken over all load-steps for each of the methods.
+%
+\begin{table}[tbp]    
   \centering
@@
     \end{tabular}
-\end{table*}
-
-Table \ref{tab:ConvergenceStudy} indicates that when using AD or the 'actual'
-analytically derived Jacobian, a minimum number of iterations is achieved
-corresponding to a maximum convergence rate. CS and FD are similarly affected
-by a grossly large step-size, while subtractive cancellation invalidates the
-results for FD for very small step-sizes. Inaccuracy for large step-sizes is is
-predicted by the FD and CS formulas, which require a small
-step-size to allow the neglect of higher order 12 terms. The results of this
-test show that step-size selection is not unimportant when using FD, but when
-properly configured any of the methods may return the same maximum convergence
-rate for this problem.  Overall conclusions regarding the methods that the accuracy of the
-Jacobian should not be confused with the accuracy of the equilibrium
-solution, since each of the methods discussed here may be equally capable given
-enough iterations. Numerical measurements of accuracy besides number of iterations taken
-are not possible here because there is not basis for comparison, as the methods are solving
-the problem separately and producing difference Newton iterates. Additionally, while the inaccuracy of the Jacobian may
-strongly affect convergence rates, this relationship was not precisely quantified here.
+\end{table}
+%
+Table~\ref{tab:ConvergenceStudy} indicates that when using AD or the `actual' analytically derived tangent-stiffness, a minimum number of iterations is achieved corresponding to a maximum convergence rate.
+CS and FD are similarly affected by a grossly large step-size, while subtractive cancellation invalidates the results for FD for very small step-sizes.
+Inaccuracy for large step-sizes is as predicted by the FD and CS formulas, which require a small step-size to allow the neglect of higher order terms.
+The results of this test show that step-size selection is not unimportant when using FD, but when properly configured any of the methods may return the same maximum convergence rate for this problem.
+Overall conclusions regarding the methods that the accuracy of the tangent-stiffness should not be confused with the accuracy of the equilibrium solution, since each of the methods discussed here may be equally capable given enough iterations.
+Numerical measurements of accuracy besides number of iterations taken are not possible here because there is not basis for comparison, as the methods are solving the problem separately and producing different local Newton iterates.
+Additionally, while the inaccuracy of the tangent-stiffness may strongly affect convergence rates, this relationship was not precisely quantified here.
 
@@
 %
-\subsection{\emph{Peridigm} and peridynamics}
+\subsection{Peridigm and peridynamics}
 \label{subsec:PaP}
 %
-Comparisons of the different tangent-stiffness matrix calculation techniques were
-carried out through code-development and instrumentation of the
-computational peridynamics code \emph{Peridigm} \cite{peridigm}.
-\emph{Peridigm} is distributed by Sandia National Laboratories as its primary
-open-source computational peridynamics code. It is a massively parallel
-simulation code for implicit and explicit multi-physics simulations primarily
-used for solid mechanics and material failure. \emph{Peridigm} is a C++ code
-utilizing agile software components from Sandia's  \emph{Trilinos} project
-\cite{trilinos}. It's important to note that there is no specialization of the
-tangent-stiffness matrix calculation methods described previously to this particular
-code or more generally to a computational peridynamics approach and the
-analysis carried out in this study should provide insight into the speed and
-accuracy of these methods when implemented into other computational mechanics
-analysis tools as well.  \emph{Peridigm} was chosen primarily because the FD,
-CD, and AD methods for tangent-stiffness matrix calculation where already implemented
-in the code, leaving only the CS method for development. Also, because
-peridynamics is a nonlocal theory, the tangent-stiffness matrices have a much
-higher bandwidth (i.e. less sparsity) than traditional finite-element
-tangent-stiffness matrix computations and the time required to construct the
-tangent-stiffness matrix is a majority of the total computation time in any
-quasi-static or implicit dynamics analysis; therefore, the \emph{Peridigm}
-development team has an interest in profiling the performance of the
-tangent-stiffness matrix computation methods.
-
-\todo[disable]{Re: I.2} Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic}
-is a nonlocal reformulation of the partial-differential equations that provide
-the statement of momentum balance in classical continuum mechanics. Its primary
-goal is to avoid the use of spatial derivatives in the balance equations or
-constitutive laws such that discontinuous displacements, i.e. cracks, are
-mathematically consistent with the governing equations. It has demonstrated
-great promise in modeling problems with pervasive failure
-\cite{littlewood2010}, interesting phenomenon such as dynamic crack branching
-\cite{ha2010sod}, and is being extended to multiphyics phenomena
-\cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale
-coupling to molecular simulations \cite{seleson2009peridynamics,seleson2014atom}.
-
-The type of material model used in this study is referred to as a linear
-peridynamic solid, having what is called a state based constitutive model. For
-the convenience of the reader, vector states will be introduced. Then the state
-based peridynamic equation of motion will be shown, followed by a description
-of the linear peridynamic solid model used in this work. After the peridynamic
-equation of motion and a constitutive model have been discussed, a definition of
-the tangent stiffness matrix is given. The following discussion is largely
-taken from \cite{silling:psa} but reproduced here for clarity.
-
-\subsubsection{Vector states for state based peridynamic theory}
-Because the vector state, which is fair to say an uncommon notation, is used in the literature discussing state based peridynamic theory
-it will be enumerated here for the sake of continuity. 
-%
-Paraphrasing, let a spherical region, centered about a
-location in threespace, called a horizon or neighborhood with a certain radius
-be defined. A vector state of order $m$ is a mapping of a vector contained
-within the horizon to a tensor of order $m$ \cite{silling:psa}. Vector states
-have have their own mathematical properties and definitions for sums, composition,
-point product, reference state, identity scalar state, null reference state, null scalar state,
-dot product, magnitude state, unit state, and direction state which are detailed in \cite{silling:psa}. 
-It is important to think of vector states as non-linear, not necessarily continuous maps of vectors to real Euclidian space, where second order tensors are linear maps of three dimensional vectors to threespace.
-%
-One important tool mechanics gains from the concept of vector states is a way
-to represent complicated deformations within a body in the form of a state
-function. Whereas a symmetric second order tensor maps a spherical region in a
-body in the reference configuration to an ellipsoid in the deformed
-configuration, a vector state can map a spherical region in a body into a
-complicated shape \cite{silling:psa}.
-
-(PLACE PERIDYNAMIC POTATO DIAGRAM HERE)
-
-%
-The properties and definitions of vector states allow interaction with vectors
-and second order tensors to be defined which includes a tensor vector product
-in terms of a state function called an expansion, the tensor product of two states in terms of the
-volume integral over a horizon of the dyadic product of those states times a
-scalar function, the shape tensor as the tensor product of the reference vector
-state with itself, and a tensor reduced form of a state as the conventionally
-defined tensor product of the state tensor product of the state and the
-reference state with the inverse of the shape tensor, and the inheritance of
-the properties of a proper orthogonal second order tensor in its state
-expansion. These ideas are detailed in \cite{silling:psa}. 
-%
-Because of its importance in this work, the definition of a derivative of a vector state
-will be given a more complete explanation.
-
-\subsubsection{The state based peridynamic equation of motion}
-
-In \cite{silling:psa}, the primary reference for state based peridynamics, the
-equation of motion for state based peridynamic theory is derived. No
-development of the theory is claimed; the equation of motion is reproduced here for the reader's convenience.
-%
-The general equation is written as it appears in \cite{silling:psa}:
+Comparisons of the different tangent-stiffness matrix calculation techniques were carried out through code-development and instrumentation of the computational peridynamics code \emph{Peridigm} \cite{peridigm}.
+\emph{Peridigm} is distributed by Sandia National Laboratories as its primary open-source computational peridynamics code.
+It is a massively parallel simulation code for implicit and explicit multi-physics simulations primarily used for solid mechanics and material failure.
+\emph{Peridigm} is a C++ code utilizing agile software components from Sandia's \emph{Trilinos} project \cite{trilinos}.
+It's important to note that there is no specialization of the tangent-stiffness matrix calculation methods described previously to this particular code or more generally to a computational peridynamics approach and the analysis carried out in this study should provide insight into the speed and accuracy of these methods when implemented into other computational mechanics analysis tools as well.
+\emph{Peridigm} was chosen primarily because the FD, CD, and AD methods for tangent-stiffness matrix calculation where already implemented in the code, leaving only the CS method for development.
+Also, because peridynamics is a nonlocal theory, the tangent-stiffness matrices have a much higher bandwidth (i.e.
+less sparsity) than traditional finite element tangent-stiffness matrix computations and the time required to construct the tangent-stiffness matrix is a majority of the total computation time in any quasi-static or implicit dynamics analysis; therefore, the \emph{Peridigm} development team has an interest in profiling the performance of the tangent-stiffness matrix computation methods.
+
+\todo[]{Re: I.2} Peridynamics \cite{silling2000ret,silling:psa,silling2010peridynamic} is a nonlocal reformulation of the partial-differential equations that provide the statement of momentum balance in classical continuum mechanics.
+Its primary goal is to avoid the use of spatial derivatives in the balance equations or constitutive laws such that discontinuous displacements, i.e.
+cracks, are mathematically consistent with the governing equations.
+It has demonstrated great promise in modeling problems with pervasive failure \cite{littlewood2010}, interesting phenomenon such as dynamic crack branching \cite{ha2010sod}, and is being extended to multiphyics phenomena \cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale coupling to molecular simulations \cite{seleson2009peridynamics,seleson2014atom,rahman2013b}.
+The material model used in this study was a linear peridynamic solid which replicates linear elasticity in the \emph{local limit}, i.e.\ when the non-local length scale vanishes.
+While \emph{linear} may appear in the name of the material model, it is geometrically non-linear under large deformations and requires a nonlinear solver to converge to equalibrium conditions in these situations.
+The peridynamic equation of motion is shown in equation~(\ref{eqn:PDMotion})
+%
 \begin{equation} 
+    \rho(\mathbf{x})\mathbf{\ddot{u}(x)} = \int_{\mathcal{H_{\delta(\mathbf{x})}}} \mathbf{f(x,x'}) {\rm d}V_{\mathbf{x'}} + \mathbf{b(x)},
     \label{eqn:PDMotion}
-    \rho(\boldsymbol{x})\boldsymbol{\ddot{u}}(\boldsymbol{x}, t) = \int_{\mathcal{H}_{\boldsymbol{x}}} \{\boldsymbol{\underline{T}}[\boldsymbol{x}, t]\langle \boldsymbol{x}^{\prime} - \boldsymbol{x} \rangle - \boldsymbol{\underline{T}}[\boldsymbol{x}^{\prime}, t]\langle \boldsymbol{x} - \boldsymbol{x}^{\prime} \rangle \}dV_{\boldsymbol{x}^{\prime}} + \boldsymbol{b}(\boldsymbol{x}, t),
 \end{equation}
 
-where "$\mathcal{H}_{\boldsymbol{x}}$ is a spherical neighborhood" with a
-radius, centered at the original coordinates of a peridynamic node,
-$\boldsymbol{x}$, and $\boldsymbol{\underline{T}}$ is the "force vector state
-field". The force vector state field maps vectors in the original coordinate system 
-measured from the peridynamic node at the center of the neighborhood or `horizon' to
-neighboring peridynamic nodes that fall within the radius of the horizon.
-$\boldsymbol{b}$ denotes a body force density.
-Equation~(\ref{eqn:PDMotion}) may be simplified based on the assumption that $\boldsymbol{\underline{T}}$ is Riemann integrable and that integral is uniformly
-convergent \cite{silling:psa} as,
-
-\begin{equation} 
-    \label{eqn:PDMotion_simple}
-    \rho\boldsymbol{\ddot{u}} = \int_{\mathcal{H}_{\boldsymbol{x}}} \{\boldsymbol{\underline{T}}\langle \boldsymbol{x}^{\prime} - \boldsymbol{x} \rangle - \boldsymbol{\underline{T}}^{\prime}\langle \boldsymbol{x} - \boldsymbol{x}^{\prime} \rangle \}dV_{\boldsymbol{x}^{\prime}} + \boldsymbol{b},
-\end{equation}
-
-where,
-
-\begin{equation} 
-    \label{eqn:PDMotion_simple}
-    \boldsymbol{\underline{T}} = \boldsymbol{\underline{T}}[\boldsymbol{x}, t], 
-    \boldsymbol{\underline{T}}^{\prime} = \boldsymbol{\underline{T}}[\boldsymbol{x}^{\prime}, t]. 
-\end{equation}
-
-\subsubsection{The linear peridynamic solid}
-
-\todo[disable]{Re: I.2}
-\subsubsection{The tangent stiffness matrix in a quasi-static peridynamic simulation}
-The tangent-stiffness matrix is the Fr\'{e}chet derivative of the residual wrt the
-deformation state.  For the purposes of this work, the residual represents the unbalanced forces in the
-static peridynamic momentum equation. The internal force
-density is computed in the peridynamic momentum equation by computing the
-volume integral over a horizon of the Fr\'{e}chet derivative of the 
-strain energy density state wrt to the deformed configuration inner product with the deformation state, per node,
-for each of the three directions in the original reference frame.
-
-For the finite difference type methods in \emph{Peridigm}, the perturbation
-method is employed by probing the deformed configuration one degree of freedom
-at a time and so allowing the derivative of the force density state wrt the
-deformation state for all nodes to be computed.
+where, $\rho$ is the mass density, $\mathbf{\ddot{u}}$ is the second time derivative of displacement (i.e.\ acceleration), $\mathcal{H}_{\delta(\mathbf{x})}$ is a neighborhood containing remote material points, $\mathbf{x'}$, with characteristic length $\delta$, centered at $\mathbf{x}$.
+The integrand $\mathbf{f}$ contains the constitutive response of the material.
+Physically, $\delta$ is a material length-scale and computationally it, along with the discretization, will set the bandwidth of the tangent-stiffness matrix.
+$\mathbf{b}$ is a body force density.
+An explicit dependence on time $t$ has been supressed from the functional arguments, but is implied.
+The reader is refered to the seminal work on peridynamics \cite{silling:psa} for detailed derivations of the momentum equation and example consitutive models.
+
+\subsubsection{\todo[]{Re: I.2}The tangent stiffness matrix in a quasi-static peridynamic simulation}
+The tangent-stiffness matrix is the Fr\'{e}chet derivative of the residual with respect to the deformed positions of the discrete computational nodes.
+For the purposes of this work, the residual represents the unbalanced forces in the peridynamic equalibrium equation.
+For the finite difference type methods in \emph{Peridigm}, a perturbation method is employed by probing the deformed configuration one degree-of-freedom at a time and so allowing the derivative of the residual for all nodes to be computed.
+The resulting spatial rate of change in the residual for each of these independent probes becomes one column of the tangent-stiffness matrix.
 
@@
 %
-The CS method was implemented in \emph{Peridigm} in a manner that follows
-closely the implementation of the FD technique.  However, because of
-\emph{Peridigm}'s reliance on \emph{Trilinos} and specifically the
-\emph{Epetra} package, special steps had to be followed to allow the use of
-complex number data types.  This is because \emph{Epetra} vectors, which can be
-thought of as distributed memory parallel cousins to standard C++
-Standard Template Library vectors, are hard coded to be
-of type {\tt double} and can not be simply declared as having a complex data
-type. \todo[disable]{Re: I.3}For the reader's convenience, \emph{Epetra} vectors and STL vectors are what are known as 'container classes', whose purpose in object oriented computer programming is to
-    automate the creation, accessing, modification and destruction of the
-    underlying contained data. Templated container classes allow for user
-    choice in what properties that underlying data exhibits.  For example, in
-    this work, complex arithmetic cannot be performed on the data contained
-    within an \emph{Epetra} vector because that data, of type {\tt double} does
-    not have the property of following the rules of complex arithmetic. The
-    reason the non-templated \emph{Epetra} vector could not be dispensed with
-    for the purposes of this work is because the \emph{Epetra} vector was
-tightly integrated with \emph{Peridigm's} parallel communication strategy and
-linear algebra code. Program methods that were used in the course of applying
-the FD method were copied, renamed and re-written to follow the CS formula.
-Because of restrictions imposed by the container class being used, this
-involved changing the data type of intermediate variables which were locally
-scoped, and dynamically casting persistent memory variables to complex data
-types where necessary. \todo[disable]{Re: I.3}For the reader's
-convenience, 'locally scoped' in the C++ programming language refers to
-data which is created within a context more specific than the context of
-the entire program, or 'translation unit'.  Most data created this way is
-destroyed when that specific context is left. The topic of contexts in C++
-is very broad, so for the purposes of this work context refers to a program
-method or function that begins, concludes and yields a result, also
-destroying locally scoped data. Dynamic casting for the purposes of this
-work refers to copying the value of data contained within an \emph{Epetra}
-vector into locally scoped data that has the property of following the rules of
-complex arithmetic, where a mapping called 'type casting' allows the
-translation of data between different types. No attempt was made to ``tune''
-the CS code in any way for performance. The algorithm  was implemented in such
-as way that it was in one-to-one correspondence with the FD and CD methods;
-however, it must be noted that the overhead associated with the dynamic casts
-in the CS method likely hurts its overall performance and this issue could be
-alleviated by using the next generation of templated \emph{TPetra} vectors
-available in \emph{Trilinos} that are capable of taking an explicit
-instantiation of any data type including {\tt complex}. \todo[disable][]{Re:
-I.3}For the purposes of this work, an explicit instantiation
-refers to requesting the creation of an object, or instance, of a templated
-class or container class that holds data which has the property of
-following desired mathematic or logic rules. The version of \emph{Peridigm}
-modified to implement CS can be found in the paper repository.
+The CS method was implemented in \emph{Peridigm} in a manner that follows closely the implementation of the FD technique.
+However, because of \emph{Peridigm}'s reliance on \emph{Trilinos} and specifically the \emph{Epetra} package, special steps had to be followed to allow the use of complex number data types.
+This is because \emph{Epetra} vectors, which can be thought of as distributed memory parallel cousins to standard C++ Standard Template Library (STL) vectors, are hard coded to be of type {\tt double} and can not be simply declared as having a complex data type.
+\todo[]{Re: I.3} For the reader's convenience, \emph{Epetra} vectors and STL vectors are what are known as 'container classes', whose purpose in object oriented computer programming is to automate the creation, accessing, modification and destruction of the underlying contained data.
+Templated container classes allow for user choice in what properties that underlying data exhibits.
+For example, in this work, complex arithmetic cannot be performed on the data contained within an \emph{Epetra} vector because that data, of type {\tt double} does not have the property of following the rules of complex arithmetic.
+The reason the non-templated \emph{Epetra} vector could not be dispensed with for the purposes of this work is because the \emph{Epetra} vector was tightly integrated with \emph{Peridigm's} parallel communication strategy and linear algebra code.
+Program methods that were used in the course of applying the FD method were copied, renamed and re-written to follow the CS formula.
+Because of restrictions imposed by the container class being used, this involved changing the data type of intermediate variables which were locally scoped, and dynamically casting persistent memory variables to complex data types where necessary.
+\todo[]{Re: I.3}For the reader's convenience, 'locally scoped' in the C++ programming language refers to data which is created within a context more specific than the context of the entire program, or 'translation unit'.
+Most data created this way is destroyed when that specific context is left.
+The topic of contexts in C++ is very broad, so for the purposes of this work context refers to a program method or function that begins, concludes and yields a result, also destroying locally scoped data.
+Dynamic casting for the purposes of this work refers to copying the value of data contained within an \emph{Epetra} vector into locally scoped data that has the property of following the rules of complex arithmetic, where a mapping called 'type casting' allows the translation of data between different types.
+No attempt was made to ``tune'' the CS code in any way for performance.
+The algorithm  was implemented in such as way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurts its overall performance and this issue could be alleviated by using the next generation of templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any data type including {\tt complex}.
+\todo[]{Re: I.3}For the purposes of this work, an explicit instantiation refers to requesting the creation of an object, or instance, of a templated class or container class that holds data which has the property of following desired mathematic or logic rules.
+The version of \emph{Peridigm} modified to implement CS can be found in the paper repository.
 
@@
 many advanced techniques in C++ such as multiple virtual inheritance, pointer
-arithmetic, and distributed data structures, \todo[disable]{Re: I.3}for example \emph{Epetra} vectors, were used in the basic
+arithmetic, and distributed data structures, \todo[]{Re: I.3}for example \emph{Epetra} vectors, were used in the basic
     code and by necessity in the modified code. These concepts are not vital to
@@
 
-\begin{figure}[tbp] \centering \includegraphics[width=1.0\textwidth]{./figs/mesh2000_replacement.png}
-\caption{2000 node mesh} \label{fig:SCMesh}
+\begin{figure}[tbp] 
+    \centering 
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{./figs/mesh2000_replacement.png}
+        \caption{2000 node mesh.}
+        \label{fig:MCMesh}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \includegraphics[width=\textwidth]{./figs/mesh1000000_replacement.png}
+        \caption{1000000 node mesh.}
+        \label{fig:SCMesh}
+    \end{subfigure}
+    \caption{Peridynamic discritizations used in study.}
+    \label{fig:pd_discrete}
 \end{figure}
@@
 terms of accuracy and speed. This series would comprise the single core runs. 
-\todo[disable]{Re: II.16}
+\todo[]{Re: II.16}
 The specific parameters of these tests can be found in \emph{Peridigm} input
@@
 
-\begin{figure}[tbp] \centering \includegraphics[width=1.0\textwidth]{./figs/mesh1000000_replacement.png}
-\caption{1000000 node mesh} \label{fig:MCMesh}
-\end{figure}
-
-\todo[disable]{Re: I.3}For the purposes of this work message passing is what
+\todo[]{Re: I.3}For the purposes of this work message passing is what
 synchronizes multiple computers that each are working on a separate section
@@
 %
-\todo[disable][]{Re: II.22}
-It can be seen in Figs.~\ref{fig:serial_accuracy} and \ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods.
+\todo[]{Re: II.22} It can be seen in Figs.~\ref{fig:serial_accuracy} and \ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods.
 Those figures are depicting absolute error.
@@
 
-\todo[disable][]{Re: I.5}
-
-\section{Convergence and displacement accuracy studies} 
+\section{\todo[][]{Re: I.5}Convergence and displacement accuracy studies} 
 \label{sec:PeridigmConvergenceStudy}
@@
 
-\todo[disable]{Re: I.5}
-\section{Conclusions}
+\section{\todo[]{Re: I.5}Conclusions}
 \label{sec:Conc}

### 2014-04-24 jacobian_calc_CMAME.tex
Numerous changes

@@
 
-\subsubsection{Implementing the complex-step method in \emph{Peridigm}} 
+\subsubsection{Implementing the complex-step method in Peridigm} 
 \label{subsubsec:Impl}
@@
 This is because \emph{Epetra} vectors, which can be thought of as distributed memory parallel cousins to standard C++ Standard Template Library (STL) vectors, are hard coded to be of type {\tt double} and can not be simply declared as having a complex data type.
-\todo[]{Re: I.3} For the reader's convenience, \emph{Epetra} vectors and STL vectors are what are known as 'container classes', whose purpose in object oriented computer programming is to automate the creation, accessing, modification and destruction of the underlying contained data.
-Templated container classes allow for user choice in what properties that underlying data exhibits.
-For example, in this work, complex arithmetic cannot be performed on the data contained within an \emph{Epetra} vector because that data, of type {\tt double} does not have the property of following the rules of complex arithmetic.
-The reason the non-templated \emph{Epetra} vector could not be dispensed with for the purposes of this work is because the \emph{Epetra} vector was tightly integrated with \emph{Peridigm's} parallel communication strategy and linear algebra code.
-Program methods that were used in the course of applying the FD method were copied, renamed and re-written to follow the CS formula.
-Because of restrictions imposed by the container class being used, this involved changing the data type of intermediate variables which were locally scoped, and dynamically casting persistent memory variables to complex data types where necessary.
-\todo[]{Re: I.3}For the reader's convenience, 'locally scoped' in the C++ programming language refers to data which is created within a context more specific than the context of the entire program, or 'translation unit'.
-Most data created this way is destroyed when that specific context is left.
-The topic of contexts in C++ is very broad, so for the purposes of this work context refers to a program method or function that begins, concludes and yields a result, also destroying locally scoped data.
-Dynamic casting for the purposes of this work refers to copying the value of data contained within an \emph{Epetra} vector into locally scoped data that has the property of following the rules of complex arithmetic, where a mapping called 'type casting' allows the translation of data between different types.
+\todo[]{Re: I.3} \emph{Epetra} vectors and STL vectors are what are known as \emph{container classes}, whose purpose in object-oriented computer programming is to automate the creation, accessing, modification,  and destruction of the underlying contained data.
+\emph{Templated} container classes allow for user choice in what properties that underlying data exhibits.
+For example, in this work, complex arithmetic cannot be performed on the data contained within an \emph{Epetra} vector because that data, of type {\tt double},  does not have the property of following the rules of complex arithmetic.
+The reason the non-templated \emph{Epetra} vector could not be dispensed with for the purposes of this work is because the \emph{Epetra} vectors are tightly integrated with \emph{Peridigm}'s parallel communication strategy and linear algebra code.
+Because of restrictions imposed by the container class being used, this involved changing the data type of intermediate variables which were locally scoped (i.e.
+only allocated to memory within the specific CS method and later destroyed), and dynamically casting persistent memory variables to complex data types where necessary.
 No attempt was made to ``tune'' the CS code in any way for performance.
 The algorithm  was implemented in such as way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurts its overall performance and this issue could be alleviated by using the next generation of templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any data type including {\tt complex}.
-\todo[]{Re: I.3}For the purposes of this work, an explicit instantiation refers to requesting the creation of an object, or instance, of a templated class or container class that holds data which has the property of following desired mathematic or logic rules.
 The version of \emph{Peridigm} modified to implement CS can be found in the paper repository.
 
-\emph{Peridigm} was primarily written by computer scientists; therefore,
-many advanced techniques in C++ such as multiple virtual inheritance, pointer
-arithmetic, and distributed data structures, \todo[]{Re: I.3}for example \emph{Epetra} vectors, were used in the basic
-    code and by necessity in the modified code. These concepts are not vital to
-understanding the complex step method itself, but may be necessary to understand complex
-step in \emph{Peridigm}. To see CS and other methods in a more narrow context,
-it is advised that readers also take a look at the simple examples provided on
-the corresponding author's website.
+\emph{Peridigm} was primarily written by computational scientists and engineers; therefore, many advanced techniques in C++ such as multiple virtual inheritance, pointer arithmetic, and distributed data structures are employed.
+These concepts are not vital to understanding the CS method itself, but may be necessary to understand CS in \emph{Peridigm}.
+To see CS and other methods in a more narrow context, it is advised that readers also take a look at the simple examples, such as that described in Section~\ref{sec:WorkingExamples},  provided on the corresponding author's website.
 
@@
 
-The methods were compared by running two sets of test problems where each
-method would solve a problem concurrently. In this context, concurrently means
-successively, drawing from the same independent variables, yet within the same
-computational process; this definition is characterized in further detail in
-Section~\ref{JGAM}. Each test problem simulated a $\SI{4}{\meter}$ by
-$\SI{0.5}{\meter}$ by $\SI{0.5}{\meter}$ meter block of a material undergoing
-tension along the axis parallel to the long dimension of the block. The test
-problems were set up in \emph{Peridigm} as quasi-static equilibrium problems,
-where the displacement boundary conditions used to apply tension were applied
-gradually in ``load steps'', and an equilibrium solution for each load step was
-achieved before applying the next load step. Applying the load
-in gradual steps is a widely used technique to ensure the Newton's method is
-always initialized near the actual solution and therefore likely to converge.
-The fictional material represented in the model had values of
-$\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$
-for shear modulus. The peridynamic horizon, a length scale that effectively
-sets the bandwidth of tangent-stiffness matrix in the computation, used in the test
-problems was always set to three times the nominal discretization size (i.e.
-node spacing in the particle discretization scheme). This results in a minimum
-tangent-stiffness matrix bandwidth of 21 for a 3D problem (for an optimal node
-numbering scheme). Pictured in Fig.~\ref{fig:SCMesh} is a viewgraph for one of
-the peridynamic meshes used in this study. The coloration in the image
-indicates node numbering, where a smooth color transition means that the nodes
-have close ID numbers. The abrupt color changes in the center and at the ends
-of the body show that numbering is not contiguous and is assigned with respect to four regions. The purpose of designating
-these regions, for this study, was to apply displacement boundary
-conditions called \emph{volume constraints} in peridynamics.
-
+The methods were compared by running two sets of test problems where each method would solve a problem concurrently.
+In this context, concurrently means successively, drawing from the same independent variables, yet within the same computational process; this definition is characterized in further detail in Section~\ref{JGAM}.
+Each test problem simulated a $\SI{4}{\meter}$ by $\SI{0.5}{\meter}$ by $\SI{0.5}{\meter}$ meter block of a material undergoing tension along the axis parallel to the long dimension of the block.
+The test problems were set up in \emph{Peridigm} as quasi-static equilibrium problems, where the displacement boundary conditions used to apply tension were applied gradually in ``load steps'', and an equilibrium solution for each load step was achieved before applying the next load step.
+Applying the load in gradual steps is a widely used technique to ensure the Newton's method is always initialized near the actual solution and therefore likely to converge.
+The fictional material represented in the model had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus.
+The peridynamic horizon, a length scale that effectively sets the bandwidth of tangent-stiffness matrix in the computation, used in the test problems was always set to three times the nominal discretization size (i.e.
+node spacing in the particle discretization scheme).
+Pictured in Figure.~\ref{fig:pd_discrete} is an illustration of the peridynamic discretizations used in this study.
+The coloration in the images indicate material regions that were defined to assist in post-processing application of boundary conditions.
+%
 \begin{figure}[tbp] 
@@
 
-A series of nine single compute core (i.e.\ serial) test problems were run with
-the discretization size being refined with every test in the series. The aim of
-increasing the refinement was to examine differences between the methods in
-terms of accuracy and speed. This series would comprise the single core runs. 
-\todo[]{Re: II.16}
-The specific parameters of these tests can be found in \emph{Peridigm} input
-{\tt xml} files included in the paper repository, however the parameters
-are reproduced in Table~\ref{tab:TestParams}. These {\tt xml} files allow
-users with the appropriately modified version of \emph{Peridigm} to reproduce
-the results found in this paper.
-
-\begin{table*}[!ht]    
+A series of nine single compute core (i.e.\ serial) test problems were run with the discretization size being refined with every test in the series.
+The aim of increasing the refinement was to examine differences between the methods in terms of accuracy and speed.
+This series would comprise the single core runs.
+\todo[]{Re: II.16} The specific parameters of these tests can be found in \emph{Peridigm} input {\tt xml} files included in the paper repository, however the parameters are reproduced in Table~\ref{tab:TestParams}.
+These {\tt xml} files allow users with the appropriately modified version of \emph{Peridigm} to reproduce the results found in this paper.
+%
+\begin{table*}[!tbp]    
   \scriptsize
@@
          \toprule
-         Peri. nodes & Nozero TS elems. & Horizon & Load rate & Number Load steps & Abs. tolerance \\ 
-                     & $\times 10^6$    & \si{\meter}& \si{\centi\meter\per\second} &   &     \\
+         Peri. nodes & Nozero TS elems. & Horizon ($\delta$) & Load rate & Number Load steps & Abs. tolerance \\ 
+                     & $\times 10^6$    & \si{\centi\meter}& \si{\centi\meter\per\second} &   &     \\
         \midrule
-        1000 & $1.99$             & $.3$   & $2.0$ & $2$ & $\num{10E-8}$ \\
-        2000 & $4.25$             & $.24$  & $2.0$ & $2$ & $\num{10E-8}$ \\
-        3000 & $7.79$             & $.21$  & $2.0$ & $2$ & $\num{10E-8}$ \\
-        4000 & $15.4$             & $.19$  & $2.0$ & $2$ & $\num{10E-8}$ \\
-        6000 & $20.0$             & $.16$  & $2.0$ & $2$ & $\num{10E-8}$ \\
-        8000 & $31.4$             & $.15$  & $2.0$ & $2$ & $\num{10E-8}$ \\
-        12000 & $40.1$            & $.131$ & $2.0$ & $2$ & $\num{10E-8}$ \\
-        16000 & $83.9$            & $.12$  & $2.0$ & $2$ & $\num{10E-8}$ \\
-        32000 & $165.0$           & $.095$ & $2.0$ & $2$ & $\num{10E-8}$ \\
-        1000000 & $1670$          & $.03$  & $2.0$ & $2$ & $\num{10E-8}$ \\
-        1000000 & $1670$          & $.03$  & $2.0$ & $2$ & $\num{10E-8}$ \\
-        1000000 & $1670$          & $.03$  & $2.0$ & $2$ & $\num{10E-8}$ \\
-        1000000 & $1670$          & $.03$  & $2.0$ & $2$ & $\num{10E-8}$ \\
+        1000 & $1.99$             & $30$   & $2$ & $2$ & $\num{10E-8}$ \\
+        2000 & $4.25$             & $24$  & $2$ & $2$ & $\num{10E-8}$ \\
+        3000 & $7.79$             & $21$  & $2$ & $2$ & $\num{10E-8}$ \\
+        4000 & $15.4$             & $19$  & $2$ & $2$ & $\num{10E-8}$ \\
+        6000 & $20.0$             & $16$  & $2$ & $2$ & $\num{10E-8}$ \\
+        8000 & $31.4$             & $15$  & $2$ & $2$ & $\num{10E-8}$ \\
+        12000 & $40.1$            & $13.1$ & $2$ & $2$ & $\num{10E-8}$ \\
+        16000 & $83.9$            & $12$  & $2$ & $2$ & $\num{10E-8}$ \\
+        32000 & $165.0$           & $9.5$ & $2$ & $2$ & $\num{10E-8}$ \\
+        1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
         \bottomrule
@@
 
-Another series of four test problems were simulated, however this time the
-number of compute cores used to solve the problem was increased from test to
-test while the discretization level was held constant at 1 million
-computational nodes (3 million degrees of freedom). This allowed for sufficient
-computational nodes per core even at the highest level of parallelization such
-that message passing computation did not overwhelm the simulation.  The
-specific parameters of these tests can be found in \emph{Peridigm} input {\tt
-xml} files also included in the paper repository, but a summary of the most relevant
-ones appears in Table~\ref{tab:TestParams}. A viewgraph of the 1 million node mesh is
-shown in Fig.~\ref{fig:MCMesh}. The nodes in the image, although actually discrete points
-in space are dense enough to appear as a continuous body at a distance.
-
-\todo[]{Re: I.3}For the purposes of this work message passing is what
-synchronizes multiple computers that each are working on a separate section
-of the problem domain. This operation involves the time-consuming use of
-the data network. To reduce the proportion of simulation time spent on
-message passing, good practice is to maximize the amount of information
-contained within a subdomain solely managed by a single computer, relative
-to the amount of information shared across multiple subdomains and
-therefor multiple computers. A useful experiment to illustrate this idea
-concretely would be to implement a code to solve a 2D Laplace equation for
-simulating a simple heat conduction problem on two processors. Use the
-textbook finite difference method, which in this case does not refer just
-to the differentiation technique.  First decompose the domain into two
-contiguous chunks, having one shared border, and distribute these chunks,
-or subdomains to the two processors. Perform the simulation a multitude of
-times to calculate an average runtime.  Next implement the solution a
-different way, instead partitioning the problem domain into many narrow,
-alternating strips, and distribute those to again two processors, where adjacent
-strips in the graph of the domain should be sent to different processors. This way
-of decomposing the domain will cause much overlap between the subdomains
-managed by the two processors and much time will be spent synchronizing the
-subdomain overlap or 'ghost nodes' between the processors. Run the modified
-simulation a multitude of times and take the average runtime. The larger
-the unpartitioned domain, the worse the discrepancy should be.
-
-\subsubsection{Quantities of interest} \label{JGAM}
-
-As stated in the Introduction, a goal of this study was to compare CS, FD, and
-CD on the basis of accuracy. AD was omitted from the comparison because it
-served as the standard of accuracy for the other methods in the absence of
-appropriate analytical forms for the tangent-stiffness matrix associated with
-the system solved in this study. The assumption that AD is accurate enough to
-serve as a standard is supported by AD's implementation as a computerized chain
-rule as explained in Section~\ref{ADsubsection}.
-
-It would be a poor comparative study to compare tangent-stiffness matrices from
-different problems, load steps that start with different current
-configurations, or from different iterations; because of this, it was necessary
-to solve one load step and conclude each Newton iteration within that load step
-by updating the displacement iterate with only the results of the AD method and
-to subsequently feed all four methods the same updated displacement in the
-following iteration. This decision precluded a comparison of Newton iteration
-convergence rate, since if the methods were allowed to solve a problem at their
-individual pace, differences in accuracy would produce differences in guess
-updates and therefore the number and nature of Newton iterations performed
-(e.g. lower accuracy predictions of the algorithmically correct
-tangent-stiffness matrix could cause a loss of quadratic convergence).  Different
-guesses from iterations started with different previous guesses could not be
-data for a valid comparison of tangent-stiffness matrix accuracy between methods.
-Additionally, having each of the methods physically operate in the same
-process, serial or one of associated parallel, allowed the comparison of
-tangent-stiffness matrix calculation time as cached from random access memory
-rather than from the hard-disk.  Caching from RAM gives the benefit of vastly
-greater speed and simpler programming compared to some other solution
-speculatively involving dynamic file management on files which for one of the
-components of this study would be on the order of a terabyte in size. However,
-the price of running the methods together in the same process and avoiding the
-use of the hard-disk was that at least two tangent-stiffness matrices had to be
-stored in RAM during the simulation, which meant that special high-memory
-compute nodes were needed for the 1 million peridynamic node test series.
-
-The other goal was to compare the four methods, including AD, on the basis of
-speed. Speed is defined as the total compute-time of one iteration of a Newton
-step which includes the tangent-stiffness matrix calculation and parallel assembly.
-The speed of iteration was measured because it could be so done at the same
-time as accuracy was being measured given a single tangent-stiffness
-calculation.  It was assumed that the order that each method was evaluated in
-parallel was unimportant, that evaluating each method successively within each
-solver iteration did not affect their performance individually and that speed
-of computation did not change with time. These assumptions allowed the test
-program to run the same problem with each of the methods at effectively the
-same time and generate an equal volume of data from each method. It was also
-assumed that the tangent-stiffness matrix calculation time for each of the
-methods did not vary based on the current configuration, as it changes slightly
-from iteration to iteration as the Newton solution iterate is updated.  This
-assumption allowed calculation time measurements to be averaged over all
-iterations within a single load step. The purpose of averaging calculation time
-measurements was to informally address the extraneous variables of parallel
-evaluation order and computer system load due to system processes not
-associated with this study. While not part of this study, the example programs
-available in the paper repository allow the reader to make a comparison of the
-methods for themselves on the basis of accuracy and convergence rate for two
-example nonlinear systems. In these examples, one will notice that the finite
-difference methods will lose quadratic convergence when the step size selected
-is too large such that accuracy relative to AD is decreased.
-
-The goals of collecting accuracy and speed measurements were achieved by
-developing and implementing metrics within the simulation program used in the
-study. The metric used to measure the accuracy of a tangent-stiffness matrix
-was the Frobenius norm (analogous to an $l^2$-norm except defined on a matrix)
-of the element-wise difference between the tangent-stiffness matrix produced by
-the method being evaluated and the tangent-stiffness matrix produced by the AD
-based method (i.e., the \emph{exact} derivatives), given that both methods were
-set upon the same problem. The lower the value of this metric, the more
-accurate the method. The expression for the accuracy metric was
-%
-\begin{equation} D = \sqrt{\sum_i \sum_j(K^{AD}_{ij} - K^{M}_{ij})^2},
-\label{eqn:accuracy} \end{equation}
-%
-where $D$ is distance, $K^{AD}$ is the tangent-stiffness matrix produced by the
-AD based method, and $M$ is replaced with either CS, FD, or CD depending on the
-method being compared.  The metric used to compare the speed of the different
-methods was the time in seconds required to calculate the tangent-stiffness
-matrix. A final basis of comparison used in this study called
-\emph{computational efficiency} is based on specific calculation time per
-tangent-stiffness matrix element.
+Another series of four test problems were simulated, however this time the number of compute cores used to solve the problem was increased from test to test while the discretization level was held constant at 1 million computational nodes (3 million degrees of freedom).
+This allowed for sufficient computational nodes per core even at the highest level of parallelization such that message passing computation did not overwhelm the simulation.
+The specific parameters of these tests can be found in \emph{Peridigm} input {\tt xml} files also included in the paper repository, but a summary of the most relevant ones appears in Table~\ref{tab:TestParams}.
+An image of the 1 million node discretization is shown in Figure~\ref{fig:MCMesh}.
+The nodes in the image, although actually discrete points in space are dense enough to appear as a continuous body at a distance.
+
+\subsubsection{Quantities-of-interest} 
+\label{JGAM}
+
+As stated in Section~\ref{sec:intro}, a goal of this study was to compare CS, FD, and CD on the basis of accuracy.
+AD was omitted from the comparison because it served as the standard of accuracy for the other methods in the absence of appropriate analytical forms for the tangent-stiffness matrix associated with the system solved in this study.
+The assumption that AD is accurate enough to serve as a standard is supported by AD's implementation as a computerized chain rule as explained in Section~\ref{ADsubsection}.
+
+It would be a poor comparative study to compare tangent-stiffness matrices from different problems, load steps that start with different current configurations, or from different iterations; because of this, it was necessary to solve one load step and conclude each Newton iteration within that load step by updating the displacement iterate with only the results of the AD method and to subsequently feed all four methods the same updated displacement in the following iteration.
+This decision precluded a comparison of Newton iteration convergence rate, since if the methods were allowed to solve a problem at their individual pace, differences in accuracy would produce differences in guess updates and therefore the number and nature of Newton iterations performed (e.g.
+lower accuracy predictions of the algorithmically correct tangent-stiffness matrix could cause a loss of quadratic convergence).
+Different guesses from iterations started with different previous guesses could not be data for a valid comparison of tangent-stiffness matrix accuracy between methods.
+Additionally, having each of the methods physically operate in the same process, serial or one of associated parallel, allowed the comparison of tangent-stiffness matrix calculation time as cached from random access memory rather than from the hard-disk.
+Caching from RAM gives the benefit of vastly greater speed and simpler programming compared to some other solution speculatively involving dynamic file management on files which for one of the components of this study would be on the order of a terabyte in size.
+However, the price of running the methods together in the same process and avoiding the use of the hard-disk was that at least two tangent-stiffness matrices had to be stored in RAM during the simulation, which meant that special high-memory compute nodes were needed for the 1 million peridynamic node test series.
+
+The other goal was to compare the four methods, including AD, on the basis of speed.
+Speed is defined as the total compute-time of one iteration of a Newton step which includes the tangent-stiffness matrix calculation and parallel assembly.
+The speed of iteration was measured because it could be so done at the same time as accuracy was being measured given a single tangent-stiffness calculation.
+It was assumed that the order that each method was evaluated in parallel was unimportant, that evaluating each method successively within each solver iteration did not affect their performance individually and that speed of computation did not change with time.
+These assumptions allowed the test program to run the same problem with each of the methods at effectively the same time and generate an equal volume of data from each method.
+It was also assumed that the tangent-stiffness matrix calculation time for each of the methods did not vary based on the current configuration, as it changes slightly from iteration to iteration as the Newton solution iterate is updated.
+This assumption allowed calculation time measurements to be averaged over all iterations within a single load step.
+The purpose of averaging calculation time measurements was to informally address the extraneous variables of parallel evaluation order and computer system load due to system processes not associated with this study.
+While not part of this study, the example programs available in the paper repository allow the reader to make a comparison of the methods for themselves on the basis of accuracy and convergence rate for two example nonlinear systems.
+In these examples, one will notice that the finite difference methods will lose quadratic convergence when the step size selected is too large such that accuracy relative to AD is decreased as shown in Table~\ref{tab:ConvergenceStudy}.
+
+The goals of collecting accuracy and speed measurements were achieved by developing and implementing metrics within the simulation program used in the study.
+The metric used to measure the accuracy of a tangent-stiffness matrix was the Frobenius norm (analogous to an $l^2$-norm except defined on a matrix) of the element-wise difference between the tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix produced by the AD based method (i.e., the \emph{exact} derivatives to numerical precision), given that both methods were set upon the same problem.
+The lower the value of this metric, the more accurate the method.
+The expression for the accuracy metric was
+%
+\begin{equation} 
+    D = \sqrt{\sum_i \sum_j(K^{AD}_{ij} - K^{M}_{ij})^2},
+    \label{eqn:accuracy} 
+\end{equation}
+%
+where $D$ is distance, $K^{AD}$ is the tangent-stiffness matrix produced by the AD based method, and $M$ is replaced with either CS, FD, or CD depending on the method being compared.
+The metric used to compare the speed of the different methods was the time in seconds required to calculate the tangent-stiffness matrix.
+A final basis of comparison used in this study called \emph{computational efficiency} is based on specific calculation time per tangent-stiffness matrix element.
 
@@
 %
-Discretization for the test problems run in this study was done using the
-software package \emph{Cubit} developed at Sandia National Laboratories
-\cite{ref-Cubit}. \emph{Cubit} generates a finite element mesh that is
-internally converted to a ``particle'' discretization internally when calling
-\emph{Peridigm}. The number of computational nodes in \emph{Peridigm}
-corresponds to the number of finite elements in the discretization, not the
-number of finite element nodes.  An example journal script can be found on the
-corresponding author's website.
-
-The test problems were run on the \emph{Stampede} HPC cluster computer housed
-at \emph{TACC} (Texas Advanced Computing Center) at The University of Texas at
-Austin. The single core test series was run using the {\tt normal} queue
-compute nodes, while the parallel test series was run using one to four {\tt
-large memory} nodes each having one terabyte of local RAM. This large amount of
-memory was required for reasons explained in Section~\ref{JGAM}, second paragraph.
+Discretization for the test problems run in this study was done using the software package \emph{Cubit} developed at Sandia National Laboratories \cite{ref-Cubit}.
+\emph{Cubit} generates a finite element mesh that is converted to a ``particle'' discretization internally when running \emph{Peridigm}.
+The number of computational nodes in \emph{Peridigm} corresponds to the number of finite elements in the discretization, not the number of finite element nodes.
+An example journal script can be found in the corresponding paper repository.
+
+The test problems were run on the \emph{Stampede} HPC cluster computer housed at \emph{TACC} (Texas Advanced Computing Center) at The University of Texas at Austin.
+The single core test series was run using the {\tt normal} queue compute nodes, while the parallel test series was run using one to four {\tt large memory} nodes each having one terabyte of local RAM.
+This large amount of memory was required for reasons explained in Section~\ref{JGAM}.
 
 \subsubsection{Data reduction} 
-
-The output produced by \emph{Peridigm} including the additional accuracy and
-speed measurements was redirected from the console to text files by the
-resource manager, \emph{SLURM} (Simple Linux User Resource Manager).
-Directories and filenames were chosen so that data would be easy to sort by
-individual test run.  \emph{Python} scripts were then used to post-process and
-plot the data. These scripts averaged accuracy or speed data for a tests
-corresponding to a particular peridynamic node density over all iterations for
-that run, and then plotted this averaged data  as a function of peridynamic
-node density as it varied from test to test in the series.  Similarly, for the
-parallel tests, the data was averaged and plotted in much the same way, but as
-a function of number of compute cores used to solve the problem. Confidence
-intervals were not calculated on the measurements. The main reason is that
-network traffic from other users makes running tests on the HPC cluster a time
-varying process. These data reduction and plotting scripts can be found on the
-corresponding author's website.
+The output produced by \emph{Peridigm} including the additional accuracy and speed measurements was redirected from the console to text files by the resource manager, \emph{SLURM} (Simple Linux User Resource Manager).
+Directories and filenames were chosen so that data would be easy to sort by individual test run.
+\emph{Python} scripts were then used to post-process and plot the data.
+These scripts averaged accuracy or speed data for a tests corresponding to a particular peridynamic node density over all iterations for that run, and then plotted this averaged data  as a function of peridynamic node density as it varied from test to test in the series.
+Similarly, for the parallel tests, the data was averaged and plotted in much the same way, but as a function of number of compute cores used to solve the problem.
+Confidence intervals were not calculated on the measurements.
+The main reason is that network traffic from other users makes running tests on the HPC cluster a time varying process.
+These data reduction and plotting scripts can be found in the paper repository.
 
@@
 Table~\ref{tab:results} shows the compute core count, number of nonzero tangent-stiffness (TS) matrix elements, load step average calculation time and load step average accuracy.
-The accuracy measure is given by Eqn.~(\ref{eqn:accuracy}) and the choice of elastic modulii mentioned in Section~\ref{tcs}.
-%
-\begin{table*}[!ht]    
+The accuracy measure is given by equation~(\ref{eqn:accuracy}) and the choice of elastic modulii mentioned in Section~\ref{tcs}.
+%
+\begin{table*}[!tbp]    
   \scriptsize
@@
 \subsection{Speed data} 
-
-Speed measurements were taken according to the methods
-mentioned in Section~\ref{tcs}. The averaged results of these speed
-measurements for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
-Average time in seconds to compute a TS matrix is plotted as a function of the
-number of nonzero TS elements.  CS was the slowest of the four methods, taking
-the longest average time per iteration,  followed by CD, FD and AD in that
-order. Calculation time has a power-law relationship with problem size, with a
-power-law index of roughly $\num{1.0E-6}$ for each of the methods. This bodes well
-for the scalability of each of the methods by themselves in \emph{Peridigm}. If we
-assumed every individual operation, generally speaking, took the same amount of
-time on average such that compute time were a measure of number of operations,
-this suggests that the methods as in \emph{Peridigm} have the same complexity class
-when considering computing a TS matrix alone, that is
-computation time is $\mathcal{O}(10^{(10^{-6} \times \log_{10}n)})$. The definition of 
-complexity class used here is the informal one given in \cite[pp. 1059]{cormen2009introduction}. This would
-suggest that the methods are equally well suited to the task of computing a TS matrix, if it were further assumed
-that the TS matrices produced by each method were identical. This is not the case
-as seen in the following subsection.
-%
-\begin{figure}[tbp] \centering \scalebox{1.0}{\input{./figs/serial_speed.pgf}}
-\caption{Serial test series speed measurements.} \label{fig:serial_speed}
+%
+Speed measurements were taken according to the methods mentioned in Section~\ref{tcs}.
+The averaged results of these speed measurements for the serial tests can be seen in Fig.~\ref{fig:serial_speed}.
+Average time in seconds to compute a TS matrix is plotted as a function of the number of nonzero TS elements.
+CS was the slowest of the four methods, taking the longest average time per iteration, followed by CD, FD and AD in that order.
+Calculation time has a power-law relationship with problem size, with a power-law index of roughly $\num{1.0E-6}$ for each of the methods.
+This bodes well for the scalability of each of the methods by themselves in \emph{Peridigm}.
+If we assumed every individual operation, generally speaking, took the same amount of time on average such that compute time were a measure of number of operations, this suggests that the methods as in \emph{Peridigm} have the same complexity class when considering computing a TS matrix alone, that is computation time is $\mathcal{O}(10^{(10^{-6} \times \log_{10}n)})$.
+The definition of complexity class used here is the informal one given in \cite[pp.
+1059]{cormen2009introduction}.
+This would suggest that the methods are equally well suited to the task of computing a TS matrix, if it were further assumed that the TS matrices produced by each method were identical.
+This is not the case as seen in the following subsection.
+%
+\begin{figure}[tbp] 
+    \centering 
+    \scalebox{1.0}{\input{./figs/serial_speed.pgf}}
+    \caption{Serial test series speed measurements.} 
+    \label{fig:serial_speed}
 \end{figure}
 %
-For the parallel simulations, averaged speed results appear in
-Fig.~\ref{fig:multi_speed}. For this series, rather than varying the number of
-nonzero TS matrix elements, the number of compute cores used to solve the
-problem is increased, while the number of nonzero TS matrix elements remained
-constant at $\num{1.67E9}$ .  CS again was the slowest method, followed by CD,
-AD and then FD.  There is a somewhat surprise reversal of FD and AD for these
-parallel simulations.  Figure~\ref{fig:multi_speed} indicates that when looking
-at average iteration time, the factor of speed improvement scales inversely with
-the number of compute cores used for the given conditions of this group of
-tests. Looking at Fig.~\ref{fig:multi_speed}, the reader can see that when core
-count doubles, compute time halves, for each method. Again, these results show
-that each of the methods as implemented in \emph{Peridigm} scale well when
-additional processors are used. 
-%
-\begin{figure}[tbp] \centering \scalebox{1.0}{\input{./figs/multi_speed.pgf}}
-\caption{Multicore test series speed measurements.} \label{fig:multi_speed}
+For the parallel simulations, averaged speed results appear in Figure~\ref{fig:multi_speed}.
+For this series, rather than varying the number of nonzero TS matrix elements, the number of compute cores used to solve the problem is increased, while the number of nonzero TS matrix elements remained constant at $\num{1.67E9}$ .
+CS again was the slowest method, followed by CD, AD and then FD.
+There is a somewhat surprise reversal of FD and AD for these parallel simulations.
+Figure~\ref{fig:multi_speed} indicates that when looking at average iteration time, the factor of speed improvement scales inversely with the number of compute cores used for the given conditions of this group of tests.
+Figure~\ref{fig:multi_speed} clearly shows that when core count doubles, compute time halves, for each method.
+Again, these results show that each of the methods as implemented in \emph{Peridigm} scale well when additional processors are used.
+%
+\begin{figure}[tbp] 
+    \centering 
+    \scalebox{1.0}{\input{./figs/multi_speed.pgf}}
+    \caption{Multicore test series speed measurements.} 
+    \label{fig:multi_speed}
 \end{figure}
 %
-A speculative explanation as to why FD and AD have reversed places in the speed
-tests with respect to the serial results is that the AD data-structures carry a
-list of partial derivatives of the functional evaluations along side the
-functional evaluations themselves only to be processed at the end of the
-tangent-stiffness matrix calculation when the derivatives are requested. This
-represents additional load on the hardware memory subsystem, such that may
-exceed that of FD yet still falls under a supposed memory bandwidth threshold
-during serial operation. Cores within a multicore processor use the same shared
-bus to access memory, such that cores compete for memory access. Supposing that
-exceeding the memory bandwidth threshold penalizes additional requested
-bandwidth, it is likely that the decrease in available memory bandwidth per
-core caused by parallel operation coupled with AD's additional resource
-requirements causes the AD method to be slower than FD.
+A speculative explanation as to why FD and AD have reversed places in the speed tests with respect to the serial results is that the AD data-structures carry a list of partial derivatives of the functional evaluations along side the functional evaluations themselves only to be processed at the end of the tangent-stiffness matrix calculation when the derivatives are requested.
+This represents additional load on the hardware memory subsystem, such that may exceed that of FD yet still falls under a supposed memory bandwidth threshold during serial operation.
+Cores within a multicore processor use the same shared bus to access memory, such that cores compete for memory access.
+Supposing that exceeding the memory bandwidth threshold penalizes additional requested bandwidth, it is likely that the decrease in available memory bandwidth per core caused by parallel operation coupled with AD's additional resource requirements causes the AD method to be slower than FD.
 
 \subsection{Accuracy data}
-
-Data for accuracy measurements were taken according to the methods mentioned in
-Section~\ref{tcs}, and reduced using Eqn.~(\ref{eqn:accuracy}). The
-averaged results of these accuracy measurements for the serial tests can be
-seen in Fig.~\ref{fig:serial_accuracy}. Averaged Frobenius norm (referred to as
-$l^2$-norm in the figure labels for conciseness) of the element-wise difference
-between a given TS matrix and the TS matrix produced by the AD method is
-plotted as a function of the number of nonzero TS matrix elements. CS was shown
-to be the most accurate of the methods when compared to AD and was roughly
-\emph{$6$ orders of magnitude more accurate than CD} and \emph{$9$ orders of
-magnitude more accurate than FD}. This accuracy ranking order was maintained
-for each test in the series. It should be pointed out that compared to the
-$l^2$-norms of any of the TS matrices by themselves, the magnitude of the
-accuracy metric calculated with Eqn.~\ref{eqn:accuracy} for any of the methods
-is relatively small. Relative error was computed from these measurements and is reported in
-Table~\ref{tab:RelAcc}. The values for the $l^2$-norms of the TS matrices appear
-in the simulation data which is available in the paper repository. The effect
-of these accuracy differences on convergence rate of the Newton solver was studied and the
-results are reported in section~\ref{sec:PeridigmConvergenceStudy}.  
-%
-The accuracy of FD and CD could be improved by using a smaller probe distance,
-$h$; however, if $h$ is too small, then the methods can suffer from severe
-effects, including solution divergence, of the round-off error associated with
-subtracting two numbers that are very close to one another.  The value of $h$
-used in this study is the default value used in \emph{Peridigm} which is
-heuristically derived based on computational node spacing using the method explained in \cite[pp. 90]{ref-Adaggio}.  
-To summarize, a spacing of $1.E-6$ times grid spacing is used to provide acceptable accuracy but prevent roundoff error, 
-as is not a user determined setting in \emph{Peridigm}. One of the beauties
-of the CS method is that since dependence on $h$ decreases for small $h$, which
-a numerical example in \cite[Table 1]{squire1998using} demonstrates, $h$,
-has no lower limit for effective values, therefore it can be set to machine epsilon and the
-analysis performed without the issues associated with round-off error. The
-value chosen for $h$ for CS used in this study was $\num{1.E-100}$ to
-demonstrate this.
-%
-\begin{figure}[tbp] \centering
-\scalebox{1.0}{\input{./figs/serial_accuracy.pgf}} \caption{Serial test series
-accuracy measurements.} \label{fig:serial_accuracy} \end{figure}
-%
-The averaged results of the accuracy measurements taken for the parallel test
-series can be seen in Fig.~\ref{fig:multi_accuracy}. The ordinate axis units
-remain unchanged from Fig.~\ref{fig:serial_accuracy}, but the abscissa now
-indicates number of compute cores used to solve the test problem. Accuracy
-trends matched those for the single core tests. 
-%
-\begin{figure}[tbp] \centering
-\scalebox{1.0}{\input{./figs/multi_accuracy.pgf}} \caption{Multicore test
-series accuracy measurements.} \label{fig:multi_accuracy} \end{figure}
-%
-\todo[]{Re: II.22} It can be seen in Figs.~\ref{fig:serial_accuracy} and \ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods.
+%
+Data for accuracy measurements were taken according to the methods mentioned in Section~\ref{tcs}, and reduced using equation~(\ref{eqn:accuracy}).
+The averaged results of these accuracy measurements for the serial tests can be seen in Figure~\ref{fig:serial_accuracy}.
+Averaged Frobenius norm (referred to as $l^2$-norm in the figure labels for conciseness) of the element-wise difference between a given TS matrix and the TS matrix produced by the AD method is plotted as a function of the number of nonzero TS matrix elements.
+CS was shown to be the most accurate of the methods when compared to AD and was roughly \emph{$6$ orders of magnitude more accurate than CD} and \emph{$9$ orders of magnitude more accurate than FD}.
+This accuracy ranking order was maintained for each test in the series.
+It should be pointed out that compared to the $l^2$-norms of any of the TS matrices by themselves, the magnitude of the accuracy metric calculated with equation~\ref{eqn:accuracy} for any of the methods is relatively small.
+Relative error was computed from these measurements and is reported in Table~\ref{tab:RelAcc}.
+The values for the $l^2$-norms of the TS matrices appear in the simulation data which is available in the paper repository.
+The effect of these accuracy differences on convergence rate of the Newton solver was studied and the results are reported in Section~\ref{sec:PeridigmConvergenceStudy}.
+%
+The accuracy of FD and CD could be improved by using a smaller probe distance, $h$; however, if $h$ is too small, then the methods can suffer from severe effects, including solution divergence, of the round-off error associated with subtracting two numbers that are very close to one another.
+The value of $h$ used in this study is the default value used in \emph{Peridigm} which is heuristically derived based on computational node spacing using the method explained in \cite[pp.
+90]{ref-Adaggio}.
+To summarize, a spacing of $1.E-6$ times grid spacing is used to provide acceptable accuracy but prevent roundoff error, as is not a user determined setting in \emph{Peridigm}.
+One of the beauties of the CS method is that since dependence on $h$ decreases for small $h$, which a numerical example in \cite[Table 1]{squire1998using} demonstrates, $h$, has no lower limit for effective values, therefore it can be set to machine epsilon and the analysis performed without the issues associated with round-off error.
+The value chosen for $h$ for CS used in this study was $\num{1.E-100}$ to demonstrate this.
+%
+\begin{figure}[tbp] 
+    \centering
+    \scalebox{1.0}{\input{./figs/serial_accuracy.pgf}} 
+    \caption{Serial test series accuracy measurements.} 
+    \label{fig:serial_accuracy} 
+\end{figure}
+%
+The averaged results of the accuracy measurements taken for the parallel test series can be seen in Figure~\ref{fig:multi_accuracy}.
+The ordinate axis units remain unchanged from Figure~\ref{fig:serial_accuracy}, but the abscissa now indicates number of compute cores used to solve the test problem.
+Accuracy trends matched those for the single core tests.
+%
+\begin{figure}[tbp] 
+    \centering
+    \scalebox{1.0}{\input{./figs/multi_accuracy.pgf}} 
+    \caption{Multicore test series accuracy measurements.} 
+    \label{fig:multi_accuracy} 
+\end{figure}
+%
+\todo[]{Re: II.22} It can be seen in Figures~\ref{fig:serial_accuracy} and \ref{fig:multi_accuracy} that there appears to be some fluctuating behavior in the accuracy of the methods.
 Those figures are depicting absolute error.
 If relative error is computed from the absolute error measurements by dividing them by the Frobenius norm of the appropriate AD produced TS matrix, it can be seen as in Table~\ref{tab:RelAcc}, that there is little variation in the accuracy of the methods between tests or between serial and parallel test groups.
-
-\begin{table*}[!ht]    
+%
+\begin{table*}[tbp]    
   \scriptsize
@@
 \subsection{Efficiency data}
-
+%
 Efficiency, calculated as the average number of seconds taken per nonzero Jacobian matrix element, was plotted as a function of the number of nonzero TS matrix elements.
-The single core results are shown in Fig.~\ref{fig:serial_efficiency}.
-The ranking order of the methods
-matched their speed ranking in Fig.~\ref{fig:serial_speed}. This plot indicates that efficiency is insensitive to the number of nonzero TS elements.  
+The single core results are shown in Figure~\ref{fig:serial_efficiency}.
+The ranking order of the methods matched their speed ranking in Figure~\ref{fig:serial_speed}.
+This plot indicates that efficiency is insensitive to the number of nonzero TS elements.
 %
@@
 %
-The parallel efficiency results are shown in Fig.~\ref{fig:multi_efficiency}.
+The parallel efficiency results are shown in Figure~\ref{fig:multi_efficiency}.
 The parallel efficiency plot shows that AD was slower than FD because for the given test problem it had a lower efficiency.
@@
 
-\section{\todo[][]{Re: I.5}Convergence and displacement accuracy studies} 
+\section{\todo[]{Re: I.5}Convergence and displacement accuracy studies.} 
 \label{sec:PeridigmConvergenceStudy}
 %
-While the previous sections profiled the per-iteration performance of each of
-the methods, it was necessary to conduct convergence rate studies in order to
-get a complete picture. It is necessary for an analyst to know if the
-accuracy differences measured in the per-iteration tests would amount to
-differences in convergence rates among the methods if they were allowed to solve problems
-normally. The question is posed, given that step-size $h$ is heuristically
-selected by \emph{Peridigm} for the finite difference methods and presumable
-makes a good choice, do the single-iteration Jacobian accuracy results predict
-the ranking of the methods in terms of convergence rate for a given study? For
-brevity, convergence rate was measured for each of the methods in solving just
-the largest problem of the single-core test series. That problem was solved with each of the methods
-both on a single core and all sixteen cores of a 'normal' node on \emph{Stampede}. As was seen in the per-iteration results,
-Jacobian accuracy did not vary greatly from test to test, such that any pair of tests
-from the two series could have been considered reasonably representative. As in
-the validation problem shown in ~\ref{subsec:Validation}, would each of the methods return an
-equal convergence rate so long as $h$ were optimally selected?
-
-The results shown in Table ~\ref{tab:ConvergenceStudy2} indicate that this was the case. 
-The difference in Jacobian accuracy between the methods seen in the per-iteration tests
-was not sufficient to lead to differences in the convergence rate of the methods.
-
-\begin{table*}[!ht]   
-    \centering \caption{Average iterations/ load-step} 
+While the previous sections profiled the per-iteration performance of each of the methods, it was necessary to conduct convergence rate studies in order to get a complete picture.
+It is necessary for an analyst to know if the accuracy differences measured in the per-iteration tests would amount to differences in convergence rates among the methods if they were allowed to solve problems normally.
+The question is posed, given that step-size $h$ is heuristically selected by \emph{Peridigm} for the finite difference methods and presumable makes a good choice, do the single-iteration Jacobian accuracy results predict the ranking of the methods in terms of convergence rate for a given study?
+For brevity, convergence rate was measured for each of the methods in solving just the largest problem of the single-core test series.
+That problem was solved with each of the methods both on a single core and all sixteen cores of a 'normal' node on \emph{Stampede}.
+As was seen in the per-iteration results, Jacobian accuracy did not vary greatly from test to test, such that any pair of tests from the two series could have been considered reasonably representative.
+As in the validation problem shown in Section~\ref{subsec:Validation}, would each of the methods return an equal convergence rate so long as $h$ were optimally selected?
+
+The results shown in Table~\ref{tab:ConvergenceStudy2} indicate that this was the case.
+The difference in Jacobian accuracy between the methods seen in the per-iteration tests was not sufficient to lead to differences in the convergence rate of the methods.
+%
+\begin{table}[htb]   
+    \centering 
+    \caption{Average iterations / load-step} 
     \label{tab:ConvergenceStudy2}   
     \begin{tabular}{c c c c c}
-    \toprule Test type & AD & CS & CD & FD\\
+        \toprule 
+        Test type & AD & CS & CD & FD\\
         \midrule 
@@
         multi-core  & 63 & " & " & "\\ 
-\bottomrule \end{tabular} \end{table*}
-
-The end displacement results shown in ~\ref{subsec:Validation} indicated that up to a certain precision the
-solutions produced by each of the methods were identical. It was examined if this was the case for the 
-solutions produced in \emph{Peridigm}. Here each of the solutions produced by each of the 
-methods were compared to the AD produced solutions, under the assumption that the AD produced solutions
-would be closest to the solution produced by using an 'exact' Jacobian. Table ~\ref{tab:PeridigmSolutionAccuracy} shows that the methods
-did indeed produce differing solutions in terms of nodal force density, although only in decimal places at least $9$ orders 
-of magnitude smaller than the force density value at the node where the difference was detected.
-
-\begin{table*}[!ht]   
-    \centering \caption{Nodal force density maximum difference} 
+        \bottomrule 
+    \end{tabular} 
+\end{table}
+
+The end displacement results shown in Section~\ref{subsec:Validation} indicated that up to a certain precision the solutions produced by each of the methods were identical.
+It was examined if this was the case for the solutions produced in \emph{Peridigm}.
+Here each of the solutions produced by each of the methods were compared to the AD produced solutions, under the assumption that the AD produced solutions would be closest to the solution produced by using an 'exact' Jacobian.
+Table~\ref{tab:PeridigmSolutionAccuracy} shows that the methods did indeed produce differing solutions in terms of nodal force density, although only in decimal places at least $9$ orders of magnitude smaller than the force density value at the node where the difference was detected.
+%
+\begin{table}[hbp]   
+    \centering 
+    \caption{Nodal force density maximum difference} 
     \label{tab:PeridigmSolutionAccuracy}   
@@
         \midrule  Order of magnitude & $\num{1.E-8}$ & $\num{1.E-8}$ & $\num{1.E-5}$ \\ 
-\bottomrule \end{tabular} \end{table*}
-
-The solution accuracy results do follow the same ranking seen in the Jacobian accuracy results, however,
-again as in the validation problem the results differ only slightly. Speculatively this degree of error
-may be important to those running probabilistic simulations such that the propagation of error may be 
-a concern. Also it is conceivable that if this error is a fixed quantity, insensitive to material 
-properties or loading conditions, it may have a greater impact on simulations involving very low
-force densities such as for small displacements or very lax materials. In this case it could be recommended
-that forward difference be avoided.
-
+        \bottomrule 
+    \end{tabular} 
+\end{table}
+
+The solution accuracy results do follow the same ranking seen in the Jacobian accuracy results; however, again as in the validation problem the results differ only slightly.
+Speculatively this degree of error may be important to those running probabilistic simulations such that the propagation of error may be a concern.
+Also it is conceivable that if this error is a fixed quantity, insensitive to material properties or loading conditions, it may have a greater impact on simulations involving very low force densities such as for small displacements or very compliant materials.
+In this case it could be recommended that forward difference be avoided.
+%
 \section{\todo[]{Re: I.5}Conclusions}
@@
 \label{subsec:FurtherStudies}
-The question is raised; what is the significance of Jacobian estimation
-accuracy to the convergence of Newton's method?  The per-iteration and even
-convergence studies shown here indicate the effect of Jacobian accuracy for
-only a single datum, a very accurate Jacobian estimate, while the validation
-problem indicates the effect of using a grossly inaccurate Jacobian. The results here show that
-for very small Jacobian estimate inaccuracies, convergence rate is not
-affected, while for grossly inaccurate Jacobians Newton's method fails. 
-Additional confounding factors include that a Jacobian is only first
-order approximation of an often curved energy surface and that a non-naive implementation of Newton's
-method includes a line search phase that helps in choosing an optimal iterate given a Jacobian.
-Is it worthwhile then to investigate the effect of using moderately inaccurate Jacobians, 
-produced deliberately or otherwise, on the convergence rate of Newton's method?
-
-The answer may be no at least unless a more highly non-linear class of problem
-is selected for study or a controllable method of producing a moderately
-inaccurate Jacobian is found along with a reason for using it. While it is possible
-to incorrectly apply finite differencing and CS by choosing an inappropriate
-step-size, there is no reason that this should be done. The experiments show
-that step-size $h$ can be properly selected and that the methods are capable of
-producing accurate Jacobians. It should be noted that the model used to perform
-the simulation could always be substituted for a more highly non-linear one and
-perhaps this would somewhat increase the importance of Jacobian accuracy. The
-study presented here failed to pose the question of how wrong a correctly
-implemented Jacobian calculation scheme could go for plastic or otherwise
-non-linear material models, here instead non-linearity was imposed by the
-nonlocal interaction between material points essential to peridynamic models,
-like the linear model used in this study.
-
-Nevertheless, Newton's method is not defined for using an inaccurate or 'wrong'
-Jacobian, so using an inaccurate Jacobian is failing to perform Newton's method
-by definition. But it is also true that it should not be expected that every
-differentiation technique performs flawlessly under all situations. The
-pragmatic strategy to study this may be to compare a prospective method's
-Jacobian accuracy to that produced by AD and develop a rule of thumb based on
-observed convergence rates from prior simulations as they relate to the percent
-difference of the prospective and AD produced Jacobians. Also it is possible to
-calculate a sensitivity of Jacobian error to step-size alongside a sensitivity
-of convergence rate to step-size, and then to relate these two quantities as a
-map of Jacobian error to convergence rate change from ideal. Additionally it is
-still possible to look at the convergence conditions for Newton's method and
-suggest a relationship between Jacobian accuracy and convergence rate in a
-theoretical manner. To this point, there are a set of convergence conditions derived in
-\cite[Chap. 2, p.41]{burkeLectures} which relate quality of Jacobian inverse approximations
-to convergence rate classification, here reproduced,
+The question is raised; what is the significance of Jacobian estimation accuracy to the convergence of Newton's method?
+The per-iteration and even convergence studies shown here indicate the effect of Jacobian accuracy for only a single datum, a very accurate Jacobian estimate, while the validation problem of Section~indicates the effect of using a grossly inaccurate Jacobian.
+The results here show that for very small Jacobian estimate inaccuracies, convergence rate is not affected, while for grossly inaccurate Jacobians Newton's method fails.
+Additional confounding factors include that a Jacobian is only first order approximation of an often curved energy surface and that a non-naive implementation of Newton's method includes a line search phase that helps in choosing an optimal iterate given a Jacobian.
+Is it worthwhile then to investigate the effect of using moderately inaccurate Jacobians, produced deliberately or otherwise, on the convergence rate of Newton's method?
+
+The answer may be no at least unless a more highly non-linear class of problem is selected for study or a controllable method of producing a moderately inaccurate Jacobian is found along with a reason for using it.
+While it is possible to incorrectly apply finite differencing and CS by choosing an inappropriate step-size, there is no reason that this should be done.
+The experiments show that step-size $h$ can be properly selected and that the methods are capable of producing accurate Jacobians.
+It should be noted that the model used to perform the simulation could always be substituted for a more highly non-linear one and perhaps this would somewhat increase the importance of Jacobian accuracy.
+The study presented here failed to pose the question of how wrong a correctly implemented Jacobian calculation scheme could go for plastic or otherwise non-linear material models, here instead non-linearity was imposed by the nonlocal interaction between material points essential to peridynamic models, like the linear model used in this study.
+
+Nevertheless, Newton's method is not defined for using an inaccurate or 'wrong' Jacobian, so using an inaccurate Jacobian is failing to perform Newton's method by definition.
+But it is also true that it should not be expected that every differentiation technique performs flawlessly under all situations.
+The pragmatic strategy to study this may be to compare a prospective method's Jacobian accuracy to that produced by AD and develop a rule of thumb based on observed convergence rates from prior simulations as they relate to the percent difference of the prospective and AD produced Jacobians.
+Also it is possible to calculate a sensitivity of Jacobian error to step-size alongside a sensitivity of convergence rate to step-size, and then to relate these two quantities as a map of Jacobian error to convergence rate change from ideal.
+Additionally it is still possible to look at the convergence conditions for Newton's method and suggest a relationship between Jacobian accuracy and convergence rate in a theoretical manner.
+To this point, there are a set of convergence conditions derived in \cite[Chap.
+2, p.41]{burkeLectures} which relate quality of Jacobian inverse approximations to convergence rate classification, here reproduced,
 %

### 2014-04-24 jacobian_calc_CMAME.tex
Fixed validation section reference, some spelling

@@
 \section{\todo[]{Re: I.5}Simple example of CS, AD and FD for solution of nonlinear problems.}
-\label{sec:WorkingExamples}
+\label{subsec:Validation}
 
@@
 All the source code to reproduce the results described here are available in the associated paper repository.
-This source describes the implemenation and interface to the \emph{Trilinos Secado library} for automatic differentiation.
+This source describes the implementation and interface to the \emph{Trilinos Secado library} for automatic differentiation.
 
@@
 %
-For this system, the equalibrium reaction force associated with a displacement $D$ of the spring is shown in equation~(\ref{eqn:TrussForce}) 
+For this system, the equilibrium reaction force associated with a displacement $D$ of the spring is shown in equation~(\ref{eqn:TrussForce}) 
 %
@@
 The material model used in this study was a linear peridynamic solid which replicates linear elasticity in the \emph{local limit}, i.e.\ when the non-local length scale vanishes.
-While \emph{linear} may appear in the name of the material model, it is geometrically non-linear under large deformations and requires a nonlinear solver to converge to equalibrium conditions in these situations.
+While \emph{linear} may appear in the name of the material model, it is geometrically non-linear under large deformations and requires a nonlinear solver to converge to equilibrium conditions in these situations.
 The peridynamic equation of motion is shown in equation~(\ref{eqn:PDMotion})
@@
 The tangent-stiffness matrix is the Fr\'{e}chet derivative of the residual with respect to the deformed positions of the discrete computational nodes.
-For the purposes of this work, the residual represents the unbalanced forces in the peridynamic equalibrium equation.
+For the purposes of this work, the residual represents the unbalanced forces in the peridynamic equilibrium equation.
 For the finite difference type methods in \emph{Peridigm}, a perturbation method is employed by probing the deformed configuration one degree-of-freedom at a time and so allowing the derivative of the residual for all nodes to be computed.
@@
 These concepts are not vital to understanding the CS method itself, but may be necessary to understand CS in \emph{Peridigm}.
-To see CS and other methods in a more narrow context, it is advised that readers also take a look at the simple examples, such as that described in Section~\ref{sec:WorkingExamples},  provided on the corresponding author's website.
+To see CS and other methods in a more narrow context, it is advised that readers also take a look at the simple examples, such as that described in Section~\ref{subsec:Validation},  provided on the corresponding author's website.
 
@@
     \end{subfigure}
-    \caption{Peridynamic discritizations used in study.}
+    \caption{Peridynamic discretizations used in study.}
     \label{fig:pd_discrete}

### 2014-04-25 jacobian_calc_CMAME.tex
added displacement label to truss figure
L_0 labeled in the truss figure
Final major edits

@@
 %% keywords here, in the form: keyword \sep keyword
-Newton's method \sep Newton-Raphson \sep numerical differentiation \sep complex-step \sep finite difference \sep automatic differentiation \sep Jacobian \sep tangent-stiffness
+Newton's method \sep Newton-Raphson \sep numerical differentiation \sep complex-step \sep finite difference \sep automatic differentiation \sep tangent-stiffness \sep tangent-stiffness
 %% MSC codes here, in the form: \MSC code \sep code
@@
 A goal of this study was to develop and evaluate a \todo[]{Re: I.1} relatively unexplored, accurate, and practical method for calculating tangent-stiffness matrices against established methods.
-This  \todo[]{Re: I.1} method is based on a complex number Taylor series expansion and referred to as CTSE or the ``complex-step'' method.
-The practical accuracy of CTSE is determined by comparing its measured accuracy relative to the measured accuracy of popular finite differencing techniques used for computing tangent-stiffness matrices.
+This  \todo[]{Re: I.1} method is based on a complex number Taylor series expansion and referred to as the \emph{complex-step} (CS) method.
+The practical accuracy of CS is determined by comparing its measured accuracy relative to the measured accuracy of popular finite differencing techniques used for computing tangent-stiffness matrices.
 For this comparison the,  exact to machine precision, derivatives computed with \emph{automatic differentiation} serve as the standard.
 
-Comparative data regarding the accuracy and compute cycle timing for the
-complex step, forward difference, central difference, and automatic differentiation methods are included to aid developers and analysts in computational mechanics code design who are faced with implementing a general
-method for tangent-stiffness matrix calculation.  A comparison of the methods was achieved through instrumentation of a massively-parallel computational mechanics code.
+Comparative data regarding the accuracy and compute cycle timing for the complex step, forward difference, central difference, and automatic differentiation methods are included to aid developers and analysts in computational mechanics code design who are faced with implementing a general method for tangent-stiffness matrix calculation.
+A comparison of the methods was achieved through instrumentation of a massively-parallel computational mechanics code.
 
@@
 Papers by Newman et al.\ \cite{newman1998} and Anderson et al.\ \cite{anderson2001sensitivity} discuss computing sensitivities in a CFD analysis using CS and compares the results to sensitivities produced with central difference derivatives.
-The work is also notable for identifying automatic-differentiation as a competing method to CS and suggesting that complex-step could be used to compute Jacobians of residuals for use in Newton-Krylov schemes for solving non-linear systems.
+The work is also notable for identifying automatic-differentiation as a competing method to CS and suggesting that complex-step could be used to compute tangent-stiffnesss of residuals for use in Newton-Krylov schemes for solving non-linear systems.
 Perez-Foguet et al.\ \cite{perez2000numerical, perez2012numerical} used the CS method to compute algorithmically consistent elastoplastic moduli in non-linear solid mechanics problems.
@@
 In a follow-on work,  Millwater et al.\  \cite{millwater2013application} extended this concept, calling it ZFEM, and implemented the method into user defined element types for use with the popular ABAQUS \cite{systemes2012abaqus} FEA software package.
-Nishida et al.\ \cite{nishida2013} used the complex-step method to compute Jacobian matrices for implementation in a Levenberg-Marquardt optimization routine for design-of-experiments purposes.
+Nishida et al.\ \cite{nishida2013} used the complex-step method to compute tangent-stiffness matrices for implementation in a Levenberg-Marquardt optimization routine for design-of-experiments purposes.
 Finally, Abreu et al.\ \cite{abreu2013generalization} discusses computing derivatives with a version of the CS method which features the combined use of real and imaginary valued step components, as in Fourier differentiation.
@@
 Mathematically speaking, the tangent-stiffness represents the gradient of a high-dimensional energy surface that ``points'' in the minimum direction.
-While the terminology of solid mechanics is used here, it's important to note that
-these operators appear in other physical settings and are known by other names, e.g.
+While the terminology of solid mechanics is used here, it's important to note that these operators appear in other physical settings and are known by other names, e.g.
 a \emph{transmissibility matrix} in the context of a Poisson problem or, more generally, a \emph{Jacobian matrix} in mathematical optimization.
@@
 \centering %%\includegraphics[width=0.7\textwidth]{./figs/truss.png}
-
 \begin{tikzpicture}
@@
 \tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=6.1cm,minimum height=.2cm];
+\tikzstyle{displacement_reference}=[fill,pattern=north east lines,draw=none,minimum width=5.cm, minimum height=.6cm];
 \coordinate (bar_1) at (0,0);
@@
 \coordinate (mm) at (1.2, 0.0);
+\coordinate (tm) at (2.5, 2.0);
 \coordinate (rht) at (5.0, 4.0);
 \coordinate (rhtt) at (5.0, 4.2);
+\coordinate (dispref) at (6.5, 3.06);
+\coordinate (bmdisref) at (6.5, 3.0);
+\coordinate (bldisref) at (6.0,3.0);
+\coordinate (brdisref) at (7.0,3.0);
+\coordinate (bbdisref) at (6.5,2.5);
+\coordinate (bbrdisref) at (6.8,2.6);
+\node (displacement_reference) at (dispref) [rectangle, draw, style=displacement_reference, scale=.2]{};
 \node (wall_slider) at (bar_2_pin) [circular sector, draw, rotate=180] {};
@@
 \node (load_label) at (rhtt) [] {$P$};
+\node (bar_length_label) at (tm) [] {$L_0$};
 \node (angle_label) at ($ (mm)!.4cm!(bar_2) $) [] {$\phi$};
+\node (displacement_label) at (bbrdisref) [] {$D$};
 \path (bar_1) -- (bar_2) [draw, very thick];
@@
 \path (bar_1) -- (mm) [draw, -latex];
+\path (bldisref) -- (brdisref) [draw];
+\path (bmdisref) -- (bbdisref) [draw, -latex];
 --cycle
@@
 \end{equation} 
-
+%
 where, $\rho$ is the mass density, $\mathbf{\ddot{u}}$ is the second time derivative of displacement (i.e.\ acceleration), $\mathcal{H}_{\delta(\mathbf{x})}$ is a neighborhood containing remote material points, $\mathbf{x'}$, with characteristic length $\delta$, centered at $\mathbf{x}$.
@@
 \subsubsection{Data reduction} 
+%
 The output produced by \emph{Peridigm} including the additional accuracy and speed measurements was redirected from the console to text files by the resource manager, \emph{SLURM} (Simple Linux User Resource Manager).
@@
     \centering 
-    \scalebox{1.0}{\input{./figs/serial_speed.pgf}}
+    \scalebox{0.93}{\input{./figs/serial_speed.pgf}}
     \caption{Serial test series speed measurements.} 
@@
     \centering 
-    \scalebox{1.0}{\input{./figs/multi_speed.pgf}}
+    \scalebox{1.1}{\input{./figs/multi_speed.pgf}}
     \caption{Multicore test series speed measurements.} 
@@
     \centering
-    \scalebox{1.0}{\input{./figs/multi_accuracy.pgf}} 
+    \scalebox{0.9}{\input{./figs/multi_accuracy.pgf}} 
     \caption{Multicore test series accuracy measurements.} 
@@
 %
-Efficiency, calculated as the average number of seconds taken per nonzero Jacobian matrix element, was plotted as a function of the number of nonzero TS matrix elements.
+Efficiency, calculated as the average number of seconds taken per nonzero tangent-stiffness matrix element, was plotted as a function of the number of nonzero TS matrix elements.
 The single core results are shown in Figure~\ref{fig:serial_efficiency}.
@@
   \centering
-  \scalebox{1.0}{\input{./figs/serial_speed_rel.pgf}}
+  \scalebox{0.9}{\input{./figs/serial_speed_rel.pgf}}
   \caption{Serial test series efficiency measurements.}
@@
   \centering
-  \scalebox{1.0}{\input{./figs/multi_speed_rel.pgf}}
+  \scalebox{0.9}{\input{./figs/multi_speed_rel.pgf}}
   \caption{Multicore test series efficiency measurements.}
@@
 It is necessary for an analyst to know if the accuracy differences measured in the per-iteration tests would amount to differences in convergence rates among the methods if they were allowed to solve problems normally.
-The question is posed, given that step-size $h$ is heuristically selected by \emph{Peridigm} for the finite difference methods and presumable makes a good choice, do the single-iteration Jacobian accuracy results predict the ranking of the methods in terms of convergence rate for a given study?
+The question is posed, given that step-size $h$ is heuristically selected by \emph{Peridigm} for the finite difference methods and presumable makes a good choice, do the single-iteration tangent-stiffness accuracy results predict the ranking of the methods in terms of convergence rate for a given study?
 For brevity, convergence rate was measured for each of the methods in solving just the largest problem of the single-core test series.
-That problem was solved with each of the methods both on a single core and all sixteen cores of a 'normal' node on \emph{Stampede}.
-As was seen in the per-iteration results, Jacobian accuracy did not vary greatly from test to test, such that any pair of tests from the two series could have been considered reasonably representative.
+That problem was solved with each of the methods both on a single core and all sixteen cores of a  node on \emph{Stampede}.
+As was seen in the per-iteration results, tangent-stiffness accuracy did not vary greatly from test to test, such that any pair of tests from the two series could have been considered reasonably representative.
 As in the validation problem shown in Section~\ref{subsec:Validation}, would each of the methods return an equal convergence rate so long as $h$ were optimally selected?
@@
 The results shown in Table~\ref{tab:ConvergenceStudy2} indicate that this was the case.
-The difference in Jacobian accuracy between the methods seen in the per-iteration tests was not sufficient to lead to differences in the convergence rate of the methods.
+The difference in tangent-stiffness accuracy between the methods seen in the per-iteration tests was not sufficient to lead to differences in the convergence rate of the methods.
 %
@@
 It was examined if this was the case for the solutions produced in \emph{Peridigm}.
-Here each of the solutions produced by each of the methods were compared to the AD produced solutions, under the assumption that the AD produced solutions would be closest to the solution produced by using an 'exact' Jacobian.
+Here each of the solutions produced by each of the methods were compared to the AD produced solutions, under the assumption that the AD produced solutions would be closest to the solution produced by using an 'exact' tangent-stiffness.
 Table~\ref{tab:PeridigmSolutionAccuracy} shows that the methods did indeed produce differing solutions in terms of nodal force density, although only in decimal places at least $9$ orders of magnitude smaller than the force density value at the node where the difference was detected.
@@
 
-The solution accuracy results do follow the same ranking seen in the Jacobian accuracy results; however, again as in the validation problem the results differ only slightly.
+The solution accuracy results do follow the same ranking seen in the tangent-stiffness accuracy results; however, again as in the validation problem the results differ only slightly.
 Speculatively this degree of error may be important to those running probabilistic simulations such that the propagation of error may be a concern.
@@
 %
-\subsection{For further studies of Jacobian accuracy in Newton's method}
-\label{subsec:FurtherStudies}
-The question is raised; what is the significance of Jacobian estimation accuracy to the convergence of Newton's method?
-The per-iteration and even convergence studies shown here indicate the effect of Jacobian accuracy for only a single datum, a very accurate Jacobian estimate, while the validation problem of Section~indicates the effect of using a grossly inaccurate Jacobian.
-The results here show that for very small Jacobian estimate inaccuracies, convergence rate is not affected, while for grossly inaccurate Jacobians Newton's method fails.
-Additional confounding factors include that a Jacobian is only first order approximation of an often curved energy surface and that a non-naive implementation of Newton's method includes a line search phase that helps in choosing an optimal iterate given a Jacobian.
-Is it worthwhile then to investigate the effect of using moderately inaccurate Jacobians, produced deliberately or otherwise, on the convergence rate of Newton's method?
-
-The answer may be no at least unless a more highly non-linear class of problem is selected for study or a controllable method of producing a moderately inaccurate Jacobian is found along with a reason for using it.
+%Let's leave this section out and see what the reviewer says.  The paper is getting pretty long.
+%
+%\subsection{For further studies of tangent-stiffness accuracy in Newton's method}
+%\label{subsec:FurtherStudies}
+%The question is raised; what is the significance of tangent-stiffness estimation accuracy to the convergence of Newton's method?
+The per-iteration and even convergence studies shown here indicate the effect of tangent-stiffness accuracy for only a single datum, a very accurate tangent-stiffness estimate, while the validation problem of Section~XX indicates the effect of using a grossly inaccurate tangent-stiffness.
+The results here show that for very small tangent-stiffness estimate inaccuracies, convergence rate is not affected, while for grossly inaccurate tangent-stiffnesss Newton's method fails.
+Additional confounding factors include that a tangent-stiffness is only first order approximation of an often curved energy surface and that a non-naive implementation of Newton's method includes a line search phase that helps in choosing an optimal iterate given a tangent-stiffness.
+Is it worthwhile then to investigate the effect of using moderately inaccurate tangent-stiffnesss, produced deliberately or otherwise, on the convergence rate of Newton's method?
+
+%The answer may be no at least unless a more highly non-linear class of problem is selected for study or a controllable method of producing a moderately inaccurate tangent-stiffness is found along with a reason for using it.
 While it is possible to incorrectly apply finite differencing and CS by choosing an inappropriate step-size, there is no reason that this should be done.
-The experiments show that step-size $h$ can be properly selected and that the methods are capable of producing accurate Jacobians.
-It should be noted that the model used to perform the simulation could always be substituted for a more highly non-linear one and perhaps this would somewhat increase the importance of Jacobian accuracy.
-The study presented here failed to pose the question of how wrong a correctly implemented Jacobian calculation scheme could go for plastic or otherwise non-linear material models, here instead non-linearity was imposed by the nonlocal interaction between material points essential to peridynamic models, like the linear model used in this study.
-
-Nevertheless, Newton's method is not defined for using an inaccurate or 'wrong' Jacobian, so using an inaccurate Jacobian is failing to perform Newton's method by definition.
+The experiments show that step-size $h$ can be properly selected and that the methods are capable of producing accurate tangent-stiffnesss.
+It should be noted that the model used to perform the simulation could always be substituted for a more highly non-linear one and perhaps this would somewhat increase the importance of tangent-stiffness accuracy.
+The study presented here failed to pose the question of how wrong a correctly implemented tangent-stiffness calculation scheme could go for plastic or otherwise non-linear material models, here instead non-linearity was imposed by the nonlocal interaction between material points essential to peridynamic models, like the linear model used in this study.
+
+%Nevertheless, Newton's method is not defined for using an inaccurate or 'wrong' tangent-stiffness, so using an inaccurate tangent-stiffness is failing to perform Newton's method by definition.
 But it is also true that it should not be expected that every differentiation technique performs flawlessly under all situations.
-The pragmatic strategy to study this may be to compare a prospective method's Jacobian accuracy to that produced by AD and develop a rule of thumb based on observed convergence rates from prior simulations as they relate to the percent difference of the prospective and AD produced Jacobians.
-Also it is possible to calculate a sensitivity of Jacobian error to step-size alongside a sensitivity of convergence rate to step-size, and then to relate these two quantities as a map of Jacobian error to convergence rate change from ideal.
-Additionally it is still possible to look at the convergence conditions for Newton's method and suggest a relationship between Jacobian accuracy and convergence rate in a theoretical manner.
+The pragmatic strategy to study this may be to compare a prospective method's tangent-stiffness accuracy to that produced by AD and develop a rule of thumb based on observed convergence rates from prior simulations as they relate to the percent difference of the prospective and AD produced tangent-stiffnesss.
+Also it is possible to calculate a sensitivity of tangent-stiffness error to step-size alongside a sensitivity of convergence rate to step-size, and then to relate these two quantities as a map of tangent-stiffness error to convergence rate change from ideal.
+Additionally it is still possible to look at the convergence conditions for Newton's method and suggest a relationship between tangent-stiffness accuracy and convergence rate in a theoretical manner.
 To this point, there are a set of convergence conditions derived in \cite[Chap.
-2, p.41]{burkeLectures} which relate quality of Jacobian inverse approximations to convergence rate classification, here reproduced,
-%
-\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq K,
-\end{equation}
-
-for worse than linear convergence,
-
-\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq \Theta_{1}^{k}K; \Theta_{1} \in (0,1),
-\end{equation}
-
-for linear convergence,
-
-\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq min{M_{2} \mid \mid x_{k} - x_{k-1} \mid \mid, K}; M_{2} > 0
-\end{equation}
-
-for two step quadratic,
-
-\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq min{M_{3} \mid \mid r(x_{k}) \mid \mid, K}; M_{3} > 0
-\end{equation}
-
-and lastly for quadratic convergence. Here $r \prime$ is the true system
-Jacobian and $J$ is the approximation. Essentially the conditions measure how
-well the difference of the true Jacobian inverse and approximation works as
-a contraction map within the closure of a radius of convergence about the true
-solution. This property is necessary for the solution method using the
-approximation to converge rather than to diverge, as is true of the exact Newton
-method.
-
-It may be interesting to artificially perturb the Jacobian calculations with
-random error in order to better control experimentally the onset of inaccuracy
-to develop a sensitivity of convergence rate to this type of error. Perhaps an
-analyst could use this knowledge to devise an adaptive differentiation
-technique selection method which economizes on accuracy until a significant
-Jacobian error is measured during a special doubly calculated iteration.
-
-Additionally, there are Quasi-Newton methods such as Broyden's method which
-uses an approximate inverse Jacobian and updates it with information from the 
-iterate and residual \cite{dennis1971convergence}. A method called
-Incomplete Jacobian Newton's method may the closest to the idea of deliberately
-making use of an inaccurate Jacobian in the sense that arbitrary elements of
-the full Jacobian are set to zero to reduce computational effort
-\cite{liu2008incomplete}. The Quasi-Newton methods represent a compromise of
-Newton's method for the promise of faster iterations in exchange for a lower
-convergence rate. For both of these methods, the iterate and residual are
-computed as in Newton's method. As to the question of algorithmic consistency,
-these numerical methods evaluate the same linearized constitutive law and
-governing equation as for the exact Newton's method which in the case of a
-plastic simulation are involved in iteratively determining algorithmic tangent moduli. 
-In fact, a secant equation based method is appropriate for Newton's method with algorithmic
-tangent modulii since the Jacobian ought only depend on known converged
-values such that the Newton iterate could not artificially drive the plasticity
-of the model \cite[Chap. 6]{belytschko1999nonlinear}. It should be noted that
-constraints or a line search algorithm to minimize the residual are necessary
-whereas the pure secant method does not return unique solutions in
-multi-dimensions.
-%
-
-\subsection{Grain of salt for speed results}
-The per-iteration speed results shown here indicate that the AD method is
-faster than CD or FD.  While this may be the case for the implementation in the
-stock version of \emph{Peridigm} this is not reflective of any limitation of
-these methods. In a side experiment, the finite difference code in
-\emph{Peridigm} was re-written to take advantage of the benefits of storing the
-results of duplicated intermediate calculations. These modifications were 
-general, low memory footprint and could be adopted in any numerical simulation
-simulation software as good practice. An informal performance test was done
-using a personal AMD Phenom-2 X4 workstation using GCC 4.7 compilers rather than a
-Intel Xeon E5 node with Intel Cluster Studio 13.0 compilers as provided by TACC. The problem
-being run was the largest single core test from the main study, instead run on four 
-compute cores. Table ~\ref{tab:PerformanceMod} shows the timing
-results for AD as reference along with an unmodified CD, and then modified CD
-and FD.
-
-\begin{table*}[!ht]   
-\centering \caption{Total time spent computing Jacobian} 
+2, p.41]{burkeLectures} which relate quality of tangent-stiffness inverse approximations to convergence rate classification, here reproduced,
+%%
+%\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq K,
+%\end{equation}
+
+%for worse than linear convergence,
+
+%\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq \Theta_{1}^{k}K; \Theta_{1} \in (0,1),
+%\end{equation}
+
+%for linear convergence,
+
+%\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq min{M_{2} \mid \mid x_{k} - x_{k-1} \mid \mid, K}; M_{2} > 0
+%\end{equation}
+
+%for two step quadratic,
+
+%\begin{equation} \mid \mid (r \prime (x_{k})^{-1} - J_{k}^{-1})y_{k} \mid \mid \leq min{M_{3} \mid \mid r(x_{k}) \mid \mid, K}; M_{3} > 0
+%\end{equation}
+
+%and lastly for quadratic convergence. Here $r \prime$ is the true system
+%tangent-stiffness and $J$ is the approximation. Essentially the conditions measure how
+%well the difference of the true tangent-stiffness inverse and approximation works as
+%a contraction map within the closure of a radius of convergence about the true
+%solution. This property is necessary for the solution method using the
+%approximation to converge rather than to diverge, as is true of the exact Newton
+%method.
+
+%It may be interesting to artificially perturb the tangent-stiffness calculations with
+%random error in order to better control experimentally the onset of inaccuracy
+%to develop a sensitivity of convergence rate to this type of error. Perhaps an
+%analyst could use this knowledge to devise an adaptive differentiation
+%technique selection method which economizes on accuracy until a significant
+%tangent-stiffness error is measured during a special doubly calculated iteration.
+
+%Additionally, there are Quasi-Newton methods such as Broyden's method which
+%uses an approximate inverse tangent-stiffness and updates it with information from the 
+%iterate and residual \cite{dennis1971convergence}. A method called
+%Incomplete tangent-stiffness Newton's method may the closest to the idea of deliberately
+%making use of an inaccurate tangent-stiffness in the sense that arbitrary elements of
+%the full tangent-stiffness are set to zero to reduce computational effort
+%\cite{liu2008incomplete}. The Quasi-Newton methods represent a compromise of
+%Newton's method for the promise of faster iterations in exchange for a lower
+%convergence rate. For both of these methods, the iterate and residual are
+%computed as in Newton's method. As to the question of algorithmic consistency,
+%these numerical methods evaluate the same linearized constitutive law and
+%governing equation as for the exact Newton's method which in the case of a
+%plastic simulation are involved in iteratively determining algorithmic tangent moduli. 
+%In fact, a secant equation based method is appropriate for Newton's method with algorithmic
+%tangent modulii since the tangent-stiffness ought only depend on known converged
+%values such that the Newton iterate could not artificially drive the plasticity
+%of the model \cite[Chap. 6]{belytschko1999nonlinear}. It should be noted that
+%constraints or a line search algorithm to minimize the residual are necessary
+%whereas the pure secant method does not return unique solutions in
+%multi-dimensions.
+%
+
+\subsection{Remarks on speed results}
+%
+The per-iteration speed results shown here indicate that the AD method is faster than CD or FD.
+While this may be the case for the implementation in the release version of \emph{Peridigm} this is not reflective of any limitation of these methods.
+In a side experiment, the finite difference code in \emph{Peridigm} was re-written to take advantage of the benefits of storing the results of duplicated intermediate calculations.
+These modifications were general, low memory footprint, and could be adopted in any numerical simulation simulation software as good practice.
+An informal performance test run was performed on the largest single core test from the main study, instead run on four compute cores.
+Table~\ref{tab:PerformanceMod} shows the timing results for AD as reference along with an unmodified CD, and then modified CD and FD.
+%
+\begin{table}[tbp]   
+\centering 
+\caption{Total time spent computing tangent-stiffness} 
 \label{tab:PerformanceMod}   
@@
 \midrule  Total seconds & $~430$ & $~570$ & $~430$ &$~320$ \\ 
-\bottomrule \end{tabular} \end{table*}
-
-The results indicated that in exchange for a modest programming effort, a
-performance increase of around a quarter was available for the
-finite difference methods and that FD could eclipse AD. What allowed the performance modifications to
-enhance the finite difference methods was their lack of sophistication
-compared to AD in that their explicitly laid out and
-repetitive code could be readily optimized.
-
-\subsection{Remarks}
-Based on the computational experiments, it appears that for users of \emph{Peridigm}
-there is a case to be made for using AD for reasons of speed. However, certain
-parallel tests done in this study showed a small performance gain over AD with
-the FD method.  The results showed that CS produced more accurate
-tangent-stiffness matrices than CD and FD under the parameters of the tests,
-however it was determined that this difference in accuracy was not sufficient
-to alter convergence rate for a test problem. In order for Jacobian accuracy to
-have an impact on convergence rate, and therefor serve as a predictor of
-convergence rate in per-iteration results, the amount of error detected between
-a supposed perfectly accurate Jacobian and the test Jacobian needs to be
-greater than allowed by the conditions seen in \cite{burkeLectures}, 
-reproduced in subsection~\ref{subsec:FurtherStudies}.
-
-It should be mentioned that another work, \cite{perez2012numerical}, compares
-the speed and accuracy of complex-step to finite difference for producing
-tangent-stiffness matrices in solid mechanics simulations using the 'MRS-Lade'
-material model, where the researchers similarly found CS to be accurate yet
-relatively computationally expensive compared to FD and CD. In that work, full
-convergence studies were also done and indicated that quadratic convergence was
-achieved with the methods \cite[p.28]{perez2012numerical}.
-
-However, at least for the models producing the results shown here, each of the
-methods returned quadratic convergence, indicating each was sufficiently
-accurate. 'Marching orders' for the analyst would be to select the swiftest
-method, while it would seem accuracy above what is adequate does not enhance
-the convergence rate or iteration speed until either classical Newton's convergence
-conditions or the approximate case conditions are violated.
-
-The CS method was shown to be highly accurate in calculating tangent-stiffness
-matrices by the standards of this study, and it is slower than other methods
-for a naive implementation. In the context of future use of CS in
-\emph{Peridigm}, CS holds the advantage of relying on byte-copyable data types,
-while AD requires a complicated serialization and deserialization in order to
-function as byte-copyable. This is fine for copying in between MPI processes,
-but deserialization on an accelerator or GPU is very difficult, as this would
-require compiling a version of the AD library being used that were compatible
-with the accelerator or GPU being used. A working definition of byte-copyable
-is that the collection of data to transfer is made of identically sized pieces.
-Byte-copyability is significant for example because it is a requirement for the
-basic offload mode provided for Intel MIC accelerator boards which are
-installed in \emph{Stampede}, where the alternative mode requires "more
-extensive" effort \cite{intel_byte_copyable}. What this means is that if
-\emph{Peridigm} or any simulation software were to be reformulated to take
-advantage of accelerators or GPUs, the flexibility of complex-step owed to its
-simple implementation may make it more viable for that application given its
-high accuracy. Additional performance gains may be achieved through the use of
-the templated \emph{TPetra} package in \emph{Trilinos} which would allow the
-use of complex data types natively, and alleviate the expense of the
-dynamic-casts that were utilized in this study.  As a final comment, it should
-be noted that from a code development perspective the CS method may be the
-easiest of the methods to implement; it only requires one functional
-evaluation, there is no heuristic choice of probe distance, and there is no
-dependence on an external AD library.  From this perspective, the method may
-still be an attractive method for TS matrix evaluation. However,
-at least for the models producing the results shown here, each of the
-methods returned quadratic convergence, indicating each was sufficiently
-accurate. 'Marching orders' for the analyst would be to select the swiftest
-method, while it would seem accuracy above what is adequate does not
-enhance the convergence rate or iteration speed until either classical
-Newton's convergence conditions or the approximate case conditions are
-violated.
+\bottomrule 
+\end{tabular} 
+\end{table}
+
+The results indicated that in exchange for a modest programming effort, a performance increase of around a quarter was available for the finite difference methods and that FD could eclipse AD.
+What allowed the performance modifications to enhance the finite difference methods was their lack of sophistication compared to AD in that their explicitly laid out and repetitive code could be readily optimized.
+
+\subsection{Final Remarks}
+%
+Based on the computational experiments, it appears that for users of \emph{Peridigm} there is a case to be made for using AD for reasons of speed.
+However, certain parallel tests done in this study showed a small performance gain over AD with the FD method.
+The results showed that CS produced more accurate tangent-stiffness matrices than CD and FD under the parameters of the tests; however, it was determined that this difference in accuracy was not sufficient to alter convergence rate for a test problem.
+%In order for tangent-stiffness accuracy to have an impact on convergence rate, and therefore serve as a predictor of convergence rate in per-iteration results, the amount of error detected between a supposed perfectly accurate tangent-stiffness and the test tangent-stiffness needs to be greater than allowed by the conditions seen in \cite{burkeLectures}, reproduced in subsection~\ref{subsec:FurtherStudies}.
+
+The CS method was shown to be highly accurate in calculating tangent-stiffness matrices by the standards of this study, and it is slower than other methods for a naive implementation.
+In the context of future use of CS in \emph{Peridigm}, CS holds the advantage of relying on byte-copyable\footnote{In this context,  byte-copyable means that the collection of data to transfer is made of identically sized pieces.} data types, while AD requires a complicated serialization and deserialization in order to function as byte-copyable.
+This is fine for copying in between MPI processes, but deserialization on an accelerator or GPU is very difficult, as this would require compiling a version of the AD library being used that were compatible with the accelerator or GPU being used.
+Byte-copyability is significant for example because it is a requirement for the basic offload mode provided for Intel MIC accelerator boards which are installed in \emph{Stampede}, where the alternative mode requires more extensive effort \cite{intel_byte_copyable}.
+What this means is that if \emph{Peridigm} or any simulation software were to be reformulated to take advantage of accelerators or GPUs, the flexibility of complex-step owed to its simple implementation may make it more viable for that application given its high accuracy.
+Additional performance gains may be achieved through the use of the templated \emph{TPetra} package in \emph{Trilinos} which would allow the use of complex data types natively, and alleviate the expense of the dynamic-casts that were utilized in this study.
+As a final comment, it should be noted that from a code development perspective the CS method may be the easiest of the methods to implement; it only requires one functional evaluation, there is no heuristic choice of probe distance, and there is no dependence on an external AD library.
+From this perspective, the method may still be an attractive method for tangent-stiffness matrix evaluation.
+However, at least for the models producing the results shown here, each of the methods returned quadratic convergence, indicating each was sufficiently accurate.
+'Marching orders' for the analyst would be to select the swiftest method, while it would seem accuracy above what is adequate does not enhance the convergence rate or iteration speed until either classical Newton's convergence conditions or the approximate case conditions are violated.
 
@@
 The authors acknowledge the Texas Advanced Computing Center (TACC) at The University of Texas at Austin for providing HPC resources that have contributed to the research results reported within this paper.
-URL: http://www.tacc.utexas.edu
+URL: \url{http://www.tacc.utexas.edu}.
 %% References

### 2014-04-25 jacobian_calc_CMAME.tex
Small edits

@@
 
-\section{Introduction}
+\section{Introduction.}
 %\subsection{Motivation}
@@
   \centering
-        \caption{Displacement, D, after load-step 12} \label{tab:Verification}   
+  \caption{Displacement, D, after load-step 12.} \label{tab:Verification}   
         \begin{tabular}{c c c c c}
@@
   \centering
-        \caption{Summed iterations over all load steps} \label{tab:ConvergenceStudy}   
+  \caption{Summed iterations over all load steps.} \label{tab:ConvergenceStudy}   
         \begin{tabular}{c c c c c}
@@
 
-\section{Description of analysis tools and approach} 
-%
-\subsection{Peridigm and peridynamics}
+\section{Description of analysis tools and approach.} 
+%
+\subsection{Peridigm and peridynamics.}
 \label{subsec:PaP}
@@
   \centering
-        \caption{Test Parameters} \label{tab:TestParams}   
+  \caption{Test Parameters.} \label{tab:TestParams}   
        \begin{tabular}{c c c c c c}
@@
 
-\section{Results and discussion}
+\section{Results and discussion.}
 \label{sec:Res}
@@
   \centering
-        \caption{Averaged Results, Each Test} \label{tab:results}   
+  \caption{Averaged Results, Each Test.} \label{tab:results}   
        \begin{tabular}{c c c c c c c c c}
@@
 %
-\section{\todo[]{Re: I.5}Conclusions}
+\section{\todo[]{Re: I.5}Conclusions.}
 \label{sec:Conc}
@@
 
-\section{Acknowledgements}
+\section{Acknowledgements.}
 \label{sec:ack}
@@
 
-\section{References}
+\section{References.}
 
@@
 %
-\section{Complex-Step Example}
+\section{Complex-Step Example.}
 \label{sec:appendixA}
@@
 
-\section{Automatic differentiation example}
+\section{Automatic differentiation example.}
 \label{sec:appendixB}

### 2014-04-25 response/response.tex
AS SUBMITTED FOR REVISION

@@
 \begin{enumerate}
-    \item
-         Some parts of the manuscript suggest that the authors introduced a new method, i.e., the complex-step (CS) method.
+    \item Some parts of the manuscript suggest that the authors introduced a new method, i.e., the complex-step (CS) method.
 For example, in the introduction, they say ``A goal of this study was to develop and evaluate a new, accurate, and practical method ...'' Another example appears in page 3, where the authors say ``The aim of this paper is not only to introduce the new complex-step method in the context of evaluating tangent-stiffness matrices ...'' However, the CS method has already appeared in many publications since 1998 [1].
@@
     \item 
-        We have revised the brief description of peridynamics in page 9 to be more complete:
-    \item
-        Material from silling2007 is interpreted including: the definition of vector states,
-        the derivation of conservation laws in terms of state-based peridynamics, the definition
-        of the isotropic elastic solid AKA the Linear elastic Peridynamic Solid.
-    \item
-        It is the authors' opinion that the additional material mentioned above explains the equations involved
-        in the peridynamic models for mechanics, used in relation to the peridynamics code
-        studied in the work.
-
-  \end{itemize}}
+        We have revised the brief description of peridynamics in page 9 to be more complete.
+  \end{itemize}}
+
     \item 
@@
   \begin{itemize}
-    \item 
-        We have revised certain parts of the manuscript for clarity, adding additional information for the reader's
-        convenience:
-    \item
-        In subsection 'Implementing the complex-step method in Peridigm', explanation of what is meant by datatype,
-        container class, template, scope and instantiation are included since these concepts may be unfamiliar to
-        a computational mechanics audience at least as denoted by these terms. 
-    \item 
-        In subsection 'The Comparative Study', the way message passing computations can slow runtime is explained
-        by an analogy to an experiment given as an exercise for the reader. 
-    \item 
-        In section 'Conclusions', the term byte-copiable is explained. 
-    \item 
-        It is the authors' opinion that the above revisions to the language of the manuscript allow the intended audience
-        to follow the discussion where concepts from computer science are relevant.
-  \end{itemize}}
-
-  \item
-     In Section 1.1.1, please include a more comprehensive literature review for
+    \item We have revised certain parts of the manuscript for clarity, adding additional information for the reader's convenience.
+  \end{itemize}
+  }
+
+  \item In Section 1.1.1, please include a more comprehensive literature review for
      the CS method, including further references (some of them are listed below).
@@
   \begin{itemize}
-     \item 
-         The authors have appended a more comprehensive literature review for the CS method to section 1.1.1.
-         The aspects of CS investigated by the prior work, as well as any particularly interesting
-         developments introduced by those works, have been listed. To argue for the novelty of the instant work,
-         each prior work is compared to the instant work in terms of the use of the CS method and the similarity of the fields of research of the works.
-     \item
-         The literature review concludes with an argument for the novelty of the instant
-         work where the prior works are taken as a whole and compared to the instant work. It is the authors' opinion
-         that the instant work is novel for at least the reasons set out in the literature review material, where
-         the contribution to the CS method is in the comparison to other methods in terms of computational
-         time,  parallel implementation, and implementation within \emph{Peridigm} .
+     \item The authors have appended a more comprehensive literature review for the CS method to section 1.1.1.
+The aspects of CS investigated by the prior work, as well as any particularly interesting developments introduced by those works, have been listed.
+The literature review concludes with an argument for the novelty of this work.
   \end{itemize}}
@@
     \item 
-        To address the reviewers comment, material from work done after the
-        initial submission of the manuscript has been included. New sections
-        include 'Working examples of CS, AD and FD for Newton's method' with
+        To address the reviewers comment, additional simulation results are now included. New sections
+        include "Simple example of CS, AD and FD for solution of nonlinear problems" with
         new subsections. To existing sections were appended new subsections
-        including 'Convergence and displacement accuracy studies', 'For further
-        studies of Jacobian accuracy in Newton's method', 'Grain of salt for
-        speed results' and 'Remarks'. Specifically, the added material addresses
-        the reviewers comments:
+        including 'Convergence and displacement accuracy studies', 'Remarks for
+        speed results' and 'Final Remarks'.
     \end{itemize}}
@@
     \item 
-        The added section 'Working examples of CS, AD and FD for Newton's method' includes
-        the two examples alluded to in the manuscript and requested by the reviewer. 
-        % 
-        To summarize, in this section is an example of interest: a displacement accuracy study and 
-        a convergence rate study on a non-linear beam problem, where each of the methods
-        discussed in the manuscript are used individually to solve the problem. 
+        The added section "Simple example of CS, AD and FD for solution of nonlinear problems" includes one of two examples (the second was omitted for brevity) alluded to in the manuscript and requested by the reviewer.
+To summarize, in this section is an example of interest: a displacement accuracy study and a convergence rate study on a non-linear beam problem, where each of the methods discussed in the manuscript are used individually to solve the problem.
  \end{itemize}}
@@
       includes convergence rate and displacement accuracy studies similar to
-      those shown in 'Working examples ...', but with the models and tools of
+      those shown in `Simple example ...', but with the models and tools of
       the main study. This subsection provides an example of interest. 
-  \item 
-      To summarize, these studies are done in \emph{Peridigm} using the most refined mesh
-      from the 'serial' test runs from the original manuscript. Measured was the performance of the methods in terms of speed
-      (convergence rate) and accuracy (displacement accuracy).  Iteration speed
-      was reported in the original manuscript.
-  \item
-      It was determined that the accuracy of CS over FD resulted in no appreciable
-      advantage in terms of displacement accuracy or convergence rate for the problems
-      solved in this study. 
-    \item
-      Additional comment on the significance of the accuracy of the Jacobian
-      to Newton's method is included in the subsections 'For further studies of
-      Jacobian accuracy in Newton's method', 'Grain of salt for speed results'
-      and 'Remarks'. 
-  \item 
-      To summarize, Newton's method is insensitive to the accuracy differences
-      shown by the methods, for the problems shown, when the methods are not
-      deliberately misused with inappropriate step sizes. For completeness,
-      formulas describing a relationship between Jacobian accuracy and
-      convergence rate order are included. 
-  \item
-      It is the authors' opinion that the above revisions offer the discussion of relative problem solving
-      performance of the methods that the reviewer has requested, including consideration
-      of speed and accuracy in solving real, non-linear problems.
-  \end{itemize}}
-
+  \end{itemize}}
   \end{enumerate}
@@
 
-      The word, 'discrimination' has been removed from the manuscript.
-      'Discrimination' was meant to mean 'measuring differences and making
-      judgements on relative usefulness'. Since this meaning is probably
-      understood from the word 'comparison' alone, the word 'discrimination'
+      The word, `discrimination' has been removed from the manuscript.
+      'Discrimination' was meant to mean ``measuring differences and making
+      judgements on relative usefulness''. Since this meaning is probably
+      understood from the word `comparison' alone, the word `discrimination'
       has been removed from the manuscript for clarity.
@@
 ` \item	
-	Pg. 2: In line 37, explain what do you mean by “in-situ instrumentation”.
+	Pg. 2: In line 37, explain what do you mean by ``in-situ instrumentation''.
 
@@
   \item 
-Pg. 4: In Eq. (1), “$\partial f^2$” should be “$\frac{\partial^2 f}{\partial x^2}$”; similarly, for the third derivative.
+Pg. 4: In Eq. (1), ``$\partial f^2$'' should be ``$\frac{\partial^2 f}{\partial x^2}$''; similarly, for the third derivative.
 
@@
 \item
-Pg. 6: In lines 9-11, please add references to the “Poisson problem” and to the
-“mathematical optimization”.
-
-{\color{red}  
-\begin{itemize}
-     \item
-         [TODO]
-
-  \end{itemize}}
-
-\item
-Pg. 7: Should Eqs. (4), (5), and (6) use “$\approx$” instead of “$=$”?
+Pg. 6: In lines 9-11, please add references to the ``Poisson problem'' and to the
+``mathematical optimization''.
+
+{\color{red}  
+\begin{itemize}
+     \item
+      The correction is made in the manuscript.
+
+  \end{itemize}}
+
+\item
+Pg. 7: Should Eqs. (4), (5), and (6) use ``$\approx$'' instead of “$=$”?
 
@@
 Pg. 7: In line 28, note that (cid:126)u is not the current conﬁguration, but the displacement
-ﬁeld. Please revise the sentence.
+field. Please revise the sentence.
 
@@
 \item
-Pg. 7: Please provide an expression for ${F^int}_i$ in the model used.
-
-{\color{red}  
-\begin{itemize}
-    \item
-        [TODO]
-        Manuscript corrected.
-In the section 'Peridigm and peridynamics' a constitutive model for the LPS is shown.
-On the page that was originally page 7, the force density corresponding to this constitutive model is expressed in descrete form.
-  \end{itemize}}
-
-\item
-Pg. 9: In line 26, please add the reference [12] below to “coupling to molecular
-simulations [29].”
+Pg. 7: Please provide an expression for $F^{int}_i$ in the model used.
+
+{\color{red}  
+\begin{itemize}
+    \item This is explained in words in Section 3.1.1
+  \end{itemize}}
+
+\item
+Pg. 9: In line 26, please add the reference [12] below to ``coupling to molecular
+simulations [29].''
 
@@
     \item
-        The manuscript has been corrected.
-Bandwidth was referring to connections between nodes rather than components and degrees of freedom.
-It is more clear to say a bandwidth of 21 for the 3D case.
-The formula is number of nodes in a neighborhood times degrees of freedom.
-  \end{itemize}}
-
-    \item
-        In lines 47-49, you say “The speciﬁc parameters ... can be found ... in the
-        paper repository.” It seems a good idea to have the paper repository with ﬁles that a
+        This statement has been removed from the manuscript, it will not hold in 3D.
+  \end{itemize}}
+
+    \item
+        In lines 47-49, you say ``The speciﬁc parameters ... can be found ... in the
+        paper repository.'' It seems a good idea to have the paper repository with ﬁles that a
         user can eventually use. However, the content of the paper should be independent
@@
      \item
-         Manuscript is corrected. Mesh sizes from smallest to largest were 1000, 2000, 3000, 4000, 6000, 8000, 12000,
-         16000, 32000, 1000000 peridynamic nodes.
+         Manuscript is corrected. Mesh sizes from smallest to largest were 1000, 2000, 3000, 4000, 6000, 8000, 12000, 16000, 32000, 1000000 peridynamic nodes.
       \end{itemize}}
@@
     %
-    The probe distances used in Peridigm are automatically determined with a
-    heuristic algorithm and are not set by the user. The algorithm for this in
-    \emph{Peridigm} is the used for another Sandia software project called
-    \emph{Adagio} \cite{ref-Adaggio}. The probe distances are 1.E-6 times the node spacing for FD and CD which is indended to prevent the onset of roundoff error.
+    The probe distances used in Peridigm are automatically determined with a heuristic algorithm and are not set by the user.
+The algorithm for this in \emph{Peridigm} is the used for another Sandia software project called \emph{Adagio} \cite{ref-Adaggio}.
+The probe distances are 1.E-6 times the node spacing for FD and CD which is indended to prevent the onset of roundoff error.
 1.E-100 was hard-coded for CS into the CS method implementation, while AD had no probe distance.
@@
 \bibliographystyle{elsarticle-num}
-\bibliography{all}
+\bibliography{../all}
 \end{document}

### 2014-04-25 jacobian_calc_CMAME.tex
AS SUBMITTED FOR REVISION

@@
 While the terminology of solid mechanics is used here, it's important to note that these operators appear in other physical settings and are known by other names, e.g.
-a \emph{transmissibility matrix} in the context of a Poisson problem or, more generally, a \emph{Jacobian matrix} in mathematical optimization.
+a \emph{transmissibility matrix} in the context of a Poisson problem or, more generally, a \emph{Jacobian matrix} in mathematical optimization \cite{morton2005numerical}.
 
 While, for simplicity, the examples shown in the following refer only to linear problems, tangent-stiffness matrices generally arise in the context of non-linear analysis where Newton's method or quasi-Newton methods are used in the minimization of a given residual function.
-In this context, the tangent-stiffness matrix can be thought of as the linearization of the system
-about a particular configuration.
+In this context, the tangent-stiffness matrix can be thought of as the linearization of the system about a particular configuration.
 It has been shown \cite{hughes1978consistent,hughes1978unconditionally} that in order to preserve the quadratic convergence properties of the Newton's methods, it is necessary that this linearization is carried out in a manner that is \emph{consistent} with the algorithmic constraint equations used when computing the internal forces that arise due to deformations.
@@
 One then evaluates the expression for each combination $(i, j)$ corresponding to a (row, column) location in the tangent-stiffness matrix.
-The
-elements of a tangent-stiffness matrix can be estimated using any of the complex-step, AD, or common finite difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the argument of the function constant, and each element of the function can be evaluated independently of the others.
+The elements of a tangent-stiffness matrix can be estimated using any of the complex-step, AD, or common finite difference techniques for functions $F:R^1 \rightarrow R^1$, since taking partial derivatives entails holding all but a single independent vector component of the argument of the function constant, and each element of the function can be evaluated independently of the others.
 \\

### 2014-06-10 jacobian_calc_CMAME.tex
Fix editorial comments and grammatical corrections

@@
 %Package for adding notes/comments/and tracking changes
-\usepackage[color=red!40, textsize=scriptsize]{todonotes}
+\usepackage[disable, color=red!40, textsize=scriptsize]{todonotes}
 
@@
 The methods were compared through profiling of the code for accuracy, speed, efficiency, and parallel scalability.
-The research provides data that can serve as practical guide for code developers and analysts faced with choosing which method best suits the needs of their application code.
+This research provides data that can serve as practical guide for code developers and analysts faced with choosing which method best suits the needs of their application code.
 Additionally, motivated by the reproducible research movement, all of the code, examples, and workflow to regenerate the data and figures in this paper are provided as open source.
@@
 %% keywords here, in the form: keyword \sep keyword
-Newton's method \sep Newton-Raphson \sep numerical differentiation \sep complex-step \sep finite difference \sep automatic differentiation \sep tangent-stiffness \sep tangent-stiffness
+Newton's method \sep Newton-Raphson \sep numerical differentiation \sep complex-step \sep finite difference \sep automatic differentiation \sep tangent-stiffness
 %% MSC codes here, in the form: \MSC code \sep code
@@
 
-In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent\todo[]{Re: II.1}\footnote{In the context of plasticity thoery, the algorithmic consistency of the tangent-stiffness matrix refers to consistency with the integration of a plasticity model under the Khun-Tucker constraints or similar where the tangent-stiffness matrix is computed using the elastoplastic modulii that are determined from the integration \cite{simo1998}.} tangent-stiffness matrices.
+In order to maintain the quadratic convergence properties of the first-order Newton's method \cite{belytschko1999nonlinear} \cite[Ch.~13]{young2009} in quasi-static nonlinear analysis of solid structures it is crucial to obtain accurate, algorithmically consistent\todo[]{Re: II.1}\footnote{In the context of plasticity theory, the algorithmic consistency of the tangent-stiffness matrix refers to consistency with the integration of a plasticity model under the Khun-Tucker constraints or similar where the tangent-stiffness matrix is computed using the elastoplastic modulii that are determined from the integration \cite{simo1998}.} tangent-stiffness matrices.
 For an extremely small class of nonlinear material models, these consistent tangent-stiffness operators can be derived analytically; however, most often in practice, they are found through numerical approximation of derivatives.
@@
 Agile-components is a software development term meaning to structure a software project for continuous development to respond to evolving specifications by making much use of techniques from object-oriented programming and reusing proven pre-existing code to reduce development time and increase reliability.
-The modularity of \emph{Peridigm} allowed the adding of new features, such as those required by the study.
-
-The aim of this paper is not only to \todo[]{Re: I.1} study the complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques useful for solving non-linear systems with Newton's method.
-After presenting background information on the underlying methods to be discussed, presented in this paper are: In Section~\ref{subsec:TS} a description of tangent-stiffness matrices, detailed directions for producing tangent-stiffness matrices with each of the methods identified above, and a description and justification of the  \todo[]{Re: I.1} complex-step method for calculating tangent-stiffness matrices, in Section~\ref{subsubsec:Impl} a description of implementing the complex-step method in the \emph{Peridigm} software, in Section~\ref{JGAM} a description of the quantities-of-interest used to rank the methods, in Section~\ref{sec:Res} a presentation and analysis of the results of the comparative study, and in section~\ref{sec:Conc} conclusions and thoughts on potential future work.
+The modularity of \emph{Peridigm} allowed the adding of new features, such as those required by this study.
+
+The aim of this paper is not only to \todo[]{Re: I.1} study the complex-step method in the context of evaluating tangent-stiffness matrices, but to serve as a review for other differentiation techniques useful for solving nonlinear systems with Newton's method.
+After presenting background information on the underlying methods to be discussed, presented in this paper are: In Section~\ref{subsec:TS} a description of tangent-stiffness matrices, detailed directions for producing tangent-stiffness matrices with each of the methods identified above, and a description and justification of the  \todo[]{Re: I.1} complex-step method for calculating tangent-stiffness matrices.
+Section~\ref{sec:Validation} provides a simple illustrative example for using each of the methods for calculating tangent-stiffness matrices in the solution of nonlinear problems.
+In Section~\ref{subsubsec:Impl} a description of implementing the complex-step method in the \emph{Peridigm} software.
+In Section~\ref{JGAM} a description of the quantities-of-interest used to rank the methods.
+In Sections~\ref{sec:Res} and \ref{sec:PeridigmConvergenceStudy} a presentation and analysis of the results of the comparative study, and in Section~\ref{sec:Conc} conclusions and thoughts on potential future work.
 Finally, in the interest of replicable research, the reader is referred to the corresponding author's website for \emph{Peridigm} C++ source-code and data necessary to reproduce the results presented here.
-Additionally, included on the website is an example serial C++ library with classes for solving non-linear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
+Additionally, included on the website is an example serial C++ library with classes for solving nonlinear systems using the techniques discussed here, as well as two validation and verification example problems which show how to use the library.
 The purpose of the serial C++ library is to demonstrate in a less complex manner than in the \emph{Peridigm} code how one would use automatic-differentiation or complex-step to better encourage their use and discussion.
@@
 
-This section contains background information on the differentiation techniques underlying the tangent-stiffness matrix calculation methods compared in the study.
+This section contains background information on the differentiation techniques underlying the tangent-stiffness matrix calculation methods compared in this study.
 Since the derivation of first-order finite difference techniques from Taylor series expansion is considered well-known, it is not described here.
@@
 %
-Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor}) and solving for the first derivative and ignoring terms of $\mathcal{O}\left(h^2 \right)$ yields an estimate
+Taking the imaginary part of both sides of equation (\ref{eqn:complexTaylor}) and solving for the first derivative and ignoring terms of $\mathcal{O}(h^2)$ yields an estimate
 %
@@
 %
-The order of the truncation error in the forward difference formula (equation~(\ref{eqn:forwardFirstDeriv})) indicates that for a given $h$ this method is less accurate an approximation than either central difference or complex-step which have errors of $\mathcal{O}(h^2)$.
+The order of the truncation error in the forward difference formula (equation~(\ref{eqn:forwardFirstDeriv})) indicates that for a given $h$ this method is less accurate than either central difference or complex-step which have errors of $\mathcal{O}(h^2)$.
 
 \todo[]{Re: I.4}As mentioned earlier, the CS method was originally introduced in a paper by Lyness and Moler \cite{lyness1967numerical}.
-A follow-on contribution by Lyness demonstrated a ``truncated'' approximate version of the original CS method to make it suitable for use on digital computers \cite{lyness1968differentiation}.
-After these original papers, little can be found in the literature regarding the method until Squire and Trapp showed through numerical experiments that CS offers superior accuracy to FD and CD \cite{squire1998using} in computing numerical derivatives.
-Papers by Newman et al.\ \cite{newman1998} and Anderson et al.\ \cite{anderson2001sensitivity} discuss computing sensitivities in a CFD analysis using CS and compares the results to sensitivities produced with central difference derivatives.
-The work is also notable for identifying automatic-differentiation as a competing method to CS and suggesting that complex-step could be used to compute tangent-stiffnesss of residuals for use in Newton-Krylov schemes for solving non-linear systems.
-Perez-Foguet et al.\ \cite{perez2000numerical, perez2012numerical} used the CS method to compute algorithmically consistent elastoplastic moduli in non-linear solid mechanics problems.
+A follow-on contribution by Lyness demonstrated a ``truncated'' approximate version of the original complex-step (CS) method to make it suitable for use on digital computers \cite{lyness1968differentiation}.
+After these original papers, little can be found in the literature regarding the method until Squire and Trapp showed through numerical experiments that CS offers superior accuracy to finite-difference (FD) and central-difference (CD) \cite{squire1998using} in computing numerical derivatives.
+Papers by Newman et al.\ \cite{newman1998} and Anderson et al.\ \cite{anderson2001sensitivity} discuss computing sensitivities in a computational fluid dynamics (CFD) analysis using CS and compares the results to sensitivities produced with central difference derivatives.
+The work is also notable for identifying automatic-differentiation as a competing method to CS and suggesting that complex-step could be used to compute tangent-stiffnesss of residuals for use in Newton-Krylov schemes for solving nonlinear systems.
+Perez-Foguet et al.\ \cite{perez2000numerical, perez2012numerical} used the CS method to compute algorithmically consistent elastoplastic moduli in nonlinear solid mechanics problems.
 A paper by Joaquim et al.\ \cite{martins2000automated} investigates using CS for computing aerodynamic sensitivities.
@@
 Al-Mohy and Higham \cite{al2010complex} investigate the use of CS in computing Fr\'{e}chet derivatives of matrix functions.
-Jin et al.\ \cite{jin2010improved} discuss using CS to compute sensitivities in a solid mechanics FEM analysis.
+Jin et al.\ \cite{jin2010improved} discuss using CS to compute sensitivities in a solid mechanics finite element method (FEM) analysis.
 Voorhees et al.\ \cite{voorhees2011complex} discusses the background of the CS method and frames it as a subset of Fourier Differentiation.
@@
 They implemented the complex-values as additional degrees of freedom in the finite element stiffness matrices using a real-valued matrix representation of the complex-numbers where required.
-In a follow-on work,  Millwater et al.\  \cite{millwater2013application} extended this concept, calling it ZFEM, and implemented the method into user defined element types for use with the popular ABAQUS \cite{systemes2012abaqus} FEA software package.
+In a follow-on work,  Millwater et al.\  \cite{millwater2013application} extended this concept, calling it ZFEM, and implemented the method into user defined element types for use with the popular ABAQUS \cite{systemes2012abaqus} finite element analysis (FEA) software package.
 Nishida et al.\ \cite{nishida2013} used the complex-step method to compute tangent-stiffness matrices for implementation in a Levenberg-Marquardt optimization routine for design-of-experiments purposes.
@@
 They suggests that the modified CS method can achieve extended approximation accuracy up to fourth order.
-This recent literature has introduced the application of CS to single-dimension and multidimensional derivatives, sensitivity analysis, non-linear solution schemes, optimization, the extension of CS for higher accuracy, and implementation strategies for CS in new and existing codes.
+This recent literature has introduced the application of CS to single-dimension and multidimensional derivatives, sensitivity analysis, nonlinear solution schemes, optimization, the extension of CS for higher accuracy, and implementation strategies for CS in new and existing codes.
 The current paper presents a comparison of the complex-step method to the other methods for numerical differentiation in as close to a one-to-one implementation as possible for direct comparison purposes in a large-scale engineering analysis code.
@@
 
-An AD system evaluates composition functions in the expected manner: it works by first evaluating the innermost function of the composition, then presenting that function's output as input to the next level function until all levels are complete, observing commonly expected order of operations.
+An AD system evaluates a composition of functions in the expected manner: it works by first evaluating the innermost function of the composition, then presenting that function's output as input to the next level function until all levels are complete, observing commonly expected order of operations.
 However, an AD system does additional work during a function evaluation in that as each nested function is evaluated, the function's partial derivative with respect to the designated variables of the given input is also calculated.
@@
 
-While, for simplicity, the examples shown in the following refer only to linear problems, tangent-stiffness matrices generally arise in the context of non-linear analysis where Newton's method or quasi-Newton methods are used in the minimization of a given residual function.
+While, for simplicity, the examples shown in the following refer only to linear problems, tangent-stiffness matrices generally arise in the context of nonlinear analysis where Newton's method or quasi-Newton methods are used in the minimization of a given residual function.
 In this context, the tangent-stiffness matrix can be thought of as the linearization of the system about a particular configuration.
@@
 An example of these constraint equations would be the Kuhn-Tucker conditions \cite{simo1998} that are used in integration of a flow rule for plasticity modeling.
-If the continuum tangent-modulii are used in place of the \emph{consistent} or \emph{algorithmic tangent moduli} in the solution of a non-linear solid mechanics problem, convergence may still be achieved, but not in the quadratic manner that makes Newton's method attractive for this class of problems.
-In the setting of non-linear solid mechanics, there is a very small class of material model algorithms in which consistent tangent-stiffness operators can be derived analytically.
+If the continuum tangent-modulii are used in place of the \emph{consistent} or \emph{algorithmic tangent moduli} in the solution of a nonlinear solid mechanics problem, convergence may still be achieved, but not in the quadratic manner that makes Newton's method attractive for this class of problems.
+In the setting of nonlinear solid mechanics, there is a very small class of material model algorithms in which consistent tangent-stiffness operators can be derived analytically.
 Perfect plasticity and plasticity with isotropic hardening used in combination with a general nearest-point projection or \emph{radial return algorithm} are a few examples.
@@
 \section{\todo[]{Re: I.5}Simple example of CS, AD and FD for solution of nonlinear problems.}
-\label{subsec:Validation}
+\label{sec:Validation}
 
@@
 \end{tikzpicture}
-\caption{Beam and spring simulated in example program.} 
+\caption{Beam and spring simulated in the example program.} 
 \label{fig:TrussSchematic}
@@
 %
-For this system, the equilibrium reaction force associated with a displacement $D$ of the spring is shown in equation~(\ref{eqn:TrussForce}) 
+For this system, the equilibrium reaction force associated with a displacement $D$ of the spring is 
 %
@@
     \label{eqn:TrussForce}
-    f(D) = \frac 1 2 AE(\cos^{2}\phi)(\frac{D}{L_{0}})^{2}[\frac{D}{L_{0}}\cos^{2}\phi - 3\sin\phi] + k_{s}D + (AE\frac{D}{L_{0}})\sin^{2}\phi,
+    f(D) = \frac 1 2 AE(\cos^{2}\phi)\left(\frac{D}{L_{0}}\right)^{2}\left[\frac{D}{L_{0}}\cos^{2}\phi - 3\sin\phi\right] + k_{s}D + \left(AE\frac{D}{L_{0}}\right)\sin^{2}\phi,
 \end{equation} 
 %
-and the analytical tangent-stiffness is given by equation~(\ref{eqn:TrussStiffness})
+and the analytical tangent-stiffness is 
 %
@@
     \label{eqn:TrussStiffness}
-    S_{T}(D) = \frac 3 2 AE(\cos^{2}\phi)[\frac{D}{L_{0}}\cos^{2}\phi - 2sin\phi](\frac{D}{L_{0}^{2}}) + k_{s} + \frac{AE\sin^{2}\phi}{L_{0}}.
+    S_{T}(D) = \frac 3 2 AE(\cos^{2}\phi)\left[\frac{D}{L_{0}}\cos^{2}\phi - 2sin\phi\right]\left(\frac{D}{L_{0}^{2}}\right) + k_{s} + \frac{AE\sin^{2}\phi}{L_{0}}.
 \end{equation}
 %
-where $\phi$ is a constant truss angle, $AE$ is the axial rigidity of the truss, $L_{0}$ is original truss member length, and $k_{s}$ is the Hooke's constant for the spring attached to the truss member.
+where $\phi$ is a constant truss angle, $AE$ is the axial rigidity of the truss, $L_{0}$ is original truss member length, $k_{s}$ is the Hooke's constant for the spring attached to the truss member, and $P$ is the applied load.
 For a demonstration and verification of each of the methods as well as a convergence rate study, this truss and spring model was analyzed with each of the methods and compared with the analytic solution.
 As in the reference \cite{rezaiee2010dynamic}, equilibrium solutions for 12 incremental load steps of $\Delta P=\SI{4.448}{\newton}$ were found.
-Solver parameters common to all tests were a maximum limit of 100 iterations for convergence, a residual change tolerance of $\SI{1e-9}{\meter}$.
+Solver parameters common to all tests were a maximum limit of 100 iterations for convergence, and a convergence tolerance for the change in displacement over an iteration $\Delta D \le \SI{1e-9}{\meter}$.
 Data collected included number of iterations to converge summed over all load-steps and final displacement $D$, after the final load-step.
-Additionally finite difference step-size was varied in order to measure the impact on each of the approximate tangent-stiffness calculation methods.
-Naturally, varying step size does not affect AD or the analytically derived tangent-stiffness since those are not functions of step size.
+Additionally finite difference and complex-step step sizes were varied in order to measure the impact on each of the approximate tangent-stiffness calculation methods.
+Naturally, varying the step size does not affect AD or the analytically derived tangent-stiffness since those are not functions of the step size.
 The accuracy results for verifying the tangent-stiffness calculation methods are summarized in Table~\ref{tab:ConvergenceStudy}.
@@
   \centering
-  \caption{Displacement, D, after load-step 12.} \label{tab:Verification}   
+  \caption{Displacement, D, after load-step 12} \label{tab:Verification}   
         \begin{tabular}{c c c c c}
          \toprule
-         Step exponent & Actual $\si{\centi\meter}$ & AD $\si{\centi\meter}$ & CS $\si{\centi\meter}$ & FD $\si{\centi\meter}$\\ 
+         Step exponent & Analytical ($\si{\centi\meter}$) & AD ($\si{\centi\meter}$) & CS ($\si{\centi\meter}$) & FD ($\si{\centi\meter}$)\\ 
         \midrule
@@
 %
+The label ``Step Exponent'' refers to the value of $n$ in $h=10^{n}$ where $h$ is the finite or complex-step step size, and the symbol (") indicates that there is no change in the value from the value directly to its left.
 Table.~\ref{tab:Verification} shows that each of the methods solved the problem as accurately as one another within the precision reported.
@@
          \toprule
-         Step exponent & Actual $\si{\centi\meter}$ & AD $\si{\centi\meter}$ & CS $\si{\centi\meter}$ & FD $\si{\centi\meter}$\\ 
+         Step exponent & Actual ($\si{\centi\meter}$) & AD ($\si{\centi\meter}$) & CS ($\si{\centi\meter}$) & FD ($\si{\centi\meter}$)\\ 
         \midrule
@@
 The results of this test show that step-size selection is not unimportant when using FD, but when properly configured any of the methods may return the same maximum convergence rate for this problem.
-Overall conclusions regarding the methods that the accuracy of the tangent-stiffness should not be confused with the accuracy of the equilibrium solution, since each of the methods discussed here may be equally capable given enough iterations.
+Conclusions regarding the methods that the accuracy of the tangent-stiffness should not be confused with the accuracy of the equilibrium solution, since each of the methods discussed here may be equally capable given enough iterations.
 Numerical measurements of accuracy besides number of iterations taken are not possible here because there is not basis for comparison, as the methods are solving the problem separately and producing different local Newton iterates.
@@
 Also, because peridynamics is a nonlocal theory, the tangent-stiffness matrices have a much higher bandwidth (i.e.
-less sparsity) than traditional finite element tangent-stiffness matrix computations and the time required to construct the tangent-stiffness matrix is a majority of the total computation time in any quasi-static or implicit dynamics analysis; therefore, the \emph{Peridigm} development team has an interest in profiling the performance of the tangent-stiffness matrix computation methods.
+less sparsity) than traditional finite element tangent-stiffness matrix computations and the time required to construct the tangent-stiffness matrix is the majority of the total computation time in any quasi-static or implicit dynamics analysis; therefore, the \emph{Peridigm} development team has an interest in profiling the performance of the tangent-stiffness matrix computation methods.
 
@@
 It has demonstrated great promise in modeling problems with pervasive failure \cite{littlewood2010}, interesting phenomenon such as dynamic crack branching \cite{ha2010sod}, and is being extended to multiphyics phenomena \cite{bobaru2011peridynamic,katiyar2013} and the problem of multi-scale coupling to molecular simulations \cite{seleson2009peridynamics,seleson2014atom,rahman2013b}.
-The material model used in this study was a linear peridynamic solid which replicates linear elasticity in the \emph{local limit}, i.e.\ when the non-local length scale vanishes.
-While \emph{linear} may appear in the name of the material model, it is geometrically non-linear under large deformations and requires a nonlinear solver to converge to equilibrium conditions in these situations.
-The peridynamic equation of motion is shown in equation~(\ref{eqn:PDMotion})
+The material model used in this study was a linear peridynamic solid (LPS) which replicates classical elasticity in the \emph{local limit}, i.e.\ when the non-local length scale vanishes.
+While \emph{linear} may appear in the name of the material model, it is geometrically nonlinear under large deformations and requires a nonlinear solver to converge to equilibrium conditions in these situations.
+The peridynamic equation of motion is
 %
@@
 %
-where, $\rho$ is the mass density, $\mathbf{\ddot{u}}$ is the second time derivative of displacement (i.e.\ acceleration), $\mathcal{H}_{\delta(\mathbf{x})}$ is a neighborhood containing remote material points, $\mathbf{x'}$, with characteristic length $\delta$, centered at $\mathbf{x}$.
-The integrand $\mathbf{f}$ contains the constitutive response of the material.
+where, $\rho$ is the mass density, $\mathbf{\ddot{u}}$ is the second time derivative of displacement field (i.e.\ acceleration), and $\mathcal{H}_{\delta(\mathbf{x})}$ is a neighborhood containing remote material points, $\mathbf{x'}$ that are closer to $\mathbf{x}$ than the characteristic length $\delta$, called the peridynamic \emph{horizon}.
+The integrand $\mathbf{f}$ contains the constitutive response of the material (see \cite{silling:psa} for the exact form of $\mathbf{f}$ in the LPS model).
 Physically, $\delta$ is a material length-scale and computationally it, along with the discretization, will set the bandwidth of the tangent-stiffness matrix.
@@
 No attempt was made to ``tune'' the CS code in any way for performance.
-The algorithm  was implemented in such as way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurts its overall performance and this issue could be alleviated by using the next generation of templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any data type including {\tt complex}.
+The algorithm  was implemented in such a way that it was in one-to-one correspondence with the FD and CD methods; however, it must be noted that the overhead associated with the dynamic casts in the CS method likely hurts its overall performance and this issue could be alleviated by using the next generation of templated \emph{TPetra} vectors available in \emph{Trilinos} that are capable of taking an explicit instantiation of any data type including {\tt complex}.
 The version of \emph{Peridigm} modified to implement CS can be found in the paper repository.
@@
 \emph{Peridigm} was primarily written by computational scientists and engineers; therefore, many advanced techniques in C++ such as multiple virtual inheritance, pointer arithmetic, and distributed data structures are employed.
-These concepts are not vital to understanding the CS method itself, but may be necessary to understand CS in \emph{Peridigm}.
-To see CS and other methods in a more narrow context, it is advised that readers also take a look at the simple examples, such as that described in Section~\ref{subsec:Validation},  provided on the corresponding author's website.
+These concepts are not vital to understanding the CS method itself, but may be necessary to understand its implementation in \emph{Peridigm}.
+To see CS and other methods in a more narrow context, it is advised that readers also take a look at the simple examples, such as that described in Section~\ref{sec:Validation},  provided on the corresponding author's website.
 
@@
 The fictional material represented in the model had values of $\SI{1.515e4}{\mega\pascal}$ for bulk modulus, and $\SI{7.813e4}{\mega\pascal}$ for shear modulus.
-The peridynamic horizon, a length scale that effectively sets the bandwidth of tangent-stiffness matrix in the computation, used in the test problems was always set to three times the nominal discretization size (i.e.
+The peridynamic horizon  used in the test problems was always set to three times the nominal discretization size (i.e.
 node spacing in the particle discretization scheme).
 Pictured in Figure.~\ref{fig:pd_discrete} is an illustration of the peridynamic discretizations used in this study.
-The coloration in the images indicate material regions that were defined to assist in post-processing application of boundary conditions.
+The coloration in the images indicate material regions that were defined to assist in post-processing and application of boundary conditions.
 %
@@
         \caption{2000 node mesh.}
-        \label{fig:MCMesh}
+        \label{fig:SCMesh}
     \end{subfigure}
@@
         \caption{1000000 node mesh.}
-        \label{fig:SCMesh}
+        \label{fig:MCMesh}
     \end{subfigure}
@@
   \centering
-  \caption{Test Parameters.} \label{tab:TestParams}   
+  \caption{Test parameters.} \label{tab:TestParams}   
        \begin{tabular}{c c c c c c}
@@
 
-The goals of collecting accuracy and speed measurements were achieved by developing and implementing metrics within the simulation program used in the study.
+The goals of collecting accuracy and speed measurements were achieved by developing and implementing metrics within the simulation program used in this study.
 The metric used to measure the accuracy of a tangent-stiffness matrix was the Frobenius norm (analogous to an $l^2$-norm except defined on a matrix) of the element-wise difference between the tangent-stiffness matrix produced by the method being evaluated and the tangent-stiffness matrix produced by the AD based method (i.e., the \emph{exact} derivatives to numerical precision), given that both methods were set upon the same problem.
@@
 \emph{Python} scripts were then used to post-process and plot the data.
-These scripts averaged accuracy or speed data for a tests corresponding to a particular peridynamic node density over all iterations for that run, and then plotted this averaged data  as a function of peridynamic node density as it varied from test to test in the series.
+These scripts averaged accuracy or speed data for a test corresponding to a particular peridynamic node density over all iterations for that run, and then plotted this averaged data  as a function of peridynamic node density as it varied from test to test in the series.
 Similarly, for the parallel tests, the data was averaged and plotted in much the same way, but as a function of number of compute cores used to solve the problem.
@@
   \centering
-  \caption{Averaged Results, Each Test.} \label{tab:results}   
+  \caption{Averaged results, each test.} \label{tab:results}   
        \begin{tabular}{c c c c c c c c c}
@@
 CS was the slowest of the four methods, taking the longest average time per iteration, followed by CD, FD and AD in that order.
-Calculation time has a power-law relationship with problem size, with a power-law index of roughly $\num{1.0E-6}$ for each of the methods.
+The calculation time has a power-law relationship with problem size, with a power-law index of roughly $\num{1.0E-6}$ for each of the methods.
 This bodes well for the scalability of each of the methods by themselves in \emph{Peridigm}.
-If we assumed every individual operation, generally speaking, took the same amount of time on average such that compute time were a measure of number of operations, this suggests that the methods as in \emph{Peridigm} have the same complexity class when considering computing a TS matrix alone, that is computation time is $\mathcal{O}(10^{(10^{-6} \times \log_{10}n)})$.
+If we assume that every individual operation takes the same amount of time on average, such that compute time is a measure of number of operations, then this suggests that the methods as implemented in \emph{Peridigm} have the same complexity class when considering computing a TS matrix alone, i.e.\ computation time is $\mathcal{O}(10^{(10^{-6} \times \log_{10}n)})$.
 The definition of complexity class used here is the informal one given in \cite[pp.
@@
 There is a somewhat surprise reversal of FD and AD for these parallel simulations.
-Figure~\ref{fig:multi_speed} indicates that when looking at average iteration time, the factor of speed improvement scales inversely with the number of compute cores used for the given conditions of this group of tests.
+Figure~\ref{fig:multi_speed} indicates that when looking at the average iteration time, the factor of speed improvement scales inversely with the number of compute cores used for the given conditions of this group of tests.
 Figure~\ref{fig:multi_speed} clearly shows that when core count doubles, compute time halves, for each method.
@@
 The averaged results of these accuracy measurements for the serial tests can be seen in Figure~\ref{fig:serial_accuracy}.
-Averaged Frobenius norm (referred to as $l^2$-norm in the figure labels for conciseness) of the element-wise difference between a given TS matrix and the TS matrix produced by the AD method is plotted as a function of the number of nonzero TS matrix elements.
+The averaged Frobenius norm (referred to as $l^2$-norm in the figure labels for conciseness) of the element-wise difference between a given TS matrix and the TS matrix produced by the AD method is plotted as a function of the number of nonzero TS matrix elements.
 CS was shown to be the most accurate of the methods when compared to AD and was roughly \emph{$6$ orders of magnitude more accurate than CD} and \emph{$9$ orders of magnitude more accurate than FD}.
@@
 The value of $h$ used in this study is the default value used in \emph{Peridigm} which is heuristically derived based on computational node spacing using the method explained in \cite[pp.
-90]{ref-Adaggio}.
-To summarize, a spacing of $1.E-6$ times grid spacing is used to provide acceptable accuracy but prevent roundoff error, as is not a user determined setting in \emph{Peridigm}.
+90]{ref-Adaggio}, as it is not a user determined setting in \emph{Peridigm}.
+To summarize, a spacing of $1.E-6$ times the grid spacing is used to provide acceptable accuracy while preventing roundoff error.
 One of the beauties of the CS method is that since dependence on $h$ decreases for small $h$, which a numerical example in \cite[Table 1]{squire1998using} demonstrates, $h$, has no lower limit for effective values, therefore it can be set to machine epsilon and the analysis performed without the issues associated with round-off error.
@@
 It is necessary for an analyst to know if the accuracy differences measured in the per-iteration tests would amount to differences in convergence rates among the methods if they were allowed to solve problems normally.
-The question is posed, given that step-size $h$ is heuristically selected by \emph{Peridigm} for the finite difference methods and presumable makes a good choice, do the single-iteration tangent-stiffness accuracy results predict the ranking of the methods in terms of convergence rate for a given study?
-For brevity, convergence rate was measured for each of the methods in solving just the largest problem of the single-core test series.
-That problem was solved with each of the methods both on a single core and all sixteen cores of a  node on \emph{Stampede}.
+The question is posed, given that the step-size $h$ is heuristically selected by \emph{Peridigm} for the finite difference methods and presumably makes a good choice, do the single-iteration tangent-stiffness accuracy results predict the ranking of the methods in terms of convergence rate for a given study?
+For brevity, the convergence rate was measured for each of the methods in solving just the largest problem of the single-core test series.
+That problem was solved with each of the methods both on a single core and on all sixteen cores of a  node on \emph{Stampede}.
 As was seen in the per-iteration results, tangent-stiffness accuracy did not vary greatly from test to test, such that any pair of tests from the two series could have been considered reasonably representative.
-As in the validation problem shown in Section~\ref{subsec:Validation}, would each of the methods return an equal convergence rate so long as $h$ were optimally selected?
+As in the validation problem shown in Section~\ref{sec:Validation}, would each of the methods return an equal convergence rate so long as $h$ were optimally selected?
 
@@
 
-The end displacement results shown in Section~\ref{subsec:Validation} indicated that up to a certain precision the solutions produced by each of the methods were identical.
+The end displacement results shown in Section~\ref{sec:Validation} indicated that up to a certain precision the solutions produced by each of the methods were identical.
 It was examined if this was the case for the solutions produced in \emph{Peridigm}.
-Here each of the solutions produced by each of the methods were compared to the AD produced solutions, under the assumption that the AD produced solutions would be closest to the solution produced by using an 'exact' tangent-stiffness.
+Here each of the solutions produced by each of the methods were compared to the AD produced solutions, under the assumption that the AD produced solutions would be closest to the solution produced by using an `exact' tangent-stiffness.
 Table~\ref{tab:PeridigmSolutionAccuracy} shows that the methods did indeed produce differing solutions in terms of nodal force density, although only in decimal places at least $9$ orders of magnitude smaller than the force density value at the node where the difference was detected.
@@
 
-%The answer may be no at least unless a more highly non-linear class of problem is selected for study or a controllable method of producing a moderately inaccurate tangent-stiffness is found along with a reason for using it.
+%The answer may be no at least unless a more highly nonlinear class of problem is selected for study or a controllable method of producing a moderately inaccurate tangent-stiffness is found along with a reason for using it.
 While it is possible to incorrectly apply finite differencing and CS by choosing an inappropriate step-size, there is no reason that this should be done.
 The experiments show that step-size $h$ can be properly selected and that the methods are capable of producing accurate tangent-stiffnesss.
-It should be noted that the model used to perform the simulation could always be substituted for a more highly non-linear one and perhaps this would somewhat increase the importance of tangent-stiffness accuracy.
-The study presented here failed to pose the question of how wrong a correctly implemented tangent-stiffness calculation scheme could go for plastic or otherwise non-linear material models, here instead non-linearity was imposed by the nonlocal interaction between material points essential to peridynamic models, like the linear model used in this study.
+It should be noted that the model used to perform the simulation could always be substituted for a more highly nonlinear one and perhaps this would somewhat increase the importance of tangent-stiffness accuracy.
+The study presented here failed to pose the question of how wrong a correctly implemented tangent-stiffness calculation scheme could go for plastic or otherwise nonlinear material models, here instead nonlinearity was imposed by the nonlocal interaction between material points essential to peridynamic models, like the linear model used in this study.
 
@@
 The per-iteration speed results shown here indicate that the AD method is faster than CD or FD.
-While this may be the case for the implementation in the release version of \emph{Peridigm} this is not reflective of any limitation of these methods.
+While this may be the case for the implementation in the released version of \emph{Peridigm} this is not reflective of any limitation of these methods.
 In a side experiment, the finite difference code in \emph{Peridigm} was re-written to take advantage of the benefits of storing the results of duplicated intermediate calculations.
-These modifications were general, low memory footprint, and could be adopted in any numerical simulation simulation software as good practice.
+These modifications were general, low memory footprint, and could be adopted in any numerical simulation software as good practice.
 An informal performance test run was performed on the largest single core test from the main study, instead run on four compute cores.
@@
 \begin{tabular}{c c c c c}
-\toprule & AD & CD & Modded CD & Modded FD\\
+\toprule & AD & CD & Modified CD & Modified FD\\
 \midrule  Total seconds & $~430$ & $~570$ & $~430$ &$~320$ \\ 
@@
 
-\subsection{Final Remarks}
+\subsection{Final remarks}
 %
@@
 However, certain parallel tests done in this study showed a small performance gain over AD with the FD method.
-The results showed that CS produced more accurate tangent-stiffness matrices than CD and FD under the parameters of the tests; however, it was determined that this difference in accuracy was not sufficient to alter convergence rate for a test problem.
+The results showed that CS produced more accurate tangent-stiffness matrices than CD and FD under the parameters of the tests; however, it was determined that this difference in accuracy was not sufficient to alter the convergence rate for a test problem.
 %In order for tangent-stiffness accuracy to have an impact on convergence rate, and therefore serve as a predictor of convergence rate in per-iteration results, the amount of error detected between a supposed perfectly accurate tangent-stiffness and the test tangent-stiffness needs to be greater than allowed by the conditions seen in \cite{burkeLectures}, reproduced in subsection~\ref{subsec:FurtherStudies}.
@@
 In the context of future use of CS in \emph{Peridigm}, CS holds the advantage of relying on byte-copyable\footnote{In this context,  byte-copyable means that the collection of data to transfer is made of identically sized pieces.} data types, while AD requires a complicated serialization and deserialization in order to function as byte-copyable.
-This is fine for copying in between MPI processes, but deserialization on an accelerator or GPU is very difficult, as this would require compiling a version of the AD library being used that were compatible with the accelerator or GPU being used.
+This is fine for copying in between MPI processes, but deserialization on an accelerator or GPU is very difficult, as this would require compiling a version of the AD library being used that is compatible with the accelerator or GPU being used.
 Byte-copyability is significant for example because it is a requirement for the basic offload mode provided for Intel MIC accelerator boards which are installed in \emph{Stampede}, where the alternative mode requires more extensive effort \cite{intel_byte_copyable}.
@@
 From this perspective, the method may still be an attractive method for tangent-stiffness matrix evaluation.
-However, at least for the models producing the results shown here, each of the methods returned quadratic convergence, indicating each was sufficiently accurate.
-'Marching orders' for the analyst would be to select the swiftest method, while it would seem accuracy above what is adequate does not enhance the convergence rate or iteration speed until either classical Newton's convergence conditions or the approximate case conditions are violated.
+However, at least for the models producing the results shown here, each of the methods returned quadratic convergence, indicating that each was sufficiently accurate.
+``Marching orders'' for the analyst would be to select the swiftest method, because it appears that a greater accuracy than what is adequate does not enhance the convergence rate or iteration speed until either classical Newton's convergence conditions or the approximate case conditions are violated.
 
@@
 %
-\section{Complex-Step Example.}
+\section{Complex-Step example.}
 \label{sec:appendixA}
@@
 Because the computer knew the analytical forms of the partial derivatives of each of $u,v,w$ beforehand, which of course needed to be specified using an AD data structure from a library like Trilinos::Sacado \cite{trilinos} or ADIFOR \cite{bischof1995adifor},
-all it needed to do (for 'backward' AD) was:
+all it needed to do (for `backward' AD) was:
 \begin{enumerate}

### 2014-06-10 jacobian_calc_CMAME.tex
addressed revisions

@@
 In order to illustrate a comparison of the numerical differentiation techniques in a simpler form than it appears in the large-scale analysis code covered in the following sections, a simple example borrowed from \cite{rezaiee2010dynamic} was implemented with each of the methods, CS, AD, and FD.
+The reason for including FD was to provide an example of implementing a familiar differentiation technique alongside the more unusual CS and AD examples.
+CD is omitted for brevity in the code development for the example.
 All the source code to reproduce the results described here are available in the associated paper repository.
@@
 where $\phi$ is a constant truss angle, $AE$ is the axial rigidity of the truss, $L_{0}$ is original truss member length, $k_{s}$ is the Hooke's constant for the spring attached to the truss member, and $P$ is the applied load.
+$\phi$ is considered constant for small displacement for initial truss angles close to zero.
 For a demonstration and verification of each of the methods as well as a convergence rate study, this truss and spring model was analyzed with each of the methods and compared with the analytic solution.
@@
   \caption{Test parameters.} \label{tab:TestParams}   
-       \begin{tabular}{c c c c c c}
+       \begin{tabular}{c c c c c c c}
          \toprule
-         Peri. nodes & Nozero TS elems. & Horizon ($\delta$) & Load rate & Number Load steps & Abs. tolerance \\ 
-                     & $\times 10^6$    & \si{\centi\meter}& \si{\centi\meter\per\second} &   &     \\
+				 Num. cores &Peri. nodes & Nozero TS elems. & Horizon ($\delta$) & Load rate & Number Load steps & Abs. tolerance \\ 
+         &            & $\times 10^6$    & \si{\centi\meter}& \si{\centi\meter\per\second} &   &     \\
         \midrule
-        1000 & $1.99$             & $30$   & $2$ & $2$ & $\num{10E-8}$ \\
-        2000 & $4.25$             & $24$  & $2$ & $2$ & $\num{10E-8}$ \\
-        3000 & $7.79$             & $21$  & $2$ & $2$ & $\num{10E-8}$ \\
-        4000 & $15.4$             & $19$  & $2$ & $2$ & $\num{10E-8}$ \\
-        6000 & $20.0$             & $16$  & $2$ & $2$ & $\num{10E-8}$ \\
-        8000 & $31.4$             & $15$  & $2$ & $2$ & $\num{10E-8}$ \\
-        12000 & $40.1$            & $13.1$ & $2$ & $2$ & $\num{10E-8}$ \\
-        16000 & $83.9$            & $12$  & $2$ & $2$ & $\num{10E-8}$ \\
-        32000 & $165.0$           & $9.5$ & $2$ & $2$ & $\num{10E-8}$ \\
-        1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
-        1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
-        1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
-        1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 1000 & $1.99$             & $30$   & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 2000 & $4.25$             & $24$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 3000 & $7.79$             & $21$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 4000 & $15.4$             & $19$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 6000 & $20.0$             & $16$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 8000 & $31.4$             & $15$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 12000 & $40.1$            & $13.1$ & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 16000 & $83.9$            & $12$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 32000 & $165.0$           & $9.5$ & $2$ & $2$ & $\num{10E-8}$ \\
+        32 & 1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
+        64 & 1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
+        96 & 1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
+        128 & 1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
         \bottomrule
@@
 Here each of the solutions produced by each of the methods were compared to the AD produced solutions, under the assumption that the AD produced solutions would be closest to the solution produced by using an `exact' tangent-stiffness.
-Table~\ref{tab:PeridigmSolutionAccuracy} shows that the methods did indeed produce differing solutions in terms of nodal force density, although only in decimal places at least $9$ orders of magnitude smaller than the force density value at the node where the difference was detected.
+Table~\ref{tab:PeridigmSolutionAccuracy} shows maximum force density discrepancy, indicating that the methods did indeed produce differing solutions in terms of nodal force density.
+The values for Table~\ref{tab:PeridigmSolutionAccuracy} were generated using a helper program included with \emph{Trilinos} called \emph{exodiff} which is used for automatically comparing \emph{exodus} output files.
+While differences in force density were detected, the magnitude of the greatest of them was $9$ orders of magnitude smaller than the relevant nodal force density values being differenced.
 %
@@
 Table~\ref{tab:PerformanceMod} shows the timing results for AD as reference along with an unmodified CD, and then modified CD and FD.
+Unmodified FD and the CS methods do not appear in the comparison since the modification was not implemented for the CS method and timing results for the unmodified methods appear, as for a different computer system, in Table~\ref{tab:results}.
 %
@@
 Because the computer knew the analytical forms of the partial derivatives of each of $u,v,w$ beforehand, which of course needed to be specified using an AD data structure from a library like Trilinos::Sacado \cite{trilinos} or ADIFOR \cite{bischof1995adifor},
-all it needed to do (for `backward' AD) was:
+all it needed to do was:
 \begin{enumerate}
@@
 Lastly, to compute the partial derivative of the entire composition function, the computer multiplies
-the individual partials together in observance of the chain-rule.
+the individual partials together in observance of the chain-rule. It is important to note that this example
+was simplistic in that only the case of a single independent variable was covered, making the distinction of type of AD
+less important.
 

### 2014-06-10 data/multicore/plotter.py

@@ -60,7 +60,7 @@ ad = ax1.scatter(speed_averages['AD'][0][:], speed_averages['AD'][1][:], marker=
 #ax1.set_yscale('log')
 plt.xlim(24.0, 160.0)
 plt.ylim(0.0, 350.0)
-plt.xlabel("Number of cores at 1TB RAM per 32 cores", fontsize=font_size)
+plt.xlabel("Number of cores", fontsize=font_size)
 plt.ylabel("Avg. time to compute TS (s)", fontsize=font_size)
 #plt.title("Average time to compute a Jacobian matrix vs. number of cores")
 plt.legend([cs, cd, fd, ad],['CS', 'CD', 'FD', 'AD'], scatterpoints = 1, loc='upper right',

### 2014-06-10 jacobian_calc_CMAME.tex
Address item 18

@@
 %
-where $\phi$ is a constant truss angle, $AE$ is the axial rigidity of the truss, $L_{0}$ is original truss member length, $k_{s}$ is the Hooke's constant for the spring attached to the truss member, and $P$ is the applied load.
+where $\phi$ is a constant truss angle (a modeling decision made in \cite{rezaiee2010dynamic} that we mimic for consistency), $AE$ is the axial rigidity of the truss, $L_{0}$ is original truss member length, $k_{s}$ is the Hooke's constant for the spring attached to the truss member, and $P$ is the applied load.
 $\phi$ is considered constant for small displacement for initial truss angles close to zero.

### 2014-06-10 jacobian_calc_CMAME.tex
Final changes

@@
 where $\phi$ is a constant truss angle (a modeling decision made in \cite{rezaiee2010dynamic} that we mimic for consistency), $AE$ is the axial rigidity of the truss, $L_{0}$ is original truss member length, $k_{s}$ is the Hooke's constant for the spring attached to the truss member, and $P$ is the applied load.
-$\phi$ is considered constant for small displacement for initial truss angles close to zero.
 For a demonstration and verification of each of the methods as well as a convergence rate study, this truss and spring model was analyzed with each of the methods and compared with the analytic solution.
@@
 \todo[]{Re: II.16} The specific parameters of these tests can be found in \emph{Peridigm} input {\tt xml} files included in the paper repository, however the parameters are reproduced in Table~\ref{tab:TestParams}.
+All of the tests used a loading rate of $\SI{2}{\centi\meter\per\second}$ for $\num{2}$ load steps, with an absolute residual convergence tolerence of $\num{10E-8}$.
 These {\tt xml} files allow users with the appropriately modified version of \emph{Peridigm} to reproduce the results found in this paper.
@@
          \toprule
-				 Num. cores &Peri. nodes & Nozero TS elems. & Horizon ($\delta$) & Load rate & Number Load steps & Abs. tolerance \\ 
-         &            & $\times 10^6$    & \si{\centi\meter}& \si{\centi\meter\per\second} &   &     \\
+				 Num. cores &Peri. nodes & Nozero TS elems. & Horizon ($\delta$)  \\ 
+         &            & $\times 10^6$    & \si{\centi\meter}   \\
         \midrule
-        1 & 1000 & $1.99$             & $30$   & $2$ & $2$ & $\num{10E-8}$ \\
-        1 & 2000 & $4.25$             & $24$  & $2$ & $2$ & $\num{10E-8}$ \\
-        1 & 3000 & $7.79$             & $21$  & $2$ & $2$ & $\num{10E-8}$ \\
-        1 & 4000 & $15.4$             & $19$  & $2$ & $2$ & $\num{10E-8}$ \\
-        1 & 6000 & $20.0$             & $16$  & $2$ & $2$ & $\num{10E-8}$ \\
-        1 & 8000 & $31.4$             & $15$  & $2$ & $2$ & $\num{10E-8}$ \\
-        1 & 12000 & $40.1$            & $13.1$ & $2$ & $2$ & $\num{10E-8}$ \\
-        1 & 16000 & $83.9$            & $12$  & $2$ & $2$ & $\num{10E-8}$ \\
-        1 & 32000 & $165.0$           & $9.5$ & $2$ & $2$ & $\num{10E-8}$ \\
-        32 & 1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
-        64 & 1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
-        96 & 1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
-        128 & 1000000 & $1670$          & $3$  & $2$ & $2$ & $\num{10E-8}$ \\
+        1 & 1000 & $1.99$             & $30$    \\
+        1 & 2000 & $4.25$             & $24$   \\
+        1 & 3000 & $7.79$             & $21$   \\
+        1 & 4000 & $15.4$             & $19$   \\
+        1 & 6000 & $20.0$             & $16$   \\
+        1 & 8000 & $31.4$             & $15$   \\
+        1 & 12000 & $40.1$            & $13.1$ \\
+        1 & 16000 & $83.9$            & $12$   \\
+        1 & 32000 & $165.0$           & $9.5$  \\
+        32 & 1000000 & $1670$          & $3$   \\
+        64 & 1000000 & $1670$          & $3$   \\
+        96 & 1000000 & $1670$          & $3$   \\
+        128 & 1000000 & $1670$          & $3$  \\
         \bottomrule
@@
 Table~\ref{tab:PeridigmSolutionAccuracy} shows maximum force density discrepancy, indicating that the methods did indeed produce differing solutions in terms of nodal force density.
-The values for Table~\ref{tab:PeridigmSolutionAccuracy} were generated using a helper program included with \emph{Trilinos} called \emph{exodiff} which is used for automatically comparing \emph{exodus} output files.
+The values for Table~\ref{tab:PeridigmSolutionAccuracy} were generated using a helper program included with \emph{Trilinos} called {\tt exodiff} which is used for automatically comparing \emph{Exodus} database output files.
 While differences in force density were detected, the magnitude of the greatest of them was $9$ orders of magnitude smaller than the relevant nodal force density values being differenced.
